### 2018-01-30 Draft/Draft.Rmd
First commit

---
author:
  - name: Ursula Laa
    affiliation: Monash University
    address: >
      First line
      Second line
    email: \email{ursula.laa@monash.edu}
    url: http://rstudio.com
  - name: Second Author
    affiliation: Affiliation
title:
  formatted: "Scagnostics and Projection Pursuit"
  # If you use tex in the formatted title, also supply version without
  #plain:     "A Capitalized Title: Something about a Package foo"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [scagnostics, tour, projection pursuit]
  #plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE, 
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
)
read_chunk('src/main.R')
```

```{r load}
```

# Introduction

A grand tour provides dynamic projections from the high dimensions into some low dimensional representation, in particular also as a
two dimensional scatter plot ADDREF.
The idea of projection persuit can be coupled to the tour algorithm, to guide the tour towards more interesting views of the projected data (the so-called guided tour), by maximising a given projection persuit index ADDREF (Cook, Buja, Cabrera, Hurley, 1995).
The guided tour thus allows the user to view interesting projections (local maxima or minima of given index) as well as their neighborhoods, allowing the user to see the projection in the context of the full parameter space.
Typically projection persuit indices are built to detect deviations of the projected data from a normal distribution, and the nonnormalness is considered equivalent to interestingness of a given projection ADDREF (Cook, Buja, Cabrera, 1993).
Moreover, indices for detecting holes, concentration of mass in the center or skewness have also been designed.

Here we aim to extend the definition of interesting projections and available projection persuit indices by studying potential
new index functions.
As a starting point we consdier the scagnostics measures that were proposed for selecting bivariate scatterplots in high-dimensional datasets ADDREF (??).
Rather than considering the scagnostics measures for all possible bivariate combinations, we want to use one of the measures as an index function for a guided tour of 2 dimensional projections.
We first study the feasability of this approach and discuss potential pitfalls.
We then aim to define more robust measures ... ???

# Scagnostics measures

Eight scagnostics measures have been defined, the definitions are based on the convex and alpha hull as well as the minimum spanning tree.
Give definitions here...

Note that in practice the measures are calculated after binning the data to avoid large computing times.

# Analysis setup

To study the feasablility of using scagnostics measures as an index fuction for guided tours, we first consider the variation of each of the measures, as calculated from the projected scatter plot, when smoothly changing the projection via the tour algorithm.
Moreover, we test the dependence on the sample size and on the outlier removal typically performed to make the scagnostics measures more robust.
In addition the computing time is also important, as it might limit the dynamic viewing and require to first record a guided tour path which can
later be viewed in a smooth display.

As we will see the scagnostics measures often exhibit a significant jump despite changing smoothly between projections, a consequence of the discreete nature of the definition, but possibly also due to the binning performed as a first step of calculation.
We therefore explore options of ameliorating this situation using different smoothing methods, and by redefining the measures in a more robust fashion.

# Application to particle physics data

## Data description

For the following illustrations we work with the particle physics dataset (ATLAS Collaboration, 2015).
This data includes a number of predictions for observables (in agreement with current limits) and the so-called finetuning measure for a large set of parameter points in a supersymmetric model.
The observables describe predictions for decay branching ratios and other precision observables that are generally used to constrain new models of particle physics.
We use the log transformed values for the variables XYZ but not the finetuning, and we rescale all variables to a [0,1] interval
to obtain comparable scales.
Contributions to the variable gminus2 can be negative or positive, we therefore add a small shift before the log transformation of this variable.

## Software setup

Use tourr and scagnostics package (ADDREF)

## Scagnositics of grand tour projections

```{r scagnostics-summary, fig.height= 14, fig.width=9, fig.cap = "(ref:scagnostics-summary)"}
```

(ref:scagnostics-summary) Summary of evolution of scagnostics measures over 100 interpolated grand tour projections
We see that three of the considered measures, Clumpy, Sparse and Striated, are not relevant for the considered data and projections,
as the valuse are consistently low (clearly below 0.1).
The remaining measures show large variation depending on the projection, sample size and also often strong dependence on the
outlier removal before calculation of scagnostics measures.
Moreover, we note that despite considering interpolated projections, i.e. a smooth transition between the randomly selected projection planes in the grand tour, the variation of the scagnostics measures does not follow a smooth evolution but exhibits several spikes.
This behaviour is expected to result in difficulites when using scagnostics measures as a projection pursuit index in the context of a dynamic guided tour.
To further understand this behaviour let us study the projected data points and also the binned version for some of the projections exhibiting a spike.
We first show in Figure ADDREF projections t=18/19/20 of the small data set (500 data points).
Not that t=19 has been marked by a vertical line in Figure ADDREF.
The left column shows the projected data points, and the right column the binned data used to calculate the scagnositcs measures.
The scagnostics measures in the three projections are summarised in Table ADDREF.
We see that despite similar distribution of the points in all three projections, the outlying measure is similar for t=18 and t=20 but jumpig down for t=19.
Differences can however be observed in the binned data (right column of Figure ADDREF), likely causing the behaviour of the measures.

(This is blueprint of how to show scatter and binning plots but maybe other projections will be more interesting.)

FIXME need to arrange plots better! Fix common axes for all corresponding plots
FIXME binned plot needs better formatting so we can see relevant features easily

```{r scatter-binning, fig.height=2, fig.cap = "(ref:scatter-binning)"}
```

```{r scatter-table}
```

Another consideration relevant for the purpose of the guided tour is the computing time, as mentioned above it may limit the tour display to previously recorded tours.
We study here the computing time depending on the sample size with and without the outlier removal when calculating the measures.
Figure ADDREF shows this variation for 3 different sample selections.
The time is measured over the 100 projections shown before and is the sum of computing time for projecting the data and calculating the scagnostics measures on the projected data.
We see that overall the outlier removal increases the runtime by about a factor two.
However we have seen above that skipping the step of outlier removal often leads to very different values of some of the scagnostics measures. (What to do with this information? Should not skip this step?)
Interestingly the computing times are longest for intermediate sample sizes of a few hundred data points, and converge to lower run times for larger samples.
This is because the binning of the input data is much faster than the computation of the measures which is increased for smaller samples.

```{r scagnositcs-timing, fig.cap = "(ref:scagnostics-timing)"}
```

## Smooth transitions

We next turn to the question of how we can obtain smooth transitions of the scagnostics measures.
(Below just an example of using ggplot smoothing to display what kind of behaviour we want to find. It follows the general trend without capturing the spikes.)
```{r smoothing, fig.cap = "(ref:smooting)"}
```

Want to explore jittering either the projection or the points, but this does not give the desired smooth behaviour (why?)
Maybe show how it compares to the sketch of smooth polynomial function? Select parameters from previous study without including the detailed plots here..

### Binning in high dimensions?
### Outlier removal before binning?

## Test with guided tour
try out scagnostics measures as guided tour index
FIXME still need to write script for this

## Additional indices to consider
FIXME load Katrins package and try the index functions she implemented

### 2018-01-30 Draft/Draft.tex
First commit

--- 2018-01-30 Draft/Draft.Rmd
@@
----
-author:
-  - name: Ursula Laa
-    affiliation: Monash University
-    address: >
-      First line
-      Second line
-    email: \email{ursula.laa@monash.edu}
-    url: http://rstudio.com
-  - name: Second Author
-    affiliation: Affiliation
-title:
-  formatted: "Scagnostics and Projection Pursuit"
-  # If you use tex in the formatted title, also supply version without
-  #plain:     "A Capitalized Title: Something about a Package foo"
-  # For running headers, if needed
-  #short:     "\\pkg{foo}: A Capitalized Title"
-abstract: >
+\documentclass[article]{jss}
+\usepackage[utf8]{inputenc}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\author{
+Ursula Laa\\Monash University \And Second Author\\Affiliation
+}
+\title{Scagnostics and Projection Pursuit}
+
+\Plainauthor{Ursula Laa, Second Author}
+
+\Abstract{
 The abstract of the article.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [scagnostics, tour, projection pursuit]
-  #plain:     [keywords, not capitalized, Java]
-preamble: >
+}
+
+\Keywords{scagnostics, tour, projection pursuit}
+
+%% publication information
+%% \Volume{50}
+%% \Issue{9}
+%% \Month{June}
+%% \Year{2012}
+%% \Submitdate{}
+%% \Acceptdate{2012-06-04}
+
+\Address{
+    Ursula Laa\\
+  Monash University\\
+  First line Second line\\
+  E-mail: \email{ursula.laa@monash.edu}\\
+  URL: \url{http://rstudio.com}\\~\\
+    }
+
 \usepackage{amsmath}
-output: rticles::jss_article
----
 
-```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
-library(knitr)
-opts_chunk$set(
-  warning = FALSE, message = FALSE, echo = FALSE, 
-  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
-  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
-)
-read_chunk('src/main.R')
-```
+\begin{document}
 
-```{r load}
-```
+\section{Introduction}\label{introduction}
 
-# Introduction
+A grand tour provides dynamic projections from the high dimensions into
+some low dimensional representation, in particular also as a two
+dimensional scatter plot ADDREF. The idea of projection persuit can be
+coupled to the tour algorithm, to guide the tour towards more
+interesting views of the projected data (the so-called guided tour), by
+maximising a given projection persuit index ADDREF (Cook, Buja, Cabrera,
+Hurley, 1995). The guided tour thus allows the user to view interesting
+projections (local maxima or minima of given index) as well as their
+neighborhoods, allowing the user to see the projection in the context of
+the full parameter space. Typically projection persuit indices are built
+to detect deviations of the projected data from a normal distribution,
+and the nonnormalness is considered equivalent to interestingness of a
+given projection ADDREF (Cook, Buja, Cabrera, 1993). Moreover, indices
+for detecting holes, concentration of mass in the center or skewness
+have also been designed.
 
-A grand tour provides dynamic projections from the high dimensions into some low dimensional representation, in particular also as a
-two dimensional scatter plot ADDREF.
-The idea of projection persuit can be coupled to the tour algorithm, to guide the tour towards more interesting views of the projected data (the so-called guided tour), by maximising a given projection persuit index ADDREF (Cook, Buja, Cabrera, Hurley, 1995).
-The guided tour thus allows the user to view interesting projections (local maxima or minima of given index) as well as their neighborhoods, allowing the user to see the projection in the context of the full parameter space.
-Typically projection persuit indices are built to detect deviations of the projected data from a normal distribution, and the nonnormalness is considered equivalent to interestingness of a given projection ADDREF (Cook, Buja, Cabrera, 1993).
-Moreover, indices for detecting holes, concentration of mass in the center or skewness have also been designed.
+Here we aim to extend the definition of interesting projections and
+available projection persuit indices by studying potential new index
+functions. As a starting point we consdier the scagnostics measures that
+were proposed for selecting bivariate scatterplots in high-dimensional
+datasets ADDREF (??). Rather than considering the scagnostics measures
+for all possible bivariate combinations, we want to use one of the
+measures as an index function for a guided tour of 2 dimensional
+projections. We first study the feasability of this approach and discuss
+potential pitfalls. We then aim to define more robust measures \ldots{}
+???
 
-Here we aim to extend the definition of interesting projections and available projection persuit indices by studying potential
-new index functions.
-As a starting point we consdier the scagnostics measures that were proposed for selecting bivariate scatterplots in high-dimensional datasets ADDREF (??).
-Rather than considering the scagnostics measures for all possible bivariate combinations, we want to use one of the measures as an index function for a guided tour of 2 dimensional projections.
-We first study the feasability of this approach and discuss potential pitfalls.
-We then aim to define more robust measures ... ???
+\section{Scagnostics measures}\label{scagnostics-measures}
 
-# Scagnostics measures
+Eight scagnostics measures have been defined, the definitions are based
+on the convex and alpha hull as well as the minimum spanning tree. Give
+definitions here\ldots{}
 
-Eight scagnostics measures have been defined, the definitions are based on the convex and alpha hull as well as the minimum spanning tree.
-Give definitions here...
+Note that in practice the measures are calculated after binning the data
+to avoid large computing times.
 
-Note that in practice the measures are calculated after binning the data to avoid large computing times.
+\section{Analysis setup}\label{analysis-setup}
 
-# Analysis setup
+To study the feasablility of using scagnostics measures as an index
+fuction for guided tours, we first consider the variation of each of the
+measures, as calculated from the projected scatter plot, when smoothly
+changing the projection via the tour algorithm. Moreover, we test the
+dependence on the sample size and on the outlier removal typically
+performed to make the scagnostics measures more robust. In addition the
+computing time is also important, as it might limit the dynamic viewing
+and require to first record a guided tour path which can later be viewed
+in a smooth display.
 
-To study the feasablility of using scagnostics measures as an index fuction for guided tours, we first consider the variation of each of the measures, as calculated from the projected scatter plot, when smoothly changing the projection via the tour algorithm.
-Moreover, we test the dependence on the sample size and on the outlier removal typically performed to make the scagnostics measures more robust.
-In addition the computing time is also important, as it might limit the dynamic viewing and require to first record a guided tour path which can
-later be viewed in a smooth display.
+As we will see the scagnostics measures often exhibit a significant jump
+despite changing smoothly between projections, a consequence of the
+discreete nature of the definition, but possibly also due to the binning
+performed as a first step of calculation. We therefore explore options
+of ameliorating this situation using different smoothing methods, and by
+redefining the measures in a more robust fashion.
 
-As we will see the scagnostics measures often exhibit a significant jump despite changing smoothly between projections, a consequence of the discreete nature of the definition, but possibly also due to the binning performed as a first step of calculation.
-We therefore explore options of ameliorating this situation using different smoothing methods, and by redefining the measures in a more robust fashion.
+\section{Application to particle physics
+data}\label{application-to-particle-physics-data}
 
-# Application to particle physics data
+\subsection{Data description}\label{data-description}
 
-## Data description
+For the following illustrations we work with the particle physics
+dataset (ATLAS Collaboration, 2015). This data includes a number of
+predictions for observables (in agreement with current limits) and the
+so-called finetuning measure for a large set of parameter points in a
+supersymmetric model. The observables describe predictions for decay
+branching ratios and other precision observables that are generally used
+to constrain new models of particle physics. We use the log transformed
+values for the variables XYZ but not the finetuning, and we rescale all
+variables to a {[}0,1{]} interval to obtain comparable scales.
+Contributions to the variable gminus2 can be negative or positive, we
+therefore add a small shift before the log transformation of this
+variable.
 
-For the following illustrations we work with the particle physics dataset (ATLAS Collaboration, 2015).
-This data includes a number of predictions for observables (in agreement with current limits) and the so-called finetuning measure for a large set of parameter points in a supersymmetric model.
-The observables describe predictions for decay branching ratios and other precision observables that are generally used to constrain new models of particle physics.
-We use the log transformed values for the variables XYZ but not the finetuning, and we rescale all variables to a [0,1] interval
-to obtain comparable scales.
-Contributions to the variable gminus2 can be negative or positive, we therefore add a small shift before the log transformation of this variable.
-
-## Software setup
+\subsection{Software setup}\label{software-setup}
 
@@
 
-## Scagnositics of grand tour projections
+\subsection{Scagnositics of grand tour
+projections}\label{scagnositics-of-grand-tour-projections}
 
-```{r scagnostics-summary, fig.height= 14, fig.width=9, fig.cap = "(ref:scagnostics-summary)"}
-```
+\begin{CodeChunk}
+\begin{figure}
 
-(ref:scagnostics-summary) Summary of evolution of scagnostics measures over 100 interpolated grand tour projections
-We see that three of the considered measures, Clumpy, Sparse and Striated, are not relevant for the considered data and projections,
-as the valuse are consistently low (clearly below 0.1).
-The remaining measures show large variation depending on the projection, sample size and also often strong dependence on the
-outlier removal before calculation of scagnostics measures.
-Moreover, we note that despite considering interpolated projections, i.e.
-a smooth transition between the randomly selected projection planes in the grand tour, the variation of the scagnostics measures does not follow a smooth evolution but exhibits several spikes.
-This behaviour is expected to result in difficulites when using scagnostics measures as a projection pursuit index in the context of a dynamic guided tour.
-To further understand this behaviour let us study the projected data points and also the binned version for some of the projections exhibiting a spike.
-We first show in Figure ADDREF projections t=18/19/20 of the small data set (500 data points).
-Not that t=19 has been marked by a vertical line in Figure ADDREF.
-The left column shows the projected data points, and the right column the binned data used to calculate the scagnositcs measures.
-The scagnostics measures in the three projections are summarised in Table ADDREF.
-We see that despite similar distribution of the points in all three projections, the outlying measure is similar for t=18 and t=20 but jumpig down for t=19.
-Differences can however be observed in the binned data (right column of Figure ADDREF), likely causing the behaviour of the measures.
+{\centering \includegraphics[width=\textwidth]{figure/scagnostics-summary-1}
 
-(This is blueprint of how to show scatter and binning plots but maybe other projections will be more interesting.)
+}
 
-FIXME need to arrange plots better! Fix common axes for all corresponding plots
-FIXME binned plot needs better formatting so we can see relevant features easily
+\caption[(ref:scagnostics-summary)]{(ref:scagnostics-summary)}\label{fig:scagnostics-summary}
+\end{figure}
+\end{CodeChunk}
 
-```{r scatter-binning, fig.height=2, fig.cap = "(ref:scatter-binning)"}
-```
+(ref:scagnostics-summary) Summary of evolution of scagnostics measures
+over 100 interpolated grand tour projections We see that three of the
+considered measures, Clumpy, Sparse and Striated, are not relevant for
+the considered data and projections, as the valuse are consistently low
+(clearly below 0.1). The remaining measures show large variation
+depending on the projection, sample size and also often strong
+dependence on the outlier removal before calculation of scagnostics
+measures. Moreover, we note that despite considering interpolated
+projections, i.e.~a smooth transition between the randomly selected
+projection planes in the grand tour, the variation of the scagnostics
+measures does not follow a smooth evolution but exhibits several spikes.
+This behaviour is expected to result in difficulites when using
+scagnostics measures as a projection pursuit index in the context of a
+dynamic guided tour. To further understand this behaviour let us study
+the projected data points and also the binned version for some of the
+projections exhibiting a spike. We first show in Figure ADDREF
+projections t=18/19/20 of the small data set (500 data points). Not that
+t=19 has been marked by a vertical line in Figure ADDREF. The left
+column shows the projected data points, and the right column the binned
+data used to calculate the scagnositcs measures. The scagnostics
+measures in the three projections are summarised in Table ADDREF. We see
+that despite similar distribution of the points in all three
+projections, the outlying measure is similar for t=18 and t=20 but
+jumpig down for t=19. Differences can however be observed in the binned
+data (right column of Figure ADDREF), likely causing the behaviour of
+the measures.
 
-```{r scatter-table}
-```
+(This is blueprint of how to show scatter and binning plots but maybe
+other projections will be more interesting.)
 
-Another consideration relevant for the purpose of the guided tour is the computing time, as mentioned above it may limit the tour display to previously recorded tours.
-We study here the computing time depending on the sample size with and without the outlier removal when calculating the measures.
-Figure ADDREF shows this variation for 3 different sample selections.
-The time is measured over the 100 projections shown before and is the sum of computing time for projecting the data and calculating the scagnostics measures on the projected data.
-We see that overall the outlier removal increases the runtime by about a factor two.
-However we have seen above that skipping the step of outlier removal often leads to very different values of some of the scagnostics measures.
-(What to do with this information?
-Should not skip this step?)
-Interestingly the computing times are longest for intermediate sample sizes of a few hundred data points, and converge to lower run times for larger samples.
-This is because the binning of the input data is much faster than the computation of the measures which is increased for smaller samples.
+FIXME need to arrange plots better! Fix common axes for all
+corresponding plots FIXME binned plot needs better formatting so we can
+see relevant features easily
 
-```{r scagnositcs-timing, fig.cap = "(ref:scagnostics-timing)"}
-```
+\begin{CodeChunk}
+\begin{figure}
 
-## Smooth transitions
+{\centering \includegraphics[width=\textwidth]{figure/scatter-binning-1} \includegraphics[width=\textwidth]{figure/scatter-binning-2} \includegraphics[width=\textwidth]{figure/scatter-binning-3}
 
-We next turn to the question of how we can obtain smooth transitions of the scagnostics measures.
-(Below just an example of using ggplot smoothing to display what kind of behaviour we want to find. It follows the general trend without capturing the spikes.)
-```{r smoothing, fig.cap = "(ref:smooting)"}
-```
+}
 
-Want to explore jittering either the projection or the points, but this does not give the desired smooth behaviour (why?)
-Maybe show how it compares to the sketch of smooth polynomial function? Select parameters from previous study without including the detailed plots here..
+\caption[(ref:scatter-binning)]{(ref:scatter-binning)}\label{fig:scatter-binning}
+\end{figure}
+\end{CodeChunk}
 
-### Binning in high dimensions?
-### Outlier removal before binning?
+\begin{CodeChunk}
 
-## Test with guided tour
-try out scagnostics measures as guided tour index
-FIXME still need to write script for this
+\begin{CodeOutput}
+                   s1          s2          s3
+Outlying  0.275676305 0.163548110 0.270907044
+Skewed    0.549038842 0.575422054 0.583569727
+Clumpy    0.004123711 0.008163265 0.004175365
+Sparse    0.038765964 0.042779580 0.040551326
+Striated  0.059405941 0.060000000 0.037433155
+Convex    0.611714549 0.602591732 0.546399191
+Skinny    0.483884014 0.495557112 0.564627720
+Stringy   0.424567811 0.368764181 0.329434667
+Monotonic 0.077330543 0.094217913 0.103036354
+\end{CodeOutput}
+\end{CodeChunk}
 
-## Additional indices to consider
+Another consideration relevant for the purpose of the guided tour is the
+computing time, as mentioned above it may limit the tour display to
+previously recorded tours. We study here the computing time depending on
+the sample size with and without the outlier removal when calculating
+the measures. Figure ADDREF shows this variation for 3 different sample
+selections. The time is measured over the 100 projections shown before
+and is the sum of computing time for projecting the data and calculating
+the scagnostics measures on the projected data. We see that overall the
+outlier removal increases the runtime by about a factor two. However we
+have seen above that skipping the step of outlier removal often leads to
+very different values of some of the scagnostics measures. (What to do
+with this information? Should not skip this step?) Interestingly the
+computing times are longest for intermediate sample sizes of a few
+hundred data points, and converge to lower run times for larger samples.
+This is because the binning of the input data is much faster than the
+computation of the measures which is increased for smaller samples.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/scagnositcs-timing-1}
+
+}
+
+\caption[(ref:scagnostics-timing)]{(ref:scagnostics-timing)}\label{fig:scagnositcs-timing}
+\end{figure}
+\end{CodeChunk}
+
+\subsection{Smooth transitions}\label{smooth-transitions}
+
+We next turn to the question of how we can obtain smooth transitions of
+the scagnostics measures. (Below just an example of using ggplot
+smoothing to display what kind of behaviour we want to find. It follows
+the general trend without capturing the spikes.)
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/smoothing-1}
+
+}
+
+\caption[(ref:smooting)]{(ref:smooting)}\label{fig:smoothing}
+\end{figure}
+\end{CodeChunk}
+
+Want to explore jittering either the projection or the points, but this
+does not give the desired smooth behaviour (why?) Maybe show how it
+compares to the sketch of smooth polynomial function? Select parameters
+from previous study without including the detailed plots here..
+
+\subsubsection{Binning in high
+dimensions?}\label{binning-in-high-dimensions}
+
+\subsubsection{Outlier removal before
+binning?}\label{outlier-removal-before-binning}
+
+\subsection{Test with guided tour}\label{test-with-guided-tour}
+
+try out scagnostics measures as guided tour index FIXME still need to
+write script for this
+
+\subsection{Additional indices to
+consider}\label{additional-indices-to-consider}
+
 FIXME load Katrins package and try the index functions she implemented
+
+\end{document}

### 2018-01-30 Draft/src/main.R

@@ -1,0 +1,157 @@
+## ---- load
+
+library(tourr)
+library(tidyverse)
+library(reshape2)
+library(binostics)
+library(gridExtra)
+library(wesanderson) #color palette
+library(tictoc) #timer
+
+# function to log transform part of data
+asLog <- function(dt){
+  dtOut <- dt %>% mutate(BRbsg_ln = log(BRbsg)) %>%
+    mutate(BrBsmumu_ln = log(BrBsmumu)) %>%
+    mutate(Omegahsq_ln = log(Omegahsq)) %>%
+    mutate(Xenonxsec_ln = log(Xenonxsec)) %>%
+    mutate(pSDxsec_ln = log(pSDxsec)) %>%
+    mutate(gminus2_shiftln = log(gminus2 + 1.8e-9)) %>%
+    select(-BRbsg, -BrBsmumu, -Omegahsq, -Xenonxsec, -pSDxsec, -gminus2)
+  return(dtOut)
+}
+
+#function to plot data and binning of requested projection
+plotProj <- function(nProj, fullPath, dMatrix){
+  p <- fullPath[nProj][[1]]
+  d <- dMatrix %*% p
+  s <- scagnostics.default(d[,1],d[,2], bins= 50)
+  p1 <- ggplot()+
+    geom_point(data = as.data.frame(d), mapping=aes(x=V1, y=V2))
+  p2 <- ggplot()+
+    geom_point(data = as.data.frame(s$bins), mapping = aes(x=V1, y=V2, color=V3, size=V3))+
+    xlim(0,1000)+
+    ylim(0,1000)+
+    lims(colour = c(1,25), size = c(1,25))+
+    scale_color_gradientn(colors=wes_palette(name="Zissou"))+
+    scale_size(range = c(1,3))
+    grid.arrange(arrangeGrob(p1, widths = 2),arrangeGrob(p2, widths = 7),nrow=1)
+    return(s$s)
+}
+
+
+# get tour path
+load("data/fullPath_nObs7.RData")
+
+# read data file
+dt <- read.csv("data/obs_Bino.csv")
+
+## ---- scagnostics-summary
+
+n <- 4*100
+dfSummary = data.frame( Outlying=rep(0, n), Skewed=rep(0,n), Clumpy=rep(0,n),
+                 Sparse=rep(0,n),Striated=rep(0,n), Convex=rep(0,n),
[…]

### 2018-01-31 Draft/Draft.Rmd
Added guided tour example.

@@
 try out scagnostics measures as guided tour index
-FIXME still need to write script for this
+
+Example below uses Convex as index function.
+We show the evolution of the index on the interpolated guided tour (top) and on the selected planes without interpolation (bottom).
+Note that the interpolated projections do not smoothly increase the index value from one selected plane to the next, but shows the unstable nature of the index computation.
+
+```{r guided-tour, include=FALSE}
+```
+```{r guided-tour-plot, fig.height = 3, fig.cap = "(ref:guided-tour-plot)"}
+```
+
+Figure (ref:guided-tour-scatter) compares the scatter plots (left unbinned, right binned) of the start projection, the projection with the lowest index value (t=19 of the interpolated projections) and the final projection (t=28).
+```{r guided-tour-scatter, fig.height = 3, fig.cap = "(ref:guided-tour-scatter)"}
+```
+
+show computing time? for each index as function of sample size? (long to prepare such a plot..)
 

### 2018-01-31 Draft/Draft.tex
Added guided tour example.

@@
 
-try out scagnostics measures as guided tour index FIXME still need to
-write script for this
+try out scagnostics measures as guided tour index
+
+Example below uses Convex as index function. We show the evolution of
+the index on the interpolated guided tour (top) and on the selected
+planes without interpolation (bottom). Note that the interpolated
+projections do not smoothly increase the index value from one selected
+plane to the next, but shows the unstable nature of the index
+computation.
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/guided-tour-plot-1} \includegraphics[width=\textwidth]{figure/guided-tour-plot-2}
+
+}
+
+\caption[(ref:guided-tour-plot)]{(ref:guided-tour-plot)}\label{fig:guided-tour-plot}
+\end{figure}
+\end{CodeChunk}
+
+Figure (ref:guided-tour-scatter) compares the scatter plots (left
+unbinned, right binned) of the start projection, the projection with the
+lowest index value (t=19 of the interpolated projections) and the final
+projection (t=28).
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/guided-tour-scatter-1} \includegraphics[width=\textwidth]{figure/guided-tour-scatter-2} \includegraphics[width=\textwidth]{figure/guided-tour-scatter-3}
+
+}
+
+\caption[(ref:guided-tour-scatter)]{(ref:guided-tour-scatter)}\label{fig:guided-tour-scatter}
+\end{figure}
+\end{CodeChunk}
+
+show computing time? for each index as function of sample size? (long to
+prepare such a plot..)
 

### 2018-01-31 Draft/src/main.R

@@ -38,6 +38,13 @@ plotProj <- function(nProj, fullPath, dMatrix){
     return(s$s)
 }
 
+scagIndex <- function(scagType){
+  function(mat){
+    sR <- scagnostics.default(mat[,1],mat[,2])$s
+    return(sR[scagType])
+  }
+}
+
 
 # get tour path
 load("data/fullPath_nObs7.RData")
@@ -155,3 +162,51 @@ dfSmoothing <- dfSummary %>%
 ggplot(dfSmoothing) + 
   geom_line(aes(x = t, y = Outlying, colour = "red")) +
   stat_smooth(aes(x = t, y = Outlying), method = lm, formula = y ~ poly(x, 10), se = FALSE)
+
+## ---- guided-tour
+
+#reload the 500 point sample
+set.seed(22012018)
+dt_sample <- asLog(sample_n(dt, 500))
+thisMatrix <- rescale(as.matrix(select(dt_sample, -modelName)))
+
+#record guided tour path based on convex measure
+gT <- guided_tour(scagIndex("Convex"))
+guidedPath <- save_history(thisMatrix, gT, max_bases = 100)
+guidedOriginalOnly <- as.list(guidedPath)
+guidedPath <- as.list(interpolate(guidedPath))
+
+## ---- guided-tour-plot
+
+n <- length(guidedPath)
+convexDf <- data.frame(Convex=rep(0, n), t=rep(0,n))
+i <- 1
+for(pMatrix in guidedPath){
+  dProj <- thisMatrix %*% pMatrix
+  convexDf[i,] <- c(scagnostics.default(dProj[,1],dProj[,2], bins= 50)$s["Convex"],i)
+  i = i+1
+}
+
+i <- 1
+n <- length(guidedOriginalOnly)
+convexOriginalOnly <- data.frame(Convex=rep(0, n), t=rep(0,n))
+for(pMatrix in guidedOriginalOnly){
+  dProj <- thisMatrix %*% pMatrix
+  convexOriginalOnly[i,] <- c(scagnostics.default(dProj[,1],dProj[,2], bins= 50)$s["Convex"],i)
+  i = i+1
+}
+
+ggplot(convexDf, mapping = aes(x=t, y=Convex)) +
+  geom_line()+
+  ggtitle("Interpolated path")
+
+ggplot(convexOriginalOnly, mapping = aes(x=t, y=Convex)) +
+  geom_line()+
[…]

### 2018-01-31 Draft/Draft.Rmd
Adding mbgraphic examples.

@@
 FIXME load Katrins package and try the index functions she implemented
+
+Additions to the standard scagnostics measures have been implemented in the package mbgraphics, based on splines (spline2d) and distance correlation (dcor2d).
+Give definitions here..
+We repeat the study from before, we first test in Figure ADDREF the behaviour of the index values over 100 interpolated projections that are randomly selected via the grand tour algorithm.
+Note that here the measures are calculated on non-binned data.
+The observed behaviour is smooth (much in contrast to the behaviour of the scagnostics measures) and we next test the behaviour in a guided tour, comparing again also the first and final projection.
+(which index to select, will try both below...)
+```{r mbgraphics-indices, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
+```
+```{r guided-spline2d, include=FALSE}
+```
+```{r guided-spline2d-plot, fig.height=3, fig.cap="(ref:guided-spline2d-plot)"}
+```
+maybe move timing to the end, add also these indices in the same plot (instead of testing several samples, behaviour anyway similar)

### 2018-01-31 Draft/Draft.tex
Adding mbgraphic examples.

@@
 
+Additions to the standard scagnostics measures have been implemented in
+the package mbgraphics, based on splines (spline2d) and distance
+correlation (dcor2d). Give definitions here.. We repeat the study from
+before, we first test in Figure ADDREF the behaviour of the index values
+over 100 interpolated projections that are randomly selected via the
+grand tour algorithm. Note that here the measures are calculated on
+non-binned data. The observed behaviour is smooth (much in contrast to
+the behaviour of the scagnostics measures) and we next test the
+behaviour in a guided tour, comparing again also the first and final
+projection. (which index to select, will try both below\ldots{})
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/mbgraphics-indices-1}
+
+}
+
+\caption[(ref:mbgraphics-indices)]{(ref:mbgraphics-indices)}\label{fig:mbgraphics-indices}
+\end{figure}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-1} \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-2} \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-3}
+
+}
+
+\caption[(ref:guided-spline2d-plot)]{(ref:guided-spline2d-plot)}\label{fig:guided-spline2d-plot}
+\end{figure}
+\end{CodeChunk}
+
+maybe move timing to the end, add also these indices in the same plot
+(instead of testing several samples, behaviour anyway similar)
+
 \end{document}

### 2018-01-31 Draft/src/main.R

@@ -7,6 +7,7 @@ library(binostics)
 library(gridExtra)
 library(wesanderson) #color palette
 library(tictoc) #timer
+library(mbgraphic) #Katrins package
 
 # function to log transform part of data
 asLog <- function(dt){
@@ -42,6 +43,18 @@ scagIndex <- function(scagType){
   function(mat){
     sR <- scagnostics.default(mat[,1],mat[,2])$s
     return(sR[scagType])
+  }
+}
+
+splineIndex <- function(){
+  function(mat){
+    return(splines2d(mat[,1], mat[,2]))
+  }
+}
+
+dcorIndex <- function(){
+  function(mat){
+    return(dcor2d(mat[,1], mat[,2]))
   }
 }
 
@@ -210,3 +223,78 @@ s1 <- plotProj(1, guidedPath, thisMatrix)
 s2 <- plotProj(13, guidedPath, thisMatrix)
 s3 <- plotProj(28, guidedPath, thisMatrix)
 
+## ---- mbgraphics-indices
+
+n <- 200
+dfKatrin = data.frame( dcor2d=rep(0, n), splines2d=rep(0,n), t=rep(0,n), s=rep(0,n))
+i <- 1
+for(sampleSize in c(500,5000)){ #fix sample size here
+  t <- 1
+  set.seed(22012018) # all samples include same (smaller) dataset
+  dtSample <- asLog(sample_n(dt, sampleSize))
+  thisMatrix <- rescale(as.matrix(select(dtSample, -modelName)))
+  for(pMatrix in fullPath){
+    dProj <- thisMatrix %*% pMatrix
+    dcor2dV <- dcor2d(dProj[,1],dProj[,2])
+    splines2dV <- splines2d(dProj[,1],dProj[,2])
+    dfKatrin[i,] = c(dcor2dV, splines2dV, t, sampleSize)
+    i = i+1
+    t = t+1
+    if(t > 100) break # only consider first 100 projections
+  }
+}
+
+meltR = melt(dfKatrin, id = c("t", "s"))
+
+ggplot(data=meltR) + 
+  geom_line(aes(x=t,y=value, color=as.factor(s))) +
+  facet_wrap(~ variable,ncol=1,scales = "free") +
+  labs(x = "Projection", y = "Value", color = "Sample Size")
+
+## ---- guided-spline2d
+#reload the 500 point sample
+set.seed(22012018)
[…]

### 2018-01-31 Draft/Draft.Rmd
Added dcor2d eample.

@@
 (which index to select, will try both below...)
+It seems they both maximise correlation, this might not be too useful here...
 ```{r mbgraphics-indices, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
@@
 ```
+```{r guided-dcor2d, include=FALSE}
+```
+```{r guided-dcor2d-plot, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
+```
 maybe move timing to the end, add also these indices in the same plot (instead of testing several samples, behaviour anyway similar)

### 2018-01-31 Draft/Draft.tex
Added dcor2d eample.

@@
 behaviour in a guided tour, comparing again also the first and final
-projection. (which index to select, will try both below\ldots{})
+projection. (which index to select, will try both below\ldots{}) It
+seems they both maximise correlation, this might not be too useful
+here\ldots{}
 
@@
 
+\begin{CodeChunk}
+
+\begin{CodeOutput}
+            [,1]        [,2]
+[1,]  0.47851413 -0.54459454
+[2,]  0.01632788 -0.00642453
+[3,]  0.20956852 -0.15615632
+[4,] -0.21613392  0.22043552
+[5,] -0.26952654  0.31203441
+[6,] -0.27310129  0.26333539
+[7,] -0.72999721 -0.68094632
+\end{CodeOutput}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-1} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-2} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-3}
+
+}
+
+\caption[(ref:guided-dcor2d-plot)]{(ref:guided-dcor2d-plot)}\label{fig:guided-dcor2d-plot}
+\end{figure}
+\end{CodeChunk}
+
 maybe move timing to the end, add also these indices in the same plot

### 2018-01-31 Draft/src/main.R

@@ -257,7 +257,7 @@ set.seed(22012018)
 dt_sample <- asLog(sample_n(dt, 500))
 thisMatrix <- rescale(as.matrix(select(dt_sample, -modelName)))
 
-#record guided tour path based on convex measure
+#record guided tour path based on spline measure
 gT <- guided_tour(splineIndex())
 guidedPath <- save_history(thisMatrix, gT, max_bases = 100)
 guidedOriginalOnly <- as.list(guidedPath)
@@ -298,3 +298,54 @@ p2 <- ggplot()+
   geom_point(data = as.data.frame(dProj), mapping=aes(x=V1, y=V2))
 
 grid.arrange(arrangeGrob(p1),arrangeGrob(p2),nrow=1)
+
+## ---- guided-dcor2d
+#reload the 500 point sample
+set.seed(22012018)
+dt_sample <- asLog(sample_n(dt, 500))
+thisMatrix <- rescale(as.matrix(select(dt_sample, -modelName)))
+
+#record guided tour path based on spline measure
+gT <- guided_tour(dcorIndex())
+guidedPath <- save_history(thisMatrix, gT, max_bases = 100)
+guidedOriginalOnly <- as.list(guidedPath)
+guidedPath <- as.list(interpolate(guidedPath))
+
+## ---- guided-dcor2d-plot
+n <- length(guidedPath)
+dcorDf <- data.frame(dcor=rep(0, n), t=rep(0,n))
+i <- 1
+for(pMatrix in guidedPath){
+  dProj <- thisMatrix %*% pMatrix
+  dcorDf[i,] <- c(dcor2d(dProj[,1],dProj[,2]),i)
+  i = i+1
+}
+
+i <- 1
+n <- length(guidedOriginalOnly)
+dcorOriginalOnly <- data.frame(dcor=rep(0, n), t=rep(0,n))
+for(pMatrix in guidedOriginalOnly){
+  dProj <- thisMatrix %*% pMatrix
+  if(i==1) firstProj <- dProj
+  dcorOriginalOnly[i,] <- c(dcor2d(dProj[,1],dProj[,2]),i)
+  i = i+1
+}
+
+ggplot(dcorDf, mapping = aes(x=t, y=dcor)) +
+  geom_line()+
+  ggtitle("Interpolated path")
+
+ggplot(dcorOriginalOnly, mapping = aes(x=t, y=dcor)) +
+  geom_line()+
[…]

### 2018-01-31 Draft/Draft.Rmd
Small additions in the text.

--- 2018-01-31 Draft/Draft.Rmd
@@
 Contributions to the variable gminus2 can be negative or positive, we therefore add a small shift before the log transformation of this variable.
+Maybe include GGally pairs plot here?
 
@@
 ## Additional indices to consider
-FIXME load Katrins package and try the index functions she implemented
+
+Additions to the standard scagnostics measures have been implemented in the package mbgraphics, based on splines (spline2d) and distance correlation (dcor2d).
+Give definitions here..
+We repeat the study from before, we first test in Figure ADDREF the behaviour of the index values over 100 interpolated projections that are randomly selected via the grand tour algorithm (the same projections as considered above).
+Note that here the measures are calculated on non-binned data.
+The general behaviour seems similar to the scagnostics monotonic index.
+The observed behaviour is smooth (much in contrast to the behaviour of the scagnostics measures) and we next test the behaviour in a guided tour, comparing again also the first and final projection.
+(which index to select, will try both below...)
+It seems they both maximise correlation, this might not be too useful?
+```{r mbgraphics-indices, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
+```
+```{r guided-spline2d, include=FALSE}
+```
+```{r guided-spline2d-plot, fig.height=3, fig.cap="(ref:guided-spline2d-plot)"}
+```
+```{r guided-dcor2d, include=FALSE}
+```
+```{r guided-dcor2d-plot, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
+```
+
+Physics: what is final projection, what is special about outlying points?
+
+maybe move timing to the end, add also these indices in the same plot (instead of testing several samples, behaviour anyway similar)

### 2018-01-31 Draft/Draft.tex
Small additions in the text.

@@
 therefore add a small shift before the log transformation of this
-variable.
+variable. Maybe include GGally pairs plot here?
 
@@
 
-FIXME load Katrins package and try the index functions she implemented
-
 Additions to the standard scagnostics measures have been implemented in
@@
 over 100 interpolated projections that are randomly selected via the
-grand tour algorithm. Note that here the measures are calculated on
-non-binned data. The observed behaviour is smooth (much in contrast to
-the behaviour of the scagnostics measures) and we next test the
-behaviour in a guided tour, comparing again also the first and final
-projection. (which index to select, will try both below\ldots{}) It
-seems they both maximise correlation, this might not be too useful
-here\ldots{}
+grand tour algorithm (the same projections as considered above). Note
+that here the measures are calculated on non-binned data. The general
+behaviour seems similar to the scagnostics monotonic index. The observed
+behaviour is smooth (much in contrast to the behaviour of the
+scagnostics measures) and we next test the behaviour in a guided tour,
+comparing again also the first and final projection. (which index to
+select, will try both below\ldots{}) It seems they both maximise
+correlation, this might not be too useful?
 
@@
 
+Physics: what is final projection, what is special about outlying
+points?
+
 maybe move timing to the end, add also these indices in the same plot

### 2018-02-01 Draft/Draft.Rmd
Small study to check dependence on gminus2 variable.

@@
 Moreover, we test the dependence on the sample size and on the outlier removal typically performed to make the scagnostics measures more robust.
-In addition the computing time is also important, as it might limit the dynamic viewing and require to first record a guided tour path which can
-later be viewed in a smooth display.
+In addition the computing time is also important, as it might limit the dynamic viewing and require to first record a guided tour path which can later be viewed in a smooth display.
 
@@
 It seems they both maximise correlation, this might not be too useful?
+Seems like [,1] basically -[,2] in the final projection (except for [7,])?
+probably we see only structure of gminus2 variable, which has very narrow peak, only few outlying points
+is there a better way of transforming it to spread out the structure?
+
 ```{r mbgraphics-indices, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
@@
 maybe move timing to the end, add also these indices in the same plot (instead of testing several samples, behaviour anyway similar)
+
+```{r gminus2-box, fig.height=3}
+```
+
+```{r guided-dcor2d-no-gminus2, include=FALSE}
+```
+```{r guided-dcor2d-plot-no-gminus2, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
+```

### 2018-02-01 Draft/Draft.tex
Small study to check dependence on gminus2 variable.

@@
 select, will try both below\ldots{}) It seems they both maximise
-correlation, this might not be too useful?
+correlation, this might not be too useful? Seems like {[},1{]} basically
+-{[},2{]} in the final projection (except for {[}7,{]})? probably we see
+only structure of gminus2 variable, which has very narrow peak, only few
+outlying points is there a better way of transforming it to spread out
+the structure?
 
@@
 
+\begin{CodeChunk}
+
+\begin{center}\includegraphics[width=\textwidth]{figure/gminus2-box-1} \end{center}
+
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{CodeOutput}
+            [,1]        [,2]
+[1,]  0.48848925 -0.37880896
+[2,]  0.09066938  0.10280711
+[3,]  0.54333417  0.80184027
+[4,] -0.05279536  0.00809942
+[5,] -0.47070661  0.10044669
+[6,] -0.48331484  0.43912584
+\end{CodeOutput}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-no-gminus2-1} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-no-gminus2-2} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-no-gminus2-3}
+
+}
+
+\caption[(ref:guided-dcor2d-plot)]{(ref:guided-dcor2d-plot)}\label{fig:guided-dcor2d-plot-no-gminus2}
+\end{figure}
+\end{CodeChunk}
+
 \end{document}

### 2018-02-01 Draft/src/main.R

@@ -349,3 +349,55 @@ grid.arrange(arrangeGrob(p1),arrangeGrob(p2),nrow=1)
 
 print(pMatrix)
 
+## ---- gminus2-box
+ggplot(dt_sample) + geom_boxplot(aes(x="", y=gminus2_shiftln))
+
+## ---- guided-dcor2d-no-gminus2
+#reload the 500 point sample
+set.seed(22012018)
+dt_sample <- asLog(sample_n(dt, 500))
+thisMatrix <- rescale(as.matrix(select(dt_sample, -modelName, -gminus2_shiftln)))
+
+#record guided tour path based on spline measure
+gT <- guided_tour(dcorIndex())
+guidedPath <- save_history(thisMatrix, gT, max_bases = 100)
+guidedOriginalOnly <- as.list(guidedPath)
+guidedPath <- as.list(interpolate(guidedPath))
+
+## ---- guided-dcor2d-plot-no-gminus2
+n <- length(guidedPath)
+dcorDf <- data.frame(dcor=rep(0, n), t=rep(0,n))
+i <- 1
+for(pMatrix in guidedPath){
+  dProj <- thisMatrix %*% pMatrix
+  dcorDf[i,] <- c(dcor2d(dProj[,1],dProj[,2]),i)
+  i = i+1
+}
+
+i <- 1
+n <- length(guidedOriginalOnly)
+dcorOriginalOnly <- data.frame(dcor=rep(0, n), t=rep(0,n))
+for(pMatrix in guidedOriginalOnly){
+  dProj <- thisMatrix %*% pMatrix
+  if(i==1) firstProj <- dProj
+  dcorOriginalOnly[i,] <- c(dcor2d(dProj[,1],dProj[,2]),i)
+  i = i+1
+}
+
+ggplot(dcorDf, mapping = aes(x=t, y=dcor)) +
+  geom_line()+
+  ggtitle("Interpolated path")
+
+ggplot(dcorOriginalOnly, mapping = aes(x=t, y=dcor)) +
+  geom_line()+
+  ggtitle("Non-interpolated guided tour planes")
+
+p1 <- ggplot()+
+  geom_point(data = as.data.frame(firstProj), mapping=aes(x=V1, y=V2))
+
+p2 <- ggplot()+
+  geom_point(data = as.data.frame(dProj), mapping=aes(x=V1, y=V2))
+
+grid.arrange(arrangeGrob(p1),arrangeGrob(p2),nrow=1)
+
+print(pMatrix)

### 2018-02-14 Draft/Draft.Rmd
restructuring paper

@@
 title:
-  formatted: "Scagnostics and Projection Pursuit"
+  formatted: "Using Scagnostics to find Interesting Projections of Multivariate Data"
   # If you use tex in the formatted title, also supply version without
@@
   # at least one keyword must be supplied
-  formatted: [scagnostics, tour, projection pursuit]
+  formatted: [scagnostics, tour, projection pursuit, multivariate data, statistical graphics, data visualisation]
   #plain:     [keywords, not capitalized, Java]
@@
 
+*** Motivating example
+
+```{r motivation}
+library(mvtnorm)
+x1 <- rmvnorm(100, mean=c(0,0), 
+             sigma=matrix(c(1, 0.8, 0.8, 1), ncol=2, byrow=T))
+x2 <- rmvnorm(100, mean=c(3,0), 
+             sigma=matrix(c(1, 0.8, 0.8, 1), ncol=2, byrow=T))
+x3 <- rmvnorm(100, mean=c(0,3), 
+             sigma=matrix(c(1, 0.8, 0.8, 1), ncol=2, byrow=T))
+df <- data.frame(x1=rnorm(300),
+                 x2=rnorm(300),
+                 x3=rnorm(300),
+                 x4=rnorm(300), 
+                 x5=scale(c(x1[,1], x2[,1], x3[,1])), 
+                 x6=scale(c(x1[,2], x2[,2], x3[,2])), 
+                 cl=c(rep("A",100), rep("B", 100), rep("C",100)))
+ggplot(df, aes(x=x5, y=x6, colour=cl)) + geom_point() + theme(aspect.ratio=1)
+```
+
+```{r}
+library(tourr)
+quartz()
+col <- c(rep("#f0a800", 100), rep("#486030", 100), rep("#c03018", 100))
+animate(df[,1:6], guided_tour(lda_pp(df$cl)),
+        display_xy(axes="bottomleft", col=col))
+```
+
+```{r}
+library(geozoo)
+
+```
+
+*** Explanation of tour algorithm and projection pursuit and software
+
+*** Explanation of scagnostics and software
+
 A grand tour provides dynamic projections from the high dimensions into some low dimensional representation, in particular also as a
@@
 
-# Scagnostics measures
+# Investigation of scagnostics
+
+## Scagnostic discreteness
+
+*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference
+
+*** katrin's work
 
@@
 
-# Analysis setup
+## Speed
+
+# Modification to scagnostics
+
+## Smoothing
+
+## Jittering
+
+## Tour optimisation
+
+# Scagnostics indices
+
+## Individual
+
+## Two group differences
+
+## Multiple group differences
+
+# Application to particle physics data
 
@@
 
-## Scagnositics of grand tour projections
+## Scagnostics of grand tour projections
 
-```{r scagnostics-summary, fig.height= 14, fig.width=9, fig.cap = "(ref:scagnostics-summary)"}
+```{r scagnostics-summary, fig.height= 14, fig.width=9, fig.cap = "(ref:scagnostics-summary)", eval=FALSE}
 ```
@@
 
-```{r scatter-binning, fig.height=2, fig.cap = "(ref:scatter-binning)"}
+```{r scatter-binning, eval=FALSE, fig.height=2, fig.cap = "(ref:scatter-binning)"}
 ```
 
-```{r scatter-table}
+```{r scatter-table, eval=FALSE}
 ```
@@
 
-```{r scagnositcs-timing, fig.cap = "(ref:scagnostics-timing)"}
+```{r scagnositcs-timing, eval=FALSE, fig.cap = "(ref:scagnostics-timing)"}
 ```
@@
 (Below just an example of using ggplot smoothing to display what kind of behaviour we want to find. It follows the general trend without capturing the spikes.)
-```{r smoothing, fig.cap = "(ref:smooting)"}
+```{r smoothing, eval=FALSE, fig.cap = "(ref:smooting)"}
 ```
@@
 
-```{r guided-tour, include=FALSE}
+```{r guided-tour, eval=FALSE, include=FALSE}
 ```
-```{r guided-tour-plot, fig.height = 3, fig.cap = "(ref:guided-tour-plot)"}
+```{r guided-tour-plot, eval=FALSE, fig.height = 3, fig.cap = "(ref:guided-tour-plot)"}
 ```
@@
 Figure (ref:guided-tour-scatter) compares the scatter plots (left unbinned, right binned) of the start projection, the projection with the lowest index value (t=19 of the interpolated projections) and the final projection (t=28).
-```{r guided-tour-scatter, fig.height = 3, fig.cap = "(ref:guided-tour-scatter)"}
+```{r guided-tour-scatter, eval=FALSE, fig.height = 3, fig.cap = "(ref:guided-tour-scatter)"}
 ```
@@
 
-```{r mbgraphics-indices, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
+```{r mbgraphics-indices, eval=FALSE, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
 ```
-```{r guided-spline2d, include=FALSE}
+```{r guided-spline2d, eval=FALSE, include=FALSE}
 ```
-```{r guided-spline2d-plot, fig.height=3, fig.cap="(ref:guided-spline2d-plot)"}
+```{r guided-spline2d-plot, eval=FALSE, fig.height=3, fig.cap="(ref:guided-spline2d-plot)"}
 ```
-```{r guided-dcor2d, include=FALSE}
+```{r guided-dcor2d, eval=FALSE, include=FALSE}
 ```
-```{r guided-dcor2d-plot, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
+```{r guided-dcor2d-plot, eval=FALSE, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
 ```
@@
 
-```{r gminus2-box, fig.height=3}
+```{r gminus2-box, eval=FALSE, fig.height=3}
 ```
 
-```{r guided-dcor2d-no-gminus2, include=FALSE}
+```{r guided-dcor2d-no-gminus2, eval=FALSE, include=FALSE}
 ```
-```{r guided-dcor2d-plot-no-gminus2, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
+```{r guided-dcor2d-plot-no-gminus2, eval=FALSE, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
 ```

### 2018-02-14 Draft/Draft.tex
restructuring paper

--- 2018-01-31 Draft/Draft.tex
@@
 }
-\title{Scagnostics and Projection Pursuit}
+\title{Using Scagnostics to find Interesting Projections of Multivariate Data}
 
@@
 
-\Keywords{scagnostics, tour, projection pursuit}
+\Keywords{scagnostics, tour, projection pursuit, multivariate data, statistical graphics, data visualisation}
 
@@
 
-\subsection{Scagnositics of grand tour
-projections}\label{scagnositics-of-grand-tour-projections}
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/scagnostics-summary-1}
-
-}
-
-\caption[(ref:scagnostics-summary)]{(ref:scagnostics-summary)}\label{fig:scagnostics-summary}
-\end{figure}
-\end{CodeChunk}
+\subsection{Scagnostics of grand tour
+projections}\label{scagnostics-of-grand-tour-projections}
 
@@
 
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/scatter-binning-1} \includegraphics[width=\textwidth]{figure/scatter-binning-2} \includegraphics[width=\textwidth]{figure/scatter-binning-3}
-
-}
-
-\caption[(ref:scatter-binning)]{(ref:scatter-binning)}\label{fig:scatter-binning}
-\end{figure}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-
-\begin{CodeOutput}
-                   s1          s2          s3
-Outlying  0.275676305 0.163548110 0.270907044
-Skewed    0.549038842 0.575422054 0.583569727
-Clumpy    0.004123711 0.008163265 0.004175365
-Sparse    0.038765964 0.042779580 0.040551326
-Striated  0.059405941 0.060000000 0.037433155
-Convex    0.611714549 0.602591732 0.546399191
-Skinny    0.483884014 0.495557112 0.564627720
-Stringy   0.424567811 0.368764181 0.329434667
-Monotonic 0.077330543 0.094217913 0.103036354
-\end{CodeOutput}
-\end{CodeChunk}
-
 Another consideration relevant for the purpose of the guided tour is the
@@
 
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/scagnositcs-timing-1}
-
-}
-
-\caption[(ref:scagnostics-timing)]{(ref:scagnostics-timing)}\label{fig:scagnositcs-timing}
-\end{figure}
-\end{CodeChunk}
-
 \subsection{Smooth transitions}\label{smooth-transitions}
@@
 the general trend without capturing the spikes.)
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/smoothing-1}
-
-}
-
-\caption[(ref:smooting)]{(ref:smooting)}\label{fig:smoothing}
-\end{figure}
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/guided-tour-plot-1} \includegraphics[width=\textwidth]{figure/guided-tour-plot-2}
-
-}
-
-\caption[(ref:guided-tour-plot)]{(ref:guided-tour-plot)}\label{fig:guided-tour-plot}
-\end{figure}
-\end{CodeChunk}
-
 Figure (ref:guided-tour-scatter) compares the scatter plots (left
@@
 projection (t=28).
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/guided-tour-scatter-1} \includegraphics[width=\textwidth]{figure/guided-tour-scatter-2} \includegraphics[width=\textwidth]{figure/guided-tour-scatter-3}
-
-}
-
-\caption[(ref:guided-tour-scatter)]{(ref:guided-tour-scatter)}\label{fig:guided-tour-scatter}
-\end{figure}
-\end{CodeChunk}
 
@@
 select, will try both below\ldots{}) It seems they both maximise
-correlation, this might not be too useful?
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/mbgraphics-indices-1}
-
-}
-
-\caption[(ref:mbgraphics-indices)]{(ref:mbgraphics-indices)}\label{fig:mbgraphics-indices}
-\end{figure}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-1} \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-2} \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-3}
-
-}
-
-\caption[(ref:guided-spline2d-plot)]{(ref:guided-spline2d-plot)}\label{fig:guided-spline2d-plot}
-\end{figure}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-
-\begin{CodeOutput}
-            [,1]        [,2]
-[1,]  0.47851413 -0.54459454
-[2,]  0.01632788 -0.00642453
-[3,]  0.20956852 -0.15615632
-[4,] -0.21613392  0.22043552
-[5,] -0.26952654  0.31203441
-[6,] -0.27310129  0.26333539
-[7,] -0.72999721 -0.68094632
-\end{CodeOutput}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-1} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-2} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-3}
-
-}
-
-\caption[(ref:guided-dcor2d-plot)]{(ref:guided-dcor2d-plot)}\label{fig:guided-dcor2d-plot}
-\end{figure}
-\end{CodeChunk}
+correlation, this might not be too useful? Seems like {[},1{]} basically
+-{[},2{]} in the final projection (except for {[}7,{]})? probably we see
+only structure of gminus2 variable, which has very narrow peak, only few
+outlying points is there a better way of transforming it to spread out
+the structure?
 

### 2018-02-14 Draft/src/main.R

@@ -3,7 +3,7 @@
 library(tourr)
 library(tidyverse)
 library(reshape2)
-library(binostics)
+library(binostics) #https://github.com/uschiLaa/binostics
 library(gridExtra)
 library(wesanderson) #color palette
 library(tictoc) #timer

### 2018-02-14 Draft/Draft.Rmd
got some optimisation working

@@
 
-```{r}
+```{r eval=FALSE}
 library(tourr)
@@
 
-```{r}
+```{r eval=FALSE}
+numbat <- read_csv("data/numbat.csv") %>% filter(group=="A")
 library(geozoo)
+x <- sphere.hollow(p=2, n=2000)$points
+colnames(x) <- c("x4", "x7")
+x <- data.frame(x1=scale(rnorm(2000)), 
+                x2=scale(rnorm(2000)), 
+                x3=scale(rnorm(2000)),
+                x4=scale(x[,1]), 
+                x5=scale(rnorm(2000)), 
+                x6=scale(rnorm(2000)),
+                x7=scale(x[,2]), 
+                x8=scale(rnorm(2000)), 
+                x9=scale(rnorm(2000)),
+                x10=scale(rnorm(2000)), group="B")
+numbat <- bind_rows(numbat, x)
+ggplot(numbat, aes(x=x4, y=x7, col=group)) + 
+  geom_point() + facet_wrap(~group) + scale_colour_manual(values=c("#609048", "#c03018"))
+write_csv(numbat, path="data/numbat_circle.csv")
+```
 
+```{r eval=FALSE}
+library(tourr)
+quartz()
+col <- c(rep("#f0a800", 100), rep("#609048", 2100), rep("#c03018", 2000))
+animate(numbat[,1:10], 
+        guided_tour(lda_pp(numbat$group), search_f = search_better_random),
+        display_xy(axes="bottomleft", col=col))
 ```

### 2018-02-14 Draft/src/main.R

@@ -46,6 +46,23 @@ scagIndex <- function(scagType){
   }
 }
 
+data <- as.matrix(numbat_only[,c(4,6,7)])
+scagIndex2 <- function(proj) {
+  proj <- matrix(proj, ncol=2, byrow=FALSE)
+  tourr:::orthonormalise(proj)
+  mat <- as.matrix(numbat_only[,c(4,6,7)]) %*% proj
+  scagType <- "Clumpy"
+  sR <- scagnostics.default(mat[,1],mat[,2])$s
+  return(sR[scagType])
+}
+
+par0 <- c(1,0,0,0,1,0)
+lower <- rep(-1,6)
+upper <- rep(1, 6)
+hjkb(par0, scagIndex2, lower=lower, upper=upper,
+     control = list(maximize=TRUE, tol=0.001, 
+                    target=0.08, info=TRUE))
+  
 splineIndex <- function(){
   function(mat){
     return(splines2d(mat[,1], mat[,2]))
@@ -184,10 +201,53 @@ dt_sample <- asLog(sample_n(dt, 500))
 thisMatrix <- rescale(as.matrix(select(dt_sample, -modelName)))
 
 #record guided tour path based on convex measure
-gT <- guided_tour(scagIndex("Convex"))
+gT <- guided_tour(scagIndex("Convex"), search_f = tourr:::search_better, method="geodesic")
 guidedPath <- save_history(thisMatrix, gT, max_bases = 100)
 guidedOriginalOnly <- as.list(guidedPath)
 guidedPath <- as.list(interpolate(guidedPath))
+
+quartz()
+animate(thisMatrix, guided_tour(scagIndex("Convex"), search_f = tourr:::search_better, method="geodesic"),
+        display_xy(axes="bottomleft"))
+
+numbat_circle <- read_csv("data/numbat.csv")
+numbat_only <- numbat %>% filter(group=="A") %>% select(-group)
+circle_only <- numbat %>% filter(group=="B") %>% select(-group)
+animate(numbat_only[,c(4,6,7)], guided_tour(scagIndex("Clumpy"), search_f = tourr:::search_better, method="geodesic", alpha=2, cooling=0.8),
+        display_xy(axes="bottomleft"))
[…]

### 2018-02-14 Draft/src/main.R
Check to avoid crash in optimisation

@@ -49,7 +49,8 @@ scagIndex <- function(scagType){
 data <- as.matrix(numbat_only[,c(4,6,7)])
 scagIndex2 <- function(proj) {
   proj <- matrix(proj, ncol=2, byrow=FALSE)
-  tourr:::orthonormalise(proj)
+  proj <- tourr:::orthonormalise(proj)
+  if(any(is.nan(proj))) return(0)
   mat <- as.matrix(numbat_only[,c(4,6,7)]) %*% proj
   scagType <- "Clumpy"
   sR <- scagnostics.default(mat[,1],mat[,2])$s

### 2018-02-15 Draft/src/main.R
tweaks to optimisation algorithm

@@ -8,6 +8,7 @@ library(gridExtra)
 library(wesanderson) #color palette
 library(tictoc) #timer
 library(mbgraphic) #Katrins package
+library(dfoptim)
 
 # function to log transform part of data
 asLog <- function(dt){
@@ -51,19 +52,23 @@ scagIndex2 <- function(proj) {
   proj <- matrix(proj, ncol=2, byrow=FALSE)
   proj <- tourr:::orthonormalise(proj)
   if(any(is.nan(proj))) return(0)
-  mat <- as.matrix(numbat_only[,c(4,6,7)]) %*% proj
-  scagType <- "Clumpy"
+  mat <- as.matrix(numbat_only) %*% proj
+  scagType <- "Monotonic"
   sR <- scagnostics.default(mat[,1],mat[,2])$s
   return(sR[scagType])
 }
 
-par0 <- c(1,0,0,0,1,0)
-lower <- rep(-1,6)
-upper <- rep(1, 6)
-hjkb(par0, scagIndex2, lower=lower, upper=upper,
+par0 <- c(1,0,0,0,0,0,0,0,0,0, 0,1,0,0,0,0,0,0,0,0)
+lower <- rep(-1,20)
+upper <- rep(1, 20)
+opt <- hjkb(par0, scagIndex2, lower=lower, upper=upper,
      control = list(maximize=TRUE, tol=0.001, 
-                    target=0.08, info=TRUE))
+                    info=TRUE))
   
+prj <- tourr:::orthonormalise(matrix(opt$par, ncol=2, byrow=FALSE))
+df <- as.data.frame(as.matrix(numbat_only) %*% prj)
+ggplot(df, aes(x=V1, y=V2)) + geom_point()
+
 splineIndex <- function(){
   function(mat){
     return(splines2d(mat[,1], mat[,2]))
@@ -212,8 +217,8 @@ animate(thisMatrix, guided_tour(scagIndex("Convex"), search_f = tourr:::search_b
         display_xy(axes="bottomleft"))
 
 numbat_circle <- read_csv("data/numbat.csv")
-numbat_only <- numbat %>% filter(group=="A") %>% select(-group)
-circle_only <- numbat %>% filter(group=="B") %>% select(-group)
+numbat_only <- numbat_circle %>% filter(group=="A") %>% select(-group)
+circle_only <- numbat_circle %>% filter(group=="B") %>% select(-group)
[…]

### 2018-02-15 scagnostics.Rmd
start on ASA formatted paper

---
title: Using Scagnostics to find Interesting Projections of Multivariate Data

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Ursula Laa
  thanks: The authors gratefully acknowledge the support of the Australian Research Council
  affiliation: Department of Physics, Monash University
  email: \email{ursula.laa@monash.edu}

- name: Dianne Cook
  affiliation: Department of Econometrics and Business Statistics, Monash University

keywords:
- tour
- projection pursuit 
- statistical graphics
- data visualisation
- exploratory data analysis
- high-dimensional data

abstract: |
  Projection pursuit describes a procedure for searching high-dimensional data for "interesting" low-dimensional projections via the optimization of a criterion function called the projection pursuit index. Most indexes developed focus on finding projections with cluster structure, outliers, or separations between known groups. Here, we are interested in finding projections where the shapes are odd, or where known groups have different shapes despite having the same mean and variance. The new indexes are based on scagnostics, which were originally developed to choose pairs of interesting variables from a large collection. An examination of the properties of the scagnostics is included, and code in the form of an R package is provided.

bibliography: bibliography.bib
output: rticles::asa_article
---

# Introduction

The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high-(say $p$)dimensional data for "interesting" low- dimensional projections ($k=1$ or $2$ usually). The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $k$-dimensional projection of $p$-dimensional data. This criterion function is optimized over the space of all $k$-dimensional projections of $p$-space, searching for both global and local maxima. It is hoped that the resulting solutions reveal low-dimensional structure in the data not necessarily found by methods such as principal component analysis.

A large number of projection pursuit indices have been developed [@f87] [@CBC92] [@Lee05] [@AHC02] [@CEM:CEM2568] [@JS87] [@5508437].

*** Motivating example

*** Explanation of tour algorithm and projection pursuit and software

*** Explanation of scagnostics and software

# Investigation of scagnostics

## Scagnostic discreteness

*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference

*** katrin's work

## Speed

# Modification to scagnostics

## Smoothing

## Jittering

## Tour optimisation

# Scagnostics indices

## Individual

## Two group differences

## Multiple group differences

# Application to particle physics data

## Data description

## Software setup

## Scagnostics of grand tour projections

# Discussion

# References

### 2018-02-15 scagnostics.Rmd
adding to intro

@@
   affiliation: Department of Econometrics and Business Statistics, Monash University
+
+- name: Heike Hofmann
+  affiliation: Department of Statistics, Iowa State University
+
+- name: Hadley Wickham
+  affiliation: RStudio
+
+- name: Antony Unwin
+  affiliation: Department of Mathematics, Augsburg University
+
+- name: Katrin Grimm
+  affiliation: ???
 
@@
 A large number of projection pursuit indices have been developed [@f87] [@CBC92] [@Lee05] [@AHC02] [@CEM:CEM2568] [@JS87] [@5508437].
+Many of these are developed to detect departure from multivariate normal, which includes clusters or outliers, or nonlinearity.
+Some have been developed to detect separations between known groups.
+However, if one wants to find unusual shapes, or shape differences between groups there are no options.
+One reason, is that it is hard, because defining what is meant by shape is not straightforward.
+An approach that seems plausible begins with scagnostics, developed by @scag, and explored in @WW08.
 
-*** Motivating example
+*** Motivating example ??
+
+Scagnostics are scatterplot diagnostics.
+They were primarily developed to extract pairs of interesting variables from a large collection.
+XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia)
 

### 2018-02-15 scagnostics.Rmd
updates to intro

@@
 They were primarily developed to extract pairs of interesting variables from a large collection.
-XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia)
+XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia).
+Scagnostics are available in two R packages [@LWscagR] and [@HWscagR].
+This work is built upon the scagnostics package by @HWscagR because the underlying base is built using C, works across platforms.
 
+Projection pursuit is typically 
 *** Explanation of tour algorithm and projection pursuit and software
@@
 *** Explanation of scagnostics and software
+
+The projectionpursuitindex, a function of all possible projections
+of the data, invariablyhas many "hills and valleys" and "knife-edgeridges"because of
+the varying shape of underlyingdensity estimates from one projectionto the next.
+
+The paper is accompanied by the R package "binostics" available on CRAN, and new indices have been implemented in the "tourr" package.
 

### 2018-02-15 scagnostics.Rmd
done with intro

@@
 XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia).
-Scagnostics are available in two R packages [@LWscagR] and [@HWscagR].
+Scagnostics are available in two R packages @LWscagR and @HWscagR.
 This work is built upon the scagnostics package by @HWscagR because the underlying base is built using C, works across platforms.
 
-Projection pursuit is typically 
-*** Explanation of tour algorithm and projection pursuit and software
+A projection pursuit index, a function of all possible projections of the data, invariably has many "hills and valleys" and "knife-edge ridges" because of the varying shapes in the underlying density of observations from one projection to the next.
+From an exploratory data analysis perpective it is interesting to combine numerical optimisation with visualisation to watch the structure of the data moving into a maximum, to jump away from this projection and follow the optimisation again, invariably moving from local maxima to local maxima, and perhaps even a global maximum.
+Each of these maxima can reveal different information about the data.
+Projection pursuit optimisation combined with geodesic interpolations between projections provided by a tour [@As85], is called a projection pursuit guided tour [@CBCH94].
+Software to run a projection pursuit guided tour is available in the R package, "tourr" [@tourr].
 
-*** Explanation of scagnostics and software
-
-The projectionpursuitindex, a function of all possible projections
-of the data, invariablyhas many "hills and valleys" and "knife-edgeridges"because of
-the varying shape of underlyingdensity estimates from one projectionto the next.
-
-The paper is accompanied by the R package "binostics" available on CRAN, and new indices have been implemented in the "tourr" package.
+The paper investigates the nature of scagnostics, particularly when considered as candidates for projection pursuit indices (Section \ref{sec:investigate})).
+It develops indexes for finding shapes as defined by the scagnostics, and for finding different shapes between multiple groups (Section \ref{sec:indexes}).
+Changes to the optimisation of the projection pursuit guided tour were needed to accommodate some of the complexities of the scagnostic indexes.
+The new guided tour methods are applied to contemporary problems from paarticle physics (Section \ref{sec:phys}).
+The paper is accompanied by the R package "binostics" available on CRAN, and the new indices have been implemented in the "tourr" package.
 
 # Investigation of scagnostics
+\label{sec:investigate}
 
@@
 # Scagnostics indices
+\label{sec:indexes}
 
@@
 # Application to particle physics data
+\label{sec:phys}
 

### 2018-02-15 testMultipleFits.R
small script generalising and testing optimisation

@@ -1,0 +1,43 @@
+
+numbat_circle <- read_csv("Draft/data/numbat.csv")
+numbat_only <- numbat_circle %>% filter(group=="A") %>% select(-group)
+circle_only <- numbat_circle %>% filter(group=="B") %>% select(-group)
+
+scagIndex2 <- function(proj,dataF,scagType) {
+  proj <- matrix(proj, ncol=2, byrow=FALSE)
+  proj <- tourr:::orthonormalise(proj)
+  if(any(is.nan(proj))) return(0)
+  mat <- as.matrix(dataF) %*% proj
+  #scagType <- "Monotonic"
+  sR <- scagnostics.default(mat[,1],mat[,2])$s
+  return(sR[scagType])
+}
+
+ix <- length(circle_only)
+ntries <- 10
+
+lower <- rep(-1,2*ix)
+upper <- rep(1, 2*ix)
+
+idxV <- rep(0,ntries)
+prjL <- list()
+
+i <- 1
+while(i<=ntries){
+  par0 <- runif(2*ix, min = 0, max = 1)
+  opt <- hjkb(par0, scagIndex2, lower=lower, upper=upper,
+              control = list(maximize=TRUE, tol=0.001, 
+                             info=TRUE),dataF=circle_only, scagType="Skinny")
+  idxV[i] <- opt$value
+  prjL[[i]] <- tourr:::orthonormalise(matrix(opt$par, ncol=2, byrow=FALSE))
+  i <- i+1
+}
+
+i1 <- order(idxV)
+
+print(idxV[i1])
+
+plannedT <- planned_tour(prjL[i1])
+quartz()
+animate(numbat_only, plannedT)
+

### 2018-02-15 scagnostics.Rmd
filling in some of the investigation

@@
 
+```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+library(knitr)
+opts_chunk$set(
+  warning = FALSE, message = FALSE, echo = FALSE, 
+  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
+  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
+  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
+)
+```
+
 # Introduction
@@
 
-## Scagnostic discreteness
+The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
+If there are 100 variables, the dimension will be $4950x9$.
+It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting.
+It can also be used to do more in-depth exploratory analysis, and is interesting to study the distribution of the scagnostics resulting from the data, as a data set in itself.
+In a sense this the data's "fingerprint".
+
+To use scagnostics for projection pursuit requires that each of the measures for functions that are relatively smooth over the space of all 2D projections, and that they can be computed rapidly enough that the display can be updated many times a second during optimisation.
+
+## Smoothness
 
@@
 
+XXX Probably use a simple benchmark data to illustrate the issues, maybe mtcars that comes with the scagnostics package
+
+```{r load}
+library(tourr)
+library(tidyverse)
+library(reshape2)
+library(binostics) #https://github.com/uschiLaa/binostics
+library(gridExtra)
+library(wesanderson) #color palette
+library(tictoc) #timer
+library(mbgraphic) #Katrins package
+library(dfoptim)
+```
+
+```{r util}
+scagIndex <- function(scagType){
+  function(mat){
+    sR <- scagnostics.default(mat[,1],mat[,2])$s
+    return(sR[scagType])
+  }
+}
+```
+
+```{r tourpath}
+data(mtcars)
+t1 <- save_history(mtcars, grand_tour(2), max=4)
+t1full <- as.list(interpolate(t1))
+n <- length(t1full)
+sc <- NULL
+for (i in 1:n) {
+  dprj <- as.matrix(mtcars) %*% t1full[[i]]
+  sc <- rbind(sc, scagnostics(dprj))
+}
+sc <- as.data.frame(sc)
+sc$t <- 1:n
+sc <- sc %>% gather(index, value, -t)
+```
+
+```{r plotscag, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+ggplot(sc, aes(x=t, y=value)) + geom_line() + 
+  facet_wrap(~index, ncol=1, scales = "free_y")
+```
+
+## Effect of binning
+
 ## Speed
 
-# Modification to scagnostics
+# Possible modifications
 

### 2018-02-27 scagnostics.Rmd
Started adding Geozoo as example.

@@
 
-The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
-If there are 100 variables, the dimension will be $4950x9$.
-It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting.
-It can also be used to do more in-depth exploratory analysis, and is interesting to study the distribution of the scagnostics resulting from the data, as a data set in itself.
-In a sense this the data's "fingerprint".
-
-To use scagnostics for projection pursuit requires that each of the measures for functions that are relatively smooth over the space of all 2D projections, and that they can be computed rapidly enough that the display can be updated many times a second during optimisation.
-
-## Smoothness
-
-*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference
-
-*** katrin's work
-
-XXX Probably use a simple benchmark data to illustrate the issues, maybe mtcars that comes with the scagnostics package
-
 ```{r load}
@@
 library(dfoptim)
+library(GGally)
+library(geozoo)
 ```
+
+The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
+If there are 100 variables, the dimension will be $4950x9$.
+It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting.
+It can also be used to do more in-depth exploratory analysis, and is interesting to study the distribution of the scagnostics resulting from the data, as a data set in itself.
+In a sense this the data's "fingerprint".
+
+To use scagnostics for projection pursuit requires that each of the measures for functions that are relatively smooth over the space of all 2D projections, and that they can be computed rapidly enough that the display can be updated many times a second during optimisation.
+
+## Dataset
+
+As a test dataset we generate high dimensional geometrical shapes using the geozoo package (REF).
+Concretely we consider five dimensional objects, and we combine a hollow shpere with radius 1 and a solid torus with radii (0.5, 0.5, 0.1,0.1).
+The resulting distributions can be seen in the scatter plot matrix below.
+In the following we combine both groups to study the resulting non-standard distribution.
+
+```{r geozooDist, fig.width=6, fig.height=7, fig.cap="Scatter plot matrix of the five dimensional shapes, randomly selecting 100 points in each group.
+Group 1 (2) contains points selected on a hollow sphere (solid torus)."}
+hSphere <- sphere.hollow(p=5,n=100)$points
+sTorus <- torus(5,100, radius = c(0.5,0.5,0.1,0.1))$points
+geoDF <- bind_rows(as.data.frame(hSphere),as.data.frame(sTorus),.id = "Group")
+ggscatmat(geoDF, columns = 2:6, color="Group")
+```
+
+## Smoothness
+
+We first study the smoothness of the scagnostics measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path.
+The temporal trace of all nine measures is shown below, where each time step refers to an interpolation step with a target distance between interpolated bases is set to 0.05 radians.
+We observe sharp jumps in all nine measures, note however the large differences in overall scale between the measures.
+In particular "Clumpy", "Monotonic", "Sparse" and "Striated" show little sensitivity, as could be expected from the distributions in Figure (REF).
 
@@
 
-```{r tourpath}
-data(mtcars)
-t1 <- save_history(mtcars, grand_tour(2), max=4)
+```{r tourpathGeozoo}
+geoCombined <- select(geoDF,-Group)
+t1 <- save_history(geoCombined, grand_tour(2), max=4)
 t1full <- as.list(interpolate(t1))
@@
 for (i in 1:n) {
-  dprj <- as.matrix(mtcars) %*% t1full[[i]]
+  dprj <- as.matrix(geoCombined) %*% t1full[[i]]
   sc <- rbind(sc, scagnostics(dprj))
@@
 sc$t <- 1:n
+scPairs <- sc
 sc <- sc %>% gather(index, value, -t)
@@
 
-```{r plotscag, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+```{r plotscagGeozoo, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
 ggplot(sc, aes(x=t, y=value)) + geom_line() + 
-  facet_wrap(~index, ncol=1, scales = "free_y")
+  facet_wrap(~index, ncol=1, scales = "free_y") +
+  geom_vline(mapping = aes(xintercept=15, color="red")) +
+  geom_vline(mapping = aes(xintercept=25, color="red")) +
+  guides(color=FALSE)
 ```
+
+```{r plotscagPairs, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections, as above.
+Scatter plot of scagnostics calculated on each projection.
+FIXME is this relevant?
+Maybe remove this plot."}
+ggpairs(scPairs, columns = 1:9)
+```
+
+Two projections, t=15 and t=25, have been marked in Figure (can I add a reference?), as examples exhibiting jumps in several measures with respect to the previous and/or next projection.
+We compare the observed behaviour to the scatter plots obtained from the projected data in the Figure below.
+
+ADD Figure next
+
+## Smoothness
+
+*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference
+
+*** katrin's work
+
+XXX Probably use a simple benchmark data to illustrate the issues, maybe mtcars that comes with the scagnostics package
 

### 2018-03-13 scagnostics.Rmd
Bigger dataset to study temporal trace, differences in projections and binning

@@
 
-Scagnostics are scatterplot diagnostics.
-They were primarily developed to extract pairs of interesting variables from a large collection.
+Scagnostics are scatterplot diagnostics. They were primarily developed to extract pairs of interesting variables from a large collection.
 XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia).
-Scagnostics are available in two R packages @LWscagR and @HWscagR.
-This work is built upon the scagnostics package by @HWscagR because the underlying base is built using C, works across platforms.
+Nine graph-theoretic measures have been defined to characterise scatterplots.
+They are calculated from a set of elementary building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST).
+For definiteness all variables are rescaled to a [0,1] intervall, such that all measures will take values between 0 and 1 as well.
+The following measures have been defined:
+
+* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantil edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long.
+The outlying measure is then calculated as $c_{outlying} = \frac{\mathrm{Total\ length\ of\ long\ edges}}{\mathrm{Total\ length\ of\ all\ edges}}$.
+Note that outlying points are not considered in the calculation of the other measures.
+* Skewed: Skewness is defined via the distribution of edge length in the MST, $q_{skew}=\frac{q_{90}-q{50}}{q_{90}-q_{10}}$.
+This is inverted as $c_{skew}= 1-w(1-q_{skew})$ to account for binning effects that result in decreasing values of $q_{skew}$ with increasing n.
+* Sparse:
+
+Note that in practice, for efficiency and robustness, the measures are calculated from binned data, and after removing outliers.
+Moreover, scaling factors are used to mitigate dependence on sample size.
+Scagnostics are available in two R packages @LWscagR and @HWscagR. This work is built upon the scagnostics package by @HWscagR because the underlying base is built using C, works across platforms.
 
@@
 
-```{r geozooDist, fig.width=6, fig.height=7, fig.cap="Scatter plot matrix of the five dimensional shapes, randomly selecting 100 points in each group.
+```{r geozooDist, fig.width=6, fig.height=7, fig.cap="Scatter plot matrix of the five dimensional shapes, randomly selecting 500 points in each group.
 Group 1 (2) contains points selected on a hollow sphere (solid torus)."}
-hSphere <- sphere.hollow(p=5,n=100)$points
-sTorus <- torus(5,100, radius = c(0.5,0.5,0.1,0.1))$points
+hSphere <- sphere.hollow(p=5,n=500)$points
+sTorus <- torus(5,500, radius = c(0.5,0.5,0.1,0.1))$points
 geoDF <- bind_rows(as.data.frame(hSphere),as.data.frame(sTorus),.id = "Group")
@@
 ## Smoothness
+*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference
+
+*** katrin's work
+
+XXX Probably use a simple benchmark data to illustrate the issues, maybe mtcars that comes with the scagnostics package
 
@@
 The temporal trace of all nine measures is shown below, where each time step refers to an interpolation step with a target distance between interpolated bases is set to 0.05 radians.
-We observe sharp jumps in all nine measures, note however the large differences in overall scale between the measures.
-In particular "Clumpy", "Monotonic", "Sparse" and "Striated" show little sensitivity, as could be expected from the distributions in Figure (REF).
+We observe sharp jumps in almost all measures (with "Monotonic" being the only exception), note however the large differences in overall scale between the measures.
+In particular "Clumpy", "Monotonic", "Outlying", "Sparse" and "Striated" show little sensitivity, as could be expected from the distributions in Figure (REF).
 
@@
 }
+
+plotProj <- function(nProj, fullPath, dMatrix,legendOn, lim=75){
+  p <- fullPath[nProj][[1]]
+  d <- as.matrix(dMatrix) %*% p
+  s <- scagnostics.default(d[,1],d[,2], bins= 50)
+  p1 <- ggplot()+
+    geom_point(data = as.data.frame(d), mapping=aes(x=V1, y=V2))
+  p2 <- ggplot()+
+    geom_point(data = as.data.frame(s$bins), mapping = aes(x=V1, y=V2, color=V3, size=V3))+
+    xlim(0,1000)+
+    ylim(0,1000)+
+    expand_limits(colour = c(1,lim), size = c(1,lim))+
+    scale_color_gradientn(colors=wes_palette(name="Zissou"))+
+    scale_size(range = c(1,3))
+    if(legendOn==0){p2 = p2 + guides(color=FALSE, size=FALSE)}
+    grid.arrange(arrangeGrob(p1, widths = 2),arrangeGrob(p2, widths = 7),nrow=1)
+    return(s$s)
+}
 ```
@@
 ```{r tourpathGeozoo}
+set.seed(272829)
 geoCombined <- select(geoDF,-Group)
@@
   facet_wrap(~index, ncol=1, scales = "free_y") +
-  geom_vline(mapping = aes(xintercept=15, color="red")) +
-  geom_vline(mapping = aes(xintercept=25, color="red")) +
+  geom_vline(mapping = aes(xintercept=24, color="red")) +
+  geom_vline(mapping = aes(xintercept=39, color="red")) +
   guides(color=FALSE)
@@
 Scatter plot of scagnostics calculated on each projection.
+Observations: discreetness of outlying and striated, anti-correlation of convex and skinny.
 FIXME is this relevant?
@@
 
-Two projections, t=15 and t=25, have been marked in Figure (can I add a reference?), as examples exhibiting jumps in several measures with respect to the previous and/or next projection.
+Two projections, t=24 and t=39, have been marked in Figure (can I add a reference?), as examples exhibiting jumps in several measures with respect to the previous and/or next projection.
+In particular the outlying measure indicates that at least one outlier has been identified in the projection at t=24, but not the projections before or after.
+The removal of the point(s) will likely impact the calculation of other measures, and indeed we even see a small bump in the otherwise smooth measure "monotonic" (note however that the same is not true in other projections with outlying points).
+The projection at t=39 has been selected because of the sharp jumps in the measures "skewed", "stringy" and "striated".
 We compare the observed behaviour to the scatter plots obtained from the projected data in the Figure below.
+As expected the projections look similar over the 3 interpolation steps, small differences in the binned distribution can however be observed.
+Certainly the sometimes large differences (jumps of 0.1 absolute change in a measure from one projection to the next) are not indicating interesting differences in the projected view of the data.
 
-ADD Figure next
+```{r plotProj15, fig.width=6, fig.height=2, fig.cap="Projected data at t = 23, 24, 25 (top to bottom)."}
+s1 <- plotProj(23, t1full, geoCombined, legendOn=0)
+s2 <- plotProj(24, t1full, geoCombined, legendOn=0)
+s3 <- plotProj(25, t1full, geoCombined, legendOn=0)
+```
 
-## Smoothness
-
-*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference
-
-*** katrin's work
-
-XXX Probably use a simple benchmark data to illustrate the issues, maybe mtcars that comes with the scagnostics package
+```{r plotProj25, fig.width=6, fig.height=2, fig.cap="Projected data at t = 38, 39, 40 (top to bottom).
+The original data points are shown on the left, the binned data, used for the calculation of the scagnostics measures is shown on the right."}
+s1 <- plotProj(38, t1full, geoCombined, legendOn=0)
+s2 <- plotProj(39, t1full, geoCombined, legendOn=0)
+s3 <- plotProj(40, t1full, geoCombined, legendOn=0)
+```
 
 ## Effect of binning
+
+Two possible sources of discreet jumps, from binning and discreetness of the underlying objects, in particular the minimal spanning tree.
+Can we see how mst looks for each projection to understand this?
 

### 2018-03-13 testMultipleFits.R

@@ -3,6 +3,10 @@ numbat_circle <- read_csv("Draft/data/numbat.csv")
 numbat_only <- numbat_circle %>% filter(group=="A") %>% select(-group)
 circle_only <- numbat_circle %>% filter(group=="B") %>% select(-group)
 
+physEx <- read_csv("~/smodels/smodels-master/testStop/summaryRvalExpected.csv")
+ex1 <- physEx %>% filter(group=="BRt_75_BRb_25") %>% select(-group,-mst,-mchi,-cat)
+
+
 scagIndex2 <- function(proj,dataF,scagType) {
   proj <- matrix(proj, ncol=2, byrow=FALSE)
   proj <- tourr:::orthonormalise(proj)
@@ -13,7 +17,8 @@ scagIndex2 <- function(proj,dataF,scagType) {
   return(sR[scagType])
 }
 
-ix <- length(circle_only)
+#ix <- length(circle_only)
+ix <- length(ex1)
 ntries <- 10
 
 lower <- rep(-1,2*ix)
@@ -26,8 +31,9 @@ i <- 1
 while(i<=ntries){
   par0 <- runif(2*ix, min = 0, max = 1)
   opt <- hjkb(par0, scagIndex2, lower=lower, upper=upper,
-              control = list(maximize=TRUE, tol=0.001, 
-                             info=TRUE),dataF=circle_only, scagType="Skinny")
+              control = list(maximize=TRUE, tol=0.001,
+                             info=TRUE),dataF=ex1, scagType="Skinny")
+  #                           info=TRUE),dataF=circle_only, scagType="Skinny")
   idxV[i] <- opt$value
   prjL[[i]] <- tourr:::orthonormalise(matrix(opt$par, ncol=2, byrow=FALSE))
   i <- i+1
@@ -39,5 +45,6 @@ print(idxV[i1])
 
 plannedT <- planned_tour(prjL[i1])
 quartz()
-animate(numbat_only, plannedT)
+animate(ex1,plannedT)
+#animate(circle_only, plannedT)
 

### 2018-03-13 scagnostics.Rmd
Added first version of measure definitions

@@
 *** Motivating example ??
+This could be the pdf fit data, more structure and shape than other datasets we have?
 
@@
 This is inverted as $c_{skew}= 1-w(1-q_{skew})$ to account for binning effects that result in decreasing values of $q_{skew}$ with increasing n.
-* Sparse:
+* Sparse: A second measure based on the distribution of edge lenths, we define $c_{sparse} = w q_{90}$.
+It takes large values if points are concentrated in well separated parts of the plane, and low values if they are distributed approximately on a lattice.
+* Clumpy: To indicate the clustering of points the Clumpy measure is defined as $c_{clumpy}=\max\limits_{j}[1-\max\limits_{k}[\frac{length(e_k)}{length(e_j)}]]$ where we maximise the edge length $length(e_k)$ in the smaller of two subgraphs generated by removing a single edge $e_j$ from the MST.
+Note that $c_{clumpy}$ will take large values (close to one) if a long edge is separating clustered points connected by short edges.
+* Striated: We define an edge as striated if they have a large angle with one of the neighboring edges, concretely $cos(angle) < -0.75$, and the corresponding measure $c_{striate} = \frac{Number of striated edges}{Number of all edges}$.
+FIXME is this really correct?
+* Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
+* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimiter to the area.
+We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a circle, and values close to one for skinny polygons.
+* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{lenght(MST)}$ where diameter is longest connected path, and lenght the total lenght (sum of all edgest), i.e.
+we are measuring branch structure, no branches means $c_{stringy}=1$.
 

### 2018-09-06 scagnostics.Rmd
Commiting old changes before rewrite.

@@
 FIXME is this really correct?
+Maybe reason for discreet behaviour observed below..
 * Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
@@
 # Possible modifications
+We next study possible modifications that will enable the use of scagnostics as an index for guided tours.
+First, given the often large variation in scagnositcs measures between neighbouring projections, we consider smoothing the calculation, either via averaging over $n$ previous projections, or by jittering either the points or the projection matrix and averaging the resulting scagnostics measures over jittered projections.
+We then consider how different optimizers may be more adapt for an optimization over the erratic scagnostics measures.
 
 ## Smoothing
+First we consider smoothing the scagnostics index by averaging over $n$ previous projections, the resulting behaviour is shown in Figure...
 
 ## Jittering
+Another method of smoothing the scagnostics index would consist in jittering either the projected data points, or the projection itself.
+We show the impact of both methods in Figure ...
 
 ## Tour optimisation
+How does current optimizer work? What is the advantege of the alternative optimizer? Show results we get using each optimizer
 

### 2018-09-06 scagnostics.Rmd
Adding measure descriptions to the introduction.

@@
 One reason, is that it is hard, because defining what is meant by shape is not straightforward.
-An approach that seems plausible begins with scagnostics, developed by @scag, and explored in @WW08.
 
-*** Motivating example ??
-This could be the pdf fit data, more structure and shape than other datasets we have?
+A related question is the selection of pair-wise scatter plots in datasets with large numbers of parameters, for which different methods have been developed.
+One approch are scatterplot diagnostics, "scagnostics", developed by @scag, and explored in @WW08.
+They are calculated from a set of elementary graph-theoretic building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST).
+These are used to define eight measures: Outlying, Skewed, Sparse, Clumpy, Striated, Convex, Skinny and Stringy.
+While in principle each measure could be used on its own to define a ranking, the original idea suggested considering pair-wise scatter plots of the scagnostics measures for identification of "characteristic" and "unusual" 2-d views in the original parameter space.
+Alternative methods suggested by (Katrins Thesis or maybe her R package?) suggest using parallel coordinates, glyphs or a so-called scagramm (based on the idea of corrgrams) that shouold be coupled to interactive brushing for selection of pair-wise scatter plots to be displayed.
+(FIXME just give reference rather than all definitions here?)
+(Issues with scagnostics already for selection of pairs have been described in Katrins Thesis, maybe have short summary of that here?)
+Given the various shortcomings of the scagnostics approach we consider several other measures that can directly be interpreted as ranking of information contained in a given 2-d view.
+Two such additional measures have been defined in @mbgraphic.
+The first is based on smoothing splines, and is defined such that it will take maximum values when the data distribution can be described by a functional form.
+For a scatterplot with variable X and Y we calculate it as
+\begin{equation}
+splines_{2d} = max(1-\frac{Var(res_{X\sim Y})}{Var(X)}, 1-\frac{Var(res_{Y\sim X})}{Var(Y)})
+\end{equation}
+where $Var(.)$ is the variance and $res_{X\sim Y}$ are the residuals in the spline model where $X$ is considered a function of $Y$.
+The second new measure introduced in @mbgraphic is aiming to detect also non-functional relations and is based on the idea of distance correlation @szekely2007. (should I repeat definitions here?)
+Finally we will also consider the maximal and total information coefficient (MIC and TIC) defined in @Reshef1518 which are based on mutual information found when imposing grids on pair-wise variable plots.
+These functions are availabel in R via the interface implemented in the minerva package @minerva.
 

### 2018-09-07 scagnostics.Rmd
Motivation from physics?

@@
 One reason, is that it is hard, because defining what is meant by shape is not straightforward.
+
+(Motivation - something along the lines of the following paragraph, but more understandable, maybe giving concrete example - e.g. dec vs time from neutron star posterior?)
+One motivation to consider projection pursuit indices beyond the ones considered so far is coming from physics applications where a standard method is the determination of physcis parameters by means of global fits to large experimental datasets.
+Typically models will come with a handful of free parameters, model predictions will be calculated by means of numeric tools and the best fit parameters are evaluated by comparing predictions to a large number of observations.
+Fit results will encode relations between the model parameters dictaded by the available observations, for example if two parameters control predictions for a well constrained observable, looking at the distribution of parameter combinations which are a good fit will be instructive.
+A standard example is to study the likelihood or posterior distribution in pair-wise 2-d plots, where the selection of parameters if often guided by knowledge of the underlying physics.
+A major challenge is the progressing complexity in all three aspects: models with larger parameter spaces are being studied, computation of observable predictions is increasingly complex and the number of experimental observations relevant to a given model is growing.
+Methods for the detection of "interesting" views that can aid physics intuition are therefore more and more relevant.
 

### 2018-09-07 scagnostics.Rmd
Adding comments for setting up structure.

@@
 These functions are availabel in R via the interface implemented in the minerva package @minerva.
+(read paper and add more detail.
+how are they similar/different compared to scagnostics and distance correlation?
+the way that binning is being built into it may make difference in how MIC/TIC may be smoother?)
 
@@
 Software to run a projection pursuit guided tour is available in the R package, "tourr" [@tourr].
+(Even just for the optimisation part the guided tour implementation is useful.)
+(Traditional projection pursuit indices are fast to compute and therefore the optimisation would be done in "real time" while displaying the tour.
+For the measures considered here this generally does not apply, as a consequence we "disentangle" the two into subsequent steps.)
 
@@
 library(geozoo)
+library(minerva)
 ```
@@
 The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
-If there are 100 variables, the dimension will be $4950x9$.
+If there are 100 variables, the dimension will be $4950\times 9$.
 It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting.

### 2018-09-10 scagnostics.Rmd
Updated example data sets (code broken).

@@
 library(reshape2)
-library(binostics) #https://github.com/uschiLaa/binostics
+library(scagnostics)
 library(gridExtra)
-library(wesanderson) #color palette
 library(tictoc) #timer
 library(mbgraphic) #Katrins package
-library(dfoptim)
 library(GGally)
 library(geozoo)
-library(minerva)
+library(minerva) #MINE indices
 ```
@@
 
-As a test dataset we generate high dimensional geometrical shapes using the geozoo package (REF).
-Concretely we consider five dimensional objects, and we combine a hollow shpere with radius 1 and a solid torus with radii (0.5, 0.5, 0.1,0.1).
-The resulting distributions can be seen in the scatter plot matrix below.
-In the following we combine both groups to study the resulting non-standard distribution.
-
-```{r geozooDist, fig.width=6, fig.height=7, fig.cap="Scatter plot matrix of the five dimensional shapes, randomly selecting 500 points in each group.
-Group 1 (2) contains points selected on a hollow sphere (solid torus)."}
-hSphere <- sphere.hollow(p=5,n=500)$points
-sTorus <- torus(5,500, radius = c(0.5,0.5,0.1,0.1))$points
-geoDF <- bind_rows(as.data.frame(hSphere),as.data.frame(sTorus),.id = "Group")
-ggscatmat(geoDF, columns = 2:6, color="Group")
-```
+We consider three example distributions in n=6 dimensions with p randomly selected points $x_i, i = 1,..,n$:
+1. p points on an n-dimensional sphere as an example of dataset without special views
+2.
+p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not showing functional dependence, and where sampling boundaries are apparant in the distribution
+3.
+p points where n-1 points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function
 
 ## Smoothness
-*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference
-
-*** katrin's work
-
-XXX Probably use a simple benchmark data to illustrate the issues, maybe mtcars that comes with the scagnostics package
-
 We first study the smoothness of the scagnostics measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path.
-The temporal trace of all nine measures is shown below, where each time step refers to an interpolation step with a target distance between interpolated bases is set to 0.05 radians.
-We observe sharp jumps in almost all measures (with "Monotonic" being the only exception), note however the large differences in overall scale between the measures.
-In particular "Clumpy", "Monotonic", "Outlying", "Sparse" and "Striated" show little sensitivity, as could be expected from the distributions in Figure (REF).
+The temporal trace of a selected set of measures is shown below, where each time step refers to an interpolation step.
 
 ```{r util}
+#defining index functions to be used with the tour
 scagIndex <- function(scagType){
@@
 
-plotProj <- function(nProj, fullPath, dMatrix,legendOn, lim=75){
-  p <- fullPath[nProj][[1]]
-  d <- as.matrix(dMatrix) %*% p
-  s <- scagnostics.default(d[,1],d[,2], bins= 50)
-  p1 <- ggplot()+
-    geom_point(data = as.data.frame(d), mapping=aes(x=V1, y=V2))
-  p2 <- ggplot()+
-    geom_point(data = as.data.frame(s$bins), mapping = aes(x=V1, y=V2, color=V3, size=V3))+
-    xlim(0,1000)+
-    ylim(0,1000)+
-    expand_limits(colour = c(1,lim), size = c(1,lim))+
-    scale_color_gradientn(colors=wes_palette(name="Zissou"))+
-    scale_size(range = c(1,3))
-    if(legendOn==0){p2 = p2 + guides(color=FALSE, size=FALSE)}
-    grid.arrange(arrangeGrob(p1, widths = 2),arrangeGrob(p2, widths = 7),nrow=1)
-    return(s$s)
+splineIndex <- function(){
+  function(mat){
+    return(splines2d(mat[,1], mat[,2]))
+  }
+}
+
+dcorIndex <- function(){
+  function(mat){
+    return(dcor2d(mat[,1], mat[,2]))
+  }
+}
+
+mineIndex <- function(mineIndex){
+  function(mat){
+    return(mine(mat[,1], mat[,2])[[mineIndex]])
+  }
 }
@@
 
-```{r tourpathGeozoo}
-set.seed(272829)
-geoCombined <- select(geoDF,-Group)
-t1 <- save_history(geoCombined, grand_tour(2), max=4)
+```{r datasetFunctions}
+sphereData <- function(n, p){
+  dRet <- geozoo::sphere.solid.random(n,p)
+  return(as.tibble(dRet$points))
+}
+
+pipeData <- function(n, p){
+  i <- 1
+  dRet <- NULL
+  while(i <= p){
+    v <- runif(n, -1, 1)
+    if(abs(v[n-1]*v[n-1] + v[n]*v[n] - 1) < 0.1){
+      dRet <- rbind(dRet, v)
+      i <- i+1
+    }
+  }
+  return(as.tibble(dRet))
+}
+
+sinData <- function(n, p){
+  vName <- paste0("V",n)
+  vNameM1 <- paste0("V",n-1)
+  expr <- paste0(vName,"=sin(",vNameM1,")") # need string expression if I want to use tibble here
+  dRet <- as.tibble(matrix(rnorm((n-1)*p), ncol=(n-1))) #generate normal distributed n-1 dim data
+  dRet <- mutate_(dRet, expr) #string evaluation calculates var(n) as tan(var(n-1))
+  colnames(dRet)[n] <- vName #correct name of new variable
+  dRet[vName] <- jitter(dRet[[vName]]) #adding noise
+  return(dRet)
+}
+```
+
+```{r datasets}
+sphere100 <- sphereData(6, 100)
+sphere1000 <- sphereData(6, 1000)
+
+pipe100 <- pipeData(6, 100)
+pipe1000 <- pipeData(6,1000)
+
+sin100 <- sinData(6, 100)
+sin1000 <- sinData(6, 1000)
+```
+
+```{r getProj}
+getProj <- function(df, tPath, nameStr){
+  sc <- NULL
+  n <- length(tPath)
+  df <- rescale(df)
+  for (i in 1:n) {
+  dprj <- df %*% tPath[[i]]
+  #FIXME not only scagnostics but all relevant!
+  #also keep track of t=i, and nameStr
+  #after want to be able to directly plot full comparison.
+  sc <- rbind(sc, scagnostics(dprj))
+}
+}
+```
+
+```{r grandtourpath}
+set.seed(1988)
+t1 <- save_history(sphere100, grand_tour(2), max=4)
 t1full <- as.list(interpolate(t1))
@@
 }
-sc <- as.data.frame(sc)
-sc$t <- 1:n
-scPairs <- sc
-sc <- sc %>% gather(index, value, -t)
-```
+#sc <- as.data.frame(sc)
+#sc$t <- 1:n
+#scPairs <- sc
+#sc <- sc %>% gather(index, value, -t)
+#```
 
-```{r plotscagGeozoo, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
-ggplot(sc, aes(x=t, y=value)) + geom_line() + 
-  facet_wrap(~index, ncol=1, scales = "free_y") +
-  geom_vline(mapping = aes(xintercept=24, color="red")) +
-  geom_vline(mapping = aes(xintercept=39, color="red")) +
-  guides(color=FALSE)
-```
-
-```{r plotscagPairs, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections, as above.
-Scatter plot of scagnostics calculated on each projection.
-Observations: discreetness of outlying and striated, anti-correlation of convex and skinny.
-FIXME is this relevant?
-Maybe remove this plot."}
-ggpairs(scPairs, columns = 1:9)
-```
-
-Two projections, t=24 and t=39, have been marked in Figure (can I add a reference?), as examples exhibiting jumps in several measures with respect to the previous and/or next projection.
-In particular the outlying measure indicates that at least one outlier has been identified in the projection at t=24, but not the projections before or after.
-The removal of the point(s) will likely impact the calculation of other measures, and indeed we even see a small bump in the otherwise smooth measure "monotonic" (note however that the same is not true in other projections with outlying points).
-The projection at t=39 has been selected because of the sharp jumps in the measures "skewed", "stringy" and "striated".
-We compare the observed behaviour to the scatter plots obtained from the projected data in the Figure below.
-As expected the projections look similar over the 3 interpolation steps, small differences in the binned distribution can however be observed.
-Certainly the sometimes large differences (jumps of 0.1 absolute change in a measure from one projection to the next) are not indicating interesting differences in the projected view of the data.
-
-```{r plotProj15, fig.width=6, fig.height=2, fig.cap="Projected data at t = 23, 24, 25 (top to bottom)."}
-s1 <- plotProj(23, t1full, geoCombined, legendOn=0)
-s2 <- plotProj(24, t1full, geoCombined, legendOn=0)
-s3 <- plotProj(25, t1full, geoCombined, legendOn=0)
-```
-
-```{r plotProj25, fig.width=6, fig.height=2, fig.cap="Projected data at t = 38, 39, 40 (top to bottom).
-The original data points are shown on the left, the binned data, used for the calculation of the scagnostics measures is shown on the right."}
-s1 <- plotProj(38, t1full, geoCombined, legendOn=0)
-s2 <- plotProj(39, t1full, geoCombined, legendOn=0)
-s3 <- plotProj(40, t1full, geoCombined, legendOn=0)
-```
+#```{r plotscagGeozoo, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on #projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+#ggplot(sc, aes(x=t, y=value)) + geom_line() + 
+#  facet_wrap(~index, ncol=1, scales = "free_y") +
+#  geom_vline(mapping = aes(xintercept=24, color="red")) +
+#  geom_vline(mapping = aes(xintercept=39, color="red")) +
+#  guides(color=FALSE)
+#```
 

### 2018-09-10 scagnostics.Rmd
Temporal traces working.

@@
 The temporal trace of a selected set of measures is shown below, where each time step refers to an interpolation step.
+Note that each parameter range has been rescaled to be between [0,1] for direct comparison.
 
@@
 ```{r getProj}
-getProj <- function(df, tPath, nameStr){
-  sc <- NULL
+getProj <- function(df, tPath, nameStr, size){
+  sc <- tibble(convex=numeric(),
+               clumpy=numeric(),
+               skinny=numeric(),
+               stringy=numeric(),
+               dcor2d=numeric(),
+               splines2d=numeric(),
+               MIC=numeric(),
+               TIC1=numeric(),
+               t=numeric(),
+               name=character(),
+               size=numeric())
   n <- length(tPath)
@@
     dprj <- df %*% tPath[[i]]
-  #FIXME not only scagnostics but all relevant!
-  #also keep track of t=i, and nameStr
-  #after want to be able to directly plot full comparison.
-  sc <- rbind(sc, scagnostics(dprj))
+    scagRes <- scagnostics(dprj)
+    dcorRes <- dcor2d(dprj[,1], dprj[,2])
+    splineRes <- splines2d(dprj[,1], dprj[,2])
+    mineRes <- mine(dprj[,1], dprj[,2])
+    sc <- add_row(sc, convex=scagRes["Convex"],
+                  clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
+                  stringy=scagRes["Stringy"], dcor2d=dcorRes,
+                  splines2d=splineRes, MIC=mineRes$MIC,
+                  TIC1=mineRes$TIC, t=i, name=nameStr, size=size)
   }
+  maxTIC <- max(sc$TIC1)
+  sc <- sc %>%
+    mutate(TIC = TIC1/maxTIC) %>%
+    select(-TIC1)
+
+  return(sc)
 }
@@
 t1full <- as.list(interpolate(t1))
-n <- length(t1full)
-sc <- NULL
-for (i in 1:n) {
-  dprj <- as.matrix(geoCombined) %*% t1full[[i]]
-  sc <- rbind(sc, scagnostics(dprj))
-}
-#sc <- as.data.frame(sc)
-#sc$t <- 1:n
-#scPairs <- sc
-#sc <- sc %>% gather(index, value, -t)
-#```
+fullRes <- getProj(sin100, t1full, "sin", 100) %>%
+  rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
+  rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
+  rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
+  rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
+  rbind(getProj(pipe1000, t1full, "pipe", 1000))
+```
 
-#```{r plotscagGeozoo, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on #projections. Most of the indices exhibit sharp jumps in scagnostic value."}
-#ggplot(sc, aes(x=t, y=value)) + geom_line() + 
-#  facet_wrap(~index, ncol=1, scales = "free_y") +
-#  geom_vline(mapping = aes(xintercept=24, color="red")) +
-#  geom_vline(mapping = aes(xintercept=39, color="red")) +
-#  guides(color=FALSE)
-#```
+```{r plotgrandtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
+ggplot(fullResMelt, aes(x=t, y=value)) +
+  geom_line(aes(color=name, linetype=factor(size))) + 
+  facet_wrap(~variable, ncol=1, scales = "free_y")
+```
 

### 2018-09-10 scagnostics.Rmd
More structure updates.

@@
 output: rticles::asa_article
+header-includes:
+  - \usepackage{booktabs}
+  - \usepackage{longtable}
 ---
@@
 ## Smoothness
-We first study the smoothness of the scagnostics measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path.
+We first study the smoothness of a subset of measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path between 4 anchor planes and using the default interpolation angle of 0.05.
+For this comparison we rescale the data such that each parameter takes values between [0,1].
 The temporal trace of a selected set of measures is shown below, where each time step refers to an interpolation step.
-Note that each parameter range has been rescaled to be between [0,1] for direct comparison.
+Note that the value of TIC is strongly dependent on the input distribution and we have therefore rescaled the range for each individual dataset to [0,1], all other measures are reproduced as reported by the calculation.
 
@@
 
-## Effect of binning
+Sharp jumps can be observed in all measures except the distance correlation measure when considering small datasets of only 100 points.
+Generally we observe smoother behaviour for larger datasets, notably all considered scagnostics measures show only small variation around a flat baseline.
+Interestingly all other indices indicate preferred viewing directions for dataset 3 when studying large samples, but not for dataset 2 which contains a special view in the same direction.
+We therefore show 2-d view defined by the tour step $t=58$, with projection matrix
+```{r t58matrix, echo=FALSE}
+knitr::kable(t1full[[58]] ,  caption = "Matrix representation of projection 58.")
+```
+i.e. while the last two parameters enter significantly, they will not dominate the view and we do not expect to fully observe the built in features.
+```{r plot58, fig.width=6, fig.height=7, fig.cap="View of all datasets defined by at tour step 58."}
+p58 <- t1full[[58]]
+pL <- list()
+i <- 1
+pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+  dm <- rescale(ds)
+  pData <- as.tibble(dm %*% p58)
+  pC <- ggplot(pData, aes(V1, V2)) +
+    geom_point() +
+    ggtitle(pLabels[i])
+  pL[[i]] <- pC
+  i <- i+1
+}
+grid.arrange(pL[[1]], pL[[2]],
+             pL[[3]], pL[[4]],
+             pL[[5]], pL[[6]],
+             ncol=2)
 
-Two possible sources of discreet jumps, from binning and discreetness of the underlying objects, in particular the minimal spanning tree.
-Can we see how mst looks for each projection to understand this?
+```
+We see from Fig. ((how to add cross references?)) that the built in special views cannot be seen in this projection, but confirm that the sine 1000 dataset presents some correlation in this view.
 
-## Speed
+## Guided tour
+Given the behaviour of the various index measures under an interpolated tour path we next try to use the tourr implementation of the guided tour together with selected index functions to uncover the special views in datasets 2 and 3 with 1000 datapoints.
 
-# Possible modifications
-We next study possible modifications that will enable the use of scagnostics as an index for guided tours.
-First, given the often large variation in scagnositcs measures between neighbouring projections, we consider smoothing the calculation, either via averaging over $n$ previous projections, or by jittering either the points or the projection matrix and averaging the resulting scagnostics measures over jittered projections.
-We then consider how different optimizers may be more adapt for an optimization over the erratic scagnostics measures.
+### Review of guided tour optimisation
 
-## Smoothing
-First we consider smoothing the scagnostics index by averaging over $n$ previous projections, the resulting behaviour is shown in Figure...
+### Looking down the pipe
+TIC should be able to uncover special view, compare performance of all indices on guided tour path. (also cmass and holes?)
 
-## Jittering
-Another method of smoothing the scagnostics index would consist in jittering either the projected data points, or the projection itself.
-We show the impact of both methods in Figure ...
+### Finding sine waves
+Use spline2d to try and find sine shape, compare performance of all indices on guided tour path.
 
-## Tour optimisation
-How does current optimizer work? What is the advantege of the alternative optimizer? Show results we get using each optimizer
+### Note on calculation time
+Considerations: time to calculate index for single 2-d plot, parameters to turn for guided tour (max.tries, alpha, cooling)
 

### 2018-09-11 scagnostics.Rmd
First attempt at guided tour.

@@
 The temporal trace of a selected set of measures is shown below, where each time step refers to an interpolation step.
-Note that the value of TIC is strongly dependent on the input distribution and we have therefore rescaled the range for each individual dataset to [0,1], all other measures are reproduced as reported by the calculation.
+Note that the value of TIC is strongly dependent on the input distribution and we have therefore normalised the TIC measure by its maximum for each individual dataset, all other measures are reproduced as reported by the calculation.
 
@@
 ### Review of guided tour optimisation
+We use the default method "search_geodesic" implemented in the tourr package. (Should I try other optimisation methods, i.e. search_better, search_better_random?)
+Search geodesic first collects $n=5$ samples by default, amongst which the most promising direction is selected.
+In a second step a linear search along the geodesic is performed in that direction.
+The optimisation is stopped when ???
 
@@
 TIC should be able to uncover special view, compare performance of all indices on guided tour path. (also cmass and holes?)
+Not currently finding "best" view. Options: compare results from n different random starting planes, try with smaller space (e.g. omitting variable 1 and 2), turn some of the guided tour parameters
+```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset."}
+set.seed(2018)
+pipeResc <- rescale(pipe1000)
+pipeTour <- save_history(pipeResc,
+                         guided_tour(mineIndex("TIC")), start = p58)
+pipeTourFull <- as.list(interpolate(pipeTour))
+pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
+pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
+ggplot(pipeTourResMelt, aes(x=t, y=value)) +
+  geom_line() + 
+  facet_wrap(~variable, ncol=1, scales = "free_y")
+```
+
+```{r testpipe}
+iLast <- length(pipeTourFull)
+fProj <- pipeTourFull[[iLast]]
+dProj <- as.tibble(pipeResc %*% fProj)
+ggplot(dProj, aes(V1,V2)) + geom_point()
+```
 

### 2018-09-11 scagnostics.Rmd
Working out some details for example distributions.

@@
 
-The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
-If there are 100 variables, the dimension will be $4950\times 9$.
-It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting.
-It can also be used to do more in-depth exploratory analysis, and is interesting to study the distribution of the scagnostics resulting from the data, as a data set in itself.
-In a sense this the data's "fingerprint".
-
-To use scagnostics for projection pursuit requires that each of the measures for functions that are relatively smooth over the space of all 2D projections, and that they can be computed rapidly enough that the display can be updated many times a second during optimisation.
-
-## Dataset
-
-We consider three example distributions in n=6 dimensions with p randomly selected points $x_i, i = 1,..,n$:
-1. p points on an n-dimensional sphere as an example of dataset without special views
-2.
-p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not showing functional dependence, and where sampling boundaries are apparant in the distribution
-3.
-p points where n-1 points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function
-
-## Smoothness
-We first study the smoothness of a subset of measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path between 4 anchor planes and using the default interpolation angle of 0.05.
-For this comparison we rescale the data such that each parameter takes values between [0,1].
-The temporal trace of a selected set of measures is shown below, where each time step refers to an interpolation step.
-Note that the value of TIC is strongly dependent on the input distribution and we have therefore normalised the TIC measure by its maximum for each individual dataset, all other measures are reproduced as reported by the calculation.
-
 ```{r util}
@@
 ```
+
+The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
+If there are 100 variables, the dimension will be $4950\times 9$.
+It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting.
+It can also be used to do more in-depth exploratory analysis, and is interesting to study the distribution of the scagnostics resulting from the data, as a data set in itself.
+In a sense this the data's "fingerprint".
+
+To use scagnostics for projection pursuit requires that each of the measures for functions that are relatively smooth over the space of all 2D projections, and that they can be computed rapidly enough that the display can be updated many times a second during optimisation.
+
+## Dataset
+
+We consider three example distributions in n=6 dimensions with p randomly selected points $x_i, i = 1,..,n$:
+1. p points on an n-dimensional sphere as an example of dataset without special views
+2.
+p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not showing functional dependence, and where sampling boundaries are apparant in the distribution
+3.
+p points where n-1 points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function
 
@@
 ```
+
+To get an overview of the dataset we show views of the first two parameters showing the independent distribution and the last two parameters showing the special view.
+The enforces relationships will also affect the overall distribution as can be seen from (fig:dataPlotsDensity).
+
+```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Data views of V1 vs V2."}
+pL <- list()
+i <- 1
+pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+  pC <- ggplot(ds, aes(V1, V2)) +
+    geom_point() +
+    ggtitle(pLabels[i])
+  pL[[i]] <- pC
+  i <- i+1
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
+             ncol=2)
+```
+
+```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Data views of V5 vs V6."}
+pL <- list()
+i <- 1
+pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+  pC <- ggplot(ds, aes(V5, V6)) +
+    geom_point() +
+    ggtitle(pLabels[i])
+  pL[[i]] <- pC
+  i <- i+1
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
+             ncol=2)
+```
+
+```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="Data views of V1 vs V6."}
+pL <- list()
+i <- 1
+pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+  distData <- melt(ds)
+  pC <- ggplot(distData, aes(value, color=variable)) +
+    geom_density() +
+    ggtitle(pLabels[i])
+  pL[[i]] <- pC
+  i <- i+1
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
+             ncol=2)
+```
+
+We next compare the various indexvalues for three views: V1 vs V2, V5 vs V6 and V1 vs V6 (capturing differences from the different distribution in V6).
+
+```{r indexTable, echo=FALSE}
+i <- 1
+pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+indexOverview <- tibble(convex=numeric(),
+               clumpy=numeric(),
+               skinny=numeric(),
+               stringy=numeric(),
+               dcor2d=numeric(),
+               splines2d=numeric(),
+               MIC=numeric(),
+               TIC=numeric(),
+               name=character(),
+               vars=character())
+for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+  ds <- as.tibble(rescale(ds))
+  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
+    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
+    scagRes <- scagnostics(dprj)
+    dcorRes <- dcor2d(dprj[,1], dprj[,2])
+    splineRes <- splines2d(dprj[,1], dprj[,2])
+    mineRes <- mine(dprj[,1], dprj[,2])
+    indexOverview <- add_row(indexOverview, convex=scagRes["Convex"],
+                    clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
+                    stringy=scagRes["Stringy"], dcor2d=dcorRes,
+                    splines2d=splineRes, MIC=mineRes$MIC,
+                    TIC=mineRes$TIC, name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
+  }
+  i <- i+1
+}
+knitr::kable(indexOverview ,  caption = "Summary of index values.", digits=2) 
+```
+
+```
+
+## Smoothness
+We first study the smoothness of a subset of measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path between 4 anchor planes and using the default interpolation angle of 0.05.
+For this comparison we rescale the data such that each parameter takes values between [0,1].
+The temporal trace of a selected set of measures is shown below, where each time step refers to an interpolation step.
+Note that the value of TIC is strongly dependent on the input distribution and we have therefore normalised the TIC measure by its maximum for each individual dataset, all other measures are reproduced as reported by the calculation.
 
@@
 pipeTour <- save_history(pipeResc,
-                         guided_tour(mineIndex("TIC")), start = p58)
+                         guided_tour(mineIndex("TIC"), n=10, cooling = 0.8), start = p58)
 pipeTourFull <- as.list(interpolate(pipeTour))

### 2018-09-11 scagnostics.Rmd
More optimisation, testing change in index when moving into V5 vs V6.

@@
 We next compare the various indexvalues for three views: V1 vs V2, V5 vs V6 and V1 vs V6 (capturing differences from the different distribution in V6).
+For the Sphere distribution by construction we see only random fluctuations between the different views, which can be sizable in particular for small datasets.
+Considering next the Pipe distribution, we see that skinny, stringy and dcor2d are increased for the special view, but in particular MIC and TIC index are much larger and may be used to detect this type of structure in the dataset.
+Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fuctional dependence.
 
@@
 knitr::kable(indexOverview ,  caption = "Summary of index values.", digits=2) 
-```
-
 ```
@@
 
+## Transition
+We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6.
+```{r plannedtourpath}
+m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
+m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
+#this is silly but seems that first two entries are being ignored, so need some fake entries
+m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
+m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
+t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
+t2full <- as.list(interpolate(t2))
+fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
+  rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
+  rbind(getProj(sphere100, t2full, "sphere", 100)) %>%
+  rbind(getProj(sphere1000, t2full, "sphere", 1000)) %>%
+  rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
+  rbind(getProj(pipe1000, t2full, "pipe", 1000))
+```
+
+```{r plannedgrandtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
+ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
+  geom_line(aes(color=name, linetype=factor(size))) + 
+  facet_wrap(~variable, ncol=1, scales = "free_y")
+```
+
+We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise.
+On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparant much closer to the final projection, making this structure more challenging to detect with a guided tour.
+(What does this mean for optimisation? How can I detect such "smaller scale increase"? Maybe try "search_better" instead of search_geodesic? increase alpha and decrease cooling? increase max.tries?)
+Can force it when using search_better with large alpha, cooling=1, max.tries=1000, but not ideal..
+Difficulties also from "fake" structure from the sampling boundaries?
+Possible issue with search_geodesic is that the search window is fixed in the code..
+
 ## Guided tour
@@
 ### Looking down the pipe
-TIC should be able to uncover special view, compare performance of all indices on guided tour path. (also cmass and holes?)
-Not currently finding "best" view. Options: compare results from n different random starting planes, try with smaller space (e.g. omitting variable 1 and 2), turn some of the guided tour parameters
-```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset."}
+As seen before (reference figure) MIC or TIC should be able to find the view down the pipe.
+The optimisation itself is however expected to be challenging because one has to be fairly close to the optimal projection to observe the increase in index value.
+We therefore suggest a two step procedure for the uncovering of such "small scale structures".
+The first step consists in "scouting" projections.
+Here we use the "search_better" optimisation available with the tourr package, with a standard search window $\alpha = 0.5$ and disabeling the cooling (in practice this can be done by fixing the cooling parameter to $1$), together with a large number of tries (setting max.tries to 5000 here).
+The resulting search will quickly steer away from the most unintresting projections, followed by a broad scan of the thus identified region.
+While this does not maximise the index function, it provides an informed starting plane that can be used in a second step where we use the "search_geodesic" optimisatino to improve the optimisation.
+The described procedure is CPU time intensive, however it is promising that we manage to uncover such small scale structure in the distribution.
+On the other hand no "guarantee" can be provided for uncovering the special view.
+Parameters for the scouting phase will need to be adapted to different examples, and with increasing number of dimensions it may be better to consider several runs with different starting projections.
+This gives some indication on how to handle difficult datasets. As we will see below most of the time standard "search_geodesic" optimisation will be efficient enough to uncover the interesting views.
+
+```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset.", echo=FALSE}
 set.seed(2018)
-pipeResc <- rescale(pipe1000)
+pipeResc <- rescale(pipe1000) #unselect 2 variables to make optimisation easier
 pipeTour <- save_history(pipeResc,
-                         guided_tour(mineIndex("TIC"), n=10, cooling = 0.8), start = p58)
+                         guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
+                                     cooling=1, alpha = 0.5, max.tries = 5000))
 pipeTourFull <- as.list(interpolate(pipeTour))
@@
 ```{r testpipe}
+
+iLast <- length(pipeTourFull)
+fProj <- pipeTourFull[[iLast]]
+dProj <- as.tibble(pipeResc %*% fProj)
+ggplot(dProj, aes(V1,V2)) + geom_point()
+```
+
+```{r refine}
+pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
+pipeTourFull2 <- as.list(interpolate(pipeTour2))
+pipeTourRes2 <- getProj(pipe1000, pipeTourFull2, "Pipe", 1000)
+pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
+ggplot(pipeTourResMelt2, aes(x=t, y=value)) +
+  geom_line() + 
+  facet_wrap(~variable, ncol=1, scales = "free_y")
+```
+
+```{r testpipe2}
 iLast <- length(pipeTourFull)
@@
 ### Finding sine waves
-Use spline2d to try and find sine shape, compare performance of all indices on guided tour path.
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Sine 1000 dataset.", echo=FALSE}
+set.seed(2018)
+sineResc <- rescale(sin1000) #unselect 2 variables to make optimisation easier
+sineTour <- save_history(sineResc,
+                         guided_tour(splineIndex()))
+sineTourFull <- as.list(interpolate(sineTour))
+sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
+sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
+ggplot(sineTourResMelt, aes(x=t, y=value)) +
+  geom_line() + 
+  facet_wrap(~variable, ncol=1, scales = "free_y")
+```
+
+```{r testsine}
+iLast <- length(sineTourFull)
+fProj <- sineTourFull[[iLast]]
+dProj <- as.tibble(sineResc %*% fProj)
+ggplot(dProj, aes(V1,V2)) + geom_point()
+```
 

### 2018-09-12 scagnostics.Rmd
Small fixes only.

@@
 
-```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset.", echo=FALSE}
+```{r findpipe, echo=FALSE}
 set.seed(2018)
@@
 pipeTourFull <- as.list(interpolate(pipeTour))
+```
+
+```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Guided tour optimising MIC index for Pipe 1000 dataset.", echo=FALSE}
 pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
@@
 ```{r testpipe}
-
 iLast <- length(pipeTourFull)
@@
 ```{r testpipe2}
-iLast <- length(pipeTourFull)
-fProj <- pipeTourFull[[iLast]]
+iLast <- length(pipeTourFull2)
+fProj <- pipeTourFull2[[iLast]]
 dProj <- as.tibble(pipeResc %*% fProj)

### 2018-09-13 scagnostics.Rmd
Moved scagnostics definitions to appendix.
Hiding guided tour printout.

--- 2018-09-11 scagnostics.Rmd
@@
 For definiteness all variables are rescaled to a [0,1] intervall, such that all measures will take values between 0 and 1 as well.
-The following measures have been defined:
-
-* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantil edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long.
-The outlying measure is then calculated as $c_{outlying} = \frac{\mathrm{Total\ length\ of\ long\ edges}}{\mathrm{Total\ length\ of\ all\ edges}}$.
-Note that outlying points are not considered in the calculation of the other measures.
-* Skewed: Skewness is defined via the distribution of edge length in the MST, $q_{skew}=\frac{q_{90}-q{50}}{q_{90}-q_{10}}$.
-This is inverted as $c_{skew}= 1-w(1-q_{skew})$ to account for binning effects that result in decreasing values of $q_{skew}$ with increasing n.
-* Sparse: A second measure based on the distribution of edge lenths, we define $c_{sparse} = w q_{90}$.
-It takes large values if points are concentrated in well separated parts of the plane, and low values if they are distributed approximately on a lattice.
-* Clumpy: To indicate the clustering of points the Clumpy measure is defined as $c_{clumpy}=\max\limits_{j}[1-\max\limits_{k}[\frac{length(e_k)}{length(e_j)}]]$ where we maximise the edge length $length(e_k)$ in the smaller of two subgraphs generated by removing a single edge $e_j$ from the MST.
-Note that $c_{clumpy}$ will take large values (close to one) if a long edge is separating clustered points connected by short edges.
-* Striated: We define an edge as striated if they have a large angle with one of the neighboring edges, concretely $cos(angle) < -0.75$, and the corresponding measure $c_{striate} = \frac{Number of striated edges}{Number of all edges}$.
-FIXME is this really correct?
-Maybe reason for discreet behaviour observed below..
-* Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
-* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimiter to the area.
-We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a circle, and values close to one for skinny polygons.
-* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{lenght(MST)}$ where diameter is longest connected path, and lenght the total lenght (sum of all edgest), i.e.
-we are measuring branch structure, no branches means $c_{stringy}=1$.
 
@@
 
-```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset.", echo=FALSE}
+```{r findpipe, results="hide"}
 set.seed(2018)
@@
 pipeTourFull <- as.list(interpolate(pipeTour))
+```
+
+```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Guided tour optimising MIC index for Pipe 1000 dataset."}
 pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
@@
 
-```{r refine}
+```{r refine, results="hide"}
 pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
@@
   facet_wrap(~variable, ncol=1, scales = "free_y")
+
 ```
@@
 ```{r testpipe2}
-iLast <- length(pipeTourFull)
-fProj <- pipeTourFull[[iLast]]
+
+iLast <- length(pipeTourFull2)
+fProj <- pipeTourFull2[[iLast]]
 dProj <- as.tibble(pipeResc %*% fProj)
@@
 ### Finding sine waves
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Sine 1000 dataset.", echo=FALSE}
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Sine 1000 dataset.", results="hide"}
 set.seed(2018)
@@
 
+# Apendix
+The following measures have been defined:
+
+* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantil edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long.
+The outlying measure is then calculated as $c_{outlying} = \frac{\mathrm{Total\ length\ of\ long\ edges}}{\mathrm{Total\ length\ of\ all\ edges}}$.
+Note that outlying points are not considered in the calculation of the other measures.
+* Skewed: Skewness is defined via the distribution of edge length in the MST, $q_{skew}=\frac{q_{90}-q{50}}{q_{90}-q_{10}}$.
+This is inverted as $c_{skew}= 1-w(1-q_{skew})$ to account for binning effects that result in decreasing values of $q_{skew}$ with increasing n.
+* Sparse: A second measure based on the distribution of edge lenths, we define $c_{sparse} = w q_{90}$.
+It takes large values if points are concentrated in well separated parts of the plane, and low values if they are distributed approximately on a lattice.
+* Clumpy: To indicate the clustering of points the Clumpy measure is defined as $c_{clumpy}=\max\limits_{j}[1-\max\limits_{k}[\frac{length(e_k)}{length(e_j)}]]$ where we maximise the edge length $length(e_k)$ in the smaller of two subgraphs generated by removing a single edge $e_j$ from the MST.
+Note that $c_{clumpy}$ will take large values (close to one) if a long edge is separating clustered points connected by short edges.
+* Striated: We define an edge as striated if they have a large angle with one of the neighboring edges, concretely $cos(angle) < -0.75$, and the corresponding measure $c_{striate} = \frac{Number of striated edges}{Number of all edges}$.
+FIXME is this really correct?
+Maybe reason for discreet behaviour observed below..
+* Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
+* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimiter to the area.
+We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a circle, and values close to one for skinny polygons.
+* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{lenght(MST)}$ where diameter is longest connected path, and lenght the total lenght (sum of all edgest), i.e.
+we are measuring branch structure, no branches means $c_{stringy}=1$.
+
 # References

### 2018-09-17 scagnostics.Rmd
Also comparing holes and cmass index.

@@
 getProj <- function(df, tPath, nameStr, size){
-  sc <- tibble(convex=numeric(),
+  sc <- tibble(
+    holes=numeric(),
+    cmass=numeric(),
+    convex=numeric(),
     clumpy=numeric(),
@@
     mineRes <- mine(dprj[,1], dprj[,2])
-    sc <- add_row(sc, convex=scagRes["Convex"],
+    holesRes <- holes(dprj)
+    cmassRes <- cmass(dprj)
+    sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
+                  convex=scagRes["Convex"],
                   clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
@@
   rbind(getProj(pipe1000, t1full, "pipe", 1000))
+
 ```
@@
   rbind(getProj(pipe1000, t2full, "pipe", 1000))
+
 ```

### 2018-09-17 scagnostics.Rmd
New reference for MIC.
Some comments added.

@@
 
+TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full information of the characteristics is used, implying that more information is considered.
+We also see that TIC is smoother when moving into the special view.
+
+Here we also see how the existing indices in the tourr package are inadequat for the detection of these type of dependence, the holes index taking maximum values before the final view and decreasing when moving into the special view.
+On the other hand, surprisingly the cmass index increases when moving into the special view, but the intermediate minimum makes it unsuitable as a projection pursuit index here.
+QUESTION: how come they are all almost the same for the various distributions?
+Especially pipe and sphere seem to give almost identical behaviour, can it be understood from the index definition?
+
 ## Guided tour

### 2018-09-17 scagnostics.Rmd
More comments.
Updated plot appearance.

@@
 ggplot(fullResMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=name, linetype=factor(size))) + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")
+  geom_line(aes(color=factor(size))) + 
+  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
+  guides(color=FALSE)
 ```
@@
 ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
-  geom_line(aes(color=name, linetype=factor(size))) + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")
-
+  geom_line(aes(color=factor(size))) + 
+  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
+  guides(color=FALSE)
 ```
@@
 We also see that TIC is smoother when moving into the special view.
+(As I understand TIC is however not equitable, i.e. it will prefer linear relations, and it may be instructive to optimise MIC and TIC)
+
+FIXME need to check the MIC_e and TIC_e, supposedly faster computation and better discrimination!
 

### 2018-09-18 scagnostics.Rmd
Added CKM example.

@@
 
-(Motivation - something along the lines of the following paragraph, but more understandable, maybe giving concrete example - e.g. dec vs time from neutron star posterior?)
-One motivation to consider projection pursuit indices beyond the ones considered so far is coming from physics applications where a standard method is the determination of physcis parameters by means of global fits to large experimental datasets.
-Typically models will come with a handful of free parameters, model predictions will be calculated by means of numeric tools and the best fit parameters are evaluated by comparing predictions to a large number of observations.
-Fit results will encode relations between the model parameters dictaded by the available observations, for example if two parameters control predictions for a well constrained observable, looking at the distribution of parameter combinations which are a good fit will be instructive.
-A standard example is to study the likelihood or posterior distribution in pair-wise 2-d plots, where the selection of parameters if often guided by knowledge of the underlying physics.
-A major challenge is the progressing complexity in all three aspects: models with larger parameter spaces are being studied, computation of observable predictions is increasingly complex and the number of experimental observations relevant to a given model is growing.
-Methods for the detection of "interesting" views that can aid physics intuition are therefore more and more relevant.
+Here we want to extend the list of projection pursuit indices with the aim of detecting non-trivial dependence between parameter combinations, i.e.
+we want to find complicated functional dependence as well as association not described by a function (e.g.
+an ellipse or non-coexistence).
+Our main motivation stems from particle physics applications where such type of association is commonly implied by experimental and theoretical constraints.
+A well-studied example would be the so-called CKM triangle.
+Here an instructive parametrisation is known, and measurements can pin-point the physical value of the parameters.
+Fig.
+\ref{fig:CKM} (left) shows a schematic view of the 2-parameter dependence for various constraints.
+Fig.
+\ref{fig:CKM} (right) shows the current state of the art constraints.
+We expect this type of relations are common in particle physics, and their detection will aid understanding of the relevant physics, and as a consequence improve modelling and parametrisation.
+The task of uncovering the relevant features is however becoming more challenging as more complicated models are fit to larger sets of experimental results.
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figure/CKM-0104062.pdf}
+\includegraphics[width=0.5\textwidth]{figure/rhoeta_large_ckmfitter.png}
+\caption{CKM triangle constraints, schematic view shown on the left, taken from \cite{Hocker2001xe}, current picture shown on the right, from \cite{Charles2004jd}.}\label{fig:CKM}
+\end{figure}
 

### 2018-09-19 scagnostics.Rmd
References added, start using neutron star data as example.

@@
 
-# Scagnostics indices
-\label{sec:indexes}
-
-## Individual
-
-## Two group differences
-
-## Multiple group differences
-
-# Application to particle physics data
+# Application to physics data
 \label{sec:phys}
 
+We now want to apply the developped tools to distributions from realistic physics application.
+Here we choose as an example to use posterior samples obtained when fitting source parameters to the observed gravitatioal wave signal GW170817 @PhysRevLett119161101.
+The fit details were described in @Abbott2018exr and the corresponding data is available online @ligoData.
+
 ## Data description
+The model describes a neutron star merger and contains 6 free parameters, i.e.
+each neutron star is described by its mass and radius, and a so-called tidal deformability parameter $\Lambda$.
+A posterior sample containing 2538 points is available @ligoData and can be used to constrain the parameters.
+For details about the modelling and statistical procedure see @Abbott2018exr.
+Show corner plot here, comment on well constrained relation between m1 and m2 ("chirp mass")
+For further study we therefore remove m2 from the dataset and use the guided tour working with the remaining 5 dimensions.
 
-## Software setup
+Use spline2d index, find relation between 3 parameters, ask Rory for interpretation
 
-## Scagnostics of grand tour projections
+## Higher D
+To complicate things can work with bigger dataset that Rory sent me, need to check if this is public, also need help with interpretation, but here optimising TIC finds interesting projection
 

### 2018-09-19 scagnostics.Rmd
Filling in example description.

@@
 A posterior sample containing 2538 points is available @ligoData and can be used to constrain the parameters.
+Here we use the "Parametrized-EoS\_nomaxmass\_posterior\_samples.dat" dataset.
 For details about the modelling and statistical procedure see @Abbott2018exr.
-Show corner plot here, comment on well constrained relation between m1 and m2 ("chirp mass")
-For further study we therefore remove m2 from the dataset and use the guided tour working with the remaining 5 dimensions.
+A common representation of posterior samples used for interpretatin are so-called "corner plots", i.e.
+density displays for all 2-d parameter combinations as well as the 1-d profile, highlighting empirically evaluated regions of highest posterior density, which correspond to credibility regions at a given level.
+Use Rorys script to make nice version of his corner plot.
+Here we generate such a corner plot using the python package "ChainConsumer", highlighting the x and y % CIs.
+We point out the following features seen from Fig.
+XX: the combination of the two mass paremters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
+We highlight the non-linear relation between the radius $R_1$ and $\Lambda$ parameter and the two-pronged structure of the radius $R_1$.
+For further study we therefore remove m2 from the dataset.
+We first look at the distribution in a grand tour display, where we use a scatter plot display combined with alpha blending to study the density distribution in the 5-d parameter space.
+The grand tour indicates that the data points are found on a curved surface in the five dimensional space.
+This may stem from a non-linear relation between the parameters, which we will try to uncover using the above developped formalism for new guided tour index functions.
 
-Use spline2d index, find relation between 3 parameters, ask Rory for interpretation
+## Using the guided tour
+Given the observations from the grand tour display we expect a functional relation between some of the parameters, and we therefore optimise the "splines2d" index function.
+Note that all parameters are standardised to the range $[0,1]$ before running the optimisation.
+Show final view found from the guided tour, give corresponding projection matrix, make corresponding plot from combination of original parameters.
 

### 2018-09-20 scagnostics.Rmd
No significant changes.

@@
 For further study we therefore remove m2 from the dataset.
-We first look at the distribution in a grand tour display, where we use a scatter plot display combined with alpha blending to study the density distribution in the 5-d parameter space.
+We first look at the distribution in a grand tour display, where we use a scatter plot display combined with alpha transperancy to study the density distribution in the 5-d parameter space.
 The grand tour indicates that the data points are found on a curved surface in the five dimensional space.

### 2018-09-26 scagnostics.Rmd
Added first neutron star example.

@@
 density displays for all 2-d parameter combinations as well as the 1-d profile, highlighting empirically evaluated regions of highest posterior density, which correspond to credibility regions at a given level.
-Use Rorys script to make nice version of his corner plot.
-Here we generate such a corner plot using the python package "ChainConsumer", highlighting the x and y % CIs.
+The information can similarly be presented in a scatter plot matrix using transparency to convey the density information.
+Since in the following we work with 2-d scatter plots we chooose to show the data in this fashion, see Fig.
+
+```{r neutronStarSPLOM}
+nsD <- read_csv("data/samples.csv") 
+ggpairs(nsD, lower=list(continuous = wrap("points", alpha = 0.05)))
+```
+
 We point out the following features seen from Fig.
@@
 Show final view found from the guided tour, give corresponding projection matrix, make corresponding plot from combination of original parameters.
+```{r neutronStarEq,  results="hide"}
+nsD <- read_csv("data/samples.csv") 
+nsM <- nsD %>%
+  select(-m2) %>% # remove m2 to avoid finding well known correlation
+  rescale() # rescaling, this returns a matrix
+nseTour <- save_history(nsM, guided_tour(splineIndex()))
+nseTourFull <- as.list(interpolate(nseTour))
+iLast <- length(nseTourFull)
+fProj <- nseTourFull[[iLast]]
+dProj <- as.tibble(nsM %*% fProj)
+ggplot(dProj, aes(V1,V2)) + geom_point()
+```
+
+```{r nseMatrix, echo=FALSE}
+knitr::kable(fProj ,  caption = "Matrix representation of best projection.")
+```
+
+The resulting projection clearly shows how the data selects a region that is well defined as a combination of several parameters.
+From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection.
+Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{nsePlotOrig} (right), which we compare to the next closest bivariate plot shown on the left.
+It is clear that in this example additonal information is found when considering general projections rather than focussing only on bivariate relations.
+
+```{r nsePlotOrig}
+p1 <- ggplot(nsD, aes(R1, L1)) + geom_point(alpha = 0.05)
+p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05)
+grid.arrange(p1, p2, ncol=2)
+```
 

### 2018-09-26 scagnostics.Rmd
Small changes.

@@
 On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparant much closer to the final projection, making this structure more challenging to detect with a guided tour.
+Note that the various indices appear less noisy than for the grand tour shown before, but should note that the tour path here is shorter, and the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplfied in the previous views).
 (What does this mean for optimisation? How can I detect such "smaller scale increase"? Maybe try "search_better" instead of search_geodesic? increase alpha and decrease cooling? increase max.tries?)
@@
 ```{r testpipe2}
-
 iLast <- length(pipeTourFull2)
@@
 ```{r neutronStarEq,  results="hide"}
+set.seed(2018)
 nsD <- read_csv("data/samples.csv") 
@@
 ```
+
+This is reflecting the dependence of the deformability parameter $\Lambda$ on the so-called compactness which is proportional to $m/R$.
+QUESTIONS: can see this when plotting e.g.
+R1/m1 vs L1, why also with the linear combination used in the plot (only works if I get the factor about right)?
+not clear how this was built into the simulation, seems not strict equality mentioned in the paper
 

### 2018-09-26 scagnostics.Rmd
Some text restructuring.

@@
 Most indexes developed focus on finding projections with cluster structure, outliers, or separations between known groups.
-Here, we are interested in finding projections where the shapes are odd, or where known groups have different shapes despite having the same mean and variance.
-The new indexes are based on scagnostics, which were originally developed to choose pairs of interesting variables from a large collection.
-An examination of the properties of the scagnostics is included, and code in the form of an R package is provided.
+Here, we are interested in finding projections revealing potentially complex relations between combinations of parameters, by combining the notion of projection pursuit with index functions developed for the detection of interesting bivariate views in the data, including scagnostics and maximum information coefficient.
 
@@
 Some have been developed to detect separations between known groups.
-However, if one wants to find unusual shapes, or shape differences between groups there are no options.
-One reason, is that it is hard, because defining what is meant by shape is not straightforward.
-
+However, if one wants to find complicated codependence between parameters there are no options.
+On the other hand, for the related question of the selection of pair-wise scatter plots in datasets with large numbers of parameters, a number of different methods have been developed, from detection of linear relations by correlation coefficients, to more complex measures aiming to capture generally the information in a given view.
 Here we want to extend the list of projection pursuit indices with the aim of detecting non-trivial dependence between parameter combinations, i.e.
@@
 an ellipse or non-coexistence).
-Our main motivation stems from particle physics applications where such type of association is commonly implied by experimental and theoretical constraints.
-A well-studied example would be the so-called CKM triangle.
-Here an instructive parametrisation is known, and measurements can pin-point the physical value of the parameters.
-Fig.
-\ref{fig:CKM} (left) shows a schematic view of the 2-parameter dependence for various constraints.
-Fig.
-\ref{fig:CKM} (right) shows the current state of the art constraints.
-We expect this type of relations are common in particle physics, and their detection will aid understanding of the relevant physics, and as a consequence improve modelling and parametrisation.
-The task of uncovering the relevant features is however becoming more challenging as more complicated models are fit to larger sets of experimental results.
-
-\begin{figure}
-\includegraphics[width=0.5\textwidth]{figure/CKM-0104062.pdf}
-\includegraphics[width=0.5\textwidth]{figure/rhoeta_large_ckmfitter.png}
-\caption{CKM triangle constraints, schematic view shown on the left, taken from \cite{Hocker2001xe}, current picture shown on the right, from \cite{Charles2004jd}.}\label{fig:CKM}
-\end{figure}
+Our motivation is based on potential physics applications aiding the interpretation of model fits to experimental results.
+Let us consider a physical model with a set of $n$ free parameters that cannot be measured directly but should be determined from fitting a set of $p$ experimental observations for which predictions can be calculated when fixing all $n$ parameters.
+Note here that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numeric calculations.
+Moreover, a single prediction can be a complicated function of all free parameters.
+In addition we often deal with large number of observations, $p \sim 100-1000$ as well as sizable parameter spaces $n \sim 10$.
+The results of a fit are generally interpreted using visualisations selected based on physics intuition and previous knowledge, but the above stated observations suggest that some correlations clearly dictated by the data may be missed, as we are typically only studying selected views of the results.
+The idea of finding special views in scatter plots fits naturally with the concept of using posterior samples obtained in a Bayesian fit.
+The idea is that by finding special views of the posterior samples we gain insight into the physics relevant to accurately describe the experimental data.
+Apart from improved intuition, finding special regions in parameter space is important as it allows us to further scrutinize the model.
+There are many examples of this, for example in particle physics detailed studies have been performed e.g.
+in the alingment limit of the 2HDM or in coannihilation regions in models of dark matter.
 
@@
 One approch are scatterplot diagnostics, "scagnostics", developed by @scag, and explored in @WW08.
-They are calculated from a set of elementary graph-theoretic building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST).
+\footnote{Scagnostics are available in two R packages @LWscagR and @HWscagR.
+This work is built upon the scagnostics package by @HWscagR because the underlying base is built using C, works across platforms.} They are calculated from a set of elementary graph-theoretic building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST).
 These are used to define eight measures: Outlying, Skewed, Sparse, Clumpy, Striated, Convex, Skinny and Stringy.
@@
 Alternative methods suggested by (Katrins Thesis or maybe her R package?) suggest using parallel coordinates, glyphs or a so-called scagramm (based on the idea of corrgrams) that shouold be coupled to interactive brushing for selection of pair-wise scatter plots to be displayed.
-(FIXME just give reference rather than all definitions here?)
-(Issues with scagnostics already for selection of pairs have been described in Katrins Thesis, maybe have short summary of that here?)
+(moved scagnostics definitions to appendix, not sure if needed at all.)
+(Issues with scagnostics already for selection of pairs have been described in Katrins Thesis, should have short summary of that here?)
+
 Given the various shortcomings of the scagnostics approach we consider several other measures that can directly be interpreted as ranking of information contained in a given 2-d view.
@@
 where $Var(.)$ is the variance and $res_{X\sim Y}$ are the residuals in the spline model where $X$ is considered a function of $Y$.
-The second new measure introduced in @mbgraphic is aiming to detect also non-functional relations and is based on the idea of distance correlation @szekely2007. (should I repeat definitions here?)
+The second new measure introduced in @mbgraphic is aiming to detect also non-functional relations and is based on the idea of distance correlation [@szekely2007]. (need to add definition)
 Finally we will also consider the maximal and total information coefficient (MIC and TIC) defined in @Reshef1518 which are based on mutual information found when imposing grids on pair-wise variable plots.
-These functions are availabel in R via the interface implemented in the minerva package @minerva.
+These functions are availabel in R via the interface implemented in the minerva package [@minerva].
 (read paper and add more detail.
@@
 the way that binning is being built into it may make difference in how MIC/TIC may be smoother?)
-
-Scagnostics are scatterplot diagnostics. They were primarily developed to extract pairs of interesting variables from a large collection.
-XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia).
-Nine graph-theoretic measures have been defined to characterise scatterplots.
-They are calculated from a set of elementary building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST).
-For definiteness all variables are rescaled to a [0,1] intervall, such that all measures will take values between 0 and 1 as well.
-
-Note that in practice, for efficiency and robustness, the measures are calculated from binned data, and after removing outliers.
-Moreover, scaling factors are used to mitigate dependence on sample size.
-Scagnostics are available in two R packages @LWscagR and @HWscagR. This work is built upon the scagnostics package by @HWscagR because the underlying base is built using C, works across platforms.
 
@@
 Software to run a projection pursuit guided tour is available in the R package, "tourr" [@tourr].
+(Motivation strongly relates to intuition, i.e. guided tour, local tour and manual tour may be useful to understand structure in the multivariate context.)
 (Even just for the optimisation part the guided tour implementation is useful.)
@@
 
-The paper investigates the nature of scagnostics, particularly when considered as candidates for projection pursuit indices (Section \ref{sec:investigate})).
-It develops indexes for finding shapes as defined by the scagnostics, and for finding different shapes between multiple groups (Section \ref{sec:indexes}).
-Changes to the optimisation of the projection pursuit guided tour were needed to accommodate some of the complexities of the scagnostic indexes.
-The new guided tour methods are applied to contemporary problems from paarticle physics (Section \ref{sec:phys}).
-The paper is accompanied by the R package "binostics" available on CRAN, and the new indices have been implemented in the "tourr" package.
+The paper investigates the nature of the above defined measures, particularly when considered as candidates for projection pursuit indices (Section \ref{sec:investigate})).
+The new guided tour methods are applied to posterior distributions from physics examples (Section \ref{sec:phys}).
 
-# Investigation of scagnostics
+# Investigation of indices
 \label{sec:investigate}
@@
 
+((REVISE - where should measure definitions go? now basically in introduction.))
 The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
@@
 
-To use scagnostics for projection pursuit requires that each of the measures for functions that are relatively smooth over the space of all 2D projections, and that they can be computed rapidly enough that the display can be updated many times a second during optimisation.
+((maybe better to start with something like this, though maybe too obvious?))
+The various measures defined in the introduciton are designed to be computed for all pairs of variables, followed either by selection based on comparing scagnostics scores between all pairs, or by an absolute ranking from most to least interesting (e.g.
+when using TIC).
+Here we modify the usage as follows.
+The tour implementation is used to generate projections of the multivariate dataset down to two dimensions, i.e.
+the projected data can be represented as a scatter plot.
+In this projection the data is represented by two parameters that are defined as combinations of the input parameters, and this parametrisation is passed into the index function calculation to obtain a score for the selected view.
+Using the guided tour implementation the index functions are optimized such that we find the generalised 2-d parametrisation yielding the most intersting view of the data.
+
+To use an index function for projection pursuit requires that the measure be relatively smooth over the space of all 2D projections, and we will first investigate the smoothness over randomly selected projections.
+An additional consideration is the computing time, ideally fast computation is preferred as the guided tour can be viewed as the index is being optimized.
+However, a compuationally more costly index may be preferred for better detection properties, and one can consider recording the guided tour for viewing after the optimisation has converged.
+Finally we note that another important consideration is how the index function changes as the tour moves into the "optimal" view, this will however depend mainly on the distribution rather than the selected index function.
+(we will see this in the examples)
 
@@
 We consider three example distributions in n=6 dimensions with p randomly selected points $x_i, i = 1,..,n$:
-1. p points on an n-dimensional sphere as an example of dataset without special views
+
+1. p points on an n-dimensional sphere randomly selected using the geozoo package, as an example of dataset without special views
 2.
-p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not showing functional dependence, and where sampling boundaries are apparant in the distribution
+p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparant in the distribution
 3.
@@
 To get an overview of the dataset we show views of the first two parameters showing the independent distribution and the last two parameters showing the special view.
-The enforces relationships will also affect the overall distribution as can be seen from (fig:dataPlotsDensity).
+The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractios of V6 (and V5 for the Pipe distribution) will be distinct from uninformed views of distributions 2 and 3.
+We select two samples for each distribution, with 100 and 1000 sampled points respectively.
 
@@
 
-We next compare the various indexvalues for three views: V1 vs V2, V5 vs V6 and V1 vs V6 (capturing differences from the different distribution in V6).
+We next compare the various index values for three views: V1 vs V2, V5 vs V6 and V1 vs V6 (capturing differences from the different distribution in V6).
 For the Sphere distribution by construction we see only random fluctuations between the different views, which can be sizable in particular for small datasets.
 Considering next the Pipe distribution, we see that skinny, stringy and dcor2d are increased for the special view, but in particular MIC and TIC index are much larger and may be used to detect this type of structure in the dataset.
+Moreover, minimising the convex index (or maximising 1-convex) should also allow us to detect the special view.
 Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fuctional dependence.

### 2018-09-27 scagnostics.Rmd
Filling in text.

@@
 The resulting search will quickly steer away from the most unintresting projections, followed by a broad scan of the thus identified region.
-While this does not maximise the index function, it provides an informed starting plane that can be used in a second step where we use the "search_geodesic" optimisatino to improve the optimisation.
+While the lack of cooling means that this will not converge towards the maximum, each step in this first search will increase the index value, thus it can be used to provide an informed starting plane that can be used in a second step where we use the "search_geodesic" optimisatino to improve the index further towards the maximum.
 The described procedure is CPU time intensive, however it is promising that we manage to uncover such small scale structure in the distribution.
@@
 This gives some indication on how to handle difficult datasets. As we will see below most of the time standard "search_geodesic" optimisation will be efficient enough to uncover the interesting views.
+
+The results are shown below, first Fig.
+\ref{fig:pipeFirstRun} shows how the considered index values evolve in the scouting phase.
+Note that this is the interpolated path and there are small local maxima and minima along the index value, especially because we keep the step size in the guided tour large.
+We can see the clear maximum of the TIC index in the final projection, as well as clear minimum of the convex index that carries similar information On the other hand we see that MIC or the dcor2d index are not suitable for uncovering the special view here.
 
@@
 
-```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Guided tour optimising MIC index for Pipe 1000 dataset."}
+```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Index value on tour path when scouting for high TIC index values."}
 pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
@@
 
-```{r testpipe}
+We next show in Fig \ref{fig:testpipe} the final view corresponding to the highest index found in the TIC scouting phase.
+While this is clearly not the optimal view of the pipe feature, we can already guess the type of dependence in the distribution, and we expect to be able to find the maximum when using this plane as a starting point of the geodesic optimisation.
+Indeed performing a standard guided tour with the TIC index and the identified starting plane efficiently identifies the optimal view shown in Fig ...
+((Probably should remove second trace plot as it does not give much information, maybe show final view in same figure as final scouting view...))
+
+```{r testpipe, out.width=".5\\textwidth", fig.cap="Final view discovered by the scouting phase."}
 iLast <- length(pipeTourFull)
@@
 
-```{r testpipe2}
+```{r testpipe2, out.width=".5\\textwidth"}
 iLast <- length(pipeTourFull2)
@@
 ### Finding sine waves
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Sine 1000 dataset.", results="hide"}
+We next work with distribution 3 and aim to identify the view showing functional dependence using the splines2d index.
+Given the results from Fig 6 ((add proper ref)), we expect to be able to efficiently identify this view without relying on the two step procedure described above.
+Indeed the guided tour quickly converges in this case, the evolution of the index functions is shown in Fig \ref{fig:findsine}.
+Here there are four index functions that appear to be suitable for the optimisation: splines2d, dcor2d, MIC and TIC.
+On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimisation when considering them as index functions.
+The final view identified by the guided tour is shown in Fig \ref{}
+
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide"}
 set.seed(2018)
@@
 
-```{r testsine}
+```{r testsine, out.width=".5\\textwidth", fig.cap="Final view identified by the optimisation."}
 iLast <- length(sineTourFull)

### 2018-09-28 scagnostics.Rmd
added some outlines for sections

@@
 ---
-title: Using Scagnostics to find Interesting Projections of Multivariate Data
+title: An investigation into using scagnostics interesting projections of high-dimensional data
 
@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
+  fig.show = 'hold', cache = FALSE, external = TRUE, dev = "pdf",
   fig.height = 5, fig.width = 8, out.width = "\\textwidth"
@@
 
+<!--
+Motivation
+PP
+Scagnostics
+(section needs shortening)
+-->
 The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high-(say $p$)dimensional data for "interesting" low- dimensional projections ($k=1$ or $2$ usually).
@@
 Some have been developed to detect separations between known groups.
-However, if one wants to find complicated codependence between parameters there are no options.
+However, if one wants to find complicated co-dependence between parameters there are no options.
 On the other hand, for the related question of the selection of pair-wise scatter plots in datasets with large numbers of parameters, a number of different methods have been developed, from detection of linear relations by correlation coefficients, to more complex measures aiming to capture generally the information in a given view.
@@
 There are many examples of this, for example in particle physics detailed studies have been performed e.g.
-in the alingment limit of the 2HDM or in coannihilation regions in models of dark matter.
+in the alignment limit of the 2HDM or in co-annihilation regions in models of dark matter.
 
 A related question is the selection of pair-wise scatter plots in datasets with large numbers of parameters, for which different methods have been developed.
-One approch are scatterplot diagnostics, "scagnostics", developed by @scag, and explored in @WW08.
+One approach are scatterplot diagnostics, "scagnostics", developed by @scag, and explored in @WW08.
 \footnote{Scagnostics are available in two R packages @LWscagR and @HWscagR.
@@
 While in principle each measure could be used on its own to define a ranking, the original idea suggested considering pair-wise scatter plots of the scagnostics measures for identification of "characteristic" and "unusual" 2-d views in the original parameter space.
-Alternative methods suggested by (Katrins Thesis or maybe her R package?) suggest using parallel coordinates, glyphs or a so-called scagramm (based on the idea of corrgrams) that shouold be coupled to interactive brushing for selection of pair-wise scatter plots to be displayed.
+Alternative methods suggested by (Katrins Thesis or maybe her R package?) suggest using parallel coordinates, glyphs or a so-called scagramm (based on the idea of corrgrams) that should be coupled to interactive brushing for selection of pair-wise scatter plots to be displayed.
 (moved scagnostics definitions to appendix, not sure if needed at all.)
@@
 Finally we will also consider the maximal and total information coefficient (MIC and TIC) defined in @Reshef1518 which are based on mutual information found when imposing grids on pair-wise variable plots.
-These functions are availabel in R via the interface implemented in the minerva package [@minerva].
+These functions are available in R via the interface implemented in the minerva package [@minerva].
 (read paper and add more detail.
@@
 A projection pursuit index, a function of all possible projections of the data, invariably has many "hills and valleys" and "knife-edge ridges" because of the varying shapes in the underlying density of observations from one projection to the next.
-From an exploratory data analysis perpective it is interesting to combine numerical optimisation with visualisation to watch the structure of the data moving into a maximum, to jump away from this projection and follow the optimisation again, invariably moving from local maxima to local maxima, and perhaps even a global maximum.
+From an exploratory data analysis perspective it is interesting to combine numerical optimisation with visualisation to watch the structure of the data moving into a maximum, to jump away from this projection and follow the optimisation again, invariably moving from local maxima to local maxima, and perhaps even a global maximum.
 Each of these maxima can reveal different information about the data.
@@
 
+# Constructing a projection pursuit index
+\label{sec:construct}
+
+<!--
+Definition of pp index
+Optimisation
+Illustration of what is already available in tourr
+New index explanations
+-->
+
 # Investigation of indices
 \label{sec:investigate}
+
+<!--
+Criteria for a good index
+  - smoothness
+  - speed
+  - fine and broad structure detection
+Simulation study
+  - example data 
+  - sample size, and other factors used
+Overview procedures 
+  - look at index values of pairs of variables
+  - tour betwen pairs of variables
+  - optimisation path
+-->
 
@@
 ((maybe better to start with something like this, though maybe too obvious?))
-The various measures defined in the introduciton are designed to be computed for all pairs of variables, followed either by selection based on comparing scagnostics scores between all pairs, or by an absolute ranking from most to least interesting (e.g.
+The various measures defined in the introduction are designed to be computed for all pairs of variables, followed either by selection based on comparing scagnostics scores between all pairs, or by an absolute ranking from most to least interesting (e.g.
 when using TIC).
@@
 In this projection the data is represented by two parameters that are defined as combinations of the input parameters, and this parametrisation is passed into the index function calculation to obtain a score for the selected view.
-Using the guided tour implementation the index functions are optimized such that we find the generalised 2-d parametrisation yielding the most intersting view of the data.
+Using the guided tour implementation the index functions are optimized such that we find the generalised 2-d parametrisation yielding the most interesting view of the data.
 
@@
 An additional consideration is the computing time, ideally fast computation is preferred as the guided tour can be viewed as the index is being optimized.
-However, a compuationally more costly index may be preferred for better detection properties, and one can consider recording the guided tour for viewing after the optimisation has converged.
+However, a computationally more costly index may be preferred for better detection properties, and one can consider recording the guided tour for viewing after the optimisation has converged.
 Finally we note that another important consideration is how the index function changes as the tour moves into the "optimal" view, this will however depend mainly on the distribution rather than the selected index function.
@@
 2.
-p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparant in the distribution
+p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution
 3.
@@
 To get an overview of the dataset we show views of the first two parameters showing the independent distribution and the last two parameters showing the special view.
-The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractios of V6 (and V5 for the Pipe distribution) will be distinct from uninformed views of distributions 2 and 3.
+The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of V6 (and V5 for the Pipe distribution) will be distinct from uninformed views of distributions 2 and 3.
 We select two samples for each distribution, with 100 and 1000 sampled points respectively.
@@
     geom_point() +
-    ggtitle(pLabels[i])
+    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
   pL[[i]] <- pC
@@
 ```
+
+<!--
+its possible to do a single figure legend - need to look up code for this
+-->
 
@@
 Moreover, minimising the convex index (or maximising 1-convex) should also allow us to detect the special view.
-Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fuctional dependence.
+Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fictional dependence.
 
@@
     mineRes <- mine(dprj[,1], dprj[,2])
-    indexOverview <- add_row(indexOverview, convex=scagRes["Convex"],
-                    clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
-                    stringy=scagRes["Stringy"], dcor2d=dcorRes,
+    indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
+                    clumpy=scagRes[,"Clumpy"], skinny=scagRes[,"Skinny"],
+                    stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
                     splines2d=splineRes, MIC=mineRes$MIC,
@@
 Note that the value of TIC is strongly dependent on the input distribution and we have therefore normalised the TIC measure by its maximum for each individual dataset, all other measures are reproduced as reported by the calculation.
+
+<!-- index values shouldn't be affected by sample size, on average it should be
+same value. Maybe some need to be normalised by n
+-->
 
@@
     sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
-                  convex=scagRes["Convex"],
-                  clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
-                  stringy=scagRes["Stringy"], dcor2d=dcorRes,
+                  convex=scagRes[,"Convex"],
+                  clumpy=scagRes[,"Clumpy"], skinny=scagRes[,"Skinny"],
+                  stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
                   splines2d=splineRes, MIC=mineRes$MIC,
@@
 We therefore show 2-d view defined by the tour step $t=58$, with projection matrix
+
 ```{r t58matrix, echo=FALSE}
@@
 ```
+
 i.e. while the last two parameters enter significantly, they will not dominate the view and we do not expect to fully observe the built in features.
+
 ```{r plot58, fig.width=6, fig.height=7, fig.cap="View of all datasets defined by at tour step 58."}
@@
 ```
+
 We see from Fig. ((how to add cross references?)) that the built in special views cannot be seen in this projection, but confirm that the sine 1000 dataset presents some correlation in this view.
@@
   rbind(getProj(pipe1000, t2full, "pipe", 1000))
-
 ```
@@
 We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise.
-On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparant much closer to the final projection, making this structure more challenging to detect with a guided tour.
-Note that the various indices appear less noisy than for the grand tour shown before, but should note that the tour path here is shorter, and the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplfied in the previous views).
+On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
+Note that the various indices appear less noisy than for the grand tour shown before, but should note that the tour path here is shorter, and the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
 (What does this mean for optimisation? How can I detect such "smaller scale increase"? Maybe try "search_better" instead of search_geodesic? increase alpha and decrease cooling? increase max.tries?)
@@
 
-Here we also see how the existing indices in the tourr package are inadequat for the detection of these type of dependence, the holes index taking maximum values before the final view and decreasing when moving into the special view.
+Here we also see how the existing indices in the tourr package are inadequate for the detection of these type of dependence, the holes index taking maximum values before the final view and decreasing when moving into the special view.
 On the other hand, surprisingly the cmass index increases when moving into the special view, but the intermediate minimum makes it unsuitable as a projection pursuit index here.
@@
 The first step consists in "scouting" projections.
-Here we use the "search_better" optimisation available with the tourr package, with a standard search window $\alpha = 0.5$ and disabeling the cooling (in practice this can be done by fixing the cooling parameter to $1$), together with a large number of tries (setting max.tries to 5000 here).
-The resulting search will quickly steer away from the most unintresting projections, followed by a broad scan of the thus identified region.
-While the lack of cooling means that this will not converge towards the maximum, each step in this first search will increase the index value, thus it can be used to provide an informed starting plane that can be used in a second step where we use the "search_geodesic" optimisatino to improve the index further towards the maximum.
+Here we use the "search_better" optimisation available with the tourr package, with a standard search window $\alpha = 0.5$ and disabling the cooling (in practice this can be done by fixing the cooling parameter to $1$), together with a large number of tries (setting max.tries to 5000 here).
+The resulting search will quickly steer away from the most uninteresting projections, followed by a broad scan of the thus identified region.
+While the lack of cooling means that this will not converge towards the maximum, each step in this first search will increase the index value, thus it can be used to provide an informed starting plane that can be used in a second step where we use the "search_geodesic" optimization to improve the index further towards the maximum.
 The described procedure is CPU time intensive, however it is promising that we manage to uncover such small scale structure in the distribution.
@@
 
-We now want to apply the developped tools to distributions from realistic physics application.
-Here we choose as an example to use posterior samples obtained when fitting source parameters to the observed gravitatioal wave signal GW170817 @PhysRevLett119161101.
+We now want to apply the developed tools to distributions from realistic physics application.
+Here we choose as an example to use posterior samples obtained when fitting source parameters to the observed gravitational wave signal GW170817 @PhysRevLett119161101.
 The fit details were described in @Abbott2018exr and the corresponding data is available online @ligoData.
@@
 The model describes a neutron star merger and contains 6 free parameters, i.e.
-each neutron star is described by its mass and radius, and a so-called tidal deformability parameter $\Lambda$.
+each neutron star is described by its mass and radius, and a so-called tidal deform ability parameter $\Lambda$.
 A posterior sample containing 2538 points is available @ligoData and can be used to constrain the parameters.
@@
 For details about the modelling and statistical procedure see @Abbott2018exr.
-A common representation of posterior samples used for interpretatin are so-called "corner plots", i.e.
+A common representation of posterior samples used for interpretation are so-called "corner plots", i.e.
 density displays for all 2-d parameter combinations as well as the 1-d profile, highlighting empirically evaluated regions of highest posterior density, which correspond to credibility regions at a given level.
 The information can similarly be presented in a scatter plot matrix using transparency to convey the density information.
-Since in the following we work with 2-d scatter plots we chooose to show the data in this fashion, see Fig.
+Since in the following we work with 2-d scatter plots we choose to show the data in this fashion, see Fig.
 
@@
 We point out the following features seen from Fig.
-XX: the combination of the two mass paremters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
+XX: the combination of the two mass parameters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
 We highlight the non-linear relation between the radius $R_1$ and $\Lambda$ parameter and the two-pronged structure of the radius $R_1$.
 For further study we therefore remove m2 from the dataset.
-We first look at the distribution in a grand tour display, where we use a scatter plot display combined with alpha transperancy to study the density distribution in the 5-d parameter space.
+We first look at the distribution in a grand tour display, where we use a scatter plot display combined with alpha transparency to study the density distribution in the 5-d parameter space.
 The grand tour indicates that the data points are found on a curved surface in the five dimensional space.
-This may stem from a non-linear relation between the parameters, which we will try to uncover using the above developped formalism for new guided tour index functions.
+This may stem from a non-linear relation between the parameters, which we will try to uncover using the above developed formalism for new guided tour index functions.
 
@@
 Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{nsePlotOrig} (right), which we compare to the next closest bivariate plot shown on the left.
-It is clear that in this example additonal information is found when considering general projections rather than focussing only on bivariate relations.
+It is clear that in this example additional information is found when considering general projections rather than focusing only on bivariate relations.
 
@@
 
-* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantil edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long.
+* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantile edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long.
 The outlying measure is then calculated as $c_{outlying} = \frac{\mathrm{Total\ length\ of\ long\ edges}}{\mathrm{Total\ length\ of\ all\ edges}}$.
@@
 * Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
-* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimiter to the area.
+* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimeter to the area.
 We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a circle, and values close to one for skinny polygons.
-* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{lenght(MST)}$ where diameter is longest connected path, and lenght the total lenght (sum of all edgest), i.e.
+* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{length(MST)}$ where diameter is longest connected path, and length the total length (sum of all edges), i.e.
 we are measuring branch structure, no branches means $c_{stringy}=1$.

### 2018-09-28 scagnostics.Rmd
cleaned up intro

@@
 <!-- 
-Motivation
-PP
-Scagnostics
-(section needs shortening)
+Main point of paper:
+
+- Some metrics for exploring high-dimensional data by finding interesting pairs of variables, can be used to find interesting projections, but some have problems. 
+- What are the ways to evaluate.
+- What adjustments can be done.
+- How it can be applied
 -->
-The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high-(say $p$)dimensional data for "interesting" low- dimensional projections ($k=1$ or $2$ usually).
+
+Organisation of this section:
+
+- Motivation first?
+- PP
+- Scagnostics
+- section needs shortening
+
+The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($k=1$ or $2$ usually).
 The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $k$-dimensional projection of $p$-dimensional data.
 This criterion function is optimized over the space of all $k$-dimensional projections of $p$-space, searching for both global and local maxima.
-It is hoped that the resulting solutions reveal low-dimensional structure in the data not necessarily found by methods such as principal component analysis.
+It is hoped that the resulting solutions reveal low-dimensional structure in the data not found by methods such as principal component analysis.
 
 A large number of projection pursuit indices have been developed [@f87] [@CBC92] [@Lee05] [@AHC02] [@CEM:CEM2568] [@JS87] [@5508437].
-Many of these are developed to detect departure from multivariate normal, which includes clusters or outliers, or nonlinearity.
-Some have been developed to detect separations between known groups.
-However, if one wants to find complicated co-dependence between parameters there are no options.
-On the other hand, for the related question of the selection of pair-wise scatter plots in datasets with large numbers of parameters, a number of different methods have been developed, from detection of linear relations by correlation coefficients, to more complex measures aiming to capture generally the information in a given view.
-Here we want to extend the list of projection pursuit indices with the aim of detecting non-trivial dependence between parameter combinations, i.e.
-we want to find complicated functional dependence as well as association not described by a function (e.g.
-an ellipse or non-coexistence).
-Our motivation is based on potential physics applications aiding the interpretation of model fits to experimental results.
-Let us consider a physical model with a set of $n$ free parameters that cannot be measured directly but should be determined from fitting a set of $p$ experimental observations for which predictions can be calculated when fixing all $n$ parameters.
-Note here that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numeric calculations.
+Many of these are developed to detect departure from multivariate normal, which includes clusters or outliers.
+Some have been developed to detect separations between known groups [REF].
+Less work has been done on indexes to find nonlinear dependence between variables.
+
+<!-- check Perisic and Posse (2012) and recent paper by Austrian guys -->
+
+A related topic is variable selection.
+With high-dimensional data, even plotting all pairs can be daunting, so "scagnostics" [@scag] [@WW08] have been developed to find the most interesting pairs of variables.
+There are 8 scagnostics, vividly named: "Outlying", "Skewed", "Sparse", "Clumpy", "Striated", "Convex", "Skinny" and "Stringy".
+Our question is whether these can be extended into projection pursuit indexes that can be used to find two-dimensional projections of high-dimensional data with unusual features.
+
+Our motivation is based on applications in physics, to aid the interpretation of model fits on experimental results.
+Consider a physical model with a set of $m$ free parameters, that cannot be measured directly and are determined by fitting a set of $p$ experimental observations, for which predictions can be calculated when fixing all $m$ parameters.
+Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.
 Moreover, a single prediction can be a complicated function of all free parameters.
-In addition we often deal with large number of observations, $p \sim 100-1000$ as well as sizable parameter spaces $n \sim 10$.
-The results of a fit are generally interpreted using visualisations selected based on physics intuition and previous knowledge, but the above stated observations suggest that some correlations clearly dictated by the data may be missed, as we are typically only studying selected views of the results.
+In addition, we often deal with large number of observations, $p \sim 100-1000$ as well as sizable parameter spaces $m \sim 10$.
+The results of a fit are generally interpreted using combinations of variables selected by intuition and prior knowledge.
+But this begs the question, whether important associations are missed because tools to search are not available.
+
+<!--
+suggest that some correlations clearly dictated by the data may be missed, as we are typically only studying selected views of the results.
 The idea of finding special views in scatter plots fits naturally with the concept of using posterior samples obtained in a Bayesian fit.
+
 The idea is that by finding special views of the posterior samples we gain insight into the physics relevant to accurately describe the experimental data.
@@
 There are many examples of this, for example in particle physics detailed studies have been performed e.g.
-in the alignment limit of the 2HDM or in co-annihilation regions in models of dark matter.
+in the alignment limit of the 2HDM or in co-annihilation regions in models of dark matter.-->
 
+[REF] (Katrin's thesis) explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
+In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
+These indices are all available in R packages, scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+The projection pursuit guided tour is available in the R package, tourr [@tourr].
+
+<!-- 
 A related question is the selection of pair-wise scatter plots in datasets with large numbers of parameters, for which different methods have been developed.
@@
 For the measures considered here this generally does not apply, as a consequence we "disentangle" the two into subsequent steps.)
+-->
 
-The paper investigates the nature of the above defined measures, particularly when considered as candidates for projection pursuit indices (Section \ref{sec:investigate})).
-The new guided tour methods are applied to posterior distributions from physics examples (Section \ref{sec:phys}).
+The paper investigates the behaviour of these newly-defined indexes.
+Section \ref{sec:construct} discusses index construction, and how they fit into the guided tour.
+Section \ref{sec:investigate} investigates the behaviour of the indexes, particularly in relation to optimisation.
+The new guided tour methods are applied to an example where the posterior distributions from physics models (Section \ref{sec:phys}) are explored.
 
@@
 
-<!--
-Definition of pp index
-Optimisation
-Illustration of what is already available in tourr
-New index explanations
--->
+Explain here how to construct the pp index
+
+- Definition of pp index
+- Optimisation
+- Illustration of what is already available in tourr
+- New index explanations
 
@@
 
-<!--
+Organisation of this section:
+
 Criteria for a good index
@@
   - optimisation path
--->
 
@@
 
-We consider three example distributions in n=6 dimensions with p randomly selected points $x_i, i = 1,..,n$:
+We consider three example distributions in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$:
 
@@
 2.
-p points where n-1 points are independently drawn from a uniform distribution between [-1,1], and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution
+p points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution
 3.
-p points where n-1 points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function
+$p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function
 
@@
 
+<!-- I think you need to standardise the scale of the data -->
+
 ```{r datasets}
-sphere100 <- sphereData(6, 100)
-sphere1000 <- sphereData(6, 1000)
+sphere100 <- sphereData(6, 100) %>% rescale() %>% as_tibble()
+sphere1000 <- sphereData(6, 1000) %>% rescale() %>% as_tibble()
 
-pipe100 <- pipeData(6, 100)
-pipe1000 <- pipeData(6,1000)
+pipe100 <- pipeData(6, 100) %>% rescale() %>% as_tibble()
+pipe1000 <- pipeData(6,1000) %>% rescale() %>% as_tibble()
 
-sin100 <- sinData(6, 100)
-sin1000 <- sinData(6, 1000)
+sin100 <- sinData(6, 100) %>% rescale() %>% as_tibble()
+sin1000 <- sinData(6, 1000) %>% rescale() %>% as_tibble()
 ```
@@
 }
-knitr::kable(indexOverview ,  caption = "Summary of index values.", digits=2) 
+knitr::kable(indexOverview,  caption = "Summary of index values.", digits=2) 
 ```
@@
 <!-- index values shouldn't be affected by sample size, on average it should be
-same value. Maybe some need to be normalised by n
+same value. Maybe some need to be normalised by n. Something needs fixing here.
 -->
@@
   rbind(getProj(pipe1000, t1full, "pipe", 1000))
-
 ```
@@
 ## Guided tour
+
 Given the behaviour of the various index measures under an interpolated tour path we next try to use the tourr implementation of the guided tour together with selected index functions to uncover the special views in datasets 2 and 3 with 1000 datapoints.
@@
 ### Review of guided tour optimisation
+
 We use the default method "search_geodesic" implemented in the tourr package. (Should I try other optimisation methods, i.e. search_better, search_better_random?)
@@
 ### Looking down the pipe
+
 As seen before (reference figure) MIC or TIC should be able to find the view down the pipe.
@@
 ### Finding sine waves
+
 We next work with distribution 3 and aim to identify the view showing functional dependence using the splines2d index.
@@
 
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide"}
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide", cache=TRUE}
 set.seed(2018)
@@
 
-```{r neutronStarSPLOM}
+```{r neutronStarSPLOM, fig.height=8, fig.width=8}
 nsD <- read_csv("data/samples.csv") 
@@
 ## Using the guided tour
+
 Given the observations from the grand tour display we expect a functional relation between some of the parameters, and we therefore optimise the "splines2d" index function.
@@
 Show final view found from the guided tour, give corresponding projection matrix, make corresponding plot from combination of original parameters.
+
 ```{r neutronStarEq,  results="hide"}
@@
 dProj <- as.tibble(nsM %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point()
+ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
 ```
@@
 ```{r nsePlotOrig}
-p1 <- ggplot(nsD, aes(R1, L1)) + geom_point(alpha = 0.05)
-p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05)
+p1 <- ggplot(nsD, aes(R1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
+p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
 grid.arrange(p1, p2, ncol=2) 
@@
 ## Higher D
+
 To complicate things can work with bigger dataset that Rory sent me, need to check if this is public, also need help with interpretation, but here optimising TIC finds interesting projection
+
+# Applications to collections of time series
+
+This could be another example.
+
+```{r eval=FALSE}
+music <- read_csv("data/tigs_music.csv")
+music <- apply(music, 2, function(x) (x-mean(x))/sd(x))
+musicTour <- save_history(music,
+                         guided_tour(mineIndex("TIC"), 
+                            search_f = tourr:::search_better,
+                            cooling=1, alpha = 0.5, max.tries = 5000))
+musicTour <- save_history(music,
+                         guided_tour(dcorIndex(), 
+                            search_f = tourr:::search_better,
+                            cooling=1, alpha = 0.5, max.tries = 5000))
+musicTourFull <- as.list(interpolate(musicTour))
+iLast <- length(musicTourFull)
+fProj <- musicTourFull[[iLast]]
+dProj <- as.tibble(as.matrix(music) %*% fProj)
+ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+fullRes <- getProj(as.matrix(music), musicTourFull, "TIC", 100)
+fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
+ggplot(fullResMelt, aes(x=t, y=value)) +
+  geom_line(aes(color=factor(size))) + 
+  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
+  guides(color=FALSE)
+```
 

### 2018-09-28 scagnostics.Rmd
organisation of topics

@@
 ---
-title: An investigation into using scagnostics interesting projections of high-dimensional data
+title: An investigation into using scagnostics to find interesting projections of high-dimensional data
 
@@
 
-```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+```{r initial, echo = FALSE, cache = TRUE, include = FALSE}
 library(knitr)
@@
 
-A large number of projection pursuit indices have been developed [@f87] [@CBC92] [@Lee05] [@AHC02] [@CEM:CEM2568] [@JS87] [@5508437].
-Many of these are developed to detect departure from multivariate normal, which includes clusters or outliers.
-Some have been developed to detect separations between known groups [REF].
+A large number of projection pursuit indices have been developed, to detect departure from multivariate normal, which includes clusters or outliers or separations between known groups (e.g.
+@f87, @CBC92, @Lee05, @AHC02, @CEM:CEM2568, @JS87, @5508437).
 Less work has been done on indexes to find nonlinear dependence between variables.
@@
 Criteria for a good index
+
   - smoothness
@@
   - fine and broad structure detection
+
 Simulation study
+
   - example data 
   - sample size, and other factors used
+
 Overview procedures
+
   - look at index values of pairs of variables
@@
 We select two samples for each distribution, with 100 and 1000 sampled points respectively.
+
+<!-- 
+Table and figure captions need to have three components: (1) what is the plot about, (2) specific details of plot, like what type of display and how variables are mapped, (3) the most important thing that the reader should learn.
+
+Be sure to cite R, as well as packages used in the work.
+-->
 

### 2018-10-01 scagnostics.Rmd
tweaks
References added, but something is broken, caching needs to be fixed first.

@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = FALSE, external = TRUE, dev = "pdf",
+  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
   fig.height = 5, fig.width = 8, out.width = "\\textwidth"
@@
 - How it can be applied
--->
 
@@
 - section needs shortening
+-->
 
@@
 
-[REF] (Katrin's thesis) explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
-In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
-These indices are all available in R packages, scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+[@Grimm2018] explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
+In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see @nason1992, @NIPS1997_1408.} These indices are all available in R packages, scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
 The projection pursuit guided tour is available in the R package, tourr [@tourr].
@@
 same value. Maybe some need to be normalised by n. Something needs fixing here.
+
+can we get "theoretical" min/max values, to assess the scale of the index for any particular data set.
 -->

### 2018-10-02 scagnostics.Rmd
Turned off caching, instead manually save results with long calculation time.

--- 2018-09-28 scagnostics.Rmd
@@
 
-```{r initial, echo = FALSE, cache = TRUE, include = FALSE}
+```{r initial, echo = FALSE, include = FALSE}
 library(knitr)
@@
 
-[REF] (Katrin's thesis) explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
+[@Grimm2018] explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
-These indices are all available in R packages, scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
-The projection pursuit guided tour is available in the R package, tourr [@tourr].
+\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see \cite{nason1992}, \cite{NIPS1997_1408}.}
+These indices are all available in R packages, scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva]. The projection pursuit guided tour is available in the R package, tourr [@tourr].
 
@@
 ```{r datasets}
+set.seed(1984)
 sphere100 <- sphereData(6, 100) %>% rescale() %>% as_tibble()
@@
 set.seed(1988)
-t1 <- save_history(sphere100, grand_tour(2), max=4)
-t1full <- as.list(interpolate(t1))
-fullRes <- getProj(sin100, t1full, "sin", 100) %>%
-  rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
-  rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
-  rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
-  rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
-  rbind(getProj(pipe1000, t1full, "pipe", 1000))
+if(!file.exists("cache/grandtourpath.rda")){
+  grandTourHist <- save_history(sphere100, grand_tour(2), max=4)
+  grandTourInt <- as.list(interpolate(grandTourHist))
+  fullRes <- getProj(sin100, grandTourInt, "sin", 100) %>%
+    rbind(getProj(sin1000, grandTourInt, "sin", 1000)) %>%
+    rbind(getProj(sphere100, grandTourInt, "sphere", 100)) %>%
+    rbind(getProj(sphere1000, grandTourInt, "sphere", 1000)) %>%
+    rbind(getProj(pipe100, grandTourInt, "pipe", 100)) %>%
+    rbind(getProj(pipe1000, grandTourInt, "pipe", 1000))
+  save(grandTourHist, grandTourInt, fullRes, file = "cache/grandtourpath.rda")
+} else {
+  load("cache/grandtourpath.rda")
+}
 ```
@@
 ```{r t58matrix, echo=FALSE}
-knitr::kable(t1full[[58]] ,  caption = "Matrix representation of projection 58.")
+knitr::kable(grandTourInt[[58]] ,  caption = "Matrix representation of projection 58.")
 ```
@@
 ```{r plot58, fig.width=6, fig.height=7, fig.cap="View of all datasets defined by at tour step 58."}
-p58 <- t1full[[58]]
+p58 <- grandTourInt[[58]]
 pL <- list()
@@
 ```{r plannedtourpath}
+if(!file.exists("cache/plannedtourpath.rda")){
   m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
@@
     rbind(getProj(pipe1000, t2full, "pipe", 1000))
+  save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
+} else {
+  load("cache/plannedtourpath.rda")
+}
 ```
 
-```{r plannedgrandtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+```{r plannedtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
 fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
@@
 ```{r findpipe, results="hide"}
-set.seed(2018)
-pipeResc <- rescale(pipe1000) #unselect 2 variables to make optimisation easier
+if(!file.exists("cache/findpipe.rda")){
+  set.seed(1995)
+  pipeResc <- rescale(pipe1000) 
   pipeTour <- save_history(pipeResc,
@@
   pipeTourFull <- as.list(interpolate(pipeTour))
+  save(pipeResc, pipeTour, pipeTourFull, file = "cache/findpipe.rda")
+} else {
+  load("cache/findpipe.rda")
+}
 ```
@@
 
-```{r refine, results="hide"}
+```{r refinepipe, results="hide"}
+if(!file.exists("cache/refinepipe.rda")){
   pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
@@
   pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
+  save(pipeTour2, pipeTourFull2, pipeTourRes2, pipeTourResMelt2, file = "cache/refinepipe.rda")
+} else {
+  load("cache/refinepipe.rda")
+}
 ggplot(pipeTourResMelt2, aes(x=t, y=value)) +
@@
 
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide", cache=TRUE}
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide"}
 set.seed(2018)
-sineResc <- rescale(sin1000) #unselect 2 variables to make optimisation easier
+if(!file.exists("cache/findsine.rda")){
+  sineResc <- rescale(sin1000)
   sineTour <- save_history(sineResc,
@@
   sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
+  save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
+} else {
+  load("cache/findsine.rda")
+}
 ggplot(sineTourResMelt, aes(x=t, y=value)) +
@@
   rescale() # rescaling, this returns a matrix
+if(!file.exists("cache/neutronStarEq.rda")){
   nseTour <- save_history(nsM, guided_tour(splineIndex()))
   nseTourFull <- as.list(interpolate(nseTour))
+  save(nseTour, nseTourFull, file = "cache/neutronStarEq.rda")
+} else {
+  load("cache/neutronStarEq.rda")
+}
 iLast <- length(nseTourFull)

### 2018-10-02 scagnostics.Rmd
More manual caching, now reasonably fast to compile.

@@
 ```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Index value on tour path when scouting for high TIC index values."}
+if(!file.exists("cache/pipeFirstRun.rda")){
   pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
   pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
+  save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
+} else {
+  load("cache/pipeFirstRun.rda")
+}
 ggplot(pipeTourResMelt, aes(x=t, y=value)) +

### 2018-10-02 scagnostics.Rmd
More text.

@@
 A large number of projection pursuit indices have been developed, to detect departure from multivariate normal, which includes clusters or outliers or separations between known groups (e.g.
-@f87, @CBC92, @Lee05, @AHC02, @CEM:CEM2568, @JS87, @5508437).
+@f87, @CBC92, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437).
 Less work has been done on indexes to find nonlinear dependence between variables.
@@
 
-[@Grimm2018] explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
+[@Grimm2016] explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
@@
 
+<!--
 Explain here how to construct the pp index
@@
 - New index explanations
+-->
+A projection pursuit index (ppi) is a scalar function $f$ defined on an $n$-dimensional data distribution.
+Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $n$ dimensional projections of a data set with $p>n$ parameters will find the most interesting projections.
+The standard approach is then to map the empirical data distribution onto a density estimate for which the departure from a normal distribution can be characterised in various ways, see e.g.
+@CBC93.
+When departing from the idea of only characterising differences to the normal distribution it may however be preferred to work directly with the empirical data distribution to better capture the true distribution without bias from fitting a densitiy function.
+
+## Optimization
+Given a ppi we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
+An important consideration here is the necessity to avoid getting trapped in local maxima that are only a result of sampling fluctuatins \footnote{This is not only relevant for the choice of optimisation method but also an important consideration when designing an efficient ppi.}, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
+A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimisation for the identification of the maximum.
+On the other hand the global maximum may not be the only projection of interest, and one may want to observe the selected views in the context of the full distribution rather than a static view.
+These issues are addressed when combining projection pursuit with the grand tour [@CBCH94].
+In this case the properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshoothing and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
+A possible alogrithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
+In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
+
+## New index functions
+
+Focusing now on $n=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
+We therefore draw on index functions that have been developped for the selection of variable pairs that show interesting features in a scatter plot, but we generalise the notion such that instead of variable pairs we consider any $2$ dimensional projection of the multivariate data distribution.
+It means that we consider the projection axes as input to the index functions, which will thus score the interestingness of the given projection.
+Concretely here we consider the following types of index functions:
+
+* Scagnostics
+    + Convex
+    + Skinny
+    + Stringy
+* New index function from @Grimm2016
+    + Distance correlation
+    + Spline based measure
+* Information based index functions from @Reshef1518
+    + Maximum Information Coefficient (MIC)
+    + Total Information Coefficient (TIC)
 

### 2018-10-02 scagnostics.Rmd
Section on data standardization?

@@
 
+## Standardization
+scaling/sphering of data?
+what about outlier removal?
+it is build in for calculation of scagnostics measures, but not the other index functions, can generate "fake linear behaviour" by highlighting direction with outliers
+consider using log transformed parameter as appropriate? e.g. timeseries examples...
+
 ## Optimization
@@
 It means that we consider the projection axes as input to the index functions, which will thus score the interestingness of the given projection.
-Concretely here we consider the following types of index functions:
+What we are optimising over are then the entries of the $p\times2$ dimensional projection matrix, considering that they obey orthonormality conditions.
+Concretely here we consider the following types of index functions (give definitions in Appendix??):
 

### 2018-10-03 scagnostics.Rmd
Index definitions added.

--- 2018-10-02 scagnostics.Rmd
@@
   - \usepackage{longtable}
+  - \usepackage{amssymb}
 ---
@@
 Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $n$ dimensional projections of a data set with $p>n$ parameters will find the most interesting projections.
-The standard approach is then to map the empirical data distribution onto a density estimate for which the departure from a normal distribution can be characterised in various ways, see e.g.
+As most indices characterise deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations may be defined in different ways, see e.g.
 @CBC93.
-When departing from the idea of only characterising differences to the normal distribution it may however be preferred to work directly with the empirical data distribution to better capture the true distribution without bias from fitting a densitiy function.
+When departing from the idea of characterising differences to the normal distribution it may however be preferred to work directly with the empirical data distribution to better capture the true distribution, and instead characterise the distribution by measures based on e.g.
+the minimal spanning tree or the mutual information.
+
+## Standardization
+Lower dimensional projections are highly sensitive to standardization performed on the original distribution.
+We want to avoid e.g.
+highlighting directions based only on different scales, or artifically extending the parameter range by including outliers, which may result in seemingly linear behaviour for most points.
+Methods to consider include rescaling of individual directions to a common interval (e.g.
+to fall between $[0,1]$), sphering, outlier removal or transformation to a logarithmic scale.
+The method(s) of choice should be informed by the distributions found in the data set as well as aims of the analysis.
+((particular challenges: noise only directions, identification of outliers?))
 
@@
 Given a ppi we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
-An important consideration here is the necessity to avoid getting trapped in local maxima that are only a result of sampling fluctuatins \footnote{This is not only relevant for the choice of optimisation method but also an important consideration when designing an efficient ppi.}, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
+An important challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
 A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimisation for the identification of the maximum.
@@
 ## New index functions
-
 Focusing now on $n=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
@@
 It means that we consider the projection axes as input to the index functions, which will thus score the interestingness of the given projection.
+What we are optimising over are then the entries of the $p\times2$ dimensional projection matrix, considering that they obey orthonormality conditions.
 Concretely here we consider the following types of index functions:
 
-* Scagnostics
-    + Convex
-    + Skinny
-    + Stringy
-* New index function from @Grimm2016
-    + Distance correlation
-    + Spline based measure
-* Information based index functions from @Reshef1518
-    + Maximum Information Coefficient (MIC)
-    + Total Information Coefficient (TIC)
+* From the scagnostics family select those best suited to detect shapes:
+    + Convex: The convex measure is defined as $c_{convex}= \frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
+This is the only measure where interesting projections will take low values and one may either minimize the index function or maximize $1-c_{convex}$.
+    + Skinny: The skinnyness of a polygon can be measured as the ratio of the perimeter to the area.
+We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a full circle, and values close to one or skinny polygons.
+    + Stringy: The stringy index is a measure of the branching structure of the minimal spanning tree (MST), $c_{stringy} = \frac{diameter(MST)}{length(MST)}$ where the diameter is the longest connected path, and the length is the total length (sum of all edges), i.e.
+$c_{stringy}=1$ if the MST contains no branches.
+
+* New index functions available in @mbgraphic:
+    + Distance correlation defined in @szekely2007: The definition is based on empiric measures of distance covariance and variance defined on the dataset 
+    \begin{equation}
+(X,Y) = \{(x_1,y_1), ..., (x_n, y_n) \in \mathbb{R}^p \times \mathbb{R}^q\}
+\end{equation}
+ as
+ \begin{equation}
+ Cov^2(X,Y) = \frac{1}{n^2} \sum_{k,l=1}^{n} A_{kl}B_{kl},\\
+ Var^2(X) = \frac{1}{n^2} \sum_{k,l=1}^{n} A_{kl}^2
+ \end{equation}
+ where $A_{kl}=a_{kl}-\bar{a}_{k.}-\bar{a}_{l.}+\bar{a}_{..}$ with
+ \begin{equation}
+ a_{kl} = | x_k - x_l|_p,\\
+ \bar{a}_{k.} = \frac{1}{n} \sum_{l=1}^n a_{kl},\\
+  \bar{a}_{.l} = \frac{1}{n} \sum_{k=1}^n a_{kl},\\
+ \bar{a}_{..} = \frac{1}{n^2}\sum_{k,l=1}^n a_{kl}
+ \end{equation}
+ and reads
+ \begin{equation}
+ dCor(X,Y) = \begin{cases}
+ \frac{Cov(X,Y)}{\sqrt{Var(X)Var(Y)}} & Var(X)Var(Y) > 0\\
+ 0 & Var(X)Var(Y) = 0
+ \end{cases}
+ \end{equation}
+    + Spline based measure: Given a spline model of the data distribution that is obtained via the R implentation in the \texttt{gam} function in the mgcv R package [@w16], we define the index based on the variance of the residuals:
+    \begin{equation}
+    splines_{2d} = max(1- \frac{Var(res_{X\sim Y})}{Var(X)}, 1-\frac{Var(res_{Y\sim X})}{Var(Y)}),
+    \end{equation}
+    which takes large values if the distribution is well described by the spline model, indicating functional dependence.
+
+* Information based index functions from @Reshef1518 are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.
+With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as
+\begin{equation}
+M(D)_{x,y} = \frac{I^{\ast} (D, x, y)}{\log \min{x,y}}.
+\end{equation}
+We can now define the following metrics:
+    + Maximum Information Coefficient (MIC):
+    \begin{equation}
+    MIC(D) = \max_{xy<B(n)}\{M(D)_{x,y}\},
+    \end{equation}
+    where by default $B(n) = n^{0.6}$
+    + Total Information Coefficient (TIC), see @JMLRv1715308:
+    \begin{equation}
+    TIC(D) = \sum_{xy<B(n)} M(D)_{x,y}
+    \end{equation}
+  While $MIC$ is normalised to fall between $[0,1]$, the upper bound of $TIC$ depends on the sample size via the bounding function $B(n)$.
+$TIC$ was designed to be more stable test against null hypothesis of independence, using the full information of the characteristic matrix rather than just the maximum, and as such is expected to be better suited as a ppi.
 

### 2018-10-03 scagnostics.Rmd
Small changes only.

@@
 \footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see \cite{nason1992}, \cite{NIPS1997_1408}.}
-These indices are all available in R packages, scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva]. The projection pursuit guided tour is available in the R package, tourr [@tourr].
+These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+The projection pursuit guided tour is available in the R package, tourr [@tourr].
 
@@
   - optimisation path
+
+In general we require that a good ppi should be smooth over geodesically interpolated projections, and be fast to calculate for dynamic viewing of the optimization with the guided tour.
+Note that there is a trade-off between smoothness requirements and the efficiency of the optization, and index functions sensitive to random small fluctuations may be viable in conjunction with an approriate optimization strategy, as noted above.
+Moreover, larger calculation times may be acceptable when we do not require observing the optimization in real-time, but instead record the guided tour which can be replayed once the optimization has converged.
 

### 2018-10-03 scagnostics.Rmd
Starting work on Sec 3.

@@
 ## New index functions
+\label{sec:indexDef}
 Focusing now on $n=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
@@
 \label{sec:investigate}
+We now examine the indices listed in Section \ref{sec:indexDef} to test which of them are suitable as projection pursuit measures, and how we may modify them for improved performance.
+We start by considering the relevant criteria:
 
-Organisation of this section:
+  - smoothness: the index function should evolve smoothly between nearby projections, i.e.
+on an interpolated path we expect to observe slow and consistent change in the index function.
+On the other hand if two similar projections produce very different index values, i.e.
+we observe sharp jumps when looking at the evolution of the index function, and in particular if those fluctuations dominate the overall evolution, the index is not suitable for projection pursuit.
+  - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
+Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified.
+  - ability to find fine and broad structure: We consider a broad structure one that is roughly observable over a large angle around the optimal projection, it should in general be easy to identify and a good ppi should show smooth increase over a large number of projections as we are approaching the optimal view.
+On the other hand fine structures are difficult to identify as they are only visible close to the optimal view.
+This poses problems in particular for the optimization, but we may use index definitions to amplify larger index values while suppressing smaller ones, such that the optimization will be less sensitive to noise.
+  - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
 
-Criteria for a good index
+##Simulation study
+To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
+They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
 
-  - smoothness
-  - speed
-  - fine and broad structure detection
+1. $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
+2.
+$p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+3.
+$p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
 
-Simulation study
+For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
 
-  - example data 
-  - sample size, and other factors used
+To understand the potential for the various indices defined we use the following overview procedures on the simulated data sets:
 
-Overview procedures
+- index values for pair of variables, explicitly comparing the uninformative and special view of each dataset, as well as discussing the dependence on the sample size
+- tour between pairs of variables, considering a planned tour that moves from the $x_1 - x_2$ view either to the $x_3 - x_4$ view (no special structure encountered over the full path) or to the $x_5 - x_6$ (moving from uninformed to special view)
+- optimisation path, starting from $x_1 - x_2$, how can we use the available index functions and optimisation algorithms to identify the special view
 
-  - look at index values of pairs of variables
-  - tour betwen pairs of variables
-  - optimisation path
-
-In general we require that a good ppi should be smooth over geodesically interpolated projections, and be fast to calculate for dynamic viewing of the optimization with the guided tour.
-Note that there is a trade-off between smoothness requirements and the efficiency of the optization, and index functions sensitive to random small fluctuations may be viable in conjunction with an approriate optimization strategy, as noted above.
-Moreover, larger calculation times may be acceptable when we do not require observing the optimization in real-time, but instead record the guided tour which can be replayed once the optimization has converged.
+###Dataset overview
 
@@
 ```
-
-((REVISE - where should measure definitions go? now basically in introduction.))
-The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics.
-If there are 100 variables, the dimension will be $4950\times 9$.
-It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting.
-It can also be used to do more in-depth exploratory analysis, and is interesting to study the distribution of the scagnostics resulting from the data, as a data set in itself.
-In a sense this the data's "fingerprint".
-
-((maybe better to start with something like this, though maybe too obvious?))
-The various measures defined in the introduction are designed to be computed for all pairs of variables, followed either by selection based on comparing scagnostics scores between all pairs, or by an absolute ranking from most to least interesting (e.g.
-when using TIC).
-Here we modify the usage as follows.
-The tour implementation is used to generate projections of the multivariate dataset down to two dimensions, i.e.
-the projected data can be represented as a scatter plot.
-In this projection the data is represented by two parameters that are defined as combinations of the input parameters, and this parametrisation is passed into the index function calculation to obtain a score for the selected view.
-Using the guided tour implementation the index functions are optimized such that we find the generalised 2-d parametrisation yielding the most interesting view of the data.
-
-To use an index function for projection pursuit requires that the measure be relatively smooth over the space of all 2D projections, and we will first investigate the smoothness over randomly selected projections.
-An additional consideration is the computing time, ideally fast computation is preferred as the guided tour can be viewed as the index is being optimized.
-However, a computationally more costly index may be preferred for better detection properties, and one can consider recording the guided tour for viewing after the optimisation has converged.
-Finally we note that another important consideration is how the index function changes as the tour moves into the "optimal" view, this will however depend mainly on the distribution rather than the selected index function.
-(we will see this in the examples)
-
-## Dataset
-
-We consider three example distributions in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$:
-
-1. p points on an n-dimensional sphere randomly selected using the geozoo package, as an example of dataset without special views
-2.
-p points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution
-3.
-$p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function
 
@@
 The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of V6 (and V5 for the Pipe distribution) will be distinct from uninformed views of distributions 2 and 3.
-We select two samples for each distribution, with 100 and 1000 sampled points respectively.
 

### 2018-10-04 scagnostics.Rmd
Captions and text updated.

@@
 
+((What is upper bound of TIC as function of sample size?
+If we define $n' = n^{0.6}$, rounded to integer values only, can we calculate it as $TIC_{max} = 2 \sum_{i=1}^{n'} \frac{n'}{i}$, keeping in mind that each entry of the characteristic matrix should be between [0,1].
+Each entry in the sum should be rounded also, only integer number of bins enter actual calculation, so this should be approximation only.))
+
 # Investigation of indices
@@
 
-Be sure to cite R, as well as packages used in the work.
 -->
 
-```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Data views of V1 vs V2."}
+```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters V1 and V2, i.e. showing a typical uninformed view for each dataset."}
 pL <- list()
@@
 
-```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Data views of V5 vs V6."}
+```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters V5 and V6, i.e.
+showing the special view built into the Pipe and Sine datasets."}
 pL <- list()
@@
 
-```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="Data views of V1 vs V6."}
+```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
+Apart from statistical fluctuations we see that for the Pipe datasets the parameters V5 and V6 are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine dataset the parameter V6 follows a very distinct distribution.
+Note also that the first five paramters of the Sine dataset appear to be drawn from different normal distributions, which is a result of the rescaling, shifting the modes to different values for each direction."}
 pL <- list()
@@
 indexOverview <- tibble(convex=numeric(),
-               clumpy=numeric(),
                skinny=numeric(),
@@
     indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
-                    clumpy=scagRes[,"Clumpy"], skinny=scagRes[,"Skinny"],
-                    stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
-                    splines2d=splineRes, MIC=mineRes$MIC,
-                    TIC=mineRes$TIC, name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
+                    skinny=scagRes[,"Skinny"], stringy=scagRes[,"Stringy"],
+                    dcor2d=dcorRes, splines2d=splineRes,
+                    MIC=mineRes$MIC, TIC=mineRes$TIC,
+                    name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
   }
@@
 ## Smoothness
-We first study the smoothness of a subset of measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path between 4 anchor planes and using the default interpolation angle of 0.05.
-For this comparison we rescale the data such that each parameter takes values between [0,1].
-The temporal trace of a selected set of measures is shown below, where each time step refers to an interpolation step.
-Note that the value of TIC is strongly dependent on the input distribution and we have therefore normalised the TIC measure by its maximum for each individual dataset, all other measures are reproduced as reported by the calculation.
+We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto V1-V2 to the projection onto V3-V4 and using the default interpolation angle of 0.05.
+Each time step refers to an interpolation step.
+Since the value of the TIC index is strongly dependent on the input distribution sample size we have normalised the TIC measure by its maximum for each individual dataset, all other measures are reproduced as reported by the calculation.
+The results are shown in Figure \ref{fig:plotV1V2toV3V4}.
+The standard index functions "holes" and "cmass" show the desired behaviour in that the change is smooth over interpolated views ((it is however curious to see the actual evolution, which seems not to depend on the underlying random distributions, but is clearely evolving over the path of the planned tour)).
+Considering now the indices from the scagnostics family, we first note stronge dependence on the sample size.
+The convex index is largely sensitive to random fluctuations when considering small samples (p=100), while being mostly flat for larger samples, especially in the case of an underlying uniform distribution.
+The skinny index on the other hand is systematically larger for the smaller samples, while stringy appears not to be sensitive to the sample size, but resulting in large fluctuations and sharp jumps for all datasets.
+Distance correlation and the spline based index show similar behaviour, i.e.
+when considering large enough data samples (p=1000) the values are consistently close to zero, as expected for a distribution without structure.
+On the other hand somewhat larger values (but still far from the maximum value of 1) are found when considering small datasets, where distributions are however mostly smooth, with few kinks depending on index function and underlying distribution.
+Finally the information based metrics also show systematic differences based on the sample size, with sampling fluctuations being significant for small datasets.
+Note here that fluctuations present in Figure \ref{fig:plotV1V2toV3V4} may be amplified, as maximum values for most index functions are far from the theoretic maximum.
 
@@
     convex=numeric(),
-    clumpy=numeric(),
     skinny=numeric(),
@@
     sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
-                  convex=scagRes[,"Convex"],
-                  clumpy=scagRes[,"Clumpy"], skinny=scagRes[,"Skinny"],
+                  convex=scagRes[,"Convex"], skinny=scagRes[,"Skinny"],
                   stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
@@
 
-```{r grandtourpath}
-set.seed(1988)
-if(!file.exists("cache/grandtourpath.rda")){
-  grandTourHist <- save_history(sphere100, grand_tour(2), max=4)
-  grandTourInt <- as.list(interpolate(grandTourHist))
-  fullRes <- getProj(sin100, grandTourInt, "sin", 100) %>%
-    rbind(getProj(sin1000, grandTourInt, "sin", 1000)) %>%
-    rbind(getProj(sphere100, grandTourInt, "sphere", 100)) %>%
-    rbind(getProj(sphere1000, grandTourInt, "sphere", 1000)) %>%
-    rbind(getProj(pipe100, grandTourInt, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, grandTourInt, "pipe", 1000))
-  save(grandTourHist, grandTourInt, fullRes, file = "cache/grandtourpath.rda")
+```{r V1V2toV3V4}
+if(!file.exists("cache/V1V2toV3V4.rda")){
+  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
+  m2 <- matrix(c(0,0,1,0,0,0,0,0,0,1,0,0),ncol=2)
+  #this is silly but seems that first two entries are being ignored, so need some fake entries
+  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
+  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
+  t1 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
+  t1full <- as.list(interpolate(t1))
+  fullRes <- getProj(sin100, t1full, "sin", 100) %>%
+    rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
+    rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
+    rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
+    rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
+    rbind(getProj(pipe1000, t1full, "pipe", 1000))
+  save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
 } else {
-  load("cache/grandtourpath.rda")
+  load("cache/V1V2toV3V4.rda")
 }
@@
 
-```{r plotgrandtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto V1-V2 to the projection onto V3-V4, for all considered datasets and index functions.
+Small samples (p=100) are shown in red, large samples (p=1000) in blue."}
 fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
@@
 
-Sharp jumps can be observed in all measures except the distance correlation measure when considering small datasets of only 100 points.
-Generally we observe smoother behaviour for larger datasets, notably all considered scagnostics measures show only small variation around a flat baseline.
-Interestingly all other indices indicate preferred viewing directions for dataset 3 when studying large samples, but not for dataset 2 which contains a special view in the same direction.
-We therefore show 2-d view defined by the tour step $t=58$, with projection matrix
+## Transition
+We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6, which will also allow us to view the picture between typical low and high values of the index function.
+Results are shown in Figure \ref{fig:plannedtour}.
+We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise.
+On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
+Note that the various indices appear less noisy than for the grand tour shown before, but should note that the tour path here is shorter, and the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
+(What does this mean for optimisation? How can I detect such "smaller scale increase"? Maybe try "search_better" instead of search_geodesic? increase alpha and decrease cooling? increase max.tries?)
+Can force it when using search_better with large alpha, cooling=1, max.tries=1000, but not ideal..
+Difficulties also from "fake" structure from the sampling boundaries?
+Possible issue with search_geodesic is that the search window is fixed in the code..
 
-```{r t58matrix, echo=FALSE}
-knitr::kable(grandTourInt[[58]] ,  caption = "Matrix representation of projection 58.")
-```
+TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full information of the characteristics is used, implying that more information is considered.
+We also see that TIC is smoother when moving into the special view.
+(As I understand TIC is however not equitable, i.e. it will prefer linear relations, and it may be instructive to optimise MIC and TIC)
 
-i.e. while the last two parameters enter significantly, they will not dominate the view and we do not expect to fully observe the built in features.
+FIXME need to check the MIC_e and TIC_e, supposedly faster computation and better discrimination!
 
-```{r plot58, fig.width=6, fig.height=7, fig.cap="View of all datasets defined by at tour step 58."}
-p58 <- grandTourInt[[58]]
-pL <- list()
-i <- 1
-pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
-  dm <- rescale(ds)
-  pData <- as.tibble(dm %*% p58)
-  pC <- ggplot(pData, aes(V1, V2)) +
-    geom_point() +
-    ggtitle(pLabels[i])
-  pL[[i]] <- pC
-  i <- i+1
-}
-grid.arrange(pL[[1]], pL[[2]],
-             pL[[3]], pL[[4]],
-             pL[[5]], pL[[6]],
-             ncol=2)
+Here we also see how the existing indices in the tourr package are inadequate for the detection of these type of dependence, the holes index taking maximum values before the final view and decreasing when moving into the special view.
+On the other hand, surprisingly the cmass index increases when moving into the special view, but the intermediate minimum makes it unsuitable as a projection pursuit index here.
+QUESTION: how come they are all almost the same for the various distributions?
+Especially pipe and sphere seem to give almost identical behaviour, can it be understood from the index definition?
 
-```
-
-We see from Fig. ((how to add cross references?)) that the built in special views cannot be seen in this projection, but confirm that the sine 1000 dataset presents some correlation in this view.
-
-## Transition
-We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6.
 ```{r plannedtourpath}
@@
 ```
-
-We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise.
-On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
-Note that the various indices appear less noisy than for the grand tour shown before, but should note that the tour path here is shorter, and the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
-(What does this mean for optimisation? How can I detect such "smaller scale increase"? Maybe try "search_better" instead of search_geodesic? increase alpha and decrease cooling? increase max.tries?)
-Can force it when using search_better with large alpha, cooling=1, max.tries=1000, but not ideal..
-Difficulties also from "fake" structure from the sampling boundaries?
-Possible issue with search_geodesic is that the search window is fixed in the code..
-
-TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full information of the characteristics is used, implying that more information is considered.
-We also see that TIC is smoother when moving into the special view.
-(As I understand TIC is however not equitable, i.e. it will prefer linear relations, and it may be instructive to optimise MIC and TIC)
-
-FIXME need to check the MIC_e and TIC_e, supposedly faster computation and better discrimination!
-
-Here we also see how the existing indices in the tourr package are inadequate for the detection of these type of dependence, the holes index taking maximum values before the final view and decreasing when moving into the special view.
-On the other hand, surprisingly the cmass index increases when moving into the special view, but the intermediate minimum makes it unsuitable as a projection pursuit index here.
-QUESTION: how come they are all almost the same for the various distributions?
-Especially pipe and sphere seem to give almost identical behaviour, can it be understood from the index definition?
 

### 2018-10-04 scagnostics.Rmd
Updated pipe guided tour.

@@
 
+Double challengind: fine structure + (coherent) noise
+
 As seen before (reference figure) MIC or TIC should be able to find the view down the pipe.
@@
 
+UPDATE: highlight times corresponding to new basis planes by red vertical line, as we approach the special view higher index values of TIC are often found on interpolated bases, this is why generally we want cooling but here we are trying to find overall maximum, but instead have good chance of finding a plane close to the special view allowing optimisation in a second step
+By no means stable procedure, when changing seed we easily end up not finding special view but rather end up with plane that is not useful for further optimisation to identify pipe view.
+
 ```{r findpipe, results="hide"}
 if(!file.exists("cache/findpipe.rda")){
-  set.seed(1995)
+  set.seed(1984)
   pipeResc <- rescale(pipe1000) 
@@
 }
+newBasisPipeTour <- which(attributes(interpolate(pipeTour))$new_basis)
 ggplot(pipeTourResMelt, aes(x=t, y=value)) +
   geom_line() + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")
+  facet_wrap(~variable, ncol=1, scales = "free_y")+
+  geom_vline(data=as.tibble(newBasisPipeTour), mapping=aes(xintercept=value), color="red")
 ```

### 2018-10-05 scagnostics.Rmd
Considerations on holes index.

@@
 
+((Notes on holes/cmass indices)):
+
+Holes index in $d=2$ dimensional projections, we have projected data given by 
+\begin{equation}
+m = \begin{pmatrix}
+x_1 & y_1 \\
+x_2 & y_2 \\
+\vdots & \vdots \\
+x_n & y_n
+\end{pmatrix}
+\end{equation}
+In tourr calculate holes index as $num/den$ where for $d=2$ we have
+\begin{equation}
+num = 1 - \frac{1}{n} \sum_{i=1}^n e^{(x_i^2+y_i^2)/2}
+\end{equation}
+and $den$ is fixed for fixed number of dimensions, here $den = 1 - \frac{1}{e}$.
+Considering now the case where all $x_i, y_i \leq 1$ we can see that points at the edge may lead to larger values than points on a circle, $x_i^2+y_i^2 = 1$.
+Note also that we cannot give generic upper limit on the index values, it will depend on the range, in particular values $>1$ also possible.
+Cmass calculated as $1-holes$, will take negative values if $holes>1$, e.g.
+the case for entries $x_i, y_i > 1$, example:
+
+```{r load}
+library(tourr)
+library(tidyverse)
+library(reshape2)
+library(scagnostics)
+library(gridExtra)
+library(tictoc) #timer
+library(mbgraphic) #Katrins package
+library(GGally)
+library(geozoo)
+library(minerva) #MINE indices
+```
+
+```{r sillyExample, echo=TRUE}
+
+mX <- matrix(c(2,2,2,2), nrow=2)
+holes(mX)
+cmass(mX)
+# compare random points on circle to extreme positions
+mC <- matrix(c(cos(1.2*pi), sin(1.2*pi), cos(pi/3), sin(pi/3), cos(0.6*pi), sin(0.6*pi)), ncol = 2)
+mT <- matrix(c(1,0,0,1,1,1), ncol=2)
+holes(mC)
+holes(mT)
+```
+
 # Investigation of indices
@@
 ###Dataset overview
-
-```{r load}
-library(tourr)
-library(tidyverse)
-library(reshape2)
-library(scagnostics)
-library(gridExtra)
-library(tictoc) #timer
-library(mbgraphic) #Katrins package
-library(GGally)
-library(geozoo)
-library(minerva) #MINE indices
-```
 

### 2018-10-05 scagnostics.Rmd
Adding density plots for index distribution.

@@
 
+## Distributions of index values
+
+Compare typical index values over the two tour paths. Notation: path1 / 2 labels tour from V1-V2 to V3-V4 / V5-V6.
+
+```{r indexDist, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+fullResMelt1 <- fullResMelt %>% add_column(path="path1")
+fullResMeltPlanned1 <- fullResMeltPlanned %>% add_column(path="path2")
+allRes <- rbind(fullResMelt1, fullResMeltPlanned1)
+ggplot(allRes, aes(x=value)) +
+  geom_density(aes(color=path, linetype = factor(size))) +
+  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
+```
+
 ## Guided tour

### 2018-10-05 scagnostics.Rmd
Small changes throughout.

--- 2018-10-05 scagnostics.Rmd
@@
 * Information based index functions from @Reshef1518 are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.
-With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as
+With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as \footnote{We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.}
 \begin{equation}
@@
 
-```{r plannedtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
+```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \ref{fig:plotV1V2toV3V4} but moving from the projection onto V1-V2 to the projection onto V5-V6, i.e.
+starting from random distribution view, and moving into the special view."}
 fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
@@
   guides(color=FALSE)
+```
+
+## Distributions of index values
+
+Compare typical index values over the two tour paths.
+Notation: path 1 / 2 labels tour from V1-V2 to V3-V4 / V5-V6.
+Typically find broader distributions for small samples because of the large fluctuations in this case.
+Some indices show preferred behaviour, i.e.
+sharp peaks around low values for path 1, and very broad distributions leading up to maximum values for path 2.
+For other indices we can see the strong dependence of the index on sample size, the most extreme example is the skinny index.
+
+No clear interpretation without trace for comparison? See e.g. TIC for path2 pipe vs sine
+
+```{r indexDist, fig.width=6, fig.height=7, fig.cap="Comparing distributions of index values found in the interpolated tour paths 1 and 2 as described in the text."}
+fullResMelt1 <- fullResMelt %>% add_column(path="path1")
+fullResMeltPlanned1 <- fullResMeltPlanned %>% add_column(path="path2")
+allRes <- rbind(fullResMelt1, fullResMeltPlanned1)
+ggplot(allRes, aes(x=value)) +
+  geom_density(aes(color=path, linetype = factor(size))) +
+  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
 ```
@@
 
+There is some issue still with search\_better not actually giving monotonically increasing results, my guess is it is because of rigid assignment of x and y axis, i.e.
+calculating index on starting plane, compare to randomly selected new planes, if new index > starting index and with step size = inf (default when using save history) this means we move to the new plane, but in the interpolation framework, i.e.
+may be rotated wrt new plane found by search\_better, and since now the distribution is the same, but x/y values are not, we get somewhat different values, this may result in somewhat smaller values than anticipated in the new starting plane, which may result in small drops rather than monotonic increase.
+
 ```{r findpipe, results="hide"}
@@
 
-We next show in Fig \ref{fig:testpipe} the final view corresponding to the highest index found in the TIC scouting phase.
-While this is clearly not the optimal view of the pipe feature, we can already guess the type of dependence in the distribution, and we expect to be able to find the maximum when using this plane as a starting point of the geodesic optimisation.
-Indeed performing a standard guided tour with the TIC index and the identified starting plane efficiently identifies the optimal view shown in Fig ...
-((Probably should remove second trace plot as it does not give much information, maybe show final view in same figure as final scouting view...))
-
-```{r testpipe, out.width=".5\\textwidth", fig.cap="Final view discovered by the scouting phase."}
-iLast <- length(pipeTourFull)
-fProj <- pipeTourFull[[iLast]]
-dProj <- as.tibble(pipeResc %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point()
-```
+We next show in Fig \ref{fig:testpipe} (left) the final view corresponding to the highest index found in the TIC scouting phase.
+While this is clearly not the optimal view of the pipe feature, we can already identify the feature in the distribution, and we expect to be able to find the maximum when using this plane as a starting point of the geodesic optimisation.
+Indeed performing a standard guided tour with the TIC index and the identified starting plane efficiently identifies the optimal view shown in the right plot.
+In this case the optimisation quickly converges, following a smooth, monotonic increas of the TIC index.
+((not shown because that is already full informtion basically))
 
@@
 }
-ggplot(pipeTourResMelt2, aes(x=t, y=value)) +
-  geom_line() + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")
-
 ```
 
-```{r testpipe2, out.width=".5\\textwidth"}
+```{r testpipe,  fig.height=4, fig.cap="Final view discovered by the scouting phase (left) and after second optimisation (right)."}
+iLast <- length(pipeTourFull)
+fProj <- pipeTourFull[[iLast]]
+dProj <- as.tibble(pipeResc %*% fProj)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point()
 iLast <- length(pipeTourFull2)
@@
 dProj <- as.tibble(pipeResc %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point()
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point()
+grid.arrange(p1, p2, ncol=2)
 ```

### 2018-10-05 scagnostics.Rmd
Added code for calculation time comparison, results not yet available.

@@
 
+## Note on calculation time
+Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example the 6-d sphere and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
+
+```{r getTimer}
+timeThis <- function(d, t, idx, pmax){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections
+    i <- 1
+    tic.clearlog()
+    tic() #start timer
+    for(pMatrix in t){
+      if(i>pmax) break
+      dProj <- d %*% pMatrix
+      sgnst <- idx(dProj[,1],dProj[,2])
+      i = i+1
+    }
+    toc(log=TRUE,quiet=TRUE)
+    scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
+}
+
+```
+
+```{r timer}
+if(!file.exists("cache/timer.rda")){
+  set.seed(2018)
+  grandTour100 <- save_history(sphere100, grand_tour(2), max=4) %>%
+    interpolate() %>%
+    as.list()
+
+  sizeL <- seq(0,10000,by=100)
+  n <- length(sizeL)
+  dfTimer = data.frame( n=rep(0, n), scagT=rep(0,n), dcorT=rep(0,n), splineT=rep(0,n), mineT=rep(0,n))
+  t <- 1
+  for(sampleSize in sizeL){
+    set.seed(sampleSize + 11) # new seed for each sample size
+    if(sampleSize == 0) sampleSize = 10 # smallest considered sample has 10 points
+    thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
+    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100)
+    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100)
+    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100)
+    mineT <- timeThis(thisMatrix, grandTour100, mine, 100)
+    dfTimer[t,] <- c(sampleSize, scagT, dcorT, splineT, mineT)
+    t <-  t+1
+  }
+  save(dfTimer, file = "cache/timer.rda")
+} else {
+  load("cache/timer.rda")
+}
+
+timerMelt <- melt(dfTimer, id="n")
+ggplot(timerMelt, aes(x = n, y = value, color=variable)) + 
+  geom_line() +
+  geom_line(data=subset(df, n==2),aes(linetype = as.factor(n))) +
+  geom_line(data=subset(df, n==3),aes(linetype = as.factor(n))) +
+  labs(x = "Sample Size", y = "Time [s]", color = "Index family")
+```
+
 ## Guided tour
@@
 Given the behaviour of the various index measures under an interpolated tour path we next try to use the tourr implementation of the guided tour together with selected index functions to uncover the special views in datasets 2 and 3 with 1000 datapoints.
-
-### Review of guided tour optimisation
-
-We use the default method "search_geodesic" implemented in the tourr package. (Should I try other optimisation methods, i.e. search_better, search_better_random?)
-Search geodesic first collects $n=5$ samples by default, amongst which the most promising direction is selected.
-In a second step a linear search along the geodesic is performed in that direction.
-The optimisation is stopped when ???
 
@@
 ```
-
-### Note on calculation time
-Considerations: time to calculate index for single 2-d plot, parameters to turn for guided tour (max.tries, alpha, cooling)
 

### 2018-10-08 scagnostics.Rmd
Changed timing code, added result.

--- 2018-10-05 scagnostics.Rmd
@@
 
+## Note on calculation time
+Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example the 6-d sphere and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
+
+```{r getTimer}
+
+mineE <- function(x,y){
+  return(mine(x=x, y=y, est = "mic_e"))
+}
+
+timeThis <- function(d, t, idx, pmax){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections
+    i <- 1
+    tic.clearlog()
+    tic() #start timer
+    for(pMatrix in t){
+      if(i>pmax) break
+      dProj <- d %*% pMatrix
+      sgnst <- idx(dProj[,1],dProj[,2])
+      i = i+1
+    }
+    toc(log=TRUE,quiet=TRUE)
+    scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
+}
+
+```
+
+```{r timer}
+if(!file.exists("cache/timer.rda")){
+  set.seed(2018)
+  grandTour100 <- save_history(sphere100, grand_tour(2), max=4) %>%
+    interpolate() %>%
+    as.list()
+
+  sizeL <- c(10, 100, 500, 1000, 5000, 10000)
+  n <- length(sizeL)
+  dfTimer = data.frame( n=rep(0, n), scagT=rep(0,n), dcorT=rep(0,n), splineT=rep(0,n), mineT=rep(0,n),
+                        mineeT=rep(0,n))
+  t <- 1
+  for(sampleSize in sizeL){
+    set.seed(sampleSize + 11) # new seed for each sample size
+    if(sampleSize == 0) sampleSize = 10 # smallest considered sample has 10 points
+    thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
+    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100)
+    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100)
+    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100)
+    mineT <- timeThis(thisMatrix, grandTour100, mine, 100)
+    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100)
+    dfTimer[t,] <- c(sampleSize, scagT, dcorT, splineT, mineT,mineeT)
+    t <-  t+1
+  }
+  save(dfTimer, file = "cache/timer.rda")
+} else {
+  load("cache/timer.rda")
+}
+
+timerMelt <- melt(dfTimer, id="n")
+ggplot(timerMelt, aes(x = n, y = value, color=variable)) + 
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  labs(x = "Sample Size", y = "Time [s]", color = "Index family")
+```
+
 ## Guided tour

### 2018-10-08 scagnostics.Rmd
Added second gravitational waves example.

@@
 Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example the 6-d sphere and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
+We see that the evaluation of the spline index shows weak dependence on the sample size, making it the least efficient for very small data sets, but one of the fastest when considering large data sets.
+The situation is similar for calculating the scagnositics measures, which are however faster to compute on datasets of around 10 points.
+This is because the implementation is first binning the data points, which is fast compared to the evaluation that requires triangulation of the binned points.
+The distance correlation based measure exhibits strong increase with sample size, we note however that the implementation also allowes for evaluation on binned data points which we have not considered.
+Finally for the MINE family the computing time increases exponentially as can be expected from the dependence on the sample size, where we see that the improved $MIC_e$ algorithm is somewhat faster than the original implementation, see also @JMLRv1715308.
 
@@
 p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
-grid.arrange(p1, p2, ncol=2) 
-```
-
-This is reflecting the dependence of the deformability parameter $\Lambda$ on the so-called compactness which is proportional to $m/R$.
-QUESTIONS: can see this when plotting e.g.
-R1/m1 vs L1, why also with the linear combination used in the plot (only works if I get the factor about right)?
-not clear how this was built into the simulation, seems not strict equality mentioned in the paper
+p3 <- ggplot(nsD, aes(R1/m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
+grid.arrange(p1, p2, p3, ncol=3) 
+```
+
+This is reflecting the dependence of the deformability parameter $\Lambda$ on the so-called compactness which is approximately proportional to $(m/R)^{-5}$ [@Abbott2018exr].
+QUESTION: what is relation between displaying as function of ratio or linear combination? can I see this from doing taylor expansion?
 
 ## Higher D
-
-To complicate things can work with bigger dataset that Rory sent me, need to check if this is public, also need help with interpretation, but here optimising TIC finds interesting projection
+The above example is using only a reduced set of the free parameters relevant to the physical interpretation.
+In general many additional nuissance parameters will also enter the fit, see e.g.
+@Smith:2016qas.
+We use a simulation study (simulating a binary black hole (BBH) merger event...need more info maybe, no reference but Rory offered to write short description) including the full set of nuissance parameters as a test case of larger parameter space in which we aim to detect structure.
+The dataset is summarised in \ref{fig:bbhSimulation} where we have removed a subset of less informative parameters to obtain a clear picture.
+Interesting correlations can be seen between the parameters "ra", "dec" and "time", where the first two describe the position of the BBH event in the sky, and the "time" being the merging time (in GPS units).
+Note in particular the relation between "dec" and "time" is not functional, which motivates us to use the TIC index for this example.
+Some other interesting patterns can also be identified, and we next optimise the index in a guided tour.
+
+```{r bbhSimulation, fig.height=8, fig.width=8, dev = "png", dpi=300}
+bbhD <- read_csv("data/posterior_samples.csv")
+bbhDsmall <- select(bbhD, -phi_jl, -m2, -psi, -chi_eff)
+ggpairs(bbhDsmall, lower=list(continuous = wrap("points", alpha = 0.02)))
+```
 

### 2018-10-09 scagnostics.Rmd
Added guided tour results for BBH merger data.

@@
 
+```{r bbhGuided, results="hide", fig.cap="Final view identified optimising the TIC index for the posterior obtained in the BBH merger simulation study."}
+bbhM <- rescale(select(bbhD,-m2))
+if(!file.exists("cache/bbhGuided.rda")){
+  bbhTour <- save_history(bbhM, guided_tour(mineIndex("TIC")))
+  bbhTourFull <- as.list(interpolate(bbhTour))
+  save(bbhTour, bbhTourFull, file = "cache/bbhGuided.rda")
+} else {
+  load("cache/bbhGuided.rda")
+}
+iLast <- length(bbhTourFull)
+fProj <- bbhTourFull[[iLast]]
+dProj <- as.tibble(bbhM %*% fProj)
+ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+```
+
+Indeed the TIC index guided tour found an informative view between the three parameters time, ra and dec whithin the full datase (after once again removing the second mass parameter, see explanation above), which is not described by a function.
+We can approximate it by showing the sum of the rescaled ra and dec variable against time, or when using the original parameter values by approximate scaling, see Figure \ref{fig:bbhRes}, but clearly the view in Fig.
+\ref{fig:bbhGuided} has higher TIC score.
+((interesting to show guided tour animation here?
+draw axes on final view?))
+
+```{r bbhRes, fig.cap="Approximating the final view found in the TIC guided tour optimisation by simple parameter combinations."}
+ggplot(as.tibble(bbhM), aes(time, ra+dec)) + geom_point()
+ggplot(bbhD, aes(time, ra+pi*dec)) + geom_point()
+```
+
 # Applications to collections of time series

### 2018-10-10 scagnostics.Rmd
Compare final view for 3 different index functions.

@@
     return(sR[scagType])
+  }
+}
+
+invConvexIndex <- function(){
+  function(mat){
+    sR <- scagnostics.default(mat[,1],mat[,2])$s
+    return(1 - sR["Convex"])
   }
@@
 ```{r bbhGuided, results="hide", fig.cap="Final view identified optimising the TIC index for the posterior obtained in the BBH merger simulation study."}
-bbhM <- rescale(select(bbhD,-m2))
-if(!file.exists("cache/bbhGuided.rda")){
-  bbhTour <- save_history(bbhM, guided_tour(mineIndex("TIC")))
-  bbhTourFull <- as.list(interpolate(bbhTour))
-  save(bbhTour, bbhTourFull, file = "cache/bbhGuided.rda")
+bbhM <- rescale(select(bbhD,-m2, -chi_eff))
+if(!file.exists("cache/bbhGuidedTIC.rda")){
+  set.seed(2018)
+  bbhTour1 <- save_history(bbhM, guided_tour(mineIndex("TIC")))
+  bbhTourFull1 <- as.list(interpolate(bbhTour1))
+  save(bbhTour1, bbhTourFull1, file = "cache/bbhGuidedTIC.rda")
 } else {
-  load("cache/bbhGuided.rda")
-}
-iLast <- length(bbhTourFull)
-fProj <- bbhTourFull[[iLast]]
-dProj <- as.tibble(bbhM %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+  load("cache/bbhGuidedTIC.rda")
+}
+if(!file.exists("cache/bbhGuidedSpline.rda")){
+  set.seed(2018)
+  bbhTour2 <- save_history(bbhM, guided_tour(splineIndex()))
+  bbhTourFull2 <- as.list(interpolate(bbhTour2))
+  save(bbhTour2, bbhTourFull2, file = "cache/bbhGuidedSpline.rda")
+} else {
+  load("cache/bbhGuidedSpline.rda")
+}
+if(!file.exists("cache/bbhGuidedConvex.rda")){
+  set.seed(2018)
+  bbhTour3 <- save_history(bbhM, guided_tour(invConvexIndex()))
+  bbhTourFull3 <- as.list(interpolate(bbhTour3))
+  save(bbhTour3, bbhTourFull3, file = "cache/bbhGuidedConvex.rda")
+} else {
+  load("cache/bbhGuidedConvex.rda")
+}
+
+iLast <- length(bbhTourFull1)
+fProj1 <- bbhTourFull1[[iLast]]
+dProj <- as.tibble(bbhM %*% fProj1)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+iLast <- length(bbhTourFull2)
+fProj2 <- bbhTourFull2[[iLast]]
+dProj <- as.tibble(bbhM %*% fProj2)
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+iLast <- length(bbhTourFull3)
+fProj3 <- bbhTourFull3[[iLast]]
+dProj <- as.tibble(bbhM %*% fProj3)
+p3 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+grid.arrange(p1, p2, p3, ncol=3) 
 ```

### 2018-10-10 scagnostics.Rmd
Paragraph on rotation invariance added.

@@
   - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
+  - rotation invariance: in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity.
+A projection pursuit index should therefore be invariant under rotation of the basis in the projection plane.
+By costruction this is not the case for most index functions defined above.
+(One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case.) (Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g.
+https://doi.org/10.1016/S0146-664X(75)80008-6, other example is interpolating over angles when building NN, see e.g.
+https://arxiv.org/pdf/1805.12301.pdf)
 

### 2018-10-10 scagnostics.Rmd
Starting cleanup.

@@
   - \usepackage{amssymb}
+  - \usepackage{xcolor}
 ---
@@
 <!--
-suggest that some correlations clearly dictated by the data may be missed, as we are typically only studying selected views of the results.
-The idea of finding special views in scatter plots fits naturally with the concept of using posterior samples obtained in a Bayesian fit.
-
-The idea is that by finding special views of the posterior samples we gain insight into the physics relevant to accurately describe the experimental data.
-Apart from improved intuition, finding special regions in parameter space is important as it allows us to further scrutinize the model.
-There are many examples of this, for example in particle physics detailed studies have been performed e.g.
-in the alignment limit of the 2HDM or in co-annihilation regions in models of dark matter.-->
+Suggestion German:
+We are motivated by physics applications, specifically to assist the interpretation of models fit to experimental results.
+In these problems one has a physical model with a set of m free parameters which cannot be measured directly, but that are determined instead by fitting a set of p experimental observations.
+The model predicts all the observations in terms of the m parameters, often analytically, but occasionally only numerically.
+Moreover, a single prediction can be a complicated function of all the free parameters requiring intensive numerical resources.
+
+These problems often deal with a large number of observations and parameters, with p in the hundreds or even thousands and m of order 10. 
+The results of a fit have been generally interpreted in terms of combinations of parameters selected by intuition or prior knowledge.
+Recently, these selections have been also assisted by machine learning techniques including BDT and neural networks.
+
+Our visualisation tools complement these procedures providing a high dimensional picture which assists with the interpretation of the results,  serves as a check for other methods and possibly revealing new associations.
+-->
 
@@
 The projection pursuit guided tour is available in the R package, tourr [@tourr].
-
-<!-- 
-A related question is the selection of pair-wise scatter plots in datasets with large numbers of parameters, for which different methods have been developed.
-One approach are scatterplot diagnostics, "scagnostics", developed by @scag, and explored in @WW08.
-\footnote{Scagnostics are available in two R packages @LWscagR and @HWscagR.
-This work is built upon the scagnostics package by @HWscagR because the underlying base is built using C, works across platforms.} They are calculated from a set of elementary graph-theoretic building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST).
-These are used to define eight measures: Outlying, Skewed, Sparse, Clumpy, Striated, Convex, Skinny and Stringy.
-While in principle each measure could be used on its own to define a ranking, the original idea suggested considering pair-wise scatter plots of the scagnostics measures for identification of "characteristic" and "unusual" 2-d views in the original parameter space.
-Alternative methods suggested by (Katrins Thesis or maybe her R package?) suggest using parallel coordinates, glyphs or a so-called scagramm (based on the idea of corrgrams) that should be coupled to interactive brushing for selection of pair-wise scatter plots to be displayed.
-(moved scagnostics definitions to appendix, not sure if needed at all.)
-(Issues with scagnostics already for selection of pairs have been described in Katrins Thesis, should have short summary of that here?)
-
-Given the various shortcomings of the scagnostics approach we consider several other measures that can directly be interpreted as ranking of information contained in a given 2-d view.
-Two such additional measures have been defined in @mbgraphic.
-The first is based on smoothing splines, and is defined such that it will take maximum values when the data distribution can be described by a functional form.
-For a scatterplot with variable X and Y we calculate it as
-\begin{equation}
-splines_{2d} = max(1-\frac{Var(res_{X\sim Y})}{Var(X)}, 1-\frac{Var(res_{Y\sim X})}{Var(Y)})
-\end{equation}
-where $Var(.)$ is the variance and $res_{X\sim Y}$ are the residuals in the spline model where $X$ is considered a function of $Y$.
-The second new measure introduced in @mbgraphic is aiming to detect also non-functional relations and is based on the idea of distance correlation [@szekely2007]. (need to add definition)
-Finally we will also consider the maximal and total information coefficient (MIC and TIC) defined in @Reshef1518 which are based on mutual information found when imposing grids on pair-wise variable plots.
-These functions are available in R via the interface implemented in the minerva package [@minerva].
-(read paper and add more detail.
-how are they similar/different compared to scagnostics and distance correlation?
-the way that binning is being built into it may make difference in how MIC/TIC may be smoother?)
-
-A projection pursuit index, a function of all possible projections of the data, invariably has many "hills and valleys" and "knife-edge ridges" because of the varying shapes in the underlying density of observations from one projection to the next.
-From an exploratory data analysis perspective it is interesting to combine numerical optimisation with visualisation to watch the structure of the data moving into a maximum, to jump away from this projection and follow the optimisation again, invariably moving from local maxima to local maxima, and perhaps even a global maximum.
-Each of these maxima can reveal different information about the data.
-Projection pursuit optimisation combined with geodesic interpolations between projections provided by a tour [@As85], is called a projection pursuit guided tour [@CBCH94].
-Software to run a projection pursuit guided tour is available in the R package, "tourr" [@tourr].
-(Motivation strongly relates to intuition, i.e. guided tour, local tour and manual tour may be useful to understand structure in the multivariate context.)
-(Even just for the optimisation part the guided tour implementation is useful.)
-(Traditional projection pursuit indices are fast to compute and therefore the optimisation would be done in "real time" while displaying the tour.
-For the measures considered here this generally does not apply, as a consequence we "disentangle" the two into subsequent steps.)
--->
 
@@
 
-((What is upper bound of TIC as function of sample size?
+\textcolor{red}{What is upper bound of TIC as function of sample size?
 If we define $n' = n^{0.6}$, rounded to integer values only, can we calculate it as $TIC_{max} = 2 \sum_{i=1}^{n'} \frac{n'}{i}$, keeping in mind that each entry of the characteristic matrix should be between [0,1].
-Each entry in the sum should be rounded also, only integer number of bins enter actual calculation, so this should be approximation only.))
-
-((Notes on holes/cmass indices)):
-
-Holes index in $d=2$ dimensional projections, we have projected data given by 
-\begin{equation}
-m = \begin{pmatrix}
-x_1 & y_1 \\
-x_2 & y_2 \\
-\vdots & \vdots \\
-x_n & y_n
-\end{pmatrix}
-\end{equation}
-In tourr calculate holes index as $num/den$ where for $d=2$ we have
-\begin{equation}
-num = 1 - \frac{1}{n} \sum_{i=1}^n e^{(x_i^2+y_i^2)/2}
-\end{equation}
-and $den$ is fixed for fixed number of dimensions, here $den = 1 - \frac{1}{e}$.
-Considering now the case where all $x_i, y_i \leq 1$ we can see that points at the edge may lead to larger values than points on a circle, $x_i^2+y_i^2 = 1$.
-Note also that we cannot give generic upper limit on the index values, it will depend on the range, in particular values $>1$ also possible.
-Cmass calculated as $1-holes$, will take negative values if $holes>1$, e.g.
-the case for entries $x_i, y_i > 1$, example:
+Each entry in the sum should be rounded also, only integer number of bins enter actual calculation, so this should be approximation only.}
+
+\textcolor{red}{Holes and cmass index require standardisation of data, in general PCA, here rescaling to mean zero, standard deviation one should be enough, need to investigate implementation for standardisation in the background.
+Should standardisation be used when working with new indices as well?}
 
@@
 library(minerva) #MINE indices
-```
-
-```{r sillyExample, echo=TRUE}
-
-mX <- matrix(c(2,2,2,2), nrow=2)
-holes(mX)
-cmass(mX)
-# compare random points on circle to extreme positions
-mC <- matrix(c(cos(1.2*pi), sin(1.2*pi), cos(pi/3), sin(pi/3), cos(0.6*pi), sin(0.6*pi)), ncol = 2)
-mT <- matrix(c(1,0,0,1,1,1), ncol=2)
-holes(mC)
-holes(mT)
 ```
@@
 The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of V6 (and V5 for the Pipe distribution) will be distinct from uninformed views of distributions 2 and 3.
-
-<!-- 
-Table and figure captions need to have three components: (1) what is the plot about, (2) specific details of plot, like what type of display and how variables are mapped, (3) the most important thing that the reader should learn.
-
--->
 
@@
 Note that the various indices appear less noisy than for the grand tour shown before, but should note that the tour path here is shorter, and the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
-(What does this mean for optimisation? How can I detect such "smaller scale increase"? Maybe try "search_better" instead of search_geodesic? increase alpha and decrease cooling? increase max.tries?)
-Can force it when using search_better with large alpha, cooling=1, max.tries=1000, but not ideal..
-Difficulties also from "fake" structure from the sampling boundaries?
-Possible issue with search_geodesic is that the search window is fixed in the code..
 
@@
 We also see that TIC is smoother when moving into the special view.
-(As I understand TIC is however not equitable, i.e. it will prefer linear relations, and it may be instructive to optimise MIC and TIC)
-
-FIXME need to check the MIC_e and TIC_e, supposedly faster computation and better discrimination!
 
@@
 On the other hand, surprisingly the cmass index increases when moving into the special view, but the intermediate minimum makes it unsuitable as a projection pursuit index here.
-QUESTION: how come they are all almost the same for the various distributions?
-Especially pipe and sphere seem to give almost identical behaviour, can it be understood from the index definition?
 
@@
 
-No clear interpretation without trace for comparison? See e.g. TIC for path2 pipe vs sine
+\textcolor{red}{Need better representation, in particular normalise TIC to upper limit for given sample size, adapt scales so sharp peaks don't dominate some of the plots.}
 
@@
 
-Double challengind: fine structure + (coherent) noise
+\textcolor{red}{Double challenging: fine structure + (coherent) noise}
 
@@
 
-UPDATE: highlight times corresponding to new basis planes by red vertical line, as we approach the special view higher index values of TIC are often found on interpolated bases, this is why generally we want cooling but here we are trying to find overall maximum, but instead have good chance of finding a plane close to the special view allowing optimisation in a second step
-By no means stable procedure, when changing seed we easily end up not finding special view but rather end up with plane that is not useful for further optimisation to identify pipe view.
-
-There is some issue still with search\_better not actually giving monotonically increasing results, my guess is it is because of rigid assignment of x and y axis, i.e.
-calculating index on starting plane, compare to randomly selected new planes, if new index > starting index and with step size = inf (default when using save history) this means we move to the new plane, but in the interpolation framework, i.e.
-may be rotated wrt new plane found by search\_better, and since now the distribution is the same, but x/y values are not, we get somewhat different values, this may result in somewhat smaller values than anticipated in the new starting plane, which may result in small drops rather than monotonic increase.
+\textcolor{red}{Sometimes index value decreases, because index is not rotationally invariant.}
 
@@
 In this case the optimisation quickly converges, following a smooth, monotonic increas of the TIC index.
-((not shown because that is already full informtion basically))
 
@@
 We next work with distribution 3 and aim to identify the view showing functional dependence using the splines2d index.
-Given the results from Fig 6 ((add proper ref)), we expect to be able to efficiently identify this view without relying on the two step procedure described above.
+Given the results from the planned tour we expect to be able to efficiently identify this view without relying on the two step procedure described above.
 Indeed the guided tour quickly converges in this case, the evolution of the index functions is shown in Fig \ref{fig:findsine}.
@@
 
-```{r neutronStarEq,  results="hide"}
+```{r neutronStarEq,  results="hide", fig.height=3, fig.cap="Final view when optimising the splines2d index for the neutron star merger dataset, showing the functional dependence between the mass, radius and lambda parameter of one of the neutron stars."}
 set.seed(2018)
@@
 ```{r nseMatrix, echo=FALSE}
-knitr::kable(fProj ,  caption = "Matrix representation of best projection.")
-```
+knitr::kable(fProj ,  caption = "Matrix representation of the final projection identified by the guided tour.")
+```
+
+\textcolor{red}{should have function formatting projectino matrices for readable output.}
 
@@
 This is reflecting the dependence of the deformability parameter $\Lambda$ on the so-called compactness which is approximately proportional to $(m/R)^{-5}$ [@Abbott2018exr].
-QUESTION: what is relation between displaying as function of ratio or linear combination? can I see this from doing taylor expansion?
+\textcolor{red}{Move up, this is part of the data description, we get back what was put in.}
 
@@
 @Smith:2016qas.
-We use a simulation study (simulating a binary black hole (BBH) merger event...need more info maybe, no reference but Rory offered to write short description) including the full set of nuissance parameters as a test case of larger parameter space in which we aim to detect structure.
+We use a simulation study \textcolor{red}{simulating a binary black hole (BBH) merger event...need more info maybe, no reference but Rory offered to write short description} including the full set of nuissance parameters as a test case of larger parameter space in which we aim to detect structure.
 The dataset is summarised in \ref{fig:bbhSimulation} where we have removed a subset of less informative parameters to obtain a clear picture.
@@
 
-```{r bbhSimulation, fig.height=8, fig.width=8, dev = "png", dpi=300}
+```{r bbhSimulation, fig.height=8, fig.width=8, dev = "png", dpi=300, fig.cap="Scatter plot matrix showing most of the variables included in the BBH dataset.
+Strong correlation between the parameters time, dec and ra can be observed."}
 bbhD <- read_csv("data/posterior_samples.csv")
@@
 
-```{r bbhGuided, results="hide", fig.cap="Final view identified optimising the TIC index for the posterior obtained in the BBH merger simulation study."}
+```{r bbhGuided, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
 bbhM <- rescale(select(bbhD,-m2, -chi_eff))
@@
 \ref{fig:bbhGuided} has higher TIC score.
-((interesting to show guided tour animation here?
-draw axes on final view?))
+
+\textcolor{red}{Things to add here: comparison of final projection matrices for all three, comparison of index evolution between the three index functions over the TIC guided tour path, plotly animation of the guided tour (check with Nick S.
+about function to readily to this)}
 
@@
 
-# Apendix
-The following measures have been defined:
-
-* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantile edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long.
-The outlying measure is then calculated as $c_{outlying} = \frac{\mathrm{Total\ length\ of\ long\ edges}}{\mathrm{Total\ length\ of\ all\ edges}}$.
-Note that outlying points are not considered in the calculation of the other measures.
-* Skewed: Skewness is defined via the distribution of edge length in the MST, $q_{skew}=\frac{q_{90}-q{50}}{q_{90}-q_{10}}$.
-This is inverted as $c_{skew}= 1-w(1-q_{skew})$ to account for binning effects that result in decreasing values of $q_{skew}$ with increasing n.
-* Sparse: A second measure based on the distribution of edge lenths, we define $c_{sparse} = w q_{90}$.
-It takes large values if points are concentrated in well separated parts of the plane, and low values if they are distributed approximately on a lattice.
-* Clumpy: To indicate the clustering of points the Clumpy measure is defined as $c_{clumpy}=\max\limits_{j}[1-\max\limits_{k}[\frac{length(e_k)}{length(e_j)}]]$ where we maximise the edge length $length(e_k)$ in the smaller of two subgraphs generated by removing a single edge $e_j$ from the MST.
-Note that $c_{clumpy}$ will take large values (close to one) if a long edge is separating clustered points connected by short edges.
-* Striated: We define an edge as striated if they have a large angle with one of the neighboring edges, concretely $cos(angle) < -0.75$, and the corresponding measure $c_{striate} = \frac{Number of striated edges}{Number of all edges}$.
-FIXME is this really correct?
-Maybe reason for discreet behaviour observed below..
-* Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
-* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimeter to the area.
-We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a circle, and values close to one for skinny polygons.
-* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{length(MST)}$ where diameter is longest connected path, and length the total length (sum of all edges), i.e.
-we are measuring branch structure, no branches means $c_{stringy}=1$.
-
 # References

### 2018-10-10 scagnostics.Rmd
Removing unnecessary plots.

@@
 
-```{r neutronStarEq,  results="hide", fig.height=3, fig.cap="Final view when optimising the splines2d index for the neutron star merger dataset, showing the functional dependence between the mass, radius and lambda parameter of one of the neutron stars."}
+```{r neutronStarEq,  results="hide"}
 set.seed(2018)
@@
 dProj <- as.tibble(nsM %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
 ```
@@
 From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection.
-Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{nsePlotOrig} (right), which we compare to the next closest bivariate plot shown on the left.
+Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{fig:nsePlotOrig} (right), which we compare to the next closest bivariate plot shown on the left.
 It is clear that in this example additional information is found when considering general projections rather than focusing only on bivariate relations.
 
-```{r nsePlotOrig}
-p1 <- ggplot(nsD, aes(R1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
+```{r nsePlotOrig, fig.cap="Comparison of guided tour final view (left), approximation based on original parameters (middle) and expected relation based on analysis setup (right)."}
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
 p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
@@
 
-```{r bbhRes, fig.cap="Approximating the final view found in the TIC guided tour optimisation by simple parameter combinations."}
-ggplot(as.tibble(bbhM), aes(time, ra+dec)) + geom_point()
-ggplot(bbhD, aes(time, ra+pi*dec)) + geom_point()
-```
-
 # Applications to collections of time series

### 2018-10-11 scagnostics.Rmd
Going through sec. 2.

@@
 
-[@Grimm2016] explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines that have some nicer properties.
+[@Grimm2016] explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines and distance correlation, that have some nicer properties.
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
@@
 Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $n$ dimensional projections of a data set with $p>n$ parameters will find the most interesting projections.
-As most indices characterise deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations may be defined in different ways, see e.g.
+As most indices characterise deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
 @CBC93.
-When departing from the idea of characterising differences to the normal distribution it may however be preferred to work directly with the empirical data distribution to better capture the true distribution, and instead characterise the distribution by measures based on e.g.
+When departing from the idea of characterising differences to the normal distribution it may however be preferred to work directly with the empirical data distribution and instead characterise it by measures based on e.g.
 the minimal spanning tree or the mutual information.
@@
 The method(s) of choice should be informed by the distributions found in the data set as well as aims of the analysis.
-((particular challenges: noise only directions, identification of outliers?))
+\textcolor{red}{particular challenges: noise only directions, identification of outliers?}
+\textcolor{red}{for holes/cmass index they are only well defined when data is standardised, how about the other index functions?}
+\textcolor{red}{When keeping in mind the concrete example of posterior samples outliers should not be an issue, if extreme outliers are found one should probably investigate origin.
+Also noise only directions, e.g.
+nuissance parameters, are also not an issue as they contain no structure, see BBH example.}
 
@@
 Given a ppi we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
-An important challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
+One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
 A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimisation for the identification of the maximum.
@@
 What we are optimising over are then the entries of the $p\times2$ dimensional projection matrix, considering that they obey orthonormality conditions.
-Concretely here we consider the following types of index functions:
+Concretely here we consider the following types of index functions:\footnote{We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.}
 
-* From the scagnostics family select those best suited to detect shapes:
+* From the scagnostics family select those best suited to detect shapes, with definitions given below.
+They are defined to take values between [0,1].
+\textcolor{red}{The definition of these should be rotation invariant, however, some small dependence is introduced via the hex binning that is done before calculating the scagnostics indexes.}
     + Convex: The convex measure is defined as $c_{convex}= \frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
@@
 
-* New index functions available in @mbgraphic:
+* New index functions available in @mbgraphic, defined to fall between [0,1] \textcolor{red}{and not rotation invariant, as I think Var(X)*Var(Y) introduces rotation invariance even for distance correlation?}:
     + Distance correlation defined in @szekely2007: The definition is based on empiric measures of distance covariance and variance defined on the dataset 
@@
 * Information based index functions from @Reshef1518 are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.
-With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as \footnote{We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.}
+With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as
 \begin{equation}
-M(D)_{x,y} = \frac{I^{\ast} (D, x, y)}{\log \min{x,y}}.
+M(D)_{x,y} = \frac{I^{\ast} (D, x, y)}{\log \min{x,y}},
 \end{equation}
+where the denominator is chosen such that each entry of $M$ will take values between [0,1].
 We can now define the following metrics:
@@
 Each entry in the sum should be rounded also, only integer number of bins enter actual calculation, so this should be approximation only.}
-
-\textcolor{red}{Holes and cmass index require standardisation of data, in general PCA, here rescaling to mean zero, standard deviation one should be enough, need to investigate implementation for standardisation in the background.
-Should standardisation be used when working with new indices as well?}
 

### 2018-10-12 scagnostics.Rmd
Comparing impact of scale, something not right.

@@
 sphere1000 <- sphereData(6, 1000) %>% rescale() %>% as_tibble()
+sphere100S <- scale(sphere100) %>% as_tibble()
+sphere1000S <- scale(sphere1000) %>% as_tibble()
 
@@
 pipe1000 <- pipeData(6,1000) %>% rescale() %>% as_tibble()
+pipe100S <- scale(pipe100) %>% as_tibble()
+pipe1000S <- scale(pipe1000) %>% as_tibble()
 
@@
 sin1000 <- sinData(6, 1000) %>% rescale() %>% as_tibble()
+sin100S <- scale(sin100) %>% as_tibble()
+sin1000S <- scale(sin1000) %>% as_tibble()
 ```
@@
 i <- 1
-pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000",
+             "Sphere 100 S", "Sphere 1000 S", "Pipe 100 S", "Pipe 1000 S", "Sine 100 S", "Sine 1000 S")
 indexOverview <- tibble(convex=numeric(),
@@
                TIC=numeric(),
+               holes=numeric(),
                name=character(),
                vars=character())
-for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
-  ds <- as.tibble(rescale(ds))
+dsL <- list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000,
+            sphere100S, sphere1000S, pipe100S, pipe1000S, sin100S, sin1000S)
+for(ds in dsL){
   for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
@@
     mineRes <- mine(dprj[,1], dprj[,2])
+    holesRes <- holes(dprj)
     indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
@@
                     dcor2d=dcorRes, splines2d=splineRes,
-                    MIC=mineRes$MIC, TIC=mineRes$TIC,
+                    MIC=mineRes$MIC, TIC=mineRes$TIC, holes=holesRes,
                     name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
@@
   n <- length(tPath)
-  df <- rescale(df)
   for (i in 1:n) {
-    dprj <- df %*% tPath[[i]]
+    dprj <- as.matrix(df) %*% tPath[[i]]
     scagRes <- scagnostics(dprj)
@@
     rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, t1full, "pipe", 1000))
+    rbind(getProj(pipe1000, t1full, "pipe", 1000)) %>%
+    rbind(getProj(sin100S, t1full, "sin-S", 100)) %>%
+    rbind(getProj(sin1000S, t1full, "sin-S", 1000)) %>%
+    rbind(getProj(sphere100S, t1full, "sphere-S", 100)) %>%
+    rbind(getProj(sphere1000S, t1full, "sphere-S", 1000)) %>%
+    rbind(getProj(pipe100S, t1full, "pipe-S", 100)) %>%
+    rbind(getProj(pipe1000S, t1full, "pipe-S", 1000))
   save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
@@
 Small samples (p=100) are shown in red, large samples (p=1000) in blue."}
-fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
+fullResMelt <- melt(fullRes, id=c("t", "name", "size")) %>%
+  mutate(sphered = str_detect(name,"-")) %>%
+  mutate(name=str_replace(name,"-S",""))
 ggplot(fullResMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size))) + 
+  geom_line(aes(color=factor(size), linetype = factor(sphered))) + 
   facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
@@
     rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, t2full, "pipe", 1000))
+    rbind(getProj(pipe1000, t2full, "pipe", 1000)) %>%
+    rbind(getProj(sin100S, t1full, "sin-S", 100)) %>%
+    rbind(getProj(sin1000S, t1full, "sin-S", 1000)) %>%
+    rbind(getProj(sphere100S, t1full, "sphere-S", 100)) %>%
+    rbind(getProj(sphere1000S, t1full, "sphere-S", 1000)) %>%
+    rbind(getProj(pipe100S, t1full, "pipe-S", 100)) %>%
+    rbind(getProj(pipe1000S, t1full, "pipe-S", 1000))
   save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
@@
 starting from random distribution view, and moving into the special view."}
-fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
+fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size")) %>%
+  mutate(sphered = str_detect(name,"-")) %>%
+  mutate(name=str_replace(name,"-S",""))
 ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size))) + 
+  geom_line(aes(color=factor(size),linetype = factor(sphered))) + 
   facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +

### 2018-10-12 scagnostics.Rmd
Fixed Fig 5.

@@
     rbind(getProj(pipe1000, t2full, "pipe", 1000)) %>%
-    rbind(getProj(sin100S, t1full, "sin-S", 100)) %>%
-    rbind(getProj(sin1000S, t1full, "sin-S", 1000)) %>%
-    rbind(getProj(sphere100S, t1full, "sphere-S", 100)) %>%
-    rbind(getProj(sphere1000S, t1full, "sphere-S", 1000)) %>%
-    rbind(getProj(pipe100S, t1full, "pipe-S", 100)) %>%
-    rbind(getProj(pipe1000S, t1full, "pipe-S", 1000))
+    rbind(getProj(sin100S, t2full, "sin-S", 100)) %>%
+    rbind(getProj(sin1000S, t2full, "sin-S", 1000)) %>%
+    rbind(getProj(sphere100S, t2full, "sphere-S", 100)) %>%
+    rbind(getProj(sphere1000S, t2full, "sphere-S", 1000)) %>%
+    rbind(getProj(pipe100S, t2full, "pipe-S", 100)) %>%
+    rbind(getProj(pipe1000S, t2full, "pipe-S", 1000))
   save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
@@
 
-```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \ref{fig:plotV1V2toV3V4} but moving from the projection onto V1-V2 to the projection onto V5-V6, i.e.
+```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto V1-V2 to the projection onto V5-V6, i.e.
 starting from random distribution view, and moving into the special view."}

### 2018-10-12 scagnostics.Rmd
findings bullet points
Remark added.

@@
 \textcolor{red}{Double challenging: fine structure + (coherent) noise}
+\textcolor{red}{simple standard guided tour does not find pipe easily with holes index, did not try brute force approach used with TIC index yet}
 
@@
 \label{sec:phys}
+
+<!--
+- First example
+    - Automatically find the functional relationship between three parameters
+- Second example
+    - Stronger relationship between already nonlinearly associated variables, when additional variable is incorporated
+    - 
+-->
 

### 2018-10-12 scagnostics.Rmd
Some additions to bbh examples.

@@
 
-## Data description
+## Example 1
 The model describes a neutron star merger and contains 6 free parameters, i.e.
-each neutron star is described by its mass and radius, and a so-called tidal deform ability parameter $\Lambda$.
+each neutron star is described by its mass $m$ and radius $R$, and a so-called tidal deform ability parameter $\Lambda$ which is a function of the mass and radius, approximately proportional to $(m/R)^{-5}$.
 A posterior sample containing 2538 points is available @ligoData and can be used to constrain the parameters.
-Here we use the "Parametrized-EoS\_nomaxmass\_posterior\_samples.dat" dataset.
+Here we use the "Parametrized-EoS\_nomaxmass\_posterior\_samples.dat" dataset and aim to recover the dependence between the three parameters of the better constrianed system out of a larger parameter set.
 For details about the modelling and statistical procedure see @Abbott2018exr.
@@
 Since in the following we work with 2-d scatter plots we choose to show the data in this fashion, see Fig.
-
-```{r neutronStarSPLOM, fig.height=8, fig.width=8}
+\ref{fig:neutronStarSPLOM}.
+
+```{r neutronStarSPLOM, fig.height=8, fig.width=8, fig.cap="Scatter plot matrix of the neutron star dataset, darker regions represent higher marginalised posterior densities."}
 nsD <- read_csv("data/samples.csv") 
@@
 
-We point out the following features seen from Fig.
-XX: the combination of the two mass parameters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
-We highlight the non-linear relation between the radius $R_1$ and $\Lambda$ parameter and the two-pronged structure of the radius $R_1$.
-For further study we therefore remove m2 from the dataset.
-We first look at the distribution in a grand tour display, where we use a scatter plot display combined with alpha transparency to study the density distribution in the 5-d parameter space.
-The grand tour indicates that the data points are found on a curved surface in the five dimensional space.
-This may stem from a non-linear relation between the parameters, which we will try to uncover using the above developed formalism for new guided tour index functions.
-
-## Using the guided tour
-
-Given the observations from the grand tour display we expect a functional relation between some of the parameters, and we therefore optimise the "splines2d" index function.
-Note that all parameters are standardised to the range $[0,1]$ before running the optimisation.
-Show final view found from the guided tour, give corresponding projection matrix, make corresponding plot from combination of original parameters.
+The following features can be observed: the combination of the two mass parameters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
+We also see the non-linear relation between the radius $R_1$ and $\Lambda$ parameter $L1$ and the two-pronged disribution for the radius $R_1$.
+For further study we therefore remove the second mass $m2$ from the dataset.
+\textcolor{red}{Looking at the distribution in a grand tour display, using a scatter plot display combined with alpha transparency to study the density distribution in the 5-d parameter space, indicates that the data points are found on a curved surface in the five dimensional space.
+This stems from the non-linear relation between the parameters, which we will try to uncover using the above developed formalism for new guided tour index functions.}
+
+### Results
+
+Since we expect a functional relation between some of the parameters, we optimise the "splines2d" index function.
+Note that all parameters are standardised to the range $[0,1]$ before running the optimisation, and we do not sphere the data for ease of interpretation of the results.
 
@@
 
-```{r nseMatrix, echo=FALSE}
-knitr::kable(fProj ,  caption = "Matrix representation of the final projection identified by the guided tour.")
-```
-
-\textcolor{red}{should have function formatting projectino matrices for readable output.}
-
-The resulting projection clearly shows how the data selects a region that is well defined as a combination of several parameters.
+The resulting projection shown in Fig \ref{fig:nsePlotOrig}(left) clearly recovers functional relations between some of the parameters.
 From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection.
-Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{fig:nsePlotOrig} (right), which we compare to the next closest bivariate plot shown on the left.
-It is clear that in this example additional information is found when considering general projections rather than focusing only on bivariate relations.
+Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{fig:nsePlotOrig} (middle), which we compare to the dependence of L1 on the ratio R1/m1 (right).
+For this example the linear combination can thus recover the approximate relation between these parameters.
 
@@
 
-This is reflecting the dependence of the deformability parameter $\Lambda$ on the so-called compactness which is approximately proportional to $(m/R)^{-5}$ [@Abbott2018exr].
-\textcolor{red}{Move up, this is part of the data description, we get back what was put in.}
-
-## Higher D
-The above example is using only a reduced set of the free parameters relevant to the physical interpretation.
+## Example 2
+The above example is using only a reduced set of the free parameters relevant to the physical interpretation, and we already know the expected result by construction.
 In general many additional nuissance parameters will also enter the fit, see e.g.
@@
 Note in particular the relation between "dec" and "time" is not functional, which motivates us to use the TIC index for this example.
-Some other interesting patterns can also be identified, and we next optimise the index in a guided tour.
+Some other interesting patterns can also be identified.
+\textcolor{red}{Question for Rory: is dependence between dec, time, ra built into prior distribution?}
+Show three results:
+ - given all variables, including nuissance, recover dec, time, ra from guided tour (TIC, splines, tried 1-convex but maybe another example is better? skinny?)
+ - removing those three, do we find interesting views?
+ - using smaller subset, find improved description when using more than 2 parameters
 
@@
 dProj <- as.tibble(bbhM %*% fProj1)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 iLast <- length(bbhTourFull2)
@@
 dProj <- as.tibble(bbhM %*% fProj2)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 iLast <- length(bbhTourFull3)
@@
 dProj <- as.tibble(bbhM %*% fProj3)
-p3 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+p3 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 grid.arrange(p1, p2, p3, ncol=3) 
+```
+
+```{r bbhGuided2, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
+bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
+if(!file.exists("cache/bbhGuidedTIC2.rda")){
+  set.seed(2018)
+  bbhTour12 <- save_history(bbhM2, guided_tour(mineIndex("TIC")))
+  bbhTourFull12 <- as.list(interpolate(bbhTour12))
+  save(bbhTour12, bbhTourFull12, file = "cache/bbhGuidedTIC2.rda")
+} else {
+  load("cache/bbhGuidedTIC2.rda")
+}
+if(!file.exists("cache/bbhGuidedSpline2.rda")){
+  set.seed(2018)
+  bbhTour22 <- save_history(bbhM2, guided_tour(splineIndex()))
+  bbhTourFull22 <- as.list(interpolate(bbhTour22))
+  save(bbhTour22, bbhTourFull22, file = "cache/bbhGuidedSpline2.rda")
+} else {
+  load("cache/bbhGuidedSpline2.rda")
+}
+if(!file.exists("cache/bbhGuidedConvex2.rda")){
+  set.seed(2018)
+  bbhTour32 <- save_history(bbhM2, guided_tour(invConvexIndex()))
+  bbhTourFull32 <- as.list(interpolate(bbhTour32))
+  save(bbhTour32, bbhTourFull32, file = "cache/bbhGuidedConvex2.rda")
+} else {
+  load("cache/bbhGuidedConvex2.rda")
+}
+
+iLast <- length(bbhTourFull12)
+fProj1 <- bbhTourFull12[[iLast]]
+dProj <- as.tibble(bbhM2 %*% fProj1)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+iLast <- length(bbhTourFull22)
+fProj2 <- bbhTourFull22[[iLast]]
+dProj <- as.tibble(bbhM2 %*% fProj2)
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+iLast <- length(bbhTourFull32)
+fProj3 <- bbhTourFull32[[iLast]]
+dProj <- as.tibble(bbhM2 %*% fProj3)
+p3 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+grid.arrange(p1, p2, p3, ncol=3) 
+```
+
+```{r bbhGuided3, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
+bbhM3 <- rescale(select(bbhD, alpha, chi_tot, chi_p, ra, dec, distance))
+if(!file.exists("cache/bbhGuidedSpline3.rda")){
+  set.seed(2018)
+  bbhTour23 <- save_history(bbhM3, guided_tour(splineIndex()))
+  bbhTourFull23 <- as.list(interpolate(bbhTour23))
+  save(bbhTour23, bbhTourFull23, file = "cache/bbhGuidedSpline3.rda")
+} else {
+  load("cache/bbhGuidedSpline3.rda")
+}
+iLast <- length(bbhTourFull23)
+fProj2 <- bbhTourFull23[[iLast]]
+dProj <- as.tibble(bbhM3 %*% fProj2)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+p2 <- ggplot(as.tibble(bbhM3), aes(ra, dec)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+grid.arrange(p1, p2, ncol=2)
 ```

### 2018-10-15 scagnostics.Rmd
More examples for bbh dataset.

@@
 ```{r bbhGuided3, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
-bbhM3 <- rescale(select(bbhD, alpha, chi_tot, chi_p, ra, dec, distance))
+bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
+getResWithSeed <- function(s){
+  set.seed(s)
+  startM <- orthonormalise(matrix(runif(14),ncol = 2))
+  tourRes <- save_history(bbhM3, guided_tour(splineIndex()))
+}
+
 if(!file.exists("cache/bbhGuidedSpline3.rda")){
   set.seed(2018)
-  bbhTour23 <- save_history(bbhM3, guided_tour(splineIndex()))
-  bbhTourFull23 <- as.list(interpolate(bbhTour23))
-  save(bbhTour23, bbhTourFull23, file = "cache/bbhGuidedSpline3.rda")
+  seedVals <- sample(1:10000, 6)
+  bbhRes1 <- getResWithSeed(seedVals[1])
+  bbhRes2 <- getResWithSeed(seedVals[2])
+  bbhRes3 <- getResWithSeed(seedVals[3])
+  bbhRes4 <- getResWithSeed(seedVals[4])
+  bbhRes5 <- getResWithSeed(seedVals[5])
+  bbhRes6 <- getResWithSeed(seedVals[6])
+  save(bbhRes1, bbhRes2, bbhRes3, bbhRes4, bbhRes5, bbhRes6, file = "cache/bbhGuidedSpline3.rda")
 } else {
@@
 }
-iLast <- length(bbhTourFull23)
-fProj2 <- bbhTourFull23[[iLast]]
-dProj <- as.tibble(bbhM3 %*% fProj2)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
-p2 <- ggplot(as.tibble(bbhM3), aes(ra, dec)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
-grid.arrange(p1, p2, ncol=2)
+resList <- list(as.list(bbhRes1), as.list(bbhRes2), as.list(bbhRes3),
+                as.list(bbhRes4), as.list(bbhRes5), as.list(bbhRes6))
+pList <- list()
+i <- 1
+for (r in resList){
+  iLast <- length(r)
+  fProj <- r[[iLast]]
+  dProj <- as.tibble(bbhM3 %*% fProj)
+  pList[[i]] <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+  i <- i+1
+}
+grid.arrange(pList[[1]], pList[[2]], pList[[3]],pList[[4]], pList[[5]], pList[[6]], ncol=3)
+```
+
+```{r constructedExample, results="hide", fig.cap="Selecting subset of 6 parameters, for which spline index can be used to improve constraint on parameter combination wrt ra vs dec."}
+bbhC <- rescale(select(bbhD, m1, ra, chi_tot, alpha, distance, dec))
+if(!file.exists("cache/constructedExample.rda")){
+  set.seed(1999)
+  tC <- save_history(bbhC, guided_tour(splineIndex()))
+  save(tC, file = "cache/constructedExample.rda")
+} else {
+  load("cache/constructedExample.rda")
+}
+iL <- length(as.list(tC))
+dProj <- bbhC %*% as.list(tC)[[iL]]
+p1 <- ggplot(as.tibble(dProj), aes(V1, V2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)
+p2 <- ggplot(as.tibble(bbhC),aes(dec-alpha/5, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+p3 <- ggplot(as.tibble(bbhC),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+grid.arrange(p1, p2, p3, ncol=3)
 ```

### 2018-10-15 scagnostics.Rmd
Additonal NSM example.

@@
 
+We now
+
+```{r nseRemL1, fig.cap="Removing L1 and optimise again of the remaining parameters, where m2 remains removed from the set.
+Because of parameter correlations we can recover clear description of L2 as a function of the other parameters, despite m2 missing."}
+nsM2 <- nsD %>%
+  select(-m2, -L1) %>% # remove m2 to avoid finding well known correlation
+  rescale()
+if(!file.exists("cache/neutronRemL1.rda")){
+  set.seed(2018)
+  nseTour2 <- save_history(nsM2, guided_tour(splineIndex()))
+  save(nseTour2, file = "cache/neutronRemL1.rda")
+} else{
+  load("cache/neutronRemL1.rda")
+}
+tL <- as.list(nseTour2)
+iLast <- length(tL)
+fProj <- tL[[iLast]]
+dProj <- as.tibble(nsM2 %*% fProj)
+p1 <- ggplot(as.tibble(dProj), aes(V1, V2))+geom_point(alpha=0.1)+ theme(aspect.ratio=1)
+p2 <- ggplot(as.tibble(nsM2), aes(m1+R1+1.5*R2, L2)) + geom_point(alpha=0.1)+ theme(aspect.ratio=1)
+p3 <- ggplot(nsD, aes(R2/m2, L2))+ geom_point(alpha=0.1)+ theme(aspect.ratio=1)
+grid.arrange(p1, p2, p3, ncol=3)
+```
+
 ## Example 2

### 2018-10-17 scagnostics.Rmd
More text.

@@
 
-We now
+To further investigate the dataset we remove the $L1$ parameter and again use the spline index to find functional dependence between the remaining four parameters.
+For this example, because of the strong correlation between the various parameters, we recover the dependence of $L2$ on the mass and radius of the lighter neutron star, despite having removed its mass from the parameter set.
+In this setting $L2$ is best described as a function of all 3 parameters still considered, see Figure \ref{fig:nseRemL1}.
 
@@
 Interesting correlations can be seen between the parameters "ra", "dec" and "time", where the first two describe the position of the BBH event in the sky, and the "time" being the merging time (in GPS units).
-Note in particular the relation between "dec" and "time" is not functional, which motivates us to use the TIC index for this example.
+Note in particular the relation between "dec" and "time" is not functional, which motivates us to use the TIC index for this example (drawback is however the run time, this data sample contains 10k points, making optimisation of the MINE index functions very inefficient).
 Some other interesting patterns can also be identified.
-\textcolor{red}{Question for Rory: is dependence between dec, time, ra built into prior distribution?}
-Show three results:
- - given all variables, including nuissance, recover dec, time, ra from guided tour (TIC, splines, tried 1-convex but maybe another example is better? skinny?)
- - removing those three, do we find interesting views?
- - using smaller subset, find improved description when using more than 2 parameters
+(These patterns are representative of the likelihood function, since most priors are flat, or build such to capture growth with volume rahter than distance.
+This is still efficient thanks to special sampling techniques that allow jumps in parameter space.)
+
+As a first test we use the full dataset (with exception of $m2$ for the same reasons as stated in the first example), and use different index functions.
+Here we expect to recover some relation between the correlated variables dec, time and ra.
+We compare the final view using three different metrics in Fig.
+\ref{fig:bbhGuided}, showing the views with (local?) maximum of the TIC index (left) spines2d index (middle) and for comparison maximum of 1-convex (right).
+Indeed the TIC and spline index identify dec, time and ra parameters and show views of the 2-d band that is embedded in three dimensions \textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
+Notably, the exact views differ based on the index function selected, optimising TIC is showing us the band "from the side", resulting in clearly non-functional dependence.
+On the other hand, the index based on the scagnostics member convex is too noisy to find any views of interest, and the final  view is essentially equal to the starting plane in this case.
+\textcolor{red}{different scagnostics member may be better suited, but I expect the result will be the same.
+might still be useful to try skinny instead of 1-convex?}.
 
@@
 ```
+
+We next follow a similar strategy as above and remove the variables with the strongest association from the dataset to understand if additional structure is encoded in the distribution.
+Removing thus time, dec and ra variable from the input, we optimise again the TIC, splines2d and 1-convex index functions.
+The results are shown in Fig.
+\ref{fig:bbhGuided2}.
+In this case the TIC index (left) identifies a view presenting seemingly linear correlation, originating however from one well constrained direction orthogonal to the main distribution of points.
+This direction is defined by the chi\_tot and chi\_p variables, for which we can read off the strong correlation in Fig.
+\ref{fig:bbhSimulation}.
+This direction is also found by the spline index (middle view), which in addition identified the bimodal structure in the data stemming from the degeneracy between two plausible values for theta\_jn.
+(Interestingly the hole index in two dimensions does not identify this view, but is expected to find it in a 1-d guided tour.) Finally, as shown on the right, 1-convex does not identify any views of interest here.
 

### 2018-10-17 scagnostics.Rmd
Some more text.

@@
 
-```{r bbhGuided3, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
+We next investigate the dependence of the final result on the input starting plane.
+In particular if there are many local maxima the effect may be large, and many random starting planes may be needed to find a global maximum.
+To this end we first reduce the dataset dimensionality to include only two types of parameters: first, spin related parameters (i.e.
+alpha, thetajn, chitot and chip), and second position related parameters (i.e.
+ra, dec and distance).
+Note that ra and dec are included here, for which we have established strong correlation above.
+We then generate a small set of random starting planes and optimise the splines index (note that it is prefered here over the TIC index as it is much more efficient to compute on large data sets).
+We compare six final views identified in this manner (i.e.
+differing only in randomly selected starting plane) in Fig.
+\ref{fig:bbhGuided3}.
+We note large variation in final views, only one of the six tries identified the view of ra vs dec.
+\textcolor{red}{list corresponding index values for each plot, maybe in the caption.}
+```{r bbhGuided3, results="hide", fig.cap="Final views identified in the dataset considering the seven dimensional parameter space (alpha, thetajn, chitot, chip, ra, dec, distance), differing only by randomly selected starting plane."}
 bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))

### 2018-10-17 scagnostics.Rmd
example 1 revised

@@
 
+This section describes the application of these projection pursuit indices to find two-dimensional structure in two multidimensional gravitational waves problems.
+
+The first example shows 2538 posterior samples obtained by fitting source parameters to the observed gravitational wave signal GW170817 from a neutron star merger [@PhysRevLett119161101].
+Data has been downloaded from @ligoData.
+The fitting procedures are described in detail in @Abbott2018exr.
+There are six parameters of physical interest (6D) with some known relationships.
+Projection pursuit is used to find the known relationships.
+
+The second example shows data from @Smith:2016qas, generated from a simulation study of a binary black hole (BBH) merger event.
+(XXX More info needed.) There are 12 parameters (12-D), with multiple nuisance parameters.
+Projection pursuit uncovers several new relationships between parameters.
+
+## Neutron star merger
+
+Figure \ref{fig:neutronStarSPLOM} shows the scatterplot matrix (with transparency) of the six parameters.
+(In astrophysics, scatterplot matrices are often called "corner plots" REF.) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables.
+From this it can be seen that m1 and m2 are strongly, and slightly, nonlinearly associated.
+Between the other variables, some linear association (R1, R2), some nonlinear association (L1, L2, R1, R2),  heteroskedastic variance in most variables and some bimodality (R1, L1, L2, m1, m2).
+
+The model describes a neutron star merger and contains 6 free parameters, with each neutron star described by its mass $m$ (m1, m2) and radius $R$ (R1, R2), and a so-called tidal deform ability parameter $\Lambda$ (L1, L2) which is a function of the mass and radius, approximately proportional to $(m/R)^{-5}$.
+
+### Data pre-pocessing
+
+Because m1 and m2 are very strongly associated, m2 is dropped before doing PP. This relationship is obvious from the scatterplots of pairs of variables and does not need to be re-discovered by PP.
+
+All variables are scaled to range between 0 and 1.
+The purpose is that range differences in individual variables should not affect the detection  of relationships between multiple variables.
+Standardising the range will still leave differences between the standard deviations of the variables, and for this problem this is preferred.
+Differences in the standard deviations can be important for keeping the non-linear relationships visible to PP.
+
+### Applying PP
+
+With only five parameters, a reasonable start is to examine the 5D space using a grand tour.
+This quickly shows that the strong nonlinear relationships between the parameters.
+PP is then used to extract these relationships.
+The best index for this sort of problem is the splines2d, and it is fast to compute.
+
+Figure \ref{fig:nsePlotOrig} shows the optimal projection found by splines2d, a reconstructed view obtained by manually combining parameters, and a plot of the known relationship between parameters.
+
 <!--
-- First example
-    - Automatically find the functional relationship between three parameters
-- Second example
-    - Stronger relationship between already nonlinearly associated variables, when additional variable is incorporated
-    - 
+(left) clearly recovers functional relations between some of the parameters.
+From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection.
+Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{fig:nsePlotOrig} (middle), which we compare to the dependence of L1 on the ratio R1/m1 (right).
+For this example the linear combination can thus recover the approximate relation between these parameters.
 -->
 
-We now want to apply the developed tools to distributions from realistic physics application.
-Here we choose as an example to use posterior samples obtained when fitting source parameters to the observed gravitational wave signal GW170817 @PhysRevLett119161101.
-The fit details were described in @Abbott2018exr and the corresponding data is available online @ligoData.
-
-## Example 1
-The model describes a neutron star merger and contains 6 free parameters, i.e.
-each neutron star is described by its mass $m$ and radius $R$, and a so-called tidal deform ability parameter $\Lambda$ which is a function of the mass and radius, approximately proportional to $(m/R)^{-5}$.
+<!--
  A posterior sample containing 2538 points is available @ligoData and can be used to constrain the parameters.
-Here we use the "Parametrized-EoS\_nomaxmass\_posterior\_samples.dat" dataset and aim to recover the dependence between the three parameters of the better constrianed system out of a larger parameter set.
+Here we use the "Parametrized-EoS\_nomaxmass\_posterior\_samples.dat" dataset and aim to recover the dependence between the three parameters of the better constrained system out of a larger parameter set.
 For details about the modelling and statistical procedure see @Abbott2018exr.
@@
 \ref{fig:neutronStarSPLOM}.
+-->
 
@@
 
-The following features can be observed: the combination of the two mass parameters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
+<!-- The following features can be observed: the combination of the two mass parameters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
 We also see the non-linear relation between the radius $R_1$ and $\Lambda$ parameter $L1$ and the two-pronged disribution for the radius $R_1$.
@@
 Note that all parameters are standardised to the range $[0,1]$ before running the optimisation, and we do not sphere the data for ease of interpretation of the results.
+-->
 
 ```{r neutronStarEq,  results="hide"}
-set.seed(2018)
-nsD <- read_csv("data/samples.csv") 
 nsM <- nsD %>%
@@
 if(!file.exists("cache/neutronStarEq.rda")){
+  set.seed(2018)
   nseTour <- save_history(nsM, guided_tour(splineIndex()))
@@
 
-The resulting projection shown in Fig \ref{fig:nsePlotOrig}(left) clearly recovers functional relations between some of the parameters.
-From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection.
-Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{fig:nsePlotOrig} (middle), which we compare to the dependence of L1 on the ratio R1/m1 (right).
-For this example the linear combination can thus recover the approximate relation between these parameters.
+```{r}
+theta <- seq(0, 2 * pi, length = 50)
+circ <- tibble(x=cos(theta), y=sin(theta))
+cntrx <- 0
+cntry <- 0.6
+sclx <- 0.1
+scly <- 0.1
+circ <- circ %>% 
+  mutate(x=x*sclx+cntrx, y=y*scly+cntry)
+colnames(fProj) <- c("PP1", "PP2")
+fProj <- as_tibble(fProj)
+```
 
 ```{r nsePlotOrig, fig.cap="Comparison of guided tour final view (left), approximation based on original parameters (middle) and expected relation based on analysis setup (right)."}
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1) + 
+  geom_path(data=circ, aes(x=x, y=y)) + 
+  xlim(-1.1, 0.5) + ylim(-0.2, 1.4) +
+  geom_segment(data=fProj, aes(x=cntrx, xend=PP1*sclx+cntrx, y=cntry, yend=PP2*scly+cntry))
 p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
@@
 
-To further investigate the dataset we remove the $L1$ parameter and again use the spline index to find functional dependence between the remaining four parameters.
-For this example, because of the strong correlation between the various parameters, we recover the dependence of $L2$ on the mass and radius of the lighter neutron star, despite having removed its mass from the parameter set.
-In this setting $L2$ is best described as a function of all 3 parameters still considered, see Figure \ref{fig:nseRemL1}.
+To further investigate relationships between parameters, $L1$ is removed and PP with the splines2D is applied to the remaining four parameters.
+The dependence of $L2$ on the mass and radius of the lighter neutron star, is revealed (Figure \ref{fig:nseRemL1} left plot).
+A manual reconstruction shows this is a relationship between L2, R1, R2 and m1 (middle plot), but it is effectively the known relationship between L2, R2 and m2 (right plot) -- m2 is latently in the relationship though m1.
 
@@
 
-## Example 2
+## Black hole simulation
+
 The above example is using only a reduced set of the free parameters relevant to the physical interpretation, and we already know the expected result by construction.

### 2018-10-17 scagnostics.Rmd
Reference added.

@@
 Figure \ref{fig:neutronStarSPLOM} shows the scatterplot matrix (with transparency) of the six parameters.
-(In astrophysics, scatterplot matrices are often called "corner plots" REF.) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables.
+(In astrophysics, scatterplot matrices are often called "corner plots" [@corner].) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables.
 From this it can be seen that m1 and m2 are strongly, and slightly, nonlinearly associated.

### 2018-10-17 scagnostics.Rmd
Index matrix added for bbh example.

@@
 
-The above example is using only a reduced set of the free parameters relevant to the physical interpretation, and we already know the expected result by construction.
-In general many additional nuissance parameters will also enter the fit, see e.g.
-@Smith:2016qas.
-We use a simulation study \textcolor{red}{simulating a binary black hole (BBH) merger event...need more info maybe, no reference but Rory offered to write short description} including the full set of nuissance parameters as a test case of larger parameter space in which we aim to detect structure.
-The dataset is summarised in \ref{fig:bbhSimulation} where we have removed a subset of less informative parameters to obtain a clear picture.
-Interesting correlations can be seen between the parameters "ra", "dec" and "time", where the first two describe the position of the BBH event in the sky, and the "time" being the merging time (in GPS units).
-Note in particular the relation between "dec" and "time" is not functional, which motivates us to use the TIC index for this example (drawback is however the run time, this data sample contains 10k points, making optimisation of the MINE index functions very inefficient).
-Some other interesting patterns can also be identified.
-(These patterns are representative of the likelihood function, since most priors are flat, or build such to capture growth with volume rahter than distance.
-This is still efficient thanks to special sampling techniques that allow jumps in parameter space.)
-
-As a first test we use the full dataset (with exception of $m2$ for the same reasons as stated in the first example), and use different index functions.
-Here we expect to recover some relation between the correlated variables dec, time and ra.
-We compare the final view using three different metrics in Fig.
-\ref{fig:bbhGuided}, showing the views with (local?) maximum of the TIC index (left) spines2d index (middle) and for comparison maximum of 1-convex (right).
+Figure \ref{fig:bbhSimulation} shows a scatterplot matrix analogue to that shown in Figure \ref{fig:neutronStarSPLOM} but where we have removed a subset of less informative parameters to obtain a clear picture.
+Strong association can be seen between the parameters ra, dec and time, where the first two describe the position of the event in the sky, and time is the merging time (in GPS units).
+Moreover, the relation between dec and time is not functional, which motivates us to use the TIC index for this example (a drawback is however the run time, this data sample contains 10k points, making optimisation of the MINE index functions very inefficient).
+In general we observe non-standard distributions for all variables, as well as various patterns of association between sets of variables.
+These patterns are representative of the likelihood function, since most priors are flat, or build such to capture growth with volume rather than distance.
+This is still efficient thanks to special sampling techniques that allow jumps in parameter space.
+
+### Data pre-pocessing
+We follow the same steps of data pre-processing as described above for the first example.
+First, since constraints on the masses of two systems observed via gravitational waves are strongly correltated, we again remove the smaller mass m2 from the set of parameters.
+We then rescale all parameters to the range from 0 to 1.
+
+### Applying PP
+
+As a first test we use the full dataset and compare results obtained with different index functions.
+Here we expect to recover the relation between the correlated variables dec, time and ra.
+We compare the final view using three different metrics in Figure \ref{fig:bbhGuided}, showing the views with (local?) maximum of the TIC index (left), spines2D index (middle) and for comparison the maximum of 1-convex (right).
 Indeed the TIC and spline index identify dec, time and ra parameters and show views of the 2-d band that is embedded in three dimensions \textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
-Notably, the exact views differ based on the index function selected, optimising TIC is showing us the band "from the side", resulting in clearly non-functional dependence.
+Notably, the exact views differ based on the index function selected, optimising TIC is showing us the band "from the side".
 On the other hand, the index based on the scagnostics member convex is too noisy to find any views of interest, and the final  view is essentially equal to the starting plane in this case.
-\textcolor{red}{different scagnostics member may be better suited, but I expect the result will be the same.
-might still be useful to try skinny instead of 1-convex?}.
+All index values for the three final views are listed in Table \ref{tab:indexTableBBH}, where we see that indeed the splines2D index does not change between the first two views, while the TIC index is increased for the first view wrt the second.
+Finally the value of 1-convex is clearly larger for the first view as compared to the final view, but the optimisation did not succeed for this noisy index function.
 
@@
 fProj1 <- bbhTourFull1[[iLast]]
-dProj <- as.tibble(bbhM %*% fProj1)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+dProj1 <- as.tibble(bbhM %*% fProj1)
+p1 <- ggplot(dProj1, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 iLast <- length(bbhTourFull2)
 fProj2 <- bbhTourFull2[[iLast]]
-dProj <- as.tibble(bbhM %*% fProj2)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+dProj2 <- as.tibble(bbhM %*% fProj2)
+p2 <- ggplot(dProj2, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 iLast <- length(bbhTourFull3)
 fProj3 <- bbhTourFull3[[iLast]]
-dProj <- as.tibble(bbhM %*% fProj3)
-p3 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+dProj3 <- as.tibble(bbhM %*% fProj3)
+p3 <- ggplot(dProj3, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 grid.arrange(p1, p2, p3, ncol=3) 
+```
+
+```{r indexTableBBH}
+getIdxComp <- function(dProj, n){
+  ticI <- mine(dProj$V1, dProj$V2)$TIC
+  splineI <- splines2d(dProj$V1, dProj$V2)
+  invConvI <- 1-scagnostics(as.matrix(dProj))[,"Convex"]
+  return(tibble(TIC=ticI, splines2D=splineI, invConv=invConvI, nView=n))
+}
+idxT <- tibble(TIC=numeric(),
+               splines2D=numeric(),
+               invConv=numeric(),
+               nView=numeric())
+idxT <- rbind(idxT, getIdxComp(dProj1, 1),
+      getIdxComp(dProj2, 2),
+      getIdxComp(dProj3, 3))
+colnames(idxT) <- c("TIC", "splines2D", "1-convex", "Projection")
+knitr::kable(idxT,  caption = "Matrix of index values, for all three indices and all three final projections. \\label{tab:indexTableBBH}", digits=2) 
 ```

### 2018-10-17 scagnostics.Rmd
Index values added to caption.

@@
 )
+opts_knit$set(eval.after = "fig.cap") #so I can use chunck output in the caption
 ```
@@
 
+We next follow a similar strategy as above and remove the variables with the strongest association from the dataset to understand if additional structure is encoded in the distribution.
+Removing thus time, dec and ra variable from the input, we optimise again the TIC, splines2d and 1-convex index functions.
+The results are shown in Fig.
+\ref{fig:bbhGuided2}.
+In this case the TIC index (left) identifies a view presenting seemingly linear correlation, originating however from one well constrained direction orthogonal to the main distribution of points.
+This direction is defined by the chi\_tot and chi\_p variables, for which we can read off the strong correlation in Fig.
+\ref{fig:bbhSimulation}.
+This direction is also found by the splines2D index (middle), which in addition identified the bimodal structure in the data stemming from the degeneracy between two plausible values for theta\_jn.
+(Interestingly the hole index in two dimensions does not identify this view, but is expected to find it in a 1-d guided tour.) Finally, as shown on the right, 1-convex does not identify any views of interest here.
+
+We next investigate the dependence of the final result on the input starting plane.
+In particular if there are many local maxima the effect may be large, and many random starting planes may be needed to find a global maximum.
+To this end we first reduce the dataset dimensionality to include only two types of parameters: first, spin related parameters (i.e.
+alpha, theta\_jn, chi\_tot and chi\_p), and second position related parameters (i.e.
+ra, dec and distance).
+Note that this includes a number of variables for which we have established association and one dimensional structure above, i.e.
+ra and dec; chi\_tot and chi\_p; and theta\_jn.
+We generate a small set of random starting planes and optimise the splines2D index (it is prefered here over the TIC index as it is much more efficient to compute on large data sets).
+We compare six final views identified in this manner (i.e.
+differing only in randomly selected starting plane) in Fig.
+\ref{fig:bbhGuided3}.
+We note large variation in final views, only one of the six tries identified the view of ra vs dec.
+\textcolor{red}{list corresponding index values for each plot, maybe in the caption.}
+
 ```{r bbhSimulation, fig.height=8, fig.width=8, dev = "png", dpi=300, fig.cap="Scatter plot matrix showing most of the variables included in the BBH dataset.
@@
 
-We next follow a similar strategy as above and remove the variables with the strongest association from the dataset to understand if additional structure is encoded in the distribution.
-Removing thus time, dec and ra variable from the input, we optimise again the TIC, splines2d and 1-convex index functions.
-The results are shown in Fig.
-\ref{fig:bbhGuided2}.
-In this case the TIC index (left) identifies a view presenting seemingly linear correlation, originating however from one well constrained direction orthogonal to the main distribution of points.
-This direction is defined by the chi\_tot and chi\_p variables, for which we can read off the strong correlation in Fig.
-\ref{fig:bbhSimulation}.
-This direction is also found by the spline index (middle view), which in addition identified the bimodal structure in the data stemming from the degeneracy between two plausible values for theta\_jn.
-(Interestingly the hole index in two dimensions does not identify this view, but is expected to find it in a 1-d guided tour.) Finally, as shown on the right, 1-convex does not identify any views of interest here.
-
 ```{r bbhGuided2, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
@@
 
-We next investigate the dependence of the final result on the input starting plane.
-In particular if there are many local maxima the effect may be large, and many random starting planes may be needed to find a global maximum.
-To this end we first reduce the dataset dimensionality to include only two types of parameters: first, spin related parameters (i.e.
-alpha, thetajn, chitot and chip), and second position related parameters (i.e.
-ra, dec and distance).
-Note that ra and dec are included here, for which we have established strong correlation above.
-We then generate a small set of random starting planes and optimise the splines index (note that it is prefered here over the TIC index as it is much more efficient to compute on large data sets).
-We compare six final views identified in this manner (i.e.
-differing only in randomly selected starting plane) in Fig.
-\ref{fig:bbhGuided3}.
-We note large variation in final views, only one of the six tries identified the view of ra vs dec.
-\textcolor{red}{list corresponding index values for each plot, maybe in the caption.}
-```{r bbhGuided3, results="hide", fig.cap="Final views identified in the dataset considering the seven dimensional parameter space (alpha, thetajn, chitot, chip, ra, dec, distance), differing only by randomly selected starting plane."}
+```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane.
+The correponding index values for the six views are: (", toString(format(idxL, digits=2)), ").")}
 bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
@@
 pList <- list()
+idxL <- NULL
 i <- 1
@@
   dProj <- as.tibble(bbhM3 %*% fProj)
+  idxL <- c(idxL, splines2d(dProj$V1, dProj$V2))
   pList[[i]] <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)

### 2018-10-17 scagnostics.Rmd
Added short discussion of final example result.

@@
 \ref{fig:bbhGuided3}.
-We note large variation in final views, only one of the six tries identified the view of ra vs dec.
-\textcolor{red}{list corresponding index values for each plot, maybe in the caption.}
+We note large variation in final views, only one of the six tries identified the view of ra vs dec, which however clearly has the maximum index value amongst all views.
+\textcolor{red}{add axes to plot, then we can make statement about how "far" the other views are from the "best" one.}
+
+Finally, given the results above, it makes sense to also consider smaller parameter-subspaces to look for stronger association.
+(A procedure to first scout the parameter space for large index values is not efficent in this case because of the strong dependence on rotation angles in this paricular setup.) One subspace for which we found strong association between three variables is using the randomly selected parameter set (m1, ra, chi\_tot, alpha, distance, dec).
+Optimising again the splines2D index value allows us to find the final view shown in Figure \ref{constructedExample} (left), which we can reproduce from the original parameters as shown in the middle (here we work with the rescaled parameter values), as a combination of the position parameters ra and dec with the spin parameter alpha.
+For comparison we also show the dependence between ra and dec only, which is fuzzy compared to the ideal view identified by the PP optimisation.
+\textcolor{red}{This result is unexpected, it can be understood from ?????}
 
@@
 ```
-
-Indeed the TIC index guided tour found an informative view between the three parameters time, ra and dec whithin the full datase (after once again removing the second mass parameter, see explanation above), which is not described by a function.
-We can approximate it by showing the sum of the rescaled ra and dec variable against time, or when using the original parameter values by approximate scaling, see Figure \ref{fig:bbhRes}, but clearly the view in Fig.
-\ref{fig:bbhGuided} has higher TIC score.
-
-\textcolor{red}{Things to add here: comparison of final projection matrices for all three, comparison of index evolution between the three index functions over the TIC guided tour path, plotly animation of the guided tour (check with Nick S.
-about function to readily to this)}
 

### 2018-10-18 scagnostics.Rmd
All the axes.

@@
 dProj <- as.tibble(nsM %*% fProj)
-```
-
-```{r}
+colnames(fProj) <- c("PP1", "PP2")
+colnames(dProj) <- c("PP1", "PP2")
+fProj <- as_tibble(fProj) %>%
+  add_column(label=colnames(nsM))
+```
+
+```{r axesDrawing}
+#select cntrx, cntry for placement
+#sclx and scly should be 20% of overall range
+getCircle <- function(cntrx, cntry, sclx, scly){
   theta <- seq(0, 2 * pi, length = 50)
   circ <- tibble(x=cos(theta), y=sin(theta))
-cntrx <- 0
-cntry <- 0.6
-sclx <- 0.1
-scly <- 0.1
   circ <- circ %>% 
     mutate(x=x*sclx+cntrx, y=y*scly+cntry)
-colnames(fProj) <- c("PP1", "PP2")
-fProj <- as_tibble(fProj)
+}
+getAxes <- function(fProj, cntrx, cntry, sclx, scly){
+  x1 <- rep(cntrx, length(fProj$PP1))
+  y1 <- rep(cntry, length(fProj[,1]))
+  x2 <- fProj$PP1*sclx+cntrx
+  y2 <- fProj$PP2*scly+cntry
+  lab <- fProj$label
+  axes <- tibble(x1=x1, y1=y1, x2=x2, y2=y2, label=lab)
+}
 ```
@@
 ```{r nsePlotOrig, fig.cap="Comparison of guided tour final view (left), approximation based on original parameters (middle) and expected relation based on analysis setup (right)."}
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1) + 
-  geom_path(data=circ, aes(x=x, y=y)) + 
-  xlim(-1.1, 0.5) + ylim(-0.2, 1.4) +
-  geom_segment(data=fProj, aes(x=cntrx, xend=PP1*sclx+cntrx, y=cntry, yend=PP2*scly+cntry))
+p1 <- ggplot(dProj, aes(PP1, PP2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1) + 
+  geom_path(data=getCircle(0,0.6,0.32,0.22), aes(x=x, y=y)) + 
+  xlim(-1.1, 0.5) + ylim(-0.2, 0.9) +
+  geom_segment(data=getAxes(fProj, 0,0.6,0.32,0.22), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj, 0,0.6,0.32,0.22), aes(x=x2, y=y2, label=label), size=2.5)
 p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
@@
 dProj <- as.tibble(nsM2 %*% fProj)
-p1 <- ggplot(as.tibble(dProj), aes(V1, V2))+geom_point(alpha=0.1)+ theme(aspect.ratio=1)
+colnames(fProj) <- c("PP1", "PP2")
+colnames(dProj) <- c("PP1", "PP2")
+fProj <- as_tibble(fProj) %>%
+  add_column(label=colnames(nsM2))
+p1 <- ggplot(as.tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.1)+ theme(aspect.ratio=1) + 
+  geom_path(data=getCircle(0.5,0.6,0.3,0.18), aes(x=x, y=y)) + 
+  xlim(0, 1.5) + ylim(-0.1, 0.8) +
+  geom_segment(data=getAxes(fProj, 0.5,0.6,0.3,0.18), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj, 0.5,0.6,0.3,0.18), aes(x=x2, y=y2, label=label), size=2.5)
 p2 <- ggplot(as.tibble(nsM2), aes(m1+R1+1.5*R2, L2)) + geom_point(alpha=0.1)+ theme(aspect.ratio=1)
@@
 dProj1 <- as.tibble(bbhM %*% fProj1)
-p1 <- ggplot(dProj1, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+colnames(fProj1) <- c("PP1", "PP2")
+colnames(dProj1) <- c("PP1", "PP2")
+fProj1 <- as_tibble(fProj1) %>%
+  add_column(label=colnames(bbhM))
+p1 <- ggplot(dProj1, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
+  geom_path(data=getCircle(0.4,0.5,0.26,0.24), aes(x=x, y=y)) + 
+  xlim(0.1, 1.4) + ylim(-0.3, 0.9) +
+  geom_segment(data=getAxes(fProj1, 0.4,0.5,0.26,0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj1, 0.4,0.5,0.26,0.24), aes(x=x2, y=y2, label=label), size=2.5)
 iLast <- length(bbhTourFull2)
@@
 dProj2 <- as.tibble(bbhM %*% fProj2)
-p2 <- ggplot(dProj2, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+colnames(fProj2) <- c("PP1", "PP2")
+colnames(dProj2) <- c("PP1", "PP2")
+fProj2 <- as_tibble(fProj2) %>%
+  add_column(label=colnames(bbhM))
+p2 <- ggplot(dProj2, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
+  geom_path(data=getCircle(0.3,-1,0.24,0.22), aes(x=x, y=y)) + 
+  xlim(-0.1, 1.1) + ylim(-1.3, -0.2) +
+  geom_segment(data=getAxes(fProj2,0.3,-1,0.24,0.22), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj2, 0.3,-1,0.24,0.22), aes(x=x2, y=y2, label=label), size=2.5)
 iLast <- length(bbhTourFull3)
@@
 dProj3 <- as.tibble(bbhM %*% fProj3)
-p3 <- ggplot(dProj3, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+colnames(fProj3) <- c("PP1", "PP2")
+colnames(dProj3) <- c("PP1", "PP2")
+fProj3 <- as_tibble(fProj3) %>%
+  add_column(label=colnames(bbhM))
+p3 <- ggplot(dProj3, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
+  geom_path(data=getCircle(0.75, 1, 0.3, 0.28), aes(x=x, y=y)) + 
+  xlim(-0.2, 1.3) + ylim(-0.1, 1.3) +
+  geom_segment(data=getAxes(fProj3,0.75, 1, 0.3, 0.28), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj3, 0.75, 1, 0.3, 0.28), aes(x=x2, y=y2, label=label), size=2.5)
 grid.arrange(p1, p2, p3, ncol=3) 
@@
 getIdxComp <- function(dProj, n){
-  ticI <- mine(dProj$V1, dProj$V2)$TIC
-  splineI <- splines2d(dProj$V1, dProj$V2)
+  ticI <- mine(dProj$PP1, dProj$PP2)$TIC
+  splineI <- splines2d(dProj$PP1, dProj$PP2)
   invConvI <- 1-scagnostics(as.matrix(dProj))[,"Convex"]
@@
 dProj <- as.tibble(bbhM2 %*% fProj1)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+colnames(fProj1) <- c("PP1", "PP2")
+colnames(dProj) <- c("PP1", "PP2")
+fProj1 <- as_tibble(fProj1) %>%
+  add_column(label=colnames(bbhM2))
+p1 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
+  geom_path(data=getCircle(1, 0, 0.34, 0.25), aes(x=x, y=y)) + 
+  xlim(-0.2, 1.5) + ylim(-0.25, 1) +
+  geom_segment(data=getAxes(fProj1,1, 0, 0.34, 0.25), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj1, 1, 0, 0.34, 0.25), aes(x=x2, y=y2, label=label), size=2.5)
 iLast <- length(bbhTourFull22)
@@
 dProj <- as.tibble(bbhM2 %*% fProj2)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+colnames(fProj2) <- c("PP1", "PP2")
+colnames(dProj) <- c("PP1", "PP2")
+fProj2 <- as_tibble(fProj2) %>%
+  add_column(label=colnames(bbhM2))
+p2 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
+  geom_path(data=getCircle(0.75, 0, 0.25, 0.24), aes(x=x, y=y)) + 
+  xlim(-0.25, 1) + ylim(-0.9, 0.3) +
+  geom_segment(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
 iLast <- length(bbhTourFull32)
@@
 dProj <- as.tibble(bbhM2 %*% fProj3)
-p3 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+colnames(fProj3) <- c("PP1", "PP2")
+colnames(dProj) <- c("PP1", "PP2")
+fProj3 <- as_tibble(fProj3) %>%
+  add_column(label=colnames(bbhM2))
+p3 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
+  geom_path(data=getCircle(0, 1.25, 0.32, 0.32), aes(x=x, y=y)) + 
+  xlim(-0.4, 1.2) + ylim(0, 1.6) +
+  geom_segment(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x2, y=y2, label=label), size=2.5)
 grid.arrange(p1, p2, p3, ncol=3) 
@@
 i <- 1
+xmins <- c(-0.5,-0.5,-0.4,-0.9,-1,0.25)
+xmaxs <- c(0.75,0.75,0.8,0.4,0.25,1.5)
+ymins <- c(-0.8,-0.8,-0.7,-0.8,-1.2,-0.6)
+ymaxs <- c(0.7,0.6,0.7,0.7,0.,0.9)
+xscales <- c(0.25,0.25,0.24,0.26,0.25,0.25)
+yscales <- c(0.3,0.28,0.28,0.3,0.24,0.3)
+xcents <- c(0.25,-0.25,-0.1,0.1,-0.75,1.25)
+ycents <- c(0.4,-0.4,-0.4,-0.4,-0.75,0.5)
 for (r in resList){
@@
   dProj <- as.tibble(bbhM3 %*% fProj)
-  idxL <- c(idxL, splines2d(dProj$V1, dProj$V2))
-  pList[[i]] <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+  colnames(fProj) <- c("PP1", "PP2")
+  colnames(dProj) <- c("PP1", "PP2")
+  fProj <- as_tibble(fProj) %>%
+    add_column(label=colnames(bbhM3))
+  idxL <- c(idxL, splines2d(dProj$PP1, dProj$PP2))
+  pList[[i]] <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
+  geom_path(data=getCircle(xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x, y=y)) + 
+  xlim(xmins[i], xmaxs[i]) + ylim(ymins[i], ymaxs[i]) +
+  geom_segment(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x2, y=y2, label=label), size=2.5)
   i <- i+1
@@
 dProj <- bbhC %*% as.list(tC)[[iL]]
-p1 <- ggplot(as.tibble(dProj), aes(V1, V2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)
+fProj <- as.list(tC)[[iL]]
+colnames(fProj) <- c("PP1", "PP2")
+colnames(dProj) <- c("PP1", "PP2")
+fProj <- as_tibble(fProj) %>%
+  add_column(label=colnames(bbhC))
+p1 <- ggplot(as.tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)+ 
+  geom_path(data=getCircle(-0.7, 0.6, 0.2, 0.24), aes(x=x, y=y)) + 
+  xlim(-0.9, 0.1) + ylim(-0.1, 1.1) +
+  geom_segment(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
+  geom_text(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
 p2 <- ggplot(as.tibble(bbhC),aes(dec-alpha/5, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)

### 2018-10-18 scagnostics.Rmd
Large plots should be png format.

@@
 
-```{r bbhGuided, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
+```{r bbhGuided, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right).", dev = "png", dpi=300}
 bbhM <- rescale(select(bbhD,-m2, -chi_eff))
@@
 
-```{r bbhGuided2, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right)."}
+```{r bbhGuided2, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right).", dev = "png", dpi=300}
 bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
@@
 ```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane.
-The correponding index values for the six views are: (", toString(format(idxL, digits=2)), ").")}
+The correponding index values for the six views are: (", toString(format(idxL, digits=2)), ")."), dev = "png", dpi=300}
 bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
@@
 
-```{r constructedExample, results="hide", fig.cap="Selecting subset of 6 parameters, for which spline index can be used to improve constraint on parameter combination wrt ra vs dec."}
+```{r constructedExample, results="hide", fig.cap="Selecting subset of 6 parameters, for which spline index can be used to improve constraint on parameter combination wrt ra vs dec.", dev = "png", dpi=300}
 bbhC <- rescale(select(bbhD, m1, ra, chi_tot, alpha, distance, dec))

### 2018-10-18 scagnostics.Rmd
Plot not rescaled in Fig 18.

@@
 (A procedure to first scout the parameter space for large index values is not efficent in this case because of the strong dependence on rotation angles in this paricular setup.) One subspace for which we found strong association between three variables is using the randomly selected parameter set (m1, ra, chi\_tot, alpha, distance, dec).
-Optimising again the splines2D index value allows us to find the final view shown in Figure \ref{constructedExample} (left), which we can reproduce from the original parameters as shown in the middle (here we work with the rescaled parameter values), as a combination of the position parameters ra and dec with the spin parameter alpha.
+Optimising again the splines2D index value allows us to find the final view shown in Figure \ref{constructedExample} (left) \textcolor{red}{(The small constribution of alpha is hidden in the axes here, middle plot is without rescaling, therefore same size contribution from dec an alpha to y axis)}, which we can reproduce from the original parameters as shown in the middle, as a combination of the position parameters ra and dec with the spin parameter alpha.
 For comparison we also show the dependence between ra and dec only, which is fuzzy compared to the ideal view identified by the PP optimisation.
@@
   geom_text(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
-p2 <- ggplot(as.tibble(bbhC),aes(dec-alpha/5, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
-p3 <- ggplot(as.tibble(bbhC),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+p2 <- ggplot(as.tibble(bbhD),aes(dec-alpha, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+p3 <- ggplot(as.tibble(bbhD),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 grid.arrange(p1, p2, p3, ncol=3)

### 2018-10-18 scagnostics.Rmd
Comment added.

@@
 For comparison we also show the dependence between ra and dec only, which is fuzzy compared to the ideal view identified by the PP optimisation.
-\textcolor{red}{This result is unexpected, it can be understood from ?????}
+\textcolor{red}{This result is unexpected, it can be understood from ?????
+(Does it mean that knowing the precession angle gives more precise information about the location?
+Knowing ra uniquely determines dec-alpha rather than dec it seems.)}
 

### 2018-10-18 scagnostics.Rmd
Minor changes in the text.

@@
 In general we observe non-standard distributions for all variables, as well as various patterns of association between sets of variables.
-These patterns are representative of the likelihood function, since most priors are flat, or build such to capture growth with volume rather than distance.
-This is still efficient thanks to special sampling techniques that allow jumps in parameter space.
+\textcolor{red}{These patterns are representative of the likelihood function, since most priors are flat, or build such to capture growth with volume rather than distance.
+This is still efficient thanks to special sampling techniques that allow jumps in parameter space.}
 
@@
 We follow the same steps of data pre-processing as described above for the first example.
-First, since constraints on the masses of two systems observed via gravitational waves are strongly correltated, we again remove the smaller mass m2 from the set of parameters.
+First, since constraints on the masses of two systems observed via gravitational waves are strongly correlated, we again remove the smaller mass m2 from the set of parameters.
 We then rescale all parameters to the range from 0 to 1.
@@
 Here we expect to recover the relation between the correlated variables dec, time and ra.
-We compare the final view using three different metrics in Figure \ref{fig:bbhGuided}, showing the views with (local?) maximum of the TIC index (left), spines2D index (middle) and for comparison the maximum of 1-convex (right).
-Indeed the TIC and spline index identify dec, time and ra parameters and show views of the 2-d band that is embedded in three dimensions \textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
+We compare the final view using three different metrics in Figure \ref{fig:bbhGuided}, showing the views with maximum of the TIC index (left), spines2D index (middle) and for comparison the local maximum found when optimising 1-convex (right).
+Indeed the TIC and splines2D index identify dec, time and ra parameters and show views of the 2-d band that is embedded in three dimensions \textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
 Notably, the exact views differ based on the index function selected, optimising TIC is showing us the band "from the side".
 On the other hand, the index based on the scagnostics member convex is too noisy to find any views of interest, and the final  view is essentially equal to the starting plane in this case.
-All index values for the three final views are listed in Table \ref{tab:indexTableBBH}, where we see that indeed the splines2D index does not change between the first two views, while the TIC index is increased for the first view wrt the second.
+All index values for the three final views are listed in Table \ref{tab:indexTableBBH}, where we see that the splines2D index does not change between the first two views, while the TIC index is increased for the first view compared the second.
 Finally the value of 1-convex is clearly larger for the first view as compared to the final view, but the optimisation did not succeed for this noisy index function.
@@
 We note large variation in final views, only one of the six tries identified the view of ra vs dec, which however clearly has the maximum index value amongst all views.
-\textcolor{red}{add axes to plot, then we can make statement about how "far" the other views are from the "best" one.}
-
-Finally, given the results above, it makes sense to also consider smaller parameter-subspaces to look for stronger association.
-(A procedure to first scout the parameter space for large index values is not efficent in this case because of the strong dependence on rotation angles in this paricular setup.) One subspace for which we found strong association between three variables is using the randomly selected parameter set (m1, ra, chi\_tot, alpha, distance, dec).
+\textcolor{red}{We see that all final fiews pick up some combination of parameters for which we expect structure, but the structure of the data distribution and corresponding index function dependence is such that one easily gets stuck in local maxima.
+One way out could be the scouting phase discussed above, however this requires some additional improvements, because non-rotation-invariance of the distribution and index function make this approach intractable here.}
+
+Finally, given the results above, it makes sense to also consider smaller parameter-subspaces to look for new association.
+One subspace for which we found strong association between three variables is using the randomly selected parameter set (m1, ra, chi\_tot, alpha, distance, dec).
 Optimising again the splines2D index value allows us to find the final view shown in Figure \ref{constructedExample} (left) \textcolor{red}{(The small constribution of alpha is hidden in the axes here, middle plot is without rescaling, therefore same size contribution from dec an alpha to y axis)}, which we can reproduce from the original parameters as shown in the middle, as a combination of the position parameters ra and dec with the spin parameter alpha.

### 2018-10-19 scagnostics.Rmd
application 2 cleaned up

--- 2018-10-18 scagnostics.Rmd
@@
 
-Figure \ref{fig:bbhSimulation} shows a scatterplot matrix analogue to that shown in Figure \ref{fig:neutronStarSPLOM} but where we have removed a subset of less informative parameters to obtain a clear picture.
-Strong association can be seen between the parameters ra, dec and time, where the first two describe the position of the event in the sky, and time is the merging time (in GPS units).
-Moreover, the relation between dec and time is not functional, which motivates us to use the TIC index for this example (a drawback is however the run time, this data sample contains 10k points, making optimisation of the MINE index functions very inefficient).
-In general we observe non-standard distributions for all variables, as well as various patterns of association between sets of variables.
-These patterns are representative of the likelihood function, since most priors are flat, or build such to capture growth with volume rather than distance.
-This is still efficient thanks to special sampling techniques that allow jumps in parameter space.
-
-### Data pre-pocessing
-We follow the same steps of data pre-processing as described above for the first example.
-First, since constraints on the masses of two systems observed via gravitational waves are strongly correltated, we again remove the smaller mass m2 from the set of parameters.
-We then rescale all parameters to the range from 0 to 1.
-
-### Applying PP
-
-As a first test we use the full dataset and compare results obtained with different index functions.
-Here we expect to recover the relation between the correlated variables dec, time and ra.
-We compare the final view using three different metrics in Figure \ref{fig:bbhGuided}, showing the views with (local?) maximum of the TIC index (left), spines2D index (middle) and for comparison the maximum of 1-convex (right).
-Indeed the TIC and spline index identify dec, time and ra parameters and show views of the 2-d band that is embedded in three dimensions \textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
-Notably, the exact views differ based on the index function selected, optimising TIC is showing us the band "from the side".
-On the other hand, the index based on the scagnostics member convex is too noisy to find any views of interest, and the final  view is essentially equal to the starting plane in this case.
-All index values for the three final views are listed in Table \ref{tab:indexTableBBH}, where we see that indeed the splines2D index does not change between the first two views, while the TIC index is increased for the first view wrt the second.
-Finally the value of 1-convex is clearly larger for the first view as compared to the final view, but the optimisation did not succeed for this noisy index function.
-
-We next follow a similar strategy as above and remove the variables with the strongest association from the dataset to understand if additional structure is encoded in the distribution.
-Removing thus time, dec and ra variable from the input, we optimise again the TIC, splines2d and 1-convex index functions.
-The results are shown in Fig.
-\ref{fig:bbhGuided2}.
-In this case the TIC index (left) identifies a view presenting seemingly linear correlation, originating however from one well constrained direction orthogonal to the main distribution of points.
-This direction is defined by the chi\_tot and chi\_p variables, for which we can read off the strong correlation in Fig.
-\ref{fig:bbhSimulation}.
-This direction is also found by the splines2D index (middle), which in addition identified the bimodal structure in the data stemming from the degeneracy between two plausible values for theta\_jn.
-(Interestingly the hole index in two dimensions does not identify this view, but is expected to find it in a 1-d guided tour.) Finally, as shown on the right, 1-convex does not identify any views of interest here.
-
-We next investigate the dependence of the final result on the input starting plane.
-In particular if there are many local maxima the effect may be large, and many random starting planes may be needed to find a global maximum.
-To this end we first reduce the dataset dimensionality to include only two types of parameters: first, spin related parameters (i.e.
-alpha, theta\_jn, chi\_tot and chi\_p), and second position related parameters (i.e.
-ra, dec and distance).
-Note that this includes a number of variables for which we have established association and one dimensional structure above, i.e.
-ra and dec; chi\_tot and chi\_p; and theta\_jn.
-We generate a small set of random starting planes and optimise the splines2D index (it is prefered here over the TIC index as it is much more efficient to compute on large data sets).
-We compare six final views identified in this manner (i.e.
-differing only in randomly selected starting plane) in Fig.
-\ref{fig:bbhGuided3}.
-We note large variation in final views, only one of the six tries identified the view of ra vs dec, which however clearly has the maximum index value amongst all views.
-\textcolor{red}{add axes to plot, then we can make statement about how "far" the other views are from the "best" one.}
-
-Finally, given the results above, it makes sense to also consider smaller parameter-subspaces to look for stronger association.
-(A procedure to first scout the parameter space for large index values is not efficent in this case because of the strong dependence on rotation angles in this paricular setup.) One subspace for which we found strong association between three variables is using the randomly selected parameter set (m1, ra, chi\_tot, alpha, distance, dec).
-Optimising again the splines2D index value allows us to find the final view shown in Figure \ref{constructedExample} (left) \textcolor{red}{(The small constribution of alpha is hidden in the axes here, middle plot is without rescaling, therefore same size contribution from dec an alpha to y axis)}, which we can reproduce from the original parameters as shown in the middle, as a combination of the position parameters ra and dec with the spin parameter alpha.
-For comparison we also show the dependence between ra and dec only, which is fuzzy compared to the ideal view identified by the PP optimisation.
-\textcolor{red}{This result is unexpected, it can be understood from ?????
-(Does it mean that knowing the precession angle gives more precise information about the location?
-Knowing ra uniquely determines dec-alpha rather than dec it seems.)}
+This data contains posterior samples for simulation from a model describing a binary black hole (BBH) merger event. There are twelve model parameters. Flat priors are used for most model parameters.
+
+Figure \ref{fig:bbhSimulation} shows a scatterplot matrix, of nine of the twelve parameters.
+(Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time.
+The first two describe the position of the event in the sky, and time is the merging time (in GPS units).
+Because of the elliptical relationship between dec and time, the TIC index is used for PP, even though it is slow to compute.
+Between the other parameters, the main structure seen is multimodality and some skewness.
+These patterns are representative of the likelihood function, since most priors are flat, or built to capture growth with volume rather than distance.
 
@@
 
-```{r bbhGuided, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right).", dev = "png", dpi=300}
+### Data pre-processing
+
+The analysis is conducted on 11 of the twelve parameters. One variable is removed, m2, because it is so strongly associated with m1. All parameters are scaled into the range 0 to 1.
+
+### Applying PP
+
+#### Exploring 11D with all PP indexes
+
+All seven PP indexes are applied to the data.
+Figure \ref{fig:bbhGuided} showing the projections that maximise three of the indexes.
+TIC and splines2D indexes identify very similar projections, that are based on the three parameters, dec, time and ra.
+This is ot be expected based on the pairwise scatterplots.
+\textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
+On the other hand, the 1-convex index finds a very different view, but actually the problem is that the optimisation doesn't adequately reach a maximum for this index.
+
+```{r bbhGuided, results="hide", fig.cap="Projections corresponding to the maxima of three indices: TIC, splines2D and 1-convex.
+Projections a, b found by TIC and splines3D are very similar, and involve the same three parameters, ra, dec and time.
+The 1-convex index finds a very different view.", dev = "png", dpi=300}
 bbhM <- rescale(select(bbhD,-m2, -chi_eff))
@@
   geom_segment(data=getAxes(fProj1, 0.4,0.5,0.26,0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj1, 0.4,0.5,0.26,0.24), aes(x=x2, y=y2, label=label), size=2.5)
+  geom_text(data=getAxes(fProj1, 0.4,0.5,0.26,0.24), aes(x=x2, y=y2, label=label), size=2.5) + ggtitle("a. TIC: 779.46")
 iLast <- length(bbhTourFull2)
@@
   geom_segment(data=getAxes(fProj2,0.3,-1,0.24,0.22), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj2, 0.3,-1,0.24,0.22), aes(x=x2, y=y2, label=label), size=2.5)
+  geom_text(data=getAxes(fProj2, 0.3,-1,0.24,0.22), aes(x=x2, y=y2, label=label), size=2.5) + ggtitle("b. splines2D: 0.97")
 iLast <- length(bbhTourFull3)
@@
   geom_segment(data=getAxes(fProj3,0.75, 1, 0.3, 0.28), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj3, 0.75, 1, 0.3, 0.28), aes(x=x2, y=y2, label=label), size=2.5)
+  geom_text(data=getAxes(fProj3, 0.75, 1, 0.3, 0.28), aes(x=x2, y=y2, label=label), size=2.5) + ggtitle("c. 1-convex: 0.50")
 grid.arrange(p1, p2, p3, ncol=3) 
@@
 
-```{r indexTableBBH}
+```{r indexTableBBH, eval=FALSE}
 getIdxComp <- function(dProj, n){
@@
 
-```{r bbhGuided2, results="hide", fig.cap="Final views identified by the guided tour using the TIC index (left), the splines2d index (middle) or 1-convex (right).", dev = "png", dpi=300}
+#### Exploring reduced space
+
+The variables time, dec and ra are dropped from the data, and PP is applied to the remaining 8D space.
+Figure \ref{fig:bbhGuided2} shows the projections which maximise the TIC, splines2D and 1-convex indices.
+The results provide similar information as laready learned from the scatterplot matrix (Figure \ref{bbhSimulation}).
+The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2D index.
+The 1-convex index finds nothing interesting.
+
+```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2D and 1-convex.", dev = "png", dpi=300}
 bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
@@
 
+#### Effect of random starts, and subsets used
+
+The initial conditions for the optimisation, and the subset of variables used, can have a large effect on the projections returned.
+We illustrate this using only the splines2D index, and find that there is one more association that can be learned that was masked earlier.
+
+Fig.
+\ref{fig:bbhGuided3} shows six maxima obtained by different starts, for two types of parameters: first, spin related parameters (i.e.
+alpha, theta\_jn, chi\_tot and chi\_p), and second position related parameters (i.e.
+ra, dec and distance).
+Four of the six (a-d) are almost identical, but not interesting projections.
+Projection f has the highest PP index value but it is primarily the view seen in the bivariate plot of dec and ra.
+
+Choosing a different subset of variables reveals something new.
+The subspace of m1, ra, chi\_tot, alpha, distance, dec produces a more refined view of Fig.
+\ref{fig:bbhGuided3} projection f.
+When alpha contributes in contrast to dec, the relationship between the points is almost perfectly on a curve.
+This is shown in Figure \ref{constructedExample}.
+Manually reconstructing this can be done by differencing the two parameters, in their original units.
+\textcolor{red}{This result is unexpected, it can be understood from ?????
+(Does it mean that knowing the precision angle gives more precise information about the location?
+Knowing ra uniquely determines dec-alpha rather than dec it seems.)}
+
 ```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane.
@@
 
+<!--
+
 # Applications to collections of time series
@@
 
+-->
+
 # Discussion

### 2018-10-19 scagnostics.Rmd
Added small study on rotation dependence.

--- 2018-10-18 scagnostics.Rmd
@@
   facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
+```
+
+## Rotation dependence
+\textcolor{red}{testin rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
+Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binnibng (and outlier removal?) here.
+Observe huge drops for some of the indexes!
+```{r rotationDep, fig.width=6, fig.height=7, fig.cap="Rotation dependence of the considered index functions tested on the sine 1000 dataset, by rotating the special view by an angle alpha."}
+sineM <- as.matrix(select(sin1000, V5, V6))
+sc <- tibble(
+    holes=numeric(),
+    cmass=numeric(),
+    convex=numeric(),
+    skinny=numeric(),
+    stringy=numeric(),
+    dcor2d=numeric(),
+    splines2d=numeric(),
+    MIC=numeric(),
+    TIC=numeric(),
+    alpha=numeric()
+)
+for (a in seq(0,2*pi, pi/20)){
+  rotM <- matrix(c(cos(a), sin(a), -sin(a), cos(a)), ncol = 2)
+  dprj <- sineM %*% rotM
+  scagRes <- scagnostics(dprj)
+  dcorRes <- dcor2d(dprj[,1], dprj[,2])
+  splineRes <- splines2d(dprj[,1], dprj[,2])
+  mineRes <- mine(dprj[,1], dprj[,2])
+  holesRes <- holes(dprj)
+  cmassRes <- cmass(dprj)
+  sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
+                convex=scagRes[,"Convex"], skinny=scagRes[,"Skinny"],
+                stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
+                splines2d=splineRes, MIC=mineRes$MIC,
+                TIC=mineRes$TIC, alpha=a)
+  }
+
+scMelt <- melt(sc, id=c("alpha"))
+ggplot(scMelt, aes(x=alpha, y=value)) +
+  geom_line() + 
+  facet_wrap(~variable, scales = "free_y", ncol = 1)
 ```

### 2018-10-21 scagnostics.Rmd
Example data scaled everywhere.

@@
 set.seed(1984)
-sphere100 <- sphereData(6, 100) %>% rescale() %>% as_tibble()
-sphere1000 <- sphereData(6, 1000) %>% rescale() %>% as_tibble()
-sphere100S <- scale(sphere100) %>% as_tibble()
-sphere1000S <- scale(sphere1000) %>% as_tibble()
-
-pipe100 <- pipeData(6, 100) %>% rescale() %>% as_tibble()
-pipe1000 <- pipeData(6,1000) %>% rescale() %>% as_tibble()
-pipe100S <- scale(pipe100) %>% as_tibble()
-pipe1000S <- scale(pipe1000) %>% as_tibble()
-
-sin100 <- sinData(6, 100) %>% rescale() %>% as_tibble()
-sin1000 <- sinData(6, 1000) %>% rescale() %>% as_tibble()
-sin100S <- scale(sin100) %>% as_tibble()
-sin1000S <- scale(sin1000) %>% as_tibble()
+sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
+sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()
+
+pipe100 <- pipeData(6, 100) %>% scale() %>% as_tibble()
+pipe1000 <- pipeData(6,1000) %>% scale() %>% as_tibble()
+
+sin100 <- sinData(6, 100) %>% scale() %>% as_tibble()
+sin1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
 ```
@@
 i <- 1
-pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000",
-             "Sphere 100 S", "Sphere 1000 S", "Pipe 100 S", "Pipe 1000 S", "Sine 100 S", "Sine 1000 S")
+pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
 indexOverview <- tibble(convex=numeric(),
@@
                vars=character())
-dsL <- list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000,
-            sphere100S, sphere1000S, pipe100S, pipe1000S, sin100S, sin1000S)
+dsL <- list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)
 for(ds in dsL){
@@
     rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, t1full, "pipe", 1000)) %>%
-    rbind(getProj(sin100S, t1full, "sin-S", 100)) %>%
-    rbind(getProj(sin1000S, t1full, "sin-S", 1000)) %>%
-    rbind(getProj(sphere100S, t1full, "sphere-S", 100)) %>%
-    rbind(getProj(sphere1000S, t1full, "sphere-S", 1000)) %>%
-    rbind(getProj(pipe100S, t1full, "pipe-S", 100)) %>%
-    rbind(getProj(pipe1000S, t1full, "pipe-S", 1000))
+    rbind(getProj(pipe1000, t1full, "pipe", 1000))
   save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
@@
 Small samples (p=100) are shown in red, large samples (p=1000) in blue."}
-fullResMelt <- melt(fullRes, id=c("t", "name", "size")) %>%
-  mutate(sphered = str_detect(name,"-")) %>%
-  mutate(name=str_replace(name,"-S",""))
+fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
 ggplot(fullResMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size), linetype = factor(sphered))) + 
+  geom_line(aes(color=factor(size))) + 
   facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
@@
     rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, t2full, "pipe", 1000)) %>%
-    rbind(getProj(sin100S, t2full, "sin-S", 100)) %>%
-    rbind(getProj(sin1000S, t2full, "sin-S", 1000)) %>%
-    rbind(getProj(sphere100S, t2full, "sphere-S", 100)) %>%
-    rbind(getProj(sphere1000S, t2full, "sphere-S", 1000)) %>%
-    rbind(getProj(pipe100S, t2full, "pipe-S", 100)) %>%
-    rbind(getProj(pipe1000S, t2full, "pipe-S", 1000))
+    rbind(getProj(pipe1000, t2full, "pipe", 1000))
   save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
@@
 starting from random distribution view, and moving into the special view."}
-fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size")) %>%
-  mutate(sphered = str_detect(name,"-")) %>%
-  mutate(name=str_replace(name,"-S",""))
+fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
 ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size),linetype = factor(sphered))) + 
+  geom_line(aes(color=factor(size))) + 
   facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +

### 2018-10-21 scagnostics.Rmd
Draw axes circles in grey.

@@
 p1 <- ggplot(dProj, aes(PP1, PP2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0,0.6,0.32,0.22), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(0,0.6,0.32,0.22), aes(x=x, y=y), color="grey") + 
   xlim(-1.1, 0.5) + ylim(-0.2, 0.9) +
@@
 p1 <- ggplot(as.tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.1)+ theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.5,0.6,0.3,0.18), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(0.5,0.6,0.3,0.18), aes(x=x, y=y), color="grey") + 
   xlim(0, 1.5) + ylim(-0.1, 0.8) +
@@
 p1 <- ggplot(dProj1, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.4,0.5,0.26,0.24), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(0.4,0.5,0.26,0.24), aes(x=x, y=y), color="grey") + 
   xlim(0.1, 1.4) + ylim(-0.3, 0.9) +
@@
 p2 <- ggplot(dProj2, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.3,-1,0.24,0.22), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(0.3,-1,0.24,0.22), aes(x=x, y=y), color="grey") + 
   xlim(-0.1, 1.1) + ylim(-1.3, -0.2) +
@@
 p3 <- ggplot(dProj3, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.75, 1, 0.3, 0.28), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(0.75, 1, 0.3, 0.28), aes(x=x, y=y), color="grey") + 
   xlim(-0.2, 1.3) + ylim(-0.1, 1.3) +
@@
 p1 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(1, 0, 0.34, 0.25), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(1, 0, 0.34, 0.25), aes(x=x, y=y), color="grey") + 
   xlim(-0.2, 1.5) + ylim(-0.25, 1) +
@@
 p2 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(0.75, 0, 0.25, 0.24), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(0.75, 0, 0.25, 0.24), aes(x=x, y=y), color="grey") + 
   xlim(-0.25, 1) + ylim(-0.9, 0.3) +
@@
 p3 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(0, 1.25, 0.32, 0.32), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(0, 1.25, 0.32, 0.32), aes(x=x, y=y), color="grey") + 
   xlim(-0.4, 1.2) + ylim(0, 1.6) +
@@
   pList[[i]] <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x, y=y), color="grey") + 
   xlim(xmins[i], xmaxs[i]) + ylim(ymins[i], ymaxs[i]) +
@@
 p1 <- ggplot(as.tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(-0.7, 0.6, 0.2, 0.24), aes(x=x, y=y)) + 
+  geom_path(data=getCircle(-0.7, 0.6, 0.2, 0.24), aes(x=x, y=y), color="grey") + 
   xlim(-0.9, 0.1) + ylim(-0.1, 1.1) +

### 2018-10-22 scagnostics.Rmd
Keep all timers, still need to combine to meaningful plot.

@@
 
-timeThis <- function(d, t, idx, pmax){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections
+timeThis <- function(d, t, idx, pmax, n, idxName){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections, n=sample size, idxName=str name of idx funciton
     i <- 1
+    dfTimer = data.frame( t= numeric(), i=numeric(), n = numeric(), name=character())
+    for(pMatrix in t){
+      if(i>pmax) break
       tic.clearlog()
       tic() #start timer
-    for(pMatrix in t){
-      if(i>pmax) break
       dProj <- d %*% pMatrix
       sgnst <- idx(dProj[,1],dProj[,2])
-      i = i+1
-    }
       toc(log=TRUE,quiet=TRUE)
       scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
+      dfTimer <- add_row(dfTimer, t=scTd, i=i, n= n, name=idxName)
+      i = i+1
+    }
+    return(dfTimer)
 }
@@
   sizeL <- c(10, 100, 500, 1000, 5000, 10000)
-  n <- length(sizeL)
-  dfTimer = data.frame( n=rep(0, n), scagT=rep(0,n), dcorT=rep(0,n), splineT=rep(0,n), mineT=rep(0,n),
-                        mineeT=rep(0,n))
   t <- 1
+  dfTimer <- NULL
   for(sampleSize in sizeL){
@@
     thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
-    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100)
-    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100)
-    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100)
-    mineT <- timeThis(thisMatrix, grandTour100, mine, 100)
-    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100)
-    dfTimer[t,] <- c(sampleSize, scagT, dcorT, splineT, mineT,mineeT)
+    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100, sampleSize, "scagnostics")
+    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100, sampleSize, "dcor2D")
+    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100, sampleSize, "splines2D")
+    mineT <- timeThis(thisMatrix, grandTour100, mine, 100, sampleSize, "MINE")
+    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100, sampleSize, "MINE E")
+    dfTimerC <- list(scagT, dcorT, splineT, mineT,mineeT) %>%
+      reduce(rbind)
+    dfTimer <- rbind(dfTimer, dfTimerC)
     t <-  t+1
@@
 
-timerMelt <- melt(dfTimer, id="n")
-ggplot(timerMelt, aes(x = n, y = value, color=variable)) + 
+#timerMelt <- melt(dfTimer, id="n")
+ggplot(dfTimer, aes(x = n, y = t, color=name)) + 
   geom_point() +
@@
 
-```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane.
-The correponding index values for the six views are: (", toString(format(idxL, digits=2)), ")."), dev = "png", dpi=300}
+```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane."), dev = "png", dpi=300}
 bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
@@
 pList <- list()
-idxL <- NULL
+idxV <- NULL
 i <- 1
@@
 ycents <- c(0.4,-0.4,-0.4,-0.4,-0.75,0.5)
+titleL <- c("a: ","b: ","c: ","d: ","f: ","e: ")
 for (r in resList){
@@
     add_column(label=colnames(bbhM3))
-  idxL <- c(idxL, splines2d(dProj$PP1, dProj$PP2))
+  idxV <- splines2d(dProj$PP1, dProj$PP2)
   pList[[i]] <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
@@
   geom_segment(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x2, y=y2, label=label), size=2.5)
+  geom_text(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x2, y=y2, label=label), size=2.5) +
+    ggtitle(paste0(titleL[i], toString(format(idxV, digits=2))))
   i <- i+1
 }
-grid.arrange(pList[[1]], pList[[2]], pList[[3]],pList[[4]], pList[[5]], pList[[6]], ncol=3)
+grid.arrange(pList[[1]], pList[[2]], pList[[3]],pList[[4]], pList[[6]], pList[[5]], ncol=3)
 ```

### 2018-10-22 scagnostics.Rmd
Added mean and connecting line to timing plot.

@@
 
+\textcolor{red}{The plot is not perfect.
+I use alpha to fade out points, but cant find value that gives nice picture here.
+In addition the timer isnt precise enough to resolve timing for some indexes on small datasets, they are zeroes, appear on bottom of the log10 y axis.}
+
 ```{r getTimer}
@@
 
-#timerMelt <- melt(dfTimer, id="n")
+timerMeans <- dfTimer %>%
+  group_by(name, n) %>%
+  summarise(t=mean(t))
+
 ggplot(dfTimer, aes(x = n, y = t, color=name)) + 
-  geom_point() +
+  geom_point(alpha=0.05) +
+  geom_point(data=timerMeans) +
+  geom_line(data=timerMeans) +
   scale_x_log10() +

### 2018-10-22 scagnostics.Rmd
Added title/values to Fig. 17

@@
   add_column(label=colnames(bbhM2))
+iv <- mine(dProj$PP1,dProj$PP2)["TIC"]
 p1 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
@@
   geom_segment(data=getAxes(fProj1,1, 0, 0.34, 0.25), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj1, 1, 0, 0.34, 0.25), aes(x=x2, y=y2, label=label), size=2.5)
+  geom_text(data=getAxes(fProj1, 1, 0, 0.34, 0.25), aes(x=x2, y=y2, label=label), size=2.5) +
+  ggtitle(paste0("a, TIC: ", toString(format(iv, digits=2))))
 iLast <- length(bbhTourFull22)
@@
   add_column(label=colnames(bbhM2))
+iv <- splines2d(dProj$PP1,dProj$PP2)
 p2 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
@@
   geom_segment(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
+  geom_text(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x2, y=y2, label=label), size=2.5) +
+  ggtitle(paste0("b, splines2D: ", toString(format(iv, digits=2))))
 iLast <- length(bbhTourFull32)
@@
   add_column(label=colnames(bbhM2))
+iv <- 1-scagnostics(as.matrix(dProj))[,"Convex"]
 p3 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
@@
   geom_segment(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x2, y=y2, label=label), size=2.5)
+  geom_text(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x2, y=y2, label=label), size=2.5) +
+  ggtitle(paste0("c, 1-convex: ", toString(format(iv, digits=2))))
 grid.arrange(p1, p2, p3, ncol=3) 

### 2018-10-23 scagnostics.Rmd
References and comments added.

@@
 
+\textcolor{red}{\cite{jones87} discusses centering and sphering of the data points.
+Centering is needed if index is not translational invariant (should not be the case for any index considered here).
+Arguments for sphering are that scale effects should be extracted, we don't want to reproduce information already obtained by PCA.
+To my mind rescaling to an overall intervall is more appropriate here.
+Rotations between parameters should not affect the outcome, but will make interpretation more difficult.
+Standardising the variance will remove information about the posterior variance.
+Overall we are interested in strong association only, if these structures were affected by the standardisation procedure, it would most likely be unwelcome effects?}
+
 ## Optimization
@@
 In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
+
+\textcolor{red}{\cite{posse95b} discusses the optimisation, in particular that for most index functions and optimisers results are too local, largely dependent on starting point.
+Fiedmans suggestion of hybrid steppind search + steepest-ascent still remains fairly local.
+\cite{posse90} suggested random search instead, which seems to perform better, but did not find pdf of that article, so not so sure what he actually does.}
 
@@
 
+\textcolor{red}{\cite{posse95b} lists five criteria: minimized by normal distribution, approximately affine invariant, consistent, resistent to features in the tail, fast computation.
+Comment that different index values for the same plane will make optimisation less efficient.
+\cite{jones87} moreover lists rotation and translation invariance (independence of choice of basis in the plane), and also mentions conditions for numeric optimisation, i.e.
+continuous and depending on optimisation methods also differentiable.
+Interesting: according to \cite{jones87} the original index proposed by Friedman and Tukey did not have property of rotation invariance.}
+
   - smoothness: the index function should evolve smoothly between nearby projections, i.e.
@@
 This poses problems in particular for the optimization, but we may use index definitions to amplify larger index values while suppressing smaller ones, such that the optimization will be less sensitive to noise.
+\textcolor{red}{there are two different considerations, one is that described above, it relates to the concept of squint angle discussed in \cite{posse95b}.
+On the other hand should also consider detectin of "fine" structure vs noise, this would be controlled e.g.
+by the allowed number of bins for the MINE index functions, additionally this will impact computing times, i.e.
+trade accuracy for efficiency}
   - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
@@
 ##Simulation study
+
+\textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
+truncated Gaussian distributions with approximately identity covariance matrix, such that they dont appear in any of the principal component planes.}
+
 To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.

### 2018-10-23 scagnostics.Rmd
Cross referencing in Sec 3.

@@
 we observe sharp jumps when looking at the evolution of the index function, and in particular if those fluctuations dominate the overall evolution, the index is not suitable for projection pursuit.
+We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
   - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
 Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified.
+Computation times for some benchmark scenarios are compared in Section \ref{sec:timer}.
   - ability to find fine and broad structure: We consider a broad structure one that is roughly observable over a large angle around the optimal projection, it should in general be easy to identify and a good ppi should show smooth increase over a large number of projections as we are approaching the optimal view.
@@
 by the allowed number of bins for the MINE index functions, additionally this will impact computing times, i.e.
-trade accuracy for efficiency}
+trade accuracy for efficiency}.
+We address these questions in Section \ref{sec:params}
   - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
+\textcolor{red}{I dont really understand this part.} We show examples in Section \ref{sec:dist}
   - rotation invariance: in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity.
@@
 By costruction this is not the case for most index functions defined above.
-(One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case.) (Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g.
+\textcolor{red}{One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case.
+Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g.
 https://doi.org/10.1016/S0146-664X(75)80008-6, other example is interpolating over angles when building NN, see e.g.
-https://arxiv.org/pdf/1805.12301.pdf)
+https://arxiv.org/pdf/1805.12301.pdf)}.
+We illustrate the impact of non invariance in Section \ref{sec:rot}
 
@@
 ## Smoothness
+\label{sec:smooth}
 We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto V1-V2 to the projection onto V3-V4 and using the default interpolation angle of 0.05.
@@
 ## Distributions of index values
-
+\label{sec:dist}
 Compare typical index values over the two tour paths.
@@
 ## Rotation dependence
+\label{sec:rot}
 \textcolor{red}{testin rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
@@
 ## Note on calculation time
+\label{sec:timer}
 Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example the 6-d sphere and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
@@
 ```
+
+## Index function parameters
+\label{sec:params}
+Some parameters must be fixed when using the index functions introduced in ...
 

### 2018-10-23 scagnostics.Rmd
Notes on index function parameters.

@@
     MIC(D) = \max_{xy<B(n)}\{M(D)_{x,y}\},
+    \label{eq:MIC}
     \end{equation}
@@
 \label{sec:params}
-Some parameters must be fixed when using the index functions introduced in ...
+Some parameters must be fixed when using the index functions considered here, controlling trade-offs between noise and fine structure detection, and between computing time and precision.
+
+- Binning:
+    - Scagnostics: the number of bins can be controlled by the user, note however that internally the implementation will reduce the number of bins if too many non-empty bins are found (more than 250).
+    - MINE: the maximum number of bins considered is fixed by the user as a function of the number of data points, see Equation \ref{eq:MIC}.
+The default is chosen such as a trade-off between resolution and noise dependence, but it may be tuned based on requirements dictated by specific datasets.
+Apart from sensitivity to noise computing time may also be a consideration here.
+- Spline knots: for the splines2D measure we need to fix the number of knots.
+By default it is fixed to be 10 (or lower if appropriate based on the data values).
+In our examples we find the number to be appropriate to identify functional dependence while rejecting noise, but some distributions may require tuning of this parameter.
 

### 2018-10-23 scagnostics.Rmd
Citation added.

@@
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
-\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see \cite{nason1992}, \cite{NIPS1997_1408}.}
+\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber1985,jones87}.}
 These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].

### 2018-10-23 scagnostics.Rmd
Make sure scales for rotation traces include 0, 1, finer binning, saved output.

@@
 \label{sec:rot}
-\textcolor{red}{testin rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
+\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
 Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binnibng (and outlier removal?) here.
@@
 ```{r rotationDep, fig.width=6, fig.height=7, fig.cap="Rotation dependence of the considered index functions tested on the sine 1000 dataset, by rotating the special view by an angle alpha."}
+if(!file.exists("cache/rotationDep.rda")){
   sineM <- as.matrix(select(sin1000, V5, V6))
@@
   )
-for (a in seq(0,2*pi, pi/20)){
+  for (a in seq(0,2*pi, pi/100)){
     rotM <- matrix(c(cos(a), sin(a), -sin(a), cos(a)), ncol = 2)
@@
   }
+  save(sc, file = "cache/rotationDep.rda")
+} else {
+  load("cache/rotationDep.rda")
+}
 
@@
   geom_line() + 
-  facet_wrap(~variable, scales = "free_y", ncol = 1)
+  facet_wrap(~variable, scales = "free_y", ncol = 1) +
+  geom_blank(aes(y = 0)) +
+  geom_blank(aes(y = 1))
 ```

### 2018-10-24 scagnostics.Rmd
Added short example of index parameter dependence.

@@
     return(mine(mat[,1], mat[,2])[[mineIndex]])
+  }
+}
+
+mineIndexAlpha <- function(mineIndex, alpha){
+  function(mat){
+    return(mine(mat[,1], mat[,2], alpha=alpha)[[mineIndex]])
   }
@@
 In our examples we find the number to be appropriate to identify functional dependence while rejecting noise, but some distributions may require tuning of this parameter.
+
+### Illustrative example
+A nice example can be constructed based on the RANDU dataset available in R.
+We use the MIC index and compare different parametrisations of the maximum binning by $\alpha$, where $B(n) = n^{\alpha}$.
+We compare in Figure \ref{fig:randuEx} the final view, index value and computing time obtained when optimising the MIC index with values $\alpha = 0.6, 0.7, 0.8$.
+Despite the low dimensionality, the small quint angle of the structure means that the identification of the maximum depends strongly on the starting conditions.
+Empirically we have found that using $\alpha = 0.7$ is most efficient in identifying the structure, while the default value $\alpha = 0.6$ is most likely to miss the structure given equal starting conditions.
+
+\textcolor{red}{Possible interpretation: 0.6 not allowing enough bins to accurately capture the structure, 0.7 good compromise, 0.8 becomes overly sensitive to noise}
+
+```{r randuEx, results="hide", fig.cap="Final view when optimising the MIC index in the RANDU dataset.
+Each plot lists the selected value of $\\alpha$, the time required by the optimisation in seconds and the MIC index value."}
+if(!file.exists("cache/randuEx.rda")){
+  pL <- list()
+  i <- 1
+  for (alpha in c( 0.6, 0.7, 0.8)){
+    set.seed(556677)
+    tic.clearlog()
+    tic() #start timer
+    tP <- save_history(randu, guided_tour(mineIndexAlpha("MIC", alpha)),max_bases = 100000)
+    toc(log=TRUE,quiet=TRUE)
+    optT <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
+    imax <- length(as.list(tP))
+    fV <- as.tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
+    colnames(fV) <- c("PP1", "PP2")
+    idxV <- mine(fV$PP1, fV$PP2, alpha = alpha)[["MIC"]]
+    pL[[i]] <- ggplot(fV, aes(PP1,PP2)) + geom_point() + theme(aspect.ratio=1) + 
+      ggtitle(paste(alpha, toString(format(optT, digits=2)), toString(format(idxV, digits=2)), sep = ", "))
+    i <- i+1
+  }
+  save(pL, file = "cache/randuEx.rda")
+} else {
+  load("cache/randuEx.rda")
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], ncol=3)
+```
 

### 2018-10-24 scagnostics.Rmd
Some editing.

@@
 on an interpolated path we expect to observe slow and consistent change in the index function.
-On the other hand if two similar projections produce very different index values, i.e.
-we observe sharp jumps when looking at the evolution of the index function, and in particular if those fluctuations dominate the overall evolution, the index is not suitable for projection pursuit.
-We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
-  - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
-Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified.
-Computation times for some benchmark scenarios are compared in Section \ref{sec:timer}.
-  - ability to find fine and broad structure: We consider a broad structure one that is roughly observable over a large angle around the optimal projection, it should in general be easy to identify and a good ppi should show smooth increase over a large number of projections as we are approaching the optimal view.
-On the other hand fine structures are difficult to identify as they are only visible close to the optimal view.
-This poses problems in particular for the optimization, but we may use index definitions to amplify larger index values while suppressing smaller ones, such that the optimization will be less sensitive to noise.
-\textcolor{red}{there are two different considerations, one is that described above, it relates to the concept of squint angle discussed in \cite{posse95b}.
-On the other hand should also consider detectin of "fine" structure vs noise, this would be controlled e.g.
+\textcolor{red}{maybe can relate this to consistency consition developed by P.
+Hall, but don't understand this at the moment.
+In addition this relates to questions of optimiser choice and efficiency.} We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
+    - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
+\textcolor{red}{I dont really understand this part.
+I think it can instead be related to the concept of squint angle.} We show examples in Section \ref{sec:dist}
+  - ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}.
+On the other hand should also consider detection of "fine" structure vs noise, this would be controlled e.g.
 by the allowed number of bins for the MINE index functions, additionally this will impact computing times, i.e.
 trade accuracy for efficiency}.
-We address these questions in Section \ref{sec:params}
-  - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
-\textcolor{red}{I dont really understand this part.} We show examples in Section \ref{sec:dist}
+We address these questions in Section \ref{sec:params} with an illustrative example.
   - rotation invariance: in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity.
@@
 https://arxiv.org/pdf/1805.12301.pdf)}.
-We illustrate the impact of non invariance in Section \ref{sec:rot}
+We illustrate the impact of non invariance in Section \ref{sec:rot}.
+\textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating.}
+    - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
+Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified.
+Computation times for some benchmark scenarios are compared in Section \ref{sec:timer}.
 
@@
 \textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
-truncated Gaussian distributions with approximately identity covariance matrix, such that they dont appear in any of the principal component planes.}
+truncated Gaussian distributions with approximately identity covariance matrix, such that they don't appear in any of the principal component planes.}
 
@@
 For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
+In the following we work with rescaled data where each direction has been centered and scaled to have variance equal to 1.
+This is neccessary for standard indices to be well defined, but does not impact the results when using the newly considered index functions.
 

### 2018-10-24 scagnostics.Rmd
More editing.

@@
 This is neccessary for standard indices to be well defined, but does not impact the results when using the newly considered index functions.
+\textcolor{red}{not yet implemented for traces of guided tour paths I think...}
 
@@
 
-## Smoothness
+## Index traces over sequence of interpolated projections
 \label{sec:smooth}
+### Random fluctuations
 We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto V1-V2 to the projection onto V3-V4 and using the default interpolation angle of 0.05.
@@
 The results are shown in Figure \ref{fig:plotV1V2toV3V4}.
-The standard index functions "holes" and "cmass" show the desired behaviour in that the change is smooth over interpolated views ((it is however curious to see the actual evolution, which seems not to depend on the underlying random distributions, but is clearely evolving over the path of the planned tour)).
+The standard index functions "holes" and "cmass" show the desired behaviour in that the change is smooth over interpolated views.
 Considering now the indices from the scagnostics family, we first note stronge dependence on the sample size.
@@
 
-## Transition
-We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6, which will also allow us to view the picture between typical low and high values of the index function.
+### Moving into informative projections
+We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6, which will also allow us to view the evolution between typical low and high values of the index function.
 Results are shown in Figure \ref{fig:plannedtour}.
@@
 On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
-Note that the various indices appear less noisy than for the grand tour shown before, but should note that the tour path here is shorter, and the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
-
-TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full information of the characteristics is used, implying that more information is considered.
-We also see that TIC is smoother when moving into the special view.
-
-Here we also see how the existing indices in the tourr package are inadequate for the detection of these type of dependence, the holes index taking maximum values before the final view and decreasing when moving into the special view.
-On the other hand, surprisingly the cmass index increases when moving into the special view, but the intermediate minimum makes it unsuitable as a projection pursuit index here.
+Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
+
+TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
+We also see that TIC is smoother when moving into the special view, in terms of the relative differences.
+
+\textcolor{red}{Holes index or even pca could discover pipe as well...}
 
@@
 \textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
-Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binnibng (and outlier removal?) here.
+Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binning (and outlier removal?) here.
 Observe huge drops for some of the indexes!

### 2018-10-24 scagnostics.Rmd
Remark about spiral.

@@
 \textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
-truncated Gaussian distributions with approximately identity covariance matrix, such that they don't appear in any of the principal component planes.}
+truncated Gaussian distributions with approximately identity covariance matrix, such that they don't appear in any of the principal component planes.
+In particular seems that spiral is hardest to identify based on results from Posse, maybe I can add this to my sphere dataset?
+or instead of it?}
 

### 2018-10-25 scagnostics.Rmd
testing my small changes merge

--- 2018-10-24 scagnostics.Rmd
@@
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
-\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber1985,jones87}.}
+\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see \cite{nason1992}, \cite{NIPS1997_1408}.}
 These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
@@
 -->
-A projection pursuit index (ppi) is a scalar function $f$ defined on an $n$-dimensional data distribution.
+A projection pursuit index (PPI) is a scalar function $f$ defined on an $n$-dimensional data distribution.
 Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $n$ dimensional projections of a data set with $p>n$ parameters will find the most interesting projections.
@@
 
-\textcolor{red}{\cite{jones87} discusses centering and sphering of the data points.
-Centering is needed if index is not translational invariant (should not be the case for any index considered here).
-Arguments for sphering are that scale effects should be extracted, we don't want to reproduce information already obtained by PCA.
-To my mind rescaling to an overall intervall is more appropriate here.
-Rotations between parameters should not affect the outcome, but will make interpretation more difficult.
-Standardising the variance will remove information about the posterior variance.
-Overall we are interested in strong association only, if these structures were affected by the standardisation procedure, it would most likely be unwelcome effects?}
-
 ## Optimization
-Given a ppi we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
+Given a PPI we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
 One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
@@
 In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
-
-\textcolor{red}{\cite{posse95b} discusses the optimisation, in particular that for most index functions and optimisers results are too local, largely dependent on starting point.
-Fiedmans suggestion of hybrid steppind search + steepest-ascent still remains fairly local.
-\cite{posse90} suggested random search instead, which seems to perform better, but did not find pdf of that article, so not so sure what he actually does.}
 
@@
     MIC(D) = \max_{xy<B(n)}\{M(D)_{x,y}\},
-    \label{eq:MIC}
     \end{equation}
@@
 \label{sec:investigate}
-We now examine the indices listed in Section \ref{sec:indexDef} to test which of them are suitable as projection pursuit measures, and how we may modify them for improved performance.
-We start by considering the relevant criteria:
-
-\textcolor{red}{\cite{posse95b} lists five criteria: minimized by normal distribution, approximately affine invariant, consistent, resistent to features in the tail, fast computation.
-Comment that different index values for the same plane will make optimisation less efficient.
-\cite{jones87} moreover lists rotation and translation invariance (independence of choice of basis in the plane), and also mentions conditions for numeric optimisation, i.e.
-continuous and depending on optimisation methods also differentiable.
-Interesting: according to \cite{jones87} the original index proposed by Friedman and Tukey did not have property of rotation invariance.}
+
+There are a number of properties that a PPI should have, which are:
 
@@
 on an interpolated path we expect to observe slow and consistent change in the index function.
-\textcolor{red}{maybe can relate this to consistency consition developed by P.
-Hall, but don't understand this at the moment.
-In addition this relates to questions of optimiser choice and efficiency.} We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
+On the other hand if two similar projections produce very different index values, i.e.
+we observe sharp jumps when looking at the evolution of the index function, and in particular if those fluctuations dominate the overall evolution, the index is not suitable for projection pursuit.
+  - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
+Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified.
+  - ability to find fine and broad structure: We consider a broad structure one that is roughly observable over a large angle around the optimal projection, it should in general be easy to identify and a good PPI should show smooth increase over a large number of projections as we are approaching the optimal view.
+On the other hand fine structures are difficult to identify as they are only visible close to the optimal view.
+This poses problems in particular for the optimization, but we may use index definitions to amplify larger index values while suppressing smaller ones, such that the optimization will be less sensitive to noise.
   - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
-\textcolor{red}{I dont really understand this part.
-I think it can instead be related to the concept of squint angle.} We show examples in Section \ref{sec:dist}
-  - ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}.
-On the other hand should also consider detection of "fine" structure vs noise, this would be controlled e.g.
-by the allowed number of bins for the MINE index functions, additionally this will impact computing times, i.e.
-trade accuracy for efficiency}.
-We address these questions in Section \ref{sec:params} with an illustrative example.
   - rotation invariance: in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity.
@@
 By costruction this is not the case for most index functions defined above.
-\textcolor{red}{One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case.
-Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g.
+(One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case.) (Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g.
 https://doi.org/10.1016/S0146-664X(75)80008-6, other example is interpolating over angles when building NN, see e.g.
-https://arxiv.org/pdf/1805.12301.pdf)}.
-We illustrate the impact of non invariance in Section \ref{sec:rot}.
-\textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating.}
-    - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
-Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified.
-Computation times for some benchmark scenarios are compared in Section \ref{sec:timer}.
-
-##Simulation study
-
-\textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
-truncated Gaussian distributions with approximately identity covariance matrix, such that they don't appear in any of the principal component planes.}
-
-To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
-They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
-
-1. $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
-2.
-$p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-3.
-$p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-
-For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
-In the following we work with rescaled data where each direction has been centered and scaled to have variance equal to 1.
-This is neccessary for standard indices to be well defined, but does not impact the results when using the newly considered index functions.
+https://arxiv.org/pdf/1805.12301.pdf)
+
+The new indexes are examined with respect to these properties, and where they are less than adequate can suggest ways to improve them.
+
+##Simulated data
+
+Three data simulations are used for examining the behaviour of the index functions. Each generates special structure in two variables, and the remaining variables are noise.
+
+- structure to be found is 2D, scagnostics could find them, point is PP would find if in a combination of variables
+- structure is 2D, noise variables are different.
 
@@
 ###Dataset overview
+
+we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
+They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
+
+- sphere:  $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
+- pipe: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+- sine:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
+
+For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
 
@@
     return(mine(mat[,1], mat[,2])[[mineIndex]])
-  }
-}
-
-mineIndexAlpha <- function(mineIndex, alpha){
-  function(mat){
-    return(mine(mat[,1], mat[,2], alpha=alpha)[[mineIndex]])
   }
@@
 ## Smoothness
-\label{sec:smooth}
 We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto V1-V2 to the projection onto V3-V4 and using the default interpolation angle of 0.05.
@@
 ## Distributions of index values
-\label{sec:dist}
+
 Compare typical index values over the two tour paths.
@@
 ## Rotation dependence
-\label{sec:rot}
-\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
+\textcolor{red}{testin rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
 Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binnibng (and outlier removal?) here.
@@
 ```{r rotationDep, fig.width=6, fig.height=7, fig.cap="Rotation dependence of the considered index functions tested on the sine 1000 dataset, by rotating the special view by an angle alpha."}
-if(!file.exists("cache/rotationDep.rda")){
 sineM <- as.matrix(select(sin1000, V5, V6))
@@
 )
-  for (a in seq(0,2*pi, pi/100)){
+for (a in seq(0,2*pi, pi/20)){
   rotM <- matrix(c(cos(a), sin(a), -sin(a), cos(a)), ncol = 2)
@@
   }
-  save(sc, file = "cache/rotationDep.rda")
-} else {
-  load("cache/rotationDep.rda")
-}
 
@@
   geom_line() + 
-  facet_wrap(~variable, scales = "free_y", ncol = 1) +
-  geom_blank(aes(y = 0)) +
-  geom_blank(aes(y = 1))
+  facet_wrap(~variable, scales = "free_y", ncol = 1)
 ```
@@
 ## Note on calculation time
-\label{sec:timer}
 Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example the 6-d sphere and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
@@
 Finally for the MINE family the computing time increases exponentially as can be expected from the dependence on the sample size, where we see that the improved $MIC_e$ algorithm is somewhat faster than the original implementation, see also @JMLRv1715308.
-
-\textcolor{red}{The plot is not perfect.
-I use alpha to fade out points, but cant find value that gives nice picture here.
-In addition the timer isnt precise enough to resolve timing for some indexes on small datasets, they are zeroes, appear on bottom of the log10 y axis.}
 
@@
 
-timerMeans <- dfTimer %>%
-  group_by(name, n) %>%
-  summarise(t=mean(t))
-
+#timerMelt <- melt(dfTimer, id="n")
 ggplot(dfTimer, aes(x = n, y = t, color=name)) + 
-  geom_point(alpha=0.05) +
-  geom_point(data=timerMeans) +
-  geom_line(data=timerMeans) +
+  geom_point() +
   scale_x_log10() +
@@
   labs(x = "Sample Size", y = "Time [s]", color = "Index family")
-```
-
-## Index function parameters
-\label{sec:params}
-Some parameters must be fixed when using the index functions considered here, controlling trade-offs between noise and fine structure detection, and between computing time and precision.
-
-- Binning:
-    - Scagnostics: the number of bins can be controlled by the user, note however that internally the implementation will reduce the number of bins if too many non-empty bins are found (more than 250).
-    - MINE: the maximum number of bins considered is fixed by the user as a function of the number of data points, see Equation \ref{eq:MIC}.
-The default is chosen such as a trade-off between resolution and noise dependence, but it may be tuned based on requirements dictated by specific datasets.
-Apart from sensitivity to noise computing time may also be a consideration here.
-- Spline knots: for the splines2D measure we need to fix the number of knots.
-By default it is fixed to be 10 (or lower if appropriate based on the data values).
-In our examples we find the number to be appropriate to identify functional dependence while rejecting noise, but some distributions may require tuning of this parameter.
-
-### Illustrative example
-A nice example can be constructed based on the RANDU dataset available in R.
-We use the MIC index and compare different parametrisations of the maximum binning by $\alpha$, where $B(n) = n^{\alpha}$.
-We compare in Figure \ref{fig:randuEx} the final view, index value and computing time obtained when optimising the MIC index with values $\alpha = 0.6, 0.7, 0.8$.
-Despite the low dimensionality, the small quint angle of the structure means that the identification of the maximum depends strongly on the starting conditions.
-Empirically we have found that using $\alpha = 0.7$ is most efficient in identifying the structure, while the default value $\alpha = 0.6$ is most likely to miss the structure given equal starting conditions.
-
-\textcolor{red}{Possible interpretation: 0.6 not allowing enough bins to accurately capture the structure, 0.7 good compromise, 0.8 becomes overly sensitive to noise}
-
-```{r randuEx, results="hide", fig.cap="Final view when optimising the MIC index in the RANDU dataset.
-Each plot lists the selected value of $\\alpha$, the time required by the optimisation in seconds and the MIC index value."}
-if(!file.exists("cache/randuEx.rda")){
-  pL <- list()
-  i <- 1
-  for (alpha in c( 0.6, 0.7, 0.8)){
-    set.seed(556677)
-    tic.clearlog()
-    tic() #start timer
-    tP <- save_history(randu, guided_tour(mineIndexAlpha("MIC", alpha)),max_bases = 100000)
-    toc(log=TRUE,quiet=TRUE)
-    optT <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
-    imax <- length(as.list(tP))
-    fV <- as.tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
-    colnames(fV) <- c("PP1", "PP2")
-    idxV <- mine(fV$PP1, fV$PP2, alpha = alpha)[["MIC"]]
-    pL[[i]] <- ggplot(fV, aes(PP1,PP2)) + geom_point() + theme(aspect.ratio=1) + 
-      ggtitle(paste(alpha, toString(format(optT, digits=2)), toString(format(idxV, digits=2)), sep = ", "))
-    i <- i+1
-  }
-  save(pL, file = "cache/randuEx.rda")
-} else {
-  load("cache/randuEx.rda")
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], ncol=3)
 ```
@@
   add_column(label=colnames(bbhM2))
-iv <- mine(dProj$PP1,dProj$PP2)["TIC"]
 p1 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
@@
   geom_segment(data=getAxes(fProj1,1, 0, 0.34, 0.25), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj1, 1, 0, 0.34, 0.25), aes(x=x2, y=y2, label=label), size=2.5) +
-  ggtitle(paste0("a, TIC: ", toString(format(iv, digits=2))))
+  geom_text(data=getAxes(fProj1, 1, 0, 0.34, 0.25), aes(x=x2, y=y2, label=label), size=2.5)
 iLast <- length(bbhTourFull22)
@@
   add_column(label=colnames(bbhM2))
-iv <- splines2d(dProj$PP1,dProj$PP2)
 p2 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
@@
   geom_segment(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x2, y=y2, label=label), size=2.5) +
-  ggtitle(paste0("b, splines2D: ", toString(format(iv, digits=2))))
+  geom_text(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
 iLast <- length(bbhTourFull32)
@@
   add_column(label=colnames(bbhM2))
-iv <- 1-scagnostics(as.matrix(dProj))[,"Convex"]
 p3 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
@@
   geom_segment(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x2, y=y2, label=label), size=2.5) +
-  ggtitle(paste0("c, 1-convex: ", toString(format(iv, digits=2))))
+  geom_text(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x2, y=y2, label=label), size=2.5)
 grid.arrange(p1, p2, p3, ncol=3) 

### 2018-10-25 scagnostics.Rmd

--- 2018-10-24 scagnostics.Rmd
@@
 ## Optimization
-Given a ppi we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
+Given a PPI we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
 One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
@@
 \label{sec:investigate}
-We now examine the indices listed in Section \ref{sec:indexDef} to test which of them are suitable as projection pursuit measures, and how we may modify them for improved performance.
-We start by considering the relevant criteria:
+
+There are a number of properties that a PPI should have, which are:
 
@@
 
-1. $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
-2.
-$p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-3.
-$p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-
-For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
-In the following we work with rescaled data where each direction has been centered and scaled to have variance equal to 1.
-This is neccessary for standard indices to be well defined, but does not impact the results when using the newly considered index functions.
-\textcolor{red}{not yet implemented for traces of guided tour paths I think...}
+Three data simulations are used for examining the behaviour of the index functions. Each generates special structure in two variables, and the remaining variables are noise.
+
+- structure to be found is 2D, scagnostics could find them, point is PP would find if in a combination of variables
+- structure is 2D, noise variables are different.
 

### 2018-10-25 scagnostics.Rmd
Reference added.

--- 2018-10-24 scagnostics.Rmd
@@
 \textcolor{red}{I dont really understand this part.
-I think it can instead be related to the concept of squint angle.} We show examples in Section \ref{sec:dist}
+I think it can instead be related to the concept of squint angle \cite{barnett1981interpreting}.} We show examples in Section \ref{sec:dist}
   - ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}.

### 2018-10-25 scagnostics.Rmd
Function for random points on a spiral.

@@
 }
+
+spiralData <- function(n, p){
+  i <- 1
+  a <- 0.1
+  b <- 0.1
+  dRet <- NULL
+  while(i <= p){
+    v <- rnorm(n-2)
+    theta <- abs(rnorm(1,0,2*pi))
+    r <- a + b * theta
+    x <- r * cos(theta)
+    y <- r * sin(theta)
+    v <- c(v, x, y)
+    dRet <- rbind(dRet, v)
+    i <- i+1
+  }
+  return(dRet)
+}
+
 ```
@@
 \label{sec:smooth}
+
 ### Random fluctuations

### 2018-10-26 scagnostics.Rmd
Some more organisation.

@@
 By costruction this is not the case for most index functions defined above.
-\textcolor{red}{One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case.
-Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g.
-https://doi.org/10.1016/S0146-664X(75)80008-6, other example is interpolating over angles when building NN, see e.g.
-https://arxiv.org/pdf/1805.12301.pdf)}.
 We illustrate the impact of non invariance in Section \ref{sec:rot}.
-\textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating.}
+\textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating, reamining issue would be optimisation efficiency.}
   - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
@@
 ##Simulation study
+
+Three data simulations are used for examining the behaviour of the index functions.
+Each generates special structure in two variables, and the remaining variables are noise.
+This construction means that variable selection would be sufficient for the identification of the structure.
+We use this setup for simplicity, and since all variable combinations are treated equally in the PP it does not affect our conclusions about the ability of identifying the structure in arbitrary linear combinations of variables.
+
+To understand the potential for the various indices defined we use the following overview procedures on the simulated data sets:
+
+- index values for pair of variables, explicitly comparing the uninformative and special view of each dataset, as well as discussing the dependence on the sample size, see Section \ref{sec:dataOv}
+- tour between pairs of variables, considering a planned tour that moves from the $V_1 - V_2$ view either to the $V_3 - V_4$ view (no special structure encountered over the full path) or to the $V_5 - V_6$ (moving from uninformed to special view), see Section \ref{sec:smooth}
+- optimisation path, starting from $V_1 - V_2$, how can we use the available index functions and optimisation algorithms to identify the special view, see Section \ref{sec:guided}
+- rotating a dataset with functional dependence to evaluate the rotation dependence of the index functions, see Section \ref{sec:rot}
+- benchmarking the computing time in Section \ref{sec:timer}
 
@@
 
+###Dataset overview
+\label{sec:dataOv}
+
 To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
-They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
-
-Three data simulations are used for examining the behaviour of the index functions. Each generates special structure in two variables, and the remaining variables are noise.
-
-- structure to be found is 2D, scagnostics could find them, point is PP would find if in a combination of variables
-- structure is 2D, noise variables are different.
-
-To understand the potential for the various indices defined we use the following overview procedures on the simulated data sets:
-
-- index values for pair of variables, explicitly comparing the uninformative and special view of each dataset, as well as discussing the dependence on the sample size
-- tour between pairs of variables, considering a planned tour that moves from the $x_1 - x_2$ view either to the $x_3 - x_4$ view (no special structure encountered over the full path) or to the $x_5 - x_6$ (moving from uninformed to special view)
-- optimisation path, starting from $x_1 - x_2$, how can we use the available index functions and optimisation algorithms to identify the special view
-
-###Dataset overview
-
-we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
 They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
@@
 ## Guided tour
+\label{sec:guided}
 

### 2018-10-26 scagnostics.Rmd
start on section 3

@@
 
-There are a number of properties that a PPI should have, which are:
-
-\textcolor{red}{\cite{posse95b} lists five criteria: minimized by normal distribution, approximately affine invariant, consistent, resistent to features in the tail, fast computation.
-Comment that different index values for the same plane will make optimisation less efficient.
-\cite{jones87} moreover lists rotation and translation invariance (independence of choice of basis in the plane), and also mentions conditions for numeric optimisation, i.e.
-continuous and depending on optimisation methods also differentiable.
-Interesting: according to \cite{jones87} the original index proposed by Friedman and Tukey did not have property of rotation invariance.}
+A useful projection pursuit index needs to have several properties.
+This has been discussed in several seminal papers (\cite{diaconis84}, \cite{huber85}, \cite{jones87}, \cite{posse95b}).
+The PPI should be  minimized by the normal distribution, because this is not interesting from a data exploration perspective.
+If all projections are normally distributed, good modeling tools already exist.
+A PPI should be approximately affine invariant, regardless how the projection is rotated the index value should be the same, and the scale of each variable shouldn't affect the index value.
+Interestingly the original index proposed by \cite{FT74} was not rotationally invariant.
+A consistent index means that small perturbations to the sample do not dramatically change the index value.
+This is particularly important to making optimisation feasible, small angles between projections correspond to small perturbations of the sample, and thus shold be small changes to index value.
+\cite{posse95b} suggests that indexes should be resistant to features in a tail of the distribution, but this is debatable because one departure from normality that is interesting to detect are anomalous observations.
+Some PPI are designed precisely for these reasons.
+Lastly, because we need to compute the PPI over many projections, it needs to be fast to compute.
+These form the basis of the crteria upon which the scagnostic indexes, and the several alternative indexes are examined:
 
@@
 
-##Simulation study
+## Simulation study
 
@@
 
-###Dataset overview
+### Dataset overview
 \label{sec:dataOv}

### 2018-10-26 scagnostics.Rmd
Summary table added.

--- 2018-10-26 scagnostics.Rmd
@@
 ```
+
+## Summary
+
+Our results can be summarised by comparing
+
+- scagnostics indices:
+    - advantages: scaling with larger datasets thanks to binning, set of index functions covering different types of interesting structures
+    - disadvantage: noisy, requires good optimisation algorithm, stringy too noisy to be used as index
+- splines2D:
+    - advantages: fast to compute, also considering scaling with larger data sets, efficient in detecting functional relations, typically smooth behaviour (but we have seen some exceptions)
+    - disadvantage: not general, highly not rotation invariant
+- dcor2D:
+    - advantages: smoothly changing, sensitive to general structure
+    - disadvantage: slow if computed without binning, large sensitivity to noise compared to MIC/TIC (example: low values obtained for the circle view of the pipe), sensitive to rotation
+    - NOTE: I have not tested this much, should probably add binned version to timing plot
+- MINE (MIC, TIC):
+    - advantages: general, good discrimination even with noisy structures
+    - disadvantages: slow to compute, noisy for small datasets, TIC sensitive to rotation
+
+\begin{table}
+\begin{center}
+\caption{Summary}
+\label{tab:summary}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Index & smooth & rotation invariant & fast & general \\
+\hline
+Convex & ! & ! & \checkmark & ! \\
+Skinny & ! & ! & \checkmark & ! \\
+Stringy & $\times$ & ! & \checkmark & ! \\
+\hline
+splines2D & \checkmark & $\times$ & \checkmark & ! \\
+dcor2D & \checkmark & $\times$ & ! & ! \\
+\hline
+MINE & ! & ! & ! & \checkmark\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2018-10-26 scagnostics.Rmd
Updated text accordingly.

--- 2018-10-26 scagnostics.Rmd
@@
 ```
+
+## Summary
+
+Our results can be summarised by evaluating and comparing the advantages and disadvanteges of each index function according to the criteria presented above.
+An overview is given in Table \ref{tab:summary}.
+The first three entries are from the scagnostics family.
+The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures (the same is however not true for each single index viewed on its own).
+On the other hand we have found that they are typically too noisy making optimisation of the function challenging.
+In particular the stringy index cannot be used with current optimisers.
+The index functions from @mbgraphic share similar behaviour as they are both found to evolve fairly smoothly between interpolated projections, but have been shown to be strongly rotation dependent.
+The splines2D index is fast to calculate and found to be the most efficient method for identifying functional dependence.
+We have only considered the non-binned implementation of dcor2D which is slow to calculate on large datasets.
+While this index is in principle general, we have found it to be too sensitive to noise to be applicable in most realistic examples.
+Finally the MINE indexes are found to be smooth only for large datasets, with some rotation dependence observed for the TIC index.
+While these indexes are slow to compute especially on large datasets, we have found them to be the most general, allowing good discrimination even with noisy structures.
+
+\begin{table}
+\begin{center}
+\caption{Summary}
+\label{tab:summary}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Index & smooth & rotation invariant & fast & general \\
+\hline
+Convex & ! & ! & \checkmark & ! \\
+Skinny & ! & ! & \checkmark & ! \\
+Stringy & $\times$ & ! & \checkmark & ! \\
+\hline
+splines2D & \checkmark & $\times$ & \checkmark & ! \\
+dcor2D & \checkmark & $\times$ & ! & ! \\
+\hline
+MINE & ! & ! & ! & \checkmark\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2018-10-26 scagnostics.Rmd
done with criteria description

--- 2018-10-26 scagnostics.Rmd
@@
 keywords:
-- tour
-- projection pursuit 
 - statistical graphics
@@
 - exploratory data analysis
-- high-dimensional data
+- data science 
+- guided tour
 
@@
 Less work has been done on indexes to find nonlinear dependence between variables.
+
+XXX Need to add something here about the guided tour
 
@@
 A useful projection pursuit index needs to have several properties.
-This has been discussed in several seminal papers (\cite{diaconis84}, \cite{huber85}, \cite{jones87}, \cite{posse95b}).
+This has been discussed in several seminal papers (\cite{diaconis84}, \cite{huber85}, \cite{jones87}, \cite{posse95b}, \cite{hall89}).
 The PPI should be  minimized by the normal distribution, because this is not interesting from a data exploration perspective.
@@
 Lastly, because we need to compute the PPI over many projections, it needs to be fast to compute.
-These form the basis of the crteria upon which the scagnostic indexes, and the several alternative indexes are examined:
-
-  - smoothness: the index function should evolve smoothly between nearby projections, i.e.
+These form the basis of the crteria upon which the scagnostic indexes, and the several alternative indexes are examined, as explained below.
+
+- **smoothness**: This is the consistency property mentioned above.
+The index function values are examined over interpolated tour paths, that is, the value is plotted against time, where time indexes the sequence of projections produced by the tour path.
+The signature of a good PPI is that the plotted function is relatively smooth.
+The interpolation path corresponds to small angle changes between projections, so the value should be very similar.
+<!-- 
+  should evolve smoothly between nearby projections, i.e.
 on an interpolated path we expect to observe slow and consistent change in the index function.
@@
 In addition this relates to questions of optimiser choice and efficiency.} We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
-  - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
+-->
+- **squintability**: \cite{barnett1981interpreting} introduced the idea of squint angle to indicate resolution of structure in the data.
+Fine structure like the parallel planes in the infamous RANDU data \ref{randu} has a small squint angle because you have to be very close the the optimal projection plane to be able to see the structure.
+Structures with small squint angle are difficult to find, because the optmisation algorithm needs to get very close to begin hill-climbing to the optimum.
+The analyst doesn't have control over the data structure, but does have control over the PPI.
+Squintability is about the shape of the PPI over all projections.
+It should have smooth low values for noise projections and a clearly larger value, ideally with a big squint angle, for structured projections.
+The optimiser should be able to clearly see the optimal projections as different from noise.
+To examine squintability, the PPI values are examined on interpolated tour paths between a noise projection and a distant structured projection.
+<!--
+distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
 \textcolor{red}{I dont really understand this part.
 I think it can instead be related to the concept of squint angle \cite{barnett1981interpreting}.} We show examples in Section \ref{sec:dist}
-  - ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}.
+-->
+- **flexibility**: An analyst can have a toolbox of indices that may cover the range of fine and broad structure, which underlies the scagnostics suite.
+Early indexes, based on density estimation could be programmed to detect fine or large structure by varying the binwidth.
+This is examined by using a range of structure in the simulated data examples.
+<!--
+ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}.
 On the other hand should also consider detection of "fine" structure vs noise, this would be controlled e.g.
@@
 trade accuracy for efficiency}.
-We address these questions in Section \ref{sec:params} with an illustrative example.
-  - rotation invariance: in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity.
+We address these questions in Section \ref{sec:params} with an illustrative example.-->
+- **rotation invariance**: The orientation of structure within a projection plane should not change the index value.
+This is especially important when using the projection pursuit guided tour, because the tour path is defined between planes, along a geodesic path, not bases within planes.
+If a particular orientation is more optimal, this will get lost as the projection shown pays no attention to orientation.
+\cite{bujaXXX} describes alternative interpolation paths based on Givens and Householder rotations which progress from basis to basis.
+It may be possible to ignore rotation invariance with these interpolations but there isn't a current implementation, primarily because the within-plane spin that is generated is distracting from a visualisation perspective.
+Rotation invariance is checked for the proposed PPIs by rotating the structured projection, within the plane.
+<!--
+in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity.
 A projection pursuit index should therefore be invariant under rotation of the basis in the projection plane.
-By costruction this is not the case for most index functions defined above.
+By construction this is not the case for most index functions defined above.
 We illustrate the impact of non invariance in Section \ref{sec:rot}.
 \textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating, reamining issue would be optimisation efficiency.}
-  - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time.
-Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified.
-Computation times for some benchmark scenarios are compared in Section \ref{sec:timer}.
+-->
+- **speed**: Being fast to compute allows the index to be used in real-time in a guided tour, where the optimisation can be watched.
+When the computations are shifted off-line, to watch in replay, computation times matter less.
+This is checked  by comparing times for benchmark scenarios with varying sample size and dimension.
 
@@
 ```
-
-## Summary
-
-Our results can be summarised by comparing
-
-- scagnostics indices:
-    - advantages: scaling with larger datasets thanks to binning, set of index functions covering different types of interesting structures
-    - disadvantage: noisy, requires good optimisation algorithm, stringy too noisy to be used as index
-- splines2D:
-    - advantages: fast to compute, also considering scaling with larger data sets, efficient in detecting functional relations, typically smooth behaviour (but we have seen some exceptions)
-    - disadvantage: not general, highly not rotation invariant
-- dcor2D:
-    - advantages: smoothly changing, sensitive to general structure
-    - disadvantage: slow if computed without binning, large sensitivity to noise compared to MIC/TIC (example: low values obtained for the circle view of the pipe), sensitive to rotation
-    - NOTE: I have not tested this much, should probably add binned version to timing plot
-- MINE (MIC, TIC):
-    - advantages: general, good discrimination even with noisy structures
-    - disadvantages: slow to compute, noisy for small datasets, TIC sensitive to rotation
-
-\begin{table}
-\begin{center}
-\caption{Summary}
-\label{tab:summary}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Index & smooth & rotation invariant & fast & general \\
-\hline
-Convex & ! & ! & \checkmark & ! \\
-Skinny & ! & ! & \checkmark & ! \\
-Stringy & $\times$ & ! & \checkmark & ! \\
-\hline
-splines2D & \checkmark & $\times$ & \checkmark & ! \\
-dcor2D & \checkmark & $\times$ & ! & ! \\
-\hline
-MINE & ! & ! & ! & \checkmark\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
 

### 2018-10-26 scagnostics.Rmd

@@
 
+## Summary
+
+Our results can be summarised by evaluating and comparing the advantages and disadvanteges of each index function according to the criteria presented above.
+An overview is given in Table \ref{tab:summary}.
+The first three entries are from the scagnostics family.
+The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures (the same is however not true for each single index viewed on its own).
+On the other hand we have found that they are typically too noisy making optimisation of the function challenging.
+In particular the stringy index cannot be used with current optimisers.
+The index functions from @mbgraphic share similar behaviour as they are both found to evolve fairly smoothly between interpolated projections, but have been shown to be strongly rotation dependent.
+The splines2D index is fast to calculate and found to be the most efficient method for identifying functional dependence.
+We have only considered the non-binned implementation of dcor2D which is slow to calculate on large datasets.
+While this index is in principle general, we have found it to be too sensitive to noise to be applicable in most realistic examples.
+Finally the MINE indexes are found to be smooth only for large datasets, with some rotation dependence observed for the TIC index.
+While these indexes are slow to compute especially on large datasets, we have found them to be the most general, allowing good discrimination even with noisy structures.
+
+\begin{table}
+\begin{center}
+\caption{Summary}
+\label{tab:summary}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Index & smooth & rotation invariant & fast & general \\
+\hline
+Convex & ! & ! & \checkmark & ! \\
+Skinny & ! & ! & \checkmark & ! \\
+Stringy & $\times$ & ! & \checkmark & ! \\
+\hline
+splines2D & \checkmark & $\times$ & \checkmark & ! \\
+dcor2D & \checkmark & $\times$ & ! & ! \\
+\hline
+MINE & ! & ! & ! & \checkmark\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 # Application to physics data

### 2018-10-26 scagnostics.Rmd
typo

@@
 - **squintability**: \cite{barnett1981interpreting} introduced the idea of squint angle to indicate resolution of structure in the data.
-Fine structure like the parallel planes in the infamous RANDU data \ref{randu} has a small squint angle because you have to be very close the the optimal projection plane to be able to see the structure.
+Fine structure like the parallel planes in the infamous RANDU data \ref{randu} has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
 Structures with small squint angle are difficult to find, because the optmisation algorithm needs to get very close to begin hill-climbing to the optimum.

### 2018-10-26 scagnostics.Rmd
Replacing sphere with spiral.

@@
 set.seed(1984)
-sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
-sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()
+spiral100 <- spiralData(6, 100) %>% scale() %>% as_tibble()
+spiral1000 <- spiralData(6, 1000) %>% scale() %>% as_tibble()
+#sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
+#sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()
 
@@
 i <- 1
-pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
   pC <- ggplot(ds, aes(V1, V2)) +
@@
 i <- 1
-pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
   pC <- ggplot(ds, aes(V5, V6)) +
@@
 i <- 1
-pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
+pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
   distData <- melt(ds)
@@
 i <- 1
-pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
 indexOverview <- tibble(convex=numeric(),
@@
                vars=character())
-dsL <- list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)
+dsL <- list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)
 for(ds in dsL){
@@
     rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
-    rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
-    rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
+    rbind(getProj(spiral100, t1full, "spiral", 100)) %>%
+    rbind(getProj(spiral1000, t1full, "spiral", 1000)) %>%
     rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
@@
     rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
-    rbind(getProj(sphere100, t2full, "sphere", 100)) %>%
-    rbind(getProj(sphere1000, t2full, "sphere", 1000)) %>%
+    rbind(getProj(spiral100, t2full, "spiral", 100)) %>%
+    rbind(getProj(spiral1000, t2full, "spiral", 1000)) %>%
     rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
@@
   set.seed(2018)
-  grandTour100 <- save_history(sphere100, grand_tour(2), max=4) %>%
+  grandTour100 <- save_history(sphereData(6,100), grand_tour(2), max=4) %>%
     interpolate() %>%

### 2018-10-26 scagnostics.Rmd
start on data section

--- 2018-10-26 scagnostics.Rmd
@@
 
-Three data simulations are used for examining the behaviour of the index functions.
-Each generates special structure in two variables, and the remaining variables are noise.
-This construction means that variable selection would be sufficient for the identification of the structure.
-We use this setup for simplicity, and since all variable combinations are treated equally in the PP it does not affect our conclusions about the ability of identifying the structure in arbitrary linear combinations of variables.
+### Data construction
+\label{sec:dataOv}
+
+Three families of data simulations are used for examining the behaviour of the index functions.
+Each generates  structure in two variables, with the remaining variables containing various types of noise.
+This is a very simple construction, because there is no need for projection pursuit to find the structure, one could simply use the scagnostics on pairs of variables.
+However, it serves the purpose to also evaluate the PPIs.
+The three data families are:
+
+- **pipe**: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+- **sine**:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
+- **sphere**:  $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
+
+Two samples sizes are also used: $p=100$ (small) and $p=1000$ (large).
+
+### Property assessment
 

### 2018-10-29 scagnostics.Rmd
Summary table updated.
Some comments.

@@
 \textcolor{red}{What is upper bound of TIC as function of sample size?
-If we define $n' = n^{0.6}$, rounded to integer values only, can we calculate it as $TIC_{max} = 2 \sum_{i=1}^{n'} \frac{n'}{i}$, keeping in mind that each entry of the characteristic matrix should be between [0,1].
-Each entry in the sum should be rounded also, only integer number of bins enter actual calculation, so this should be approximation only.}
+It seems from checking by putting in identical parameters as x, y that max(TIC) = 16 for 100 datapoints and max(TIC) = 148 for 1000 datapoints.
+Don't know how to derive this though, should depend also on c parameter and how the algorithm is approximating the true matrix?.}
 
@@
 
+\textcolor{red}{Maybe it will be better to combine the two traces? Can have tour from V1-V2 to V3-V4 to V5-V6, indicate position of V3-V4 in path, will save space, make direct comparison easier.}
+
 ### Random fluctuations
@@
 \label{tab:summary}
-\begin{tabular}{|l|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|}
 \hline
-Index & smooth & rotation invariant & fast & general \\
+Index & smooth & squintability & flexible & rotation invariant & speed \\
 \hline
-Convex & ! & ! & \checkmark & ! \\
-Skinny & ! & ! & \checkmark & ! \\
-Stringy & $\times$ & ! & \checkmark & ! \\
+Convex & ! & ! & ! & ! & \checkmark \\
+Skinny & ! & !& ! & ! & \checkmark \\
+Stringy & $\times$ & $\times$ & ! & ! & \checkmark \\
 \hline
-splines2D & \checkmark & $\times$ & \checkmark & ! \\
-dcor2D & \checkmark & $\times$ & ! & ! \\
+splines2D & \checkmark & \checkmark & ! & $\times$ & \checkmark \\
+dcor2D & \checkmark & \checkmark & ! & $\times$ & ! \\
 \hline
-MINE & ! & ! & ! & \checkmark\\
+MINE & ! & \checkmark  & \checkmark & ! & ! \\
 \hline

### 2018-10-30 scagnostics.Rmd
Some text updates in the summary.

@@
 The first three entries are from the scagnostics family.
-The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures (the same is however not true for each single index viewed on its own).
-On the other hand we have found that they are typically too noisy making optimisation of the function challenging.
+The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures, making the set flexible in applications (the same is however not true for each single index viewed on its own).
+On the other hand we have found that they are typically too noisy making optimisation of the function challenging, which also results in poor squinatbility.
 In particular the stringy index cannot be used with current optimisers.
@@
 We have only considered the non-binned implementation of dcor2D which is slow to calculate on large datasets.
-While this index is in principle general, we have found it to be too sensitive to noise to be applicable in most realistic examples.
+While this index is in principle general, we have found it to be too sensitive to noise to be applicable in most realistic examples, therefore we consider flexibility limited.
 Finally the MINE indexes are found to be smooth only for large datasets, with some rotation dependence observed for the TIC index.
 While these indexes are slow to compute especially on large datasets, we have found them to be the most general, allowing good discrimination even with noisy structures.
+In addition we observed good squinatbility for all three examples considered, given large datasets.
+Results in Section \ref{sec:params} suggest that tuning of the MINE parameters may result in improved behaviour for smaller datasets.
 
@@
 \begin{center}
-\caption{Summary}
+\caption{Summary of findings, showing to what extend the considered index functions pass the criteria for a good PPI.
+Checkmark symbols good behaviour, exclamation mark symbols some issues while cross symbols failure on the corresponding criteria.
+Each index has particular strengths and drawbacks and selection must be guided by the considered example, see text for details.}
 \label{tab:summary}

### 2018-10-30 scagnostics.Rmd
Comments on finding spiral.

@@
 
+mineIndexE <- function(mineIndex){
+  function(mat){
+    return(mine(mat[,1], mat[,2], est = "mic_e")[[mineIndex]])
+  }
+}
+
 mineIndexAlpha <- function(mineIndex, alpha){
@@
   set.seed(1984)
-  pipeResc <- rescale(pipe1000) 
+  pipeResc <- as.matrix(pipe1000) 
   pipeTour <- save_history(pipeResc,
@@
 if(!file.exists("cache/findsine.rda")){
-  sineResc <- rescale(sin1000)
+  sineResc <- as.matrix(sin1000)
   sineTour <- save_history(sineResc,
@@
 ggplot(dProj, aes(V1,V2)) + geom_point()
+```
+
+### Spiral detection
+
+\textcolor{red}{can use skinny to find sprial in 3 dimensions, but not higher, scouting method does not work because of large rotation dependence, maybe make it study of guided tour success vs number of parameters entering?
+e.g.
+given n starting planes how many times was view in plane V5-V6 identified as function of number of dimensions and index used?}
+
+\textcolor{red}{what about role of optimisation here?
+search_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
+(but may also give same issues of rotation dependence) -- tried but didnt really work...}
+
+```{r findspiral, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the skinny index for the Sprial 1000 dataset.", results="hide"}
+#getResWithSeed <- function(s){
+#  set.seed(s)
+#  startM <- orthonormalise(matrix(runif(12),ncol = 2))
+#  tourRes <- save_history(spiralResc, guided_tour(scagIndex("Skinny")), max_bases = 10000)
+#}
+#set.seed(202020)
+#if(!file.exists("cache/findsine.rda")){
+#  spiralResc <- as.matrix(spiral1000)
+
+#spiralTour <- save_history(spiralResc, guided_tour(scagIndex("Skinny"), max.tries = 100, n=10, stepS=0.5), max_bases = 10000)
+  #spiralTour <- save_history(spiralResc,guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
+  #                                   alpha = 0.5, max.tries = 500), max_bases = 10000)
+  #spiralTourFull <- as.list(interpolate(spiralTour))
+  #spiralTourRes <- getProj(spiral1000, spiralTourFull, "Spiral", 1000)
+  #spiralTourResMelt <- melt(spiralTourRes, id=c("t", "name", "size"))
+#  save(spiralResc, spiralTour, spiralTourFull, spiralTourRes, spiralTourResMelt,
+#       file = "cache/findspiral.rda")
+#} else {
+#  load("cache/findsine.rda")
+#}
+#ggplot(spiralTourResMelt, aes(x=t, y=value)) +
+#  geom_line() + 
+#  facet_wrap(~variable, ncol=1, scales = "free_y")
 ```

### 2018-10-31 scagnostics.Rmd
Testing different bins parameter for scagnostics indices.

@@
 
+scagIndexNbin <- function(scagType, n){
+  function(mat){
+    sR <- scagnostics.default(mat[,1], mat[,2], bins = n)$s
+    return(sR[scagType])
+  }
+}
+
 invConvexIndex <- function(){
@@
 
+### Scagnostics binning
+The implementation of scagnostics will always perform a binning step first, before evaluating the measures.
+The binning will introduce random fluctuations on interpolated tour paths, but we may also consider using it to reduce fluctuations from the discreteness inherent to the index definitions, see in particular the definition of the stringy index.
+This needs some care however as different values may be preferred for different structures and depending on which index is being considered.
+Moreover, choosing a binning that is too rough may result in loss of discrimination power as structures can no longer be resolved.
+As an example we compare in Figure \ref{fig:sprialScagBinning} the scagnostics index traces for the Spiral 1000 dataset when moving from V1-V2 to V5-V6, for three values of the bins parameter entering the hexagonal binning procedure.
+We find indeed that different values may be preferred depending on the index considered, with the default bins=50 being the better choice for the convex and skinny index, while bins=20 may work better for the stringy index.
+Note that neither choice results in smooth curves for any of the indexes considered here.
+
+```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.
+For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
+With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
+For the convex index we also find bins=50 to be the better choice."}
+getScagComp <- function(df, tPath){
+  sc <- tibble(
+    convex10=numeric(),
+    skinny10=numeric(),
+    stringy10=numeric(),
+    convex20=numeric(),
+    skinny20=numeric(),
+    stringy20=numeric(),
+    convex50=numeric(),
+    skinny50=numeric(),
+    stringy50=numeric(),
+    t=numeric())
+  n <- length(tPath)
+  for (i in 1:n) {
+    dprj <- as.matrix(df) %*% tPath[[i]]
+    scagRes10 <- scagnostics.default(dprj[,1], dprj[,2], bins=10)$s
+    scagRes20 <- scagnostics.default(dprj[,1], dprj[,2], bins=20)$s
+    scagRes50 <- scagnostics.default(dprj[,1], dprj[,2], bins=50)$s
+    sc <- add_row(sc,
+                  convex10=scagRes10["Convex"], skinny10=scagRes10["Skinny"],
+                  stringy10=scagRes10["Stringy"],
+                  convex20=scagRes20["Convex"], skinny20=scagRes20["Skinny"],
+                  stringy20=scagRes20["Stringy"],
+                  convex50=scagRes50["Convex"], skinny50=scagRes50["Skinny"],
+                  stringy50=scagRes50["Stringy"],
+                  t=i)
+  }
+
+  return(sc)
+}
+if(!file.exists("cache/spiralScagBinning.rda")){
+  scagBinningDf <- getScagComp(spiral1000, t2full)
+  scagMelt <- melt(scagBinningDf, id=c("t")) %>%
+    mutate(nbin = as.numeric(str_sub(variable, start = -2))) %>%
+    mutate(idx = str_sub(variable, end = -3))
+  save(scagBinningDf, scagMelt, file = "cache/spiralScagBinning.rda")
+} else {
+  load("cache/spiralScagBinning.rda")
+}
+ggplot(scagMelt, aes(x=t, y=value)) +
+  geom_line(aes(color=factor(nbin))) + 
+  facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
+
+```
+
 ### Illustrative example
@@
 \textcolor{red}{what about role of optimisation here?
-search_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
-(but may also give same issues of rotation dependence) -- tried but didnt really work...}
+search\_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
+(but may also give same issues of rotation dependence) ...
+tried  but didnt really work...}
 
@@
 #  spiralResc <- as.matrix(spiral1000)
-
+#spiralTour <- save_history(spiralResc, guided_tour(cmass))
 #spiralTour <- save_history(spiralResc, guided_tour(scagIndex("Skinny"), max.tries = 100, n=10, stepS=0.5), max_bases = 10000)

### 2018-10-31 scagnostics.Rmd
Starting section on modifications.

@@
 
+## Possible modifications
+
+Big issues: smoothness, rotation invariance
+
+### Smoothing
+
+Focusing on scagnostics indices, for which the biggest issues with noisy transitions have been observed.
+Smoothing can be performed by interpolation, or averaging similar views.
+For example we can use jittering, or we can use tour functionalities to find a set of randomly selected projections within a small angle of the current view.
+In this way we can generate a number of similar views.
+We can then select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calulate the mean from the remaining evaluations.
+
+### Rotation invariance
+Do we want to say anything here? big issue for splines2D measure, but also scagnostics
+
 ## Summary

### 2018-10-31 scagnostics.Rmd
Spiral description added.

@@
 - **sine**:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-- **sphere**:  $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
+- **spiral**:  $p$ points where $n-2$ points are independently drawn from a normal distribution, and dirction $n-1$ and $n$ have points selected on an Archimedean spiral, i.e.
+$r = a + b \theta$, with $a=b=0.1$ and we sample angles $\theta$ from a normal distribution with mean 0 and variance $2\pi$ and taking the absolute for well defined spiral shape with higher density at smaller radius.
 

### 2018-10-31 scagnostics.Rmd
Summary table symbols updated.
Notation updated.

--- 2018-10-31 scagnostics.Rmd
@@
 
-The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($k=1$ or $2$ usually).
-The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $k$-dimensional projection of $p$-dimensional data.
-This criterion function is optimized over the space of all $k$-dimensional projections of $p$-space, searching for both global and local maxima.
+The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
+The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $d$-dimensional projection of $p$-dimensional data.
+This criterion function is optimized over the space of all $d$-dimensional projections of $p$-space, searching for both global and local maxima.
 It is hoped that the resulting solutions reveal low-dimensional structure in the data not found by methods such as principal component analysis.
@@
 Our motivation is based on applications in physics, to aid the interpretation of model fits on experimental results.
-Consider a physical model with a set of $m$ free parameters, that cannot be measured directly and are determined by fitting a set of $p$ experimental observations, for which predictions can be calculated when fixing all $m$ parameters.
+Consider a physical model with a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $n$ experimental observations, for which predictions can be calculated when fixing all $p$ parameters.
 Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.
 Moreover, a single prediction can be a complicated function of all free parameters.
-In addition, we often deal with large number of observations, $p \sim 100-1000$ as well as sizable parameter spaces $m \sim 10$.
+In addition, we often deal with large number of observations, $n \sim 100-1000$ as well as sizable parameter spaces $p \sim 10$.
 The results of a fit are generally interpreted using combinations of variables selected by intuition and prior knowledge.
@@
 -->
-A projection pursuit index (PPI) is a scalar function $f$ defined on an $n$-dimensional data distribution.
-Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $n$ dimensional projections of a data set with $p>n$ parameters will find the most interesting projections.
+A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data distribution.
+Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $d$ dimensional projections of a data set with $p>d$ parameters will find the most interesting projections.
 As most indices characterise deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
@@
 ## Optimization
-Given a PPI we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
+Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
 One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
@@
 \label{sec:indexDef}
-Focusing now on $n=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
+Focusing now on $d=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
 We therefore draw on index functions that have been developped for the selection of variable pairs that show interesting features in a scatter plot, but we generalise the notion such that instead of variable pairs we consider any $2$ dimensional projection of the multivariate data distribution.
@@
 
-- **pipe**: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-- **sine**:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-- **sphere**:  $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
-
-Two samples sizes are also used: $p=100$ (small) and $p=1000$ (large).
+- **pipe**: $n$ points where all $p$ directions are independently drawn from a uniform distribution between $[-1,1]$, but enforcing $x_p^2 + x_{p-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+- **sine**:  $n$ points where $p-1$ directions are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_p = \sin(x_{p-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
+- **spiral**:  $n$ points where $p-2$ directions are independently drawn from a normal distribution, and direction $p-1$ and $p$ are sampled on an Archimedean spiral, i.e.
+$r = a + b \theta$, with $a=b=0.1$ and we sample angles $\theta$ from a normal distribution with mean 0 and variance $2\pi$, giving a spiral with higher densities at lower radii.
+We take the absolute value of $\theta$ to fix the direction of the spiral shape when sampling the points.
+
+Two samples sizes are also used: $n=100$ (small) and $n=1000$ (large).
 
@@
 - index values for pair of variables, explicitly comparing the uninformative and special view of each dataset, as well as discussing the dependence on the sample size, see Section \ref{sec:dataOv}
-- tour between pairs of variables, considering a planned tour that moves from the $V_1 - V_2$ view either to the $V_3 - V_4$ view (no special structure encountered over the full path) or to the $V_5 - V_6$ (moving from uninformed to special view), see Section \ref{sec:smooth}
-- optimisation path, starting from $V_1 - V_2$, how can we use the available index functions and optimisation algorithms to identify the special view, see Section \ref{sec:guided}
+- tour between pairs of variables, considering a planned tour that moves from the $x_1 - x_2$ view either to the $x_3 - x_4$ view (no special structure encountered over the full path) or to the $x_5 - x_6$ (moving from uninformed to special view), see Section \ref{sec:smooth}
+- optimisation path, starting from $x_1 - x_2$, how can we use the available index functions and optimisation algorithms to identify the special view, see Section \ref{sec:guided}
 - rotating a dataset with functional dependence to evaluate the rotation dependence of the index functions, see Section \ref{sec:rot}
@@
 \label{sec:dataOv}
-
-To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
-They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
-
-- sphere:  $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
-- pipe: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-- sine:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-
-For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
 
@@
 
-To get an overview of the dataset we show views of the first two parameters showing the independent distribution and the last two parameters showing the special view.
-The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of V6 (and V5 for the Pipe distribution) will be distinct from uninformed views of distributions 2 and 3.
-
-```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters V1 and V2, i.e. showing a typical uninformed view for each dataset."}
+To get an overview of the dataset we show views of the first two directions showing the independent distribution and the last two directions showing the special view.
+The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of $x_6$ (and $x_5$ for the Pipe and Spiral distribution) will be distinct from uninformed views of distributions 2 and 3.
+
+```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
+showing a typical uninformed view for each dataset."}
 pL <- list()
@@
 
-```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters V5 and V6, i.e.
-showing the special view built into the Pipe and Sine datasets."}
+```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters $x_5$ and $x_6$, i.e. showing the special view built into the datasets."}
 pL <- list()
@@
 ```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
-Apart from statistical fluctuations we see that for the Pipe datasets the parameters V5 and V6 are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine dataset the parameter V6 follows a very distinct distribution.
+Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine dataset the parameter $x_6$ follows a very distinct distribution.
 Note also that the first five paramters of the Sine dataset appear to be drawn from different normal distributions, which is a result of the rescaling, shifting the modes to different values for each direction."}
@@
 
-We next compare the various index values for three views: V1 vs V2, V5 vs V6 and V1 vs V6 (capturing differences from the different distribution in V6).
+We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
 For the Sphere distribution by construction we see only random fluctuations between the different views, which can be sizable in particular for small datasets.
@@
 ### Random fluctuations
-We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto V1-V2 to the projection onto V3-V4 and using the default interpolation angle of 0.05.
+We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$ and using the default interpolation angle of 0.05.
 Each time step refers to an interpolation step.
@@
 Considering now the indices from the scagnostics family, we first note stronge dependence on the sample size.
-The convex index is largely sensitive to random fluctuations when considering small samples (p=100), while being mostly flat for larger samples, especially in the case of an underlying uniform distribution.
+The convex index is largely sensitive to random fluctuations when considering small samples ($n=100$), while being mostly flat for larger samples, especially in the case of an underlying uniform distribution.
 The skinny index on the other hand is systematically larger for the smaller samples, while stringy appears not to be sensitive to the sample size, but resulting in large fluctuations and sharp jumps for all datasets.
 Distance correlation and the spline based index show similar behaviour, i.e.
-when considering large enough data samples (p=1000) the values are consistently close to zero, as expected for a distribution without structure.
+when considering large enough data samples ($n=1000$) the values are consistently close to zero, as expected for a distribution without structure.
 On the other hand somewhat larger values (but still far from the maximum value of 1) are found when considering small datasets, where distributions are however mostly smooth, with few kinks depending on index function and underlying distribution.
@@
 ### Moving into informative projections
-We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6, which will also allow us to view the evolution between typical low and high values of the index function.
+We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function.
 Results are shown in Figure \ref{fig:plannedtour}.
@@
 
-```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto V1-V2 to the projection onto V5-V6, i.e.
+```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto $x_1$-$x_2$ to the projection onto $x_5$-$x_6$, i.e.
 starting from random distribution view, and moving into the special view."}
@@
 Compare typical index values over the two tour paths.
-Notation: path 1 / 2 labels tour from V1-V2 to V3-V4 / V5-V6.
+Notation: path 1 / 2 labels tour from $x_1$-$x_2$ to $x_3$-$x_4$ / $x_5$-$x_6$.
 Typically find broader distributions for small samples because of the large fluctuations in this case.
@@
 \label{sec:rot}
-\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
+\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$, maybe can add second path for spiral?}
 Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binning (and outlier removal?) here.
@@
 Moreover, choosing a binning that is too rough may result in loss of discrimination power as structures can no longer be resolved.
-As an example we compare in Figure \ref{fig:sprialScagBinning} the scagnostics index traces for the Spiral 1000 dataset when moving from V1-V2 to V5-V6, for three values of the bins parameter entering the hexagonal binning procedure.
+As an example we compare in Figure \ref{fig:sprialScagBinning} the scagnostics index traces for the Spiral 1000 dataset when moving from $x_1$-$x_2$ to $x_5$-$x_6$, for three values of the bins parameter entering the hexagonal binning procedure.
 We find indeed that different values may be preferred depending on the index considered, with the default bins=50 being the better choice for the convex and skinny index, while bins=20 may work better for the stringy index.
@@
 \caption{Summary of findings, showing to what extend the considered index functions pass the criteria for a good PPI.
-Checkmark symbols good behaviour, exclamation mark symbols some issues while cross symbols failure on the corresponding criteria.
+"\checkmark" symbols good behaviour, "$\cdot$" symbols some issues while "$\times$" symbols failure on the corresponding criteria.
 Each index has particular strengths and drawbacks and selection must be guided by the considered example, see text for details.}
@@
 \hline
-Convex & ! & ! & ! & ! & \checkmark \\
-Skinny & ! & !& ! & ! & \checkmark \\
-Stringy & $\times$ & $\times$ & ! & ! & \checkmark \\
+Convex & $\cdot$ & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark \\
+Skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark \\
+Stringy & $\times$ & $\times$ & $\cdot$ & $\cdot$ & \checkmark \\
 \hline
-splines2D & \checkmark & \checkmark & ! & $\times$ & \checkmark \\
-dcor2D & \checkmark & \checkmark & ! & $\times$ & ! \\
+splines2D & \checkmark & \checkmark & $\cdot$ & $\times$ & \checkmark \\
+dcor2D & \checkmark & \checkmark & $\cdot$ & $\times$ & $\cdot$ \\
 \hline
-MINE & ! & \checkmark  & \checkmark & ! & ! \\
+MINE & $\cdot$ & \checkmark  & \checkmark & $\cdot$ & $\cdot$ \\
 \hline

### 2018-10-31 scagnostics.Rmd
Replace sphere with spiral in text.

@@
 ```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
-Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine dataset the parameter $x_6$ follows a very distinct distribution.
-Note also that the first five paramters of the Sine dataset appear to be drawn from different normal distributions, which is a result of the rescaling, shifting the modes to different values for each direction."}
+Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine (and Spiral) dataset the parameter $x_6$ (and $x_5$) follows a very distinct distribution."}
 pL <- list()
@@
 We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
-For the Sphere distribution by construction we see only random fluctuations between the different views, which can be sizable in particular for small datasets.
+For the Spiral distribution the scagnostics indices skinny and stringy are large and close to their theoretic maximum for the special view.
+On the other hand while dcor2d is larger is also increased for the special view, the value is far from the theoretic maximum and larger values may be found from random fluctuations in other views (CHECK!).
+MIC and TIC are also found to be sensitive to the spiral distribution.
 Considering next the Pipe distribution, we see that skinny, stringy and dcor2d are increased for the special view, but in particular MIC and TIC index are much larger and may be used to detect this type of structure in the dataset.
@@
 
-```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto V1-V2 to the projection onto V3-V4, for all considered datasets and index functions.
-Small samples (p=100) are shown in red, large samples (p=1000) in blue."}
+```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$, for all considered datasets and index functions.
+Small samples (n=100) are shown in red, large samples (n=1000) in blue."}
 fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
@@
 On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
+Finally the Spiral appears to be the most challenging, having an even smaller squint angle than the Pipe distribution.
+While "cmass" appears to be appropriate to identify this view we may actually find larger values of this index for random distributions, making it unfit to use in a guided tour.
+The scagnostics indices skinny (or 1-convex) appear to be sensitive to the structure at a larger angle compared to MINE indices, but will suffer from the large fluctuations in random views when used in a guided tour.
 Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
@@
 \label{sec:timer}
-Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example the 6-d sphere and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
+Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example samples of points in a 6-d solid sphere (obtained via the geozoo package [@geozoo]) and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
+\textcolor{red}{Since projection is fast compared to index evaluation, the number of original dimensions is not very important here.
+On the other hand different distributions may impact the computing time.}
 We see that the evaluation of the spline index shows weak dependence on the sample size, making it the least efficient for very small data sets, but one of the fastest when considering large data sets.

### 2018-10-31 scagnostics.Rmd
Checked dcor2d over spiral distributions, updated text.

@@
 For the Spiral distribution the scagnostics indices skinny and stringy are large and close to their theoretic maximum for the special view.
-On the other hand while dcor2d is larger is also increased for the special view, the value is far from the theoretic maximum and larger values may be found from random fluctuations in other views (CHECK!).
+On the other hand while dcor2d is larger is also increased for the special view, the value is far from the theoretic maximum and larger values may be found in other views.
 MIC and TIC are also found to be sensitive to the spiral distribution.
@@
 #  spiralResc <- as.matrix(spiral1000)
-#spiralTour <- save_history(spiralResc, guided_tour(cmass))
+#spiralTour <- save_history(spiralResc, guided_tour(dcorIndex()))
 #spiralTour <- save_history(spiralResc, guided_tour(scagIndex("Skinny"), max.tries = 100, n=10, stepS=0.5), max_bases = 10000)
@@
 #  facet_wrap(~variable, ncol=1, scales = "free_y")
+#iL <- length(as.list(spiralTour))
+#dProj <- spiralResc %*% as.list(spiralTour)[[iL]]
+#ggplot(as.tibble(dProj), aes(V1, V2)) + geom_point()
 ```

### 2018-11-01 scagnostics.Rmd
References added.

@@
 
-The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
+The term "projection pursuit" was coined by @FT74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
 The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $d$-dimensional projection of $p$-dimensional data.
@@
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
-\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber1985,jones87}.}
+\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber85,jones87}.}
 These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
@@
 - **squintability**: \cite{barnett1981interpreting} introduced the idea of squint angle to indicate resolution of structure in the data.
-Fine structure like the parallel planes in the infamous RANDU data \ref{randu} has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
+Fine structure like the parallel planes in the infamous RANDU data \cite{compaqFortran} (FIXME this is fortran reference, could not find one for the dataset) has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
 Structures with small squint angle are difficult to find, because the optmisation algorithm needs to get very close to begin hill-climbing to the optimum.
@@
 If a particular orientation is more optimal, this will get lost as the projection shown pays no attention to orientation.
-\cite{bujaXXX} describes alternative interpolation paths based on Givens and Householder rotations which progress from basis to basis.
+\cite{Buja2004ComputationalMF} describes alternative interpolation paths based on Givens and Householder rotations which progress from basis to basis.
 It may be possible to ignore rotation invariance with these interpolations but there isn't a current implementation, primarily because the within-plane spin that is generated is distracting from a visualisation perspective.

### 2018-11-01 scagnostics.Rmd
I think I might have made changes to 3.1

--- 2018-10-31 scagnostics.Rmd
@@
 
-The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
-The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $d$-dimensional projection of $p$-dimensional data.
-This criterion function is optimized over the space of all $d$-dimensional projections of $p$-space, searching for both global and local maxima.
+The term "projection pursuit" was coined by @ff74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($k=1$ or $2$ usually).
+The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $k$-dimensional projection of $p$-dimensional data.
+This criterion function is optimized over the space of all $k$-dimensional projections of $p$-space, searching for both global and local maxima.
 It is hoped that the resulting solutions reveal low-dimensional structure in the data not found by methods such as principal component analysis.
@@
 Our motivation is based on applications in physics, to aid the interpretation of model fits on experimental results.
-Consider a physical model with a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $n$ experimental observations, for which predictions can be calculated when fixing all $p$ parameters.
+Consider a physical model with a set of $m$ free parameters, that cannot be measured directly and are determined by fitting a set of $p$ experimental observations, for which predictions can be calculated when fixing all $m$ parameters.
 Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.
 Moreover, a single prediction can be a complicated function of all free parameters.
-In addition, we often deal with large number of observations, $n \sim 100-1000$ as well as sizable parameter spaces $p \sim 10$.
+In addition, we often deal with large number of observations, $p \sim 100-1000$ as well as sizable parameter spaces $m \sim 10$.
 The results of a fit are generally interpreted using combinations of variables selected by intuition and prior knowledge.
@@
 -->
-A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data distribution.
-Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $d$ dimensional projections of a data set with $p>d$ parameters will find the most interesting projections.
+A projection pursuit index (PPI) is a scalar function $f$ defined on an $n$-dimensional data distribution.
+Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $n$ dimensional projections of a data set with $p>n$ parameters will find the most interesting projections.
 As most indices characterise deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
@@
 ## Optimization
-Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
+Given a PPI we are confronted with the task of finding the maximum over all possible $n$ dimensional projections.
 One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
@@
 \label{sec:indexDef}
-Focusing now on $d=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
+Focusing now on $n=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
 We therefore draw on index functions that have been developped for the selection of variable pairs that show interesting features in a scatter plot, but we generalise the notion such that instead of variable pairs we consider any $2$ dimensional projection of the multivariate data distribution.
@@
 However, it serves the purpose to also evaluate the PPIs.
-The three data families are:
-
-- **pipe**: $n$ points where all $p$ directions are independently drawn from a uniform distribution between $[-1,1]$, but enforcing $x_p^2 + x_{p-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-- **sine**:  $n$ points where $p-1$ directions are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_p = \sin(x_{p-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-- **spiral**:  $n$ points where $p-2$ directions are independently drawn from a normal distribution, and direction $p-1$ and $p$ are sampled on an Archimedean spiral, i.e.
-$r = a + b \theta$, with $a=b=0.1$ and we sample angles $\theta$ from a normal distribution with mean 0 and variance $2\pi$, giving a spiral with higher densities at lower radii.
-We take the absolute value of $\theta$ to fix the direction of the spiral shape when sampling the points.
-
-Two samples sizes are also used: $n=100$ (small) and $n=1000$ (large).
+The three data families are explained below.
+In each set, $p$ is used for the number of points, $n$ is the number of dimensions, and $d=2$ is the projection dimension.
+
+- **pipe**: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+- **sine**:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
+- **spiral**: XXX $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
+
+Two samples sizes are also used: $p=100$ (small) and $p=1000$ (large).
 
@@
 - index values for pair of variables, explicitly comparing the uninformative and special view of each dataset, as well as discussing the dependence on the sample size, see Section \ref{sec:dataOv}
-- tour between pairs of variables, considering a planned tour that moves from the $x_1 - x_2$ view either to the $x_3 - x_4$ view (no special structure encountered over the full path) or to the $x_5 - x_6$ (moving from uninformed to special view), see Section \ref{sec:smooth}
-- optimisation path, starting from $x_1 - x_2$, how can we use the available index functions and optimisation algorithms to identify the special view, see Section \ref{sec:guided}
+- tour between pairs of variables, considering a planned tour that moves from the $V_1 - V_2$ view either to the $V_3 - V_4$ view (no special structure encountered over the full path) or to the $V_5 - V_6$ (moving from uninformed to special view), see Section \ref{sec:smooth}
+- optimisation path, starting from $V_1 - V_2$, how can we use the available index functions and optimisation algorithms to identify the special view, see Section \ref{sec:guided}
 - rotating a dataset with functional dependence to evaluate the rotation dependence of the index functions, see Section \ref{sec:rot}
@@
 \label{sec:dataOv}
+
+To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
+They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
+
+- spiral:  XXX $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
+- pipe: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+- sine:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
+
+For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
 
@@
     sR <- scagnostics.default(mat[,1],mat[,2])$s
-    return(sR[scagType])
-  }
-}
-
-scagIndexNbin <- function(scagType, n){
-  function(mat){
-    sR <- scagnostics.default(mat[,1], mat[,2], bins = n)$s
     return(sR[scagType])
@@
 
-To get an overview of the dataset we show views of the first two directions showing the independent distribution and the last two directions showing the special view.
-The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of $x_6$ (and $x_5$ for the Pipe and Spiral distribution) will be distinct from uninformed views of distributions 2 and 3.
-
-```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
-showing a typical uninformed view for each dataset."}
+To get an overview of the dataset we show views of the first two parameters showing the independent distribution and the last two parameters showing the special view.
+The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of V6 (and V5 for the Pipe distribution) will be distinct from uninformed views of distributions 2 and 3.
+
+```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters V1 and V2, i.e. showing a typical uninformed view for each dataset."}
 pL <- list()
@@
 
-```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters $x_5$ and $x_6$, i.e. showing the special view built into the datasets."}
+```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters V5 and V6, i.e.
+showing the special view built into the Pipe and Sine datasets."}
 pL <- list()
@@
 ```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
-Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine (and Spiral) dataset the parameter $x_6$ (and $x_5$) follows a very distinct distribution."}
+Apart from statistical fluctuations we see that for the Pipe datasets the parameters V5 and V6 are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine dataset the parameter V6 follows a very distinct distribution.
+Note also that the first five paramters of the Sine dataset appear to be drawn from different normal distributions, which is a result of the rescaling, shifting the modes to different values for each direction."}
 pL <- list()
@@
 
-We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
-For the Spiral distribution the scagnostics indices skinny and stringy are large and close to their theoretic maximum for the special view.
-On the other hand while dcor2d is larger is also increased for the special view, the value is far from the theoretic maximum and larger values may be found from random fluctuations in other views (CHECK!).
-MIC and TIC are also found to be sensitive to the spiral distribution.
+We next compare the various index values for three views: V1 vs V2, V5 vs V6 and V1 vs V6 (capturing differences from the different distribution in V6).
+For the Sphere distribution by construction we see only random fluctuations between the different views, which can be sizable in particular for small datasets.
 Considering next the Pipe distribution, we see that skinny, stringy and dcor2d are increased for the special view, but in particular MIC and TIC index are much larger and may be used to detect this type of structure in the dataset.
@@
 ### Random fluctuations
-We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$ and using the default interpolation angle of 0.05.
+We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto V1-V2 to the projection onto V3-V4 and using the default interpolation angle of 0.05.
 Each time step refers to an interpolation step.
@@
 Considering now the indices from the scagnostics family, we first note stronge dependence on the sample size.
-The convex index is largely sensitive to random fluctuations when considering small samples ($n=100$), while being mostly flat for larger samples, especially in the case of an underlying uniform distribution.
+The convex index is largely sensitive to random fluctuations when considering small samples (p=100), while being mostly flat for larger samples, especially in the case of an underlying uniform distribution.
 The skinny index on the other hand is systematically larger for the smaller samples, while stringy appears not to be sensitive to the sample size, but resulting in large fluctuations and sharp jumps for all datasets.
 Distance correlation and the spline based index show similar behaviour, i.e.
-when considering large enough data samples ($n=1000$) the values are consistently close to zero, as expected for a distribution without structure.
+when considering large enough data samples (p=1000) the values are consistently close to zero, as expected for a distribution without structure.
 On the other hand somewhat larger values (but still far from the maximum value of 1) are found when considering small datasets, where distributions are however mostly smooth, with few kinks depending on index function and underlying distribution.
@@
 
-```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$, for all considered datasets and index functions.
-Small samples (n=100) are shown in red, large samples (n=1000) in blue."}
+```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto V1-V2 to the projection onto V3-V4, for all considered datasets and index functions.
+Small samples (p=100) are shown in red, large samples (p=1000) in blue."}
 fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
@@
 ### Moving into informative projections
-We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function.
+We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6, which will also allow us to view the evolution between typical low and high values of the index function.
 Results are shown in Figure \ref{fig:plannedtour}.
@@
 On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
-Finally the Spiral appears to be the most challenging, having an even smaller squint angle than the Pipe distribution.
-While "cmass" appears to be appropriate to identify this view we may actually find larger values of this index for random distributions, making it unfit to use in a guided tour.
-The scagnostics indices skinny (or 1-convex) appear to be sensitive to the structure at a larger angle compared to MINE indices, but will suffer from the large fluctuations in random views when used in a guided tour.
 Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
@@
 
-```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto $x_1$-$x_2$ to the projection onto $x_5$-$x_6$, i.e.
+```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto V1-V2 to the projection onto V5-V6, i.e.
 starting from random distribution view, and moving into the special view."}
@@
 Compare typical index values over the two tour paths.
-Notation: path 1 / 2 labels tour from $x_1$-$x_2$ to $x_3$-$x_4$ / $x_5$-$x_6$.
+Notation: path 1 / 2 labels tour from V1-V2 to V3-V4 / V5-V6.
 Typically find broader distributions for small samples because of the large fluctuations in this case.
@@
 \label{sec:rot}
-\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$, maybe can add second path for spiral?}
+\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
 Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binning (and outlier removal?) here.
@@
 \label{sec:timer}
-Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example samples of points in a 6-d solid sphere (obtained via the geozoo package [@geozoo]) and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
-\textcolor{red}{Since projection is fast compared to index evaluation, the number of original dimensions is not very important here.
-On the other hand different distributions may impact the computing time.}
+Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example the 6-d sphere and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
 We see that the evaluation of the spline index shows weak dependence on the sample size, making it the least efficient for very small data sets, but one of the fastest when considering large data sets.
@@
 In our examples we find the number to be appropriate to identify functional dependence while rejecting noise, but some distributions may require tuning of this parameter.
-
-### Scagnostics binning
-The implementation of scagnostics will always perform a binning step first, before evaluating the measures.
-The binning will introduce random fluctuations on interpolated tour paths, but we may also consider using it to reduce fluctuations from the discreteness inherent to the index definitions, see in particular the definition of the stringy index.
-This needs some care however as different values may be preferred for different structures and depending on which index is being considered.
-Moreover, choosing a binning that is too rough may result in loss of discrimination power as structures can no longer be resolved.
-As an example we compare in Figure \ref{fig:sprialScagBinning} the scagnostics index traces for the Spiral 1000 dataset when moving from $x_1$-$x_2$ to $x_5$-$x_6$, for three values of the bins parameter entering the hexagonal binning procedure.
-We find indeed that different values may be preferred depending on the index considered, with the default bins=50 being the better choice for the convex and skinny index, while bins=20 may work better for the stringy index.
-Note that neither choice results in smooth curves for any of the indexes considered here.
-
-```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.
-For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
-With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
-For the convex index we also find bins=50 to be the better choice."}
-getScagComp <- function(df, tPath){
-  sc <- tibble(
-    convex10=numeric(),
-    skinny10=numeric(),
-    stringy10=numeric(),
-    convex20=numeric(),
-    skinny20=numeric(),
-    stringy20=numeric(),
-    convex50=numeric(),
-    skinny50=numeric(),
-    stringy50=numeric(),
-    t=numeric())
-  n <- length(tPath)
-  for (i in 1:n) {
-    dprj <- as.matrix(df) %*% tPath[[i]]
-    scagRes10 <- scagnostics.default(dprj[,1], dprj[,2], bins=10)$s
-    scagRes20 <- scagnostics.default(dprj[,1], dprj[,2], bins=20)$s
-    scagRes50 <- scagnostics.default(dprj[,1], dprj[,2], bins=50)$s
-    sc <- add_row(sc,
-                  convex10=scagRes10["Convex"], skinny10=scagRes10["Skinny"],
-                  stringy10=scagRes10["Stringy"],
-                  convex20=scagRes20["Convex"], skinny20=scagRes20["Skinny"],
-                  stringy20=scagRes20["Stringy"],
-                  convex50=scagRes50["Convex"], skinny50=scagRes50["Skinny"],
-                  stringy50=scagRes50["Stringy"],
-                  t=i)
-  }
-
-  return(sc)
-}
-if(!file.exists("cache/spiralScagBinning.rda")){
-  scagBinningDf <- getScagComp(spiral1000, t2full)
-  scagMelt <- melt(scagBinningDf, id=c("t")) %>%
-    mutate(nbin = as.numeric(str_sub(variable, start = -2))) %>%
-    mutate(idx = str_sub(variable, end = -3))
-  save(scagBinningDf, scagMelt, file = "cache/spiralScagBinning.rda")
-} else {
-  load("cache/spiralScagBinning.rda")
-}
-ggplot(scagMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(nbin))) + 
-  facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
-
-```
 
@@
 \textcolor{red}{what about role of optimisation here?
-search\_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
-(but may also give same issues of rotation dependence) ...
-tried  but didnt really work...}
+search_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
+(but may also give same issues of rotation dependence) -- tried but didnt really work...}
 
@@
 #  spiralResc <- as.matrix(spiral1000)
-#spiralTour <- save_history(spiralResc, guided_tour(cmass))
+
 #spiralTour <- save_history(spiralResc, guided_tour(scagIndex("Skinny"), max.tries = 100, n=10, stepS=0.5), max_bases = 10000)
@@
 
-## Possible modifications
-
-Big issues: smoothness, rotation invariance
-
-### Smoothing
-
-Focusing on scagnostics indices, for which the biggest issues with noisy transitions have been observed.
-Smoothing can be performed by interpolation, or averaging similar views.
-For example we can use jittering, or we can use tour functionalities to find a set of randomly selected projections within a small angle of the current view.
-In this way we can generate a number of similar views.
-We can then select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calulate the mean from the remaining evaluations.
-
-### Rotation invariance
-Do we want to say anything here? big issue for splines2D measure, but also scagnostics
-
 ## Summary
@@
 \caption{Summary of findings, showing to what extend the considered index functions pass the criteria for a good PPI.
-"\checkmark" symbols good behaviour, "$\cdot$" symbols some issues while "$\times$" symbols failure on the corresponding criteria.
+Checkmark symbols good behaviour, exclamation mark symbols some issues while cross symbols failure on the corresponding criteria.
 Each index has particular strengths and drawbacks and selection must be guided by the considered example, see text for details.}
@@
 \hline
-Convex & $\cdot$ & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark \\
-Skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark \\
-Stringy & $\times$ & $\times$ & $\cdot$ & $\cdot$ & \checkmark \\
+Convex & ! & ! & ! & ! & \checkmark \\
+Skinny & ! & !& ! & ! & \checkmark \\
+Stringy & $\times$ & $\times$ & ! & ! & \checkmark \\
 \hline
-splines2D & \checkmark & \checkmark & $\cdot$ & $\times$ & \checkmark \\
-dcor2D & \checkmark & \checkmark & $\cdot$ & $\times$ & $\cdot$ \\
+splines2D & \checkmark & \checkmark & ! & $\times$ & \checkmark \\
+dcor2D & \checkmark & \checkmark & ! & $\times$ & ! \\
 \hline
-MINE & $\cdot$ & \checkmark  & \checkmark & $\cdot$ & $\cdot$ \\
+MINE & ! & \checkmark  & \checkmark & ! & ! \\
 \hline

### 2018-11-01 scagnostics.Rmd
Smoothing example.

--- 2018-11-01 scagnostics.Rmd
@@
 We can then select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calulate the mean from the remaining evaluations.
+Similar to the study of scagnostics binning we now compare different methods of smoothing when moving into the special view of the Spiral 1000 dataset.
+
+```{r spiralScagSmoothing, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.
+For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
+With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
+For the convex index we also find bins=50 to be the better choice."}
+
+jitterScagnostics <-function(proj, d, alpha){
+  newProj <- tourr:::basis_nearby(proj, alpha = alpha, method = "geodesic")
+  newD <- d %*% newProj
+  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
+}
+
+jitterPointsScagnostics <-function(projData, alpha){
+  newD <- jitter(projData, amount=alpha)
+  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
+}
+
+getSgnMean <- function(proj, d, alpha, method){
+  dProj <- d %*% proj
+  orig <- as.vector(scagnostics.default(dProj[,1],dProj[,2])$s)
+  if(method == "jitterAngle"){
+    sgnVec <- replicate(10, jitterScagnostics(proj, d, alpha))
+  } else if (method=="jitterPoints"){
+    sgnVec <- replicate(10, jitterPointsScagnostics(dProj, alpha))
+  }
+  return(rowMeans(cbind(orig, sgnVec)))
+}
+
+getScagSmooth <- function(df, tPath){
+  sc <- tibble(
+    convex=numeric(),
+    skinny=numeric(),
+    stringy=numeric(),
+    t=numeric(),
+    method=character(),
+    alpha=numeric())
+  n <- length(tPath)
+  for (method in c("jitterAngle", "jitterPoints")){
+    for (alpha in c(0.01, 0.05, 0.1)){
+      for (i in 1:n) {
+        scagMean <- getSgnMean(tPath[[i]], as.matrix(df), alpha, method)
+        convex <- scagMean[6]
+        skinny <- scagMean[7]
+        stringy <- scagMean[8]
+        sc <- sc %>% add_row(convex=convex, skinny=skinny, stringy=stringy,
+                             t=i, method=method, alpha=alpha)
+      }
+    }
+  }
+
+  return(sc)
+}
+
+if(!file.exists("cache/spiralScagSmoothing.rda")){
+  scagSmooth <- getScagSmooth(spiral1000, as.list(t2full))
+  scagSmoothMelt <- melt(scagSmooth, id=c("t", "method", "alpha"))
+  save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")
+} else {
+  load("cache/spiralScagSmoothing.rda")
+}
+
+ggplot(scagSmoothMelt, aes(x=t, y=value)) +
+  geom_line(aes(color=method, linetype=factor(alpha))) + 
+  facet_wrap(~variable, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
+
+```
 

### 2018-11-01 scagnostics.Rmd

--- 2018-11-01 scagnostics.Rmd
@@
 However, it serves the purpose to also evaluate the PPIs.
-The three data families are:
-
-- **pipe**: $n$ points where all $p$ directions are independently drawn from a uniform distribution between $[-1,1]$, but enforcing $x_p^2 + x_{p-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-- **sine**:  $n$ points where $p-1$ directions are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_p = \sin(x_{p-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
+The three data families are explained below.
+In each set, $p$ is used for the number of points, $n$ is the number of dimensions, and $d=2$ is the projection dimension.
+
+- **pipe**: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+- **sine**:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
 - **spiral**:  $n$ points where $p-2$ directions are independently drawn from a normal distribution, and direction $p-1$ and $p$ are sampled on an Archimedean spiral, i.e.

### 2018-11-01 scagnostics.Rmd
data set explanation

@@
 
-- **pipe**: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-- **sine**:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-- **spiral**:  $n$ points where $p-2$ directions are independently drawn from a normal distribution, and direction $p-1$ and $p$ are sampled on an Archimedean spiral, i.e.
+- **pipe**: nuisance directions are generated by sampling independently from a uniform distribution between $[-1,1]$, and the circle is generated by sampling randomly on a 2D circle, and adding a small radial noise.
+- **sine**:  nuisance directions are generated by sampling independently from a standard normal distribution, and the sine curve is generated by $x_n = \sin(x_{n-1}) + \mathrm{jittering}$.
+- **spiral**:  nuisance directions are generated by sampling independently from a normal distribution, and the structure directions are sampled from an Archimedean spiral, i.e.
 $r = a + b \theta$, with $a=b=0.1$ and we sample angles $\theta$ from a normal distribution with mean 0 and variance $2\pi$, giving a spiral with higher densities at lower radii.
-We take the absolute value of $\theta$ to fix the direction of the spiral shape when sampling the points.
+The absolute value of $\theta$ fixes the direction of the spiral shape.
 

### 2018-11-01 scagnostics.Rmd
More facets for smoothing compairson.

@@
 However, it serves the purpose to also evaluate the PPIs.
-The three data families are explained below.
-In each set, $p$ is used for the number of points, $n$ is the number of dimensions, and $d=2$ is the projection dimension.
-
-- **pipe**: nuisance directions are generated by sampling independently from a uniform distribution between $[-1,1]$, and the circle is generated by sampling randomly on a 2D circle, and adding a small radial noise.
-- **sine**:  nuisance directions are generated by sampling independently from a standard normal distribution, and the sine curve is generated by $x_n = \sin(x_{n-1}) + \mathrm{jittering}$.
-- **spiral**:  nuisance directions are generated by sampling independently from a normal distribution, and the structure directions are sampled from an Archimedean spiral, i.e.
+The three data families are:
+
+- **pipe**: $n$ points where all $p$ directions are independently drawn from a uniform distribution between $[-1,1]$, but enforcing $x_p^2 + x_{p-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
+- **sine**:  $n$ points where $p-1$ directions are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_p = \sin(x_{p-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
+- **spiral**:  $n$ points where $p-2$ directions are independently drawn from a normal distribution, and direction $p-1$ and $p$ are sampled on an Archimedean spiral, i.e.
 $r = a + b \theta$, with $a=b=0.1$ and we sample angles $\theta$ from a normal distribution with mean 0 and variance $2\pi$, giving a spiral with higher densities at lower radii.
-The absolute value of $\theta$ fixes the direction of the spiral shape.
+We take the absolute value of $\theta$ to fix the direction of the spiral shape when sampling the points.
 
@@
 \label{sec:dataOv}
-
-To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
-They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
-
-- spiral:  XXX $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
-- pipe: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-- sine:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-
-For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
 
@@
 
-```{r spiralScagSmoothing, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.
-For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
-With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
-For the convex index we also find bins=50 to be the better choice."}
+```{r spiralScagSmoothing, fig.cap="Comparing the traces of the three scagnostics indices when smoothing the index values, either by averaging over the index value after jittering the projection by some angle alpha (left) or after jittering the projected datapoints with some amount alpha (right).
+While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered."}
 
@@
 ggplot(scagSmoothMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=method, linetype=factor(alpha))) + 
-  facet_wrap(~variable, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
+  geom_line(aes(color=factor(alpha))) + 
+  facet_wrap(variable~method, ncol=2, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
 

### 2018-11-01 scagnostics.Rmd
more on data description

@@
 - **pipe**: nuisance directions are generated by sampling independently from a uniform distribution between $[-1,1]$, and the circle is generated by sampling randomly on a 2D circle, and adding a small radial noise.
+The circle should be easy to see by some indices because it is large structure, but the nonlinearity creates a complication.
 - **sine**:  nuisance directions are generated by sampling independently from a standard normal distribution, and the sine curve is generated by $x_n = \sin(x_{n-1}) + \mathrm{jittering}$.
+The sine is medium nonlinear structure, which should be visible by multiple indices.
 - **spiral**:  nuisance directions are generated by sampling independently from a normal distribution, and the structure directions are sampled from an Archimedean spiral, i.e.
@@
 The absolute value of $\theta$ fixes the direction of the spiral shape.
-
-Two samples sizes are also used: $n=100$ (small) and $n=1000$ (large).
+This is fine structure which is only visible close to the optimal projection.
+
+Two samples sizes are also used: $n=(100, 1000)$.
 

### 2018-11-02 scagnostics.Rmd
plots of data sets

@@
 ---
-title: An investigation into using scagnostics to find interesting projections of high-dimensional data
+title: Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics
 
@@
 keywords:
+- scagnostics
 - statistical graphics
@@
 Each generates  structure in two variables, with the remaining variables containing various types of noise.
-This is a very simple construction, because there is no need for projection pursuit to find the structure, one could simply use the scagnostics on pairs of variables.
+This is a very simple construction, because there is no need for projection pursuit to find the structure, one could simply use the PPIs on pairs of variables.
 However, it serves the purpose to also evaluate the PPIs.
@@
 
-Two samples sizes are also used: $n=(100, 1000)$.
+The structured projection is in fifth and sixth variables. Two samples sizes are also used: $n=(100, 1000)$. All variables are standardized to mean 0 and standard deviation 1.
+
+```{r datasets}
+# sample(1000:9999, 1)
+set.seed(3705)
+spiral100 <- spiralData(6, 100) %>% scale() %>% as_tibble()
+spiral1000 <- spiralData(6, 1000) %>% scale() %>% as_tibble()
+#sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
+#sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()
+
+pipe100 <- pipeData(6, 100) %>% scale() %>% as_tibble()
+pipe1000 <- pipeData(6,1000) %>% scale() %>% as_tibble()
+
+sin100 <- sinData(6, 100) %>% scale() %>% as_tibble()
+sin1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
+```
+
+Figure \ref{data} shows samples from each of the families, both nuisance and structured projections.
+
+<!--
+To get an overview of the dataset we show views of the first two directions showing the independent distribution and the last two directions showing the special view.
+The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of $x_6$ (and $x_5$ for the Pipe and Spiral distribution) will be distinct from uninformed views of distributions 2 and 3.
+-->
+
+```{r data, fig.width=6, fig.height=7, fig.cap="Scatterplots of pairs of variables from samples of each family, showing the nuisance variables and structured variables."}
+dsn <- spiral1000 %>% select(V1, V2) %>%
+  rename(X1=V1, X2=V2) %>%
+  mutate(family="spiral", type="nuisance")
+dss <- spiral1000 %>% select(V5, V6) %>%
+  rename(X1=V5, X2=V6) %>%
+  mutate(family="spiral", type="structured")
+dpn <- pipe1000 %>% select(V1, V2) %>%
+  rename(X1=V1, X2=V2) %>%
+  mutate(family="pipe", type="nuisance")
+dps <- pipe1000 %>% select(V5, V6) %>%
+  rename(X1=V5, X2=V6) %>%
+  mutate(family="pipe", type="structured")
+din <- sin1000 %>% select(V1, V2) %>%
+  rename(X1=V1, X2=V2) %>%
+  mutate(family="sine", type="nuisance")
+dis <- sin1000 %>% select(V5, V6) %>%
+  rename(X1=V5, X2=V6) %>%
+  mutate(family="sine", type="structured")
+d <- bind_rows(dsn, dss, dpn, dps, din, dis)
+ggplot(d, aes(x=X1, y=X2)) + geom_point(alpha=0.3) + 
+  facet_grid(type~family) + theme(aspect.ratio=1) +
+  xlab("") + ylab("")
+```
 
@@
 
-To understand the potential for the various indices defined we use the following overview procedures on the simulated data sets:
-
-- index values for pair of variables, explicitly comparing the uninformative and special view of each dataset, as well as discussing the dependence on the sample size, see Section \ref{sec:dataOv}
-- tour between pairs of variables, considering a planned tour that moves from the $x_1 - x_2$ view either to the $x_3 - x_4$ view (no special structure encountered over the full path) or to the $x_5 - x_6$ (moving from uninformed to special view), see Section \ref{sec:smooth}
-- optimisation path, starting from $x_1 - x_2$, how can we use the available index functions and optimisation algorithms to identify the special view, see Section \ref{sec:guided}
-- rotating a dataset with functional dependence to evaluate the rotation dependence of the index functions, see Section \ref{sec:rot}
-- benchmarking the computing time in Section \ref{sec:timer}
-
+The family of data sets can be used to assess the PPIs on the desirable properties: smoothness, squintability, flexibility, rotation invariance, and speed. The procedures are as follows:
+
+- Compare the index values for structured vs nuisance variables, by simulating many examples from the model families, and recording the 90% confidence intervals.
+This checks if the PPI can detect the structure, that it has distinct differences between the sets of values.
+- Use the tour to show the index values along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
+The result is ideally a smooth change in low values.
+This checks the smoothness property.
+- Use the tour to show the index values along an interpolation between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$.
+This examines the squintability, and smoothness.
+If the function is smooth and slowly increases towards the structured projection, then the structure is visible from a distance.
+- A guided tour, showing the optimisation path, starting from $x_1 - x_2$, for one index, shows how effectively the maximum is attained.
+The values for other PPIs is computed for this path also, which allows an assessment of similarity between PPIs.
+- Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
+- Benchmarking is used to assess the computational speed.
+
+<!--
 \textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
@@
 or instead of it?}
-
-### Dataset overview
-\label{sec:dataOv}
-
-To test the behaviour of the index functions we use three types of simulated data sets based on different underlying distributions, and with different built-in features.
-They are defined in $n=6$ dimensions with $p$ randomly selected points $x_i, i = 1,..,n$ according to:
-
-- spiral:  XXX $p$ points on an n-dimensional sphere randomly selected using the geozoo package [@geozoo], as an example of dataset without special views, labelled "sphere"
-- pipe: $p$ points where $n-1$ points are independently drawn from a uniform distribution between $[-1,1]$, and enforcing $x_n^2 + x_{n-1}^2 = 1 \pm 0.1$ by rejection sampling as an example of a dataset with special view that is not described by functional dependence, and where sampling boundaries are apparent in the distribution, labelled "pipe"
-- sine:  $p$ points where $n-1$ points are independently drawn from a normal distribution with mean 0 and variance 1, and with $x_n = \sin(x_{n-1}) + \mathrm{jittering}$ as an example dataset with special view described by a function, labelled "sine"
-
-For each distribution we consider a small sample of $p=100$ and a large sample with $p=1000$.
+-->
 
@@
 
-```{r datasets}
-set.seed(1984)
-spiral100 <- spiralData(6, 100) %>% scale() %>% as_tibble()
-spiral1000 <- spiralData(6, 1000) %>% scale() %>% as_tibble()
-#sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
-#sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()
-
-pipe100 <- pipeData(6, 100) %>% scale() %>% as_tibble()
-pipe1000 <- pipeData(6,1000) %>% scale() %>% as_tibble()
-
-sin100 <- sinData(6, 100) %>% scale() %>% as_tibble()
-sin1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
-```
-
-To get an overview of the dataset we show views of the first two directions showing the independent distribution and the last two directions showing the special view.
-The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of $x_6$ (and $x_5$ for the Pipe and Spiral distribution) will be distinct from uninformed views of distributions 2 and 3.
-
 ```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
-showing a typical uninformed view for each dataset."}
+showing a typical uninformed view for each dataset.", eval=FALSE}
 pL <- list()
@@
 
-```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters $x_5$ and $x_6$, i.e. showing the special view built into the datasets."}
+```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters $x_5$ and $x_6$, i.e.
+showing the special view built into the datasets.", eval=FALSE}
 pL <- list()

### 2018-11-02 scagnostics.Rmd
plots of data sets
re-titled assessment sections

@@
 
+```{r util}
+#defining index functions to be used with the tour
+scagIndex <- function(scagType){
+  function(mat){
+    sR <- scagnostics.default(mat[,1],mat[,2])$s
+    return(sR[scagType])
+  }
+}
+
+scagIndexNbin <- function(scagType, n){
+  function(mat){
+    sR <- scagnostics.default(mat[,1], mat[,2], bins = n)$s
+    return(sR[scagType])
+  }
+}
+
+invConvexIndex <- function(){
+  function(mat){
+    sR <- scagnostics.default(mat[,1],mat[,2])$s
+    return(1 - sR["Convex"])
+  }
+}
+
+splineIndex <- function(){
+  function(mat){
+    return(splines2d(mat[,1], mat[,2]))
+  }
+}
+
+dcorIndex <- function(){
+  function(mat){
+    return(dcor2d(mat[,1], mat[,2]))
+  }
+}
+
+mineIndex <- function(mineIndex){
+  function(mat){
+    return(mine(mat[,1], mat[,2])[[mineIndex]])
+  }
+}
+
+mineIndexE <- function(mineIndex){
+  function(mat){
+    return(mine(mat[,1], mat[,2], est = "mic_e")[[mineIndex]])
+  }
+}
+
+mineIndexAlpha <- function(mineIndex, alpha){
+  function(mat){
+    return(mine(mat[,1], mat[,2], alpha=alpha)[[mineIndex]])
+  }
+}
+```
+
+```{r datasetFunctions}
+sphereData <- function(n, p){
+  dRet <- geozoo::sphere.solid.random(n,p)
+  return(as.tibble(dRet$points))
+}
+
+pipeData <- function(n, p){
+  i <- 1
+  dRet <- NULL
+  while(i <= p){
+    v <- runif(n, -1, 1)
+    if(abs(v[n-1]*v[n-1] + v[n]*v[n] - 1) < 0.1){
+      dRet <- rbind(dRet, v)
+      i <- i+1
+    }
+  }
+  return(as.tibble(dRet))
+}
+
+sinData <- function(n, p){
+  vName <- paste0("V",n)
+  vNameM1 <- paste0("V",n-1)
+  expr <- paste0(vName,"=sin(",vNameM1,")") # need string expression if I want to use tibble here
+  dRet <- as.tibble(matrix(rnorm((n-1)*p), ncol=(n-1))) #generate normal distributed n-1 dim data
+  dRet <- mutate_(dRet, expr) #string evaluation calculates var(n) as tan(var(n-1))
+  colnames(dRet)[n] <- vName #correct name of new variable
+  dRet[vName] <- jitter(dRet[[vName]]) #adding noise
+  return(dRet)
+}
+
+spiralData <- function(n, p){
+  i <- 1
+  a <- 0.1
+  b <- 0.1
+  dRet <- NULL
+  while(i <= p){
+    v <- rnorm(n-2)
+    theta <- abs(rnorm(1,0,2*pi))
+    r <- a + b * theta
+    x <- r * cos(theta)
+    y <- r * sin(theta)
+    v <- c(v, x, y)
+    dRet <- rbind(dRet, v)
+    i <- i+1
+  }
+  return(dRet)
+}
+
+```
+
 ```{r datasets}
@@
 
-Figure \ref{data} shows samples from each of the families, both nuisance and structured projections.
+Figure \ref{data} shows samples from each of the families, of the nuisance and structured pairs of variables.
 
@@
 
-```{r util}
-#defining index functions to be used with the tour
-scagIndex <- function(scagType){
-  function(mat){
-    sR <- scagnostics.default(mat[,1],mat[,2])$s
-    return(sR[scagType])
-  }
-}
-
-scagIndexNbin <- function(scagType, n){
-  function(mat){
-    sR <- scagnostics.default(mat[,1], mat[,2], bins = n)$s
-    return(sR[scagType])
-  }
-}
-
-invConvexIndex <- function(){
-  function(mat){
-    sR <- scagnostics.default(mat[,1],mat[,2])$s
-    return(1 - sR["Convex"])
-  }
-}
-
-splineIndex <- function(){
-  function(mat){
-    return(splines2d(mat[,1], mat[,2]))
-  }
-}
-
-dcorIndex <- function(){
-  function(mat){
-    return(dcor2d(mat[,1], mat[,2]))
-  }
-}
-
-mineIndex <- function(mineIndex){
-  function(mat){
-    return(mine(mat[,1], mat[,2])[[mineIndex]])
-  }
-}
-
-mineIndexE <- function(mineIndex){
-  function(mat){
-    return(mine(mat[,1], mat[,2], est = "mic_e")[[mineIndex]])
-  }
-}
-
-mineIndexAlpha <- function(mineIndex, alpha){
-  function(mat){
-    return(mine(mat[,1], mat[,2], alpha=alpha)[[mineIndex]])
-  }
-}
-```
-
-```{r datasetFunctions}
-sphereData <- function(n, p){
-  dRet <- geozoo::sphere.solid.random(n,p)
-  return(as.tibble(dRet$points))
-}
-
-pipeData <- function(n, p){
-  i <- 1
-  dRet <- NULL
-  while(i <= p){
-    v <- runif(n, -1, 1)
-    if(abs(v[n-1]*v[n-1] + v[n]*v[n] - 1) < 0.1){
-      dRet <- rbind(dRet, v)
-      i <- i+1
-    }
-  }
-  return(as.tibble(dRet))
-}
-
-sinData <- function(n, p){
-  vName <- paste0("V",n)
-  vNameM1 <- paste0("V",n-1)
-  expr <- paste0(vName,"=sin(",vNameM1,")") # need string expression if I want to use tibble here
-  dRet <- as.tibble(matrix(rnorm((n-1)*p), ncol=(n-1))) #generate normal distributed n-1 dim data
-  dRet <- mutate_(dRet, expr) #string evaluation calculates var(n) as tan(var(n-1))
-  colnames(dRet)[n] <- vName #correct name of new variable
-  dRet[vName] <- jitter(dRet[[vName]]) #adding noise
-  return(dRet)
-}
-
-spiralData <- function(n, p){
-  i <- 1
-  a <- 0.1
-  b <- 0.1
-  dRet <- NULL
-  while(i <= p){
-    v <- rnorm(n-2)
-    theta <- abs(rnorm(1,0,2*pi))
-    r <- a + b * theta
-    x <- r * cos(theta)
-    y <- r * sin(theta)
-    v <- c(v, x, y)
-    dRet <- rbind(dRet, v)
-    i <- i+1
-  }
-  return(dRet)
-}
-
-```
-
 <!-- I think you need to standardise the scale of the data -->
@@
 ```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
-Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine (and Spiral) dataset the parameter $x_6$ (and $x_5$) follows a very distinct distribution."}
+Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine (and Spiral) dataset the parameter $x_6$ (and $x_5$) follows a very distinct distribution.", eval=FALSE}
 pL <- list()
@@
 
+Table \ref{indexTable} compares the PPIs for structured projections against those for nuisance variables. XXX adjust the text below once re-working of table is done.
 We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
@@
 
-## Index traces over sequence of interpolated projections
+## Index traces over tour sequence of interpolated nuisance projections
 \label{sec:smooth}
@@
 
-### Random fluctuations
 We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$ and using the default interpolation angle of 0.05.
@@
 
-### Moving into informative projections
+## Index traces over tour sequence between nuisance and structured projections
+\label{sec:squintability}
+
 We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function.

### 2018-11-02 scagnostics.Rmd
Collecting data for updated Table 1, still needs formatting.

@@
 
+FIXME yes/no labels appear to be inverted?
+
 ```{r indexTable, echo=FALSE}
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-indexOverview <- tibble(convex=numeric(),
-               skinny=numeric(),
-               stringy=numeric(),
-               dcor2d=numeric(),
-               splines2d=numeric(),
-               MIC=numeric(),
-               TIC=numeric(),
-               holes=numeric(),
-               name=character(),
-               vars=character())
-dsL <- list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)
-for(ds in dsL){
-  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
-    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
-    scagRes <- scagnostics(dprj)
-    dcorRes <- dcor2d(dprj[,1], dprj[,2])
-    splineRes <- splines2d(dprj[,1], dprj[,2])
-    mineRes <- mine(dprj[,1], dprj[,2])
-    holesRes <- holes(dprj)
-    indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
-                    skinny=scagRes[,"Skinny"], stringy=scagRes[,"Stringy"],
-                    dcor2d=dcorRes, splines2d=splineRes,
-                    MIC=mineRes$MIC, TIC=mineRes$TIC, holes=holesRes,
-                    name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
-  }
-  i <- i+1
-}
-knitr::kable(indexOverview,  caption = "Summary of index values.", digits=2) 
+tableTemp <- function(){
+  tibble(
+    index=character(),
+    data=character(),
+    size=numeric(),
+    structured=character(),
+    value5=numeric(),
+    value95=numeric()
+  )
+}
+
+getSample <- function(dataT, size){
+  if (dataT == "Pipe"){
+    ret <- pipeData(p = size, n = 6) %>% scale() %>% as_tibble()
+  }
+  if (dataT == "Spiral"){
+    ret <- spiralData(p = size, n = 6) %>% scale() %>% as_tibble()
+  }
+  if (dataT == "Sine"){
+    ret <- sinData(p = size, n = 6) %>% scale() %>% as_tibble()
+  }
+  return(ret)
+}
+
+getIdx <- function(dataT, size, structured){
+  convex <- numeric(length = 100)
+  skinny <- numeric(length = 100)
+  stringy <- numeric(length = 100)
+  micidx <- numeric(length = 100)
+  ticidx <- numeric(length = 100)
+  dcor2D <- numeric(length = 100)
+  splines2D <- numeric(length = 100)
+  for(i in 1:100){
+    cSample <- getSample(dataT, size)
+    if(structured){
+      x = as.numeric(cSample$V1)
+      y = as.numeric(cSample$V2)
+    } else{
+      x = as.numeric(cSample$V5)
+      y = as.numeric(cSample$V6)
+    }
+    scag <- scagnostics(x=x, y=y)$s
+    midx <- mine(x, y, est = "mic_e")
+    skinny[i] <- scag["Skinny"]
+    stringy[i] <- scag["Stringy"]
+    convex[i] <- scag["Convex"]
+    micidx[i] <- midx$MIC
+    ticidx[i] <- midx$TIC
+    dcor2D[i] <- dcor2d(x=x, y=y)
+    splines2D[i] <- splines2d(x,y)
+  }
+  lapply(c(skinny, stringy, convex, micidx, ticidx, dcor2D, splines2D), sort)
+  res <- tableTemp() %>%
+    add_row(index="Skinny", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=skinny[5], value95=skinny[95]) %>%
+    add_row(index="Stringy", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=stringy[5], value95=stringy[95]) %>%
+    add_row(index="Convex", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=convex[5], value95=convex[95]) %>%
+    add_row(index="MICe", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=micidx[5], value95=micidx[95]) %>%
+    add_row(index="TICe", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=ticidx[5], value95=ticidx[95]) %>%
+    add_row(index="dcor2D", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=dcor2D[5], value95=dcor2D[95]) %>%
+    add_row(index="splines2D", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=splines2D[5], value95=splines2D[95])
+}
+
+if(!file.exists("cache/indexTable.rda")){
+  set.seed(505)
+  res <- tableTemp()
+  for(ds in c("Pipe", "Sine", "Spiral")){
+    for (size in c(100, 1000)){
+      for (structured in c(TRUE, FALSE)){
+        res <- full_join(res,getIdx(ds, size, structured))
+      }
+    }
+  }
+  save(res, file = "cache/indexTable.rda")
+} else {
+  load("cache/indexTable.rda")
+}
+knitr::kable(res,  caption = "Summary of index values.", digits=2) 
 ```

### 2018-11-02 scagnostics.Rmd
Structured labels fixed, data added.

@@
 
-FIXME yes/no labels appear to be inverted?
-
 ```{r indexTable, echo=FALSE}
@@
     cSample <- getSample(dataT, size)
-    if(structured){
+    if(!structured){
       x = as.numeric(cSample$V1)

### 2018-11-02 scagnostics.Rmd
Fixing table 1, some formatting.

--- 2018-11-02 scagnostics.Rmd
@@
 library(minerva) #MINE indices
+library(kableExtra) #table layout
 ```
@@
 ```{r indexTable, echo=FALSE}
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-indexOverview <- tibble(convex=numeric(),
-               skinny=numeric(),
-               stringy=numeric(),
-               dcor2d=numeric(),
-               splines2d=numeric(),
-               MIC=numeric(),
-               TIC=numeric(),
-               holes=numeric(),
-               name=character(),
-               vars=character())
-dsL <- list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)
-for(ds in dsL){
-  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
-    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
-    scagRes <- scagnostics(dprj)
-    dcorRes <- dcor2d(dprj[,1], dprj[,2])
-    splineRes <- splines2d(dprj[,1], dprj[,2])
-    mineRes <- mine(dprj[,1], dprj[,2])
-    holesRes <- holes(dprj)
-    indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
-                    skinny=scagRes[,"Skinny"], stringy=scagRes[,"Stringy"],
-                    dcor2d=dcorRes, splines2d=splineRes,
-                    MIC=mineRes$MIC, TIC=mineRes$TIC, holes=holesRes,
-                    name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
-  }
-  i <- i+1
-}
-knitr::kable(indexOverview,  caption = "Summary of index values.", digits=2) 
+tableTemp <- function(){
+  tibble(
+    index=character(),
+    data=character(),
+    size=numeric(),
+    structured=character(),
+    value5=numeric(),
+    value95=numeric()
+  )
+}
+
+getSample <- function(dataT, size){
+  if (dataT == "Pipe"){
+    ret <- pipeData(p = size, n = 6) %>% scale() %>% as_tibble()
+  }
+  if (dataT == "Spiral"){
+    ret <- spiralData(p = size, n = 6) %>% scale() %>% as_tibble()
+  }
+  if (dataT == "Sine"){
+    ret <- sinData(p = size, n = 6) %>% scale() %>% as_tibble()
+  }
+  return(ret)
+}
+
+getIdx <- function(dataT, size, structured){
+  convex <- numeric(length = 100)
+  skinny <- numeric(length = 100)
+  stringy <- numeric(length = 100)
+  micidx <- numeric(length = 100)
+  ticidx <- numeric(length = 100)
+  dcor2D <- numeric(length = 100)
+  splines2D <- numeric(length = 100)
+  for(i in 1:100){
+    cSample <- getSample(dataT, size)
+    if(!structured){
+      x = as.numeric(cSample$V1)
+      y = as.numeric(cSample$V2)
+    } else{
+      x = as.numeric(cSample$V5)
+      y = as.numeric(cSample$V6)
+    }
+    scag <- scagnostics(x=x, y=y)$s
+    midx <- mine(x, y, est = "mic_e")
+    skinny[i] <- scag["Skinny"]
+    stringy[i] <- scag["Stringy"]
+    convex[i] <- scag["Convex"]
+    micidx[i] <- midx$MIC
+    ticidx[i] <- midx$TIC
+    dcor2D[i] <- dcor2d(x=x, y=y)
+    splines2D[i] <- splines2d(x,y)
+  }
+  sortedList <- lapply(list(skinny, stringy, convex, micidx, ticidx, dcor2D, splines2D), sort)
+  res <- tableTemp() %>%
+    add_row(index="Skinny", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=sortedList[[1]][5], value95=sortedList[[1]][95]) %>%
+    add_row(index="Stringy", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=sortedList[[2]][5], value95=sortedList[[2]][95]) %>%
+    add_row(index="Convex", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=sortedList[[3]][5], value95=sortedList[[3]][95]) %>%
+    add_row(index="MICe", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=sortedList[[4]][5], value95=sortedList[[4]][95]) %>%
+    add_row(index="TICe", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=sortedList[[5]][5], value95=sortedList[[5]][95]) %>%
+    add_row(index="dcor2D", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=sortedList[[6]][5], value95=sortedList[[6]][95]) %>%
+    add_row(index="splines2D", data=dataT, size=size,
+            structured=ifelse(structured, "yes", "no"),
+            value5=sortedList[[7]][5], value95=sortedList[[7]][95])
+}
+
+if(!file.exists("cache/indexTable.rda")){
+  set.seed(505)
+  res <- tableTemp()
+  for(ds in c("Pipe", "Sine", "Spiral")){
+    for (size in c(100, 1000)){
+      for (structured in c(TRUE, FALSE)){
+        res <- full_join(res,getIdx(ds, size, structured))
+      }
+    }
+  }
+  save(res, file = "cache/indexTable.rda")
+} else {
+  load("cache/indexTable.rda")
+}
+
+res <- res %>%
+  mutate(Range=paste0("[",formatC(value5, digits=2, format ="f"),",",
+                      formatC(value95, digits=2, format = "f"),"]")) %>%
+  select(-value5, -value95) %>%
+  spread(key=structured, value = Range) %>%
+  rename(nuisance=no, structured=yes) %>%
+  arrange(data) %>%
+  select(-data)
+
+kable(res,  "latex", booktabs = T, longtable = T, caption = "Summary of index values.", digits=2) %>%
+  kable_styling(latex_options = c("repeat_header")) %>%
+  group_rows("Pipe", 1, 14) %>%
+  group_rows("Sine", 15, 28) %>%
+  group_rows("Spiral", 29, 42)
 ```

### 2018-11-02 scagnostics.Rmd
got table formatted nicely

--- 2018-11-02 scagnostics.Rmd
@@
 library(minerva) #MINE indices
+library(kableExtra)
+library(forcats)
 ```
@@
     size=numeric(),
-    structured=character(),
+    type=character(),
     value5=numeric(),
@@
 
-getIdx <- function(dataT, size, structured){
+getIdx <- function(dataT, size, type){
   convex <- numeric(length = 100)
@@
     cSample <- getSample(dataT, size)
-    if(!structured){
+    if(type == "noise"){
       x = as.numeric(cSample$V1)
@@
   res <- tableTemp() %>%
-    add_row(index="Skinny", data=dataT, size=size,
-            structured=ifelse(structured, "yes", "no"),
+    add_row(index="skinny", data=dataT, size=size,
+            type=type,
             value5=skinny[5], value95=skinny[95]) %>%
-    add_row(index="Stringy", data=dataT, size=size,
-            structured=ifelse(structured, "yes", "no"),
+    add_row(index="stringy", data=dataT, size=size,
+            type=type,
             value5=stringy[5], value95=stringy[95]) %>%
-    add_row(index="Convex", data=dataT, size=size,
-            structured=ifelse(structured, "yes", "no"),
+    add_row(index="convex", data=dataT, size=size,
+            type=type,
             value5=convex[5], value95=convex[95]) %>%
     add_row(index="MICe", data=dataT, size=size,
-            structured=ifelse(structured, "yes", "no"),
+            type=type,
             value5=micidx[5], value95=micidx[95]) %>%
     add_row(index="TICe", data=dataT, size=size,
-            structured=ifelse(structured, "yes", "no"),
+            type=type,
             value5=ticidx[5], value95=ticidx[95]) %>%
     add_row(index="dcor2D", data=dataT, size=size,
-            structured=ifelse(structured, "yes", "no"),
+            type=type,
             value5=dcor2D[5], value95=dcor2D[95]) %>%
     add_row(index="splines2D", data=dataT, size=size,
-            structured=ifelse(structured, "yes", "no"),
+            type=type,
             value5=splines2D[5], value95=splines2D[95])
@@
     for (size in c(100, 1000)){
-      for (structured in c(TRUE, FALSE)){
-        res <- full_join(res,getIdx(ds, size, structured))
-      }
-    }
-  }
+      for (type in c("structure", "noise")){
+        res <- bind_rows(res,getIdx(ds, size, type))
+      }
+    }
+  }
+  res <- res %>% arrange(data, size, index, type) %>% select(data, size, index, type, value5, value95)
   save(res, file = "cache/indexTable.rda")
@@
 }
-knitr::kable(res,  caption = "Summary of index values.", digits=2) 
+resp <- res %>% filter(size == 1000, data == "Pipe") %>%
+  select(index, type, value95, value5) %>%
+  rename(lower=value95, upper=value5)
+resi <- res %>% filter(size == 1000, data == "Sine") %>%
+  select(index, type, value95, value5) %>%
+  rename(lower=value95, upper=value5)
+ress <- res %>% filter(size == 1000, data == "Spiral") %>%
+  select(index, type, value95, value5) %>%
+  rename(lower=value95, upper=value5)
+res_all <- bind_cols(resp, resi, ress) %>%
+  select(-index1, -type1, -index2, -type2) %>%
+  mutate(index = fct_relevel(index, "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe")) %>%
+  mutate(index = c("convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", ""))
+# Good resource for doing tables
+# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
+kable(res_all, format = "latex",  caption = "Comparison of index values between noise projections and structured projections for sample size 1000, using 5th and 95th percentiles from 100 simulated sets.
+Most indices have much larger values for the structured projections, except for convex.",
+      col.names=c("Index","", "lower", "upper", "lower", "upper", "lower", "upper"),
+      digits=2, booktabs = T) %>% 
+      add_header_above(c(" ", " ", "pipe" = 2, "sine" = 2, "spiral"=2)) %>%
+  group_rows("", 1, 2) %>%
+  group_rows("", 3, 4) %>%
+  group_rows("", 5, 6) %>%
+  group_rows("", 7, 8) %>%
+  group_rows("", 9, 10) %>%
+  group_rows("", 11, 12) %>%
+  group_rows("", 13, 14) %>%
+  kable_styling(position = "center")
 ```

### 2018-11-02 scagnostics.Rmd

@@
   }
-  lapply(c(skinny, stringy, convex, micidx, ticidx, dcor2D, splines2D), sort)
+  sortedList <- lapply(list(skinny, stringy, convex, micidx, ticidx, dcor2D, splines2D), sort)
   res <- tableTemp() %>%

### 2018-11-02 scagnostics.Rmd
language for table

@@
 
-Table \ref{indexTable} compares the PPIs for structured projections against those for nuisance variables. XXX adjust the text below once re-working of table is done.
+Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000.
+The lower and upper show the 5th and 95th percentile of values.
+All indexes, except convex show distinctly higher values for the structured projections.
+The convex index considers the noise projections more interesting than the structured projections, which might be expected, but also provides an argument to use (1-convex) as the PPI.
+
+<!--
 We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
@@
 Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fictional dependence.
+-->
 

### 2018-11-02 scagnostics.Rmd
re-working null path images

@@
 Small samples (n=100) are shown in red, large samples (n=1000) in blue."}
-fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
-ggplot(fullResMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size))) + 
-  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
-  guides(color=FALSE)
+fullResMelt <- fullRes %>% gather(PPI, value, -t, -name, -size)
+#ggplot(fullResMelt, aes(x=t, y=value)) +
+#  geom_line(aes(color=factor(size))) + 
+#  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
+#  guides(color=FALSE)
+fullResMelt %>% 
+  ggplot(aes(x=t, y=value)) + 
+  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
+  ylim(c(0,1)) +
+  facet_grid(PPI~name) + 
+  theme(legend.position="none")
 ```

### 2018-11-04 scagnostics.Rmd
nuisance traces

--- 2018-11-02 scagnostics.Rmd
@@
 
-<!-- I think you need to standardise the scale of the data -->
-
 ```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
@@
 
-\textcolor{red}{Maybe it will be better to combine the two traces? Can have tour from V1-V2 to V3-V4 to V5-V6, indicate position of V3-V4 in path, will save space, make direct comparison easier.}
-
+Figure \ref{fig:plotV1V2toV3V4} shows the traces representing the index values calculated across a tour between a pair of nuisance projections.
+The range of each axis is set to be the limits of the index, as might be expected over many different data sets, 0 to 1.
+Each projection in the interpolation will also be noise.
+Two different sample sizes are show, $n=100$ as a dashed line, and $n=1000$ as a solid line.
+The ideal trace is a smooth function, with relatively low values, and no difference between the sample sizes.
+A major feature to notice is that the scagnostics produce noisy functions, which is problematic, because small changes in the projection result in big jumps in the value.
+This will make them difficult to optimize.
+TIC also is noisy.
+\textcolor{red}{why is TIC so high?
+did it get inverted?} On the other hand holes, dcor2d, splines2d and MIC are relatively smooth functions.
+\textcolor{red}{why is holes so high?
+it should range between 0-1, maybe needs to be better standardised, too.} Several of the indexes are sensitive to sample size also, the same structured projection with differing numbers of points, produces different values.
+
+<!--
 We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$ and using the default interpolation angle of 0.05.
@@
 Note here that fluctuations present in Figure \ref{fig:plotV1V2toV3V4} may be amplified, as maximum values for most index functions are far from the theoretic maximum.
+-->
 
@@
 
-```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$, for all considered datasets and index functions.
-Small samples (n=100) are shown in red, large samples (n=1000) in blue."}
-fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
-ggplot(fullResMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size))) + 
-  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
-  guides(color=FALSE)
+```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between two nuisance projections.
+All projections would be nuisance so the PPI are ideally low and smooth, with little difference between sample sizes (solid lines: $n=1000$; dashed: $n=100$).
+The scagnostic PPIs are noisy, as is TIC.
+Some indexes have distinct differences in values between sample sizes."}
+fullResMelt <- fullRes %>%
+  mutate(convex = 1-convex) %>% 
+  gather(PPI, value, -t, -name, -size) %>%
+  filter(PPI != "cmass") %>%
+  mutate(PPI = fct_relevel(PPI, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+fullResMelt %>% 
+  ggplot(aes(x=t, y=value)) + 
+  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
+  ylim(c(0,1)) +
+  facet_grid(PPI~name) + 
+  theme(legend.position="none") + xlab("Sequence of projections (t)") + ylab("PPI value")
 ```
@@
 starting from random distribution view, and moving into the special view."}
-fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
-ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size))) + 
-  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
-  guides(color=FALSE)
-```
-
+#fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
+fullResMeltPlanned <- fullResPlanned %>%
+  mutate(convex = 1-convex) %>% 
+  gather(PPI, value, -t, -name, -size) %>%
+  filter(PPI != "cmass") %>%
+  mutate(PPI = fct_relevel(PPI, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+fullResMeltPlanned %>% 
+  ggplot(aes(x=t, y=value)) + 
+  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
+  ylim(c(0,1)) +
+  facet_grid(PPI~name) + 
+  theme(legend.position="none") + xlab("Sequence of projections, t") + ylab("PPI value")
+```
+
+<!--
+I don't think this section adds anything to the paper
 ## Distributions of index values
@@
   geom_density(aes(color=path, linetype = factor(size))) +
-  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
-```
+  facet_wrap(PPI~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
+```
+
+-->
 
@@
 
-ggplot(dfTimer, aes(x = n, y = t, color=name)) + 
-  geom_point(alpha=0.05) +
-  geom_point(data=timerMeans) +
-  geom_line(data=timerMeans) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(x = "Sample Size", y = "Time [s]", color = "Index family")
+ggplot(dfTimer, aes(x = n, y = t)) + 
+  geom_point(alpha=0.1, aes(color=name, shape=name)) +
+  geom_point(data=timerMeans, aes(color=name, shape=name)) +
+  geom_line(data=timerMeans, aes(linetype=name, color=name)) +
+  scale_x_log10(limits=c(90,10100)) +
+#  scale_y_log10(limits=c(0.00001,1)) + 
+  labs(x = "Sample Size", y = "Time [s]", color = "Index family", 
+       linetype = "Index family", shape = "Index family") + 
+  scale_colour_brewer(palette="Dark2")
 ```

### 2018-11-04 scagnostics.Rmd
rotation invariance

@@
 
+Figure \ref{fig:plannedtour} shows the PPIs for a tour sequence between a nuisance and structured projection.
+Sample size is indicated by linetype: dashed being $n=100$ and solid is $n=1000$.
+The beginnnig of the sequence is the nuisance projection and the end is the structured projection.
+The index values for most PPIs increases substantially nearing the structured projection, indicating that they "see" the structure.
+Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
+[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
+The holes index easily detects the pipe, and can see it from a distance.
+The scagnostic index, stringy, can see the structure but is myopic, only when it is very close.
+Interestingly the scagnostic, skinny, sees the spiral from a distance.
+
+\textcolor{red}{XXX This shouldn't go here.
+The index needs to have been introduced much earlier in the paper.
+TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
+We also see that TIC is smoother when moving into the special view, in terms of the relative differences.}
+<!--
 We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function.
@@
 \textcolor{red}{Holes index or even pca could discover pipe as well...}
+-->
 
@@
 
-```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto $x_1$-$x_2$ to the projection onto $x_5$-$x_6$, i.e.
-starting from random distribution view, and moving into the special view."}
+```{r plannedtour, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between a nuisance and a structured projection (solid lines: $n=1000$; dashed: $n=100$).
+Most of the indexes have peaks at the end of the sequence, indicating that they see the structure.
+The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection."}
 #fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
@@
   facet_grid(PPI~name) + 
-  theme(legend.position="none") + xlab("Sequence of projections, t") + ylab("PPI value")
+  theme(legend.position="none") + xlab("Sequence of projections (t)") + ylab("PPI value")
 ```
@@
 
--->
-
-## Rotation dependence
-\label{sec:rot}
 \textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$, maybe can add second path for spiral?}
@@
 Observe huge drops for some of the indexes!
-```{r rotationDep, fig.width=6, fig.height=7, fig.cap="Rotation dependence of the considered index functions tested on the sine 1000 dataset, by rotating the special view by an angle alpha."}
+-->
+
+## Rotation dependence
+\label{sec:rot}
+```{r rotationDep, fig.width=6, fig.height=7, fig.cap="PPI for rotations of the sine 1000 data, to examine rotation invariance.
+Most are close to rotation invariant, except for dcor2d, splines2d and TIC."}
 if(!file.exists("cache/rotationDep.rda")){
@@
 
-scMelt <- melt(sc, id=c("alpha"))
-ggplot(scMelt, aes(x=alpha, y=value)) +
+scMelt <- sc %>% 
+  mutate(angle=alpha*360/(2*pi), TIC = TIC/148) %>%
+  select(-alpha) %>%
+  gather(PPI, value, -angle) %>%
+  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+ggplot(scMelt, aes(x=angle, y=value)) +
   geom_line() + 
-  facet_wrap(~variable, scales = "free_y", ncol = 1) +
-  geom_blank(aes(y = 0)) +
-  geom_blank(aes(y = 1))
+  ylim(c(-1, 1.1)) +
+  facet_wrap(~PPI, ncol = 3) +
+  #geom_blank(aes(y = 0)) +
+  #geom_blank(aes(y = 1)) +
+  coord_polar() +
+  xlab("") + ylab("")
 ```

### 2018-11-14 scagnostics.Rmd
Fixes for Table 1, adding holes index.

@@
   splines2D <- numeric(length = 100)
+  holesI <- numeric(length = 100)
   for(i in 1:100){
@@
     micidx[i] <- midx$MIC
-    ticidx[i] <- midx$TIC
+    if (size == 100) {ticidx[i] <- midx$TIC / 16}
+    if (size == 1000) {ticidx[i] <- midx$TIC / 148}
     dcor2D[i] <- dcor2d(x=x, y=y)
     splines2D[i] <- splines2d(x,y)
-  }
-  sortedList <- lapply(list(skinny, stringy, convex, micidx, ticidx, dcor2D, splines2D), sort)
+    holesI[i] <- holes(matrix(c(x,y), ncol = 2))
+
+  }
+  #sortedList <- lapply(list(skinny, stringy, convex, micidx, ticidx, dcor2D, splines2D, holesI), sort)
+  skinny <- sort(skinny)
+  stringy <- sort(stringy)
+  convex <- sort(convex)
+  micidx <- sort(micidx)
+  ticidx <- sort(ticidx)
+  dcor2D <- sort(dcor2D)
+  splines2D <- sort(splines2D)
+  holesI <- sort(holesI)
   res <- tableTemp() %>%
@@
             type=type,
-            value5=splines2D[5], value95=splines2D[95])
+            value5=splines2D[5], value95=splines2D[95]) %>%
+    add_row(index="holes", data=dataT, size=size,
+            type=type,
+            value5=holesI[5], value95=holesI[95])
 }
@@
 resp <- res %>% filter(size == 1000, data == "Pipe") %>%
-  select(index, type, value95, value5) %>%
-  rename(lower=value95, upper=value5)
+  select(index, type, value5, value95) %>%
+  rename(lower=value5, upper=value95)
 resi <- res %>% filter(size == 1000, data == "Sine") %>%
-  select(index, type, value95, value5) %>%
-  rename(lower=value95, upper=value5)
+  select(index, type, value5, value95) %>%
+  rename(lower=value5, upper=value95)
 ress <- res %>% filter(size == 1000, data == "Spiral") %>%
-  select(index, type, value95, value5) %>%
-  rename(lower=value95, upper=value5)
+  select(index, type, value5, value95) %>%
+  rename(lower=value5, upper=value95)
 res_all <- bind_cols(resp, resi, ress) %>%
   select(-index1, -type1, -index2, -type2) %>%
-  mutate(index = fct_relevel(index, "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe")) %>%
-  mutate(index = c("convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", ""))
+  mutate(index = fct_relevel(index, "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe", "holes")) %>%
+  arrange(index) %>%
+  mutate(index = c("convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", "", "holes", ""))
 # Good resource for doing tables
@@
   group_rows("", 13, 14) %>%
+  group_rows("", 15, 16) %>%
   kable_styling(position = "center")

### 2018-11-14 scagnostics.Rmd
TIC normalisation fixed throughout.

@@
   }
-  maxTIC <- max(sc$TIC1)
+  if (size == 100) {maxTIC <- 16}
+  if (size == 1000) {maxTIC <- 148}
   sc <- sc %>%
@@
     select(-TIC1)
-
   return(sc)
@@
                   splines2d=splineRes, MIC=mineRes$MIC,
-                  TIC=mineRes$TIC, alpha=a)
+                  TIC=mineRes$TIC/148, alpha=a)
   }
@@
 scMelt <- sc %>% 
-  mutate(angle=alpha*360/(2*pi), TIC = TIC/148) %>%
+  mutate(angle=alpha*360/(2*pi), TIC = TIC) %>%
   select(-alpha) %>%
@@
 if(!file.exists("cache/refinepipe.rda")){
+  iLast <- length(pipeTourFull)
+  fProj <- pipeTourFull[[iLast]]
   pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)

### 2018-11-15 scagnostics.Rmd
section 3 re-organisation

@@
 If the function is smooth and slowly increases towards the structured projection, then the structure is visible from a distance.
-- A guided tour, showing the optimisation path, starting from $x_1 - x_2$, for one index, shows how effectively the maximum is attained.
-The values for other PPIs is computed for this path also, which allows an assessment of similarity between PPIs.
+- Ease of optimisation depends on having a smooth function, and this is tested using a guided tour, showing the optimisation path, starting from $x_1 - x_2$.
+One index is optimised to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, which allows an assessment of similarity between PPIs.
 - Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
-- Benchmarking is used to assess the computational speed.
+- Varying the parameters for PPIs is used to examine the effect on value, smoothness, how flexible the index can be to find different structure and whether the index is fiddly to tune.
+Trace plots between noise and structured projections are used to examine this.
+- Benchmarking on a range of sample sizes is used to assess the computational speed.
 
@@
 
-## Index traces over tour sequence between nuisance and structured projections
-\label{sec:squintability}
-
-Figure \ref{fig:plannedtour} shows the PPIs for a tour sequence between a nuisance and structured projection.
-Sample size is indicated by linetype: dashed being $n=100$ and solid is $n=1000$.
-The beginnnig of the sequence is the nuisance projection and the end is the structured projection.
-The index values for most PPIs increases substantially nearing the structured projection, indicating that they "see" the structure.
-Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
-[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
-The holes index easily detects the pipe, and can see it from a distance.
-The scagnostic index, stringy, can see the structure but is myopic, only when it is very close.
-Interestingly the scagnostic, skinny, sees the spiral from a distance.
-
-\textcolor{red}{XXX This shouldn't go here.
-The index needs to have been introduced much earlier in the paper.
-TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
-We also see that TIC is smoother when moving into the special view, in terms of the relative differences.}
-<!--
-We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function.
-Results are shown in Figure \ref{fig:plannedtour}.
-We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise.
-On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
-Finally the Spiral appears to be the most challenging, having an even smaller squint angle than the Pipe distribution.
-While "cmass" appears to be appropriate to identify this view we may actually find larger values of this index for random distributions, making it unfit to use in a guided tour.
-The scagnostics indices skinny (or 1-convex) appear to be sensitive to the structure at a larger angle compared to MINE indices, but will suffer from the large fluctuations in random views when used in a guided tour.
-Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
-
-TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
-We also see that TIC is smoother when moving into the special view, in terms of the relative differences.
-
-\textcolor{red}{Holes index or even pca could discover pipe as well...}
--->
-
 ```{r plannedtourpath}
@@
 
+## Index traces over tour sequence between nuisance and structured projections
+\label{sec:squintability}
+
+Figure \ref{fig:plannedtour} shows the PPIs for a tour sequence between a nuisance and structured projection.
+Sample size is indicated by linetype: dashed being $n=100$ and solid is $n=1000$.
+The beginnnig of the sequence is the nuisance projection and the end is the structured projection.
+The index values for most PPIs increases substantially nearing the structured projection, indicating that they "see" the structure.
+Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
+[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
+The holes index easily detects the pipe, and can see it from a distance.
+The scagnostic index, stringy, can see the structure but is myopic, only when it is very close.
+Interestingly the scagnostic, skinny, sees the spiral from a distance.
+
+\textcolor{red}{XXX This shouldn't go here.
+The index needs to have been introduced much earlier in the paper.
+TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
+We also see that TIC is smoother when moving into the special view, in terms of the relative differences.}
+<!--
+We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function.
+Results are shown in Figure \ref{fig:plannedtour}.
+We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise.
+On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
+Finally the Spiral appears to be the most challenging, having an even smaller squint angle than the Pipe distribution.
+While "cmass" appears to be appropriate to identify this view we may actually find larger values of this index for random distributions, making it unfit to use in a guided tour.
+The scagnostics indices skinny (or 1-convex) appear to be sensitive to the structure at a larger angle compared to MINE indices, but will suffer from the large fluctuations in random views when used in a guided tour.
+Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
+
+TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
+We also see that TIC is smoother when moving into the special view, in terms of the relative differences.
+
+\textcolor{red}{Holes index or even pca could discover pipe as well...}
+-->
+
 <!--
@@
 -->
+
+## Ease of optimisation
+\label{sec:guided}
+
+- remind reader of the options for optimisation in guided tour
+- how you use them here
+-
+
+The guided tour combines optimisation with interpolating between pairs of planes.
+Target planes of the path are chosen by a random search method to find projection planes with high index values, which are then interpolated between.
+It operates roughly like simulated annealing that as one gets closer to an optimal value the random search range is reduced.
+
+Given the behaviour of the various index measures under an interpolated tour path we next try to use the tourr implementation of the guided tour together with selected index functions to uncover the special views in datasets 2 and 3 with 1000 datapoints.
+
+### Looking down the pipe
+
+\textcolor{red}{Double challenging: fine structure + (coherent) noise}
+\textcolor{red}{simple standard guided tour does not find pipe easily with holes index, did not try brute force approach used with TIC index yet}
+
+As seen before (reference figure) MIC or TIC should be able to find the view down the pipe.
+The optimisation itself is however expected to be challenging because one has to be fairly close to the optimal projection to observe the increase in index value.
+We therefore suggest a two step procedure for the uncovering of such "small scale structures".
+The first step consists in "scouting" projections.
+Here we use the "search_better" optimisation available with the tourr package, with a standard search window $\alpha = 0.5$ and disabling the cooling (in practice this can be done by fixing the cooling parameter to $1$), together with a large number of tries (setting max.tries to 5000 here).
+The resulting search will quickly steer away from the most uninteresting projections, followed by a broad scan of the thus identified region.
+While the lack of cooling means that this will not converge towards the maximum, each step in this first search will increase the index value, thus it can be used to provide an informed starting plane that can be used in a second step where we use the "search_geodesic" optimization to improve the index further towards the maximum.
+The described procedure is CPU time intensive, however it is promising that we manage to uncover such small scale structure in the distribution.
+On the other hand no "guarantee" can be provided for uncovering the special view.
+Parameters for the scouting phase will need to be adapted to different examples, and with increasing number of dimensions it may be better to consider several runs with different starting projections.
+This gives some indication on how to handle difficult datasets. As we will see below most of the time standard "search_geodesic" optimisation will be efficient enough to uncover the interesting views.
+
+The results are shown below, first Fig.
+\ref{fig:pipeFirstRun} shows how the considered index values evolve in the scouting phase.
+Note that this is the interpolated path and there are small local maxima and minima along the index value, especially because we keep the step size in the guided tour large.
+We can see the clear maximum of the TIC index in the final projection, as well as clear minimum of the convex index that carries similar information On the other hand we see that MIC or the dcor2d index are not suitable for uncovering the special view here.
+
+\textcolor{red}{Sometimes index value decreases, because index is not rotationally invariant.}
+
+```{r findpipe, results="hide"}
+if(!file.exists("cache/findpipe.rda")){
+  set.seed(1984)
+  pipeResc <- as.matrix(pipe1000) 
+  pipeTour <- save_history(pipeResc,
+                          guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
+                                     cooling=1, alpha = 0.5, max.tries = 5000))
+  pipeTourFull <- as.list(interpolate(pipeTour))
+  save(pipeResc, pipeTour, pipeTourFull, file = "cache/findpipe.rda")
+} else {
+  load("cache/findpipe.rda")
+}
+```
+
+```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Index value on tour path when scouting for high TIC index values."}
+if(!file.exists("cache/pipeFirstRun.rda")){
+  pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
+  pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
+  save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
+} else {
+  load("cache/pipeFirstRun.rda")
+}
+newBasisPipeTour <- which(attributes(interpolate(pipeTour))$new_basis)
+ggplot(pipeTourResMelt, aes(x=t, y=value)) +
+  geom_line() + 
+  facet_wrap(~variable, ncol=1, scales = "free_y")+
+  geom_vline(data=as.tibble(newBasisPipeTour), mapping=aes(xintercept=value), color="red")
+```
+
+We next show in Fig \ref{fig:testpipe} (left) the final view corresponding to the highest index found in the TIC scouting phase.
+While this is clearly not the optimal view of the pipe feature, we can already identify the feature in the distribution, and we expect to be able to find the maximum when using this plane as a starting point of the geodesic optimisation.
+Indeed performing a standard guided tour with the TIC index and the identified starting plane efficiently identifies the optimal view shown in the right plot.
+In this case the optimisation quickly converges, following a smooth, monotonic increas of the TIC index.
+
+```{r refinepipe, results="hide"}
+if(!file.exists("cache/refinepipe.rda")){
+  iLast <- length(pipeTourFull)
+  fProj <- pipeTourFull[[iLast]]
+  pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
+  pipeTourFull2 <- as.list(interpolate(pipeTour2))
+  pipeTourRes2 <- getProj(pipe1000, pipeTourFull2, "Pipe", 1000)
+  pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
+  save(pipeTour2, pipeTourFull2, pipeTourRes2, pipeTourResMelt2, file = "cache/refinepipe.rda")
+} else {
+  load("cache/refinepipe.rda")
+}
+```
+
+```{r testpipe,  fig.height=4, fig.cap="Final view discovered by the scouting phase (left) and after second optimisation (right)."}
+iLast <- length(pipeTourFull)
+fProj <- pipeTourFull[[iLast]]
+dProj <- as.tibble(pipeResc %*% fProj)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point()
+iLast <- length(pipeTourFull2)
+fProj <- pipeTourFull2[[iLast]]
+dProj <- as.tibble(pipeResc %*% fProj)
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point()
+grid.arrange(p1, p2, ncol=2)
+```
+
+### Finding sine waves
+
+We next work with distribution 3 and aim to identify the view showing functional dependence using the splines2d index.
+Given the results from the planned tour we expect to be able to efficiently identify this view without relying on the two step procedure described above.
+Indeed the guided tour quickly converges in this case, the evolution of the index functions is shown in Fig \ref{fig:findsine}.
+Here there are four index functions that appear to be suitable for the optimisation: splines2d, dcor2d, MIC and TIC.
+On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimisation when considering them as index functions.
+The final view identified by the guided tour is shown in Fig \ref{}
+
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide"}
+set.seed(2018)
+if(!file.exists("cache/findsine.rda")){
+  sineResc <- as.matrix(sin1000)
+  sineTour <- save_history(sineResc,
+                          guided_tour(splineIndex()))
+  sineTourFull <- as.list(interpolate(sineTour))
+  sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
+  sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
+  save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
+} else {
+  load("cache/findsine.rda")
+}
+ggplot(sineTourResMelt, aes(x=t, y=value)) +
+  geom_line() + 
+  facet_wrap(~variable, ncol=1, scales = "free_y")
+```
+
+```{r testsine, out.width=".5\\textwidth", fig.cap="Final view identified by the optimisation."}
+iLast <- length(sineTourFull)
+fProj <- sineTourFull[[iLast]]
+dProj <- as.tibble(sineResc %*% fProj)
+ggplot(dProj, aes(V1,V2)) + geom_point()
+```
+
+### Spiral detection
+
+\textcolor{red}{can use skinny to find sprial in 3 dimensions, but not higher, scouting method does not work because of large rotation dependence, maybe make it study of guided tour success vs number of parameters entering?
+e.g.
+given n starting planes how many times was view in plane V5-V6 identified as function of number of dimensions and index used?}
+
+\textcolor{red}{what about role of optimisation here?
+search\_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
+(but may also give same issues of rotation dependence) ...
+tried  but didnt really work...}
+
+```{r findspiral, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the skinny index for the Sprial 1000 dataset.", results="hide"}
+#getResWithSeed <- function(s){
+#  set.seed(s)
+#  startM <- orthonormalise(matrix(runif(12),ncol = 2))
+#  tourRes <- save_history(spiralResc, guided_tour(scagIndex("Skinny")), max_bases = 10000)
+#}
+#set.seed(202020)
+#if(!file.exists("cache/findsine.rda")){
+#  spiralResc <- as.matrix(spiral1000)
+#spiralTour <- save_history(spiralResc, guided_tour(dcorIndex()))
+#spiralTour <- save_history(spiralResc, guided_tour(scagIndex("Skinny"), max.tries = 100, n=10, stepS=0.5), max_bases = 10000)
+  #spiralTour <- save_history(spiralResc,guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
+  #                                   alpha = 0.5, max.tries = 500), max_bases = 10000)
+  #spiralTourFull <- as.list(interpolate(spiralTour))
+  #spiralTourRes <- getProj(spiral1000, spiralTourFull, "Spiral", 1000)
+  #spiralTourResMelt <- melt(spiralTourRes, id=c("t", "name", "size"))
+#  save(spiralResc, spiralTour, spiralTourFull, spiralTourRes, spiralTourResMelt,
+#       file = "cache/findspiral.rda")
+#} else {
+#  load("cache/findsine.rda")
+#}
+#ggplot(spiralTourResMelt, aes(x=t, y=value)) +
+#  geom_line() + 
+#  facet_wrap(~variable, ncol=1, scales = "free_y")
+#iL <- length(as.list(spiralTour))
+#dProj <- spiralResc %*% as.list(spiralTour)[[iL]]
+#ggplot(as.tibble(dProj), aes(V1, V2)) + geom_point()
+```
 
@@
 
-## Index function parameters
+## Effect of parameter choice in nndex value
 \label{sec:params}
@@
 grid.arrange(pL[[1]], pL[[2]], pL[[3]], ncol=3)
-```
-
-## Guided tour
-\label{sec:guided}
-
-Given the behaviour of the various index measures under an interpolated tour path we next try to use the tourr implementation of the guided tour together with selected index functions to uncover the special views in datasets 2 and 3 with 1000 datapoints.
-
-### Looking down the pipe
-
-\textcolor{red}{Double challenging: fine structure + (coherent) noise}
-\textcolor{red}{simple standard guided tour does not find pipe easily with holes index, did not try brute force approach used with TIC index yet}
-
-As seen before (reference figure) MIC or TIC should be able to find the view down the pipe.
-The optimisation itself is however expected to be challenging because one has to be fairly close to the optimal projection to observe the increase in index value.
-We therefore suggest a two step procedure for the uncovering of such "small scale structures".
-The first step consists in "scouting" projections.
-Here we use the "search_better" optimisation available with the tourr package, with a standard search window $\alpha = 0.5$ and disabling the cooling (in practice this can be done by fixing the cooling parameter to $1$), together with a large number of tries (setting max.tries to 5000 here).
-The resulting search will quickly steer away from the most uninteresting projections, followed by a broad scan of the thus identified region.
-While the lack of cooling means that this will not converge towards the maximum, each step in this first search will increase the index value, thus it can be used to provide an informed starting plane that can be used in a second step where we use the "search_geodesic" optimization to improve the index further towards the maximum.
-The described procedure is CPU time intensive, however it is promising that we manage to uncover such small scale structure in the distribution.
-On the other hand no "guarantee" can be provided for uncovering the special view.
-Parameters for the scouting phase will need to be adapted to different examples, and with increasing number of dimensions it may be better to consider several runs with different starting projections.
-This gives some indication on how to handle difficult datasets. As we will see below most of the time standard "search_geodesic" optimisation will be efficient enough to uncover the interesting views.
-
-The results are shown below, first Fig.
-\ref{fig:pipeFirstRun} shows how the considered index values evolve in the scouting phase.
-Note that this is the interpolated path and there are small local maxima and minima along the index value, especially because we keep the step size in the guided tour large.
-We can see the clear maximum of the TIC index in the final projection, as well as clear minimum of the convex index that carries similar information On the other hand we see that MIC or the dcor2d index are not suitable for uncovering the special view here.
-
-\textcolor{red}{Sometimes index value decreases, because index is not rotationally invariant.}
-
-```{r findpipe, results="hide"}
-if(!file.exists("cache/findpipe.rda")){
-  set.seed(1984)
-  pipeResc <- as.matrix(pipe1000) 
-  pipeTour <- save_history(pipeResc,
-                          guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
-                                     cooling=1, alpha = 0.5, max.tries = 5000))
-  pipeTourFull <- as.list(interpolate(pipeTour))
-  save(pipeResc, pipeTour, pipeTourFull, file = "cache/findpipe.rda")
-} else {
-  load("cache/findpipe.rda")
-}
-```
-
-```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Index value on tour path when scouting for high TIC index values."}
-if(!file.exists("cache/pipeFirstRun.rda")){
-  pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
-  pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
-  save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
-} else {
-  load("cache/pipeFirstRun.rda")
-}
-newBasisPipeTour <- which(attributes(interpolate(pipeTour))$new_basis)
-ggplot(pipeTourResMelt, aes(x=t, y=value)) +
-  geom_line() + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")+
-  geom_vline(data=as.tibble(newBasisPipeTour), mapping=aes(xintercept=value), color="red")
-```
-
-We next show in Fig \ref{fig:testpipe} (left) the final view corresponding to the highest index found in the TIC scouting phase.
-While this is clearly not the optimal view of the pipe feature, we can already identify the feature in the distribution, and we expect to be able to find the maximum when using this plane as a starting point of the geodesic optimisation.
-Indeed performing a standard guided tour with the TIC index and the identified starting plane efficiently identifies the optimal view shown in the right plot.
-In this case the optimisation quickly converges, following a smooth, monotonic increas of the TIC index.
-
-```{r refinepipe, results="hide"}
-if(!file.exists("cache/refinepipe.rda")){
-  iLast <- length(pipeTourFull)
-  fProj <- pipeTourFull[[iLast]]
-  pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
-  pipeTourFull2 <- as.list(interpolate(pipeTour2))
-  pipeTourRes2 <- getProj(pipe1000, pipeTourFull2, "Pipe", 1000)
-  pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
-  save(pipeTour2, pipeTourFull2, pipeTourRes2, pipeTourResMelt2, file = "cache/refinepipe.rda")
-} else {
-  load("cache/refinepipe.rda")
-}
-```
-
-```{r testpipe,  fig.height=4, fig.cap="Final view discovered by the scouting phase (left) and after second optimisation (right)."}
-iLast <- length(pipeTourFull)
-fProj <- pipeTourFull[[iLast]]
-dProj <- as.tibble(pipeResc %*% fProj)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point()
-iLast <- length(pipeTourFull2)
-fProj <- pipeTourFull2[[iLast]]
-dProj <- as.tibble(pipeResc %*% fProj)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point()
-grid.arrange(p1, p2, ncol=2)
-```
-
-### Finding sine waves
-
-We next work with distribution 3 and aim to identify the view showing functional dependence using the splines2d index.
-Given the results from the planned tour we expect to be able to efficiently identify this view without relying on the two step procedure described above.
-Indeed the guided tour quickly converges in this case, the evolution of the index functions is shown in Fig \ref{fig:findsine}.
-Here there are four index functions that appear to be suitable for the optimisation: splines2d, dcor2d, MIC and TIC.
-On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimisation when considering them as index functions.
-The final view identified by the guided tour is shown in Fig \ref{}
-
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide"}
-set.seed(2018)
-if(!file.exists("cache/findsine.rda")){
-  sineResc <- as.matrix(sin1000)
-  sineTour <- save_history(sineResc,
-                          guided_tour(splineIndex()))
-  sineTourFull <- as.list(interpolate(sineTour))
-  sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
-  sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
-  save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
-} else {
-  load("cache/findsine.rda")
-}
-ggplot(sineTourResMelt, aes(x=t, y=value)) +
-  geom_line() + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")
-```
-
-```{r testsine, out.width=".5\\textwidth", fig.cap="Final view identified by the optimisation."}
-iLast <- length(sineTourFull)
-fProj <- sineTourFull[[iLast]]
-dProj <- as.tibble(sineResc %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point()
-```
-
-### Spiral detection
-
-\textcolor{red}{can use skinny to find sprial in 3 dimensions, but not higher, scouting method does not work because of large rotation dependence, maybe make it study of guided tour success vs number of parameters entering?
-e.g.
-given n starting planes how many times was view in plane V5-V6 identified as function of number of dimensions and index used?}
-
-\textcolor{red}{what about role of optimisation here?
-search\_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
-(but may also give same issues of rotation dependence) ...
-tried  but didnt really work...}
-
-```{r findspiral, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the skinny index for the Sprial 1000 dataset.", results="hide"}
-#getResWithSeed <- function(s){
-#  set.seed(s)
-#  startM <- orthonormalise(matrix(runif(12),ncol = 2))
-#  tourRes <- save_history(spiralResc, guided_tour(scagIndex("Skinny")), max_bases = 10000)
-#}
-#set.seed(202020)
-#if(!file.exists("cache/findsine.rda")){
-#  spiralResc <- as.matrix(spiral1000)
-#spiralTour <- save_history(spiralResc, guided_tour(dcorIndex()))
-#spiralTour <- save_history(spiralResc, guided_tour(scagIndex("Skinny"), max.tries = 100, n=10, stepS=0.5), max_bases = 10000)
-  #spiralTour <- save_history(spiralResc,guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
-  #                                   alpha = 0.5, max.tries = 500), max_bases = 10000)
-  #spiralTourFull <- as.list(interpolate(spiralTour))
-  #spiralTourRes <- getProj(spiral1000, spiralTourFull, "Spiral", 1000)
-  #spiralTourResMelt <- melt(spiralTourRes, id=c("t", "name", "size"))
-#  save(spiralResc, spiralTour, spiralTourFull, spiralTourRes, spiralTourResMelt,
-#       file = "cache/findspiral.rda")
-#} else {
-#  load("cache/findsine.rda")
-#}
-#ggplot(spiralTourResMelt, aes(x=t, y=value)) +
-#  geom_line() + 
-#  facet_wrap(~variable, ncol=1, scales = "free_y")
-#iL <- length(as.list(spiralTour))
-#dProj <- spiralResc %*% as.list(spiralTour)[[iL]]
-#ggplot(as.tibble(dProj), aes(V1, V2)) + geom_point()
 ```

### 2018-11-15 scagnostics.Rmd
Rearranging table 1.

--- 2018-11-14 scagnostics.Rmd
@@
   select(-index1, -type1, -index2, -type2) %>%
-  mutate(index = fct_relevel(index, "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe", "holes")) %>%
+  mutate(index = fct_relevel(index, "holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe")) %>%
   arrange(index) %>%
-  mutate(index = c("convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", "", "holes", ""))
+  mutate(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", ""))
 # Good resource for doing tables

### 2018-11-21 scagnostics.Rmd
tidied data section

@@
 
-The structured projection is in fifth and sixth variables. Two samples sizes are also used: $n=(100, 1000)$. All variables are standardized to mean 0 and standard deviation 1.
+The structured projection is in variables 5 and 6. Two samples sizes are used: $n=(100, 1000)$. All variables are standardized to have mean 0 and standard deviation 1.
 
@@
 ```
-
-Figure \ref{data} shows samples from each of the families, of the nuisance and structured pairs of variables.
 
@@
 ```
-
-### Property assessment
-
-The family of data sets can be used to assess the PPIs on the desirable properties: smoothness, squintability, flexibility, rotation invariance, and speed. The procedures are as follows:
-
-- Compare the index values for structured vs nuisance variables, by simulating many examples from the model families, and recording the 90% confidence intervals.
-This checks if the PPI can detect the structure, that it has distinct differences between the sets of values.
-- Use the tour to show the index values along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
-The result is ideally a smooth change in low values.
-This checks the smoothness property.
-- Use the tour to show the index values along an interpolation between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$.
-This examines the squintability, and smoothness.
-If the function is smooth and slowly increases towards the structured projection, then the structure is visible from a distance.
-- Ease of optimisation depends on having a smooth function, and this is tested using a guided tour, showing the optimisation path, starting from $x_1 - x_2$.
-One index is optimised to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, which allows an assessment of similarity between PPIs.
-- Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
-- Varying the parameters for PPIs is used to examine the effect on value, smoothness, how flexible the index can be to find different structure and whether the index is fiddly to tune.
-Trace plots between noise and structured projections are used to examine this.
-- Benchmarking on a range of sample sizes is used to assess the computational speed.
-
-<!--
-\textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
-truncated Gaussian distributions with approximately identity covariance matrix, such that they don't appear in any of the principal component planes.
-In particular seems that spiral is hardest to identify based on results from Posse, maybe I can add this to my sphere dataset?
-or instead of it?}
--->
-
-```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
-showing a typical uninformed view for each dataset.", eval=FALSE}
-pL <- list()
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
-  pC <- ggplot(ds, aes(V1, V2)) +
-    geom_point() +
-    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
-  pL[[i]] <- pC
-  i <- i+1
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
-             ncol=2)
-```
-
-```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters $x_5$ and $x_6$, i.e.
-showing the special view built into the datasets.", eval=FALSE}
-pL <- list()
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
-  pC <- ggplot(ds, aes(V5, V6)) +
-    geom_point() +
-    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
-  pL[[i]] <- pC
-  i <- i+1
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
-             ncol=2)
-```
-
-<!--
-its possible to do a single figure legend - need to look up code for this
--->
-
-```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
-Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine (and Spiral) dataset the parameter $x_6$ (and $x_5$) follows a very distinct distribution.", eval=FALSE}
-pL <- list()
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
-  distData <- melt(ds)
-  pC <- ggplot(distData, aes(value, color=variable)) +
-    geom_density() +
-    ggtitle(pLabels[i])
-  pL[[i]] <- pC
-  i <- i+1
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
-             ncol=2)
-```
-
-Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000.
-The lower and upper show the 5th and 95th percentile of values.
-All indexes, except convex show distinctly higher values for the structured projections.
-The convex index considers the noise projections more interesting than the structured projections, which might be expected, but also provides an argument to use (1-convex) as the PPI.
-
-<!--
-We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
-For the Spiral distribution the scagnostics indices skinny and stringy are large and close to their theoretic maximum for the special view.
-On the other hand while dcor2d is larger is also increased for the special view, the value is far from the theoretic maximum and larger values may be found in other views.
-MIC and TIC are also found to be sensitive to the spiral distribution.
-Considering next the Pipe distribution, we see that skinny, stringy and dcor2d are increased for the special view, but in particular MIC and TIC index are much larger and may be used to detect this type of structure in the dataset.
-Moreover, minimising the convex index (or maximising 1-convex) should also allow us to detect the special view.
-Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fictional dependence.
--->
 
@@
 
+Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables.
+Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000.
+The lower and upper show the 5th and 95th percentile of values.
+All indexes, except convex show distinctly higher values for the structured projections.
+The convex index considers the noise projections more interesting than the structured projections, which might be expected, but also provides an argument to use (1-convex) as the PPI.
+
+### Property assessment
+
+The family of data sets can be used to assess the PPIs on the desirable properties: smoothness, squintability, flexibility, rotation invariance, and speed. The procedures are as follows:
+
+- Compare the index values for structured vs nuisance variables, by simulating many examples from the model families, and recording the 90% confidence intervals.
+This checks if the PPI can detect the structure, that it has distinct differences between the sets of values.
+- Use the tour to show the index values along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
+The result is ideally a smooth change in low values.
+This checks the smoothness property.
+- Use the tour to show the index values along an interpolation between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$.
+This examines the squintability, and smoothness.
+If the function is smooth and slowly increases towards the structured projection, then the structure is visible from a distance.
+- Ease of optimisation depends on having a smooth function, and this is tested using a guided tour, showing the optimisation path, starting from $x_1 - x_2$.
+One index is optimised to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, which allows an assessment of similarity between PPIs.
+- Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
+- Varying the parameters for PPIs is used to examine the effect on value, smoothness, how flexible the index can be to find different structure and whether the index is fiddly to tune.
+Trace plots between noise and structured projections are used to examine this.
+- Benchmarking on a range of sample sizes is used to assess the computational speed.
+
+<!--
+\textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
+truncated Gaussian distributions with approximately identity covariance matrix, such that they don't appear in any of the principal component planes.
+In particular seems that spiral is hardest to identify based on results from Posse, maybe I can add this to my sphere dataset?
+or instead of it?}
+-->
+
+```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
+showing a typical uninformed view for each dataset.", eval=FALSE}
+pL <- list()
+i <- 1
+pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
+  pC <- ggplot(ds, aes(V1, V2)) +
+    geom_point() +
+    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
+  pL[[i]] <- pC
+  i <- i+1
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
+             ncol=2)
+```
+
+```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters $x_5$ and $x_6$, i.e.
+showing the special view built into the datasets.", eval=FALSE}
+pL <- list()
+i <- 1
+pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
+  pC <- ggplot(ds, aes(V5, V6)) +
+    geom_point() +
+    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
+  pL[[i]] <- pC
+  i <- i+1
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
+             ncol=2)
+```
+
+<!--
+its possible to do a single figure legend - need to look up code for this
+-->
+
+```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
+Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine (and Spiral) dataset the parameter $x_6$ (and $x_5$) follows a very distinct distribution.", eval=FALSE}
+pL <- list()
+i <- 1
+pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
+for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
+  distData <- melt(ds)
+  pC <- ggplot(distData, aes(value, color=variable)) +
+    geom_density() +
+    ggtitle(pLabels[i])
+  pL[[i]] <- pC
+  i <- i+1
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
+             ncol=2)
+```
+
+<!--
+We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
+For the Spiral distribution the scagnostics indices skinny and stringy are large and close to their theoretic maximum for the special view.
+On the other hand while dcor2d is larger is also increased for the special view, the value is far from the theoretic maximum and larger values may be found in other views.
+MIC and TIC are also found to be sensitive to the spiral distribution.
+Considering next the Pipe distribution, we see that skinny, stringy and dcor2d are increased for the special view, but in particular MIC and TIC index are much larger and may be used to detect this type of structure in the dataset.
+Moreover, minimising the convex index (or maximising 1-convex) should also allow us to detect the special view.
+Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fictional dependence.
+-->
+
 ## Index traces over tour sequence of interpolated nuisance projections
@@
 
+<!--
 \textcolor{red}{XXX This shouldn't go here.
@@
 We also see that TIC is smoother when moving into the special view, in terms of the relative differences.}
+-->
+
 <!--
@@
 
-## Effect of parameter choice in nndex value
+## Effect of parameter choice in index value
 \label{sec:params}

### 2018-11-21 scagnostics.Rmd
tinkering with property assessment overview

@@
 
-- Compare the index values for structured vs nuisance variables, by simulating many examples from the model families, and recording the 90% confidence intervals.
+<!-- - Compare the index values for structured vs nuisance variables, by simulating many examples from the model families, and recording the 90% confidence intervals.
 This checks if the PPI can detect the structure, that it has distinct differences between the sets of values.
-- Use the tour to show the index values along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
+-->
+1.
+Use the tour to show the index values along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
 The result is ideally a smooth change in low values.
 This checks the smoothness property.
-- Use the tour to show the index values along an interpolation between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$.
+2.
+Use the tour to show the index values along an interpolation between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$.
 This examines the squintability, and smoothness.
 If the function is smooth and slowly increases towards the structured projection, then the structure is visible from a distance.
-- Ease of optimisation depends on having a smooth function, and this is tested using a guided tour, showing the optimisation path, starting from $x_1 - x_2$.
+3.
+Ease of optimisation depends on having a smooth function, and this is tested using a guided tour, showing the optimisation path, starting from $x_1 - x_2$.
 One index is optimised to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, which allows an assessment of similarity between PPIs.
-- Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
-- Varying the parameters for PPIs is used to examine the effect on value, smoothness, how flexible the index can be to find different structure and whether the index is fiddly to tune.
+4.Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
+5.
+Varying the parameters for PPIs is used to examine the effect on value, smoothness, how flexible the index can be to find different structure and whether the index is fiddly to tune.
 Trace plots between noise and structured projections are used to examine this.
-- Benchmarking on a range of sample sizes is used to assess the computational speed.
+6. Benchmarking on a range of sample sizes is used to assess the computational speed.
 
@@
 
-## Ease of optimisation
+## Optimisation
 \label{sec:guided}
@@
 On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimisation when considering them as index functions.
-The final view identified by the guided tour is shown in Fig \ref{}
+The final view identified by the guided tour is shown in Figure \ref{fig:findsine}.
 

### 2018-11-22 scagnostics.Rmd
updating index exploration section

@@
 
-The structured projection is in variables 5 and 6. Two samples sizes are used: $n=(100, 1000)$. All variables are standardized to have mean 0 and standard deviation 1.
-
 ```{r util}
@@
 
+The structured projection is in variables 5 and 6 ($x_5, x_6$).
+Two samples sizes are used: $n=(100, 1000)$.
+All variables are standardized to have mean 0 and standard deviation 1.
 Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables.
@@
 
-The family of data sets can be used to assess the PPIs on the desirable properties: smoothness, squintability, flexibility, rotation invariance, and speed. The procedures are as follows:
+The procedures for assessing PPI properties of smoothness, squintability, flexibility, rotation invariance, and speed are examined for samples from the family of data sets are as follows:
 
@@
 1.
-Use the tour to show the index values along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
+Compute the PPI values on the tour path along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
 The result is ideally a smooth change in low values.
@@
 2.
-Use the tour to show the index values along an interpolation between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$.
+Chamge to a tour path between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$, and compute the PPI along this.
 This examines the squintability, and smoothness.
@@
 3.
-Ease of optimisation depends on having a smooth function, and this is tested using a guided tour, showing the optimisation path, starting from $x_1 - x_2$.
-One index is optimised to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, which allows an assessment of similarity between PPIs.
-4.Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
+Use the guided tour to examine the ease of optimisation.
+This depends on having a relatovely smooth function, with structure visible from a distance.
+One index is optimised to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, to examine the similarity between PPIs.
+4. Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
 5.
-Varying the parameters for PPIs is used to examine the effect on value, smoothness, how flexible the index can be to find different structure and whether the index is fiddly to tune.
-Trace plots between noise and structured projections are used to examine this.
-6. Benchmarking on a range of sample sizes is used to assess the computational speed.
+Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure.
+The parameters for these PPIs is varied and examined along a tour path between noise and structured projections, to examine the effects of parameter choices and how fiddly the index is to tune.
+6. Computational speed for the selected indices is examined on a range of sample sizes.
 
@@
 
-## Index traces over tour sequence of interpolated nuisance projections
+## PPI traces over a tour sequence of interpolated nuisance projections
 \label{sec:smooth}
@@
 All projections would be nuisance so the PPI are ideally low and smooth, with little difference between sample sizes (solid lines: $n=1000$; dashed: $n=100$).
-The scagnostic PPIs are noisy, as is TIC.
+The scagnostic PPIs are noisy.
 Some indexes have distinct differences in values between sample sizes."}
@@
 
-## Index traces over tour sequence between nuisance and structured projections
+## PPI traces over a tour sequence between nuisance and structured projections
 \label{sec:squintability}
@@
 
-## Optimisation
+## Optimisation check with the guided tour
 \label{sec:guided}
@@
 
-The guided tour combines optimisation with interpolating between pairs of planes.
-Target planes of the path are chosen by a random search method to find projection planes with high index values, which are then interpolated between.
-It operates roughly like simulated annealing that as one gets closer to an optimal value the random search range is reduced.
-
-Given the behaviour of the various index measures under an interpolated tour path we next try to use the tourr implementation of the guided tour together with selected index functions to uncover the special views in datasets 2 and 3 with 1000 datapoints.
+The guided tour combines optimisation with interpolation between pairs of planes.
+Target planes of the path are chosen to maximise the PPI.
+There are three derivative-free optimisation methods available in the guided tour: \verb# search_better_random# (1), \verb# search_better# (2), and \verb# search_geodesic# (3).
+Method 1 casts a wide net randomly generating projection planes, computing the PPIs and keeping the best projection.
+Method 2 conducts a localised maximum search.
+Method 3 is quite different: a local search is conducted to determine a promising direction, and then this direction is followed until the maximum in that direction is found.
+In each case, a single target plane is chosen by the search methods, and a tour path generated to interpolate from the current projection to the target plane.
 
@@
 }
+```
+
+```{r eval=FALSE}
+# Quick glimpse of optimisations
+quartz()
+animate_xy(pipeResc, guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
+                                     cooling=1, alpha = 0.5, max.tries = 5000))
+animate_xy(pipeResc, guided_tour(holes, search_f = tourr:::search_better,
+                                     cooling=1, alpha = 0.5, max.tries = 5000))
 ```

### 2018-11-26 scagnostics.Rmd
just playig with watching tours

@@
 Method 3 is quite different: a local search is conducted to determine a promising direction, and then this direction is followed until the maximum in that direction is found.
-In each case, a single target plane is chosen by the search methods, and a tour path generated to interpolate from the current projection to the target plane.
+In each case, a single target plane is chosen by the search methods, and a tour path is generated to interpolate from the current projection to the target plane.
 
@@
 # Quick glimpse of optimisations
+myholes <- function(mat) {
+    n <- nrow(mat)
+    d <- ncol(mat)
+    num <- 1 - 1/n * sum(exp(-0.5 * rowSums(mat^2)))
+    den <- 1 - exp(-d/2)
+    (num/den - 0.7) * 1.43
+}
 quartz()
-animate_xy(pipeResc, guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
-                                     cooling=1, alpha = 0.5, max.tries = 5000))
-animate_xy(pipeResc, guided_tour(holes, search_f = tourr:::search_better,
-                                     cooling=1, alpha = 0.5, max.tries = 5000))
+set.seed(sample(10000:50000, 1))
+animate_xy(pipeResc, guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better, cooling=0.9, alpha = pi, max.tries = 500))
+animate_xy(pipeResc, guided_tour(mineIndex("MIC"), search_f = tourr:::search_better, cooling=1, alpha = pi, max.tries = 500))
+animate_xy(pipeResc, guided_tour(mineIndex("TIC"), search_f = tourr:::search_better, cooling=1, alpha = pi, max.tries = 5000))
+animate_xy(pipeResc, guided_tour(holes, search_f = tourr:::search_geodesic,
+      cooling=0.8, alpha = pi, max.tries = 500))
 ```

### 2018-11-27 scagnostics.Rmd
Comment on holes normalisation added.

@@
 TIC also is noisy.
-\textcolor{red}{why is TIC so high?
-did it get inverted?} On the other hand holes, dcor2d, splines2d and MIC are relatively smooth functions.
+On the other hand holes, dcor2d, splines2d and MIC are relatively smooth functions.
 \textcolor{red}{why is holes so high?
-it should range between 0-1, maybe needs to be better standardised, too.} Several of the indexes are sensitive to sample size also, the same structured projection with differing numbers of points, produces different values.
+it should range between 0-1, maybe needs to be better standardised, too.}
+\textcolor{blue}{Holes and cmass indexes are derived from $I_0^N$ of \cite{CBC93}.
+As noted in proposition 1 of that paper the index takes local maxima for the minimum and maximum of $a_0$ which are achieved by central hole and central mass distributions respectively.
+Ref \cite{Cook:2007:IDG:1557227} then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
+They are defined such that each one is maximised for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
+It follows that for either index both large and small values signal deviation from the normal distribution, and given a normal distribution we expect to find "average" index values rather than values close to zero.}
+Several of the indexes are sensitive to sample size also, the same structured projection with differing numbers of points, produces different values.
 

### 2018-11-28 scagnostics.Rmd
More on rescaling of holes/cmass index.

@@
 They are defined such that each one is maximised for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
-It follows that for either index both large and small values signal deviation from the normal distribution, and given a normal distribution we expect to find "average" index values rather than values close to zero.}
+It follows that for either index both large and small values signal deviation from the normal distribution, and given a normal distribution we expect to find "average" index values rather than values close to zero.
+I think we can estimate the values found for normal distribution by comparing values of $a_{00}$ of Sec 7.1 in \cite{CBC93}.
+The maximum value is $1/(2\pi)$ found for cmass type distributions, the minimum is $1/(2\pi e)$ found for hole type distributions.
+Evaluating for normal distributions gives $1/(4\pi)$, rescaling such that the index values range from 0 to 1 then puts the value for normal distributions at approximately $0.2$.
+This appears consistent with the results we have, i.e.
+for normal distributions the cmass index is about $0.2$, and the holes index (=1-cmass) is about $0.8$.
+Proposed rescaling: first have cutoff at the respective value for the normal distribution, i.e.
+any value below 0.2 (0.8) is set to zero for cmass (holes) index, rescale remaining range to be between zero and one.
+}
+
 Several of the indexes are sensitive to sample size also, the same structured projection with differing numbers of points, produces different values.

### 2018-11-30 scagnostics.Rmd
Updated pipe trace plots according to issues.

@@
 \textcolor{red}{Sometimes index value decreases, because index is not rotationally invariant.}
+\textcolor{red}{Updated the plot, makes it look like skinny, 1-convex should be better index to optimise, however in practice this is not the case, probably because TIC is less sensitive to fluctuations, making it more robust against getting stuck in local maxima.}
 
@@
 }
-newBasisPipeTour <- which(attributes(interpolate(pipeTour))$new_basis)
-ggplot(pipeTourResMelt, aes(x=t, y=value)) +
-  geom_line() + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")+
-  geom_vline(data=as.tibble(newBasisPipeTour), mapping=aes(xintercept=value), color="red")
+newBasisPipeTour <- pipeTourResMelt %>%
+  filter(t %in% which(attributes(interpolate(pipeTour))$new_basis))
+ggplot(newBasisPipeTour, aes(x=t, y=value)) +
+  geom_point(mapping=aes(color=(variable=="TIC"))) + 
+  scale_colour_manual(values = c("black", "red")) +
+  facet_wrap(~variable, ncol=1) +
+  guides(color=FALSE) +
+  ylim(c(0,1))
 ```

### 2018-11-30 scagnostics.Rmd
Sine trace plot updated.

@@
 Indeed the guided tour quickly converges in this case, the evolution of the index functions is shown in Fig \ref{fig:findsine}.
+\textcolor{red}{Notice that even for anchor planes highlighted by blue vertical lines the index sometimes decreases because of the rotation dependence, ...
+more details here?
+e.g.
+point out that such effects depend on the optimizer, here we use search\_geodesic, closely relates to how interpolation is done, therefore no big jumps in index values as found for search\_random type optimizers.}
 Here there are four index functions that appear to be suitable for the optimisation: splines2d, dcor2d, MIC and TIC.
@@
 
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset.", results="hide"}
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset. Anchor planes are marked by the blue vertical lines.", results="hide"}
 set.seed(2018)
@@
 }
+anchorIdx <- as.tibble(which(attributes(interpolate(sineTour))$new_basis))
 ggplot(sineTourResMelt, aes(x=t, y=value)) +
-  geom_line() + 
-  facet_wrap(~variable, ncol=1, scales = "free_y")
+  geom_line(mapping=aes(color=(variable=="splines2d"))) + 
+  facet_wrap(~variable, ncol=1, scales = "free_y") +
+  scale_colour_manual(values = c("black", "red")) +
+  facet_wrap(~variable, ncol=1) +
+  geom_vline(data=anchorIdx, mapping=aes(xintercept=value), color="blue", alpha=0.3) +
+  guides(color=FALSE) +
+  ylim(c(0,1))
 ```

### 2018-11-30 scagnostics.Rmd
Replacing convex with 1-convex in most plots.

@@
 }
-newBasisPipeTour <- pipeTourResMelt %>%
+newBasisPipeTour <- pipeTourRes %>%
+  mutate(convex = 1-convex) %>% 
+  melt(id=c("t", "name", "size")) %>%
   filter(t %in% which(attributes(interpolate(pipeTour))$new_basis))
@@
 anchorIdx <- as.tibble(which(attributes(interpolate(sineTour))$new_basis))
+sineTourResMelt <-sineTourRes %>%
+  mutate(convex = 1-convex) %>% 
+  melt(id=c("t", "name", "size"))
 ggplot(sineTourResMelt, aes(x=t, y=value)) +
@@
 scMelt <- sc %>% 
+  mutate(convex = 1-convex) %>%
   mutate(angle=alpha*360/(2*pi), TIC = TIC) %>%

### 2018-11-30 scagnostics.Rmd
Smaller rotation dependence plot.

@@
 
+\textcolor{red}{This isn't actually the "ideal" view, why is that?}
+
 ```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset. Anchor planes are marked by the blue vertical lines.", results="hide"}
@@
 \label{sec:rot}
-```{r rotationDep, fig.width=6, fig.height=7, fig.cap="PPI for rotations of the sine 1000 data, to examine rotation invariance.
+```{r rotationDep, out.width=".6\\textwidth", fig.cap="PPI for rotations of the sine 1000 data, to examine rotation invariance.
 Most are close to rotation invariant, except for dcor2d, splines2d and TIC."}

### 2018-12-03 scagnostics.Rmd
Added study for spiral guided tour optimisation.

@@
 
+\textcolor{red}{Probably a good idea to give stopping condition to scouting phase, since at each step we may spend a long time looking for better projections.
+E.g.
+if maximum index value is 1, stop scoutig phase when reaching index value 0.9, i.e.
+should be in squintable distance of the index maximum, switch to geodesic optimiser at this point (only works if indeed the optimal view takes index values close to the maximum though)}
+
 The results are shown below, first Fig.
@@
 
-\textcolor{red}{can use skinny to find sprial in 3 dimensions, but not higher, scouting method does not work because of large rotation dependence, maybe make it study of guided tour success vs number of parameters entering?
-e.g.
-given n starting planes how many times was view in plane V5-V6 identified as function of number of dimensions and index used?}
-
-\textcolor{red}{what about role of optimisation here?
-search\_better not working because rotation in plane makes index drop randomly, geodesic search does not work because (my guess) it picks directions based on being most promising on small step size tests, maybe need to modify code to pass step size as argument, search directions based on larger step size may result in improved scouting run?
-(but may also give same issues of rotation dependence) ...
-tried  but didnt really work...}
-
-```{r findspiral, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the skinny index for the Sprial 1000 dataset.", results="hide"}
-#getResWithSeed <- function(s){
-#  set.seed(s)
-#  startM <- orthonormalise(matrix(runif(12),ncol = 2))
-#  tourRes <- save_history(spiralResc, guided_tour(scagIndex("Skinny")), max_bases = 10000)
-#}
-#set.seed(202020)
-#if(!file.exists("cache/findsine.rda")){
-#  spiralResc <- as.matrix(spiral1000)
-#spiralTour <- save_history(spiralResc, guided_tour(dcorIndex()))
-#spiralTour <- save_history(spiralResc, guided_tour(scagIndex("Skinny"), max.tries = 100, n=10, stepS=0.5), max_bases = 10000)
-  #spiralTour <- save_history(spiralResc,guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
-  #                                   alpha = 0.5, max.tries = 500), max_bases = 10000)
-  #spiralTourFull <- as.list(interpolate(spiralTour))
-  #spiralTourRes <- getProj(spiral1000, spiralTourFull, "Spiral", 1000)
-  #spiralTourResMelt <- melt(spiralTourRes, id=c("t", "name", "size"))
-#  save(spiralResc, spiralTour, spiralTourFull, spiralTourRes, spiralTourResMelt,
-#       file = "cache/findspiral.rda")
-#} else {
-#  load("cache/findsine.rda")
-#}
-#ggplot(spiralTourResMelt, aes(x=t, y=value)) +
-#  geom_line() + 
-#  facet_wrap(~variable, ncol=1, scales = "free_y")
-#iL <- length(as.list(spiralTour))
-#dProj <- spiralResc %*% as.list(spiralTour)[[iL]]
-#ggplot(as.tibble(dProj), aes(V1, V2)) + geom_point()
+The spiral is the most challenging structure to detect amonst the distribution considered here (compare also Posses results?).
+Here we explore the role of dimensionality in the index optimisation, using the skinny index that was found to be sensitive to this structure.
+We use a setup similar to that introduced for detecting the pipe structure, i.e.\ we use the search\_better optimisation of the guided tour, disabeling the cooling and with a large search window, while sampling a large number of random projections (up to 5000) for each step.
+We consider distributions in d=4,6,8 dimensions and compare features found in the tour.
+In d=4 dimensions the optimisation identifies a view showing the spiral after 33 steps, and we limit the optimiser to consider only 33 steps for the study in 6 and 8 dimensions for direct comparison.
+Note however that the spiral is typically also not identified with much larger number of steps, the limiting factor being rather the numver of random projections sampled and rotation non-invariance of the index prohibiting the convergence, in particular in higher dimensions.
+Results are shown in the top panel Fig.~\ref{fig:findspiral}, showing the pariwise distance of the 33 planes found in 4, 6 and 8 dimensions (left), and comparing the evolution of the distance from the plane containing the spiral view with progression of the optimisation (right).
+We see that the pairwise distances grow with d (what do we learn from that?).
+Moreover we find that the distance to the special view is far from monotonically decreasing towards zero, and large jumps are possible.
+Note that the spiral itself is approximately rotation invariant, but partial views are not.
+At medium distances from the special view we often observe a jump to higher distances as consequence of in-plane rotation dependence for such partial views.
+This stabilises closer to the special view, however it relies on randomly identifying a projection within this angle.
+As a consequence increasing dimensionality dramatically reduces the chance of identifying structures with high index values.
+
+```{r findspiral, out.width="\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with d = 4, 6, 8.
+The top left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the top right shows the evolution of distance to the plane containing the spiral as the index is being optimised.
+The bottom row shows the final view obtained for d = 4 (left), d = 6 (middle) and d = 8 (right) after 33 steps in the guided tour.", results="hide"}
+
+distanceDist <- function(planes, nn=F){
+  #nn could be used to turn on only distance to nearest neighbour?
+  planes <- as.list(planes)
+  maxI <- length(planes)
+  distL <- numeric(choose(maxI, 2))
+  i <- 1
+  idx <- 1
+  while(i<maxI+1){
+    j <- i+1
+    while(j<maxI+1){
+      cDist <- proj_dist(planes[[i]], planes[[j]])
+      distL[idx] <- cDist
+      j <- j+1
+      idx <- idx+1
+    }
+    i <- i+1
+  }
+  return(distL)
+}
+
+distanceToSp <- function(planes, specialPlane){
+  planes <- as.list(planes)
+  maxI <- length(planes)
+  distL <- numeric(maxI)
+  i <- 1
+  while(i<maxI+1){
+    cDist <- proj_dist(planes[[i]], specialPlane)
+    distL[i] <- cDist
+    i <- i+1
+  }
+  return(distL)
+  }
+if(!file.exists("cache/findspiral.rda")){
+  distDf <- tibble(d=numeric(), distance=numeric())
+  distToSp <- tibble(d=numeric(), distance=numeric(), t=numeric())
+
+  maxBases <- 100
+  fViewL <- list()
+  allPlanes <- list()
+  i <- 1
+  for(d in c(4,6,8)){
+    set.seed(1779+d)
+    specialPlane <- matrix(c(rep(0,d-2),1,0,rep(0,d-1),1),ncol=2)
+    spiralD <- spiralData(d, 1000) %>% scale() %>% as_tibble()
+    spiralTour <- save_history(spiralD, guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
+                                                    max.tries = 5000, cooling = 1, alpha = pi),
+                              max_bases = maxBases)
+    if(d==4){maxBases <- min(maxBases, length(spiralTour))}
+    distDf <- rbind(distDf, tibble(d=d, distance=distanceDist(spiralTour)))
+    distToSp <- rbind(distToSp, tibble(d=d, distance=distanceToSp(spiralTour, specialPlane), t=1:maxBases))
+    fView <- as.tibble(as.matrix(spiralD) %*% as.list(spiralTour)[[maxBases]])
+    fViewL[[i]] <- ggplot(fView, aes(V1, V2)) + geom_point()
+    allPlanes[[i]] <- spiralTour
+    i <- i+1
+  }
+  save(distDf, distToSp, fViewL, allPlanes, file = "cache/findspiral.rda")
+} else {
+  load("cache/findspiral.rda")
+}
+
+p1 <- ggplot(distDf, aes(group=d, y=distance, x=d)) + geom_boxplot()
+p2 <- ggplot(distToSp, aes(color=as.factor(d),x =t, y=distance)) +
+  geom_point() +
+  labs(y = "Distance to special view", color = "d")
+
+grid.arrange(p1, p2, fViewL[[1]], fViewL[[2]], fViewL[[3]], layout_matrix = rbind(c(1,2,2),c(3,4,5)))
+
 ```

### 2018-12-03 scagnostics.Rmd
Function for rescaled cmass, holes index.

@@
 }
+
+##rescaled holes, cmass
+holesR <- function(){
+  function(mat){
+    ret <- tourr::holes(mat)
+    if(ret<0.8) ret <- 0.
+    else ret <- (ret-0.8) * 5
+    return(ret)
+  }
+}
+
+cmassR <- function(){
+  function(mat){
+    ret <- tourr::cmass(mat)
+    if(ret<0.2) ret <- 0.
+    else ret <- (ret-0.2) * 1.25
+    return(ret)
+  }
+}
+
 ```

### 2018-12-06 scagnostics.Rmd
Clean up (remove commented content).

@@
 
+## PPI traces over a tour sequence between nuisance and structured projections
+\label{sec:squintability}
+
+Figure \ref{fig:plannedtour} shows the PPIs for a tour sequence between a nuisance and structured projection.
+Sample size is indicated by linetype: dashed being $n=100$ and solid is $n=1000$.
+The beginnnig of the sequence is the nuisance projection and the end is the structured projection.
+The index values for most PPIs increases substantially nearing the structured projection, indicating that they "see" the structure.
+Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
+[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
+The holes index easily detects the pipe, and can see it from a distance.
+The scagnostic index, stringy, can see the structure but is myopic, only when it is very close.
+Interestingly the scagnostic, skinny, sees the spiral from a distance.
+
 ```{r plannedtourpath}
@@
 ```
-
-## PPI traces over a tour sequence between nuisance and structured projections
-\label{sec:squintability}
-
-Figure \ref{fig:plannedtour} shows the PPIs for a tour sequence between a nuisance and structured projection.
-Sample size is indicated by linetype: dashed being $n=100$ and solid is $n=1000$.
-The beginnnig of the sequence is the nuisance projection and the end is the structured projection.
-The index values for most PPIs increases substantially nearing the structured projection, indicating that they "see" the structure.
-Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
-[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
-The holes index easily detects the pipe, and can see it from a distance.
-The scagnostic index, stringy, can see the structure but is myopic, only when it is very close.
-Interestingly the scagnostic, skinny, sees the spiral from a distance.
-
-<!--
-\textcolor{red}{XXX This shouldn't go here.
-The index needs to have been introduced much earlier in the paper.
-TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
-We also see that TIC is smoother when moving into the special view, in terms of the relative differences.}
--->
-
-<!--
-We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function.
-Results are shown in Figure \ref{fig:plannedtour}.
-We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise.
-On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour.
-Finally the Spiral appears to be the most challenging, having an even smaller squint angle than the Pipe distribution.
-While "cmass" appears to be appropriate to identify this view we may actually find larger values of this index for random distributions, making it unfit to use in a guided tour.
-The scagnostics indices skinny (or 1-convex) appear to be sensitive to the structure at a larger angle compared to MINE indices, but will suffer from the large fluctuations in random views when used in a guided tour.
-Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).
-
-TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered.
-We also see that TIC is smoother when moving into the special view, in terms of the relative differences.
-
-\textcolor{red}{Holes index or even pca could discover pipe as well...}
--->
-
-<!--
-I don't think this section adds anything to the paper
-## Distributions of index values
-\label{sec:dist}
-Compare typical index values over the two tour paths.
-Notation: path 1 / 2 labels tour from $x_1$-$x_2$ to $x_3$-$x_4$ / $x_5$-$x_6$.
-Typically find broader distributions for small samples because of the large fluctuations in this case.
-Some indices show preferred behaviour, i.e.
-sharp peaks around low values for path 1, and very broad distributions leading up to maximum values for path 2.
-For other indices we can see the strong dependence of the index on sample size, the most extreme example is the skinny index.
-
-\textcolor{red}{Need better representation, in particular normalise TIC to upper limit for given sample size, adapt scales so sharp peaks don't dominate some of the plots.}
-
-```{r indexDist, fig.width=6, fig.height=7, fig.cap="Comparing distributions of index values found in the interpolated tour paths 1 and 2 as described in the text."}
-fullResMelt1 <- fullResMelt %>% add_column(path="path1")
-fullResMeltPlanned1 <- fullResMeltPlanned %>% add_column(path="path2")
-allRes <- rbind(fullResMelt1, fullResMeltPlanned1)
-ggplot(allRes, aes(x=value)) +
-  geom_density(aes(color=path, linetype = factor(size))) +
-  facet_wrap(PPI~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
-```
-
-\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$, maybe can add second path for spiral?}
-Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binning (and outlier removal?) here.
-Observe huge drops for some of the indexes!
--->
 

### 2018-12-06 scagnostics.Rmd
Editing sec 3.4.

@@
 
-- remind reader of the options for optimisation in guided tour
-- how you use them here
--
-
 The guided tour combines optimisation with interpolation between pairs of planes.
@@
 Method 3 is quite different: a local search is conducted to determine a promising direction, and then this direction is followed until the maximum in that direction is found.
-In each case, a single target plane is chosen by the search methods, and a tour path is generated to interpolate from the current projection to the target plane.
+For all methods the optimisation is iterative, and we record each step where a plane with increased index value has been found.
+The stopping condition is given as number of tries that did not yield an increased index value at the final step in the optimisation.
+For method 1 and 2 two parameters control the optimisation: the search window $\alpha$, giving the maximum distance from the current plane in which projections are samples, and the cooling factor, giving the incremental decrease in search window size.
+Method 3 in principle also has two free parameters, which are however fixed in the current implementation.
+The first is the small step size used when evaluating the most promising direction, it is fixed to 0.01, and the second parameter being the window over which the line search is performed, fixet to $\pm \pi/4$ away from the current plane.
+For distributions and indexes with smooth behaviour and good squinatbaility, method 3 is the most efficient method for optimisation.
+If these two criteria are not met the method may still be useful, but only given an informed starting point.
+In such cases we can follow a method similar to that proposed by Friedman, i.e.
+we break the optimisation in two distinct steps.
+A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum (note here that this also requires large number of tries for the stopping condition in particular in high dimensions, since most randomly picked planes will not be interesting).
+The second step is then to use method 3 starting from the approximate maximum, which will return the global maximum.
 
@@
 
-\textcolor{red}{Double challenging: fine structure + (coherent) noise}
-\textcolor{red}{simple standard guided tour does not find pipe easily with holes index, did not try brute force approach used with TIC index yet}
-
-As seen before (reference figure) MIC or TIC should be able to find the view down the pipe.
-The optimisation itself is however expected to be challenging because one has to be fairly close to the optimal projection to observe the increase in index value.
-We therefore suggest a two step procedure for the uncovering of such "small scale structures".
-The first step consists in "scouting" projections.
-Here we use the "search_better" optimisation available with the tourr package, with a standard search window $\alpha = 0.5$ and disabling the cooling (in practice this can be done by fixing the cooling parameter to $1$), together with a large number of tries (setting max.tries to 5000 here).
-The resulting search will quickly steer away from the most uninteresting projections, followed by a broad scan of the thus identified region.
-While the lack of cooling means that this will not converge towards the maximum, each step in this first search will increase the index value, thus it can be used to provide an informed starting plane that can be used in a second step where we use the "search_geodesic" optimization to improve the index further towards the maximum.
-The described procedure is CPU time intensive, however it is promising that we manage to uncover such small scale structure in the distribution.
-On the other hand no "guarantee" can be provided for uncovering the special view.
-Parameters for the scouting phase will need to be adapted to different examples, and with increasing number of dimensions it may be better to consider several runs with different starting projections.
-This gives some indication on how to handle difficult datasets. As we will see below most of the time standard "search_geodesic" optimisation will be efficient enough to uncover the interesting views.
+As seen in Figure \ref{fig:plannedtour} the pipe structure has a small squint angle, and indeed using only method 3 does not uncover it in the six dimensional dataset.
+We therefore adapt the two step procedure described above.
+For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimise the TIC index.
 
@@
 
-The results are shown below, first Fig.
-\ref{fig:pipeFirstRun} shows how the considered index values evolve in the scouting phase.
-Note that this is the interpolated path and there are small local maxima and minima along the index value, especially because we keep the step size in the guided tour large.
-We can see the clear maximum of the TIC index in the final projection, as well as clear minimum of the convex index that carries similar information On the other hand we see that MIC or the dcor2d index are not suitable for uncovering the special view here.
-
-\textcolor{red}{Sometimes index value decreases, because index is not rotationally invariant.}
-\textcolor{red}{Updated the plot, makes it look like skinny, 1-convex should be better index to optimise, however in practice this is not the case, probably because TIC is less sensitive to fluctuations, making it more robust against getting stuck in local maxima.}
+The results are shown below, first Fig. \ref{fig:pipeFirstRun} shows how the considered index values evolve in the scouting phase.
+\textcolor{red}{Updated the plot, makes it look like skinny, 1-convex should be better index to optimise, however in practice this is not the case, probably because TIC is less sensitive to fluctuations especially also under rotation, making it more robust against getting stuck in local maxima.
+Maybe it will be better to leave scales free for this plot, since increase in TIC is very hard to see currently.
+Or just remove plot?
+not actually that intersting...}
 

### 2018-12-06 scagnostics.Rmd
More 3.4 edits.

@@
 
-We next work with distribution 3 and aim to identify the view showing functional dependence using the splines2d index.
-Given the results from the planned tour we expect to be able to efficiently identify this view without relying on the two step procedure described above.
+Given the results from Fig.\ref{fig:plannedtour} we expect to be able to efficiently identify the sine view in the 6-d distribution, using only method 3 for the optimisation, using e.g.
+the splines2d index.
 Indeed the guided tour quickly converges in this case, the evolution of the index functions is shown in Fig \ref{fig:findsine}.
-\textcolor{red}{Notice that even for anchor planes highlighted by blue vertical lines the index sometimes decreases because of the rotation dependence, ...
-more details here?
-e.g.
-point out that such effects depend on the optimizer, here we use search\_geodesic, closely relates to how interpolation is done, therefore no big jumps in index values as found for search\_random type optimizers.}
+The only difficulties stem from the rotation dependence of the index function as well as the structured distribution, this will be discussed in some more details below.
+Here we only note that even for anchor planes highlighted by blue vertical lines the index sometimes decreases because of this dependence.
+These are however not dramatic effects because this search method is closely related to how interpolation is done (a different picture would be optained using method 2 which often leads to large drops in index values after the interpolation step).
+
 Here there are four index functions that appear to be suitable for the optimisation: splines2d, dcor2d, MIC and TIC.
@@
 
-\textcolor{red}{This isn't actually the "ideal" view, why is that?}
+\textcolor{red}{This isn't actually the "ideal" view, could be because of rotation dependence??}
 

### 2018-12-06 scagnostics.Rmd
Some text for sec 3.5.

@@
 \label{sec:rot}
+
+We study rotation dependence using the sine function distribution, and tracing the index values as a function of roation angle.
+Results are shown in Figure \ref{fig:rotationDep}.
+Several indexes are found to be invariant, holes, convex and MIC.
+The dcor2d, splines2d and TIC index show consistent patterns in rotation dependence, on the other hand the scagnostics index skinny is found to have sizable random fluctuations as the basis is rotated.
 ```{r rotationDep, out.width=".6\\textwidth", fig.cap="PPI for rotations of the sine 1000 data, to examine rotation invariance.

### 2018-12-06 scagnostics.Rmd
Fig 2,3 with rescaled holes index.

@@
 
+rescaleHoles <- function(x){
+  if(x<0.8) x <- 0.
+  else x <- (x-0.8) * 5
+  return(x)
+}
+
+rescaleCmass <- function(x){
+  if(x<0.2) x <- 0.
+  else x <- (x-0.2) * 1.25
+  return(x)
+}
+
 ```
@@
   mutate(convex = 1-convex) %>%
+  rowwise() %>%
+  mutate(holes = rescaleHoles(holes)) %>%
   gather(PPI, value, -t, -name, -size) %>%
   filter(PPI != "cmass") %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
 fullResMelt %>% 
@@
   mutate(convex = 1-convex) %>% 
+  rowwise() %>%
+  mutate(holes = rescaleHoles(holes)) %>%
   gather(PPI, value, -t, -name, -size) %>%

### 2018-12-10 scagnostics.Rmd
fixed table, made appendix with holes re-scaling

@@
   affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Heike Hofmann
-  affiliation: Department of Statistics, Iowa State University
-
-- name: Hadley Wickham
-  affiliation: RStudio
-
-- name: Antony Unwin
-  affiliation: Department of Mathematics, Augsburg University
-
-- name: Katrin Grimm
-  affiliation: ???
 
@@
 abstract: |
-  Projection pursuit describes a procedure for searching high-dimensional data for "interesting" low-dimensional projections via the optimization of a criterion function called the projection pursuit index.
-Most indexes developed focus on finding projections with cluster structure, outliers, or separations between known groups.
-Here, we are interested in finding projections revealing potentially complex relations between combinations of parameters, by combining the notion of projection pursuit with index functions developed for the detection of interesting bivariate views in the data, including scagnostics and maximum information coefficient.
+  Projection pursuit is used to find interesting low-dimensional projections of high-dimensional data by optimizing an index over all possible projections.
+Most indexes have been developed to detect departure from known distributions, such as normality, or to find separations between known groups.
+Here, we are interested in finding projections revealing potentially complex bivariate patterns, using new indexes constructed from scagnostics and a maximum information coefficient, with a purpose to detect unusual relationships between model parameters describing physics phenomena.
+The performance of these indices is examined with respect to ideal behaviour, using simulated data, and then applied to several physics problems.
+The implementation builds upon the projection pursuit tools available in the R package, tourr, with indices constructed from code in the R package, scagnostics.
 
@@
 - **squintability**: \cite{barnett1981interpreting} introduced the idea of squint angle to indicate resolution of structure in the data.
-Fine structure like the parallel planes in the infamous RANDU data \cite{compaqFortran} (FIXME this is fortran reference, could not find one for the dataset) has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
+Fine structure like the parallel planes in the infamous RANDU data \cite{Marsaglia68} has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
 Structures with small squint angle are difficult to find, because the optmisation algorithm needs to get very close to begin hill-climbing to the optimum.
@@
       add_header_above(c(" ", " ", "pipe" = 2, "sine" = 2, "spiral"=2)) %>%
-  group_rows("", 1, 2) %>%
-  group_rows("", 3, 4) %>%
-  group_rows("", 5, 6) %>%
-  group_rows("", 7, 8) %>%
-  group_rows("", 9, 10) %>%
-  group_rows("", 11, 12) %>%
-  group_rows("", 13, 14) %>%
-  group_rows("", 15, 16) %>%
+  group_rows("", 1, 2, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 3, 4, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 5, 6, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 7, 8, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 9, 10, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 11, 12, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 13, 14, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 15, 16, latex_gap_space = "-0.2cm") %>% 
+  #row_spec(1:16, font_size = 10) %>%
   kable_styling(position = "center") 
@@
 All indexes, except convex show distinctly higher values for the structured projections.
-The convex index considers the noise projections more interesting than the structured projections, which might be expected, but also provides an argument to use (1-convex) as the PPI.
+The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs.
+The scale for the holes index is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performace assessment so that all indices can be plotted on a common scale of 0-1.
 
@@
 2.
-Chamge to a tour path between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$, and compute the PPI along this.
+Change to a tour path between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$, and compute the PPI along this.
 This examines the squintability, and smoothness.
@@
 On the other hand holes, dcor2d, splines2d and MIC are relatively smooth functions.
-\textcolor{red}{why is holes so high?
-it should range between 0-1, maybe needs to be better standardised, too.}
-\textcolor{blue}{Holes and cmass indexes are derived from $I_0^N$ of \cite{CBC93}.
-As noted in proposition 1 of that paper the index takes local maxima for the minimum and maximum of $a_0$ which are achieved by central hole and central mass distributions respectively.
-Ref \cite{Cook:2007:IDG:1557227} then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
-They are defined such that each one is maximised for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
-It follows that for either index both large and small values signal deviation from the normal distribution, and given a normal distribution we expect to find "average" index values rather than values close to zero.
-I think we can estimate the values found for normal distribution by comparing values of $a_{00}$ of Sec 7.1 in \cite{CBC93}.
-The maximum value is $1/(2\pi)$ found for cmass type distributions, the minimum is $1/(2\pi e)$ found for hole type distributions.
-Evaluating for normal distributions gives $1/(4\pi)$, rescaling such that the index values range from 0 to 1 then puts the value for normal distributions at approximately $0.2$.
-This appears consistent with the results we have, i.e.
-for normal distributions the cmass index is about $0.2$, and the holes index (=1-cmass) is about $0.8$.
-Proposed rescaling: first have cutoff at the respective value for the normal distribution, i.e.
-any value below 0.2 (0.8) is set to zero for cmass (holes) index, rescale remaining range to be between zero and one.
-}
 
@@
 # References
+
+# Appendix
+
+XXX Need to make this not numbered
+
+## Re-scaling of holes index
+
+Holes and cmass indexes are derived from $I_0^N$ of \cite{CBC93}.
+As noted in proposition 1 of that paper the index takes local maxima for the minimum and maximum of $a_0$ which are achieved by central hole and central mass distributions respectively.
+Ref \cite{Cook:2007:IDG:1557227} then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
+They are defined such that each one is maximised for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
+It follows that for either index both large and small values signal deviation from the normal distribution, and given a normal distribution we expect to find "average" index values rather than values close to zero.
+
+We can estimate the values found for normal distribution by comparing values of $a_{00}$ of Sec 7.1 in \cite{CBC93}.
+The maximum value is $1/(2\pi)$ found for cmass type distributions, the minimum is $1/(2\pi e)$ found for hole type distributions.
+Evaluating for normal distributions gives $1/(4\pi)$, rescaling such that the index values range from 0 to 1 then puts the value for normal distributions at approximately $0.2$.
+This appears consistent with the results we have, i.e.
+for normal distributions the cmass index is about $0.2$, and the holes index (=1-cmass) is about $0.8$.
+
+Rescaling: first have cutoff at the respective value for the normal distribution, i.e.
+any value below 0.2 (0.8) is set to zero for cmass (holes) index, rescale remaining range to be between zero and one.

### 2018-12-10 scagnostics.Rmd
fixed table, made appendix with holes re-scaling

@@
 The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs.
-The scale for the holes index is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performace assessment so that all indices can be plotted on a common scale of 0-1.
+The scale for the holes index is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performace assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix).
+Similarly, the TIC index is re-scaled.
 

### 2018-12-11 scagnostics.Rmd
guided tour section revised

@@
       add_header_above(c(" ", " ", "pipe" = 2, "sine" = 2, "spiral"=2)) %>%
-  group_rows("", 1, 2, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 3, 4, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 5, 6, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 7, 8, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 9, 10, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 11, 12, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 13, 14, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 15, 16, latex_gap_space = "-0.2cm") %>% 
+  group_rows("", 1, 2, latex_gap_space = "-0.5cm") %>%
+  group_rows("", 3, 4, latex_gap_space = "-0.5cm") %>%
+  group_rows("", 5, 6, latex_gap_space = "-0.5cm") %>%
+  group_rows("", 7, 8, latex_gap_space = "-0.5cm") %>%
+  group_rows("", 9, 10, latex_gap_space = "-0.5cm") %>%
+  group_rows("", 11, 12, latex_gap_space = "-0.5cm") %>%
+  group_rows("", 13, 14, latex_gap_space = "-0.5cm") %>%
+  group_rows("", 15, 16, latex_gap_space = "-0.5cm") %>% 
   #row_spec(1:16, font_size = 10) %>%
@@
 There are three derivative-free optimisation methods available in the guided tour: \verb# search_better_random# (1), \verb# search_better# (2), and \verb# search_geodesic# (3).
-Method 1 casts a wide net randomly generating projection planes, computing the PPIs and keeping the best projection.
-Method 2 conducts a localised maximum search.
+Method 1 casts a wide net randomly generating projection planes, computing the PPIs and keeping the best projection, and method 2 conducts a localised maximum search.
 Method 3 is quite different: a local search is conducted to determine a promising direction, and then this direction is followed until the maximum in that direction is found.
-For all methods the optimisation is iterative, and we record each step where a plane with increased index value has been found.
-The stopping condition is given as number of tries that did not yield an increased index value at the final step in the optimisation.
-For method 1 and 2 two parameters control the optimisation: the search window $\alpha$, giving the maximum distance from the current plane in which projections are samples, and the cooling factor, giving the incremental decrease in search window size.
+For all methods the optimisation is iterative, the best projections form target planes in the tour, the tour path is the interpolation to this target, and then a new search for a better projection is made, followed by the interpolation.
+For each projection during the interpolation steps, the PPI is recorded.
+
+The stopping rule is that no better projections are found after a fixed number of tries, given a fixed tolerance value measuring difference.
+For method 1 and 2 two additional parameters control the optimisation: the search window $\alpha$, giving the maximum distance from the current plane in which projections are sampled, and the cooling factor, giving the incremental decrease in search window size.
 Method 3 in principle also has two free parameters, which are however fixed in the current implementation.
-The first is the small step size used when evaluating the most promising direction, it is fixed to 0.01, and the second parameter being the window over which the line search is performed, fixet to $\pm \pi/4$ away from the current plane.
-For distributions and indexes with smooth behaviour and good squinatbaility, method 3 is the most efficient method for optimisation.
-If these two criteria are not met the method may still be useful, but only given an informed starting point.
-In such cases we can follow a method similar to that proposed by Friedman, i.e.
-we break the optimisation in two distinct steps.
-A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum (note here that this also requires large number of tries for the stopping condition in particular in high dimensions, since most randomly picked planes will not be interesting).
-The second step is then to use method 3 starting from the approximate maximum, which will return the global maximum.
+The first is the small step size used when evaluating the most promising direction, it is fixed to 0.01, and the second parameter being the window over which the line search is performed, fixed to $\pm \pi/4$ away from the current plane.
+
+For distributions and indexes with smooth behaviour and good squintability, method 3 is the most effective method for optimisation.
+If these two criteria are not met the method may still be useful, but only given an informed starting projection.
+In such cases we can follow a method similar to that proposed by Friedman (XXX REF): we break the optimisation in two distinct steps.
+A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum.
+Note that this likely requires large number of tries, especially as dimension increases, since most randomly picked planes will not be interesting.
+The second step is uses method 3 starting from the approximate maximum, which will take small steps to refine the result to be closer to the global maximum.
 
@@
 As noted in proposition 1 of that paper the index takes local maxima for the minimum and maximum of $a_0$ which are achieved by central hole and central mass distributions respectively.
-Ref \cite{Cook:2007:IDG:1557227} then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
+\cite{Cook:2007:IDG:1557227} then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
 They are defined such that each one is maximised for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.

### 2018-12-11 scagnostics.Rmd
Rescale holes in Table 1, fixed sorting.

@@
 }
+#rescale holes index first
+res <- res %>%
+  rowwise() %>%
+  mutate(value5 = if_else(index=="holes", rescaleHoles(value5), value5)) %>%
+  rowwise() %>%
+  mutate(value95 = if_else(index=="holes", rescaleHoles(value95), value95))
 resp <- res %>% filter(size == 1000, data == "Pipe") %>%
@@
   select(-index1, -type1, -index2, -type2) %>%
-  mutate(index = fct_relevel(index, "holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe")) %>%
+  mutate(index = factor(index, c("holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe"))) %>%
   arrange(index) %>%

### 2018-12-11 scagnostics.Rmd
Rescale holes, rescale or remove cmass everywhere.

@@
   mutate(convex = 1-convex) %>% 
+  rowwise() %>%
+  mutate(holes = rescaleHoles(holes)) %>%
+  select(-cmass) %>%
   melt(id=c("t", "name", "size")) %>%
@@
   mutate(convex = 1-convex) %>%
+  rowwise() %>%
+  mutate(holes = rescaleHoles(holes)) %>%
+  select(-cmass) %>%
   melt(id=c("t", "name", "size"))
@@
   mutate(convex = 1-convex) %>%
+  rowwise() %>%
+  mutate(holes = rescaleHoles(holes)) %>%
+  rowwise() %>%
+  mutate(cmass = rescaleCmass(cmass)) %>%
   mutate(angle=alpha*360/(2*pi), TIC = TIC) %>%

### 2018-12-11 scagnostics.Rmd
Explanation of parameters in smoothing study.

@@
 
-Focusing on scagnostics indices, for which the biggest issues with noisy transitions have been observed.
-Smoothing can be performed by interpolation, or averaging similar views.
-For example we can use jittering, or we can use tour functionalities to find a set of randomly selected projections within a small angle of the current view.
-In this way we can generate a number of similar views.
-We can then select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calulate the mean from the remaining evaluations.
-Similar to the study of scagnostics binning we now compare different methods of smoothing when moving into the special view of the Spiral 1000 dataset.
-
-```{r spiralScagSmoothing, fig.cap="Comparing the traces of the three scagnostics indices when smoothing the index values, either by averaging over the index value after jittering the projection by some angle alpha (left) or after jittering the projected datapoints with some amount alpha (right).
+Different methods can be considered for smoothing the index traces over similar views for each basis:
+
+* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$
+* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction
+* Turning axes: given the rotation non-invariance, maybe rotation of basis could be useful also??
+
+Moreover we can either select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calulate the mean from the remaining evaluations.
+Only the mean will be considered here.
+
+This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
+We study the potential of smoothing these two for the tour path between normal distributed noise parameters of the Spiral 1000 dataset, varying the $\epsilon$ and $\beta$ parameters.
+
+```{r spiralScagSmoothing, fig.cap="Comparing the traces of the three scagnostics indices when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (left) or after jittering the projected datapoints with some amount $\\beta$ (right).
 While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered."}

### 2018-12-11 scagnostics.Rmd
pipe section plus spell check

@@
 
-[@Grimm2016] explored the behaviour of scagnostics for selecting variables, and proposed two more, based on smoothing splines and distance correlation, that have some nicer properties.
+[@Grimm2016] explored the behavior of scagnostics for selecting variables, and proposed two more, based on smoothing splines and distance correlation, that have some nicer properties.
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
@@
 
-The paper investigates the behaviour of these newly-defined indexes.
+The paper investigates the behavior of these newly-defined indexes.
 Section \ref{sec:construct} discusses index construction, and how they fit into the guided tour.
-Section \ref{sec:investigate} investigates the behaviour of the indexes, particularly in relation to optimisation.
+Section \ref{sec:investigate} investigates the behavior of the indexes, particularly in relation to optimization.
 The new guided tour methods are applied to an example where the posterior distributions from physics models (Section \ref{sec:phys}) are explored.
@@
 A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data distribution.
-Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximising $f$ over all possible $d$ dimensional projections of a data set with $p>d$ parameters will find the most interesting projections.
-As most indices characterise deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
+Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximizing $f$ over all possible $d$ dimensional projections of a data set with $p>d$ parameters will find the most interesting projections.
+As most indices characterize deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
 @CBC93.
-When departing from the idea of characterising differences to the normal distribution it may however be preferred to work directly with the empirical data distribution and instead characterise it by measures based on e.g.
+When departing from the idea of characterizing differences to the normal distribution it may however be preferred to work directly with the empirical data distribution and instead characterize it by measures based on e.g.
 the minimal spanning tree or the mutual information.
@@
 We want to avoid e.g.
-highlighting directions based only on different scales, or artifically extending the parameter range by including outliers, which may result in seemingly linear behaviour for most points.
+highlighting directions based only on different scales, or artificially extending the parameter range by including outliers, which may result in seemingly linear behavior for most points.
 Methods to consider include rescaling of individual directions to a common interval (e.g.
@@
 \textcolor{red}{particular challenges: noise only directions, identification of outliers?}
-\textcolor{red}{for holes/cmass index they are only well defined when data is standardised, how about the other index functions?}
+\textcolor{red}{for holes/cmass index they are only well defined when data is standardized, how about the other index functions?}
 \textcolor{red}{When keeping in mind the concrete example of posterior samples outliers should not be an issue, if extreme outliers are found one should probably investigate origin.
 Also noise only directions, e.g.
-nuissance parameters, are also not an issue as they contain no structure, see BBH example.}
+nuisance parameters, are also not an issue as they contain no structure, see BBH example.}
 
@@
 Arguments for sphering are that scale effects should be extracted, we don't want to reproduce information already obtained by PCA.
-To my mind rescaling to an overall intervall is more appropriate here.
+To my mind rescaling to an overall interval is more appropriate here.
 Rotations between parameters should not affect the outcome, but will make interpretation more difficult.
-Standardising the variance will remove information about the posterior variance.
-Overall we are interested in strong association only, if these structures were affected by the standardisation procedure, it would most likely be unwelcome effects?}
+Standardizing the variance will remove information about the posterior variance.
+Overall we are interested in strong association only, if these structures were affected by the standardization procedure, it would most likely be unwelcome effects?}
 
@@
 Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
-One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuatins or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
-A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimisation for the identification of the maximum.
+One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
+A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimization for the identification of the maximum.
 On the other hand the global maximum may not be the only projection of interest, and one may want to observe the selected views in the context of the full distribution rather than a static view.
 These issues are addressed when combining projection pursuit with the grand tour [@CBCH94].
-In this case the properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshoothing and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
-A possible alogrithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
-In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
-
-\textcolor{red}{\cite{posse95b} discusses the optimisation, in particular that for most index functions and optimisers results are too local, largely dependent on starting point.
-Fiedmans suggestion of hybrid steppind search + steepest-ascent still remains fairly local.
+In this case the properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshooting and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
+A possible algorithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
+In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimists the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
+
+\textcolor{red}{\cite{posse95b} discusses the optimization, in particular that for most index functions and optimists results are too local, largely dependent on starting point.
+Fiedmans suggestion of hybrid stepping search + steepest-ascent still remains fairly local.
 \cite{posse90} suggested random search instead, which seems to perform better, but did not find pdf of that article, so not so sure what he actually does.}
@@
 Focusing now on $d=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
-We therefore draw on index functions that have been developped for the selection of variable pairs that show interesting features in a scatter plot, but we generalise the notion such that instead of variable pairs we consider any $2$ dimensional projection of the multivariate data distribution.
+We therefore draw on index functions that have been developed for the selection of variable pairs that show interesting features in a scatter plot, but we generalist the notion such that instead of variable pairs we consider any $2$ dimensional projection of the multivariate data distribution.
 It means that we consider the projection axes as input to the index functions, which will thus score the interestingness of the given projection.
-What we are optimising over are then the entries of the $p\times2$ dimensional projection matrix, considering that they obey orthonormality conditions.
+What we are optimizing over are then the entries of the $p\times2$ dimensional projection matrix, considering that they obey orthonormality conditions.
 Concretely here we consider the following types of index functions:\footnote{We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.}
@@
     + Skinny: The skinnyness of a polygon can be measured as the ratio of the perimeter to the area.
-We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a full circle, and values close to one or skinny polygons.
+We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $c_{skinny} = 0$ for a full circle, and values close to one or skinny polygons.
     + Stringy: The stringy index is a measure of the branching structure of the minimal spanning tree (MST), $c_{stringy} = \frac{diameter(MST)}{length(MST)}$ where the diameter is the longest connected path, and the length is the total length (sum of all edges), i.e.
@@
  \end{equation}
-    + Spline based measure: Given a spline model of the data distribution that is obtained via the R implentation in the \texttt{gam} function in the mgcv R package [@w16], we define the index based on the variance of the residuals:
+    + Spline based measure: Given a spline model of the data distribution that is obtained via the R implantation in the \texttt{gam} function in the mgcv R package [@w16], we define the index based on the variance of the residuals:
     \begin{equation}
@@
     \end{equation}
-  While $MIC$ is normalised to fall between $[0,1]$, the upper bound of $TIC$ depends on the sample size via the bounding function $B(n)$.
-$TIC$ was designed to be more stable test against null hypothesis of independence, using the full information of the characteristic matrix rather than just the maximum, and as such is expected to be better suited as a ppi.
+  While $MIC$ is normalized to fall between $[0,1]$, the upper bound of $TIC$ depends on the sample size via the bounding function $B(n)$.
+$TIC$ was designed to be more stable test against null hypothesis of independence, using the full information of the characteristic matrix rather than just the maximum, and as such is expected to be better suited as a PPI.
 
@@
 A consistent index means that small perturbations to the sample do not dramatically change the index value.
-This is particularly important to making optimisation feasible, small angles between projections correspond to small perturbations of the sample, and thus shold be small changes to index value.
+This is particularly important to making optimization feasible, small angles between projections correspond to small perturbations of the sample, and thus should be small changes to index value.
 \cite{posse95b} suggests that indexes should be resistant to features in a tail of the distribution, but this is debatable because one departure from normality that is interesting to detect are anomalous observations.
@@
 Lastly, because we need to compute the PPI over many projections, it needs to be fast to compute.
-These form the basis of the crteria upon which the scagnostic indexes, and the several alternative indexes are examined, as explained below.
+These form the basis of the criteria upon which the scagnostic indexes, and the several alternative indexes are examined, as explained below.
 
@@
 Fine structure like the parallel planes in the infamous RANDU data \cite{Marsaglia68} has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
-Structures with small squint angle are difficult to find, because the optmisation algorithm needs to get very close to begin hill-climbing to the optimum.
+Structures with small squint angle are difficult to find, because the optimization algorithm needs to get very close to begin hill-climbing to the optimum.
 The analyst doesn't have control over the data structure, but does have control over the PPI.
@@
 It should have smooth low values for noise projections and a clearly larger value, ideally with a big squint angle, for structured projections.
-The optimiser should be able to clearly see the optimal projections as different from noise.
+The optimizer should be able to clearly see the optimal projections as different from noise.
 To examine squintability, the PPI values are examined on interpolated tour paths between a noise projection and a distant structured projection.
@@
 \cite{Buja2004ComputationalMF} describes alternative interpolation paths based on Givens and Householder rotations which progress from basis to basis.
-It may be possible to ignore rotation invariance with these interpolations but there isn't a current implementation, primarily because the within-plane spin that is generated is distracting from a visualisation perspective.
+It may be possible to ignore rotation invariance with these interpolations but there isn't a current implementation, primarily because the within-plane spin that is generated is distracting from a visualization perspective.
 Rotation invariance is checked for the proposed PPIs by rotating the structured projection, within the plane.
@@
 -->
-- **speed**: Being fast to compute allows the index to be used in real-time in a guided tour, where the optimisation can be watched.
+- **speed**: Being fast to compute allows the index to be used in real-time in a guided tour, where the optimization can be watched.
 When the computations are shifted off-line, to watch in replay, computation times matter less.
@@
 
-Three families of data simulations are used for examining the behaviour of the index functions.
+Three families of data simulations are used for examining the behavior of the index functions.
 Each generates  structure in two variables, with the remaining variables containing various types of noise.
@@
 The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs.
-The scale for the holes index is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performace assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix).
+The scale for the holes index is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix).
 Similarly, the TIC index is re-scaled.
@@
 3.
-Use the guided tour to examine the ease of optimisation.
-This depends on having a relatovely smooth function, with structure visible from a distance.
-One index is optimised to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, to examine the similarity between PPIs.
+Use the guided tour to examine the ease of optimization.
+This depends on having a relatively smooth function, with structure visible from a distance.
+One index is optimized to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, to examine the similarity between PPIs.
 4. Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
@@
 Figure \ref{fig:plannedtour} shows the PPIs for a tour sequence between a nuisance and structured projection.
-Sample size is indicated by linetype: dashed being $n=100$ and solid is $n=1000$.
-The beginnnig of the sequence is the nuisance projection and the end is the structured projection.
+Sample size is indicated by line type: dashed being $n=100$ and solid is $n=1000$.
+The beginning of the sequence is the nuisance projection and the end is the structured projection.
 The index values for most PPIs increases substantially nearing the structured projection, indicating that they "see" the structure.
 Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
-[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
+[@Grimm2016]'s indexes, cor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
 The holes index easily detects the pipe, and can see it from a distance.
@@
 
-The guided tour combines optimisation with interpolation between pairs of planes.
-Target planes of the path are chosen to maximise the PPI.
-There are three derivative-free optimisation methods available in the guided tour: \verb# search_better_random# (1), \verb# search_better# (2), and \verb# search_geodesic# (3).
-Method 1 casts a wide net randomly generating projection planes, computing the PPIs and keeping the best projection, and method 2 conducts a localised maximum search.
+The guided tour combines optimization with interpolation between pairs of planes.
+Target planes of the path are chosen to maximize the PPI.
+There are three derivative-free optimization methods available in the guided tour: \verb# search_better_random# (1), \verb# search_better# (2), and \verb# search_geodesic# (3).
+Method 1 casts a wide net randomly generating projection planes, computing the PPIs and keeping the best projection, and method 2 conducts a localized maximum search.
 Method 3 is quite different: a local search is conducted to determine a promising direction, and then this direction is followed until the maximum in that direction is found.
-For all methods the optimisation is iterative, the best projections form target planes in the tour, the tour path is the interpolation to this target, and then a new search for a better projection is made, followed by the interpolation.
+For all methods the optimization is iterative, the best projections form target planes in the tour, the tour path is the interpolation to this target, and then a new search for a better projection is made, followed by the interpolation.
 For each projection during the interpolation steps, the PPI is recorded.
@@
 The stopping rule is that no better projections are found after a fixed number of tries, given a fixed tolerance value measuring difference.
-For method 1 and 2 two additional parameters control the optimisation: the search window $\alpha$, giving the maximum distance from the current plane in which projections are sampled, and the cooling factor, giving the incremental decrease in search window size.
+For method 1 and 2 two additional parameters control the optimization: the search window $\alpha$, giving the maximum distance from the current plane in which projections are sampled, and the cooling factor, giving the incremental decrease in search window size.
 Method 3 in principle also has two free parameters, which are however fixed in the current implementation.
@@
 
-For distributions and indexes with smooth behaviour and good squintability, method 3 is the most effective method for optimisation.
+For distributions and indexes with smooth behavior and good squintability, method 3 is the most effective method for optimization.
 If these two criteria are not met the method may still be useful, but only given an informed starting projection.
-In such cases we can follow a method similar to that proposed by Friedman (XXX REF): we break the optimisation in two distinct steps.
+In such cases we can follow a method similar to that proposed by Friedman (XXX REF): we break the optimization in two distinct steps.
 A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum.
@@
 
-As seen in Figure \ref{fig:plannedtour} the pipe structure has a small squint angle, and indeed using only method 3 does not uncover it in the six dimensional dataset.
-We therefore adapt the two step procedure described above.
-For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimise the TIC index.
-
-\textcolor{red}{Probably a good idea to give stopping condition to scouting phase, since at each step we may spend a long time looking for better projections.
-E.g.
-if maximum index value is 1, stop scoutig phase when reaching index value 0.9, i.e.
-should be in squintable distance of the index maximum, switch to geodesic optimiser at this point (only works if indeed the optimal view takes index values close to the maximum though)}
-
-The results are shown below, first Fig. \ref{fig:pipeFirstRun} shows how the considered index values evolve in the scouting phase.
-\textcolor{red}{Updated the plot, makes it look like skinny, 1-convex should be better index to optimise, however in practice this is not the case, probably because TIC is less sensitive to fluctuations especially also under rotation, making it more robust against getting stuck in local maxima.
-Maybe it will be better to leave scales free for this plot, since increase in TIC is very hard to see currently.
-Or just remove plot?
-not actually that intersting...}
+Despite the simple structure, the pipe is relatively difficult for the PPIs to find.
+For the TIC index, there is a fairly small squint angle.
+For the holes index, there are several local maxima, that divert the optimizer.
+There is a hint of this from Figure \ref{fig:plannedtour} because the initial projection (left side of trace) of purely noise variables has a higher index value than the linear combinations of noise and structured variables along the path.
+The uniform distribution was used to generate the noise variables, which has a higher PPI value than a normal distribution, yielding the higher initial value.
+In addition, a local maximum is observed whenever the pair of variables is one structured variable and one noise variable, because there is a lighter density in the center of the projection.
+
+The optimization is done in two stages, a scouting phase using method 2, and a refinement stage using method 3.
+For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimist the TIC index.
+The results are shown in Figures \ref{fig:pipeFirstRun} and \ref{fig:testpipe}.
+
+Figure \ref{fig:pipeFirstRun} shows the target projections (points)  selected during the scouting with method 2 on the TIC index.
+The focus is on the target projections rather than the interpolation between them, because the optimization is done off-line, and only the targets are used for the next step.
+The horizontal distance between the points in the plot reflects the relative geodesic distance between the planes in the 6D space.
+All of the other indexes are shown for interest.
+The TIC index value is generally low for this data, although it successfully detects the pipe.
+The holes, convex, skinny, and to some extent MIC, mirror the TIC performance.
+The holes differs in that it has some intermediate high values which are likely the indication of multi-modality of this index on this data.
+
+Figure \ref{fig:testpipe} shows the projection returned during the scouting phase (left) and the refinement phase (right).
+It was important to start the method 3 optimizer at the best projection returned during the scouting phase, to smoothly converge more closely to the ideal projection.
 
@@
 
-```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Index value on tour path when scouting for high TIC index values."}
+```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Scouting for the pipe using optimization method 2 on the TIC index.
+Other PPI values also shown for interest.
+Only the values for the target plane are shown because the optimization is done off-line.
+The TIC maximum value for this data is small,  even though it identifies the pipe.
+The holes, convex, skinny, and to some extent MIC, mirror the performance of TIC for this structure."}
 if(!file.exists("cache/pipeFirstRun.rda")){
@@
   geom_point(mapping=aes(color=(variable=="TIC"))) + 
+  geom_line(mapping=aes(color=(variable=="TIC"))) + 
   scale_colour_manual(values = c("black", "red")) +
-  facet_wrap(~variable, ncol=1) +
+  facet_wrap(~variable, ncol=1, strip.position = "right") +
   guides(color=FALSE) +
-  ylim(c(0,1))
-```
-
-We next show in Fig \ref{fig:testpipe} (left) the final view corresponding to the highest index found in the TIC scouting phase.
-While this is clearly not the optimal view of the pipe feature, we can already identify the feature in the distribution, and we expect to be able to find the maximum when using this plane as a starting point of the geodesic optimisation.
-Indeed performing a standard guided tour with the TIC index and the identified starting plane efficiently identifies the optimal view shown in the right plot.
-In this case the optimisation quickly converges, following a smooth, monotonic increas of the TIC index.
+  ylim(c(0,1)) +
+  xlab("Number of projections") +
+  ylab("PPI value") 
+```
 
@@
 
-```{r testpipe,  fig.height=4, fig.cap="Final view discovered by the scouting phase (left) and after second optimisation (right)."}
+```{r testpipe,  fig.height=4, fig.width=8, out.width = "0.8\\textwidth",  fig.cap="Projections returned by TIC optimisation: by the scouting phase (left) and refined by  optimisation method 3 (right), starting from the scouting phase projection."}
 iLast <- length(pipeTourFull)
@@
 dProj <- as.tibble(pipeResc %*% fProj)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point()
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
 iLast <- length(pipeTourFull2)
@@
 dProj <- as.tibble(pipeResc %*% fProj)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point()
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
 grid.arrange(p1, p2, ncol=2)
@@
 
-Given the results from Fig.\ref{fig:plannedtour} we expect to be able to efficiently identify the sine view in the 6-d distribution, using only method 3 for the optimisation, using e.g.
+Given the results from Fig.\ref{fig:plannedtour} we expect to be able to efficiently identify the sine view in the 6-d distribution, using only method 3 for the optimization, using e.g.
 the splines2d index.
@@
 Here we only note that even for anchor planes highlighted by blue vertical lines the index sometimes decreases because of this dependence.
-These are however not dramatic effects because this search method is closely related to how interpolation is done (a different picture would be optained using method 2 which often leads to large drops in index values after the interpolation step).
-
-Here there are four index functions that appear to be suitable for the optimisation: splines2d, dcor2d, MIC and TIC.
-On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimisation when considering them as index functions.
+These are however not dramatic effects because this search method is closely related to how interpolation is done (a different picture would be obtained using method 2 which often leads to large drops in index values after the interpolation step).
+
+Here there are four index functions that appear to be suitable for the optimization: splines2d, dcor2d, MIC and TIC.
+On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimization when considering them as index functions.
 The final view identified by the guided tour is shown in Figure \ref{fig:findsine}.
@@
 
-The spiral is the most challenging structure to detect amonst the distribution considered here (compare also Posses results?).
-Here we explore the role of dimensionality in the index optimisation, using the skinny index that was found to be sensitive to this structure.
-We use a setup similar to that introduced for detecting the pipe structure, i.e.\ we use the search\_better optimisation of the guided tour, disabeling the cooling and with a large search window, while sampling a large number of random projections (up to 5000) for each step.
+The spiral is the most challenging structure to detect amongst the distribution considered here (compare also Posses results?).
+Here we explore the role of dimensionality in the index optimization, using the skinny index that was found to be sensitive to this structure.
+We use a setup similar to that introduced for detecting the pipe structure, i.e.\ we use the search\_better optimization of the guided tour, disabling the cooling and with a large search window, while sampling a large number of random projections (up to 5000) for each step.
 We consider distributions in d=4,6,8 dimensions and compare features found in the tour.
-In d=4 dimensions the optimisation identifies a view showing the spiral after 33 steps, and we limit the optimiser to consider only 33 steps for the study in 6 and 8 dimensions for direct comparison.
-Note however that the spiral is typically also not identified with much larger number of steps, the limiting factor being rather the numver of random projections sampled and rotation non-invariance of the index prohibiting the convergence, in particular in higher dimensions.
-Results are shown in the top panel Fig.~\ref{fig:findspiral}, showing the pariwise distance of the 33 planes found in 4, 6 and 8 dimensions (left), and comparing the evolution of the distance from the plane containing the spiral view with progression of the optimisation (right).
+In d=4 dimensions the optimization identifies a view showing the spiral after 33 steps, and we limit the optimizer to consider only 33 steps for the study in 6 and 8 dimensions for direct comparison.
+Note however that the spiral is typically also not identified with much larger number of steps, the limiting factor being rather the number of random projections sampled and rotation non-invariance of the index prohibiting the convergence, in particular in higher dimensions.
+Results are shown in the top panel Fig.~\ref{fig:findspiral}, showing the pairwise distance of the 33 planes found in 4, 6 and 8 dimensions (left), and comparing the evolution of the distance from the plane containing the spiral view with progression of the optimization (right).
 We see that the pairwise distances grow with d (what do we learn from that?).
@@
 At medium distances from the special view we often observe a jump to higher distances as consequence of in-plane rotation dependence for such partial views.
-This stabilises closer to the special view, however it relies on randomly identifying a projection within this angle.
+This stabilizes closer to the special view, however it relies on randomly identifying a projection within this angle.
 As a consequence increasing dimensionality dramatically reduces the chance of identifying structures with high index values.
@@
 
-## Rotation dependence
+## Rotational invariance or not
 \label{sec:rot}
 
-We study rotation dependence using the sine function distribution, and tracing the index values as a function of roation angle.
+We study rotation dependence using the sine function distribution, and tracing the index values as a function of rotation angle.
 Results are shown in Figure \ref{fig:rotationDep}.
@@
 The dcor2d, splines2d and TIC index show consistent patterns in rotation dependence, on the other hand the scagnostics index skinny is found to have sizable random fluctuations as the basis is rotated.
+
 ```{r rotationDep, out.width=".6\\textwidth", fig.cap="PPI for rotations of the sine 1000 data, to examine rotation invariance.
@@
 
-## Note on calculation time
+## Computational performance
+
 \label{sec:timer}
@@
 We see that the evaluation of the spline index shows weak dependence on the sample size, making it the least efficient for very small data sets, but one of the fastest when considering large data sets.
-The situation is similar for calculating the scagnositics measures, which are however faster to compute on datasets of around 10 points.
+The situation is similar for calculating the scagnostics measures, which are however faster to compute on datasets of around 10 points.
 This is because the implementation is first binning the data points, which is fast compared to the evaluation that requires triangulation of the binned points.
-The distance correlation based measure exhibits strong increase with sample size, we note however that the implementation also allowes for evaluation on binned data points which we have not considered.
+The distance correlation based measure exhibits strong increase with sample size, we note however that the implementation also allows for evaluation on binned data points which we have not considered.
 Finally for the MINE family the computing time increases exponentially as can be expected from the dependence on the sample size, where we see that the improved $MIC_e$ algorithm is somewhat faster than the original implementation, see also @JMLRv1715308.
@@
 
-### Scagnostics binning
+### Effect of binning in scagnostics
+
 The implementation of scagnostics will always perform a binning step first, before evaluating the measures.
@@
 A nice example can be constructed based on the RANDU dataset available in R.
-We use the MIC index and compare different parametrisations of the maximum binning by $\alpha$, where $B(n) = n^{\alpha}$.
-We compare in Figure \ref{fig:randuEx} the final view, index value and computing time obtained when optimising the MIC index with values $\alpha = 0.6, 0.7, 0.8$.
+We use the MIC index and compare different parametrization of the maximum binning by $\alpha$, where $B(n) = n^{\alpha}$.
+We compare in Figure \ref{fig:randuEx} the final view, index value and computing time obtained when optimizing the MIC index with values $\alpha = 0.6, 0.7, 0.8$.
 Despite the low dimensionality, the small quint angle of the structure means that the identification of the maximum depends strongly on the starting conditions.
@@
 
-Moreover we can either select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calulate the mean from the remaining evaluations.
+Moreover we can either select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calculate the mean from the remaining evaluations.
 Only the mean will be considered here.
@@
 
+<!--
 ### Rotation invariance
 Do we want to say anything here? big issue for splines2D measure, but also scagnostics
+-->
 
@@
 
-Our results can be summarised by evaluating and comparing the advantages and disadvanteges of each index function according to the criteria presented above.
+Our results can be summarized by evaluating and comparing the advantages and disadvantages of each index function according to the criteria presented above.
 An overview is given in Table \ref{tab:summary}.
@@
 The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures, making the set flexible in applications (the same is however not true for each single index viewed on its own).
-On the other hand we have found that they are typically too noisy making optimisation of the function challenging, which also results in poor squinatbility.
-In particular the stringy index cannot be used with current optimisers.
-The index functions from @mbgraphic share similar behaviour as they are both found to evolve fairly smoothly between interpolated projections, but have been shown to be strongly rotation dependent.
+On the other hand we have found that they are typically too noisy making optimization of the function challenging, which also results in poor squint ability.
+In particular the stringy index cannot be used with current optimizers.
+The index functions from @mbgraphic share similar behavior as they are both found to evolve fairly smoothly between interpolated projections, but have been shown to be strongly rotation dependent.
 The splines2D index is fast to calculate and found to be the most efficient method for identifying functional dependence.
@@
 While these indexes are slow to compute especially on large datasets, we have found them to be the most general, allowing good discrimination even with noisy structures.
-In addition we observed good squinatbility for all three examples considered, given large datasets.
-Results in Section \ref{sec:params} suggest that tuning of the MINE parameters may result in improved behaviour for smaller datasets.
+In addition we observed good squintability for all three examples considered, given large datasets.
+Results in Section \ref{sec:params} suggest that tuning of the MINE parameters may result in improved behavior for smaller datasets.
 
@@
 \caption{Summary of findings, showing to what extend the considered index functions pass the criteria for a good PPI.
-"\checkmark" symbols good behaviour, "$\cdot$" symbols some issues while "$\times$" symbols failure on the corresponding criteria.
+"\checkmark" symbols good behavior, "$\cdot$" symbols some issues while "$\times$" symbols failure on the corresponding criteria.
 Each index has particular strengths and drawbacks and selection must be guided by the considered example, see text for details.}
@@
 All seven PP indexes are applied to the data.
-Figure \ref{fig:bbhGuided} showing the projections that maximise three of the indexes.
+Figure \ref{fig:bbhGuided} showing the projections that maximize three of the indexes.
 TIC and splines2D indexes identify very similar projections, that are based on the three parameters, dec, time and ra.
-This is ot be expected based on the pairwise scatterplots.
+This is to be expected based on the pairwise scatterplots.
 \textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
-On the other hand, the 1-convex index finds a very different view, but actually the problem is that the optimisation doesn't adequately reach a maximum for this index.
+On the other hand, the 1-convex index finds a very different view, but actually the problem is that the optimization doesn't adequately reach a maximum for this index.
 
@@
 The variables time, dec and ra are dropped from the data, and PP is applied to the remaining 8D space.
-Figure \ref{fig:bbhGuided2} shows the projections which maximise the TIC, splines2D and 1-convex indices.
-The results provide similar information as laready learned from the scatterplot matrix (Figure \ref{bbhSimulation}).
+Figure \ref{fig:bbhGuided2} shows the projections which maximize the TIC, splines2D and 1-convex indices.
+The results provide similar information as already learned from the scatterplot matrix (Figure \ref{bbhSimulation}).
 The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2D index.
@@
 
-The initial conditions for the optimisation, and the subset of variables used, can have a large effect on the projections returned.
+The initial conditions for the optimization, and the subset of variables used, can have a large effect on the projections returned.
 We illustrate this using only the splines2D index, and find that there is one more association that can be learned that was masked earlier.
@@
 \cite{Cook:2007:IDG:1557227} then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
-They are defined such that each one is maximised for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
+They are defined such that each one is maximized for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
 It follows that for either index both large and small values signal deviation from the normal distribution, and given a normal distribution we expect to find "average" index values rather than values close to zero.

### 2018-12-11 scagnostics.Rmd
Replacing melt by gather everywhere.

@@
 library(tidyverse)
-library(reshape2)
 library(scagnostics)
@@
 for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
-  distData <- melt(ds)
+  distData <- gather(ds, variable, value)
   pC <- ggplot(distData, aes(value, color=variable)) +
@@
 The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection."}
-#fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
 fullResMeltPlanned <- fullResPlanned %>%
@@
   pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
-  pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
-  save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
+  #pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
+  #never used?
+  #save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
+  save(pipeTourRes, file = "cache/pipeFirstRun.rda")
 } else {
@@
   select(-cmass) %>%
-  melt(id=c("t", "name", "size")) %>%
+  gather(variable, value, -t, -name, -size) %>%
   filter(t %in% which(attributes(interpolate(pipeTour))$new_basis))
@@
   pipeTourRes2 <- getProj(pipe1000, pipeTourFull2, "Pipe", 1000)
-  pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
-  save(pipeTour2, pipeTourFull2, pipeTourRes2, pipeTourResMelt2, file = "cache/refinepipe.rda")
+  #pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
+  #save(pipeTour2, pipeTourFull2, pipeTourRes2, pipeTourResMelt2, file = "cache/refinepipe.rda")
+  save(pipeTour2, pipeTourFull2, pipeTourRes2, file = "cache/refinepipe.rda")
 } else {
@@
   sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
-  sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
-  save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
+  #sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
+  #save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
+  save(sineResc, sineTour, sineTourFull, sineTourRes, file = "cache/findsine.rda")
 } else {
@@
   select(-cmass) %>%
-  melt(id=c("t", "name", "size"))
+  gather(variable, value, -t, -name, -size)
 ggplot(sineTourResMelt, aes(x=t, y=value)) +
@@
   scagBinningDf <- getScagComp(spiral1000, t2full)
-  scagMelt <- melt(scagBinningDf, id=c("t")) %>%
+  scagMelt <- gather(scagBinningDf, variable, value, -t) %>%
     mutate(nbin = as.numeric(str_sub(variable, start = -2))) %>%
@@
   scagSmooth <- getScagSmooth(spiral1000, as.list(t2full))
-  scagSmoothMelt <- melt(scagSmooth, id=c("t", "method", "alpha"))
+  scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
   save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")

### 2018-12-11 scagnostics.Rmd
Updated smoothing plot.

@@
 
-```{r spiralScagSmoothing, fig.cap="Comparing the traces of the three scagnostics indices when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (left) or after jittering the projected datapoints with some amount $\\beta$ (right).
+\textcolor{red}{I fixed the y axis to be between 0 and 1, which highlights that indeed we get much smoother behaviour, on the other hand theres still large relative jumps observed in all traces, and a smaller y axis scale may be preferred to show this more clearely.}
+
+```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
 While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered."}
@@
 if(!file.exists("cache/spiralScagSmoothing.rda")){
-  scagSmooth <- getScagSmooth(spiral1000, as.list(t2full))
+  scagSmooth <- getScagSmooth(spiral1000, as.list(t1full))
   scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
@@
 
+scagSmoothMelt <- scagSmoothMelt %>%
+  filter(variable != "convex") %>%
+  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
 ggplot(scagSmoothMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(alpha))) + 
-  facet_wrap(variable~method, ncol=2, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
+  geom_line(aes(linetype=factor(method))) + 
+  theme(legend.position="none") +
+  ylim(0,1) +
+  facet_grid(alpha~variable) +
+  xlab("Sequence of projections (t)") +
+  ylab("PPI value")
 

### 2018-12-11 scagnostics.Rmd
Replacing Fig. by Figure in all text.

@@
 
-Given the results from Fig.\ref{fig:plannedtour} we expect to be able to efficiently identify the sine view in the 6-d distribution, using only method 3 for the optimization, using e.g.
+Given the results from Figure\ref{fig:plannedtour} we expect to be able to efficiently identify the sine view in the 6-d distribution, using only method 3 for the optimization, using e.g.
 the splines2d index.
@@
 Note however that the spiral is typically also not identified with much larger number of steps, the limiting factor being rather the number of random projections sampled and rotation non-invariance of the index prohibiting the convergence, in particular in higher dimensions.
-Results are shown in the top panel Fig.~\ref{fig:findspiral}, showing the pairwise distance of the 33 planes found in 4, 6 and 8 dimensions (left), and comparing the evolution of the distance from the plane containing the spiral view with progression of the optimization (right).
+Results are shown in the top panel Figure \ref{fig:findspiral}, showing the pairwise distance of the 33 planes found in 4, 6 and 8 dimensions (left), and comparing the evolution of the distance from the plane containing the spiral view with progression of the optimization (right).
 We see that the pairwise distances grow with d (what do we learn from that?).
@@
 
-Fig.
-\ref{fig:bbhGuided3} shows six maxima obtained by different starts, for two types of parameters: first, spin related parameters (i.e.
+Figure \ref{fig:bbhGuided3} shows six maxima obtained by different starts, for two types of parameters: first, spin related parameters (i.e.
 alpha, theta\_jn, chi\_tot and chi\_p), and second position related parameters (i.e.
@@
 Choosing a different subset of variables reveals something new.
-The subspace of m1, ra, chi\_tot, alpha, distance, dec produces a more refined view of Fig.
-\ref{fig:bbhGuided3} projection f.
+The subspace of m1, ra, chi\_tot, alpha, distance, dec produces a more refined view of Figure \ref{fig:bbhGuided3} projection f.
 When alpha contributes in contrast to dec, the relationship between the points is almost perfectly on a curve.

### 2018-12-12 scagnostics.Rmd
Updated spiral guided tour example.

@@
   }
+
 if(!file.exists("cache/findspiral.rda")){
@@
   i <- 1
-  for(d in c(4,6,8)){
-    set.seed(1779+d)
+  for(d in c(4,5,6)){
+    set.seed(58958)
     specialPlane <- matrix(c(rep(0,d-2),1,0,rep(0,d-1),1),ncol=2)
-    spiralD <- spiralData(d, 1000) %>% scale() %>% as_tibble()
-    spiralTour <- save_history(spiralD, guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
+    spiralD <- spiral1000 %>%
+      purrr::when(d<6 ~ select(., -V4), ~ .) %>%
+      purrr::when(d<5 ~ select(., -V3), ~ .)
+    spiralTour <- save_history(as.matrix(spiralD),
+                               guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
                                                     max.tries = 5000, cooling = 1, alpha = pi),
                               max_bases = maxBases)
-    if(d==4){maxBases <- min(maxBases, length(spiralTour))}
+    #if(d==4){maxBases <- min(maxBases, length(spiralTour))}
+    nBases <- length(spiralTour)
     distDf <- rbind(distDf, tibble(d=d, distance=distanceDist(spiralTour)))
-    distToSp <- rbind(distToSp, tibble(d=d, distance=distanceToSp(spiralTour, specialPlane), t=1:maxBases))
-    fView <- as.tibble(as.matrix(spiralD) %*% as.list(spiralTour)[[maxBases]])
+    distToSp <- rbind(distToSp, tibble(d=d, distance=distanceToSp(spiralTour, specialPlane), t=1:nBases))
+    fView <- as.tibble(as.matrix(spiralD) %*% as.list(spiralTour)[[nBases]])
     fViewL[[i]] <- ggplot(fView, aes(V1, V2)) + geom_point()
@@
 
-p1 <- ggplot(distDf, aes(group=d, y=distance, x=d)) + geom_boxplot()
+p1 <- ggplot(distDf, aes(group=d, y=distance, x=d)) +
+  geom_boxplot() +
+  scale_x_continuous(breaks = c(4, 5, 6))
 p2 <- ggplot(distToSp, aes(color=as.factor(d),x =t, y=distance)) +

### 2018-12-12 scagnostics.Rmd
Updated spiral guided tour example text accordingly.

@@
 We use a setup similar to that introduced for detecting the pipe structure, i.e.\ we use the search\_better optimization of the guided tour, disabling the cooling and with a large search window, while sampling a large number of random projections (up to 5000) for each step.
-We consider distributions in d=4,6,8 dimensions and compare features found in the tour.
-In d=4 dimensions the optimization identifies a view showing the spiral after 33 steps, and we limit the optimizer to consider only 33 steps for the study in 6 and 8 dimensions for direct comparison.
-Note however that the spiral is typically also not identified with much larger number of steps, the limiting factor being rather the number of random projections sampled and rotation non-invariance of the index prohibiting the convergence, in particular in higher dimensions.
-Results are shown in the top panel Figure \ref{fig:findspiral}, showing the pairwise distance of the 33 planes found in 4, 6 and 8 dimensions (left), and comparing the evolution of the distance from the plane containing the spiral view with progression of the optimization (right).
+We consider distributions in d=4,5,6 dimensions (subsets of the spiral dataset) and compare features found in the tour.
+In d=4 dimensions the optimization identifies a view showing the spiral after 39 steps.
+For larger values of d the optimization does not uncover the spiral view, and we limit the optimizer to 100 steps to compare the so obtained bases to the results found in four dimensions.
+Results are shown in the top panel Figure \ref{fig:findspiral}.
+The top left is showing the pairwise distances between all planes found by the optimization, note that while d=5,6 contains 100 planes, d=4 only contains 39 planes.
+The top right is then comparing the evolution of the distance from the plane containing the spiral view with progression of the optimization.
 We see that the pairwise distances grow with d (what do we learn from that?).
-Moreover we find that the distance to the special view is far from monotonically decreasing towards zero, and large jumps are possible.
+Moreover we find that the distance to the special view is far from monotonically decreasing towards zero, and large jumps are possible as a consequence of rotation dependence.
 Note that the spiral itself is approximately rotation invariant, but partial views are not.
@@
 
-```{r findspiral, out.width="\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with d = 4, 6, 8.
+```{r findspiral, out.width="\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with d = 4, 5, 6.
 The top left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the top right shows the evolution of distance to the plane containing the spiral as the index is being optimised.
-The bottom row shows the final view obtained for d = 4 (left), d = 6 (middle) and d = 8 (right) after 33 steps in the guided tour.", results="hide"}
+The bottom row shows the final view obtained for d = 4 (left), d = 5 (middle) and d = 6 (right).", results="hide"}
 

### 2018-12-12 scagnostics.Rmd
Tentative Fig 4 to potentially replace Fig 3.

@@
 
+## Tentative
+try how this looks going first to V1 vs V5, then to V5 vs V6
+```{r wIntermediatePath}
+if(!file.exists("cache/wIntermediate.rda")){
+  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
+  mInt <- matrix(c(1,0,0,0,0,0,0,0,0,0,1,0),ncol=2)
+  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
+  #this is silly but seems that first two entries are being ignored, so need some fake entries
+  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
+  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
+  t3 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,mInt,m2)))
+  t3full <- as.list(interpolate(t3))
+  fullResPlanned <- getProj(sin100, t3full, "sin", 100) %>%
+    rbind(getProj(sin1000, t3full, "sin", 1000)) %>%
+    rbind(getProj(spiral100, t3full, "spiral", 100)) %>%
+    rbind(getProj(spiral1000, t3full, "spiral", 1000)) %>%
+    rbind(getProj(pipe100, t3full, "pipe", 100)) %>%
+    rbind(getProj(pipe1000, t3full, "pipe", 1000))
+  save(t3, t3full, fullResPlanned, file = "cache/wIntermediate.rda")
+} else {
+  load("cache/wIntermediate.rda")
+}
+```
+
+```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between a nuisance and a structured projection (solid lines: $n=1000$; dashed: $n=100$).
+Most of the indexes have peaks at the end of the sequence, indicating that they see the structure.
+The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection."}
+fullResMeltPlanned <- fullResPlanned %>%
+  mutate(convex = 1-convex) %>% 
+  rowwise() %>%
+  mutate(holes = rescaleHoles(holes)) %>%
+  gather(PPI, value, -t, -name, -size) %>%
+  filter(PPI != "cmass") %>%
+  mutate(PPI = fct_relevel(PPI, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+fullResMeltPlanned %>% 
+  ggplot(aes(x=t, y=value)) + 
+  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
+  ylim(c(0,1)) +
+  facet_grid(PPI~name) + 
+  theme(legend.position="none") + xlab("Sequence of projections (t)") + ylab("PPI value")
+```
+
 ## Optimisation check with the guided tour

### 2018-12-12 scagnostics.Rmd
done with the new sections, next all need to revisit all sections

--- 2018-12-12 scagnostics.Rmd
@@
 [@Grimm2016] explored the behavior of scagnostics for selecting variables, and proposed two more, based on smoothing splines and distance correlation, that have some nicer properties.
-In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518] round out our collection of metrics.
+In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], and computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308], round out the collection of metrics.
 \footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber85,jones87}.}
@@
 If a particular orientation is more optimal, this will get lost as the projection shown pays no attention to orientation.
-\cite{Buja2004ComputationalMF} describes alternative interpolation paths based on Givens and Householder rotations which progress from basis to basis.
+@BCAH05 describes alternative interpolation paths based on Givens and Householder rotations which progress from basis to basis.
 It may be possible to ignore rotation invariance with these interpolations but there isn't a current implementation, primarily because the within-plane spin that is generated is distracting from a visualization perspective.
@@
 
-Figure \ref{fig:plannedtour} shows the PPIs for a tour sequence between a nuisance and structured projection.
+Figure \ref{fig:wIntermediatePath} shows the PPIs for a tour sequence between a nuisance and structured projection.
+A long sequence is generated where the path runs from V1-V2, V1-V5, V5-V6, in order to see some of the intricacies of holes index.
 Sample size is indicated by line type: dashed being $n=100$ and solid is $n=1000$.
@@
 
+<!--
 ```{r plannedtourpath}
@@
 
+## Tentative
+try how this looks going first to V1 vs V5, then to V5 vs V6
+-->
+
+```{r wIntermediatePath}
+if(!file.exists("cache/wIntermediate.rda")){
+  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
+  mInt <- matrix(c(1,0,0,0,0,0,0,0,0,0,1,0),ncol=2)
+  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
+  #this is silly but seems that first two entries are being ignored, so need some fake entries
+  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
+  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
+  t3 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,mInt,m2)))
+  t3full <- as.list(interpolate(t3))
+  fullResPlanned <- getProj(sin100, t3full, "sin", 100) %>%
+    rbind(getProj(sin1000, t3full, "sin", 1000)) %>%
+    rbind(getProj(spiral100, t3full, "spiral", 100)) %>%
+    rbind(getProj(spiral1000, t3full, "spiral", 1000)) %>%
+    rbind(getProj(pipe100, t3full, "pipe", 100)) %>%
+    rbind(getProj(pipe1000, t3full, "pipe", 1000))
+  save(t3, t3full, fullResPlanned, file = "cache/wIntermediate.rda")
+} else {
+  load("cache/wIntermediate.rda")
+}
+```
+
+```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between nuisance and structured projections, following V1-V2 to V1-V5 to V5-V6.
+Solid lines are the larger sample size, $n=1000$, and dashed are smaller sample size, $n=100$.
+Peaks at the end of the sequence indicate the index sees the structure.
+The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection.
+Holes only responds to the pipe, and is a multimodal function for this data with a local maximum at V1-V5.
+The scagnostics tend to be sensitive to sample size.
+"}
+fullResMeltPlanned <- fullResPlanned %>%
+  mutate(convex = 1-convex) %>% 
+  rowwise() %>%
+  mutate(holes = rescaleHoles(holes)) %>%
+  gather(PPI, value, -t, -name, -size) %>%
+  filter(PPI != "cmass") %>%
+  mutate(PPI = fct_relevel(PPI, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+fullResMeltPlanned %>% 
+  ggplot(aes(x=t, y=value)) + 
+  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
+  ylim(c(0,1)) +
+  facet_grid(PPI~name) + 
+  theme(legend.position="none") + xlab("Sequence of projections (t)") + ylab("PPI value")
+```
+
 ## Optimisation check with the guided tour
@@
 For the holes index, there are several local maxima, that divert the optimizer.
-There is a hint of this from Figure \ref{fig:plannedtour} because the initial projection (left side of trace) of purely noise variables has a higher index value than the linear combinations of noise and structured variables along the path.
+There is a hint of this from Figure \ref{fig:wIntermediate} because the initial projection (left side of trace) of purely noise variables has a higher index value than the linear combinations of noise and structured variables along the path.
 The uniform distribution was used to generate the noise variables, which has a higher PPI value than a normal distribution, yielding the higher initial value.
@@
 
-Given the results from Figure\ref{fig:plannedtour} we expect to be able to efficiently identify the sine view in the 6-d distribution, using only method 3 for the optimization, using e.g.
-the splines2d index.
-Indeed the guided tour quickly converges in this case, the evolution of the index functions is shown in Fig \ref{fig:findsine}.
-The only difficulties stem from the rotation dependence of the index function as well as the structured distribution, this will be discussed in some more details below.
+Given the patterns in Figure\ref{fig:wIntermediate} it would be expected that the sine could be found easily, using only optimization method 3 using the splines2d, dcor2d, MIC or TIC indexes.
+This is examined in Figure \ref{fig:findsine}.
+Optimization is conducted using the splines2d index, and the trace of the PPI over the optimisation is shown, along with the PPI values for the other indexes over that path.
+The vertical blue lines indicate anchor bases, where the optimizer stops, and does a new search.
+The distance between anchor planes is smaller as the maximum is neared.
+
+The only complications arise from a lack of rotation invariance of the splines2d index.
+It is not easily visible here, but it is possible that the best projection will have a lower PPI.
+The index changes depending on the basis used to define the plane, but the geodesic interpolation conducted by the tour uses any suitable basis to describe the plane, ignoring that which optimizes the PPI.
+This is discussed in section \ref{sec:rot}.
+
+<!--, this will be discussed in some more details below.
 Here we only note that even for anchor planes highlighted by blue vertical lines the index sometimes decreases because of this dependence.
-These are however not dramatic effects because this search method is closely related to how interpolation is done (a different picture would be obtained using method 2 which often leads to large drops in index values after the interpolation step).
-
-Here there are four index functions that appear to be suitable for the optimization: splines2d, dcor2d, MIC and TIC.
+These are however not dramatic effects because this search method is closely related to how interpolation is done (a different picture would be obtained using method 2 which often leads to large drops in index values after the interpolation step).-->
+
+<!-- Here there are four index functions that appear to be suitable for the optimization: splines2d, dcor2d, MIC and TIC.
 On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimization when considering them as index functions.
@@
 
-\textcolor{red}{This isn't actually the "ideal" view, could be because of rotation dependence??}
-
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising the splines2d index for the Sine 1000 dataset. Anchor planes are marked by the blue vertical lines.", results="hide"}
+\textcolor{red}{This isn't actually the "ideal" view, could be because of rotation dependence??}-->
+
+```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising  the splines2d index, using method 3, for the sine data, with $n=1000$.
+Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima.
+The sine is found relatively easily, by splines2D, and it si indicated that MIC, TIC, dcor2D and convex would also likely find this structure.", results="hide"}
 set.seed(2018)
@@
   guides(color=FALSE) +
-  ylim(c(0,1))
-```
-
+  ylim(c(0,1)) + xlab("Number of projections") + ylab("PPI value")
+```
+
+<!--
+Di says probably we can skip this plot
 ```{r testsine, out.width=".5\\textwidth", fig.cap="Final view identified by the optimisation."}
@@
 ```
+-->
 
@@
 
-The spiral is the most challenging structure to detect amongst the distribution considered here (compare also Posses results?).
-Here we explore the role of dimensionality in the index optimization, using the skinny index that was found to be sensitive to this structure.
+The spiral is the most challenging structure to detect because it has a small squint angle [@posse95a], especially as the ratio of noise to structure dimensions increases.
+This is explored in this section, using optimisation method 2 to scout the space for approximate maxima first.
+The skinny scagnostic index is used because it was observed (Figure \ref{fig:wIntermediate}) to be sensitive to this structure, although the noisiness of the index might be problematic.
+The stringy appears to be more sensitive to the spiral, but it has a much small squint angle.
+
+<!-- Here we explore the role of dimensionality in the index optimization, using the skinny index that was found to be sensitive to this structure.
 We use a setup similar to that introduced for detecting the pipe structure, i.e.\ we use the search\_better optimization of the guided tour, disabling the cooling and with a large search window, while sampling a large number of random projections (up to 5000) for each step.
-We consider distributions in d=4,5,6 dimensions (subsets of the spiral dataset) and compare features found in the tour.
+-->
+
+The search is conducted for $d=4,5,6$ which would correspond to 2, 3 and 4 noise dimensions respectively.
+In addition we examine the distance between planes, using a Frobenius norm, as defined by Equation 2 of @BCAH05, and available in the ```proj_dist``` function in the tourr package, to compare searches with squint angle across dimensions.
+The distance between planes is related to squint angle, how far away from the ideal projection, can the structure be glimpsed.
+In order for the optimizer to find the spiral, the distance between planes would need to be smaller than the squint angle.
+Figure \ref{fig:findspiral} summarizes the results.
+When $d=4$ the scouting method effectively finds the spiral.
+Plot (a) shows the side-by-side boxplots of distance between ideal plane and planes examined during the optimization, for $d=4,5,6$.
+These are on average smaller for the lower dimension, and gradually increase as dimension increases.
+This is an indication of the extra computation needed to brute force find the spiral as noise dimensions increase.
+Plot (b) shows these as a function of the optimization procedure, where it can be seen that only when $d=4$ does it converge to the ideal.
+Its likely that expanding the search space should result in uncovering the spiral in higher dimensions.
+
+<!-- We consider distributions in d=4,5,6 dimensions (subsets of the spiral dataset) and compare features found in the tour.
 In d=4 dimensions the optimization identifies a view showing the spiral after 39 steps.
@@
 This stabilizes closer to the special view, however it relies on randomly identifying a projection within this angle.
-As a consequence increasing dimensionality dramatically reduces the chance of identifying structures with high index values.
+As a consequence increasing dimensionality dramatically reduces the chance of identifying structures with high index values.-->
 
@@
   maxBases <- 100
+#  maxBases <- 1000
   fViewL <- list()
@@
   for(d in c(4,5,6)){
+  #d = 6
     set.seed(58958)
@@
     spiralTour <- save_history(as.matrix(spiralD),
-                               guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better,
-                                                    max.tries = 5000, cooling = 1, alpha = pi),
+                               guided_tour(scagIndex("Skinny"), 
+                                  search_f = tourr:::search_better,
+                                  max.tries = 5000, 
+                                  #max.tries = 50000, 
+                                  cooling = 1, alpha = pi),
                               max_bases = maxBases)
@@
   geom_boxplot() +
-  scale_x_continuous(breaks = c(4, 5, 6))
+  scale_x_continuous(breaks = c(4, 5, 6)) + ggtitle("(a)")
 p2 <- ggplot(distToSp, aes(color=as.factor(d),x =t, y=distance)) +
   geom_point() +
-  labs(y = "Distance to special view", color = "d")
+  labs(y = "Distance to special view", color = "d") + ggtitle("(b)")
 
@@
 
-We study rotation dependence using the sine function distribution, and tracing the index values as a function of rotation angle.
+Rotational invariance is examined using the sine data (V5-V6), computing PPI for different rotations within the 2D plane, parametrized by angle.
 Results are shown in Figure \ref{fig:rotationDep}.
-Several indexes are found to be invariant, holes, convex and MIC.
-The dcor2d, splines2d and TIC index show consistent patterns in rotation dependence, on the other hand the scagnostics index skinny is found to have sizable random fluctuations as the basis is rotated.
+Several indexes are invariant, holes, convex and MIC, because their value is constant around rotations.
+The dcor2d, splines2d and TIC index are clearly not rotationally invariant because the value changes depending on the rotation.
+The scagnostics indexes are approximately rotationally invariant, but particularly the skinny index has some random variation depending on rotation.
 
@@
 \label{sec:timer}
-Computing time differs for each index, and depends strongly on the number of data points (as well as probably some dependence on distribution type), here we use as an example samples of points in a 6-d solid sphere (obtained via the geozoo package [@geozoo]) and calculate the computing time for different indices over 100 interpolated grand tour projections as a function of the sample size.
-\textcolor{red}{Since projection is fast compared to index evaluation, the number of original dimensions is not very important here.
+Computational time is important for using the PPIs with the guided tour, online.
+Figure \ref{fig:timer} summarizes performance for each PPI.
+For simplicity, data with sample sizes ranging from 100-10000 are drawn from a 6-d solid sphere, using the geozoo package [@geozoo].
+The time to compute the PPIs over 100 interpolated grand tour projections is recorded.
+The scagnostics PPI are computed as a bundle, since this is the code base, and that major computational constraint is common to all the scagnostics.
+There are two versions of the MIC and TIC algorithm, labelled MINE and MINE E, the second being a newer algorithm which improves their computational performance.
+
+The results are interesting. The scagnostic indexes and splines2D are very fast regardless of sample size. MIC, TIC (both versions) and dcor2D slow rapidly as sample size increases.
+
+<!--\textcolor{red}{Since projection is fast compared to index evaluation, the number of original dimensions is not very important here.
 On the other hand different distributions may impact the computing time.}
@@
 The distance correlation based measure exhibits strong increase with sample size, we note however that the implementation also allows for evaluation on binned data points which we have not considered.
-Finally for the MINE family the computing time increases exponentially as can be expected from the dependence on the sample size, where we see that the improved $MIC_e$ algorithm is somewhat faster than the original implementation, see also @JMLRv1715308.
+Finally for the MINE family the computing time increases exponentially as can be expected from the dependence on the sample size, where we see that the improved $MIC_e$ algorithm is somewhat faster than the original implementation, see also @JMLRv1715308.-->
 
@@
 
-```{r timer}
+```{r timer, fig.cap="Computational perfoamce for PPIs, using sample sizes 100-10000.
+Colour indicates the PPI.
+Because the scagnostics calculation is bundled together, the values are the same for all these indexes, and they are really fast to compute.
+MINE includes the MIC and TIC indexes, and MINE E are computationally more efficient algorithms for these.
+These, along with dcor2D, are slower with larger sample sizes."}
 if(!file.exists("cache/timer.rda")){
@@
 \label{sec:params}
-Some parameters must be fixed when using the index functions considered here, controlling trade-offs between noise and fine structure detection, and between computing time and precision.
+Some parameters must be provided for some PPIs.
+This can be advantageous, allowing the index to more flexibly work for different types of structure, controlling trade-offs between noise and fine structure detection, and affecting computing time and precision.
 
@@
 
-### Effect of binning in scagnostics
-
-The implementation of scagnostics will always perform a binning step first, before evaluating the measures.
+<!--### Effect of binning in scagnostics -->
+
+The bins argument for the scagnostics might be reasonably expected to affect the smoothness of the index: a small number of bins should provide a smooth index function, but may affect its ability to detect fine structure.
+Figure \ref{fig:sprialScagBinning} examines this.
+Scagnostics PPIs are computed for the spiral1000 data on a tour path between V1-V2 to V5-V6, for number of bins equal to 10, 20, 50.
+The interesting observation to make is that even with small bin size the functions are all relatively noisy.
+The problem with the small bin size is that the spiral becomes invisible to the PPI.
+
+<!--The implementation of scagnostics will always perform a binning step first, before evaluating the measures.
 The binning will introduce random fluctuations on interpolated tour paths, but we may also consider using it to reduce fluctuations from the discreteness inherent to the index definitions, see in particular the definition of the stringy index.
@@
 We find indeed that different values may be preferred depending on the index considered, with the default bins=50 being the better choice for the convex and skinny index, while bins=20 may work better for the stringy index.
-Note that neither choice results in smooth curves for any of the indexes considered here.
+Note that neither choice results in smooth curves for any of the indexes considered here.-->
 
@@
   geom_line(aes(color=factor(nbin))) + 
-  facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE))
-
-```
-
-### Illustrative example
-A nice example can be constructed based on the RANDU dataset available in R.
-We use the MIC index and compare different parametrization of the maximum binning by $\alpha$, where $B(n) = n^{\alpha}$.
-We compare in Figure \ref{fig:randuEx} the final view, index value and computing time obtained when optimizing the MIC index with values $\alpha = 0.6, 0.7, 0.8$.
-Despite the low dimensionality, the small quint angle of the structure means that the identification of the maximum depends strongly on the starting conditions.
+  facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
+  scale_colour_discrete("Num bins") + xlab("Number of projections") + 
+  ylab("PPI value")
+
+```
+
+<!--### Illustrative example-->
+
+To examine the sensitivity of binning in the MIC PPI, the classic RANDU data [@Marsaglia68], available in R, is used.
+Binning is defined by $\alpha$, where $B(n) = n^{\alpha}$.
+Figure \ref{fig:randuEx} shows the best projection, index value and computing time obtained when optimizing the MIC index with values $\alpha = 0.6, 0.7, 0.8$.
+With small $\alpha$, less bins, the structure isn't visible, and with larger $\alpha$ the structure is confounded with noise.
+It does appear that this parameter affects the performace of the MIC index.
+
+<!--Despite the low dimensionality, the small squint angle of the structure means that the identification of the maximum depends strongly on the starting conditions.
 Empirically we have found that using $\alpha = 0.7$ is most efficient in identifying the structure, while the default value $\alpha = 0.6$ is most likely to miss the structure given equal starting conditions.
 
-\textcolor{red}{Possible interpretation: 0.6 not allowing enough bins to accurately capture the structure, 0.7 good compromise, 0.8 becomes overly sensitive to noise}
-
-```{r randuEx, results="hide", fig.cap="Final view when optimising the MIC index in the RANDU dataset.
-Each plot lists the selected value of $\\alpha$, the time required by the optimisation in seconds and the MIC index value."}
+\textcolor{red}{Possible interpretation: 0.6 not allowing enough bins to accurately capture the structure, 0.7 good compromise, 0.8 becomes overly sensitive to noise}-->
+
+```{r randuEx, results="hide", fig.cap="Best projection obtained by optimising the MIC index on the RANDU data, using different number of bins, defined by $\\alpha$.
+The smaller the value the fewer bins.
+Above each plot is written the value of $\\alpha$, time required to optimize (seconds) and the MIC index value.
+The best $\\alpha = 0.7$, and the result indicates that this parameter does affect MIC performance."}
 if(!file.exists("cache/randuEx.rda")){
@@
 
-## Possible modifications
-
-Big issues: smoothness, rotation invariance
-
-### Smoothing
-
-Different methods can be considered for smoothing the index traces over similar views for each basis:
-
-* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$
-* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction
+## Ways to refine the PPIs
+
+The biggest issues revealed by the investigation into the new PPIs are a lack of smoothness, particularly for the scagnostics indexes, and the rotation invariance of Grimm's indexes.
+To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighbourhood of projections and average the value, or alternatively average the PPI for several jittered projections.
+This is investigated in Fig.
+\ref{fig:spiralScagSmoothing}.
+Rotation invariance is more difficult to fix, but an alternative tour interpolation method could be useful.
+The geodesic interpolation transitions between planes, and it ignores the basis defining the plane, creating a problem with rotation invariant indices.
+Alternative interpolations based on Givens or Householder rotations could be implemented to transition between bases, which should alleviate the need for rotationally invariant indices.
+
+<!-- ### Smoothing-->
+
+Two different methods are considered for smoothing the index values:
+
+* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$.
+* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction.
+
+<!--
 * Turning axes: given the rotation non-invariance, maybe rotation of basis could be useful also??
-
-Moreover we can either select the mean value, or to avoid large pulls from sharp peaks, we can remove the most extreme values and then calculate the mean from the remaining evaluations.
-Only the mean will be considered here.
+-->
+
+The mean value from a sample of projections is recorded as the PPI value. This could be robsitufied by dropping the most extreme values.
 
 This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
-We study the potential of smoothing these two for the tour path between normal distributed noise parameters of the Spiral 1000 dataset, varying the $\epsilon$ and $\beta$ parameters.
-
-\textcolor{red}{I fixed the y axis to be between 0 and 1, which highlights that indeed we get much smoother behaviour, on the other hand theres still large relative jumps observed in all traces, and a smaller y axis scale may be preferred to show this more clearely.}
+Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values.
+Both methods appear to be promising in smoothing the function.
+Because the scagnistics are fast to compute, wither of these methods is feasible.
+
+<!--\textcolor{red}{I fixed the y axis to be between 0 and 1, which highlights that indeed we get much smoother behaviour, on the other hand theres still large relative jumps observed in all traces, and a smaller y axis scale may be preferred to show this more clearely.}-->
 
@@
 
-# References
-
 # Appendix

### 2018-12-12 scagnostics.Rmd
Marking V1-V5 plane in Fig 3.

@@
 "}
+intermediatePlane <- which(attributes(interpolate(t3))$new_basis)[2]
 fullResMeltPlanned <- fullResPlanned %>%
@@
   ylim(c(0,1)) +
+  geom_vline(xintercept=intermediatePlane, color="blue", alpha=0.3) +
   facet_grid(PPI~name) + 

### 2018-12-13 scagnostics.Rmd
Dataset dimension is p and not d.

@@
 
-The search is conducted for $d=4,5,6$ which would correspond to 2, 3 and 4 noise dimensions respectively.
+The search is conducted for $p=4,5,6$ which would correspond to 2, 3 and 4 noise dimensions respectively.
 In addition we examine the distance between planes, using a Frobenius norm, as defined by Equation 2 of @BCAH05, and available in the ```proj_dist``` function in the tourr package, to compare searches with squint angle across dimensions.
@@
 Figure \ref{fig:findspiral} summarizes the results.
-When $d=4$ the scouting method effectively finds the spiral.
-Plot (a) shows the side-by-side boxplots of distance between ideal plane and planes examined during the optimization, for $d=4,5,6$.
+When $p=4$ the scouting method effectively finds the spiral.
+Plot (a) shows the side-by-side boxplots of distance between ideal plane and planes examined during the optimization, for $p=4,5,6$.
 These are on average smaller for the lower dimension, and gradually increase as dimension increases.
 This is an indication of the extra computation needed to brute force find the spiral as noise dimensions increase.
-Plot (b) shows these as a function of the optimization procedure, where it can be seen that only when $d=4$ does it converge to the ideal.
+Plot (b) shows these as a function of the optimization procedure, where it can be seen that only when $p=4$ does it converge to the ideal.
 Its likely that expanding the search space should result in uncovering the spiral in higher dimensions.
@@
 
-```{r findspiral, out.width="\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with d = 4, 5, 6.
+```{r findspiral, out.width="\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with p = 4, 5, 6.
 The top left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the top right shows the evolution of distance to the plane containing the spiral as the index is being optimised.
-The bottom row shows the final view obtained for d = 4 (left), d = 5 (middle) and d = 6 (right).", results="hide"}
+The bottom row shows the final view obtained for p = 4 (left), p = 5 (middle) and p = 6 (right).", results="hide"}
 
@@
 
-p1 <- ggplot(distDf, aes(group=d, y=distance, x=d)) +
+distDf <- distDf %>%
+  mutate(p=d) %>%
+  select(-d)
+
+distToSp <- distToSp %>%
+  mutate(p=d) %>%
+  select(-d)
+
+p1 <- ggplot(distDf, aes(group=p, y=distance, x=p)) +
   geom_boxplot() +
   scale_x_continuous(breaks = c(4, 5, 6)) + ggtitle("(a)")
-p2 <- ggplot(distToSp, aes(color=as.factor(d),x =t, y=distance)) +
+p2 <- ggplot(distToSp, aes(color=as.factor(p),x =t, y=distance)) +
   geom_point() +
-  labs(y = "Distance to special view", color = "d") + ggtitle("(b)")
+  labs(y = "Distance to special view", color = "p") + ggtitle("(b)")
 

### 2018-12-13 scagnostics.Rmd
Added non-smoothed for comparison in smoothing plot, short statement about scaling of computing time.

@@
 Both methods appear to be promising in smoothing the function.
-Because the scagnistics are fast to compute, wither of these methods is feasible.
+Because the scagnistics are fast to compute, either of these methods is feasible.
+For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.
 
@@
 ```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
-While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered."}
+While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered.
+For comparison the red line in the background shows the trace without any smoothing applied."}
 
@@
 
+#getting non-smoothed result
+load("cache/V1V2toV3V4.rda")
+noSmoothing <- fullRes %>%
+  filter(size==1000) %>%
+  filter(name=="spiral") %>%
+  select(skinny, stringy, t) %>%
+  gather(variable, value, -t)
+
 scagSmoothMelt <- scagSmoothMelt %>%
@@
 ggplot(scagSmoothMelt, aes(x=t, y=value)) +
+  geom_line(data=noSmoothing, color="red", alpha=0.5) +
   geom_line(aes(linetype=factor(method))) + 

### 2018-12-13 scagnostics.Rmd
Fixed index sorting in trace plots.

@@
   gather(variable, value, -t, -name, -size) %>%
-  filter(t %in% which(attributes(interpolate(pipeTour))$new_basis))
+  filter(t %in% which(attributes(interpolate(pipeTour))$new_basis)) %>%
+  mutate(variable = fct_relevel(variable, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+
 ggplot(newBasisPipeTour, aes(x=t, y=value)) +
@@
 
-```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising  the splines2d index, using method 3, for the sine data, with $n=1000$.
+```{r findsine, fig.width=6, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Guided tour optimising  the splines2d index, using method 3, for the sine data, with $n=1000$.
 Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima.
@@
   select(-cmass) %>%
-  gather(variable, value, -t, -name, -size)
+  gather(variable, value, -t, -name, -size) %>%
+  mutate(variable = fct_relevel(variable, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
 ggplot(sineTourResMelt, aes(x=t, y=value)) +
@@
   scale_colour_manual(values = c("black", "red")) +
-  facet_wrap(~variable, ncol=1) +
+  facet_wrap(~variable, ncol=1, strip.position = "right") +
   geom_vline(data=anchorIdx, mapping=aes(xintercept=value), color="blue", alpha=0.3) +

### 2018-12-14 scagnostics.Rmd
Removed numbering in appendix.
Replace XXX by appropriate reference.
Added squint angle estimation.

@@
 If these two criteria are not met the method may still be useful, but only given an informed starting projection.
-In such cases we can follow a method similar to that proposed by Friedman (XXX REF): we break the optimization in two distinct steps.
+In such cases we can follow a method similar to that proposed by Friedman [@f87]: we break the optimization in two distinct steps.
 A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum.
@@
 
+```
+
+```{r squintAngle, out.width="0.5\\textwidth", fig.cap="Estimated squint angles for the Spiral dataset with 1000 datapoints, with p = 4, 5, 6, containing estimates evaluated for 100 randomly selected directions each.", results="hide"}
+
+squintAngleEstimat <- function(data, indexF, cutoff, structurePlane, n = 100, stepSize = 0.01){
+  # data = numerical input with p > 2 dimensions
+  # indexF = the index function used to estimate the squint angle
+  # cutoff = lower bound on the index value to be considered structured (i.e. all lower index values are outside the squint angle)
+  # structuredPlane = projection matrix onto plane containing the 2-d structure
+  # n = number of random starting directions over which the squint angle estimate is averaged
+  # stepSize = accuracy, step size determines where the index is evaluated to find the first plane above cutoff
+  data <- as.matrix(data) # make sure data is in matrix format
+  angles <- numeric(length = n) # collecting all individual estimates
+  i <- 1
+  p <- ncol(data)
+  while(i <= n){
+    # first generate random direction, make sure it is not too close to structure plane
+    dist <- 0.
+    while(dist < 0.1){
+      rBasis <- tourr::basis_random(p)
+      dist <- tourr::proj_dist(rBasis, structurePlane)
+    }
+    # now interpolate from rBasis to structure plane with selected step size
+    # since planned tour ignores first two entries, generate some random planes to be ignored
+    notUsed1 <- tourr::basis_random(p)
+    notUsed2 <- tourr::basis_random(p)
+    tourHist <- save_history(data,tour_path=planned_tour(list(notUsed1, notUsed2, rBasis, structurePlane)))
+    allBases <- as.list(interpolate(tourHist, angle = stepSize))
+    cIndex <- 0.
+    j <- 1
+    while(cIndex < cutoff){
+      cProj <- data %*% allBases[[j]]
+      cIndex <- indexF(cProj)
+      j <- j+1
+    }
+    cDist <- tourr::proj_dist(allBases[[j]], structurePlane)
+    angles[i] <- cDist
+    i <- i+1
+  }
+  return(angles)
+}
+
+if(!file.exists("cache/squintAngle.rda")){
+  allAngles <- tibble(d=numeric(), angle=numeric())
+  for(d in c(4,5,6)){
+    set.seed(58958)
+    specialPlane <- matrix(c(rep(0,d-2),1,0,rep(0,d-1),1),ncol=2)
+    spiralD <- spiral1000 %>%
+      purrr::when(d<6 ~ select(., -V4), ~ .) %>%
+      purrr::when(d<5 ~ select(., -V3), ~ .)
+    cAngles <- squintAngleEstimat(spiralD, scagIndex("Skinny"), 0.6, specialPlane)
+    allAngles <- bind_rows(allAngles, tibble(d=d, angle=cAngles))
+  }
+  save(allAngles, file = "cache/squintAngle.rda")
+} else {
+  load("cache/squintAngle.rda")
+}
+
+ggplot(allAngles, aes(d, angle, group=d)) + geom_boxplot()
 ```
@@
 
-# Appendix
-
-XXX Need to make this not numbered
-
-## Re-scaling of holes index
+# Appendix {-}
+
+## Re-scaling of holes index {-}
 

### 2018-12-14 scagnostics.Rmd
Use xi not Vi for parameter directions.

@@
 Figure \ref{fig:wIntermediatePath} shows the PPIs for a tour sequence between a nuisance and structured projection.
-A long sequence is generated where the path runs from V1-V2, V1-V5, V5-V6, in order to see some of the intricacies of holes index.
+A long sequence is generated where the path runs from $x_1$-$x_2$, $x_1$-$x_5$, $x_5$-$x_6$, in order to see some of the intricacies of holes index.
 Sample size is indicated by line type: dashed being $n=100$ and solid is $n=1000$.
@@
 
-```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between nuisance and structured projections, following V1-V2 to V1-V5 to V5-V6.
-Solid lines are the larger sample size, $n=1000$, and dashed are smaller sample size, $n=100$.
+```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between nuisance and structured projections, following $x_1$-$x_2$ to $x_1$-$x_5$ to $x_5$-$x_6$ Solid lines are the larger sample size, $n=1000$, and dashed are smaller sample size, $n=100$.
 Peaks at the end of the sequence indicate the index sees the structure.
 The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection.
-Holes only responds to the pipe, and is a multimodal function for this data with a local maximum at V1-V5.
-The scagnostics tend to be sensitive to sample size.
+Holes only responds to the pipe, and is a multimodal function for this data with a local maximum at $x_1$-$x_5$ The scagnostics tend to be sensitive to sample size.
 "}
@@
 dProj <- as.tibble(pipeResc %*% fProj)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
+  xlab("PP1") + ylab("PP2")
 iLast <- length(pipeTourFull2)
@@
 dProj <- as.tibble(pipeResc %*% fProj)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
+  xlab("PP1") + ylab("PP2")
 grid.arrange(p1, p2, ncol=2)
@@
 
+fViewL[[1]] <- fViewL[[1]] + xlab("PP1") + ylab("PP2")
+fViewL[[2]] <- fViewL[[2]] + xlab("PP1") + ylab("PP2")
+fViewL[[3]] <- fViewL[[3]] + xlab("PP1") + ylab("PP2")
+
 grid.arrange(p1, p2, fViewL[[1]], fViewL[[2]], fViewL[[3]], layout_matrix = rbind(c(1,2,2),c(3,4,5)))
@@
 
-ggplot(allAngles, aes(d, angle, group=d)) + geom_boxplot()
+allAngles <- rename(allAngles, p=d)
+
+ggplot(allAngles, aes(p, angle, group=p)) + geom_boxplot()
 ```
@@
 
-Rotational invariance is examined using the sine data (V5-V6), computing PPI for different rotations within the 2D plane, parametrized by angle.
+Rotational invariance is examined using the sine data ($x_5$-$x_6$), computing PPI for different rotations within the 2D plane, parametrized by angle.
 Results are shown in Figure \ref{fig:rotationDep}.
@@
 Figure \ref{fig:sprialScagBinning} examines this.
-Scagnostics PPIs are computed for the spiral1000 data on a tour path between V1-V2 to V5-V6, for number of bins equal to 10, 20, 50.
+Scagnostics PPIs are computed for the spiral1000 data on a tour path between $x_1$-$x_2$ to $x_5$-$x_6$, for number of bins equal to 10, 20, 50.
 The interesting observation to make is that even with small bin size the functions are all relatively noisy.

### 2018-12-17 scagnostics.Rmd
Some notation fixes.

@@
 The three data families are explained below.
-In each set, $p$ is used for the number of points, $n$ is the number of dimensions, and $d=2$ is the projection dimension.
+In each set, $n$ is used for the number of points, $p$ is the number of dimensions, and $d=2$ is the projection dimension.
 
@@
 The circle should be easy to see by some indices because it is large structure, but the nonlinearity creates a complication.
-- **sine**:  nuisance directions are generated by sampling independently from a standard normal distribution, and the sine curve is generated by $x_n = \sin(x_{n-1}) + \mathrm{jittering}$.
+- **sine**:  nuisance directions are generated by sampling independently from a standard normal distribution, and the sine curve is generated by $x_p = \sin(x_{p-1}) + \mathrm{jittering}$.
 The sine is medium nonlinear structure, which should be visible by multiple indices.
@@
 
-Figure \ref{fig:wIntermediatePath} shows the PPIs for a tour sequence between a nuisance and structured projection.
+Figure \ref{fig:wIntermediate} shows the PPIs for a tour sequence between a nuisance and structured projection.
 A long sequence is generated where the path runs from $x_1$-$x_2$, $x_1$-$x_5$, $x_5$-$x_6$, in order to see some of the intricacies of holes index.
@@
 To examine the sensitivity of binning in the MIC PPI, the classic RANDU data [@Marsaglia68], available in R, is used.
-Binning is defined by $\alpha$, where $B(n) = n^{\alpha}$.
-Figure \ref{fig:randuEx} shows the best projection, index value and computing time obtained when optimizing the MIC index with values $\alpha = 0.6, 0.7, 0.8$.
-With small $\alpha$, less bins, the structure isn't visible, and with larger $\alpha$ the structure is confounded with noise.
+Binning is defined by $\delta$, where $B(n) = n^{\delta}$.
+Figure \ref{fig:randuEx} shows the best projection, index value and computing time obtained when optimizing the MIC index with values $\delta = 0.6, 0.7, 0.8$.
+With small $\delta$, less bins, the structure isn't visible, and with larger $\delta$ the structure is confounded with noise.
 It does appear that this parameter affects the performace of the MIC index.
@@
 
-```{r randuEx, results="hide", fig.cap="Best projection obtained by optimising the MIC index on the RANDU data, using different number of bins, defined by $\\alpha$.
+```{r randuEx, results="hide", fig.cap="Best projection obtained by optimising the MIC index on the RANDU data, using different number of bins, defined by $\\delta$.
 The smaller the value the fewer bins.
-Above each plot is written the value of $\\alpha$, time required to optimize (seconds) and the MIC index value.
-The best $\\alpha = 0.7$, and the result indicates that this parameter does affect MIC performance."}
+Above each plot is written the value of $\\delta$, time required to optimize (seconds) and the MIC index value.
+The best $\\delta = 0.7$, and the result indicates that this parameter does affect MIC performance."}
 if(!file.exists("cache/randuEx.rda")){

### 2019-01-04 scagnostics.Rmd
Smaller table 1.

@@
   #row_spec(1:16, font_size = 10) %>%
-  kable_styling(position = "center") 
+  kable_styling(position = "center", font_size = 7) 
 ```

### 2019-01-04 scagnostics.Rmd
Moving Figures to appendix, remove some commented old text.

@@
 
-<!--
-We first study the smoothness of each measure calculated on a sequence of 2-d projections obtained via an interpolated planned tour path moving between the projection onto $x_1$-$x_2$ to the projection onto $x_3$-$x_4$ and using the default interpolation angle of 0.05.
-Each time step refers to an interpolation step.
-Since the value of the TIC index is strongly dependent on the input distribution sample size we have normalised the TIC measure by its maximum for each individual dataset, all other measures are reproduced as reported by the calculation.
-The results are shown in Figure \ref{fig:plotV1V2toV3V4}.
-The standard index functions "holes" and "cmass" show the desired behaviour in that the change is smooth over interpolated views.
-Considering now the indices from the scagnostics family, we first note stronge dependence on the sample size.
-The convex index is largely sensitive to random fluctuations when considering small samples ($n=100$), while being mostly flat for larger samples, especially in the case of an underlying uniform distribution.
-The skinny index on the other hand is systematically larger for the smaller samples, while stringy appears not to be sensitive to the sample size, but resulting in large fluctuations and sharp jumps for all datasets.
-Distance correlation and the spline based index show similar behaviour, i.e.
-when considering large enough data samples ($n=1000$) the values are consistently close to zero, as expected for a distribution without structure.
-On the other hand somewhat larger values (but still far from the maximum value of 1) are found when considering small datasets, where distributions are however mostly smooth, with few kinks depending on index function and underlying distribution.
-Finally the information based metrics also show systematic differences based on the sample size, with sampling fluctuations being significant for small datasets.
-Note here that fluctuations present in Figure \ref{fig:plotV1V2toV3V4} may be amplified, as maximum values for most index functions are far from the theoretic maximum.
--->
-
 <!-- index values shouldn't be affected by sample size, on average it should be
@@
 
-<!--
-```{r plannedtourpath}
-if(!file.exists("cache/plannedtourpath.rda")){
-  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
-  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
-  #this is silly but seems that first two entries are being ignored, so need some fake entries
-  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
-  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
-  t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
-  t2full <- as.list(interpolate(t2))
-  fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
-    rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
-    rbind(getProj(spiral100, t2full, "spiral", 100)) %>%
-    rbind(getProj(spiral1000, t2full, "spiral", 1000)) %>%
-    rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, t2full, "pipe", 1000))
-  save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
-} else {
-  load("cache/plannedtourpath.rda")
-}
-```
-
-```{r plannedtour, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between a nuisance and a structured projection (solid lines: $n=1000$; dashed: $n=100$).
-Most of the indexes have peaks at the end of the sequence, indicating that they see the structure.
-The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection."}
-fullResMeltPlanned <- fullResPlanned %>%
-  mutate(convex = 1-convex) %>% 
-  rowwise() %>%
-  mutate(holes = rescaleHoles(holes)) %>%
-  gather(PPI, value, -t, -name, -size) %>%
-  filter(PPI != "cmass") %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
-fullResMeltPlanned %>% 
-  ggplot(aes(x=t, y=value)) + 
-  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
-  ylim(c(0,1)) +
-  facet_grid(PPI~name) + 
-  theme(legend.position="none") + xlab("Sequence of projections (t)") + ylab("PPI value")
-```
-
-## Tentative
-try how this looks going first to V1 vs V5, then to V5 vs V6
--->
-
 ```{r wIntermediatePath}
@@
 For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimist the TIC index.
-The results are shown in Figures \ref{fig:pipeFirstRun} and \ref{fig:testpipe}.
+The results are shown in Figures \ref{fig:pipeFirstRun}.
 
@@
 
-Figure \ref{fig:testpipe} shows the projection returned during the scouting phase (left) and the refinement phase (right).
-It was important to start the method 3 optimizer at the best projection returned during the scouting phase, to smoothly converge more closely to the ideal projection.
+The final views obtained in each of the two stages are compared in the Appendix, see Fig. \ref{fig:testpipe}.
 
@@
 }
-```
-
-```{r testpipe,  fig.height=4, fig.width=8, out.width = "0.8\\textwidth",  fig.cap="Projections returned by TIC optimisation: by the scouting phase (left) and refined by  optimisation method 3 (right), starting from the scouting phase projection."}
-iLast <- length(pipeTourFull)
-fProj <- pipeTourFull[[iLast]]
-dProj <- as.tibble(pipeResc %*% fProj)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
-  xlab("PP1") + ylab("PP2")
-iLast <- length(pipeTourFull2)
-fProj <- pipeTourFull2[[iLast]]
-dProj <- as.tibble(pipeResc %*% fProj)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
-  xlab("PP1") + ylab("PP2")
-grid.arrange(p1, p2, ncol=2)
 ```
@@
 
-Figure \ref{fig:neutronStarSPLOM} shows the scatterplot matrix (with transparency) of the six parameters.
+Figure \ref{fig:neutronStarSPLOM} in the Appendix shows the scatterplot matrix (with transparency) of the six parameters.
 (In astrophysics, scatterplot matrices are often called "corner plots" [@corner].) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables.
@@
 
-<!--
-(left) clearly recovers functional relations between some of the parameters.
-From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection.
-Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{fig:nsePlotOrig} (middle), which we compare to the dependence of L1 on the ratio R1/m1 (right).
-For this example the linear combination can thus recover the approximate relation between these parameters.
--->
-
-<!--
- A posterior sample containing 2538 points is available @ligoData and can be used to constrain the parameters.
-Here we use the "Parametrized-EoS\_nomaxmass\_posterior\_samples.dat" dataset and aim to recover the dependence between the three parameters of the better constrained system out of a larger parameter set.
-For details about the modelling and statistical procedure see @Abbott2018exr.
-A common representation of posterior samples used for interpretation are so-called "corner plots", i.e.
-density displays for all 2-d parameter combinations as well as the 1-d profile, highlighting empirically evaluated regions of highest posterior density, which correspond to credibility regions at a given level.
-The information can similarly be presented in a scatter plot matrix using transparency to convey the density information.
-Since in the following we work with 2-d scatter plots we choose to show the data in this fashion, see Fig.
-\ref{fig:neutronStarSPLOM}.
--->
-
-```{r neutronStarSPLOM, fig.height=8, fig.width=8, fig.cap="Scatter plot matrix of the neutron star dataset, darker regions represent higher marginalised posterior densities."}
+```{r neutronStarEq,  results="hide"}
 nsD <- read_csv("data/samples.csv") 
-ggpairs(nsD, lower=list(continuous = wrap("points", alpha = 0.05)))
-```
-
-<!-- The following features can be observed: the combination of the two mass parameters is well constrained (a consequence of how a particular combination, the "chirp mass" enters predictions), other parameters also show patterns of dependence, they are however less pronounces.
-We also see the non-linear relation between the radius $R_1$ and $\Lambda$ parameter $L1$ and the two-pronged disribution for the radius $R_1$.
-For further study we therefore remove the second mass $m2$ from the dataset.
-\textcolor{red}{Looking at the distribution in a grand tour display, using a scatter plot display combined with alpha transparency to study the density distribution in the 5-d parameter space, indicates that the data points are found on a curved surface in the five dimensional space.
-This stems from the non-linear relation between the parameters, which we will try to uncover using the above developed formalism for new guided tour index functions.}
-
-### Results
-
-Since we expect a functional relation between some of the parameters, we optimise the "splines2d" index function.
-Note that all parameters are standardised to the range $[0,1]$ before running the optimisation, and we do not sphere the data for ease of interpretation of the results.
--->
-
-```{r neutronStarEq,  results="hide"}
 nsM <- nsD %>%
@@
 
-Figure \ref{fig:bbhSimulation} shows a scatterplot matrix, of nine of the twelve parameters.
+Figure \ref{fig:bbhSimulation} in the Appendix shows a scatterplot matrix, of nine of the twelve parameters.
 (Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time.
@@
 These patterns are representative of the likelihood function, since most priors are flat, or built to capture growth with volume rather than distance.
-
-```{r bbhSimulation, fig.height=8, fig.width=8, dev = "png", dpi=300, fig.cap="Scatter plot matrix showing most of the variables included in the BBH dataset.
-Strong correlation between the parameters time, dec and ra can be observed."}
-bbhD <- read_csv("data/posterior_samples.csv")
-bbhDsmall <- select(bbhD, -phi_jl, -m2, -psi, -chi_eff)
-ggpairs(bbhDsmall, lower=list(continuous = wrap("points", alpha = 0.02)))
-```
 
@@
 The 1-convex index finds a very different view.", dev = "png", dpi=300}
+bbhD <- read_csv("data/posterior_samples.csv")
 bbhM <- rescale(select(bbhD,-m2, -chi_eff))
@@
 
-## Re-scaling of holes index {-}
+## Re-scaling of holes index
 
@@
 any value below 0.2 (0.8) is set to zero for cmass (holes) index, rescale remaining range to be between zero and one.
+
+## Additional Figures
+
+### Final projection of pipe guided tour
+Figure \ref{fig:testpipe} shows the projection returned during the scouting phase (left) and the refinement phase (right).
+It was important to start the method 3 optimizer at the best projection returned during the scouting phase, to smoothly converge more closely to the ideal projection.
+
+```{r testpipe,  fig.height=4, fig.width=8, out.width = "0.8\\textwidth",  fig.cap="Projections returned by TIC optimisation: by the scouting phase (left) and refined by  optimisation method 3 (right), starting from the scouting phase projection."}
+iLast <- length(pipeTourFull)
+fProj <- pipeTourFull[[iLast]]
+dProj <- as.tibble(pipeResc %*% fProj)
+p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
+  xlab("PP1") + ylab("PP2")
+iLast <- length(pipeTourFull2)
+fProj <- pipeTourFull2[[iLast]]
+dProj <- as.tibble(pipeResc %*% fProj)
+p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
+  xlab("PP1") + ylab("PP2")
+grid.arrange(p1, p2, ncol=2)
+```
+
+### Dataset overview scatter plot matrices
+
+```{r neutronStarSPLOM, fig.height=8, fig.width=8, fig.cap="Scatter plot matrix of the neutron star dataset, darker regions represent higher marginalised posterior densities."}
+ggpairs(nsD, lower=list(continuous = wrap("points", alpha = 0.05)))
+```
+
+```{r bbhSimulation, fig.height=8, fig.width=8, dev = "png", dpi=300, fig.cap="Scatter plot matrix showing most of the variables included in the BBH dataset.
+Strong correlation between the parameters time, dec and ra can be observed."}
+bbhDsmall <- select(bbhD, -phi_jl, -m2, -psi, -chi_eff)
+ggpairs(bbhDsmall, lower=list(continuous = wrap("points", alpha = 0.02)))
+```

### 2019-01-04 scagnostics.Rmd
Moved randu example to appendix.

@@
 
-<!--The implementation of scagnostics will always perform a binning step first, before evaluating the measures.
-The binning will introduce random fluctuations on interpolated tour paths, but we may also consider using it to reduce fluctuations from the discreteness inherent to the index definitions, see in particular the definition of the stringy index.
-This needs some care however as different values may be preferred for different structures and depending on which index is being considered.
-Moreover, choosing a binning that is too rough may result in loss of discrimination power as structures can no longer be resolved.
-As an example we compare in Figure \ref{fig:sprialScagBinning} the scagnostics index traces for the Spiral 1000 dataset when moving from $x_1$-$x_2$ to $x_5$-$x_6$, for three values of the bins parameter entering the hexagonal binning procedure.
-We find indeed that different values may be preferred depending on the index considered, with the default bins=50 being the better choice for the convex and skinny index, while bins=20 may work better for the stringy index.
-Note that neither choice results in smooth curves for any of the indexes considered here.-->
-
 ```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.
@@
 
-<!--### Illustrative example-->
-
-To examine the sensitivity of binning in the MIC PPI, the classic RANDU data [@Marsaglia68], available in R, is used.
-Binning is defined by $\delta$, where $B(n) = n^{\delta}$.
-Figure \ref{fig:randuEx} shows the best projection, index value and computing time obtained when optimizing the MIC index with values $\delta = 0.6, 0.7, 0.8$.
-With small $\delta$, less bins, the structure isn't visible, and with larger $\delta$ the structure is confounded with noise.
-It does appear that this parameter affects the performace of the MIC index.
-
-<!--Despite the low dimensionality, the small squint angle of the structure means that the identification of the maximum depends strongly on the starting conditions.
-Empirically we have found that using $\alpha = 0.7$ is most efficient in identifying the structure, while the default value $\alpha = 0.6$ is most likely to miss the structure given equal starting conditions.
-
-\textcolor{red}{Possible interpretation: 0.6 not allowing enough bins to accurately capture the structure, 0.7 good compromise, 0.8 becomes overly sensitive to noise}-->
-
-```{r randuEx, results="hide", fig.cap="Best projection obtained by optimising the MIC index on the RANDU data, using different number of bins, defined by $\\delta$.
-The smaller the value the fewer bins.
-Above each plot is written the value of $\\delta$, time required to optimize (seconds) and the MIC index value.
-The best $\\delta = 0.7$, and the result indicates that this parameter does affect MIC performance."}
-if(!file.exists("cache/randuEx.rda")){
-  pL <- list()
-  i <- 1
-  for (alpha in c( 0.6, 0.7, 0.8)){
-    set.seed(556677)
-    tic.clearlog()
-    tic() #start timer
-    tP <- save_history(randu, guided_tour(mineIndexAlpha("MIC", alpha)),max_bases = 100000)
-    toc(log=TRUE,quiet=TRUE)
-    optT <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
-    imax <- length(as.list(tP))
-    fV <- as.tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
-    colnames(fV) <- c("PP1", "PP2")
-    idxV <- mine(fV$PP1, fV$PP2, alpha = alpha)[["MIC"]]
-    pL[[i]] <- ggplot(fV, aes(PP1,PP2)) + geom_point() + theme(aspect.ratio=1) + 
-      ggtitle(paste(alpha, toString(format(optT, digits=2)), toString(format(idxV, digits=2)), sep = ", "))
-    i <- i+1
-  }
-  save(pL, file = "cache/randuEx.rda")
-} else {
-  load("cache/randuEx.rda")
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], ncol=3)
-```
+The sensitivity to the binning for calculating the MIC index is examined for one example in the Appendix XXX (add cross reference)
 
@@
 ```
-
-<!--
-### Rotation invariance
-Do we want to say anything here? big issue for splines2D measure, but also scagnostics
--->
 
@@
 
+## Binning sensitivity of MIC index
+To examine the sensitivity of binning in the MIC PPI, the classic RANDU data [@Marsaglia68], available in R, is used.
+Binning is defined by $\delta$, where $B(n) = n^{\delta}$.
+Figure \ref{fig:randuEx} shows the best projection, index value and computing time obtained when optimizing the MIC index with values $\delta = 0.6, 0.7, 0.8$.
+With small $\delta$, less bins, the structure isn't visible, and with larger $\delta$ the structure is confounded with noise.
+It does appear that this parameter affects the performace of the MIC index.
+
+```{r randuEx, results="hide", fig.cap="Best projection obtained by optimising the MIC index on the RANDU data, using different number of bins, defined by $\\delta$.
+The smaller the value the fewer bins.
+Above each plot is written the value of $\\delta$, time required to optimize (seconds) and the MIC index value.
+The best $\\delta = 0.7$, and the result indicates that this parameter does affect MIC performance."}
+if(!file.exists("cache/randuEx.rda")){
+  pL <- list()
+  i <- 1
+  for (alpha in c( 0.6, 0.7, 0.8)){
+    set.seed(556677)
+    tic.clearlog()
+    tic() #start timer
+    tP <- save_history(randu, guided_tour(mineIndexAlpha("MIC", alpha)),max_bases = 100000)
+    toc(log=TRUE,quiet=TRUE)
+    optT <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
+    imax <- length(as.list(tP))
+    fV <- as.tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
+    colnames(fV) <- c("PP1", "PP2")
+    idxV <- mine(fV$PP1, fV$PP2, alpha = alpha)[["MIC"]]
+    pL[[i]] <- ggplot(fV, aes(PP1,PP2)) + geom_point() + theme(aspect.ratio=1) + 
+      ggtitle(paste(alpha, toString(format(optT, digits=2)), toString(format(idxV, digits=2)), sep = ", "))
+    i <- i+1
+  }
+  save(pL, file = "cache/randuEx.rda")
+} else {
+  load("cache/randuEx.rda")
+}
+grid.arrange(pL[[1]], pL[[2]], pL[[3]], ncol=3)
+```
+
 ## Additional Figures

### 2019-01-04 scagnostics.Rmd
Smaller figures.

@@
 
-<!--\textcolor{red}{Since projection is fast compared to index evaluation, the number of original dimensions is not very important here.
-On the other hand different distributions may impact the computing time.}
-We see that the evaluation of the spline index shows weak dependence on the sample size, making it the least efficient for very small data sets, but one of the fastest when considering large data sets.
-The situation is similar for calculating the scagnostics measures, which are however faster to compute on datasets of around 10 points.
-This is because the implementation is first binning the data points, which is fast compared to the evaluation that requires triangulation of the binned points.
-The distance correlation based measure exhibits strong increase with sample size, we note however that the implementation also allows for evaluation on binned data points which we have not considered.
-Finally for the MINE family the computing time increases exponentially as can be expected from the dependence on the sample size, where we see that the improved $MIC_e$ algorithm is somewhat faster than the original implementation, see also @JMLRv1715308.-->
-
 ```{r getTimer}
@@
 MINE includes the MIC and TIC indexes, and MINE E are computationally more efficient algorithms for these.
-These, along with dcor2D, are slower with larger sample sizes."}
+These, along with dcor2D, are slower with larger sample sizes.", out.width=".6\\textwidth"}
 if(!file.exists("cache/timer.rda")){
@@
 
-```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.
-For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
+XXX from caption, remove?:  For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
 With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
-For the convex index we also find bins=50 to be the better choice."}
+For the convex index we also find bins=50 to be the better choice.
+
+```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.", out.width=".7\\textwidth"}
 getScagComp <- function(df, tPath){

### 2019-01-04 scagnostics.Rmd
First version of discussion.

@@
 
+Physics examples motivate us to investigate projection pursuit indexes that detect complex patterns in two dimensional projections.
+In this context we have investigated the use of indexes developed for variable selection as projection pursuit indexes.
+Having outlined a set of criteria that define a good index function, we have established methods for the evaluation of the index properties: smoothness, squintability, flexibility, rotation invariance and speed.
+To this end we used three different sets of simulated data, defined to have different underlying noise distributions (uniform or normal) and different structures (circle, sine and spiral), and we have evaluated several index functions proposed in the literature for variable selection (scagnostics indexes, maximum information based indexes, distance correlation and a spline based index).
+While all index functions are found to fall short on at least some of the criteria, we have shown that combined with the optimisation available via the guided tour they can already be used to identify interesting views, in particular the built-in structures in the simulated data sets.
+
+We have also sketched some potential improvements of the indexes considered, in particular tuning parameters inherent to the index definition, such as binning, or different smoothing methods for the scagnostics indexes.
+
+Our primary interest is in multidimensional physics models, and we have shown two examples from gravitational wave detection.
+In the first example, studying a neutron star merger, we have sucessfully used the developped methods to re-discover the non-linear relation between three of the parameters.
+The second example, stemming from a black hole simulation, has a large parameter space.
+We used this dataset to explore the results obtained depending on parameter selection, index selection and starting conditions.
+Interestingly, when working with a small parameter space (six dimensions), projection pursuit uncovers a new relation that has not been anticipated.
+
+We conclude that these methods are promissing and may yield new insight in the inter-dependence of model parameters in the light of experimental data ... other applications can also be envisioned??
+
+There are three frontiers for future work building on our study.
+First, while we found the considered index functions useful to detect structures, several shortcomings have also been pointed out.
+While many definitions to identify clusters have been considered, there is much potential for the development of index functions that detect the types of structures considered here.
+Second, an implementation of interpolations from basis to basis, alleviating the need for rotation invariant indexes is envisioned.
+Finally, more generally, improved optimisation algorithms are required, e.g.
+for the identification of structures with small squint angle in high dimensions.
+
 # Appendix {-}

### 2019-01-07 scagnostics.Rmd
Some shuffling in the introduction, removing duplicated bibtex entry.

@@
 The performance of these indices is examined with respect to ideal behaviour, using simulated data, and then applied to several physics problems.
-The implementation builds upon the projection pursuit tools available in the R package, tourr, with indices constructed from code in the R package, scagnostics.
+The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphics.
 
@@
 
-A large number of projection pursuit indices have been developed, to detect departure from multivariate normal, which includes clusters or outliers or separations between known groups (e.g.
+A large number of projection pursuit indexes have been developed, to detect departure from multivariate normal, which includes clusters or outliers or separations between known groups (e.g.
 @f87, @CBC92, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437).
 Less work has been done on indexes to find nonlinear dependence between variables.
-
-XXX Need to add something here about the guided tour
+\footnote{\textcolor{red}{The second component in implementing a projection pursuit method is the optimisation algorithm.
+Several algorithms have been suggested, some tailored to specific applications, see e.g.
+XXX references?.
+In this work our focus is on the index functions, and we will only use optimisation algorithms available via the guided tour implementation of tourr.}}
 
 <!-- check Perisic and Posse (2012) and recent paper by Austrian guys -->
-
-A related topic is variable selection.
-With high-dimensional data, even plotting all pairs can be daunting, so "scagnostics" [@scag] [@WW08] have been developed to find the most interesting pairs of variables.
-There are 8 scagnostics, vividly named: "Outlying", "Skewed", "Sparse", "Clumpy", "Striated", "Convex", "Skinny" and "Stringy".
-Our question is whether these can be extended into projection pursuit indexes that can be used to find two-dimensional projections of high-dimensional data with unusual features.
 
@@
 
-<!--
-Suggestion German:
-We are motivated by physics applications, specifically to assist the interpretation of models fit to experimental results.
-In these problems one has a physical model with a set of m free parameters which cannot be measured directly, but that are determined instead by fitting a set of p experimental observations.
-The model predicts all the observations in terms of the m parameters, often analytically, but occasionally only numerically.
-Moreover, a single prediction can be a complicated function of all the free parameters requiring intensive numerical resources.
-
-These problems often deal with a large number of observations and parameters, with p in the hundreds or even thousands and m of order 10. 
-The results of a fit have been generally interpreted in terms of combinations of parameters selected by intuition or prior knowledge.
-Recently, these selections have been also assisted by machine learning techniques including BDT and neural networks.
-
-Our visualisation tools complement these procedures providing a high dimensional picture which assists with the interpretation of the results,  serves as a check for other methods and possibly revealing new associations.
--->
+We draw on developments from the related topic of variable selection.
+With high-dimensional data, even plotting all pairs can be daunting, so "scagnostics" [@scag] [@WW08] have been developed to find the most interesting pairs of variables.
+There are 8 scagnostics, vividly named: "Outlying", "Skewed", "Sparse", "Clumpy", "Striated", "Convex", "Skinny" and "Stringy".
+Our question is whether these can be extended into projection pursuit indexes that can be used to find two-dimensional projections of high-dimensional data with unusual features.
 
@@
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], and computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308], round out the collection of metrics.
-\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber85,jones87}.}
+\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber85,JS87}.}
 These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
@@
 
-\textcolor{red}{\cite{jones87} discusses centering and sphering of the data points.
+\textcolor{red}{\cite{JS87} discusses centering and sphering of the data points.
 Centering is needed if index is not translational invariant (should not be the case for any index considered here).
@@
 A useful projection pursuit index needs to have several properties.
-This has been discussed in several seminal papers (\cite{diaconis84}, \cite{huber85}, \cite{jones87}, \cite{posse95b}, \cite{hall89}).
+This has been discussed in several seminal papers (\cite{diaconis84}, \cite{huber85}, \cite{JS87}, \cite{posse95b}, \cite{hall89}).
 The PPI should be  minimized by the normal distribution, because this is not interesting from a data exploration perspective.

### 2019-01-08 scagnostics.Rmd
Small cleanup of Sec 2, but requires going through together still.

@@
 
-<!--
-Explain here how to construct the pp index
-
-- Definition of pp index
-- Optimisation
-- Illustration of what is already available in tourr
-- New index explanations
--->
 A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data distribution.
@@
 A possible algorithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
-In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimists the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
+In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
 
@@
 
-* From the scagnostics family select those best suited to detect shapes, with definitions given below.
-They are defined to take values between [0,1].
-\textcolor{red}{The definition of these should be rotation invariant, however, some small dependence is introduced via the hex binning that is done before calculating the scagnostics indexes.}
+* From the scagnostics family select those best suited to detect shapes, with definitions given below. They are defined to take values between [0,1].
     + Convex: The convex measure is defined as $c_{convex}= \frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
@@
 
-* New index functions available in @mbgraphic, defined to fall between [0,1] \textcolor{red}{and not rotation invariant, as I think Var(X)*Var(Y) introduces rotation invariance even for distance correlation?}:
+* New index functions available in @mbgraphic, defined to fall between [0,1]:
     + Distance correlation defined in @szekely2007: The definition is based on empiric measures of distance covariance and variance defined on the dataset 

### 2019-01-08 scagnostics.Rmd
Starting work on Sec 3.

@@
 
-# Investigation of indices
+# Investigation of indexes
 \label{sec:investigate}
@@
 The interpolation path corresponds to small angle changes between projections, so the value should be very similar.
-<!-- 
-  should evolve smoothly between nearby projections, i.e.
-on an interpolated path we expect to observe slow and consistent change in the index function.
-\textcolor{red}{maybe can relate this to consistency consition developed by P.
-Hall, but don't understand this at the moment.
-In addition this relates to questions of optimiser choice and efficiency.} We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
--->
+
 - **squintability**: \cite{barnett1981interpreting} introduced the idea of squint angle to indicate resolution of structure in the data.
@@
 To examine squintability, the PPI values are examined on interpolated tour paths between a noise projection and a distant structured projection.
-<!--
-distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions.
-\textcolor{red}{I dont really understand this part.
-I think it can instead be related to the concept of squint angle \cite{barnett1981interpreting}.} We show examples in Section \ref{sec:dist}
--->
+
 - **flexibility**: An analyst can have a toolbox of indices that may cover the range of fine and broad structure, which underlies the scagnostics suite.
@@
 This is examined by using a range of structure in the simulated data examples.
-<!--
-ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}.
-On the other hand should also consider detection of "fine" structure vs noise, this would be controlled e.g.
-by the allowed number of bins for the MINE index functions, additionally this will impact computing times, i.e.
-trade accuracy for efficiency}.
-We address these questions in Section \ref{sec:params} with an illustrative example.-->
+
 - **rotation invariance**: The orientation of structure within a projection plane should not change the index value.
@@
 Rotation invariance is checked for the proposed PPIs by rotating the structured projection, within the plane.
-<!--
-in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity.
-A projection pursuit index should therefore be invariant under rotation of the basis in the projection plane.
-By construction this is not the case for most index functions defined above.
-We illustrate the impact of non invariance in Section \ref{sec:rot}.
-\textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating, reamining issue would be optimisation efficiency.}
--->
+
 - **speed**: Being fast to compute allows the index to be used in real-time in a guided tour, where the optimization can be watched.
 When the computations are shifted off-line, to watch in replay, computation times matter less.
-This is checked  by comparing times for benchmark scenarios with varying sample size and dimension.
+This is checked  by comparing times for benchmark scenarios with varying sample size.
 
@@
 - **sine**:  nuisance directions are generated by sampling independently from a standard normal distribution, and the sine curve is generated by $x_p = \sin(x_{p-1}) + \mathrm{jittering}$.
-The sine is medium nonlinear structure, which should be visible by multiple indices.
+The sine is a medium nonlinear structure, which should be visible to multiple indices.
 - **spiral**:  nuisance directions are generated by sampling independently from a normal distribution, and the structure directions are sampled from an Archimedean spiral, i.e.
@@
 
-<!--
-To get an overview of the dataset we show views of the first two directions showing the independent distribution and the last two directions showing the special view.
-The enforced relationships will also affect the 1-d densities as can be seen from \ref{fig:dataPlotsDensity}, which means that views including sizable fractions of $x_6$ (and $x_5$ for the Pipe and Spiral distribution) will be distinct from uninformed views of distributions 2 and 3.
--->
-
-```{r data, fig.width=6, fig.height=7, fig.cap="Scatterplots of pairs of variables from samples of each family, showing the nuisance variables and structured variables."}
+```{r data, fig.width=6, fig.height=7, fig.cap="Scatterplots of pairs of variables from samples of each family, showing the nuisance variables and structured variables.", out.width = "0.8\\textwidth"}
 dsn <- spiral1000 %>% select(V1, V2) %>%
@@
 The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs.
-The scale for the holes index is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix).
-Similarly, the TIC index is re-scaled.
+The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix).
+Similarly, the TIC index is re-scaled depending on sample size.
 
@@
 
-The procedures for assessing PPI properties of smoothness, squintability, flexibility, rotation invariance, and speed are examined for samples from the family of data sets are as follows:
+The procedures for assessing the PPI properties of smoothness, squintability, flexibility, rotation invariance, and speed are examined for samples from the family of data sets are:
 

### 2019-01-08 scagnostics.Rmd
Appendix for squint angle estimation.

@@
 
-## Simulation study
+## Simulation study setup
 
@@
 
-<!-- - Compare the index values for structured vs nuisance variables, by simulating many examples from the model families, and recording the 90% confidence intervals.
-This checks if the PPI can detect the structure, that it has distinct differences between the sets of values.
--->
 1.
-Compute the PPI values on the tour path along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view either to the $x_3 - x_4$.
+Compute the PPI values on the tour path along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view to $x_3 - x_4$.
 The result is ideally a smooth change in low values.
@@
 2.
-Change to a tour path between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$, and compute the PPI along this.
+Change to a tour path between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$ via the intermediate projection onto $x_1 - x_5$, and compute the PPI along this.
 This examines the squintability, and smoothness.
@@
 
-<!--
-\textcolor{red}{\cite{posse95a} uses simulations with "strongly hidden structure", i.e.
-truncated Gaussian distributions with approximately identity covariance matrix, such that they don't appear in any of the principal component planes.
-In particular seems that spiral is hardest to identify based on results from Posse, maybe I can add this to my sphere dataset?
-or instead of it?}
--->
-
 ```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
@@
 ```
-
-<!--
-its possible to do a single figure legend - need to look up code for this
--->
 
@@
 
-<!--
-We next compare the various index values for three views: $x_1$ vs $x_2$, $x_5$ vs $x_6$ and $x_1$ vs $x_6$ (capturing differences from the different distribution in $x_6$).
-For the Spiral distribution the scagnostics indices skinny and stringy are large and close to their theoretic maximum for the special view.
-On the other hand while dcor2d is larger is also increased for the special view, the value is far from the theoretic maximum and larger values may be found in other views.
-MIC and TIC are also found to be sensitive to the spiral distribution.
-Considering next the Pipe distribution, we see that skinny, stringy and dcor2d are increased for the special view, but in particular MIC and TIC index are much larger and may be used to detect this type of structure in the dataset.
-Moreover, minimising the convex index (or maximising 1-convex) should also allow us to detect the special view.
-Finally the Sine distribution shows that several of the indices are maximised in the special view, in particular also the splines2d index designed for the detection of fictional dependence.
--->
-
 ## PPI traces over a tour sequence of interpolated nuisance projections
@@
 Several of the indexes are sensitive to sample size also, the same structured projection with differing numbers of points, produces different values.
-
-<!-- index values shouldn't be affected by sample size, on average it should be
-same value. Maybe some need to be normalised by n. Something needs fixing here.
-
-can we get "theoretical" min/max values, to assess the scale of the index for any particular data set.
--->
 
@@
 Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
-[@Grimm2016]'s indexes, cor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
-The holes index easily detects the pipe, and can see it from a distance.
+[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
+The holes index easily detects the pipe, and can see it from a distance, but also has local maxima along the tour path.
 The scagnostic index, stringy, can see the structure but is myopic, only when it is very close.
@@
 
-```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between nuisance and structured projections, following $x_1$-$x_2$ to $x_1$-$x_5$ to $x_5$-$x_6$ Solid lines are the larger sample size, $n=1000$, and dashed are smaller sample size, $n=100$.
+```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between nuisance and structured projections, following $x_1$-$x_2$ to $x_1$-$x_5$ to $x_5$-$x_6$ (solid lines: $n=1000$; dashed: $n=100$).
 Peaks at the end of the sequence indicate the index sees the structure.
 The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection.
-Holes only responds to the pipe, and is a multimodal function for this data with a local maximum at $x_1$-$x_5$ The scagnostics tend to be sensitive to sample size.
+Holes only responds to the pipe, and is a multimodal function for this data with a local maximum at $x_1$-$x_5$.
 "}
@@
 For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimist the TIC index.
-The results are shown in Figures \ref{fig:pipeFirstRun}.
 
@@
 
-```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Scouting for the pipe using optimization method 2 on the TIC index.
-Other PPI values also shown for interest.
-Only the values for the target plane are shown because the optimization is done off-line.
-The TIC maximum value for this data is small,  even though it identifies the pipe.
-The holes, convex, skinny, and to some extent MIC, mirror the performance of TIC for this structure."}
+```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Scouting for the pipe using optimization method 2 on the TIC index, other PPI values also shown for interest.
+Only the values for the target planes are shown.
+Despite the small maximum value of TIC for this data, it identifies the pipe."}
 if(!file.exists("cache/pipeFirstRun.rda")){
@@
 
-Given the patterns in Figure\ref{fig:wIntermediate} it would be expected that the sine could be found easily, using only optimization method 3 using the splines2d, dcor2d, MIC or TIC indexes.
+Given the patterns in Figure \ref{fig:wIntermediate} it would be expected that the sine could be found easily, using only optimization method 3 using the splines2d, dcor2d, MIC or TIC indexes.
 This is examined in Figure \ref{fig:findsine}.
@@
 
-<!--, this will be discussed in some more details below.
-Here we only note that even for anchor planes highlighted by blue vertical lines the index sometimes decreases because of this dependence.
-These are however not dramatic effects because this search method is closely related to how interpolation is done (a different picture would be obtained using method 2 which often leads to large drops in index values after the interpolation step).-->
-
-<!-- Here there are four index functions that appear to be suitable for the optimization: splines2d, dcor2d, MIC and TIC.
-On the other hand, while some of the scagnostics indices are reaching their maximum/minimum in the final projection, they do not show a smooth increase/decrease thus preventing efficient optimization when considering them as index functions.
-The final view identified by the guided tour is shown in Figure \ref{fig:findsine}.
-
-\textcolor{red}{This isn't actually the "ideal" view, could be because of rotation dependence??}-->
-
 ```{r findsine, fig.width=6, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Guided tour optimising  the splines2d index, using method 3, for the sine data, with $n=1000$.
 Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima.
-The sine is found relatively easily, by splines2D, and it si indicated that MIC, TIC, dcor2D and convex would also likely find this structure.", results="hide"}
+The sine is found relatively easily, by splines2D, and it is indicated that MIC, TIC, dcor2D and convex would also likely find this structure.", results="hide"}
 set.seed(2018)
@@
 
-<!--
-Di says probably we can skip this plot
-```{r testsine, out.width=".5\\textwidth", fig.cap="Final view identified by the optimisation."}
-iLast <- length(sineTourFull)
-fProj <- sineTourFull[[iLast]]
-dProj <- as.tibble(sineResc %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point()
-```
--->
-
 ### Spiral detection
@@
 The spiral is the most challenging structure to detect because it has a small squint angle [@posse95a], especially as the ratio of noise to structure dimensions increases.
-This is explored in this section, using optimisation method 2 to scout the space for approximate maxima first.
+This is explored using optimisation method 2 to scout the space for approximate maxima.
 The skinny scagnostic index is used because it was observed (Figure \ref{fig:wIntermediate}) to be sensitive to this structure, although the noisiness of the index might be problematic.
@@
 
-<!-- Here we explore the role of dimensionality in the index optimization, using the skinny index that was found to be sensitive to this structure.
-We use a setup similar to that introduced for detecting the pipe structure, i.e.\ we use the search\_better optimization of the guided tour, disabling the cooling and with a large search window, while sampling a large number of random projections (up to 5000) for each step.
--->
-
 The search is conducted for $p=4,5,6$ which would correspond to 2, 3 and 4 noise dimensions respectively.
-In addition we examine the distance between planes, using a Frobenius norm, as defined by Equation 2 of @BCAH05, and available in the ```proj_dist``` function in the tourr package, to compare searches with squint angle across dimensions.
-The distance between planes is related to squint angle, how far away from the ideal projection, can the structure be glimpsed.
+In addition we examine the distance between planes, using a Frobenius norm, as defined by Equation 2 of @BCAH05, and available in the ```proj_dist``` function in the tourr package, to compare searches across dimensions.
+The distance between planes is related to squint angle, how far away from the ideal projection can the structure be glimpsed.
+We estimate the squint angle depending on the number of noise dimensions in the appendix.
 In order for the optimizer to find the spiral, the distance between planes would need to be smaller than the squint angle.
@@
 When $p=4$ the scouting method effectively finds the spiral.
-Plot (a) shows the side-by-side boxplots of distance between ideal plane and planes examined during the optimization, for $p=4,5,6$.
+Plot (a) shows the side-by-side boxplots of pairwise distances between planes examined during the optimization, for $p=4,5,6$.
 These are on average smaller for the lower dimension, and gradually increase as dimension increases.
 This is an indication of the extra computation needed to brute force find the spiral as noise dimensions increase.
-Plot (b) shows these as a function of the optimization procedure, where it can be seen that only when $p=4$ does it converge to the ideal.
-Its likely that expanding the search space should result in uncovering the spiral in higher dimensions.
-
-<!-- We consider distributions in d=4,5,6 dimensions (subsets of the spiral dataset) and compare features found in the tour.
-In d=4 dimensions the optimization identifies a view showing the spiral after 39 steps.
-For larger values of d the optimization does not uncover the spiral view, and we limit the optimizer to 100 steps to compare the so obtained bases to the results found in four dimensions.
-Results are shown in the top panel Figure \ref{fig:findspiral}.
-The top left is showing the pairwise distances between all planes found by the optimization, note that while d=5,6 contains 100 planes, d=4 only contains 39 planes.
-The top right is then comparing the evolution of the distance from the plane containing the spiral view with progression of the optimization.
-We see that the pairwise distances grow with d (what do we learn from that?).
-Moreover we find that the distance to the special view is far from monotonically decreasing towards zero, and large jumps are possible as a consequence of rotation dependence.
-Note that the spiral itself is approximately rotation invariant, but partial views are not.
-At medium distances from the special view we often observe a jump to higher distances as consequence of in-plane rotation dependence for such partial views.
-This stabilizes closer to the special view, however it relies on randomly identifying a projection within this angle.
-As a consequence increasing dimensionality dramatically reduces the chance of identifying structures with high index values.-->
+Plot (b) shows the distance of the plane in each iteration of the optimisation to the ideal plane, where it can be seen that only when $p=4$ does it converge to the ideal.
+Its likely that expanding the search space should result in uncovering the spiral in higher dimensions, which however requires tuning of the stopping conditions and long run times.
+
+XXX update the plots, maybe resulting views on top?
 
 ```{r findspiral, out.width="\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with p = 4, 5, 6.
-The top left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the top right shows the evolution of distance to the plane containing the spiral as the index is being optimised.
+The top left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the top right shows the evolution of distance to the ideal plane as the index is being optimised.
 The bottom row shows the final view obtained for p = 4 (left), p = 5 (middle) and p = 6 (right).", results="hide"}
@@
 
-```
-
-```{r squintAngle, out.width="0.5\\textwidth", fig.cap="Estimated squint angles for the Spiral dataset with 1000 datapoints, with p = 4, 5, 6, containing estimates evaluated for 100 randomly selected directions each.", results="hide"}
-
-squintAngleEstimat <- function(data, indexF, cutoff, structurePlane, n = 100, stepSize = 0.01){
-  # data = numerical input with p > 2 dimensions
-  # indexF = the index function used to estimate the squint angle
-  # cutoff = lower bound on the index value to be considered structured (i.e. all lower index values are outside the squint angle)
-  # structuredPlane = projection matrix onto plane containing the 2-d structure
-  # n = number of random starting directions over which the squint angle estimate is averaged
-  # stepSize = accuracy, step size determines where the index is evaluated to find the first plane above cutoff
-  data <- as.matrix(data) # make sure data is in matrix format
-  angles <- numeric(length = n) # collecting all individual estimates
-  i <- 1
-  p <- ncol(data)
-  while(i <= n){
-    # first generate random direction, make sure it is not too close to structure plane
-    dist <- 0.
-    while(dist < 0.1){
-      rBasis <- tourr::basis_random(p)
-      dist <- tourr::proj_dist(rBasis, structurePlane)
-    }
-    # now interpolate from rBasis to structure plane with selected step size
-    # since planned tour ignores first two entries, generate some random planes to be ignored
-    notUsed1 <- tourr::basis_random(p)
-    notUsed2 <- tourr::basis_random(p)
-    tourHist <- save_history(data,tour_path=planned_tour(list(notUsed1, notUsed2, rBasis, structurePlane)))
-    allBases <- as.list(interpolate(tourHist, angle = stepSize))
-    cIndex <- 0.
-    j <- 1
-    while(cIndex < cutoff){
-      cProj <- data %*% allBases[[j]]
-      cIndex <- indexF(cProj)
-      j <- j+1
-    }
-    cDist <- tourr::proj_dist(allBases[[j]], structurePlane)
-    angles[i] <- cDist
-    i <- i+1
-  }
-  return(angles)
-}
-
-if(!file.exists("cache/squintAngle.rda")){
-  allAngles <- tibble(d=numeric(), angle=numeric())
-  for(d in c(4,5,6)){
-    set.seed(58958)
-    specialPlane <- matrix(c(rep(0,d-2),1,0,rep(0,d-1),1),ncol=2)
-    spiralD <- spiral1000 %>%
-      purrr::when(d<6 ~ select(., -V4), ~ .) %>%
-      purrr::when(d<5 ~ select(., -V3), ~ .)
-    cAngles <- squintAngleEstimat(spiralD, scagIndex("Skinny"), 0.6, specialPlane)
-    allAngles <- bind_rows(allAngles, tibble(d=d, angle=cAngles))
-  }
-  save(allAngles, file = "cache/squintAngle.rda")
-} else {
-  load("cache/squintAngle.rda")
-}
-
-allAngles <- rename(allAngles, p=d)
-
-ggplot(allAngles, aes(p, angle, group=p)) + geom_boxplot()
 ```
@@
 
+## Extimating the squint angle of the spiral
+
+We estimate the squint angle of the spiral (for p = 4, 5, 6) as follows.
+First pick a random starting plane and generate a tour path from the starting plane to the ideal plane containing the spiral.
+Using skinny as the reference index we fix a lower index value that is attributed to indicate squintability at $0.6$, and we move along the tour path towards the ideal plane until this value is reached.
+The distance between the thus identified plane and the ideal plane is used as an estimate of the squint angle in this direction.
+Since this will strongly depend on the random starting plane, i.e.
+the considered direction, we repeat the estimation 100 times and present the results in the form of a box plot in Figure \ref{fig:squintAngle}.
+The result shows a large drop in squint angle when going from p = 4 to higher dimensions, and generally a large spread of squint angles depending on the direction.
+
+```{r squintAngle, out.width="0.5\\textwidth", fig.cap="Estimated squint angles for the Spiral dataset with 1000 datapoints, with p = 4, 5, 6, containing estimates evaluated for 100 randomly selected directions each.", results="hide"}
+
+squintAngleEstimat <- function(data, indexF, cutoff, structurePlane, n = 100, stepSize = 0.01){
+  # data = numerical input with p > 2 dimensions
+  # indexF = the index function used to estimate the squint angle
+  # cutoff = lower bound on the index value to be considered structured (i.e. all lower index values are outside the squint angle)
+  # structuredPlane = projection matrix onto plane containing the 2-d structure
+  # n = number of random starting directions over which the squint angle estimate is averaged
+  # stepSize = accuracy, step size determines where the index is evaluated to find the first plane above cutoff
+  data <- as.matrix(data) # make sure data is in matrix format
+  angles <- numeric(length = n) # collecting all individual estimates
+  i <- 1
+  p <- ncol(data)
+  while(i <= n){
+    # first generate random direction, make sure it is not too close to structure plane
+    dist <- 0.
+    while(dist < 0.1){
+      rBasis <- tourr::basis_random(p)
+      dist <- tourr::proj_dist(rBasis, structurePlane)
+    }
+    # now interpolate from rBasis to structure plane with selected step size
+    # since planned tour ignores first two entries, generate some random planes to be ignored
+    notUsed1 <- tourr::basis_random(p)
+    notUsed2 <- tourr::basis_random(p)
+    tourHist <- save_history(data,tour_path=planned_tour(list(notUsed1, notUsed2, rBasis, structurePlane)))
+    allBases <- as.list(interpolate(tourHist, angle = stepSize))
+    cIndex <- 0.
+    j <- 1
+    while(cIndex < cutoff){
+      cProj <- data %*% allBases[[j]]
+      cIndex <- indexF(cProj)
+      j <- j+1
+    }
+    cDist <- tourr::proj_dist(allBases[[j]], structurePlane)
+    angles[i] <- cDist
+    i <- i+1
+  }
+  return(angles)
+}
+
+if(!file.exists("cache/squintAngle.rda")){
+  allAngles <- tibble(d=numeric(), angle=numeric())
+  for(d in c(4,5,6)){
+    set.seed(58958)
+    specialPlane <- matrix(c(rep(0,d-2),1,0,rep(0,d-1),1),ncol=2)
+    spiralD <- spiral1000 %>%
+      purrr::when(d<6 ~ select(., -V4), ~ .) %>%
+      purrr::when(d<5 ~ select(., -V3), ~ .)
+    cAngles <- squintAngleEstimat(spiralD, scagIndex("Skinny"), 0.6, specialPlane)
+    allAngles <- bind_rows(allAngles, tibble(d=d, angle=cAngles))
+  }
+  save(allAngles, file = "cache/squintAngle.rda")
+} else {
+  load("cache/squintAngle.rda")
+}
+
+allAngles <- rename(allAngles, p=d)
+
+ggplot(allAngles, aes(p, angle, group=p)) + geom_boxplot()
+```
+
 ## Binning sensitivity of MIC index

### 2019-01-08 scagnostics.Rmd
Less plots in Fig 6.

@@
 
-XXX update the plots, maybe resulting views on top?
-
-```{r findspiral, out.width="\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with p = 4, 5, 6.
-The top left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the top right shows the evolution of distance to the ideal plane as the index is being optimised.
-The bottom row shows the final view obtained for p = 4 (left), p = 5 (middle) and p = 6 (right).", results="hide"}
+```{r findspiral, out.width="0.9\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with p = 4, 5, 6.
+The left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the right shows the evolution of distance to the ideal plane as the index is being optimised.", results="hide", fig.width=7, fig.height=2.5}
 
@@
 
-grid.arrange(p1, p2, fViewL[[1]], fViewL[[2]], fViewL[[3]], layout_matrix = rbind(c(1,2,2),c(3,4,5)))
-
+#grid.arrange(p1, p2, fViewL[[1]], fViewL[[2]], fViewL[[3]], layout_matrix = rbind(c(1,2,2),c(3,4,5)))
+grid.arrange(p1, p2, layout_matrix = rbind(c(1,2,2)))
 ```

### 2019-01-08 scagnostics.Rmd
Shifting more things to the appendix.

@@
 The parameters for these PPIs is varied and examined along a tour path between noise and structured projections, to examine the effects of parameter choices and how fiddly the index is to tune.
-6. Computational speed for the selected indices is examined on a range of sample sizes.
+Results of this study can be found in the appendix together with preliminary evaluation of potential improvements via smoothing.
+6. Computational speed for the selected indices is examined on a range of sample sizes. Results are presented in the appendix.
 
@@
 ```{r rotationDep, out.width=".6\\textwidth", fig.cap="PPI for rotations of the sine 1000 data, to examine rotation invariance.
-Most are close to rotation invariant, except for dcor2d, splines2d and TIC."}
+Most are close to rotation invariant, except for skinny, dcor2d, splines2d and TIC."}
 if(!file.exists("cache/rotationDep.rda")){
@@
   xlab("") + ylab("")
-```
-
-## Computational performance
-
-\label{sec:timer}
-Computational time is important for using the PPIs with the guided tour, online.
-Figure \ref{fig:timer} summarizes performance for each PPI.
-For simplicity, data with sample sizes ranging from 100-10000 are drawn from a 6-d solid sphere, using the geozoo package [@geozoo].
-The time to compute the PPIs over 100 interpolated grand tour projections is recorded.
-The scagnostics PPI are computed as a bundle, since this is the code base, and that major computational constraint is common to all the scagnostics.
-There are two versions of the MIC and TIC algorithm, labelled MINE and MINE E, the second being a newer algorithm which improves their computational performance.
-
-The results are interesting. The scagnostic indexes and splines2D are very fast regardless of sample size. MIC, TIC (both versions) and dcor2D slow rapidly as sample size increases.
-
-```{r getTimer}
-
-mineE <- function(x,y){
-  return(mine(x=x, y=y, est = "mic_e"))
-}
-
-timeThis <- function(d, t, idx, pmax, n, idxName){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections, n=sample size, idxName=str name of idx funciton
-    i <- 1
-    dfTimer = data.frame( t= numeric(), i=numeric(), n = numeric(), name=character())
-    for(pMatrix in t){
-      if(i>pmax) break
-      tic.clearlog()
-      tic() #start timer
-      dProj <- d %*% pMatrix
-      sgnst <- idx(dProj[,1],dProj[,2])
-      toc(log=TRUE,quiet=TRUE)
-      scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
-      dfTimer <- add_row(dfTimer, t=scTd, i=i, n= n, name=idxName)
-      i = i+1
-    }
-    return(dfTimer)
-}
-
-```
-
-```{r timer, fig.cap="Computational perfoamce for PPIs, using sample sizes 100-10000.
-Colour indicates the PPI.
-Because the scagnostics calculation is bundled together, the values are the same for all these indexes, and they are really fast to compute.
-MINE includes the MIC and TIC indexes, and MINE E are computationally more efficient algorithms for these.
-These, along with dcor2D, are slower with larger sample sizes.", out.width=".6\\textwidth"}
-if(!file.exists("cache/timer.rda")){
-  set.seed(2018)
-  grandTour100 <- save_history(sphereData(6,100), grand_tour(2), max=4) %>%
-    interpolate() %>%
-    as.list()
-
-  sizeL <- c(10, 100, 500, 1000, 5000, 10000)
-  t <- 1
-  dfTimer <- NULL
-  for(sampleSize in sizeL){
-    set.seed(sampleSize + 11) # new seed for each sample size
-    if(sampleSize == 0) sampleSize = 10 # smallest considered sample has 10 points
-    thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
-    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100, sampleSize, "scagnostics")
-    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100, sampleSize, "dcor2D")
-    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100, sampleSize, "splines2D")
-    mineT <- timeThis(thisMatrix, grandTour100, mine, 100, sampleSize, "MINE")
-    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100, sampleSize, "MINE E")
-    dfTimerC <- list(scagT, dcorT, splineT, mineT,mineeT) %>%
-      reduce(rbind)
-    dfTimer <- rbind(dfTimer, dfTimerC)
-    t <-  t+1
-  }
-  save(dfTimer, file = "cache/timer.rda")
-} else {
-  load("cache/timer.rda")
-}
-
-timerMeans <- dfTimer %>%
-  group_by(name, n) %>%
-  summarise(t=mean(t))
-
-ggplot(dfTimer, aes(x = n, y = t)) + 
-  geom_point(alpha=0.1, aes(color=name, shape=name)) +
-  geom_point(data=timerMeans, aes(color=name, shape=name)) +
-  geom_line(data=timerMeans, aes(linetype=name, color=name)) +
-  scale_x_log10(limits=c(90,10100)) +
-#  scale_y_log10(limits=c(0.00001,1)) + 
-  labs(x = "Sample Size", y = "Time [s]", color = "Index family", 
-       linetype = "Index family", shape = "Index family") + 
-  scale_colour_brewer(palette="Dark2")
-```
-
-## Effect of parameter choice in index value
-\label{sec:params}
-Some parameters must be provided for some PPIs.
-This can be advantageous, allowing the index to more flexibly work for different types of structure, controlling trade-offs between noise and fine structure detection, and affecting computing time and precision.
-
-- Binning:
-    - Scagnostics: the number of bins can be controlled by the user, note however that internally the implementation will reduce the number of bins if too many non-empty bins are found (more than 250).
-    - MINE: the maximum number of bins considered is fixed by the user as a function of the number of data points, see Equation \ref{eq:MIC}.
-The default is chosen such as a trade-off between resolution and noise dependence, but it may be tuned based on requirements dictated by specific datasets.
-Apart from sensitivity to noise computing time may also be a consideration here.
-- Spline knots: for the splines2D measure we need to fix the number of knots.
-By default it is fixed to be 10 (or lower if appropriate based on the data values).
-In our examples we find the number to be appropriate to identify functional dependence while rejecting noise, but some distributions may require tuning of this parameter.
-
-<!--### Effect of binning in scagnostics -->
-
-The bins argument for the scagnostics might be reasonably expected to affect the smoothness of the index: a small number of bins should provide a smooth index function, but may affect its ability to detect fine structure.
-Figure \ref{fig:sprialScagBinning} examines this.
-Scagnostics PPIs are computed for the spiral1000 data on a tour path between $x_1$-$x_2$ to $x_5$-$x_6$, for number of bins equal to 10, 20, 50.
-The interesting observation to make is that even with small bin size the functions are all relatively noisy.
-The problem with the small bin size is that the spiral becomes invisible to the PPI.
-
-XXX from caption, remove?:  For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
-With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
-For the convex index we also find bins=50 to be the better choice.
-
-```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.", out.width=".7\\textwidth"}
-getScagComp <- function(df, tPath){
-  sc <- tibble(
-    convex10=numeric(),
-    skinny10=numeric(),
-    stringy10=numeric(),
-    convex20=numeric(),
-    skinny20=numeric(),
-    stringy20=numeric(),
-    convex50=numeric(),
-    skinny50=numeric(),
-    stringy50=numeric(),
-    t=numeric())
-  n <- length(tPath)
-  for (i in 1:n) {
-    dprj <- as.matrix(df) %*% tPath[[i]]
-    scagRes10 <- scagnostics.default(dprj[,1], dprj[,2], bins=10)$s
-    scagRes20 <- scagnostics.default(dprj[,1], dprj[,2], bins=20)$s
-    scagRes50 <- scagnostics.default(dprj[,1], dprj[,2], bins=50)$s
-    sc <- add_row(sc,
-                  convex10=scagRes10["Convex"], skinny10=scagRes10["Skinny"],
-                  stringy10=scagRes10["Stringy"],
-                  convex20=scagRes20["Convex"], skinny20=scagRes20["Skinny"],
-                  stringy20=scagRes20["Stringy"],
-                  convex50=scagRes50["Convex"], skinny50=scagRes50["Skinny"],
-                  stringy50=scagRes50["Stringy"],
-                  t=i)
-  }
-
-  return(sc)
-}
-if(!file.exists("cache/spiralScagBinning.rda")){
-  scagBinningDf <- getScagComp(spiral1000, t2full)
-  scagMelt <- gather(scagBinningDf, variable, value, -t) %>%
-    mutate(nbin = as.numeric(str_sub(variable, start = -2))) %>%
-    mutate(idx = str_sub(variable, end = -3))
-  save(scagBinningDf, scagMelt, file = "cache/spiralScagBinning.rda")
-} else {
-  load("cache/spiralScagBinning.rda")
-}
-ggplot(scagMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(nbin))) + 
-  facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
-  scale_colour_discrete("Num bins") + xlab("Number of projections") + 
-  ylab("PPI value")
-
-```
-
-The sensitivity to the binning for calculating the MIC index is examined for one example in the Appendix XXX (add cross reference)
-
-## Ways to refine the PPIs
-
-The biggest issues revealed by the investigation into the new PPIs are a lack of smoothness, particularly for the scagnostics indexes, and the rotation invariance of Grimm's indexes.
-To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighbourhood of projections and average the value, or alternatively average the PPI for several jittered projections.
-This is investigated in Fig.
-\ref{fig:spiralScagSmoothing}.
-Rotation invariance is more difficult to fix, but an alternative tour interpolation method could be useful.
-The geodesic interpolation transitions between planes, and it ignores the basis defining the plane, creating a problem with rotation invariant indices.
-Alternative interpolations based on Givens or Householder rotations could be implemented to transition between bases, which should alleviate the need for rotationally invariant indices.
-
-<!-- ### Smoothing-->
-
-Two different methods are considered for smoothing the index values:
-
-* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$.
-* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction.
-
-<!--
-* Turning axes: given the rotation non-invariance, maybe rotation of basis could be useful also??
--->
-
-The mean value from a sample of projections is recorded as the PPI value. This could be robsitufied by dropping the most extreme values.
-
-This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
-Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values.
-Both methods appear to be promising in smoothing the function.
-Because the scagnistics are fast to compute, either of these methods is feasible.
-For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.
-
-<!--\textcolor{red}{I fixed the y axis to be between 0 and 1, which highlights that indeed we get much smoother behaviour, on the other hand theres still large relative jumps observed in all traces, and a smaller y axis scale may be preferred to show this more clearely.}-->
-
-```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
-While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered.
-For comparison the red line in the background shows the trace without any smoothing applied."}
-
-jitterScagnostics <-function(proj, d, alpha){
-  newProj <- tourr:::basis_nearby(proj, alpha = alpha, method = "geodesic")
-  newD <- d %*% newProj
-  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
-}
-
-jitterPointsScagnostics <-function(projData, alpha){
-  newD <- jitter(projData, amount=alpha)
-  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
-}
-
-getSgnMean <- function(proj, d, alpha, method){
-  dProj <- d %*% proj
-  orig <- as.vector(scagnostics.default(dProj[,1],dProj[,2])$s)
-  if(method == "jitterAngle"){
-    sgnVec <- replicate(10, jitterScagnostics(proj, d, alpha))
-  } else if (method=="jitterPoints"){
-    sgnVec <- replicate(10, jitterPointsScagnostics(dProj, alpha))
-  }
-  return(rowMeans(cbind(orig, sgnVec)))
-}
-
-getScagSmooth <- function(df, tPath){
-  sc <- tibble(
-    convex=numeric(),
-    skinny=numeric(),
-    stringy=numeric(),
-    t=numeric(),
-    method=character(),
-    alpha=numeric())
-  n <- length(tPath)
-  for (method in c("jitterAngle", "jitterPoints")){
-    for (alpha in c(0.01, 0.05, 0.1)){
-      for (i in 1:n) {
-        scagMean <- getSgnMean(tPath[[i]], as.matrix(df), alpha, method)
-        convex <- scagMean[6]
-        skinny <- scagMean[7]
-        stringy <- scagMean[8]
-        sc <- sc %>% add_row(convex=convex, skinny=skinny, stringy=stringy,
-                             t=i, method=method, alpha=alpha)
-      }
-    }
-  }
-
-  return(sc)
-}
-
-if(!file.exists("cache/spiralScagSmoothing.rda")){
-  scagSmooth <- getScagSmooth(spiral1000, as.list(t1full))
-  scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
-  save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")
-} else {
-  load("cache/spiralScagSmoothing.rda")
-}
-
-#getting non-smoothed result
-load("cache/V1V2toV3V4.rda")
-noSmoothing <- fullRes %>%
-  filter(size==1000) %>%
-  filter(name=="spiral") %>%
-  select(skinny, stringy, t) %>%
-  gather(variable, value, -t)
-
-scagSmoothMelt <- scagSmoothMelt %>%
-  filter(variable != "convex") %>%
-  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
-ggplot(scagSmoothMelt, aes(x=t, y=value)) +
-  geom_line(data=noSmoothing, color="red", alpha=0.5) +
-  geom_line(aes(linetype=factor(method))) + 
-  theme(legend.position="none") +
-  ylim(0,1) +
-  facet_grid(alpha~variable) +
-  xlab("Sequence of projections (t)") +
-  ylab("PPI value")
-
 ```
@@
 
+## Computational performance
+
+\label{sec:timer}
+Computational time is important for using the PPIs with the guided tour, online.
+Figure \ref{fig:timer} summarizes performance for each PPI.
+For simplicity, data with sample sizes ranging from 100-10000 are drawn from a 6-d solid sphere, using the geozoo package [@geozoo].
+The time to compute the PPIs over 100 interpolated grand tour projections is recorded.
+The scagnostics PPI are computed as a bundle, since this is the code base, and that major computational constraint is common to all the scagnostics.
+There are two versions of the MIC and TIC algorithm, labelled MINE and MINE E, the second being a newer algorithm which improves their computational performance.
+
+The results are interesting. The scagnostic indexes and splines2D are very fast regardless of sample size. MIC, TIC (both versions) and dcor2D slow rapidly as sample size increases.
+
+```{r getTimer}
+
+mineE <- function(x,y){
+  return(mine(x=x, y=y, est = "mic_e"))
+}
+
+timeThis <- function(d, t, idx, pmax, n, idxName){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections, n=sample size, idxName=str name of idx funciton
+    i <- 1
+    dfTimer = data.frame( t= numeric(), i=numeric(), n = numeric(), name=character())
+    for(pMatrix in t){
+      if(i>pmax) break
+      tic.clearlog()
+      tic() #start timer
+      dProj <- d %*% pMatrix
+      sgnst <- idx(dProj[,1],dProj[,2])
+      toc(log=TRUE,quiet=TRUE)
+      scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
+      dfTimer <- add_row(dfTimer, t=scTd, i=i, n= n, name=idxName)
+      i = i+1
+    }
+    return(dfTimer)
+}
+
+```
+
+```{r timer, fig.cap="Computational perfoamce for PPIs, using sample sizes 100-10000.
+Colour indicates the PPI.
+Because the scagnostics calculation is bundled together, the values are the same for all these indexes, and they are really fast to compute.
+MINE includes the MIC and TIC indexes, and MINE E are computationally more efficient algorithms for these.
+These, along with dcor2D, are slower with larger sample sizes.", out.width=".6\\textwidth"}
+if(!file.exists("cache/timer.rda")){
+  set.seed(2018)
+  grandTour100 <- save_history(sphereData(6,100), grand_tour(2), max=4) %>%
+    interpolate() %>%
+    as.list()
+
+  sizeL <- c(10, 100, 500, 1000, 5000, 10000)
+  t <- 1
+  dfTimer <- NULL
+  for(sampleSize in sizeL){
+    set.seed(sampleSize + 11) # new seed for each sample size
+    if(sampleSize == 0) sampleSize = 10 # smallest considered sample has 10 points
+    thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
+    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100, sampleSize, "scagnostics")
+    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100, sampleSize, "dcor2D")
+    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100, sampleSize, "splines2D")
+    mineT <- timeThis(thisMatrix, grandTour100, mine, 100, sampleSize, "MINE")
+    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100, sampleSize, "MINE E")
+    dfTimerC <- list(scagT, dcorT, splineT, mineT,mineeT) %>%
+      reduce(rbind)
+    dfTimer <- rbind(dfTimer, dfTimerC)
+    t <-  t+1
+  }
+  save(dfTimer, file = "cache/timer.rda")
+} else {
+  load("cache/timer.rda")
+}
+
+timerMeans <- dfTimer %>%
+  group_by(name, n) %>%
+  summarise(t=mean(t))
+
+ggplot(dfTimer, aes(x = n, y = t)) + 
+  geom_point(alpha=0.1, aes(color=name, shape=name)) +
+  geom_point(data=timerMeans, aes(color=name, shape=name)) +
+  geom_line(data=timerMeans, aes(linetype=name, color=name)) +
+  scale_x_log10(limits=c(90,10100)) +
+#  scale_y_log10(limits=c(0.00001,1)) + 
+  labs(x = "Sample Size", y = "Time [s]", color = "Index family", 
+       linetype = "Index family", shape = "Index family") + 
+  scale_colour_brewer(palette="Dark2")
+```
+
+## Effect of parameter choice in index value
+\label{sec:params}
+Some parameters must be provided for some PPIs.
+This can be advantageous, allowing the index to more flexibly work for different types of structure, controlling trade-offs between noise and fine structure detection, and affecting computing time and precision.
+
+- Binning:
+    - Scagnostics: the number of bins can be controlled by the user, note however that internally the implementation will reduce the number of bins if too many non-empty bins are found (more than 250).
+    - MINE: the maximum number of bins considered is fixed by the user as a function of the number of data points, see Equation \ref{eq:MIC}.
+The default is chosen such as a trade-off between resolution and noise dependence, but it may be tuned based on requirements dictated by specific datasets.
+Apart from sensitivity to noise computing time may also be a consideration here.
+- Spline knots: for the splines2D measure we need to fix the number of knots.
+By default it is fixed to be 10 (or lower if appropriate based on the data values).
+In our examples we find the number to be appropriate to identify functional dependence while rejecting noise, but some distributions may require tuning of this parameter.
+
+<!--### Effect of binning in scagnostics -->
+
+The bins argument for the scagnostics might be reasonably expected to affect the smoothness of the index: a small number of bins should provide a smooth index function, but may affect its ability to detect fine structure.
+Figure \ref{fig:sprialScagBinning} examines this.
+Scagnostics PPIs are computed for the spiral1000 data on a tour path between $x_1$-$x_2$ to $x_5$-$x_6$, for number of bins equal to 10, 20, 50.
+The interesting observation to make is that even with small bin size the functions are all relatively noisy.
+The problem with the small bin size is that the spiral becomes invisible to the PPI.
+
+XXX from caption, remove?:  For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
+With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
+For the convex index we also find bins=50 to be the better choice.
+
+```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.", out.width=".7\\textwidth"}
+getScagComp <- function(df, tPath){
+  sc <- tibble(
+    convex10=numeric(),
+    skinny10=numeric(),
+    stringy10=numeric(),
+    convex20=numeric(),
+    skinny20=numeric(),
+    stringy20=numeric(),
+    convex50=numeric(),
+    skinny50=numeric(),
+    stringy50=numeric(),
+    t=numeric())
+  n <- length(tPath)
+  for (i in 1:n) {
+    dprj <- as.matrix(df) %*% tPath[[i]]
+    scagRes10 <- scagnostics.default(dprj[,1], dprj[,2], bins=10)$s
+    scagRes20 <- scagnostics.default(dprj[,1], dprj[,2], bins=20)$s
+    scagRes50 <- scagnostics.default(dprj[,1], dprj[,2], bins=50)$s
+    sc <- add_row(sc,
+                  convex10=scagRes10["Convex"], skinny10=scagRes10["Skinny"],
+                  stringy10=scagRes10["Stringy"],
+                  convex20=scagRes20["Convex"], skinny20=scagRes20["Skinny"],
+                  stringy20=scagRes20["Stringy"],
+                  convex50=scagRes50["Convex"], skinny50=scagRes50["Skinny"],
+                  stringy50=scagRes50["Stringy"],
+                  t=i)
+  }
+
+  return(sc)
+}
+if(!file.exists("cache/spiralScagBinning.rda")){
+  scagBinningDf <- getScagComp(spiral1000, t2full)
+  scagMelt <- gather(scagBinningDf, variable, value, -t) %>%
+    mutate(nbin = as.numeric(str_sub(variable, start = -2))) %>%
+    mutate(idx = str_sub(variable, end = -3))
+  save(scagBinningDf, scagMelt, file = "cache/spiralScagBinning.rda")
+} else {
+  load("cache/spiralScagBinning.rda")
+}
+ggplot(scagMelt, aes(x=t, y=value)) +
+  geom_line(aes(color=factor(nbin))) + 
+  facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
+  scale_colour_discrete("Num bins") + xlab("Number of projections") + 
+  ylab("PPI value")
+
+```
+
+The sensitivity to the binning for calculating the MIC index is examined for one example in the Appendix XXX (add cross reference)
+
+## Ways to refine the PPIs
+
+The biggest issues revealed by the investigation into the new PPIs are a lack of smoothness, particularly for the scagnostics indexes, and the rotation invariance of Grimm's indexes.
+To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighbourhood of projections and average the value, or alternatively average the PPI for several jittered projections.
+This is investigated in Fig.
+\ref{fig:spiralScagSmoothing}.
+Rotation invariance is more difficult to fix, but an alternative tour interpolation method could be useful.
+The geodesic interpolation transitions between planes, and it ignores the basis defining the plane, creating a problem with rotation invariant indices.
+Alternative interpolations based on Givens or Householder rotations could be implemented to transition between bases, which should alleviate the need for rotationally invariant indices.
+
+Two different methods are considered for smoothing the index values:
+
+* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$.
+* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction.
+
+The mean value from a sample of projections is recorded as the PPI value. This could be robsitufied by dropping the most extreme values.
+
+This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
+Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values.
+Both methods appear to be promising in smoothing the function.
+Because the scagnistics are fast to compute, either of these methods is feasible.
+For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.
+
+```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
+While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered.
+For comparison the red line in the background shows the trace without any smoothing applied."}
+
+jitterScagnostics <-function(proj, d, alpha){
+  newProj <- tourr:::basis_nearby(proj, alpha = alpha, method = "geodesic")
+  newD <- d %*% newProj
+  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
+}
+
+jitterPointsScagnostics <-function(projData, alpha){
+  newD <- jitter(projData, amount=alpha)
+  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
+}
+
+getSgnMean <- function(proj, d, alpha, method){
+  dProj <- d %*% proj
+  orig <- as.vector(scagnostics.default(dProj[,1],dProj[,2])$s)
+  if(method == "jitterAngle"){
+    sgnVec <- replicate(10, jitterScagnostics(proj, d, alpha))
+  } else if (method=="jitterPoints"){
+    sgnVec <- replicate(10, jitterPointsScagnostics(dProj, alpha))
+  }
+  return(rowMeans(cbind(orig, sgnVec)))
+}
+
+getScagSmooth <- function(df, tPath){
+  sc <- tibble(
+    convex=numeric(),
+    skinny=numeric(),
+    stringy=numeric(),
+    t=numeric(),
+    method=character(),
+    alpha=numeric())
+  n <- length(tPath)
+  for (method in c("jitterAngle", "jitterPoints")){
+    for (alpha in c(0.01, 0.05, 0.1)){
+      for (i in 1:n) {
+        scagMean <- getSgnMean(tPath[[i]], as.matrix(df), alpha, method)
+        convex <- scagMean[6]
+        skinny <- scagMean[7]
+        stringy <- scagMean[8]
+        sc <- sc %>% add_row(convex=convex, skinny=skinny, stringy=stringy,
+                             t=i, method=method, alpha=alpha)
+      }
+    }
+  }
+
+  return(sc)
+}
+
+if(!file.exists("cache/spiralScagSmoothing.rda")){
+  scagSmooth <- getScagSmooth(spiral1000, as.list(t1full))
+  scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
+  save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")
+} else {
+  load("cache/spiralScagSmoothing.rda")
+}
+
+#getting non-smoothed result
+load("cache/V1V2toV3V4.rda")
+noSmoothing <- fullRes %>%
+  filter(size==1000) %>%
+  filter(name=="spiral") %>%
+  select(skinny, stringy, t) %>%
+  gather(variable, value, -t)
+
+scagSmoothMelt <- scagSmoothMelt %>%
+  filter(variable != "convex") %>%
+  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
+ggplot(scagSmoothMelt, aes(x=t, y=value)) +
+  geom_line(data=noSmoothing, color="red", alpha=0.5) +
+  geom_line(aes(linetype=factor(method))) + 
+  theme(legend.position="none") +
+  ylim(0,1) +
+  facet_grid(alpha~variable) +
+  xlab("Sequence of projections (t)") +
+  ylab("PPI value")
+
+```
+
 ## Binning sensitivity of MIC index

### 2019-01-08 scagnostics.Rmd
Small fixes in Sec 3.6
Some comments added to Sec 4.2

@@
 In addition we observed good squintability for all three examples considered, given large datasets.
-Results in Section \ref{sec:params} suggest that tuning of the MINE parameters may result in improved behavior for smaller datasets.
+Results in the appendix \ref{sec:params} suggest that tuning of the MINE parameters may result in improved behavior for smaller datasets.
 
@@
 Convex & $\cdot$ & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark \\
-Skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark \\
+Skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\times$ & \checkmark \\
 Stringy & $\times$ & $\times$ & $\cdot$ & $\cdot$ & \checkmark \\
@@
 
-The first example shows 2538 posterior samples obtained by fitting source parameters to the observed gravitational wave signal GW170817 from a neutron star merger [@PhysRevLett119161101].
+The first example contains 2538 posterior samples obtained by fitting source parameters to the observed gravitational wave signal GW170817 from a neutron star merger [@PhysRevLett119161101].
 Data has been downloaded from @ligoData.
@@
 
-The second example shows data from @Smith:2016qas, generated from a simulation study of a binary black hole (BBH) merger event.
+The second example contains data generated from a simulation study of a binary black hole (BBH) merger event according to @Smith:2016qa.
 (XXX More info needed.) There are 12 parameters (12-D), with multiple nuisance parameters.
@@
 Projection f has the highest PP index value but it is primarily the view seen in the bivariate plot of dec and ra.
+\textcolor{red}{Actually I don't think any of the obtained projections is particularly interesting on its own, but a good point to make here is that these are local maxima and moving between them should be revealing.
+E.g.
+I looked at a planned tour moving from views a-c-e-f which shows relation of multimodality in the high-D space, so maybe that is the point to make here?}
 
@@
 (Does it mean that knowing the precision angle gives more precise information about the location?
-Knowing ra uniquely determines dec-alpha rather than dec it seems.)}
+Knowing ra uniquely determines dec-alpha rather than dec it seems.) Also highlighting why finding global maxima is important, ra, dec and alpha were present in the above example but the optimiser did not converge ot the the global maximum, thus the refined relationship was not revealed.}
 
@@
 
+```{r eval=FALSE}
+m3 <- matrix(c(1,0,0,0,1,0,0,0,1,0,0,0,1,0),ncol=2)
+m4 <- matrix(c(1,1,1,0,0,0,0,0,1,1,0,0,0,0),ncol=2)
+planeL <- list(m3, m4)
+li <- 3
+for(i in c(1, 3, 6, 5)){
+  r <- resList[[i]]
+  iLast <- length(r)
+  fProj <- r[[iLast]]
+  planeL[[li]] <- fProj
+  li <- li+1
+}
+tP <- tourr::planned_tour(planeL[3:6]) # only need dummy planes when using history tour
+dPoints <- attributes(r)$data
+quartz()
+animate_xy(dPoints, tP, pch=".")
+```
+
 ```{r constructedExample, results="hide", fig.cap="Selecting subset of 6 parameters, for which spline index can be used to improve constraint on parameter combination wrt ra vs dec.", dev = "png", dpi=300}

### 2019-01-08 scagnostics.Rmd
Some rewording of the discussion.

@@
 In this context we have investigated the use of indexes developed for variable selection as projection pursuit indexes.
-Having outlined a set of criteria that define a good index function, we have established methods for the evaluation of the index properties: smoothness, squintability, flexibility, rotation invariance and speed.
-To this end we used three different sets of simulated data, defined to have different underlying noise distributions (uniform or normal) and different structures (circle, sine and spiral), and we have evaluated several index functions proposed in the literature for variable selection (scagnostics indexes, maximum information based indexes, distance correlation and a spline based index).
+The set of relevant criteria that define a good index function includes: smoothness, squintability, flexibility, rotation invariance and speed.
+To evaluate these criteria we used three different sets of simulated data, defined to have different underlying noise distributions (uniform or normal) and different structures (circle, sine and spiral which can be considered as large, medium and fine structure respectively).
+The properties were assessed by a list of procedures introduced in Section (XXX cross reference?) and we have evaluated several index functions proposed in the literature for variable selection: scagnostics indexes, maximum information based indexes, distance correlation and a spline based index.
 While all index functions are found to fall short on at least some of the criteria, we have shown that combined with the optimisation available via the guided tour they can already be used to identify interesting views, in particular the built-in structures in the simulated data sets.
-
-We have also sketched some potential improvements of the indexes considered, in particular tuning parameters inherent to the index definition, such as binning, or different smoothing methods for the scagnostics indexes.
 
@@
 
-We conclude that these methods are promissing and may yield new insight in the inter-dependence of model parameters in the light of experimental data ... other applications can also be envisioned??
-
-There are three frontiers for future work building on our study.
+The results are encouraging, showing large potential for discovering unanticipated relations between multiple variables.
+On the other hand there is much room for improvement on several fronts, as identified in this work.
+
+Concretely we see three frontiers for future work building on our study.
 First, while we found the considered index functions useful to detect structures, several shortcomings have also been pointed out.
@@
 for the identification of structures with small squint angle in high dimensions.
+
+As a final remark we point out that small changes to the implementation of the guided tour algorithm would already be helpful, and we plan on implementing the following: allowing for a stopping condition based on index value, and making the step size for the selection of the most promising direction an argument when calling the geodesic optimiser.
 

### 2019-01-09 scagnostics.Rmd
Render with bookdown for appendix settings.

@@
 
+\appendix
+
 # Appendix {-}
 
-## Re-scaling of holes index
+# Re-scaling of holes index
 
@@
 
-## Extimating the squint angle of the spiral
+# Extimating the squint angle of the spiral
 
@@
 
-## Computational performance
+# Computational performance
 
@@
 
-## Effect of parameter choice in index value
+# Effect of parameter choice in index value
 \label{sec:params}
@@
 
-## Ways to refine the PPIs
+# Ways to refine the PPIs
 
@@
 
-## Binning sensitivity of MIC index
+# Binning sensitivity of MIC index
 To examine the sensitivity of binning in the MIC PPI, the classic RANDU data [@Marsaglia68], available in R, is used.
@@
 
-## Additional Figures
-
-### Final projection of pipe guided tour
+# Additional Figures
+
+## Final projection of pipe guided tour
 Figure \ref{fig:testpipe} shows the projection returned during the scouting phase (left) and the refinement phase (right).
@@
 
-### Dataset overview scatter plot matrices
+## Dataset overview scatter plot matrices
 

### 2019-01-09 build.R

@@ -1,0 +1,2 @@
+bookdown::render_book("scagnostics.Rmd")
+

### 2019-01-10 scagnostics.Rmd
spring erfiles

@@
 bibliography: bibliography.bib
-output: rticles::asa_article
+output: rticles::springer_article
 header-includes:

### 2019-01-10 scagnostics.Rmd
Fixed bibliography issues.
Adding short sections summarising results shown in the appendix.

@@
 4. Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
-5.
-Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure.
-The parameters for these PPIs is varied and examined along a tour path between noise and structured projections, to examine the effects of parameter choices and how fiddly the index is to tune.
-Results of this study can be found in the appendix together with preliminary evaluation of potential improvements via smoothing.
-6. Computational speed for the selected indices is examined on a range of sample sizes. Results are presented in the appendix.
+5. Computational speed for the selected indices is examined on a range of sample sizes.
 
@@
 
+## Speed
+Examining the computing time as a function of sample size we find that scagnostics and splines2D are fast even for large samples, while all other index functions slow rapidly with increasing sample size.
+Detailed results are shown in Appendix XXX (cross reference).
+
+## Parameter choices
+Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure.
+In Appendix XXX (cross reference) we examine the dependence on the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
+Sensitivity to the number of bins in the MIC index is also examined, showing that tuning the parameter can improve the final result.
+
+## Index enhancement
+We identified two potential improvements.
+First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in Appendix XXX (cross reference).
+In addition, rotation dependent indexes may be enhanced by redefining them in a rotation invariant way.
+
 ## Summary
@@
 
-The second example contains data generated from a simulation study of a binary black hole (BBH) merger event according to @Smith:2016qa.
+The second example contains data generated from a simulation study of a binary black hole (BBH) merger event according to @Smith:2016qas.
 (XXX More info needed.) There are 12 parameters (12-D), with multiple nuisance parameters.
@@
 ```
+
+# References {-}

### 2019-01-10 scagnostics.Rmd
Shorter summary.

@@
 Our results can be summarized by evaluating and comparing the advantages and disadvantages of each index function according to the criteria presented above.
-An overview is given in Table \ref{tab:summary}.
-The first three entries are from the scagnostics family.
-The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures, making the set flexible in applications (the same is however not true for each single index viewed on its own).
-On the other hand we have found that they are typically too noisy making optimization of the function challenging, which also results in poor squint ability.
-In particular the stringy index cannot be used with current optimizers.
-The index functions from @mbgraphic share similar behavior as they are both found to evolve fairly smoothly between interpolated projections, but have been shown to be strongly rotation dependent.
-The splines2D index is fast to calculate and found to be the most efficient method for identifying functional dependence.
-We have only considered the non-binned implementation of dcor2D which is slow to calculate on large datasets.
-While this index is in principle general, we have found it to be too sensitive to noise to be applicable in most realistic examples, therefore we consider flexibility limited.
-Finally the MINE indexes are found to be smooth only for large datasets, with some rotation dependence observed for the TIC index.
-While these indexes are slow to compute especially on large datasets, we have found them to be the most general, allowing good discrimination even with noisy structures.
-In addition we observed good squintability for all three examples considered, given large datasets.
-Results in the appendix \ref{sec:params} suggest that tuning of the MINE parameters may result in improved behavior for smaller datasets.
+Such an overview is given in Table \ref{tab:summary}, listing if the criteria is fully met (\checkmark), there are some shortcomings ($\cdot$) or failure ($\times$).
+We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fullfilled.
+In addition the limited flexibility of most indices highlights the importance of index selection in the projection pursuit setup.
+Table \ref{tab:summary} further suggests that there is much room for the improvement of index functions detecting unusual association between model parameters.
 
@@
 \hline
-Convex & $\cdot$ & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark \\
+Convex & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark & \checkmark \\
 Skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\times$ & \checkmark \\

### 2019-01-10 scagnostics.Rmd
Cross referencing.

@@
 
-### Property assessment
+### Property assessment {#sec:propAss}
 
@@
 Examining the computing time as a function of sample size we find that scagnostics and splines2D are fast even for large samples, while all other index functions slow rapidly with increasing sample size.
-Detailed results are shown in Appendix XXX (cross reference).
+Detailed results are shown in Appendix \ref{app:speed}.
 
@@
 Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure.
-In Appendix XXX (cross reference) we examine the dependence on the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
+In Appendix \ref{app:param} we examine the dependence on the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
 Sensitivity to the number of bins in the MIC index is also examined, showing that tuning the parameter can improve the final result.
@@
 We identified two potential improvements.
-First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in Appendix XXX (cross reference).
+First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in Appendix \ref{app:enh}.
 In addition, rotation dependent indexes may be enhanced by redefining them in a rotation invariant way.
@@
 To evaluate these criteria we used three different sets of simulated data, defined to have different underlying noise distributions (uniform or normal) and different structures (circle, sine and spiral which can be considered as large, medium and fine structure respectively).
-The properties were assessed by a list of procedures introduced in Section (XXX cross reference?) and we have evaluated several index functions proposed in the literature for variable selection: scagnostics indexes, maximum information based indexes, distance correlation and a spline based index.
+The properties were assessed by a list of procedures introduced in Section \ref{sec:propAss} and we have evaluated several index functions proposed in the literature for variable selection: scagnostics indexes, maximum information based indexes, distance correlation and a spline based index.
 While all index functions are found to fall short on at least some of the criteria, we have shown that combined with the optimisation available via the guided tour they can already be used to identify interesting views, in particular the built-in structures in the simulated data sets.
@@
 
-# Computational performance
-
-\label{sec:timer}
+# Computational performance {#app:speed}
 Computational time is important for using the PPIs with the guided tour, online.
@@
 
-# Effect of parameter choice in index value
-\label{sec:params}
+# Effect of parameter choice in index value {#app:param}
 Some parameters must be provided for some PPIs.
@@
 
-The sensitivity to the binning for calculating the MIC index is examined for one example in the Appendix XXX (add cross reference)
-
-# Ways to refine the PPIs
-
-The biggest issues revealed by the investigation into the new PPIs are a lack of smoothness, particularly for the scagnostics indexes, and the rotation invariance of Grimm's indexes.
-To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighbourhood of projections and average the value, or alternatively average the PPI for several jittered projections.
-This is investigated in Fig.
-\ref{fig:spiralScagSmoothing}.
-Rotation invariance is more difficult to fix, but an alternative tour interpolation method could be useful.
-The geodesic interpolation transitions between planes, and it ignores the basis defining the plane, creating a problem with rotation invariant indices.
-Alternative interpolations based on Givens or Householder rotations could be implemented to transition between bases, which should alleviate the need for rotationally invariant indices.
-
-Two different methods are considered for smoothing the index values:
-
-* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$.
-* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction.
-
-The mean value from a sample of projections is recorded as the PPI value. This could be robsitufied by dropping the most extreme values.
-
-This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
-Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values.
-Both methods appear to be promising in smoothing the function.
-Because the scagnistics are fast to compute, either of these methods is feasible.
-For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.
-
-```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
-While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered.
-For comparison the red line in the background shows the trace without any smoothing applied."}
-
-jitterScagnostics <-function(proj, d, alpha){
-  newProj <- tourr:::basis_nearby(proj, alpha = alpha, method = "geodesic")
-  newD <- d %*% newProj
-  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
-}
-
-jitterPointsScagnostics <-function(projData, alpha){
-  newD <- jitter(projData, amount=alpha)
-  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
-}
-
-getSgnMean <- function(proj, d, alpha, method){
-  dProj <- d %*% proj
-  orig <- as.vector(scagnostics.default(dProj[,1],dProj[,2])$s)
-  if(method == "jitterAngle"){
-    sgnVec <- replicate(10, jitterScagnostics(proj, d, alpha))
-  } else if (method=="jitterPoints"){
-    sgnVec <- replicate(10, jitterPointsScagnostics(dProj, alpha))
-  }
-  return(rowMeans(cbind(orig, sgnVec)))
-}
-
-getScagSmooth <- function(df, tPath){
-  sc <- tibble(
-    convex=numeric(),
-    skinny=numeric(),
-    stringy=numeric(),
-    t=numeric(),
-    method=character(),
-    alpha=numeric())
-  n <- length(tPath)
-  for (method in c("jitterAngle", "jitterPoints")){
-    for (alpha in c(0.01, 0.05, 0.1)){
-      for (i in 1:n) {
-        scagMean <- getSgnMean(tPath[[i]], as.matrix(df), alpha, method)
-        convex <- scagMean[6]
-        skinny <- scagMean[7]
-        stringy <- scagMean[8]
-        sc <- sc %>% add_row(convex=convex, skinny=skinny, stringy=stringy,
-                             t=i, method=method, alpha=alpha)
-      }
-    }
-  }
-
-  return(sc)
-}
-
-if(!file.exists("cache/spiralScagSmoothing.rda")){
-  scagSmooth <- getScagSmooth(spiral1000, as.list(t1full))
-  scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
-  save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")
-} else {
-  load("cache/spiralScagSmoothing.rda")
-}
-
-#getting non-smoothed result
-load("cache/V1V2toV3V4.rda")
-noSmoothing <- fullRes %>%
-  filter(size==1000) %>%
-  filter(name=="spiral") %>%
-  select(skinny, stringy, t) %>%
-  gather(variable, value, -t)
-
-scagSmoothMelt <- scagSmoothMelt %>%
-  filter(variable != "convex") %>%
-  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
-ggplot(scagSmoothMelt, aes(x=t, y=value)) +
-  geom_line(data=noSmoothing, color="red", alpha=0.5) +
-  geom_line(aes(linetype=factor(method))) + 
-  theme(legend.position="none") +
-  ylim(0,1) +
-  facet_grid(alpha~variable) +
-  xlab("Sequence of projections (t)") +
-  ylab("PPI value")
-
-```
-
-# Binning sensitivity of MIC index
+## Binning sensitivity of MIC index
 To examine the sensitivity of binning in the MIC PPI, the classic RANDU data [@Marsaglia68], available in R, is used.
@@
 
+# Ways to refine the PPIs {#app:enh}
+
+The biggest issues revealed by the investigation into the new PPIs are a lack of smoothness, particularly for the scagnostics indexes, and the rotation invariance of Grimm's indexes.
+To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighbourhood of projections and average the value, or alternatively average the PPI for several jittered projections.
+This is investigated in Fig.
+\ref{fig:spiralScagSmoothing}.
+Rotation invariance is more difficult to fix, but an alternative tour interpolation method could be useful.
+The geodesic interpolation transitions between planes, and it ignores the basis defining the plane, creating a problem with rotation invariant indices.
+Alternative interpolations based on Givens or Householder rotations could be implemented to transition between bases, which should alleviate the need for rotationally invariant indices.
+
+Two different methods are considered for smoothing the index values:
+
+* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$.
+* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction.
+
+The mean value from a sample of projections is recorded as the PPI value. This could be robsitufied by dropping the most extreme values.
+
+This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
+Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values.
+Both methods appear to be promising in smoothing the function.
+Because the scagnistics are fast to compute, either of these methods is feasible.
+For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.
+
+```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
+While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered.
+For comparison the red line in the background shows the trace without any smoothing applied."}
+
+jitterScagnostics <-function(proj, d, alpha){
+  newProj <- tourr:::basis_nearby(proj, alpha = alpha, method = "geodesic")
+  newD <- d %*% newProj
+  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
+}
+
+jitterPointsScagnostics <-function(projData, alpha){
+  newD <- jitter(projData, amount=alpha)
+  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
+}
+
+getSgnMean <- function(proj, d, alpha, method){
+  dProj <- d %*% proj
+  orig <- as.vector(scagnostics.default(dProj[,1],dProj[,2])$s)
+  if(method == "jitterAngle"){
+    sgnVec <- replicate(10, jitterScagnostics(proj, d, alpha))
+  } else if (method=="jitterPoints"){
+    sgnVec <- replicate(10, jitterPointsScagnostics(dProj, alpha))
+  }
+  return(rowMeans(cbind(orig, sgnVec)))
+}
+
+getScagSmooth <- function(df, tPath){
+  sc <- tibble(
+    convex=numeric(),
+    skinny=numeric(),
+    stringy=numeric(),
+    t=numeric(),
+    method=character(),
+    alpha=numeric())
+  n <- length(tPath)
+  for (method in c("jitterAngle", "jitterPoints")){
+    for (alpha in c(0.01, 0.05, 0.1)){
+      for (i in 1:n) {
+        scagMean <- getSgnMean(tPath[[i]], as.matrix(df), alpha, method)
+        convex <- scagMean[6]
+        skinny <- scagMean[7]
+        stringy <- scagMean[8]
+        sc <- sc %>% add_row(convex=convex, skinny=skinny, stringy=stringy,
+                             t=i, method=method, alpha=alpha)
+      }
+    }
+  }
+
+  return(sc)
+}
+
+if(!file.exists("cache/spiralScagSmoothing.rda")){
+  scagSmooth <- getScagSmooth(spiral1000, as.list(t1full))
+  scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
+  save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")
+} else {
+  load("cache/spiralScagSmoothing.rda")
+}
+
+#getting non-smoothed result
+load("cache/V1V2toV3V4.rda")
+noSmoothing <- fullRes %>%
+  filter(size==1000) %>%
+  filter(name=="spiral") %>%
+  select(skinny, stringy, t) %>%
+  gather(variable, value, -t)
+
+scagSmoothMelt <- scagSmoothMelt %>%
+  filter(variable != "convex") %>%
+  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
+ggplot(scagSmoothMelt, aes(x=t, y=value)) +
+  geom_line(data=noSmoothing, color="red", alpha=0.5) +
+  geom_line(aes(linetype=factor(method))) + 
+  theme(legend.position="none") +
+  ylim(0,1) +
+  facet_grid(alpha~variable) +
+  xlab("Sequence of projections (t)") +
+  ylab("PPI value")
+
+```
+
 # Additional Figures

### 2019-01-10 scagnostics.Rmd
Footnotes in rmd style rather than latex.

@@
 Less work has been done on indexes to find nonlinear dependence between variables.
-\footnote{\textcolor{red}{The second component in implementing a projection pursuit method is the optimisation algorithm.
+^[\textcolor{red}{The second component in implementing a projection pursuit method is the optimisation algorithm.
 Several algorithms have been suggested, some tailored to specific applications, see e.g.
 XXX references?.
-In this work our focus is on the index functions, and we will only use optimisation algorithms available via the guided tour implementation of tourr.}}
+In this work our focus is on the index functions, and we will only use optimisation algorithms available via the guided tour implementation of tourr.}]
 
@@
 In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], and computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308], round out the collection of metrics.
-\footnote{It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in \cite{huber85,JS87}.}
+^[It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in @huber85 and @JS87.]
 These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
@@
 What we are optimizing over are then the entries of the $p\times2$ dimensional projection matrix, considering that they obey orthonormality conditions.
-Concretely here we consider the following types of index functions:\footnote{We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.}
+Concretely here we consider the following types of index functions:^[We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.]
 

### 2019-01-11 scagnostics.Rmd
as.tibble warning fixed

@@
   dRet <- geozoo::sphere.solid.random(n,p)
-  return(as.tibble(dRet$points))
+  return(as_tibble(dRet$points))
 }
@@
   }
-  return(as.tibble(dRet))
+  return(as_tibble(dRet))
 }
@@
   expr <- paste0(vName,"=sin(",vNameM1,")") # need string expression if I want to use tibble here
-  dRet <- as.tibble(matrix(rnorm((n-1)*p), ncol=(n-1))) #generate normal distributed n-1 dim data
+  dRet <- as_tibble(matrix(rnorm((n-1)*p), ncol=(n-1))) #generate normal distributed n-1 dim data
   dRet <- mutate_(dRet, expr) #string evaluation calculates var(n) as tan(var(n-1))
@@
 }
-anchorIdx <- as.tibble(which(attributes(interpolate(sineTour))$new_basis))
+anchorIdx <- as_tibble(which(attributes(interpolate(sineTour))$new_basis))
 sineTourResMelt <-sineTourRes %>%
@@
     distToSp <- rbind(distToSp, tibble(d=d, distance=distanceToSp(spiralTour, specialPlane), t=1:nBases))
-    fView <- as.tibble(as.matrix(spiralD) %*% as.list(spiralTour)[[nBases]])
+    fView <- as_tibble(as.matrix(spiralD) %*% as.list(spiralTour)[[nBases]])
     fViewL[[i]] <- ggplot(fView, aes(V1, V2)) + geom_point()
@@
 fProj <- nseTourFull[[iLast]]
-dProj <- as.tibble(nsM %*% fProj)
+dProj <- as_tibble(nsM %*% fProj)
 colnames(fProj) <- c("PP1", "PP2")
@@
 fProj <- tL[[iLast]]
-dProj <- as.tibble(nsM2 %*% fProj)
+dProj <- as_tibble(nsM2 %*% fProj)
 colnames(fProj) <- c("PP1", "PP2")
@@
   add_column(label=colnames(nsM2))
-p1 <- ggplot(as.tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.1)+ theme(aspect.ratio=1) + 
+p1 <- ggplot(as_tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.1)+ theme(aspect.ratio=1) + 
   geom_path(data=getCircle(0.5,0.6,0.3,0.18), aes(x=x, y=y), color="grey") + 
@@
   geom_text(data=getAxes(fProj, 0.5,0.6,0.3,0.18), aes(x=x2, y=y2, label=label), size=2.5)
-p2 <- ggplot(as.tibble(nsM2), aes(m1+R1+1.5*R2, L2)) + geom_point(alpha=0.1)+ theme(aspect.ratio=1)
+p2 <- ggplot(as_tibble(nsM2), aes(m1+R1+1.5*R2, L2)) + geom_point(alpha=0.1)+ theme(aspect.ratio=1)
 p3 <- ggplot(nsD, aes(R2/m2, L2))+ geom_point(alpha=0.1)+ theme(aspect.ratio=1)
@@
 fProj1 <- bbhTourFull1[[iLast]]
-dProj1 <- as.tibble(bbhM %*% fProj1)
+dProj1 <- as_tibble(bbhM %*% fProj1)
 colnames(fProj1) <- c("PP1", "PP2")
@@
 fProj2 <- bbhTourFull2[[iLast]]
-dProj2 <- as.tibble(bbhM %*% fProj2)
+dProj2 <- as_tibble(bbhM %*% fProj2)
 colnames(fProj2) <- c("PP1", "PP2")
@@
 fProj3 <- bbhTourFull3[[iLast]]
-dProj3 <- as.tibble(bbhM %*% fProj3)
+dProj3 <- as_tibble(bbhM %*% fProj3)
 colnames(fProj3) <- c("PP1", "PP2")
@@
 fProj1 <- bbhTourFull12[[iLast]]
-dProj <- as.tibble(bbhM2 %*% fProj1)
+dProj <- as_tibble(bbhM2 %*% fProj1)
 colnames(fProj1) <- c("PP1", "PP2")
@@
 fProj2 <- bbhTourFull22[[iLast]]
-dProj <- as.tibble(bbhM2 %*% fProj2)
+dProj <- as_tibble(bbhM2 %*% fProj2)
 colnames(fProj2) <- c("PP1", "PP2")
@@
 fProj3 <- bbhTourFull32[[iLast]]
-dProj <- as.tibble(bbhM2 %*% fProj3)
+dProj <- as_tibble(bbhM2 %*% fProj3)
 colnames(fProj3) <- c("PP1", "PP2")
@@
   fProj <- r[[iLast]]
-  dProj <- as.tibble(bbhM3 %*% fProj)
+  dProj <- as_tibble(bbhM3 %*% fProj)
   colnames(fProj) <- c("PP1", "PP2")
@@
   add_column(label=colnames(bbhC))
-p1 <- ggplot(as.tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)+ 
+p1 <- ggplot(as_tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)+ 
   geom_path(data=getCircle(-0.7, 0.6, 0.2, 0.24), aes(x=x, y=y), color="grey") + 
@@
   geom_text(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
-p2 <- ggplot(as.tibble(bbhD),aes(dec-alpha, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
-p3 <- ggplot(as.tibble(bbhD),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+p2 <- ggplot(as_tibble(bbhD),aes(dec-alpha, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
+p3 <- ggplot(as_tibble(bbhD),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
 grid.arrange(p1, p2, p3, ncol=3)
@@
 fProj <- musicTourFull[[iLast]]
-dProj <- as.tibble(as.matrix(music) %*% fProj)
+dProj <- as_tibble(as.matrix(music) %*% fProj)
 ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
@@
     imax <- length(as.list(tP))
-    fV <- as.tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
+    fV <- as_tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
     colnames(fV) <- c("PP1", "PP2")
@@
 fProj <- pipeTourFull[[iLast]]
-dProj <- as.tibble(pipeResc %*% fProj)
+dProj <- as_tibble(pipeResc %*% fProj)
 p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
@@
 fProj <- pipeTourFull2[[iLast]]
-dProj <- as.tibble(pipeResc %*% fProj)
+dProj <- as_tibble(pipeResc %*% fProj)
 p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +

### 2019-01-11 scagnostics.Rmd
reworked intro

@@
 
-A large number of projection pursuit indexes have been developed, to detect departure from multivariate normal, which includes clusters or outliers or separations between known groups (e.g.
+A large number of projection pursuit indexes have been developed, primarily based on departure from normality, which includes clusters or outliers, and also for finding separations between known groups (e.g.
 @f87, @CBC92, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437).
-Less work has been done on indexes to find nonlinear dependence between variables.
+Less work has been done on indexes to find nonlinear dependence between variables, when $d>1$.
+
+<!-- check Perisic and Posse (2012) and recent paper by Austrian guys -->
+
+The motivation is based on physics applications, to aid the interpretation of model fits on experimental results.
+A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated.
+Different model fits ($n$) yield the data to be examined using projection pursuit.
+(Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.) A single prediction can be a complicated function of all of the free parameters, and typically $q \in [100,1000]$ and $p \sim 10$.
+Current practice is to examine pairs of parameters, or combinations produced by intuition or prior knowledge.
+This begs the question, whether important nonlinear associations are missed because tools to search are not available.
+
+To find appropriate projections pursuit indexes, literature on variable selection was examined.
+With high-dimensional data, even plotting all pairs of variables can lead to many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs.
+There are eight scagnostics, descriptively named: "outlying", "skewed", "sparse", "clumpy", "striated", "convex", "skinny" and "stringy".
+The question is whether these can be adapted into projection pursuit indexes, to search for unusual features in two-dimensional projections of high-dimensional data.
+Recent PhD research by @Grimm2016 explored the behavior of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation.
+In addition, two more indexes for measuring dependence have been proposed in the machine learning literature, based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], with computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308].
+These are related to original 1D projection pursuit indexes based on entropy (e.g.
+@huber85, @JS87).
+This provides 12 current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+The projection pursuit guided tour is available in the R package, tourr [@tourr] provides optimisation routines, and visualization.
+
+<!--
+^[It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in @huber85 and @JS87.]
+These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+The projection pursuit guided tour is available in the R package, tourr [@tourr].
+
 ^[\textcolor{red}{The second component in implementing a projection pursuit method is the optimisation algorithm.
@@
 In this work our focus is on the index functions, and we will only use optimisation algorithms available via the guided tour implementation of tourr.}]
-
-<!-- check Perisic and Posse (2012) and recent paper by Austrian guys -->
-
-Our motivation is based on applications in physics, to aid the interpretation of model fits on experimental results.
-Consider a physical model with a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $n$ experimental observations, for which predictions can be calculated when fixing all $p$ parameters.
-Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.
-Moreover, a single prediction can be a complicated function of all free parameters.
-In addition, we often deal with large number of observations, $n \sim 100-1000$ as well as sizable parameter spaces $p \sim 10$.
-The results of a fit are generally interpreted using combinations of variables selected by intuition and prior knowledge.
-But this begs the question, whether important associations are missed because tools to search are not available.
-
-We draw on developments from the related topic of variable selection.
-With high-dimensional data, even plotting all pairs can be daunting, so "scagnostics" [@scag] [@WW08] have been developed to find the most interesting pairs of variables.
-There are 8 scagnostics, vividly named: "Outlying", "Skewed", "Sparse", "Clumpy", "Striated", "Convex", "Skinny" and "Stringy".
-Our question is whether these can be extended into projection pursuit indexes that can be used to find two-dimensional projections of high-dimensional data with unusual features.
-
-[@Grimm2016] explored the behavior of scagnostics for selecting variables, and proposed two more, based on smoothing splines and distance correlation, that have some nicer properties.
-In addition the availability of another two indices based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], and computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308], round out the collection of metrics.
-^[It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in @huber85 and @JS87.]
-These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
-The projection pursuit guided tour is available in the R package, tourr [@tourr].
+-->
 
@@
 A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data distribution.
-Typically the definition is such that larger values of $f$ indicate more interesting distribution, and therefore maximizing $f$ over all possible $d$ dimensional projections of a data set with $p>d$ parameters will find the most interesting projections.
+Typically the definition is such that larger values of $f$ indicate a more interesting distribution of observations, and therefore maximizing $f$ over all possible $d$ dimensional projections of a data set with $p>d$ parameters will find the most interesting projections.
 As most indices characterize deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.

### 2019-01-11 scagnostics.Rmd
tweaks to last paragraph of intro

@@
 
-The paper investigates the behavior of these newly-defined indexes.
+The paper investigates the behavior of these indexes.
 Section \ref{sec:construct} discusses index construction, and how they fit into the guided tour.
-Section \ref{sec:investigate} investigates the behavior of the indexes, particularly in relation to optimization.
-The new guided tour methods are applied to an example where the posterior distributions from physics models (Section \ref{sec:phys}) are explored.
+Section \ref{sec:investigate} investigates the behavior of the indexes, explored primarily using tour methods (@As85, @BCAH05, @CBCH94).
+The new guided tour with these indexes are applied to two examples from physics models (Section \ref{sec:phys}).
 

### 2019-01-11 scagnostics.Rmd
Reference added.

@@
 
-A large number of projection pursuit indexes have been developed, primarily based on departure from normality, which includes clusters or outliers, and also for finding separations between known groups (e.g.
-@f87, @CBC92, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437).
+A large number of projection pursuit indexes have been developed, primarily based on departure from normality, which includes clusters, outliers and skewness, and also for finding separations between known groups (e.g.
+@f87, @hall89 @CBC92, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437, @LOPERFIDO201842).
 Less work has been done on indexes to find nonlinear dependence between variables, when $d>1$.

### 2019-01-11 scagnostics.Rmd
More references added.
start of index computation section

--- 2019-01-11 scagnostics.Rmd
@@
 
-The term "projection pursuit" was coined by @FT74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
+The term "projection pursuit" (PP) was coined by @FT74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
 The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $d$-dimensional projection of $p$-dimensional data.
@@
 It is hoped that the resulting solutions reveal low-dimensional structure in the data not found by methods such as principal component analysis.
+Projection pursuit is primarily used for visualization, with projected data always reported as plots.
 
@@
 
-# Constructing a projection pursuit index
+# Projection pursuit index construction and optimization
 \label{sec:construct}
 
-A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data distribution.
-Typically the definition is such that larger values of $f$ indicate a more interesting distribution of observations, and therefore maximizing $f$ over all possible $d$ dimensional projections of a data set with $p>d$ parameters will find the most interesting projections.
-As most indices characterize deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
+A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data set, computed by taking a $d$-dimensional projection of an $n\times p$ data matrix.
+Typically the definition is such that larger values of $f$ indicate a more interesting distribution of observations, and therefore maximizing $f$ over all possible $d$ dimensional projections of a data set with $p>d$ variables will find the most interesting projections.
+This section describes the 12 indexes that are to be used to explore bivariate association.
+Some data pre-processing, including standardization, is advisable, prior to optimizing the PPI.
+
+<!-- As most indices characterize deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
 @CBC93.
 When departing from the idea of characterizing differences to the normal distribution it may however be preferred to work directly with the empirical data distribution and instead characterize it by measures based on e.g.
-the minimal spanning tree or the mutual information.
-
-## Standardization
+the minimal spanning tree or the mutual information.-->
+
+## Scaling and standardization
+
+Making a plot always involves some choice of scaling.
+When a scatterplot is made, effectively, albeit under the hood, the data is scaled into a range of $[a, b]$ (often $a=0, b=1$) on both axes to print it on a page or display in a graphics device window.
+The range deliminates page space within which to draw.
+The upshot is that the original data scale is standardized to the range and aspect ratio on the display space.
+It maybe that the original range of X is $[1, 1000]$ and Y is $[1, 1.6]$ but the display linearly warps this to $[0, 1]$ and $[0, 1]$, giving both variables equal visual weight.
+
+With high dimensional data, and particularly projections, it is also necessary to re-scale the original range, and it is important to pay attention to what is conventional, or possible, and the effects.
+The PPIs also may require specific scaling for them to be effectively computed.
+Both of these are addressed here.
+The common pre-processing include:
+
+- Standardizing each variable, to have mean 0 and variance 1, so that individual variable scales do not affect the result.
+Different variable scales are examinable without resorting to projection pursuit, so can be handled prior to searching through high dimensions.
+- Sphering the high-dimensional data is often done to remove linear dependence.
+This is typically done using principal component analysis, and using the principal components as the variables passed to PP.
+If linear dependence is the only structure PP is not needed, and thus this is removed before PP so that the PPIs are not distracted by simple structure.
+- Transform single variables to reduce skewness.
+It is marginal structure, visible in a single variable, which doesn't need a multivariate technique to reveal.
+Skewed distributions will inadvertently affect the PPIs, distracting the search for dependence.
+- Remove outliers, which may be an iterative process, to discover, identify and delete. Extreme values will likely affect PPI performance. Outliers can be examined on a case by case basis later.
+- Possibly remove noise dimensions, which is also likely to be an iterative process.
+Directions where the distribution is purely noise make optimization of a PPI more difficult.
+If a variable is suspected to have little structure and relationship with other variables, conducting PP on the subset of variables without them may improve the efficiency of the search.
+- Centering and scaling of the projected data, can be helpful visually.
+If the data has a small amount of non-normal distribution in some directions, the projected data can appear to wander around the plot window during a tour.
+It doesn't matter what the center of the projected data is, so centering removes a wandering scatterplot.
+Less commonly, it may be useful to scale the projected data to standard values, which would be done to remove any linear dependence remaining in the data.
+
+<!--
 Lower dimensional projections are highly sensitive to standardization performed on the original distribution.
@@
 Overall we are interested in strong association only, if these structures were affected by the standardization procedure, it would most likely be unwelcome effects?}
-
-## Optimization
-Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
-One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
-A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimization for the identification of the maximum.
-On the other hand the global maximum may not be the only projection of interest, and one may want to observe the selected views in the context of the full distribution rather than a static view.
-These issues are addressed when combining projection pursuit with the grand tour [@CBCH94].
-In this case the properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshooting and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
-A possible algorithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
-In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
-
-\textcolor{red}{\cite{posse95b} discusses the optimization, in particular that for most index functions and optimists results are too local, largely dependent on starting point.
-Fiedmans suggestion of hybrid stepping search + steepest-ascent still remains fairly local.
-\cite{posse90} suggested random search instead, which seems to perform better, but did not find pdf of that article, so not so sure what he actually does.}
+-->
+
+## Notation
 
@@
 Don't know how to derive this though, should depend also on c parameter and how the algorithm is approximating the true matrix?.}
+
+## Optimization
+Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
+One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
+A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimization for the identification of the maximum.
+On the other hand the global maximum may not be the only projection of interest, and one may want to observe the selected views in the context of the full distribution rather than a static view.
+These issues are addressed when combining projection pursuit with the grand tour [@CBCH94].
+In this case the properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshooting and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
+A possible algorithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
+In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
+
+\textcolor{red}{\cite{posse95b} discusses the optimization, in particular that for most index functions and optimists results are too local, largely dependent on starting point.
+Fiedmans suggestion of hybrid stepping search + steepest-ascent still remains fairly local.
+\cite{posse90} suggested random search instead, which seems to perform better, but did not find pdf of that article, so not so sure what he actually does.}
 

### 2019-01-11 scagnostics.Rmd

@@
 
-A large number of projection pursuit indexes have been developed, primarily based on departure from normality, which includes clusters or outliers, and also for finding separations between known groups (e.g.
-@f87, @CBC92, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437).
+A large number of projection pursuit indexes have been developed, primarily based on departure from normality, which includes clusters, outliers and skewness, and also for finding separations between known groups (e.g.
+@f87, @hall89 @CBC92, @naito1997, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437, @PAN2000153, @FER13, @LOPERFIDO201842).
 Less work has been done on indexes to find nonlinear dependence between variables, when $d>1$.

### 2019-01-14 scagnostics.Rmd
notation table added

@@
   - \usepackage{xcolor}
+  - \usepackage{bm}
 ---
@@
 With high-dimensional data, even plotting all pairs of variables can lead to many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs.
-There are eight scagnostics, descriptively named: "outlying", "skewed", "sparse", "clumpy", "striated", "convex", "skinny" and "stringy".
+There are eight scagnostics, of which three ("convex", "skinny" and "stringy") are used here.
 The question is whether these can be adapted into projection pursuit indexes, to search for unusual features in two-dimensional projections of high-dimensional data.
@@
 @huber85, @JS87).
-This provides 12 current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
 The projection pursuit guided tour is available in the R package, tourr [@tourr] provides optimisation routines, and visualization.
@@
 Typically the definition is such that larger values of $f$ indicate a more interesting distribution of observations, and therefore maximizing $f$ over all possible $d$ dimensional projections of a data set with $p>d$ variables will find the most interesting projections.
-This section describes the 12 indexes that are to be used to explore bivariate association.
+This section describes the seven indexes that are to be used to explore bivariate association.
 Some data pre-processing, including standardization, is advisable, prior to optimizing the PPI.
@@
 The upshot is that the original data scale is standardized to the range and aspect ratio on the display space.
-It maybe that the original range of X is $[1, 1000]$ and Y is $[1, 1.6]$ but the display linearly warps this to $[0, 1]$ and $[0, 1]$, giving both variables equal visual weight.
+It maybe that the original range of one variable is $[1, 1000]$ and the other is $[1, 1.6]$ but the display linearly warps this to $[0, 1]$ and $[0, 1]$, say, giving both variables equal visual weight.
 
@@
 Overall we are interested in strong association only, if these structures were affected by the standardization procedure, it would most likely be unwelcome effects?}
+
+## Notation
 -->
 
-## Notation
-
-## New index functions
+\begin{table}[htp]
+\caption{Summary of notation}
+\centering
+\begin{tabular}{|l|p{8.2cm}|} \hline
+$p$ & be the number of variables in the data. For physics models the obervable space is higher dimension, say $q$, and the data examined is the fitted model space, typically less than 10. \\
+$d (=2) $ & projection dimension. For studying physics models, typically $d=2$ and this is the focus for the index definition. \\
+$n$ & Number of observations, which for the physics models is the number of fitted models being examined and compared. \\
+&\\
+$\bm{X}_{n\times p} = (\bm{X}_1, ..., \bm{X}_p)$ & $n\times p$-dimensional data matrix, where variables $\bm{X}_j$ may be scaled or standardised, and $\bm{X}$ may be sphered\\
+$\bm{Y}_{n\times 2} = (\bm{Y}_1, ..., \bm{Y}_2)$ & projected data matrix. where $Y_j = (\alpha_1 X_1, ..., \alpha_p X_p$ \\
+$\bm{F}_{p\times 2} = (\bm{\alpha}_1, ..., \bm{\alpha}_2)$ & orthonomal projection matrix\\ 
+&\\
+$H$ & Convex hull \\
+$A$ & Alpha hull \\\hline
+\end{tabular}
+\label{notation}
+\end{table}
+
+## New projection pursuit index functions
 \label{sec:indexDef}
+
+Table \ref{notation} summarises the notation used for this section.
+This an overview of the functions that are converted into projection pursuit indexes.
+Full details of the functions can be found inthe original sources.
+
+<!--
 Focusing now on $d=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
@@
 Concretely here we consider the following types of index functions:^[We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.]
+-->
 

### 2019-01-15 scagnostics.Rmd
dcor explanation changed

@@
 The performance of these indices is examined with respect to ideal behaviour, using simulated data, and then applied to several physics problems.
-The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphics.
+The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphic.
 
@@
 This an overview of the functions that are converted into projection pursuit indexes.
-Full details of the functions can be found inthe original sources.
+Full details of the functions can be found in the original sources.
 
@@
 
-* From the scagnostics family select those best suited to detect shapes, with definitions given below. They are defined to take values between [0,1].
-    + Convex: The convex measure is defined as $c_{convex}= \frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
-This is the only measure where interesting projections will take low values and one may either minimize the index function or maximize $1-c_{convex}$.
-    + Skinny: The skinnyness of a polygon can be measured as the ratio of the perimeter to the area.
-We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $c_{skinny} = 0$ for a full circle, and values close to one or skinny polygons.
-    + Stringy: The stringy index is a measure of the branching structure of the minimal spanning tree (MST), $c_{stringy} = \frac{diameter(MST)}{length(MST)}$ where the diameter is the longest connected path, and the length is the total length (sum of all edges), i.e.
-$c_{stringy}=1$ if the MST contains no branches.
-
-* New index functions available in @mbgraphic, defined to fall between [0,1]:
-    + Distance correlation defined in @szekely2007: The definition is based on empiric measures of distance covariance and variance defined on the dataset 
+* **scagnostics**: The first step to computing the scagnostics is that the bivariate data is binned, and scaled between [0,1] for calculations.
+The convex and alpha hulls, and the minimal spanning tree (MST), are computed.
+    + *convex*: The ratio of the area of alpha to convenx hull, $I_{\bm{F}, convex}= \frac{area(A(Y))}{area(H(Y))}$.
+In this is the only measure where interesting projections will take low values, with a maximum of 1 if both areas are the same.
+Thus $1-c_{convex}$ is used.
+    + *skinny*: The ratio of the perimeter to the area of the alpha hull, $I_{\bm{F}, skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $c_{skinny} = 0$ for a full circle.
+Values close to 1 indicate a skinny polygon.
+    + *stringy*: Based on the MST, $I_{\bm{F}, stringy} = \frac{diameter(MST)}{length(MST)}$ where the diameter is the longest connected path, and the length is the total length (sum of all edges).
+If the MST contains no branches $I_{\bm{F}, skinny} = 1$.
+
+* **association**: New index functions available in @mbgraphic are defined to fall between [0,1]:
+    + *dcor2D*: This function is based on distance correlation [@szekely2007], which is designed to find both linear and nonlinear dependencies between variables.
+It involves computing the distances between pairs of observations, conducting an analysis of variance type breakdown of the distances relative to each variable, and the result is then passed to the usual covariance and hence correlation formula.
+The function ***wdcor*** in the R package, extracat [@extracat], computes the statistic, and the mbgraphic package utilises this function.
+There is a possibility to bin the data before computing the statistic, which could manage computational performance with large data sets.
+    + *splines2D*: iven a spline model of the data distribution that is obtained via the R implantation in the \texttt{gam} function in the mgcv R package [@w16], we define the index based on the variance of the residuals:
     \begin{equation}
-(X,Y) = \{(x_1,y_1), ..., (x_n, y_n) \in \mathbb{R}^p \times \mathbb{R}^q\}
+    splines_{2d} = max(1- \frac{Var(res_{X\sim Y})}{Var(X)}, 1-\frac{Var(res_{Y\sim X})}{Var(Y)}),
     \end{equation}
- as
+    which takes large values if the distribution is well described by the spline model, indicating functional dependence.
+<!--
+\begin{equation}
+ I_{(\bm{F}, dcor2D)} = \begin{cases}
+ \frac{Cov(F_1,F_2)}{\sqrt{Var(F_1)Var(F_2)}} & Var(F_1)Var(F_2) > 0\\
+ 0 & Var(F_1)Var(F_2) = 0
+ \end{cases}
+ \end{equation}
  \begin{equation}
@@
  and reads
- \begin{equation}
- dCor(X,Y) = \begin{cases}
- \frac{Cov(X,Y)}{\sqrt{Var(X)Var(Y)}} & Var(X)Var(Y) > 0\\
- 0 & Var(X)Var(Y) = 0
- \end{cases}
- \end{equation}
-    + Spline based measure: Given a spline model of the data distribution that is obtained via the R implantation in the \texttt{gam} function in the mgcv R package [@w16], we define the index based on the variance of the residuals:
-    \begin{equation}
-    splines_{2d} = max(1- \frac{Var(res_{X\sim Y})}{Var(X)}, 1-\frac{Var(res_{Y\sim X})}{Var(Y)}),
-    \end{equation}
-    which takes large values if the distribution is well described by the spline model, indicating functional dependence.
+ -->
 

### 2019-01-15 scagnostics.Rmd
splines explanation changed

@@
   - \usepackage{longtable}
+  - \usepackage{amsmath}
   - \usepackage{amssymb}
@@
 Thus $1-c_{convex}$ is used.
-    + *skinny*: The ratio of the perimeter to the area of the alpha hull, $I_{\bm{F}, skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $c_{skinny} = 0$ for a full circle.
+    + *skinny*: The ratio of the perimeter to the area of the alpha hull, $I_{\bm{F}, skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $I_{\bm{F}, skinny} = 0$ for a full circle.
 Values close to 1 indicate a skinny polygon.
@@
 
-* **association**: New index functions available in @mbgraphic are defined to fall between [0,1]:
+* **association**: The index functions are available in @mbgraphic, and are defined to range in [0,1].
+Both fucntions in the mbgraphic package can bin the data before computing the index, for computational performance.
     + *dcor2D*: This function is based on distance correlation [@szekely2007], which is designed to find both linear and nonlinear dependencies between variables.
 It involves computing the distances between pairs of observations, conducting an analysis of variance type breakdown of the distances relative to each variable, and the result is then passed to the usual covariance and hence correlation formula.
-The function ***wdcor*** in the R package, extracat [@extracat], computes the statistic, and the mbgraphic package utilises this function.
-There is a possibility to bin the data before computing the statistic, which could manage computational performance with large data sets.
-    + *splines2D*: iven a spline model of the data distribution that is obtained via the R implantation in the \texttt{gam} function in the mgcv R package [@w16], we define the index based on the variance of the residuals:
+The function wdcor, in the R package, extracat [@extracat], computes the statistic, and the mbgraphic package utilises this function.
+    + *splines2D*: Measures nonlinear dependence by fitting a spline model of $\bm{Y}_2$ on $\bm{Y}_1$ and also $\bm{Y}_1$ on $\bm{Y}_2$, using the \texttt{gam} function in the R package, mgcv [@w16].
+The index compares the variance of the residuals:
     \begin{equation}
-    splines_{2d} = max(1- \frac{Var(res_{X\sim Y})}{Var(X)}, 1-\frac{Var(res_{Y\sim X})}{Var(Y)}),
+    I_{\bm{F}, splines2d} = max(1- \frac{Var(res_{\bm{Y}_1\sim \bm{Y}_2)}}{Var(\bm{Y}_1)}, 1-\frac{Var(res_{\bm{Y}_2\sim {\bm{Y}_1})}}{Var(\bm{Y}_2)}),
     \end{equation}
-    which takes large values if the distribution is well described by the spline model, indicating functional dependence.
+    which takes large values if functional dependence is strong.
 <!--
@@
 
-* Information based index functions from @Reshef1518 are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.
+* **Information**:  The index functions [@Reshef1518] measure nonlinear association using a nonparametric computation of information.
+It is strictly a 2D measure, which observations are binned and the relative count of the bin is used to compute entropy.
+are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.
 With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as

### 2019-01-15 scagnostics.Rmd
working on MIC

@@
 
-* **Information**:  The index functions [@Reshef1518] measure nonlinear association using a nonparametric computation of information.
+* **Information**:  The index functions [@Reshef1518] measure nonlinear association by finding an optimal grid to separate empty cells from cells containing observations.
 It is strictly a 2D measure, which observations are binned and the relative count of the bin is used to compute entropy.

### 2019-01-16 scagnostics.Rmd
start on the information indexes

--- 2019-01-15 scagnostics.Rmd
@@
 
-* **Information**:  The index functions [@Reshef1518] measure nonlinear association using a nonparametric computation of information.
-It is strictly a 2D measure, which observations are binned and the relative count of the bin is used to compute entropy.
+* **Information**:  The index functions [@Reshef1518] nonparametricly measure nonlinear association by finding an optimal grid to separate empty cells from cells containing observations.
+It is strictly a 2D measure, where the relative bin counts are used to compute mutual information, $\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p{by_2}))$, where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$ is the relative bin count in each cell, and $p(by_1), p(by_2)$ are the row and column relative counts.
+
 are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.

### 2019-01-16 scagnostics.Rmd
done with the information indexes

--- 2019-01-15 scagnostics.Rmd
@@
 
-* **Information**:  The index functions [@Reshef1518] measure nonlinear association using a nonparametric computation of information.
-It is strictly a 2D measure, which observations are binned and the relative count of the bin is used to compute entropy.
+* **information**:  The index functions [@Reshef1518] nonparametricly measure nonlinear association by computing the mutual information,
+
+\[
+\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p{by_2})),
+\]
+
+where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$ on a range of bin resolutions of the data.
+It is strictly a 2D measure.
+For a fixed binning, e.g.
+$2\times 2$ or $10\times 4$, the optimal binning is found by maximizing $I$.
+The values of I range between $[0,1]$ because they are normalized across bins by dividing by $log min(\# bins_{y_1}, \# bins_{y_2})$.
+
+<!--
+finding an optimal grid to separate empty cells from cells containing observations.
+It is strictly a 2D measure, and the optimum corresponds to the maximum mutual information, $\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p{by_2}))$, where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$ is the relative bin count in each cell, and $p(by_1), p(by_2)$ are the row and column relative counts.
+
 are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.
@@
 We can now define the following metrics:
-    + Maximum Information Coefficient (MIC):
+
     \begin{equation}
@@
     where by default $B(n) = n^{0.6}$
-    + Total Information Coefficient (TIC), see @JMLRv1715308:
+    see @JMLRv1715308:
     \begin{equation}
@@
     \end{equation}
+-->
+    + *Maximum Information Coefficient (MIC)*: uses the maximum normalized $I$ across all bin resolutions. 
+    + *Total Information Coefficient (TIC)*,: sums the normalized $I$ for all bin resolutions computed.
+This creates a problem of scaling - there is no upper limit, although it is related to number of bins, and number of bin resolutions used.
+In the work below we have made empirical estimates of the maximum and scaled the TIC index using this to get it in the range $[0,1]$.
+This index should be more stable than MIC.
+
+<!--
   While $MIC$ is normalized to fall between $[0,1]$, the upper bound of $TIC$ depends on the sample size via the bounding function $B(n)$.
 $TIC$ was designed to be more stable test against null hypothesis of independence, using the full information of the characteristic matrix rather than just the maximum, and as such is expected to be better suited as a PPI.
-
 \textcolor{red}{What is upper bound of TIC as function of sample size?
@@
 Don't know how to derive this though, should depend also on c parameter and how the algorithm is approximating the true matrix?.}
+-->
 

### 2019-01-16 scagnostics.Rmd
formatting fix of MIC/TIC

@@
 * **information**:  The index functions [@Reshef1518] nonparametricly measure nonlinear association by computing the mutual information, 
-
-\[
+\begin{equation}
 \bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p{by_2})),
-\]
-
+\end{equation}
 where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$ on a range of bin resolutions of the data.
@@
 The values of I range between $[0,1]$ because they are normalized across bins by dividing by $log min(\# bins_{y_1}, \# bins_{y_2})$.
+    - *Maximum Information Coefficient (MIC)*: uses the maximum normalized $I$ across all bin resolutions. 
+    - *Total Information Coefficient (TIC)*: sums the normalized $I$ for all bin resolutions computed.
+This creates a problem of scaling - there is no upper limit, although it is related to number of bins, and number of bin resolutions used.
+In the work below we have made empirical estimates of the maximum and scaled the TIC index using this to get it in the range $[0,1]$.
+This index should be more stable than MIC.
 
@@
 -->
-    + *Maximum Information Coefficient (MIC)*: uses the maximum normalized $I$ across all bin resolutions. 
-    + *Total Information Coefficient (TIC)*,: sums the normalized $I$ for all bin resolutions computed.
-This creates a problem of scaling - there is no upper limit, although it is related to number of bins, and number of bin resolutions used.
-In the work below we have made empirical estimates of the maximum and scaled the TIC index using this to get it in the range $[0,1]$.
-This index should be more stable than MIC.
 

### 2019-01-16 scagnostics.Rmd
finished polishing

@@
 Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
-One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of noisy index functions, and @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
+One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of A noisy index function.
+\cite{posse95b} discusses the optimization, in particular that for most index functions and optimal results are too local, largely dependent on starting point.
+@f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
 A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimization for the identification of the maximum.
-On the other hand the global maximum may not be the only projection of interest, and one may want to observe the selected views in the context of the full distribution rather than a static view.
-These issues are addressed when combining projection pursuit with the grand tour [@CBCH94].
-In this case the properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshooting and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
+For exploring high-dimensional data, it can be interesting to observe local maxima as well as a global maximum, and thus a hybrid algorithm that still aloows lingering but not being trapped by local maxima is ideal.
+In addition, being able to visually monitor the optimisation and see the optimal projection in the context of neighbouring projections is useful.
+This is provided by combining projection pursuit with the grand tour [@CBCH94].
+The properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshooting and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
 A possible algorithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
-In addition the tourr package provides the \texttt{search\_geodesic} search function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
-
+The tourr package also provides the \texttt{search\_geodesic} function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
+
+<!--
 \textcolor{red}{\cite{posse95b} discusses the optimization, in particular that for most index functions and optimists results are too local, largely dependent on starting point.
@@
 \cite{posse90} suggested random search instead, which seems to perform better, but did not find pdf of that article, so not so sure what he actually does.}
+
+-->
 
@@
 A useful projection pursuit index needs to have several properties.
-This has been discussed in several seminal papers (\cite{diaconis84}, \cite{huber85}, \cite{JS87}, \cite{posse95b}, \cite{hall89}).
+This has been discussed in several seminal papers, e.g.
+@diaconis84, @huber85, @JS87, @posse95b, @hall89.
 The PPI should be  minimized by the normal distribution, because this is not interesting from a data exploration perspective.

### 2019-01-17 scagnostics.Rmd
done with conclusions

@@
 
+<!--
+Reminder what the paper is about
+summarise main findings
+Future plans
+-->
+
+The motivation for this work was to discover dependencies between estimated parameters in multiple model fits in physics problems.
+This paper shows how projection pursuit with the new indexes can help address this problem.
+The results are encouraging, showing large potential for discovering unanticipated relations between multiple variables.
+
+All of the indexes fall short against some aspect of the ideal properties of smoothness, squintability, flexibility, rotation invariance and speed.
+The paper describes how these properties can be assessed using tour methodology.
+Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
+
+The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools.
+Exploring this area would help improve the guided tours.
+The tour is also lacking a portable, easy to use interface that would enable user interactions like stopping, re-starting, saving optimal projections, and development of a graphical user interface, similar to the tourrGui [@tourrGui] (now archived), would help users and projection pursuit researchers.
+
+<!--
 Physics examples motivate us to investigate projection pursuit indexes that detect complex patterns in two dimensional projections.
@@
 As a final remark we point out that small changes to the implementation of the guided tour algorithm would already be helpful, and we plan on implementing the following: allowing for a stopping condition based on index value, and making the step size for the selection of the most promising direction an argument when calling the geodesic optimiser.
+-->
 

### 2019-01-17 scagnostics.Rmd
Remove most of the red text.

@@
 This is to be expected based on the pairwise scatterplots.
-\textcolor{red}{maybe having tour of those three dimensions would be helpful here}.
 On the other hand, the 1-convex index finds a very different view, but actually the problem is that the optimization doesn't adequately reach a maximum for this index.
@@
 Projection f has the highest PP index value but it is primarily the view seen in the bivariate plot of dec and ra.
-\textcolor{red}{Actually I don't think any of the obtained projections is particularly interesting on its own, but a good point to make here is that these are local maxima and moving between them should be revealing.
-E.g.
-I looked at a planned tour moving from views a-c-e-f which shows relation of multimodality in the high-D space, so maybe that is the point to make here?}
+\textcolor{red}{While none of these projections are particularly interesting on their own, moving between them can be revealing.}
 
@@
 Manually reconstructing this can be done by differencing the two parameters, in their original units.
-\textcolor{red}{This result is unexpected, it can be understood from ?????
-(Does it mean that knowing the precision angle gives more precise information about the location?
-Knowing ra uniquely determines dec-alpha rather than dec it seems.) Also highlighting why finding global maxima is important, ra, dec and alpha were present in the above example but the optimiser did not converge ot the the global maximum, thus the refined relationship was not revealed.}
+\textcolor{red}{This result is highlighting the importance of improved optimisation, that would allow to identify this relation given the full list of input parameters.}
 

### 2019-01-17 scagnostics.Rmd
Spell checks.

@@
 The question is whether these can be adapted into projection pursuit indexes, to search for unusual features in two-dimensional projections of high-dimensional data.
-Recent PhD research by @Grimm2016 explored the behavior of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation.
+Recent PhD research by @Grimm2016 explored the behaviour of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation.
 In addition, two more indexes for measuring dependence have been proposed in the machine learning literature, based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], with computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308].
@@
 
-The paper investigates the behavior of these indexes.
+The paper investigates the behaviour of these indexes.
 Section \ref{sec:construct} discusses index construction, and how they fit into the guided tour.
-Section \ref{sec:investigate} investigates the behavior of the indexes, explored primarily using tour methods (@As85, @BCAH05, @CBCH94).
+Section \ref{sec:investigate} investigates the behaviour of the indexes, explored primarily using tour methods (@As85, @BCAH05, @CBCH94).
 The new guided tour with these indexes are applied to two examples from physics models (Section \ref{sec:phys}).
@@
 If the data has a small amount of non-normal distribution in some directions, the projected data can appear to wander around the plot window during a tour.
-It doesn't matter what the center of the projected data is, so centering removes a wandering scatterplot.
+It doesn't matter what the centre of the projected data is, so centering removes a wandering scatterplot.
 Less commonly, it may be useful to scale the projected data to standard values, which would be done to remove any linear dependence remaining in the data.
@@
 \begin{tabular}{|l|p{8.2cm}|} \hline
-$p$ & be the number of variables in the data. For physics models the obervable space is higher dimension, say $q$, and the data examined is the fitted model space, typically less than 10. \\
+$p$ & be the number of variables in the data. For physics models the observable space is higher dimension, say $q$, and the data examined is the fitted model space, typically less than 10. \\
 $d (=2) $ & projection dimension. For studying physics models, typically $d=2$ and this is the focus for the index definition. \\
@@
 The convex and alpha hulls, and the minimal spanning tree (MST), are computed.
-    + *convex*: The ratio of the area of alpha to convenx hull, $I_{\bm{F}, convex}= \frac{area(A(Y))}{area(H(Y))}$.
+    + *convex*: The ratio of the area of alpha to convex hull, $I_{\bm{F}, convex}= \frac{area(A(Y))}{area(H(Y))}$.
 In this is the only measure where interesting projections will take low values, with a maximum of 1 if both areas are the same.
@@
 A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimization for the identification of the maximum.
-For exploring high-dimensional data, it can be interesting to observe local maxima as well as a global maximum, and thus a hybrid algorithm that still aloows lingering but not being trapped by local maxima is ideal.
+For exploring high-dimensional data, it can be interesting to observe local maxima as well as a global maximum, and thus a hybrid algorithm that still allows lingering but not being trapped by local maxima is ideal.
 In addition, being able to visually monitor the optimisation and see the optimal projection in the context of neighbouring projections is useful.
@@
 The PPI should be  minimized by the normal distribution, because this is not interesting from a data exploration perspective.
-If all projections are normally distributed, good modeling tools already exist.
+If all projections are normally distributed, good modelling tools already exist.
 A PPI should be approximately affine invariant, regardless how the projection is rotated the index value should be the same, and the scale of each variable shouldn't affect the index value.
@@
 
-Three families of data simulations are used for examining the behavior of the index functions.
+Three families of data simulations are used for examining the behaviour of the index functions.
 Each generates  structure in two variables, with the remaining variables containing various types of noise.
@@
 
-Rescaling: first have cutoff at the respective value for the normal distribution, i.e.
+Rescaling: first have cut-off at the respective value for the normal distribution, i.e.
 any value below 0.2 (0.8) is set to zero for cmass (holes) index, rescale remaining range to be between zero and one.
@@
 
-XXX from caption, remove?:  For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detedction being obscured.
+XXX from caption, remove?:  For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detection being obscured.
 With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
@@
 With small $\delta$, less bins, the structure isn't visible, and with larger $\delta$ the structure is confounded with noise.
-It does appear that this parameter affects the performace of the MIC index.
+It does appear that this parameter affects the performance of the MIC index.
 

### 2019-01-17 scagnostics.Rmd
Small fixes in first two sections.

@@
 Here, we are interested in finding projections revealing potentially complex bivariate patterns, using new indexes constructed from scagnostics and a maximum information coefficient, with a purpose to detect unusual relationships between model parameters describing physics phenomena.
-The performance of these indices is examined with respect to ideal behaviour, using simulated data, and then applied to several physics problems.
+The performance of these indexes is examined with respect to ideal behaviour, using simulated data, and then applied to several physics problems.
 The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphic.
@@
 A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated.
-Different model fits ($n$) yield the data to be examined using projection pursuit.
-(Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.) A single prediction can be a complicated function of all of the free parameters, and typically $q \in [100,1000]$ and $p \sim 10$.
+(Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.) Different model fits compatible with the experimental results ($n$) yield the data to be examined using projection pursuit.
+A single prediction can be a complicated function of all of the free parameters, and typically $q \in [100,1000]$ and $p \sim 10$.
 Current practice is to examine pairs of parameters, or combinations produced by intuition or prior knowledge.
@@
 This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
-The projection pursuit guided tour is available in the R package, tourr [@tourr] provides optimisation routines, and visualization.
-
-<!--
-^[It is interesting to note that entropy, closely related to the notion of mutual information, has previously been considered as a projection pursuit index when searching for interesting one-dimensional projections, see discussion in @huber85 and @JS87.]
-These indices are all available in R packages [@rref]: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
-The projection pursuit guided tour is available in the R package, tourr [@tourr].
-
-^[\textcolor{red}{The second component in implementing a projection pursuit method is the optimisation algorithm.
-Several algorithms have been suggested, some tailored to specific applications, see e.g.
-XXX references?.
-In this work our focus is on the index functions, and we will only use optimisation algorithms available via the guided tour implementation of tourr.}]
--->
+The projection pursuit guided tour is available in the R package, tourr [@tourr], and provides optimisation routines, and visualization.
 
@@
 Some data pre-processing, including standardization, is advisable, prior to optimizing the PPI.
-
-<!-- As most indices characterize deviations from a normal distribution, one would first map the empiric data distribution onto a density for which such deviations can then be defined in different ways, see e.g.
-@CBC93.
-When departing from the idea of characterizing differences to the normal distribution it may however be preferred to work directly with the empirical data distribution and instead characterize it by measures based on e.g.
-the minimal spanning tree or the mutual information.-->
 
@@
 
-<!--
-Lower dimensional projections are highly sensitive to standardization performed on the original distribution.
-We want to avoid e.g.
-highlighting directions based only on different scales, or artificially extending the parameter range by including outliers, which may result in seemingly linear behavior for most points.
-Methods to consider include rescaling of individual directions to a common interval (e.g.
-to fall between $[0,1]$), sphering, outlier removal or transformation to a logarithmic scale.
-The method(s) of choice should be informed by the distributions found in the data set as well as aims of the analysis.
-\textcolor{red}{particular challenges: noise only directions, identification of outliers?}
-\textcolor{red}{for holes/cmass index they are only well defined when data is standardized, how about the other index functions?}
-\textcolor{red}{When keeping in mind the concrete example of posterior samples outliers should not be an issue, if extreme outliers are found one should probably investigate origin.
-Also noise only directions, e.g.
-nuisance parameters, are also not an issue as they contain no structure, see BBH example.}
-
-\textcolor{red}{\cite{JS87} discusses centering and sphering of the data points.
-Centering is needed if index is not translational invariant (should not be the case for any index considered here).
-Arguments for sphering are that scale effects should be extracted, we don't want to reproduce information already obtained by PCA.
-To my mind rescaling to an overall interval is more appropriate here.
-Rotations between parameters should not affect the outcome, but will make interpretation more difficult.
-Standardizing the variance will remove information about the posterior variance.
-Overall we are interested in strong association only, if these structures were affected by the standardization procedure, it would most likely be unwelcome effects?}
-
-## Notation
--->
-
 \begin{table}[htp]
@@
 $\bm{X}_{n\times p} = (\bm{X}_1, ..., \bm{X}_p)$ & $n\times p$-dimensional data matrix, where variables $\bm{X}_j$ may be scaled or standardised, and $\bm{X}$ may be sphered\\
-$\bm{Y}_{n\times 2} = (\bm{Y}_1, ..., \bm{Y}_2)$ & projected data matrix. where $Y_j = (\alpha_1 X_1, ..., \alpha_p X_p$ \\
+$\bm{Y}_{n\times 2} = (\bm{Y}_1, ..., \bm{Y}_2)$ & projected data matrix. where $Y_j = (\alpha_1 X_1, ..., \alpha_p X_p)$ \\
 $\bm{F}_{p\times 2} = (\bm{\alpha}_1, ..., \bm{\alpha}_2)$ & orthonomal projection matrix\\ 
@@
 
-<!--
-Focusing now on $d=2$ dimensional projections, we aim to identify projections that reveal interesting structures or correlations between combinations of variables.
-We therefore draw on index functions that have been developed for the selection of variable pairs that show interesting features in a scatter plot, but we generalist the notion such that instead of variable pairs we consider any $2$ dimensional projection of the multivariate data distribution.
-It means that we consider the projection axes as input to the index functions, which will thus score the interestingness of the given projection.
-What we are optimizing over are then the entries of the $p\times2$ dimensional projection matrix, considering that they obey orthonormality conditions.
-Concretely here we consider the following types of index functions:^[We note that all descriptions are idealized, while in practice approximations and binning are used to obtain reasonable computing time, we refer to original references and package documentation for details.]
--->
-
 * **scagnostics**: The first step to computing the scagnostics is that the bivariate data is binned, and scaled between [0,1] for calculations.
@@
     + *convex*: The ratio of the area of alpha to convex hull, $I_{\bm{F}, convex}= \frac{area(A(Y))}{area(H(Y))}$.
-In this is the only measure where interesting projections will take low values, with a maximum of 1 if both areas are the same.
+This is the only measure where interesting projections will take low values, with a maximum of 1 if both areas are the same.
 Thus $1-c_{convex}$ is used.
@@
     \begin{equation}
-    I_{\bm{F}, splines2d} = max(1- \frac{Var(res_{\bm{Y}_1\sim \bm{Y}_2)}}{Var(\bm{Y}_1)}, 1-\frac{Var(res_{\bm{Y}_2\sim {\bm{Y}_1})}}{Var(\bm{Y}_2)}),
+    I_{\bm{F}, splines2d} = max(1- \frac{Var(res_{\bm{Y}_1\sim \bm{Y}_2})}{Var(\bm{Y}_1)}, 1-\frac{Var(res_{\bm{Y}_2\sim {\bm{Y}_1}})}{Var(\bm{Y}_2)}),
     \end{equation}
     which takes large values if functional dependence is strong.
-<!--
-\begin{equation}
- I_{(\bm{F}, dcor2D)} = \begin{cases}
- \frac{Cov(F_1,F_2)}{\sqrt{Var(F_1)Var(F_2)}} & Var(F_1)Var(F_2) > 0\\
- 0 & Var(F_1)Var(F_2) = 0
- \end{cases}
- \end{equation}
- \begin{equation}
- Cov^2(X,Y) = \frac{1}{n^2} \sum_{k,l=1}^{n} A_{kl}B_{kl},\\
- Var^2(X) = \frac{1}{n^2} \sum_{k,l=1}^{n} A_{kl}^2
- \end{equation}
- where $A_{kl}=a_{kl}-\bar{a}_{k.}-\bar{a}_{l.}+\bar{a}_{..}$ with
- \begin{equation}
- a_{kl} = | x_k - x_l|_p,\\
- \bar{a}_{k.} = \frac{1}{n} \sum_{l=1}^n a_{kl},\\
-  \bar{a}_{.l} = \frac{1}{n} \sum_{k=1}^n a_{kl},\\
- \bar{a}_{..} = \frac{1}{n^2}\sum_{k,l=1}^n a_{kl}
- \end{equation}
- and reads
- -->
 
@@
 \begin{equation}
-\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p{by_2})),
+\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p(by_2))),
 \end{equation}
-where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$ on a range of bin resolutions of the data.
+where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$is the relative bin count in each cell, and $p(by_1), p(by_2)$ are the row and column relative counts, on a range of bin resolutions of the data.
 It is strictly a 2D measure.
@@
 $2\times 2$ or $10\times 4$, the optimal binning is found by maximizing $I$.
-The values of I range between $[0,1]$ because they are normalized across bins by dividing by $log min(\# bins_{y_1}, \# bins_{y_2})$.
+The values of I range between $[0,1]$ because they are normalized across bins by dividing by $log(min(\# bins_{y_1}, \# bins_{y_2}))$.
     - *Maximum Information Coefficient (MIC)*: uses the maximum normalized $I$ across all bin resolutions. 
@@
 
-<!--
-finding an optimal grid to separate empty cells from cells containing observations.
-It is strictly a 2D measure, and the optimum corresponds to the maximum mutual information, $\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p{by_2}))$, where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$ is the relative bin count in each cell, and $p(by_1), p(by_2)$ are the row and column relative counts.
-
-are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$.
-With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as
-\begin{equation}
-M(D)_{x,y} = \frac{I^{\ast} (D, x, y)}{\log \min{x,y}},
-\end{equation}
-where the denominator is chosen such that each entry of $M$ will take values between [0,1].
-We can now define the following metrics:
-
-    \begin{equation}
-    MIC(D) = \max_{xy<B(n)}\{M(D)_{x,y}\},
-    \label{eq:MIC}
-    \end{equation}
-    where by default $B(n) = n^{0.6}$
-    see @JMLRv1715308:
-    \begin{equation}
-    TIC(D) = \sum_{xy<B(n)} M(D)_{x,y}
-    \end{equation}
--->
-
-<!--
-  While $MIC$ is normalized to fall between $[0,1]$, the upper bound of $TIC$ depends on the sample size via the bounding function $B(n)$.
-$TIC$ was designed to be more stable test against null hypothesis of independence, using the full information of the characteristic matrix rather than just the maximum, and as such is expected to be better suited as a PPI.
-\textcolor{red}{What is upper bound of TIC as function of sample size?
-It seems from checking by putting in identical parameters as x, y that max(TIC) = 16 for 100 datapoints and max(TIC) = 148 for 1000 datapoints.
-Don't know how to derive this though, should depend also on c parameter and how the algorithm is approximating the true matrix?.}
--->
-
 ## Optimization
 Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
-One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of A noisy index function.
-\cite{posse95b} discusses the optimization, in particular that for most index functions and optimal results are too local, largely dependent on starting point.
+One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of a noisy index function.
+@posse95b discusses the optimization, in particular that for most index functions and optimizers results are too local, largely dependent on starting point.
 @f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.

### 2019-01-17 scagnostics.Rmd
Replace all \cite with @.

@@
 
-<!--
-\textcolor{red}{\cite{posse95b} discusses the optimization, in particular that for most index functions and optimists results are too local, largely dependent on starting point.
-Fiedmans suggestion of hybrid stepping search + steepest-ascent still remains fairly local.
-\cite{posse90} suggested random search instead, which seems to perform better, but did not find pdf of that article, so not so sure what he actually does.}
-
--->
-
 ```{r load}
@@
 A PPI should be approximately affine invariant, regardless how the projection is rotated the index value should be the same, and the scale of each variable shouldn't affect the index value.
-Interestingly the original index proposed by \cite{FT74} was not rotationally invariant.
+Interestingly the original index proposed by @FT74 was not rotationally invariant.
 A consistent index means that small perturbations to the sample do not dramatically change the index value.
 This is particularly important to making optimization feasible, small angles between projections correspond to small perturbations of the sample, and thus should be small changes to index value.
-\cite{posse95b} suggests that indexes should be resistant to features in a tail of the distribution, but this is debatable because one departure from normality that is interesting to detect are anomalous observations.
+@posse95b suggests that indexes should be resistant to features in a tail of the distribution, but this is debatable because one departure from normality that is interesting to detect are anomalous observations.
 Some PPI are designed precisely for these reasons.
@@
 
-- **squintability**: \cite{barnett1981interpreting} introduced the idea of squint angle to indicate resolution of structure in the data.
-Fine structure like the parallel planes in the infamous RANDU data \cite{Marsaglia68} has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
+- **squintability**: @barnett1981interpreting introduced the idea of squint angle to indicate resolution of structure in the data.
+Fine structure like the parallel planes in the infamous RANDU data [@Marsaglia68] has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
 Structures with small squint angle are difficult to find, because the optimization algorithm needs to get very close to begin hill-climbing to the optimum.
@@
 
-Holes and cmass indexes are derived from $I_0^N$ of \cite{CBC93}.
+Holes and cmass indexes are derived from $I_0^N$ of [@CBC93].
 As noted in proposition 1 of that paper the index takes local maxima for the minimum and maximum of $a_0$ which are achieved by central hole and central mass distributions respectively.
-\cite{Cook:2007:IDG:1557227} then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
+@Cook:2007:IDG:1557227 then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
 They are defined such that each one is maximized for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
@@
 
-We can estimate the values found for normal distribution by comparing values of $a_{00}$ of Sec 7.1 in \cite{CBC93}.
+We can estimate the values found for normal distribution by comparing values of $a_{00}$ of Sec 7.1 in [@CBC93].
 The maximum value is $1/(2\pi)$ found for cmass type distributions, the minimum is $1/(2\pi e)$ found for hole type distributions.

### 2019-01-17 scagnostics.Rmd
Fixed formatting of Tab 2 for Springer article format.

@@
       add_header_above(c(" ", " ", "pipe" = 2, "sine" = 2, "spiral"=2)) %>%
-  group_rows("", 1, 2, latex_gap_space = "-0.5cm") %>%
-  group_rows("", 3, 4, latex_gap_space = "-0.5cm") %>%
-  group_rows("", 5, 6, latex_gap_space = "-0.5cm") %>%
-  group_rows("", 7, 8, latex_gap_space = "-0.5cm") %>%
-  group_rows("", 9, 10, latex_gap_space = "-0.5cm") %>%
-  group_rows("", 11, 12, latex_gap_space = "-0.5cm") %>%
-  group_rows("", 13, 14, latex_gap_space = "-0.5cm") %>%
-  group_rows("", 15, 16, latex_gap_space = "-0.5cm") %>% 
+  group_rows("", 1, 2, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 3, 4, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 5, 6, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 7, 8, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 9, 10, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 11, 12, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 13, 14, latex_gap_space = "-0.2cm") %>%
+  group_rows("", 15, 16, latex_gap_space = "-0.2cm") %>% 
   #row_spec(1:16, font_size = 10) %>%
-  kable_styling(position = "center", font_size = 7) 
+  kable_styling(position = "center") #, font_size = 7) 
 ```

### 2019-01-17 scagnostics.Rmd
Small fixes Sec 3.

@@
 1.
-Compute the PPI values on the tour path along an interpolation between pairs of nuisance variables, $x_1 - x_2$ view to $x_3 - x_4$.
+Compute the PPI values on the tour path along an interpolation between pairs of nuisance variables, $x_1 - x_2$ to $x_3 - x_4$.
 The result is ideally a smooth change in low values.
@@
 4. Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
-5. Computational speed for the selected indices is examined on a range of sample sizes.
+5. Computational speed for the selected indexes is examined on a range of sample sizes.
 
@@
 This will make them difficult to optimize.
-TIC also is noisy.
-On the other hand holes, dcor2d, splines2d and MIC are relatively smooth functions.
+On the other hand holes, dcor2d, splines2d, MIC and TIC are relatively smooth functions.
 
@@
 Note that this likely requires large number of tries, especially as dimension increases, since most randomly picked planes will not be interesting.
-The second step is uses method 3 starting from the approximate maximum, which will take small steps to refine the result to be closer to the global maximum.
+The second step uses method 3 starting from the approximate maximum, which will take small steps to refine the result to be closer to the global maximum.
 
@@
 The only complications arise from a lack of rotation invariance of the splines2d index.
-It is not easily visible here, but it is possible that the best projection will have a lower PPI.
+It is not easily visible here, but it is possible that the best projection will have a higher PPI.
 The index changes depending on the basis used to define the plane, but the geodesic interpolation conducted by the tour uses any suitable basis to describe the plane, ignoring that which optimizes the PPI.
@@
 Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima.
-The sine is found relatively easily, by splines2D, and it is indicated that MIC, TIC, dcor2D and convex would also likely find this structure.", results="hide"}
+The sine is found relatively easily, by splines2d, and it is indicated that MIC, TIC, dcor2d and convex would also likely find this structure.", results="hide"}
 set.seed(2018)
@@
 The skinny scagnostic index is used because it was observed (Figure \ref{fig:wIntermediate}) to be sensitive to this structure, although the noisiness of the index might be problematic.
-The stringy appears to be more sensitive to the spiral, but it has a much small squint angle.
+The stringy appears to be more sensitive to the spiral, but it has a much smaller squint angle.
 
@@
 ## Speed
-Examining the computing time as a function of sample size we find that scagnostics and splines2D are fast even for large samples, while all other index functions slow rapidly with increasing sample size.
+Examining the computing time as a function of sample size we find that scagnostics and splines2d are fast even for large samples, while all other index functions slow rapidly with increasing sample size.
 Detailed results are shown in Appendix \ref{app:speed}.
@@
 Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure.
-In Appendix \ref{app:param} we examine the dependence on the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
+In Appendix \ref{app:param} we examine the dependence of the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
 Sensitivity to the number of bins in the MIC index is also examined, showing that tuning the parameter can improve the final result.
@@
 We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fullfilled.
-In addition the limited flexibility of most indices highlights the importance of index selection in the projection pursuit setup.
+In addition the limited flexibility of most indexes highlights the importance of index selection in the projection pursuit setup.
 Table \ref{tab:summary} further suggests that there is much room for the improvement of index functions detecting unusual association between model parameters.
@@
 \caption{Summary of findings, showing to what extend the considered index functions pass the criteria for a good PPI.
-"\checkmark" symbols good behavior, "$\cdot$" symbols some issues while "$\times$" symbols failure on the corresponding criteria.
+``\checkmark" symbols good behavior, ``$\cdot$" symbols some issues while ``$\times$" symbols failure on the corresponding criteria.
 Each index has particular strengths and drawbacks and selection must be guided by the considered example, see text for details.}

### 2019-01-17 scagnostics.Rmd
Small fixes Sec 5.

@@
 The fitting procedures are described in detail in @Abbott2018exr.
-There are six parameters of physical interest (6D) with some known relationships.
+There are six parameters of physical interest (6-D) with some known relationships.
 Projection pursuit is used to find the known relationships.
 
-The second example contains data generated from a simulation study of a binary black hole (BBH) merger event according to @Smith:2016qas.
-(XXX More info needed.) There are 12 parameters (12-D), with multiple nuisance parameters.
-Projection pursuit uncovers several new relationships between parameters.
+The second example contains data generated from a simulation study of a binary black hole (BBH) merger event, as described in @Smith:2016qas.
+There are 12 parameters (12-D), with multiple nuisance parameters.
+Projection pursuit uncovers new relationships between parameters.
 
@@
 Figure \ref{fig:bbhSimulation} in the Appendix shows a scatterplot matrix, of nine of the twelve parameters.
-(Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time.
+(Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they are uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time.
 The first two describe the position of the event in the sky, and time is the merging time (in GPS units).
@@
 Figure \ref{fig:bbhGuided} showing the projections that maximize three of the indexes.
-TIC and splines2D indexes identify very similar projections, that are based on the three parameters, dec, time and ra.
+TIC and splines2d indexes identify very similar projections, that are based on the three parameters, dec, time and ra.
 This is to be expected based on the pairwise scatterplots.
@@
 ```{r bbhGuided, results="hide", fig.cap="Projections corresponding to the maxima of three indices: TIC, splines2D and 1-convex.
-Projections a, b found by TIC and splines3D are very similar, and involve the same three parameters, ra, dec and time.
+Projections a, b found by TIC and splines2d are very similar, and involve the same three parameters, ra, dec and time.
 The 1-convex index finds a very different view.", dev = "png", dpi=300}
@@
 Figure \ref{fig:bbhGuided2} shows the projections which maximize the TIC, splines2D and 1-convex indices.
-The results provide similar information as already learned from the scatterplot matrix (Figure \ref{bbhSimulation}).
-The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2D index.
+The results provide similar information as already learned from the scatterplot matrix (Figure \ref{fig:bbhSimulation}).
+The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2d index.
 The 1-convex index finds nothing interesting.
 
-```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2D and 1-convex.", dev = "png", dpi=300}
+```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2d and 1-convex.", dev = "png", dpi=300}
 bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
@@
   geom_text(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x2, y=y2, label=label), size=2.5) +
-  ggtitle(paste0("b, splines2D: ", toString(format(iv, digits=2))))
+  ggtitle(paste0("b, splines2d: ", toString(format(iv, digits=2))))
 iLast <- length(bbhTourFull32)
@@
 The initial conditions for the optimization, and the subset of variables used, can have a large effect on the projections returned.
-We illustrate this using only the splines2D index, and find that there is one more association that can be learned that was masked earlier.
+We illustrate this using only the splines2d index, and find that there is one more association that can be learned that was masked earlier.
 
@@
 When alpha contributes in contrast to dec, the relationship between the points is almost perfectly on a curve.
-This is shown in Figure \ref{constructedExample}.
+This is shown in Figure \ref{fig:constructedExample}.
 Manually reconstructing this can be done by differencing the two parameters, in their original units.

### 2019-01-17 scagnostics.Rmd
Small fixes in Appendix.

@@
 
-<!--
-Physics examples motivate us to investigate projection pursuit indexes that detect complex patterns in two dimensional projections.
-In this context we have investigated the use of indexes developed for variable selection as projection pursuit indexes.
-The set of relevant criteria that define a good index function includes: smoothness, squintability, flexibility, rotation invariance and speed.
-To evaluate these criteria we used three different sets of simulated data, defined to have different underlying noise distributions (uniform or normal) and different structures (circle, sine and spiral which can be considered as large, medium and fine structure respectively).
-The properties were assessed by a list of procedures introduced in Section \ref{sec:propAss} and we have evaluated several index functions proposed in the literature for variable selection: scagnostics indexes, maximum information based indexes, distance correlation and a spline based index.
-While all index functions are found to fall short on at least some of the criteria, we have shown that combined with the optimisation available via the guided tour they can already be used to identify interesting views, in particular the built-in structures in the simulated data sets.
-
-Our primary interest is in multidimensional physics models, and we have shown two examples from gravitational wave detection.
-In the first example, studying a neutron star merger, we have sucessfully used the developped methods to re-discover the non-linear relation between three of the parameters.
-The second example, stemming from a black hole simulation, has a large parameter space.
-We used this dataset to explore the results obtained depending on parameter selection, index selection and starting conditions.
-Interestingly, when working with a small parameter space (six dimensions), projection pursuit uncovers a new relation that has not been anticipated.
-
-The results are encouraging, showing large potential for discovering unanticipated relations between multiple variables.
-On the other hand there is much room for improvement on several fronts, as identified in this work.
-
-Concretely we see three frontiers for future work building on our study.
-First, while we found the considered index functions useful to detect structures, several shortcomings have also been pointed out.
-While many definitions to identify clusters have been considered, there is much potential for the development of index functions that detect the types of structures considered here.
-Second, an implementation of interpolations from basis to basis, alleviating the need for rotation invariant indexes is envisioned.
-Finally, more generally, improved optimisation algorithms are required, e.g.
-for the identification of structures with small squint angle in high dimensions.
-
-As a final remark we point out that small changes to the implementation of the guided tour algorithm would already be helpful, and we plan on implementing the following: allowing for a stopping condition based on index value, and making the step size for the selection of the most promising direction an argument when calling the geodesic optimiser.
--->
-
 \appendix
@@
 
-Rescaling: first have cut-off at the respective value for the normal distribution, i.e.
-any value below 0.2 (0.8) is set to zero for cmass (holes) index, rescale remaining range to be between zero and one.
+We therefore rescale as follows: first have cut-off at the respective value for the normal distribution, i.e.
+any value below 0.2 (0.8) is set to zero for cmass (holes) index, and we rescale the remaining range to be between zero and one.
 
@@
 
-The results are interesting. The scagnostic indexes and splines2D are very fast regardless of sample size. MIC, TIC (both versions) and dcor2D slow rapidly as sample size increases.
+The results are interesting. The scagnostic indexes and splines2d are very fast regardless of sample size. MIC, TIC (both versions) and dcor2d slow rapidly as sample size increases.
 
@@
     - Scagnostics: the number of bins can be controlled by the user, note however that internally the implementation will reduce the number of bins if too many non-empty bins are found (more than 250).
-    - MINE: the maximum number of bins considered is fixed by the user as a function of the number of data points, see Equation \ref{eq:MIC}.
-The default is chosen such as a trade-off between resolution and noise dependence, but it may be tuned based on requirements dictated by specific datasets.
+    - MINE: the maximum number of bins considered is fixed by the user as a function of the number of data points.
+The default is chosen as a trade-off between resolution and noise dependence, but it may be tuned based on requirements dictated by specific datasets.
 Apart from sensitivity to noise computing time may also be a consideration here.
-- Spline knots: for the splines2D measure we need to fix the number of knots.
+- Spline knots: for the splines2d measure we need to fix the number of knots.
 By default it is fixed to be 10 (or lower if appropriate based on the data values).
@@
 The bins argument for the scagnostics might be reasonably expected to affect the smoothness of the index: a small number of bins should provide a smooth index function, but may affect its ability to detect fine structure.
-Figure \ref{fig:sprialScagBinning} examines this.
+Figure \ref{fig:spiralScagBinning} examines this.
 Scagnostics PPIs are computed for the spiral1000 data on a tour path between $x_1$-$x_2$ to $x_5$-$x_6$, for number of bins equal to 10, 20, 50.
@@
 The problem with the small bin size is that the spiral becomes invisible to the PPI.
-
-XXX from caption, remove?:  For too rough binning (bins=10) we observe random fluctuations from the binning, with structure detection being obscured.
-With bins=20 we see somewhat smoother behaviour for the stringy index, especially when moving into the maximum, while the skinny index shows a dip in the final view, and larger bins=50 is preferred.
-For the convex index we also find bins=50 to be the better choice.
 
@@
   facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
-  scale_colour_discrete("Num bins") + xlab("Number of projections") + 
+  scale_colour_discrete("bins") + xlab("Number of projections") + 
   ylab("PPI value")
@@
 
-* Jittering points: using the jitter function we move each points by a random amount drawn from a uniform distribution between $\pm \beta$.
+* Jittering points: using the jitter function we move each point by a random amount drawn from a uniform distribution between $\pm \beta$.
 * Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction.
@@
 ```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
-While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered.
 For comparison the red line in the background shows the trace without any smoothing applied."}
@@
 
+Figures \ref{fig:neutronStarSPLOM} and \ref{fig:bbhSimulation} show the scatter plot matrices of the gravitational wave datasets considered, see Section \ref{sec:phys} for details.
+
 ```{r neutronStarSPLOM, fig.height=8, fig.width=8, fig.cap="Scatter plot matrix of the neutron star dataset, darker regions represent higher marginalised posterior densities."}

### 2019-01-17 scagnostics.Rmd
Final spell check.

@@
 * **association**: The index functions are available in @mbgraphic, and are defined to range in [0,1].
-Both fucntions in the mbgraphic package can bin the data before computing the index, for computational performance.
+Both functions in the mbgraphic package can bin the data before computing the index, for computational performance.
     + *dcor2D*: This function is based on distance correlation [@szekely2007], which is designed to find both linear and nonlinear dependencies between variables.
@@
 
-For distributions and indexes with smooth behavior and good squintability, method 3 is the most effective method for optimization.
+For distributions and indexes with smooth behaviour and good squintability, method 3 is the most effective method for optimization.
 If these two criteria are not met the method may still be useful, but only given an informed starting projection.
@@
 The uniform distribution was used to generate the noise variables, which has a higher PPI value than a normal distribution, yielding the higher initial value.
-In addition, a local maximum is observed whenever the pair of variables is one structured variable and one noise variable, because there is a lighter density in the center of the projection.
+In addition, a local maximum is observed whenever the pair of variables is one structured variable and one noise variable, because there is a lighter density in the centre of the projection.
 
@@
 
-Rotational invariance is examined using the sine data ($x_5$-$x_6$), computing PPI for different rotations within the 2D plane, parametrized by angle.
+Rotational invariance is examined using the sine data ($x_5$-$x_6$), computing PPI for different rotations within the 2D plane, parameterized by angle.
 Results are shown in Figure \ref{fig:rotationDep}.
@@
 Such an overview is given in Table \ref{tab:summary}, listing if the criteria is fully met (\checkmark), there are some shortcomings ($\cdot$) or failure ($\times$).
-We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fullfilled.
+We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fulfilled.
 In addition the limited flexibility of most indexes highlights the importance of index selection in the projection pursuit setup.
@@
 Both methods appear to be promising in smoothing the function.
-Because the scagnistics are fast to compute, either of these methods is feasible.
+Because the scagnostics are fast to compute, either of these methods is feasible.
 For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.

### 2019-01-17 scagnostics.Rmd
Fixed labels to say 1-convex where appropriate.

@@
   mutate(convex = 1-convex) %>%
+  rename("1-convex"=convex) %>%
   rowwise() %>%
@@
   filter(PPI != "cmass") %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+
 fullResMelt %>% 
@@
   filter(PPI != "cmass") %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+  mutate(PPI = fct_relevel(PPI, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
 fullResMeltPlanned %>% 
@@
   filter(t %in% which(attributes(interpolate(pipeTour))$new_basis)) %>%
-  mutate(variable = fct_relevel(variable, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+  mutate(variable = fct_relevel(variable, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
 
@@
   gather(variable, value, -t, -name, -size) %>%
-  mutate(variable = fct_relevel(variable, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+  mutate(variable = fct_relevel(variable, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+
 ggplot(sineTourResMelt, aes(x=t, y=value)) +
@@
   gather(PPI, value, -angle) %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+
 ggplot(scMelt, aes(x=angle, y=value)) +

### 2019-01-17 scagnostics.Rmd
done with red chunks

@@
 Projections a, b found by TIC and splines2d are very similar, and involve the same three parameters, ra, dec and time.
-The 1-convex index finds a very different view.", dev = "png", dpi=300}
+The 1-convex index finds a very different view.", fig.height = 3.2, dev = "png", dpi=300}
 bbhD <- read_csv("data/posterior_samples.csv")
@@
 
-```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2d and 1-convex.", dev = "png", dpi=300}
+```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2d and 1-convex.", fig.height = 3.2, dev = "png", dpi=300}
 bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
@@
 Projection f has the highest PP index value but it is primarily the view seen in the bivariate plot of dec and ra.
-\textcolor{red}{While none of these projections are particularly interesting on their own, moving between them can be revealing.}
+While none of these projections are particularly interesting on their own, moving between them can be revealing.
 
@@
 This is shown in Figure \ref{fig:constructedExample}.
-Manually reconstructing this can be done by differencing the two parameters, in their original units.
-\textcolor{red}{This result is highlighting the importance of improved optimisation, that would allow to identify this relation given the full list of input parameters.}
+Manually reconstructing the optimal projection (left plot) can be done by differencing the two parameters, in their original units.
+This highlights the importance of improved optimisation, that would use tiny final steps to polish the view to a finer optimal projection and possibly remove noise induced by small contributions of many variables.
 
@@
 
-```{r constructedExample, results="hide", fig.cap="Selecting subset of 6 parameters, for which spline index can be used to improve constraint on parameter combination wrt ra vs dec.", dev = "png", dpi=300}
+```{r constructedExample, results="hide", fig.cap="Manual reconstruction of an optimal projection (left), constructed by differencing alpha from dec in the original units against ra (middle), compared with the two main variables (right).", fig.height = 3.2, dev = "png", dpi=300}
 bbhC <- rescale(select(bbhD, m1, ra, chi_tot, alpha, distance, dec))

### 2019-01-23 scagnostics.Rmd
Typo fixed.
Taking into account comments by German.

@@
   affiliation: Department of Physics, Monash University
-  email: \email{ursula.laa@monash.edu}
+  email: ursula.laa@monash.edu
 
@@
 Here, we are interested in finding projections revealing potentially complex bivariate patterns, using new indexes constructed from scagnostics and a maximum information coefficient, with a purpose to detect unusual relationships between model parameters describing physics phenomena.
-The performance of these indexes is examined with respect to ideal behaviour, using simulated data, and then applied to several physics problems.
+The performance of these indexes is examined with respect to ideal behaviour, using simulated data, and then applied to problems from gravitational wave astronomy.
 The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphic.
@@
 A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated.
-(Note here, that while we often have analytic expressions for the predictions, this is not always the case and we may have to rely on numerical computation.) Different model fits compatible with the experimental results ($n$) yield the data to be examined using projection pursuit.
+(Note here, that while we may have analytic expressions for the predictions, this is not always the case and we often have to rely on numerical computation.) Different sets of model parameters ($n$) found to be  compatible with the experimental results within a selected level of confidence yield the data to be examined using projection pursuit.
 A single prediction can be a complicated function of all of the free parameters, and typically $q \in [100,1000]$ and $p \sim 10$.
@@
 To find appropriate projections pursuit indexes, literature on variable selection was examined.
-With high-dimensional data, even plotting all pairs of variables can lead to many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs.
+With high-dimensional data, even plotting all pairs of variables can lead to too many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs.
 There are eight scagnostics, of which three ("convex", "skinny" and "stringy") are used here.
@@
 
-The paper investigates the behaviour of these indexes.
+This paper investigates the behaviour of these indexes.
 Section \ref{sec:construct} discusses index construction, and how they fit into the guided tour.
 Section \ref{sec:investigate} investigates the behaviour of the indexes, explored primarily using tour methods (@As85, @BCAH05, @CBCH94).
-The new guided tour with these indexes are applied to two examples from physics models (Section \ref{sec:phys}).
+The new guided tour with these indexes is applied to two examples from gravitational wave astronomy (Section \ref{sec:phys}).
 
@@
 The upshot is that the original data scale is standardized to the range and aspect ratio on the display space.
-It maybe that the original range of one variable is $[1, 1000]$ and the other is $[1, 1.6]$ but the display linearly warps this to $[0, 1]$ and $[0, 1]$, say, giving both variables equal visual weight.
+It may be that the original range of one variable is $[1, 1000]$ and the other is $[1, 1.6]$ but the display linearly warps this to $[0, 1]$ and $[0, 1]$, say, giving both variables equal visual weight.
 
@@
 If a variable is suspected to have little structure and relationship with other variables, conducting PP on the subset of variables without them may improve the efficiency of the search.
-- Centering and scaling of the projected data, can be helpful visually.
+- Centring and scaling of the projected data, can be helpful visually.
 If the data has a small amount of non-normal distribution in some directions, the projected data can appear to wander around the plot window during a tour.
-It doesn't matter what the centre of the projected data is, so centering removes a wandering scatterplot.
+It doesn't matter what the centre of the projected data is, so centring removes a wandering scatterplot.
 Less commonly, it may be useful to scale the projected data to standard values, which would be done to remove any linear dependence remaining in the data.
@@
 Table \ref{notation} summarises the notation used for this section.
-This an overview of the functions that are converted into projection pursuit indexes.
+Here we give an overview of the functions that are converted into projection pursuit indexes.
 Full details of the functions can be found in the original sources.
@@
 The lower and upper show the 5th and 95th percentile of values.
-All indexes, except convex show distinctly higher values for the structured projections.
+The holes index is sensitive only to the pipe distribution.
+All other indexes, except convex show distinctly higher values for the structured projections.
 The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs.
@@
 Some indexes have distinct differences in values between sample sizes."}
+namesFixed <- str_replace_all(fullRes$name, "sin", "sine")
 fullResMelt <- fullRes %>%
+  mutate(name = namesFixed) %>%
   mutate(convex = 1-convex) %>%
@@
 intermediatePlane <- which(attributes(interpolate(t3))$new_basis)[2]
+namesFixed <- str_replace_all(fullResPlanned$name, "sin", "sine")
 fullResMeltPlanned <- fullResPlanned %>%
+  mutate(name = namesFixed) %>%
   mutate(convex = 1-convex) %>%
@@
 
-# Application to physics data
+# Application to physics examples
 \label{sec:phys}
@@
 With only five parameters, a reasonable start is to examine the 5D space using a grand tour.
-This quickly shows that the strong nonlinear relationships between the parameters.
+This quickly shows  the strong nonlinear relationships between the parameters.
 PP is then used to extract these relationships.
@@
 
-This data contains posterior samples for simulation from a model describing a binary black hole (BBH) merger event. There are twelve model parameters. Flat priors are used for most model parameters.
+This data contains posterior samples from simulation from a model describing a binary black hole (BBH) merger event. There are twelve model parameters. Flat priors are used for most model parameters.
 
@@
 This is to be expected based on the pairwise scatterplots.
-On the other hand, the 1-convex index finds a very different view, but actually the problem is that the optimization doesn't adequately reach a maximum for this index.
+On the other hand, the 1-convex index finds a very different view, but this is because the optimization doesn't adequately reach a maximum for this index.
 
@@
 
-# Extimating the squint angle of the spiral
+# Estimating the squint angle of the spiral
 

### 2019-01-31 scagnostics.Rmd
Comments Rory.
Acknowledgments added.
Affiliation showing.

@@
 - name: Ursula Laa
-  thanks: The authors gratefully acknowledge the support of the Australian Research Council
-  affiliation: Department of Physics, Monash University
+  address: School of Physics and Astronomy, Monash University
   email: ursula.laa@monash.edu
-
 - name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
+  address: Department of Econometrics and Business Statistics, Monash University
+  email: dicook@monash.edu
 
@@
 The fitting procedures are described in detail in @Abbott2018exr.
-There are six parameters of physical interest (6-D) with some known relationships.
+We consider six parameters of physical interest (6-D) with some known relationships.
 Projection pursuit is used to find the known relationships.
@@
 
+# Acknowledgments {-}
+We thank Rory Smith for help with the gravitational wave examples, and German Valencia for useful comments.
+
 \appendix

### 2019-01-31 scagnostics.Rmd
Acknowledgements update.

@@
 # Acknowledgments {-}
+The authors gratefully acknowledge the support of the Australian Research Council.
 We thank Rory Smith for help with the gravitational wave examples, and German Valencia for useful comments.
+This article was created with knitr [@knitr], R Markdown [@rmarkdown] and bookdown [@bookdown] with embedded code.
 

### 2019-05-13 review/response.Rmd
review and response added

COMMENTS FOR THE AUTHOR:

Associate Editor:

The submission has been evaluated by two reviewers. The manuscript is well written and organized. I have some more philosophical issues in addition to those raised by the two reviewers.

The authors need to convince readers (users) this kind of environment is good for daily EDA for the real world of big data/data science.

Example(s) with very large number (at least hundreds) of dimensions should be included in the study.

**Theoretically, it might work for really big data, but there are two components that would be required: (1) fast computational engine, (2) an efficient optimiser. Both of these would require major implementation work. We have explained this now in the discusion section.**

**We have analysed an example that is high-dimensional but think that it would be too distracting for the reader of this paper. It is an analysis of physics modeled fits, using initial ideas developed in this paper. There the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability. The tour was applied on the reduced dimensionality, and the work is published in EPJC (cite).**

Who are the target users of this environment? Do you recommend naive users (users without much EDA experience) to get hand on the proposed system?

How do you encourage users to try your system?

**The methods are implemented in an R package, which makes them readily available to all. The code for this paper is all available on github, and we have now made this repo public. It depends on three other packages, mbgraphics, binostics, minerva, which are now updated and on CRAN. We have coded the diagnostic techniques into a new package, and will submit this to CRAN.**

Reviewer #1: Review report for "Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics"

This manuscript presented a simulation study to visually explore the properties of a set of projection pursuit indexes (PPIs) using a grand tour based on the five criterions. The selected PPIs were then applied to two real datasets in physics for revealing potential patterns. In general I found this article was well written and it fully addresses the problems to be solved. I have no major issue. Some specific comments are listed below:

1) Authors conducted the simulation study to evaluate the selected PPIs based on the five criterions and the results were summarized into Table 3. On the other hand, authors may consider another table that summarizes the advantages/disadvantages/characteristics of these mathematical indexes from the previous literatures. This review is helpful for users to comprehend these indexes' properties.

2) According to the aims & scope of the journal, Computational Statistics, this article should keep the content and discussion relevant more focused on the computational aspects.

3) Although the target of this article focuses on the dataset with two-dimensional structure (d=2), it is worth to discuss the possibility (or give some suggestions) of applying the proposed procedure to the situation where d > 2.

4) The procedure for assessing the PPIs' properties was listed in Section 3.1.2. In addition, it would be helpful to list the procedure step-by-step (with the detailed R functions/packages) to guide the users how to apply the tour methodology with PPIs to explore the potential patterns of the real-world datasets.

5) Why use the 5th/95th percentiles in Table 2 instead of the average/sd?

**The main reason is that it provides a range of typical values of the index, which is useful for scaling the index dynamically in the guided tour. While it is possible to use a mean, and maybe +/- 3sd's for a similar purpose, range is more useful.**

6) There was some confusion regarding the number of PPIs used in the content. There are eight indexes in Table 2; nine indexes in Fig 7; six indexes in Table 3, and seven indices in Section 4.2.2.

7) I would suggest moving the appendix A (the description of "holes/cmass" index) to Section 2.2 for readability.

8) Check the spelling of words for clarity and consistency, e.g, splines2D/splines2d; convex/1-convex, MIC/MICe, and so on.

### 2019-06-04 submission/scagnostics.Rmd
Keeping copy of submission version of rmd and bib.

--- 2019-01-31 scagnostics.Rmd
@@
 
-The procedures for assessing the PPI properties of smoothness, squintability, flexibility, rotation invariance, and speed are examined for samples from the family of data sets are:
+The procedures for assessing the PPI properties of smoothness, squintability, flexibility, rotation invariance, and speed examined for samples from the family of data sets are:
 
@@
 
-Given the patterns in Figure \ref{fig:wIntermediate} it would be expected that the sine could be found easily, using only optimization method 3 using the splines2d, dcor2d, MIC or TIC indexes.
+Given the patterns in Figure \ref{fig:wIntermediate} it would be expected that the sine could be found easily, using only optimization method 3 with the splines2d, dcor2d, MIC or TIC indexes.
 This is examined in Figure \ref{fig:findsine}.
@@
 From this it can be seen that m1 and m2 are strongly, and slightly, nonlinearly associated.
-Between the other variables, some linear association (R1, R2), some nonlinear association (L1, L2, R1, R2),  heteroskedastic variance in most variables and some bimodality (R1, L1, L2, m1, m2).
-
-The model describes a neutron star merger and contains 6 free parameters, with each neutron star described by its mass $m$ (m1, m2) and radius $R$ (R1, R2), and a so-called tidal deform ability parameter $\Lambda$ (L1, L2) which is a function of the mass and radius, approximately proportional to $(m/R)^{-5}$.
+Between the other variables we observe some linear association (R1, R2), some nonlinear association (L1, L2, R1, R2),  heteroskedastic variance in most variables and some bimodality (R1, L1, L2, m1, m2).
+
+The model describes a neutron star merger and contains 6 free parameters, with each neutron star described by its mass $m$ (m1, m2) and radius $R$ (R1, R2), and a so-called tidal deformability parameter $\Lambda$ (L1, L2) which is a function of the mass and radius, approximately proportional to $(m/R)^{-5}$.
 
@@
 Evaluating for normal distributions gives $1/(4\pi)$, rescaling such that the index values range from 0 to 1 then puts the value for normal distributions at approximately $0.2$.
-This appears consistent with the results we have, i.e.
+This is consistent with the results we found, i.e.
 for normal distributions the cmass index is about $0.2$, and the holes index (=1-cmass) is about $0.8$.
@@
 This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
-Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values.
+Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using the tour path between noise variables of the spiral1000 dataset and different $\epsilon$ and $\beta$ values.
 Both methods appear to be promising in smoothing the function.
@@
 
-```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
+```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indexes skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
 For comparison the red line in the background shows the trace without any smoothing applied."}

### 2019-06-04 review/response.Rmd
Starting work on revision.

@@
 
+**The target audience consists of two groups.
+First we address people interested in developping new index functions, who will find the introduced methods useful to diagnose new indexes as presented in the paper.
+The methods to do so are now easily accessible via the spinebil package.
+The second group would be people interested in applying the considered indexes to their datasets.
+This can be done directly using the tourr package, but we also encourage users unfamiliar with the system to try this through the GUI available in the galahr package.
+Both packages are now referenced in the discussion (or should it be mentioned somewhere else?).**
+
 **The methods are implemented in an R package, which makes them readily available to all.
@@
 
+**This comparison was the subject of the thesis referenced in the introduction (Grimm (2016)). We have now added a brief summary of these findings at the end of section 2.2.**
+
 2) According to the aims & scope of the journal, Computational Statistics, this article should keep the content and discussion relevant more focused on the computational aspects.
+
+**maybe move more things to the appendix? not sure how to reply here...**
 
@@
 
+**We thank the referee for the suggestion.
+Indeed much of the methods developped for diagnostics of the index function apply directly to scenarios where d > 2.
+While the considered index functions may in principle be generalised to higher dimensions as well, this is currently only implemented for the distance correlation function.
+A detailed assessment of the d > 2 situation will be the subject of future work.
+(as now stated in the discussion?)**
+
 4) The procedure for assessing the PPIs' properties was listed in Section 3.1.2.
 In addition, it would be helpful to list the procedure step-by-step (with the detailed R functions/packages) to guide the users how to apply the tour methodology with PPIs to explore the potential patterns of the real-world datasets.
+
+**The functions used for the evaluation of these tests are now packaged in the spinebil R package.
+The package documentation contains examples to guide users and make it easy to reproduce the results for different test datasets or different index functions.**
 
@@
 
+**Need to decide treatment of holes/cmass index!**
+
 7) I would suggest moving the appendix A (the description of "holes/cmass" index) to Section 2.2 for readability.
 
+**Mention holes/cmass with reference in Sec. 2.2, but keep rescaling information in the appendix?**
+
 8) Check the spelling of words for clarity and consistency, e.g, splines2D/splines2d; convex/1-convex, MIC/MICe, and so on.
+
+**We have made the spelling consistent where referring to the same index, and provided explanations where the different spelling indicates differences in index calculation.**

### 2019-06-04 scagnostics.Rmd
Starting work on revision.

@@
 This index should be more stable than MIC.
+
+\textcolor{red}{Here add short summary of comparisons between these functions for variable selection, taking results from Katrins thesis.}
 
@@
   select(-index1, -type1, -index2, -type2) %>%
-  mutate(index = factor(index, c("holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe"))) %>%
-  arrange(index) %>%
-  mutate(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", ""))
+  mutate(indexF = factor(index, c("holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe"))) %>%
+  arrange(indexF) %>%
+  select(-indexF, -index) %>%
+  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", ""))
 # Good resource for doing tables
@@
 
+\textcolor{red}{Mention spinebil and galahr packages, some statement about big data (addressing the AE question). Mention d>2 as subject of future work.}
+
 # Acknowledgments {-}

### 2019-06-05 review/response.Rmd
Adding comments from reviewer 2.

@@
 **We have made the spelling consistent where referring to the same index, and provided explanations where the different spelling indicates differences in index calculation.**
+
+Reviewer #2:
+Summary of the contribution:
+The authors present a simulation study to evaluate various projection pursuit indexes for detecting specific bivariate patterns in the data.
+Moreover, two data sets from physics are examined using PPI.
+
+Major remarks
+The simulation study targets three particular structures the pipe, the sine, and the spiral. Some motivation and justification for selecting these structures should be added.
+
+ **We have added a statement in Section 3.1.1 outlining the motivation for the three structures considered in the simulation study.**
+
+Second, the simulation study uses six-dimensional data of which four dimensions con- tribute nuisance and the other two the intended structure.
+Implicitly, this is clear, but the description does not make this explicit and uses the general notion of p dimensions instead.
+Please make this more explicit and also provide some justification that this set-up is adequate for targeting the problem at hand.
+
+**We have added an explicit statement that we fix p=6.
+As mentioned in the statement p=6 results in two independent planes showing the nuisance distribution and one plane showing the structured distribution.
+This set-up is adequat for most of our study where we follow interpolated paths either between nuisance only projections or from nuisance views to structured projections.**
+
+Thirdly, the simulations refer to sequences of projections (the horizontal axis in plots in Figures 2, 3, and 4.
+As I understand these sequences are a consequence of optimising the corresponding projection pursuit index.
+How can you then actually control the two different scenarios referred to in section 3.2 and 3.3?
+Or, do you extract a specified sequence of projections?
+Moreover, how do you make sure that you have the same sequence length for each trial?
+Please clarify!
+
+**Figures 2 and 3 present the trace for a specified sequence of projections (since in the simulation study the nuisance and structured views are known).
+This is explained in Section 3.1.2, and we have added a sentence clarifying this in Section 3.2 and made the wording more explicit in Section 3.3.
+Figure 4 is different as in this case the path is indeed the result of optimising an index.
+However, we only optimise the TIC index highlighted in red, while the other index values are shown for comparison along the optimisation path.**
+
+**(In my opinion all of this was already very clear before.
+When discussing this before you suggested making the Figure caption more clear, but it seems they are already very clear and I would not want to make them longer.)**
+
+The optimisation methods are further discussed in section 3.4 and depend on the struc- ture present in the data.
+What is the value of these investigations for more realistic settings in which the structure is not known.
+
+**(I would list: demonstrate potential as PPI, understand shortcomings, and based on that develop strategies, e.g. the two-step optimisation. Add something like this in Section 3.4?)**
+
+Finally, the relationship between the simulation study and the examples using the astrophysics data remains unclear.
+How are the physics applications linked to the simulation study and vice versa?
+Currently, the two aspects seem to me rather independent and could actually be presented in separate manuscripts.
+Moreover, since some rather important aspects of the simulation study have been put into an appendix.
+For the COST audience certainly the details of the simulation study are more relevant than the astrophysics application.
+
+**(Point to initial motiviation, what else should we say here?)**
+
+Minor remarks
+When introducing the various PPIs some terms and notations are used which are not properly introduced, e.g.
+alpha hull, convex.
+Check which ones actually need to be properly defined as they are needed for further steps and for which ones you could just refer to the literature avoiding unnecessary details.
+
+**(Still need to do this)**
+
+In Figure 3, each plot shows a vertical blue line. Please explain what this line represents.
+
+**We have added such a sentence to the figure caption.**
+
+p. 14, l. 26: optimist should be optimise
+
+**We have fixed the typo.**

### 2019-06-05 scagnostics.Rmd
Adding comments from reviewer 2.

@@
 In each set, $n$ is used for the number of points, $p$ is the number of dimensions, and $d=2$ is the projection dimension.
+\textcolor{red}{The three structures were selected to cover both functional and non-functional dependence, different types of nuissance distributions and different structure size and squintability properties.}
 
@@
 
-The structured projection is in variables 5 and 6 ($x_5, x_6$).
+\textcolor{red}{In the following we fix $p=6$ (this means that we have two independent planes containing the nuisance distribution, and one plane containing the structured distribution), and}
+the structured projection is in variables 5 and 6 ($x_5, x_6$).
 Two samples sizes are used: $n=(100, 1000)$.
@@
 Figure \ref{fig:plotV1V2toV3V4} shows the traces representing the index values calculated across a tour between a pair of nuisance projections.
+\textcolor{red}{The tour path is generating by interpolating between the two independent nuisance planes, i.e. from the projection onto $x_1$-$x_2$ to one onto $x_3$-$x_4$.}
 The range of each axis is set to be the limits of the index, as might be expected over many different data sets, 0 to 1.
@@
 
-Figure \ref{fig:wIntermediate} shows the PPIs for a tour sequence between a nuisance and structured projection.
-A long sequence is generated where the path runs from $x_1$-$x_2$, $x_1$-$x_5$, $x_5$-$x_6$, in order to see some of the intricacies of holes index.
+Figure \ref{fig:wIntermediate} shows the PPIs for a tour sequence between a nuisance and structured projection. A long sequence is generated where the path 
+\textcolor{red}{interpolates between projections onto}
+$x_1$-$x_2$, $x_1$-$x_5$, $x_5$-$x_6$, in order to see some of the intricacies of holes index.
 Sample size is indicated by line type: dashed being $n=100$ and solid is $n=1000$.
@@
 ```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between nuisance and structured projections, following $x_1$-$x_2$ to $x_1$-$x_5$ to $x_5$-$x_6$ (solid lines: $n=1000$; dashed: $n=100$).
+The vertical blue line indicates the position of the projection onto $x_1$-$x_5$ in the sequence.
 Peaks at the end of the sequence indicate the index sees the structure.
@@
 The optimization is done in two stages, a scouting phase using method 2, and a refinement stage using method 3.
-For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimist the TIC index.
+For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimised the TIC index.
 

### 2019-06-05 review/response.Rmd
response revised

@@
-COMMENTS FOR THE AUTHOR:
+## Author's response to reviews
 
-Associate Editor:
+The authors wish to thank the reviewers for the careful reading and comments on our first version. This is the detailed information on how we have addressed each comment from the reviewers.
+
+### Associate Editor:
 
@@
 
+**We have revised the introduction to improve the motivation to make this clear.** XXX CHECK THIS
+
 Example(s) with very large number (at least hundreds) of dimensions should be included in the study.
@@
 Both of these would require major implementation work.
-We have explained this now in the discusion section.**
+We have explained this now in the discussion section.**
 
@@
 **The target audience consists of two groups.
-First we address people interested in developping new index functions, who will find the introduced methods useful to diagnose new indexes as presented in the paper.
+First we address people interested in developing new index functions, who will find the introduced methods useful to diagnose new indexes as presented in the paper.
 The methods to do so are now easily accessible via the spinebil package.
@@
 
-Reviewer #1: Review report for "Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics"
+### Reviewer #1:
+
+Review report for "Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics"
 
@@
 
-**maybe move more things to the appendix? not sure how to reply here...**
+**We are really not sure what is being requested here.
+We have tried to focus the content in the paper on the compututational aspects of the work, and have kept it concise and to the point.
+The paper is at a reasonable length.
+The appendix has some details that broaden the discussion.
+If there are more parts that the reviewer suggests should be shifted to an appendix, we would be happy to do that.**
 
@@
 **We thank the referee for the suggestion.
-Indeed much of the methods developped for diagnostics of the index function apply directly to scenarios where d > 2.
-While the considered index functions may in principle be generalised to higher dimensions as well, this is currently only implemented for the distance correlation function.
-A detailed assessment of the d > 2 situation will be the subject of future work.
-(as now stated in the discussion?)**
+There are many index functions available for d>2, and the ones in the tourr package work for arbitrary dimension.
+The work in this paper focuses on studying a selection of indexes designed for 2D.
+However, the diagnostics cold be applied generally to any dimension projection index.
+We have commented on this in the discussion.** XXX CHECK THIS.
 
@@
 
-Reviewer #2:
+### Reviewer #2:
+
 Summary of the contribution:

### 2019-06-07 scagnostics.Rmd
Added summary of Katrins thesis.

@@
 
-\textcolor{red}{Here add short summary of comparisons between these functions for variable selection, taking results from Katrins thesis.}
+\textcolor{red}{A comparison between these indexes for the purpose of variable selection was discussed in @Grimm2016, and can be summarised as follows.
+The scagnostics measures are flexible and calculating the full set of measures provides useful guidance in variable selection, however they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning).
+Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation.
+Finally the mutual information based index functions are found to be flexible, but are sensitive to the sample size and often take relatively large values also when no association is present.
+A comparison of MIC and dcor2D was also presented in @2014arXiv1401.7645S.}
 

### 2019-06-07 review/response.Rmd
Added summary of Katrins thesis.
Working on the introduction.

@@
 
-**We have revised the introduction to improve the motivation to make this clear.** XXX CHECK THIS
+**We have revised the introduction to improve the motivation to make this clear.**
 
@@
 
-**Theoretically, it might work for really big data, but there are two components that would be required: (1) fast computational engine, (2) an efficient optimiser.
+**The simulation study and the applications reveal that the current methods are efficient only in smaller parameter spaces, there are two components that would be required to scale it to examples with hundreds of dimensions: (1) fast computational engine, (2) an efficient optimiser.
 Both of these would require major implementation work.
-We have explained this now in the discussion section.**
+We have explained this now in the discussion section.** XXX check, add to discussion
 
-**We have analysed an example that is high-dimensional but think that it would be too distracting for the reader of this paper.
-It is an analysis of physics modeled fits, using initial ideas developed in this paper.
-There the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability.
-The tour was applied on the reduced dimensionality, and the work is published in EPJC (cite).**
+**As we now explain in the introduction, we have previously considered an example where the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability (Cook, Laa, and Valencia 2018).**
 
@@
 
-**This comparison was the subject of the thesis referenced in the introduction (Grimm (2016)). We have now added a brief summary of these findings at the end of section 2.2.**
+**This comparison was the subject of the thesis Grimm (2016). We have now added a brief summary of these findings at the end of section 2.2.**
 

### 2019-06-07 scagnostics.Rmd
Working on the introduction.

@@
 This begs the question, whether important nonlinear associations are missed because tools to search are not available.
+
+\textcolor{red}{The same methods can also be used to explore larger datasets with hundreds of dimensions, as common in real world examples.
+In these scenarios it is best to reduce the number of dimensions before applying projection pursuit.
+For example in [@Cook:2018mvr] a 56 dimensional parameter space was reduced to the first six principal components before applying tour methods for the identification of multivariate relationships, revealing previously overlooked features.
+On the other hand, applying projection pursuit directly to the high-dimensional dataset is generally limited by the availability of an efficient optimiser.
+}
 

### 2019-06-07 review/response.Rmd
Working on discussion.

@@
 
-**The simulation study and the applications reveal that the current methods are efficient only in smaller parameter spaces, there are two components that would be required to scale it to examples with hundreds of dimensions: (1) fast computational engine, (2) an efficient optimiser.
+**The simulation study and the applications reveal that the current methods are efficient only in smaller parameter spaces.
+There are two components that would be required to scale it to examples with hundreds of dimensions: (1) fast computational engine, (2) an efficient optimiser.
 Both of these would require major implementation work.
@@
 This can be done directly using the tourr package, but we also encourage users unfamiliar with the system to try this through the GUI available in the galahr package.
-Both packages are now referenced in the discussion (or should it be mentioned somewhere else?).**
+Both packages are now referenced in the discussion.** XXX add references (need to put on CRAN first?)
 
-**The methods are implemented in an R package, which makes them readily available to all.
-The code for this paper is all available on github, and we have now made this repo public.
-It depends on three other packages, mbgraphics, binostics, minerva, which are now updated and on CRAN.
-We have coded the diagnostic techniques into a new package, and will submit this to CRAN.**
+**The considered index functions are all available in R packages, which makes them readily available to all. The packages are mbgraphics, binostics and minerva, which are now updated and on CRAN.**
 
@@
 The work in this paper focuses on studying a selection of indexes designed for 2D.
-However, the diagnostics cold be applied generally to any dimension projection index.
+However, the diagnostics could be applied generally to any dimension projection index.
 We have commented on this in the discussion.** XXX CHECK THIS.
@@
 
-**The functions used for the evaluation of these tests are now packaged in the spinebil R package.
-The package documentation contains examples to guide users and make it easy to reproduce the results for different test datasets or different index functions.**
+**As we now mention in the discussion, the easiest way for new users is to access these through the graphical interface available in the galahr package.
+The package documentation includes a step-by-step example using the guided tour.** XXX check!
 

### 2019-06-07 scagnostics.Rmd
Working on discussion.

@@
 Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
-
-The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools.
-Exploring this area would help improve the guided tours.
-The tour is also lacking a portable, easy to use interface that would enable user interactions like stopping, re-starting, saving optimal projections, and development of a graphical user interface, similar to the tourrGui [@tourrGui] (now archived), would help users and projection pursuit researchers.
+\textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes.
+It includes the functionalities needed to reproduce the assessements presented in this work.
+While the current focus is on two-dimensional index functions, the methodology introduced here directly applies to the assessment of index functions where $d>2$.}
+
+The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools. Exploring this area would help improve the guided tours.
+
+\textcolors{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [XXX Ref] package.
+Both packages contain examples to show how the guided tour can be used with different index functions.}
+
+#The tour is also lacking a portable, easy to use interface that would enable user interactions like stopping, re-starting, saving optimal projections, and development of a graphical user interface, similar to the tourrGui [@tourrGui] (now archived), would help users and projection pursuit researchers.
 

### 2019-06-18 scagnostics.Rmd
Removing obsolete comment.
textcolors to textcolor

@@
 
-\textcolors{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [XXX Ref] package.
+\textcolor{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [XXX Ref] package.
 Both packages contain examples to show how the guided tour can be used with different index functions.}
@@
 #The tour is also lacking a portable, easy to use interface that would enable user interactions like stopping, re-starting, saving optimal projections, and development of a graphical user interface, similar to the tourrGui [@tourrGui] (now archived), would help users and projection pursuit researchers.
-
-\textcolor{red}{Mention spinebil and galahr packages, some statement about big data (addressing the AE question). Mention d>2 as subject of future work.}
 

### 2019-07-24 review/response.Rmd
Working on revision.

@@
 Both of these would require major implementation work.
-We have explained this now in the discussion section.** XXX check, add to discussion
+We have explained this now in the discussion section.** XXX check
 
-**As we now explain in the introduction, we have previously considered an example where the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability (Cook, Laa, and Valencia 2018).**
+**However, as we now explain in the introduction, we have previously considered an example where the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability (Cook, Laa, and Valencia 2018).**
 
@@
 **We thank the referee for the suggestion.
-There are many index functions available for d>2, and the ones in the tourr package work for arbitrary dimension.
+Many of the available index functions also work for d>2, for example the ones in the tourr package work for arbitrary dimension.
 The work in this paper focuses on studying a selection of indexes designed for 2D.
-However, the diagnostics could be applied generally to any dimension projection index.
+However, the same diagnostics could be applied generally to any dimension projection index.
 We have commented on this in the discussion.** XXX CHECK THIS.

### 2019-07-24 scagnostics.Rmd
Working on revision.

@@
 The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools. Exploring this area would help improve the guided tours.
-
-\textcolor{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [XXX Ref] package.
+\textcolor{red}{In particular these developments would enable the application of guided tours to datasets with large number of dimensions.}
+
+\textcolor{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [XXX Ref] package, superseding the now archived tourrGui [@tourrGui].
 Both packages contain examples to show how the guided tour can be used with different index functions.}
-
-#The tour is also lacking a portable, easy to use interface that would enable user interactions like stopping, re-starting, saving optimal projections, and development of a graphical user interface, similar to the tourrGui [@tourrGui] (now archived), would help users and projection pursuit researchers.
 

### 2019-07-24 review/response.Rmd
Edited suggestions in replies to reviewer 2.

@@
 **(In my opinion all of this was already very clear before.
-When discussing this before you suggested making the Figure caption more clear, but it seems they are already very clear and I would not want to make them longer.)**
+When discussing this before you suggested making the Figure caption more clear, but it seems they are already very clear and I would not want to make them longer.)** XXX check and remove
 
@@
 
-**(I would list: demonstrate potential as PPI, understand shortcomings, and based on that develop strategies, e.g. the two-step optimisation. Add something like this in Section 3.4?)**
+**(I would list: demonstrate potential as PPI, understand shortcomings, and based on that develop strategies, e.g. the two-step optimisation. Add something like this in Section 3.4?)** XXX add?
 
@@
 
-**(Point to initial motiviation, what else should we say here?)**
+**(Point to initial motiviation, maybe say that physics application presents more realistic example with interesting structures in non-trivial projections)** XXX
 
@@
 
-**(Still need to do this)**
+**Suggestion: explain alpha hull, convex hull, MST, and add reference for smoothing splines?** XXX decide and add
 

### 2019-07-26 review/response.Rmd
Explicit statements on including holes index.

@@
 
-**Need to decide treatment of holes/cmass index!**
+**We are considering seven new index functions.
+In Section 3 we also include the holes index for comparison, as is now explained in the end of Section 2.2.
+The holes index is missing in Table 3 which summarises our findings on the new index functions, in addition we have combined the MINE indexes (MIC, TIC) that have very similar behaviour in all assessments, resulting in a list of six indexes.
+This is now also explained in the text.
+Finally Section 4.2.2 only considers the new index functions.**
 
@@
 
-**Mention holes/cmass with reference in Sec. 2.2, but keep rescaling information in the appendix?**
+**As mentioned above we now mention the holes index and point to the relevant references in Section 2.2. However, we believe the explanation of the rescaling should remain in the appendix.**
 

### 2019-07-26 scagnostics.Rmd
Explicit statements on including holes index.

@@
 A comparison of MIC and dcor2D was also presented in @2014arXiv1401.7645S.}
+
+\textcolor{red}{In addition to the seven indexes described above, we will also consider the holes index included in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
+This serves as a benchmark, demonstrating the desired behaviour.
+The index takes maximum values for a central hole distribution, and we rescale the values as described in the appendix such that values close to zero are found for normal distributions.}
 
@@
   mutate(holes = rescaleHoles(holes)) %>%
-  rowwise() %>%
-  mutate(cmass = rescaleCmass(cmass)) %>%
+  #rowwise() %>%
+  #mutate(cmass = rescaleCmass(cmass)) %>%
+  select(-cmass) %>%
   mutate(angle=alpha*360/(2*pi), TIC = TIC) %>%
@@
   gather(PPI, value, -angle) %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
+  mutate(PPI = fct_relevel(PPI, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
 
@@
 Such an overview is given in Table \ref{tab:summary}, listing if the criteria is fully met (\checkmark), there are some shortcomings ($\cdot$) or failure ($\times$).
+\textcolor{red}{Notice that the established holes index does not appear in the summary, and the MINE indexes (MIC and TIC) are combined as they were found to have similar behaviour.}
 We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fulfilled.

### 2019-08-05 review/response.Rmd
working on response to reviewers

@@
 
-**We have revised the introduction to improve the motivation to make this clear.**
+**We have added a paragraph to the introduction to explain the approach for higher dimensional problems, as follows:**
+
+**PP can also be used to explore data with many more variables, data that is commonly seen today.
+For example, [@Cook:2018mvr] explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit.
+There is not a single solution for working with high-dimensional data.
+PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies but preserved the nonlinear relationships that were interesting to discover.
+Some projection pursuit indexes incorporation penalty terms to naturally deal with noise dimensions.
+Having an efficient optimizer can also important with increasing dimensionality, because there are many more dimensions to search.**
 
@@
 **The simulation study and the applications reveal that the current methods are efficient only in smaller parameter spaces.
-There are two components that would be required to scale it to examples with hundreds of dimensions: (1) fast computational engine, (2) an efficient optimiser.
-Both of these would require major implementation work.
-We have explained this now in the discussion section.** XXX check
-
-**However, as we now explain in the introduction, we have previously considered an example where the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability (Cook, Laa, and Valencia 2018).**
+There are several components that would be required to scale it to examples with hundreds of dimensions: (1) fast computational engine, (2) an efficient optimiser, (3) new PP indexes that effectively use penalty terms to automatically elliminate nuisance dimensions.
+The first two would require major implementation work.
+The methods explained in this paper would be useful for assessing and evaluating new PP indexes.
+The recommended approach to high-dimensions is to first reduce the dimensionality using PCA, before applying PP.
+We have explained a sentence to this effect in the discussion section.**
 
@@
 **The target audience consists of two groups.
-First we address people interested in developing new index functions, who will find the introduced methods useful to diagnose new indexes as presented in the paper.
-The methods to do so are now easily accessible via the spinebil package.
-The second group would be people interested in applying the considered indexes to their datasets.
-This can be done directly using the tourr package, but we also encourage users unfamiliar with the system to try this through the GUI available in the galahr package.
-Both packages are now referenced in the discussion.** XXX add references (need to put on CRAN first?)
+First we hope that researchers developing new index functions, who will find our methods useful to diagnose their new indexes.
+We have implemented out methods in the R package, spinebil, to make this an attractive prospect to other researchers.
+The second group would be people interested in applying the indexes to explore nonlinear structure in their data.
+This can be done directly using the guided tour, in the tourr package, using code similarly to that distributed with this paper.
+We also encourage novice R users to try this through the tour GUI available in the galahr package.
+Both packages are now referenced in the discussion.**
 
@@
 However, the same diagnostics could be applied generally to any dimension projection index.
-We have commented on this in the discussion.** XXX CHECK THIS.
+We have commented on this in the discussion.**
 
@@
 The package documentation includes a step-by-step example using the guided tour.** XXX check!
+UP TO HERE!
 

### 2019-08-05 scagnostics.Rmd
working on response to reviewers

@@
 
-\textcolor{red}{The same methods can also be used to explore larger datasets with hundreds of dimensions, as common in real world examples.
-In these scenarios it is best to reduce the number of dimensions before applying projection pursuit.
-For example in [@Cook:2018mvr] a 56 dimensional parameter space was reduced to the first six principal components before applying tour methods for the identification of multivariate relationships, revealing previously overlooked features.
-On the other hand, applying projection pursuit directly to the high-dimensional dataset is generally limited by the availability of an efficient optimiser.
-}
+\textcolor{red}{PP can also be used to explore data with many more variables, data that is commonly seen today.
+For example, [@Cook:2018mvr] explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit.
+There is not a single solution for working with high-dimensional data.
+PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies but preserved the nonlinear relationships that were interesting to discover.
+Some projection pursuit indexes incorporation penalty terms to naturally deal with noise dimensions.
+Having an efficient optimizer can also important with increasing dimensionality, because there are many more dimensions to search.}
 
@@
 library(tidyverse)
-library(scagnostics)
+library(binostics)
 library(gridExtra)
@@
 \textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes.
-It includes the functionalities needed to reproduce the assessements presented in this work.
-While the current focus is on two-dimensional index functions, the methodology introduced here directly applies to the assessment of index functions where $d>2$.}
+It includes the functionalities needed to reproduce the assessements presented in this paper.
+While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.}
 
 The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools. Exploring this area would help improve the guided tours.
-\textcolor{red}{In particular these developments would enable the application of guided tours to datasets with large number of dimensions.}
+\textcolor{red}{As new optimization techniques become available, adapting these to the guided tour would extend the technique to a broader range of problems.
+The current recommended approach is to first reduce the dimensionality, for example by using PCA, taking care to preserve nonlinear structure, prior to applying PP.}
 
@@
 Both packages contain examples to show how the guided tour can be used with different index functions.}
+
+# Supplementary material
+
+- R code for paper
+- spinebil
+- galahr
+- outline of appendix
 

### 2019-08-06 review/response.Rmd
working on response to reviewers

@@
 The package documentation includes a step-by-step example using the guided tour.** XXX check!
-UP TO HERE!
 
@@
 **The main reason is that it provides a range of typical values of the index, which is useful for scaling the index dynamically in the guided tour.
-While it is possible to use a mean, and maybe +/- 3sd's for a similar purpose, range is more useful.**
+While it is possible to use a mean, and maybe +/- 3sd's, its possible that extreme values could occur making it less useful to scale the index for comparison  to other indexes.**
 
@@
 **We are considering seven new index functions.
-In Section 3 we also include the holes index for comparison, as is now explained in the end of Section 2.2.
-The holes index is missing in Table 3 which summarises our findings on the new index functions, in addition we have combined the MINE indexes (MIC, TIC) that have very similar behaviour in all assessments, resulting in a list of six indexes.
-This is now also explained in the text.
-Finally Section 4.2.2 only considers the new index functions.**
+In Section 3, we also include the holes index for comparison, because it is an established index with generally good, focused performance.
+This is now explained in the end of Section 2.2.
+The holes index is not included in Table 3 because the focus is on summarising our findings on the new index functions.
+Because the MINE indexes (MIC, TIC) that have very similar behaviour in all assessments,  we combined for this summary.
+We have now added another line for TIC, because it doesn't use much extra space, and saves confusion about the differing number of indexes.
+Finally Section 4.2.2 applies a selection of the new index functions to physics problems.**
 
@@
 
-**As mentioned above we now mention the holes index and point to the relevant references in Section 2.2. However, we believe the explanation of the rescaling should remain in the appendix.**
+**We now briefly describe the holes index and relevant references in Section 2.2.
+However, the explanation of the rescaling should remain in the appendix, because it distracts from the main message of Section 2.2.**
 
@@
 
- **We have added a statement in Section 3.1.1 outlining the motivation for the three structures considered in the simulation study.**
+ **We have added more explanation in Section 3.1.1 outlining the motivation for why these three structures are interesting for the simulation study.**
 
-Second, the simulation study uses six-dimensional data of which four dimensions con- tribute nuisance and the other two the intended structure.
+Second, the simulation study uses six-dimensional data of which four dimensions contribute nuisance and the other two the intended structure.
 Implicitly, this is clear, but the description does not make this explicit and uses the general notion of p dimensions instead.
@@
 As mentioned in the statement p=6 results in two independent planes showing the nuisance distribution and one plane showing the structured distribution.
-This set-up is adequat for most of our study where we follow interpolated paths either between nuisance only projections or from nuisance views to structured projections.**
+This set-up is adequate for most of our study where we follow interpolated paths either between nuisance only projections or from nuisance views to structured projections.**
 
@@
 **(In my opinion all of this was already very clear before.
-When discussing this before you suggested making the Figure caption more clear, but it seems they are already very clear and I would not want to make them longer.)** XXX check and remove
+When discussing this before you suggested making the Figure caption more clear, but it seems they are already very clear and I would not want to make them longer.)** XXX check and remove UP TO HERE!
 

### 2019-08-06 scagnostics.Rmd
working on response to reviewers

@@
 
-\textcolor{red}{In the following we fix $p=6$ (this means that we have two independent planes containing the nuisance distribution, and one plane containing the structured distribution), and}
-the structured projection is in variables 5 and 6 ($x_5, x_6$).
+\textcolor{red}{For simplicity in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
+}
+The structured projection is in variables 5 and 6 ($x_5, x_6$).
 Two samples sizes are used: $n=(100, 1000)$.

### 2019-08-06 scagnostics.Rmd
Outlining applications, filling in supplemental material section.

@@
 In each set, $n$ is used for the number of points, $p$ is the number of dimensions, and $d=2$ is the projection dimension.
-\textcolor{red}{The three structures were selected to cover both functional and non-functional dependence, different types of nuissance distributions and different structure size and squintability properties.}
+\textcolor{red}{The three structures were selected to cover both functional and non-functional dependence, different types of nuisance distributions and different structure size and squintability properties.}
 
@@
 
-\textcolor{red}{For simplicity in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
+\textcolor{red}{For simplicity in the investigations of the index behaviour, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
 }
@@
 
+Applying projection pursuit to new datasets can be done using the guided tour available in the `tourr` package. The necessary steps are:
+
+- Scaling and standardization of the data
+- Index function selection and if necessary formatting according to the guided tour requirements, i.e. a function taking matrix input and returning the index value
+- Calling the guided tour, for exploration this can be done via the `tourr::animate` function, for recording the results `tourr::save_history` should be used, examples are given in the documentation of the `tourr::guided_tour` function
+- Things to explore: different indexes, different starting planes, different optimisation options, different ways of reducing dimensionality prior to the guided tour run
+
 This section describes the application of these projection pursuit indices to find two-dimensional structure in two multidimensional gravitational waves problems.
+\textcolor{red}{These illustrate different ways of exploring real data using projection pursuit, showcasing the dependence of the results on index function, starting plane and variable selection.}
 
@@
 \textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes.
-It includes the functionalities needed to reproduce the assessements presented in this paper.
+It includes the functionalities needed to reproduce the assessments presented in this paper.
 While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.}
@@
 
-- R code for paper
-- spinebil
-- galahr
-- outline of appendix
+This article was created with R Markdown [@rmarkdown], the code for the paper is available at [XXX].
+The functions for testing new index functions as presented in this work are implemented in the R package spinebil [XXX].
+The R package galahr [XXX] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
+Additional information in the appendix is XXX
 

### 2019-08-06 scagnostics.Rmd
Fixed Table 3.

@@
 Such an overview is given in Table \ref{tab:summary}, listing if the criteria is fully met (\checkmark), there are some shortcomings ($\cdot$) or failure ($\times$).
-\textcolor{red}{Notice that the established holes index does not appear in the summary, and the MINE indexes (MIC and TIC) are combined as they were found to have similar behaviour.}
+\textcolor{red}{Notice that the established holes index does not appear in the summary.}
 We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fulfilled.
@@
 \begin{center}
-\caption{Summary of findings, showing to what extend the considered index functions pass the criteria for a good PPI.
+\caption{Summary of findings, showing to what extend the considered new index functions pass the criteria for a good PPI.
 ``\checkmark" symbols good behavior, ``$\cdot$" symbols some issues while ``$\times$" symbols failure on the corresponding criteria.
@@
 \hline
-Convex & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark & \checkmark \\
-Skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\times$ & \checkmark \\
-Stringy & $\times$ & $\times$ & $\cdot$ & $\cdot$ & \checkmark \\
+convex & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark & \checkmark \\
+skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\times$ & \checkmark \\
+stringy & $\times$ & $\times$ & $\cdot$ & $\cdot$ & \checkmark \\
 \hline
@@
 \hline
-MINE & $\cdot$ & \checkmark  & \checkmark & $\cdot$ & $\cdot$ \\
+MIC & $\cdot$ & \checkmark  & \checkmark & $\cdot$ & $\cdot$ \\
+TIC & $\cdot$ & \checkmark  & \checkmark & $\cdot$ & $\cdot$ \\
 \hline

### 2019-08-06 scagnostics.Rmd
last of reviewer comments

@@
 This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
-The projection pursuit guided tour is available in the R package, tourr [@tourr], and provides optimisation routines, and visualization.
+The projection pursuit guided tour is available in the R package, tourr [@tourr], and provides optimization routines, and visualization.
 
@@
 For exploring high-dimensional data, it can be interesting to observe local maxima as well as a global maximum, and thus a hybrid algorithm that still allows lingering but not being trapped by local maxima is ideal.
-In addition, being able to visually monitor the optimisation and see the optimal projection in the context of neighbouring projections is useful.
+In addition, being able to visually monitor the optimization and see the optimal projection in the context of neighbouring projections is useful.
 This is provided by combining projection pursuit with the grand tour [@CBCH94].
@@
 A possible algorithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
-The tourr package also provides the \texttt{search\_geodesic} function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimises the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
+The tourr package also provides the \texttt{search\_geodesic} function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimizes the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
 
@@
 
-```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between two nuisance projections.
+```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="PPIs for  projections along a geodesic interpolation between two nuisance projections.
 All projections would be nuisance so the PPI are ideally low and smooth, with little difference between sample sizes (solid lines: $n=1000$; dashed: $n=100$).
 The scagnostic PPIs are noisy.
-Some indexes have distinct differences in values between sample sizes."}
+Some indexes have distinct differences in values between sample sizes.
+(This is not an optimization path, but an interpolation containing 41 projections between two known projections.)"}
 namesFixed <- str_replace_all(fullRes$name, "sin", "sine")
@@
 Holes only responds to the pipe, and is a multimodal function for this data with a local maximum at $x_1$-$x_5$.
-"}
+(This is not an optimization path, but an interpolation containing 59 projections between three known projections.) "}
 intermediatePlane <- which(attributes(interpolate(t3))$new_basis)[2]
@@
 
-## Optimisation check with the guided tour
+## Optimization check with the guided tour
 \label{sec:guided}
@@
 The optimization is done in two stages, a scouting phase using method 2, and a refinement stage using method 3.
-For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimised the TIC index.
+For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimized the TIC index.
 
@@
 
-```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Scouting for the pipe using optimization method 2 on the TIC index, other PPI values also shown for interest.
-Only the values for the target planes are shown.
+```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="PPIs for a sequence of projections produced by scouting for the pipe using optimization method 2 on the TIC index.
+Other PPI values are shown for interest.
+Only the values on the target planes are shown.
 Despite the small maximum value of TIC for this data, it identifies the pipe."}
@@
 This is examined in Figure \ref{fig:findsine}.
-Optimization is conducted using the splines2d index, and the trace of the PPI over the optimisation is shown, along with the PPI values for the other indexes over that path.
+Optimization is conducted using the splines2d index, and the trace of the PPI over the optimization is shown, along with the PPI values for the other indexes over that path.
 The vertical blue lines indicate anchor bases, where the optimizer stops, and does a new search.
@@
 
-```{r findsine, fig.width=6, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Guided tour optimising  the splines2d index, using method 3, for the sine data, with $n=1000$.
+```{r findsine, fig.width=6, fig.height=7, out.width = "0.8\\textwidth", fig.cap="PPIs for sequence of projections produced by a guided tour optimizing  the splines2d index, using optimization method 3, for the sine data, with $n=1000$.
 Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima.
@@
 The spiral is the most challenging structure to detect because it has a small squint angle [@posse95a], especially as the ratio of noise to structure dimensions increases.
-This is explored using optimisation method 2 to scout the space for approximate maxima.
+This is explored using optimization method 2 to scout the space for approximate maxima.
 The skinny scagnostic index is used because it was observed (Figure \ref{fig:wIntermediate}) to be sensitive to this structure, although the noisiness of the index might be problematic.
@@
 This is an indication of the extra computation needed to brute force find the spiral as noise dimensions increase.
-Plot (b) shows the distance of the plane in each iteration of the optimisation to the ideal plane, where it can be seen that only when $p=4$ does it converge to the ideal.
+Plot (b) shows the distance of the plane in each iteration of the optimization to the ideal plane, where it can be seen that only when $p=4$ does it converge to the ideal.
 Its likely that expanding the search space should result in uncovering the spiral in higher dimensions, which however requires tuning of the stopping conditions and long run times.
 
-```{r findspiral, out.width="0.9\\textwidth", fig.cap="Guided tour optimising the skinny index for the Sprial dataset with 1000 datapoints, with p = 4, 5, 6.
-The left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the right shows the evolution of distance to the ideal plane as the index is being optimised.", results="hide", fig.width=7, fig.height=2.5}
+```{r findspiral, out.width="0.9\\textwidth", fig.cap="Guided tour optimizing the skinny index for the Sprial dataset with 1000 datapoints, with p = 4, 5, 6.
+The left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the right shows the evolution of distance to the ideal plane as the index is being optimized.", results="hide", fig.width=7, fig.height=2.5}
 
@@
 - Calling the guided tour, for exploration this can be done via the `tourr::animate` function, for recording the results `tourr::save_history` should be used, examples are given in the documentation of the `tourr::guided_tour` function
-- Things to explore: different indexes, different starting planes, different optimisation options, different ways of reducing dimensionality prior to the guided tour run
+- Things to explore: different indexes, different starting planes, different optimization options, different ways of reducing dimensionality prior to the guided tour run
 
@@
 
-```{r nseRemL1, fig.cap="Removing L1 and optimise again of the remaining parameters, where m2 remains removed from the set.
+```{r nseRemL1, fig.cap="Removing L1 and optimize again of the remaining parameters, where m2 remains removed from the set.
 Because of parameter correlations we can recover clear description of L2 as a function of the other parameters, despite m2 missing."}
@@
 Manually reconstructing the optimal projection (left plot) can be done by differencing the two parameters, in their original units.
-This highlights the importance of improved optimisation, that would use tiny final steps to polish the view to a finer optimal projection and possibly remove noise induced by small contributions of many variables.
+This highlights the importance of improved optimization, that would use tiny final steps to polish the view to a finer optimal projection and possibly remove noise induced by small contributions of many variables.
 
@@
 
-```{r randuEx, results="hide", fig.cap="Best projection obtained by optimising the MIC index on the RANDU data, using different number of bins, defined by $\\delta$.
+```{r randuEx, results="hide", fig.cap="Best projection obtained by optimizing the MIC index on the RANDU data, using different number of bins, defined by $\\delta$.
 The smaller the value the fewer bins.
@@
 
-```{r testpipe,  fig.height=4, fig.width=8, out.width = "0.8\\textwidth",  fig.cap="Projections returned by TIC optimisation: by the scouting phase (left) and refined by  optimisation method 3 (right), starting from the scouting phase projection."}
+```{r testpipe,  fig.height=4, fig.width=8, out.width = "0.8\\textwidth",  fig.cap="Projections returned by TIC optimization: by the scouting phase (left) and refined by  optimization method 3 (right), starting from the scouting phase projection."}
 iLast <- length(pipeTourFull)

### 2019-08-06 scagnostics.Rmd
Moved application guidelines.

--- 2019-08-06 scagnostics.Rmd
@@
 
-Applying projection pursuit to new datasets can be done using the guided tour available in the `tourr` package. The necessary steps are:
-
-- Scaling and standardization of the data
-- Index function selection and if necessary formatting according to the guided tour requirements, i.e. a function taking matrix input and returning the index value
-- Calling the guided tour, for exploration this can be done via the `tourr::animate` function, for recording the results `tourr::save_history` should be used, examples are given in the documentation of the `tourr::guided_tour` function
-- Things to explore: different indexes, different starting planes, different optimisation options, different ways of reducing dimensionality prior to the guided tour run
-
 This section describes the application of these projection pursuit indices to find two-dimensional structure in two multidimensional gravitational waves problems.
-\textcolor{red}{These illustrate different ways of exploring real data using projection pursuit, showcasing the dependence of the results on index function, starting plane and variable selection.}
 
@@
 
+## Application guide
+
+Applying projection pursuit to new datasets can be done using the guided tour available in the `tourr` package. The necessary steps are:
+
+1. Scaling and standardization of the data
+2. Index function selection and if necessary formatting according to the guided tour requirements, i.e. a function taking matrix input and returning the index value
+3. Calling the guided tour with the data and index function via:
+    + for exploration this can be done via the `tourr::animate` function
+    + for recording the results `tourr::save_history` should be used
+4. Exploring how the results depend on choices made (index function, starting planes, optimisation method, reducing dimensionality)
+
 <!--

### 2019-08-06 review/response.Rmd

@@
 
-**Figures 2 and 3 present the trace for a specified sequence of projections (since in the simulation study the nuisance and structured views are known).
-This is explained in Section 3.1.2, and we have added a sentence clarifying this in Section 3.2 and made the wording more explicit in Section 3.3.
-Figure 4 is different as in this case the path is indeed the result of optimising an index.
-However, we only optimise the TIC index highlighted in red, while the other index values are shown for comparison along the optimisation path.**
+**Captions and the text in Sections 3.2-3 have been changed to make it clearer that Figures 2 and 3 present the trace for a specified sequence of projections (since in the simulation study the nuisance and structured views are known), and Figure 4 is different as in this case the path is indeed the result of optimising an index.**
 
-**(In my opinion all of this was already very clear before.
-When discussing this before you suggested making the Figure caption more clear, but it seems they are already very clear and I would not want to make them longer.)** XXX check and remove UP TO HERE!
-
-The optimisation methods are further discussed in section 3.4 and depend on the struc- ture present in the data.
+The optimisation methods are further discussed in section 3.4 and depend on the structure present in the data.
 What is the value of these investigations for more realistic settings in which the structure is not known.
 
-**(I would list: demonstrate potential as PPI, understand shortcomings, and based on that develop strategies, e.g. the two-step optimisation. Add something like this in Section 3.4?)** XXX add?
+**(I would list: demonstrate potential as PPI, understand shortcomings, and based on that develop strategies, e.g. the two-step optimisation. Add something like this in Section 3.4?)** XXX add
 
@@
 
-**(Point to initial motiviation, maybe say that physics application presents more realistic example with interesting structures in non-trivial projections)** XXX
+**(Point to initial motiviation, maybe say that physics application presents more realistic example with interesting structures in non-trivial projections)** XXX add
 

### 2019-08-06 review/response.Rmd
Why optimisation checks

@@
 **As we now mention in the discussion, the easiest way for new users is to access these through the graphical interface available in the galahr package.
-The package documentation includes a step-by-step example using the guided tour.** XXX check!
+The package documentation includes a step-by-step example using the guided tour.
+We have also added a short guide to applications (Section 4.3).** XXX check!
 
@@
 
-**(I would list: demonstrate potential as PPI, understand shortcomings, and based on that develop strategies, e.g. the two-step optimisation. Add something like this in Section 3.4?)** XXX add
+**It is important to understand the performance of the optimisation before applying the method on new examples and we use the known examples for this.
+This is now explained in the first sentence in Section 3.4.** XXX check
 

### 2019-08-06 scagnostics.Rmd
Why optimisation checks

--- 2019-08-06 scagnostics.Rmd
@@
 
+\textcolor{red}{Before applying the new index functions with the guided tour on real examples, we test them on the simulated dataset to understand the performance of the optimization.}
 The guided tour combines optimization with interpolation between pairs of planes.
@@
 
-Applying projection pursuit to new datasets can be done using the guided tour available in the `tourr` package. The necessary steps are:
-
-- Scaling and standardization of the data
-- Index function selection and if necessary formatting according to the guided tour requirements, i.e. a function taking matrix input and returning the index value
-- Calling the guided tour, for exploration this can be done via the `tourr::animate` function, for recording the results `tourr::save_history` should be used, examples are given in the documentation of the `tourr::guided_tour` function
-- Things to explore: different indexes, different starting planes, different optimization options, different ways of reducing dimensionality prior to the guided tour run
-
 This section describes the application of these projection pursuit indices to find two-dimensional structure in two multidimensional gravitational waves problems.
-\textcolor{red}{These illustrate different ways of exploring real data using projection pursuit, showcasing the dependence of the results on index function, starting plane and variable selection.}
 
@@
 ```{r neutronStarEq,  results="hide"}
+## Step 1: read data, rescale
 nsD <- read_csv("data/samples.csv") 
@@
   rescale() # rescaling, this returns a matrix
+## Step 2: select index function (splines2d),
+##         formatting for guided tour is done in the function splineIndex
+## Step 3: recording guided tour via the save_history function
 if(!file.exists("cache/neutronStarEq.rda")){
@@
 Because of parameter correlations we can recover clear description of L2 as a function of the other parameters, despite m2 missing."}
+## Step 4: we explore reducing dimensionality, here dropping L1
+## then repeat steps 2+3 as before
 nsM2 <- nsD %>%
@@
 The 1-convex index finds a very different view.", fig.height = 3.2, dev = "png", dpi=300}
+# Step 1: data cleaning and rescaling
 bbhD <- read_csv("data/posterior_samples.csv")
 bbhM <- rescale(select(bbhD,-m2, -chi_eff))
+# Step 2: select index function (TIC) formatted in mineIndex
+# Step 3: record guided tour with save_history
 if(!file.exists("cache/bbhGuidedTIC.rda")){
@@
 }
+# Step 4: repeating steps 2+3 for different index functions (splines2d, 1-convex)
 if(!file.exists("cache/bbhGuidedSpline.rda")){
@@
 ```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2d and 1-convex.", fig.height = 3.2, dev = "png", dpi=300}
+## Step 4: exploring reduced parameter space (here simply dropping several variables)
 bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
@@
 ```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane."), dev = "png", dpi=300}
+## Step 4: exploring dependence on randomly selected starting planes
 bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
@@
 
+## Application guide
+
+Applying projection pursuit to new datasets can be done using the guided tour available in the `tourr` package. The necessary steps are:
+
+1. Scaling and standardization of the data
+2. Index function selection and if necessary formatting according to the guided tour requirements, i.e. a function taking matrix input and returning the index value
+3. Calling the guided tour with the data and index function:
+    + for exploration this can be done via the `tourr::animate` function
+    + for recording the results `tourr::save_history` should be used
+4. Exploring how the results depend on choices made (index function, starting planes, optimisation method, reducing dimensionality)
+
+These four steps are followed in the applications presented here and are documented in the comments of the source code [XXX github src].
+For simple usage examples see documentation of the `tourr::guided_tour` function.
+
 <!--

### 2019-08-06 Draft/src/main.R
Deleting obsolete files.

@@ -1,469 +1,0 @@
-## ---- load
-
-library(tourr)
-library(tidyverse)
-library(reshape2)
-library(binostics) #https://github.com/uschiLaa/binostics
-library(gridExtra)
-library(wesanderson) #color palette
-library(tictoc) #timer
-library(mbgraphic) #Katrins package
-library(dfoptim)
-
-# function to log transform part of data
-asLog <- function(dt){
-  dtOut <- dt %>% mutate(BRbsg_ln = log(BRbsg)) %>%
-    mutate(BrBsmumu_ln = log(BrBsmumu)) %>%
-    mutate(Omegahsq_ln = log(Omegahsq)) %>%
-    mutate(Xenonxsec_ln = log(Xenonxsec)) %>%
-    mutate(pSDxsec_ln = log(pSDxsec)) %>%
-    mutate(gminus2_shiftln = log(gminus2 + 1.8e-9)) %>%
-    select(-BRbsg, -BrBsmumu, -Omegahsq, -Xenonxsec, -pSDxsec, -gminus2)
-  return(dtOut)
-}
-
-#function to plot data and binning of requested projection
-plotProj <- function(nProj, fullPath, dMatrix){
-  p <- fullPath[nProj][[1]]
-  d <- dMatrix %*% p
-  s <- scagnostics.default(d[,1],d[,2], bins= 50)
-  p1 <- ggplot()+
-    geom_point(data = as.data.frame(d), mapping=aes(x=V1, y=V2))
-  p2 <- ggplot()+
-    geom_point(data = as.data.frame(s$bins), mapping = aes(x=V1, y=V2, color=V3, size=V3))+
-    xlim(0,1000)+
-    ylim(0,1000)+
-    lims(colour = c(1,25), size = c(1,25))+
-    scale_color_gradientn(colors=wes_palette(name="Zissou"))+
-    scale_size(range = c(1,3))
-    grid.arrange(arrangeGrob(p1, widths = 2),arrangeGrob(p2, widths = 7),nrow=1)
-    return(s$s)
-}
-
-scagIndex <- function(scagType){
-  function(mat){
-    sR <- scagnostics.default(mat[,1],mat[,2])$s
-    return(sR[scagType])
-  }
-}
-
-data <- as.matrix(numbat_only[,c(4,6,7)])
-scagIndex2 <- function(proj) {
-  proj <- matrix(proj, ncol=2, byrow=FALSE)
[…]

### 2019-08-06 testMultipleFits.R

@@ -1,50 +1,0 @@
-
-numbat_circle <- read_csv("Draft/data/numbat.csv")
-numbat_only <- numbat_circle %>% filter(group=="A") %>% select(-group)
-circle_only <- numbat_circle %>% filter(group=="B") %>% select(-group)
-
-physEx <- read_csv("~/smodels/smodels-master/testStop/summaryRvalExpected.csv")
-ex1 <- physEx %>% filter(group=="BRt_75_BRb_25") %>% select(-group,-mst,-mchi,-cat)
-
-
-scagIndex2 <- function(proj,dataF,scagType) {
-  proj <- matrix(proj, ncol=2, byrow=FALSE)
-  proj <- tourr:::orthonormalise(proj)
-  if(any(is.nan(proj))) return(0)
-  mat <- as.matrix(dataF) %*% proj
-  #scagType <- "Monotonic"
-  sR <- scagnostics.default(mat[,1],mat[,2])$s
-  return(sR[scagType])
-}
-
-#ix <- length(circle_only)
-ix <- length(ex1)
-ntries <- 10
-
-lower <- rep(-1,2*ix)
-upper <- rep(1, 2*ix)
-
-idxV <- rep(0,ntries)
-prjL <- list()
-
-i <- 1
-while(i<=ntries){
-  par0 <- runif(2*ix, min = 0, max = 1)
-  opt <- hjkb(par0, scagIndex2, lower=lower, upper=upper,
-              control = list(maximize=TRUE, tol=0.001,
-                             info=TRUE),dataF=ex1, scagType="Skinny")
-  #                           info=TRUE),dataF=circle_only, scagType="Skinny")
-  idxV[i] <- opt$value
-  prjL[[i]] <- tourr:::orthonormalise(matrix(opt$par, ncol=2, byrow=FALSE))
-  i <- i+1
-}
-
-i1 <- order(idxV)
-
-print(idxV[i1])
-
-plannedT <- planned_tour(prjL[i1])
-quartz()
-animate(ex1,plannedT)
-#animate(circle_only, plannedT)
-

### 2019-08-06 scagnostics.Rmd
Github references

@@
 
-This article was created with R Markdown [@rmarkdown], the code for the paper is available at [XXX].
-The functions for testing new index functions as presented in this work are implemented in the R package spinebil [XXX].
-The R package galahr [XXX] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
+This article was created with R Markdown [@rmarkdown], the code for the paper is available at [@paperSRC].
+The functions for testing new index functions as presented in this work are implemented in the R package spinebil [@spinebil].
+The R package galahr [@galahr] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
 Additional information in the appendix is XXX

### 2019-08-06 review/response.Rmd
why use physics data

@@
 
-**(Point to initial motiviation, maybe say that physics application presents more realistic example with interesting structures in non-trivial projections)** XXX add
+**As we now outline in the section overview, this paper first investigates the behaviour of the new indexes to understand the potential and challenges in using them in projection pursuit settings, before applying them to examples of the physics problems we are interested in addressing.
+While we agree that the simulation study is central to the work, we believe that showing applications of the method to unknown distributions is also relevant for the COST audience.
+Note that the application section now also contains guidelines on how to apply projection pursuit to new datasets, which are referenced in comments in the source code producing the examples, making them more useful to the reader** XXX check
 

### 2019-08-06 scagnostics.Rmd
why use physics data

@@
 
-This paper investigates the behaviour of these indexes.
+\textcolor{red}{This paper first investigates the behaviour of these indexes to understand the potential and challenges in using them in projection pursuit settings, before applying them to examples of the physics problems we are interested in addressing.}
 Section \ref{sec:construct} discusses index construction, and how they fit into the guided tour.

### 2019-08-06 scagnostics.Rmd
More references.

@@
 * **scagnostics**: The first step to computing the scagnostics is that the bivariate data is binned, and scaled between [0,1] for calculations.
-The convex and alpha hulls, and the minimal spanning tree (MST), are computed.
+The convex [@Eddy:1977] and alpha hulls [@1056714], and the minimal spanning tree (MST) [@2033241], are computed.
     + *convex*: The ratio of the area of alpha to convex hull, $I_{\bm{F}, convex}= \frac{area(A(Y))}{area(H(Y))}$.
@@
 The function wdcor, in the R package, extracat [@extracat], computes the statistic, and the mbgraphic package utilises this function.
-    + *splines2D*: Measures nonlinear dependence by fitting a spline model of $\bm{Y}_2$ on $\bm{Y}_1$ and also $\bm{Y}_1$ on $\bm{Y}_2$, using the \texttt{gam} function in the R package, mgcv [@w16].
+    + *splines2D*: Measures nonlinear dependence by fitting a spline model [@wahba1990spline] of $\bm{Y}_2$ on $\bm{Y}_1$ and also $\bm{Y}_1$ on $\bm{Y}_2$, using the \texttt{gam} function in the R package, mgcv [@w16].
 The index compares the variance of the residuals:

### 2019-08-07 scagnostics.Rmd
Fixed Table 2.

@@
   select(-indexF, -index) %>%
-  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", ""))
+  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", "")) %>%
+  select(index, everything())
 # Good resource for doing tables

### 2019-08-07 scagnostics.Rmd
Filling in supplemental material section.

@@
 
-These four steps are followed in the applications presented here and are documented in the comments of the source code [XXX github src].
+These four steps are followed in the applications presented here and are documented in the comments of the source code [@paperSRC].
 For simple usage examples see documentation of the `tourr::guided_tour` function.
@@
 Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
-\textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes.
+\textcolor{red}{We recommend to use the spinebil [@spinebil] package when developing new indexes.
 It includes the functionalities needed to reproduce the assessments presented in this paper.
@@
 
-\textcolor{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [XXX Ref] package, superseding the now archived tourrGui [@tourrGui].
+\textcolor{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [@galahr] package, superseding the now archived tourrGui [@tourrGui].
 Both packages contain examples to show how the guided tour can be used with different index functions.}
@@
 This article was created with R Markdown [@rmarkdown], the code for the paper is available at [@paperSRC].
-The functions for testing new index functions as presented in this work are implemented in the R package spinebil [@spinebil].
+Methods for testing new index functions as presented in this work are implemented in the R package spinebil [@spinebil].
 The R package galahr [@galahr] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
-Additional information in the appendix is XXX
+Additional information is available in the appendix, covering details of how the holes index was rescaled, estimations of the squint angle, a comparison of the computational performence of the index functions, testing the effect of index parameters on the results and suggestions how the index functions may be refined.
+Finally additional figures are also presented in the appendix.
 

### 2019-08-08 review/response.Rmd
More references.
final smoothing

--- 2019-08-06 review/response.Rmd
@@
 
-**As we now mention in the discussion, the easiest way for new users is to access these through the graphical interface available in the galahr package.
-The package documentation includes a step-by-step example using the guided tour.
-We have also added a short guide to applications (Section 4.3).** XXX check!
+**We have made a shiny app available in the package galahr, that will make it easier for new users to use a guided tour.
+The package documentation includes a step-by-step example.
+We have also added a step-by-step guide for the two applications in Section 4.3.**
 
@@
 **It is important to understand the performance of the optimisation before applying the method on new examples and we use the known examples for this.
-This is now explained in the first sentence in Section 3.4.** XXX check
+This is now explained in the first sentence in Section 3.4.**
 
@@
 
-**(Point to initial motiviation, maybe say that physics application presents more realistic example with interesting structures in non-trivial projections)** XXX add
+**The connection between two main sections is now explained in the introduction.
+We've added step wise instructions for applying projection pursuit to these two examples that can be used by researchers wishing to apply the techniques to their data.**
 
@@
 
-**Suggestion: explain alpha hull, convex hull, MST, and add reference for smoothing splines?** XXX decide and add
+**We have added more explanation, and appropriate references.**
 

### 2019-08-08 scagnostics.Rmd
final smoothing

@@
 This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+
+Projection pursuit can be understood and investigated more substantially when combined with a tour.
+A tour (@As85, @BCAH05) displays a smooth sequence of low dimensional projections from high dimensions to explore multivariate data for structure such as clusters, outliers, and nonlinear relationships.
+@CBCH94 provided an approach combining the tour algorithm with projection pursuit, to interactively both search for interesting projections, and examine the behavior of the indexes.
 The projection pursuit guided tour is available in the R package, tourr [@tourr], and provides optimization routines, and visualization.
 
-\textcolor{red}{This paper first investigates the behaviour of these indexes to understand the potential and challenges in using them in projection pursuit settings, before applying them to examples of the physics problems we are interested in addressing.}
-Section \ref{sec:construct} discusses index construction, and how they fit into the guided tour.
-Section \ref{sec:investigate} investigates the behaviour of the indexes, explored primarily using tour methods (@As85, @BCAH05, @CBCH94).
+\textcolor{red}{This paper is structured as follows.} Section \ref{sec:construct} discusses index construction, and how they can be used in the guided tour.
+Section \ref{sec:investigate} investigates the behaviour of the indexes, explored primarily using tour methods.
 The new guided tour with these indexes is applied to two examples from gravitational wave astronomy (Section \ref{sec:phys}).
+\textcolor{red}{The latter two parts are connected in that the application of the new indexes to these problems is the main motivation for the paper, and the simulation study, in the first part, was conducted to better understand the behavior of the indexes in general.
+The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to assess their behavior.
+The paper finishes with a discussion about the limitations of this work, and the potential future directions.}
 
@@
 
-## Application guide
-
-Applying projection pursuit to new datasets can be done using the guided tour available in the `tourr` package. The necessary steps are:
-
-1. Scaling and standardization of the data
-2. Index function selection and if necessary formatting according to the guided tour requirements, i.e. a function taking matrix input and returning the index value
-3. Calling the guided tour with the data and index function:
+## Instructions for applying to new data
+
+Applying these procedures to new datasets can be done using the guided tour available in the `tourr` package. The typical steps required are:
+
+1. Scale, standardize or sphere (principal components) the data. 
+2. Index function selection matching the type of structure that is interesting to detect. Any new function can be used as long as it takes a matrix as input and returns a single index value.
+3. Call the guided tour with the data and index function:
     + for exploration this can be done via the `tourr::animate` function
     + for recording the results `tourr::save_history` should be used
-4. Exploring how the results depend on choices made (index function, starting planes, optimisation method, reducing dimensionality)
-
-These four steps are followed in the applications presented here and are documented in the comments of the source code [@paperSRC].
-For simple usage examples see documentation of the `tourr::guided_tour` function.
+4. Explore how the results depend on choices made (index function, starting planes, optimisation method, prior dimension reduction).
+
+These are steps followed in the above two applications and are documented in the comments of the source code [@paperSRC].
+For simple usage examples, see documentation of the `tourr::guided_tour` function.
 
@@
 
-This article was created with R Markdown [@rmarkdown], the code for the paper is available at [@paperSRC].
-Methods for testing new index functions as presented in this work are implemented in the R package spinebil [@spinebil].
-The R package galahr [@galahr] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
-Additional information is available in the appendix, covering details of how the holes index was rescaled, estimations of the squint angle, a comparison of the computational performence of the index functions, testing the effect of index parameters on the results and suggestions how the index functions may be refined.
-Finally additional figures are also presented in the appendix.
+- This article was created with R Markdown [@rmarkdown], the code for the paper is available at [@paperSRC].
+- Methods for testing new index functions as presented in this work are implemented in the R package spinebil [@spinebil].
+- The R package galahr [@galahr] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
+- Additional explanations are available in the Appendix, covering details of 
+    - how the holes index was rescaled, 
+    - estimations of the squint angle, 
+    - a comparison of the computational performence of the index functions, 
+    - testing the effect of index parameters on the results and 
+    - suggestions how the index functions may be refined.
 

### 2019-08-08 scagnostics.Rmd
final smoothing

@@
 
-\textcolor{red}{A comparison between these indexes for the purpose of variable selection was discussed in @Grimm2016, and can be summarised as follows.
+\textcolor{red}{A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016, and is summarised as follows.
 The scagnostics measures are flexible and calculating the full set of measures provides useful guidance in variable selection, however they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning).
 Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation.
-Finally the mutual information based index functions are found to be flexible, but are sensitive to the sample size and often take relatively large values also when no association is present.
-A comparison of MIC and dcor2D was also presented in @2014arXiv1401.7645S.}
+The mutual information based index functions are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present.
+A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.}
 
 \textcolor{red}{In addition to the seven indexes described above, we will also consider the holes index included in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
-This serves as a benchmark, demonstrating the desired behaviour.
-The index takes maximum values for a central hole distribution, and we rescale the values as described in the appendix such that values close to zero are found for normal distributions.}
+This serves as a benchmark, demonstrating some desired behaviour.
+The index takes maximum values for a central hole in the sample distribution.}
 

### 2019-08-09 scagnostics.Rmd
Changed to AE everywhere.

@@
 The question is whether these can be adapted into projection pursuit indexes, to search for unusual features in two-dimensional projections of high-dimensional data.
-Recent PhD research by @Grimm2016 explored the behaviour of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation.
+Recent PhD research by @Grimm2016 explored the behavior of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation.
 In addition, two more indexes for measuring dependence have been proposed in the machine learning literature, based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], with computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308].
@@
 \textcolor{red}{This paper is structured as follows.} Section \ref{sec:construct} discusses index construction, and how they can be used in the guided tour.
-Section \ref{sec:investigate} investigates the behaviour of the indexes, explored primarily using tour methods.
+Section \ref{sec:investigate} investigates the behavior of the indexes, explored primarily using tour methods.
 The new guided tour with these indexes is applied to two examples from gravitational wave astronomy (Section \ref{sec:phys}).
@@
 If a variable is suspected to have little structure and relationship with other variables, conducting PP on the subset of variables without them may improve the efficiency of the search.
-- Centring and scaling of the projected data, can be helpful visually.
+- Centering and scaling of the projected data, can be helpful visually.
 If the data has a small amount of non-normal distribution in some directions, the projected data can appear to wander around the plot window during a tour.
-It doesn't matter what the centre of the projected data is, so centring removes a wandering scatterplot.
+It doesn't matter what the center of the projected data is, so centering removes a wandering scatterplot.
 Less commonly, it may be useful to scale the projected data to standard values, which would be done to remove any linear dependence remaining in the data.
@@
 &\\
-$\bm{X}_{n\times p} = (\bm{X}_1, ..., \bm{X}_p)$ & $n\times p$-dimensional data matrix, where variables $\bm{X}_j$ may be scaled or standardised, and $\bm{X}$ may be sphered\\
+$\bm{X}_{n\times p} = (\bm{X}_1, ..., \bm{X}_p)$ & $n\times p$-dimensional data matrix, where variables $\bm{X}_j$ may be scaled or standardized, and $\bm{X}$ may be sphered\\
 $\bm{Y}_{n\times 2} = (\bm{Y}_1, ..., \bm{Y}_2)$ & projected data matrix. where $Y_j = (\alpha_1 X_1, ..., \alpha_p X_p)$ \\
@@
 
-Table \ref{notation} summarises the notation used for this section.
+Table \ref{notation} summarizes the notation used for this section.
 Here we give an overview of the functions that are converted into projection pursuit indexes.
@@
     + *dcor2D*: This function is based on distance correlation [@szekely2007], which is designed to find both linear and nonlinear dependencies between variables.
-It involves computing the distances between pairs of observations, conducting an analysis of variance type breakdown of the distances relative to each variable, and the result is then passed to the usual covariance and hence correlation formula.
+It involves computing the distances between pairs of observations, conducting an analysis of variance type breakdown of the distances relative to each variable, and the result is then passed to the usual co variance and hence correlation formula.
 The function wdcor, in the R package, extracat [@extracat], computes the statistic, and the mbgraphic package utilises this function.
@@
 
-\textcolor{red}{A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016, and is summarised as follows.
+\textcolor{red}{A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016, and is summarized as follows.
 The scagnostics measures are flexible and calculating the full set of measures provides useful guidance in variable selection, however they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning).
@@
 \textcolor{red}{In addition to the seven indexes described above, we will also consider the holes index included in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
-This serves as a benchmark, demonstrating some desired behaviour.
+This serves as a benchmark, demonstrating some desired behavior.
 The index takes maximum values for a central hole in the sample distribution.}
@@
 For exploring high-dimensional data, it can be interesting to observe local maxima as well as a global maximum, and thus a hybrid algorithm that still allows lingering but not being trapped by local maxima is ideal.
-In addition, being able to visually monitor the optimization and see the optimal projection in the context of neighbouring projections is useful.
+In addition, being able to visually monitor the optimization and see the optimal projection in the context of neighboring projections is useful.
 This is provided by combining projection pursuit with the grand tour [@CBCH94].
@@
 
-\textcolor{red}{For simplicity in the investigations of the index behaviour, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
+\textcolor{red}{For simplicity in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
 }
@@
 
-For distributions and indexes with smooth behaviour and good squintability, method 3 is the most effective method for optimization.
+For distributions and indexes with smooth behavior and good squintability, method 3 is the most effective method for optimization.
 If these two criteria are not met the method may still be useful, but only given an informed starting projection.
@@
 The uniform distribution was used to generate the noise variables, which has a higher PPI value than a normal distribution, yielding the higher initial value.
-In addition, a local maximum is observed whenever the pair of variables is one structured variable and one noise variable, because there is a lighter density in the centre of the projection.
+In addition, a local maximum is observed whenever the pair of variables is one structured variable and one noise variable, because there is a lighter density in the center of the projection.
 
@@
 The purpose is that range differences in individual variables should not affect the detection  of relationships between multiple variables.
-Standardising the range will still leave differences between the standard deviations of the variables, and for this problem this is preferred.
+Standardizing the range will still leave differences between the standard deviations of the variables, and for this problem this is preferred.
 Differences in the standard deviations can be important for keeping the non-linear relationships visible to PP.
@@
     + for recording the results `tourr::save_history` should be used
-4. Explore how the results depend on choices made (index function, starting planes, optimisation method, prior dimension reduction).
+4. Explore how the results depend on choices made (index function, starting planes, optimization method, prior dimension reduction).
 
@@
     - estimations of the squint angle, 
-    - a comparison of the computational performence of the index functions, 
+    - a comparison of the computational performance of the index functions, 
     - testing the effect of index parameters on the results and 
@@
 The biggest issues revealed by the investigation into the new PPIs are a lack of smoothness, particularly for the scagnostics indexes, and the rotation invariance of Grimm's indexes.
-To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighbourhood of projections and average the value, or alternatively average the PPI for several jittered projections.
+To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighborhood of projections and average the value, or alternatively average the PPI for several jittered projections.
 This is investigated in Fig.

### 2019-08-14 scagnostics.Rmd
adjustments to introduction

@@
 It is hoped that the resulting solutions reveal low-dimensional structure in the data not found by methods such as principal component analysis.
-Projection pursuit is primarily used for visualization, with projected data always reported as plots.
+Projection pursuit is primarily used for visualization, with the projected data always reported as plots.
 
@@
 @f87, @hall89 @CBC92, @naito1997, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437, @PAN2000153, @FER13, @LOPERFIDO201842).
-Less work has been done on indexes to find nonlinear dependence between variables, when $d>1$.
+Less work has been done on indexes to find nonlinear dependence between variables, focused on $d=2$, which motivates this research.
 
@@
 
-The motivation is based on physics applications, to aid the interpretation of model fits on experimental results.
-A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated.
+The driving application is from physics, to aid the interpretation of model fits on experimental results.
+A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q~ (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated.
 (Note here, that while we may have analytic expressions for the predictions, this is not always the case and we often have to rely on numerical computation.) Different sets of model parameters ($n$) found to be  compatible with the experimental results within a selected level of confidence yield the data to be examined using projection pursuit.
@@
 Current practice is to examine pairs of parameters, or combinations produced by intuition or prior knowledge.
-This begs the question, whether important nonlinear associations are missed because tools to search are not available.
-
-\textcolor{red}{PP can also be used to explore data with many more variables, data that is commonly seen today.
-For example, [@Cook:2018mvr] explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit.
-There is not a single solution for working with high-dimensional data.
-PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies but preserved the nonlinear relationships that were interesting to discover.
-Some projection pursuit indexes incorporation penalty terms to naturally deal with noise dimensions.
-Having an efficient optimizer can also important with increasing dimensionality, because there are many more dimensions to search.}
-
-To find appropriate projections pursuit indexes, literature on variable selection was examined.
+This begs the question, whether important nonlinear associations are missed because because they are hidden in linear combinations of more than two variables.
+
+\textcolor{red}{PP can be combined with other dimension reduction methods when $p$ is very high.
+For example, it can be beneficial to first do principal component analysis prior to PP, especially to remove linear dependencies before searching for other types of association.
+This is the approach used in @Cook:2018mvr, which explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit.
+PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies while preserving the nonlinear relationships that were interesting to discover.
+Some projection pursuit indexes do incorporate penalty terms to automate removing noise dimensions.
+It can also be important to have an efficient PP optimizer, particularly when working with high dimensions, because the search space increases exponentially with dimension.}
+
+To find appropriate projections pursuit indexes for detecting nonlinear dependencies, the literature on variable selection was a starting point.
 With high-dimensional data, even plotting all pairs of variables can lead to too many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs.
@@
 
-Projection pursuit can be understood and investigated more substantially when combined with a tour.
+PP index behavior can be understood and investigated more substantially when combined with a tour.
 A tour (@As85, @BCAH05) displays a smooth sequence of low dimensional projections from high dimensions to explore multivariate data for structure such as clusters, outliers, and nonlinear relationships.
-@CBCH94 provided an approach combining the tour algorithm with projection pursuit, to interactively both search for interesting projections, and examine the behavior of the indexes.
+@CBCH94 provided an approach combining the tour algorithm with PP, to interactively both search for interesting projections, and examine the behavior of the indexes.
 The projection pursuit guided tour is available in the R package, tourr [@tourr], and provides optimization routines, and visualization.
@@
 \textcolor{red}{The latter two parts are connected in that the application of the new indexes to these problems is the main motivation for the paper, and the simulation study, in the first part, was conducted to better understand the behavior of the indexes in general.
-The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to assess their behavior.
+The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to visually assess their behavior.
+Visual methods to diagnose the index behavior is important because PP is primarily used for visualistion.
 The paper finishes with a discussion about the limitations of this work, and the potential future directions.}

### 2019-08-14 scagnostics.Rmd
finished proofing

@@
 
-\textcolor{red}{A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016, and is summarized as follows.
-The scagnostics measures are flexible and calculating the full set of measures provides useful guidance in variable selection, however they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning).
+\textcolor{red}{A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016.
+It is only available in German, we we summarise the main findings here.
+The scagnostics measures are flexible, and calculating the full set of measures provides useful guidance in variable selection.
+However, they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning).
 Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation.
-The mutual information based index functions are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present.
+The mutual information based index functions (MIC< TIC) are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present.
 A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.}
 
-\textcolor{red}{In addition to the seven indexes described above, we will also consider the holes index included in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
+\textcolor{red}{In addition to the seven indexes described above, we will also include the holes index available in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
 This serves as a benchmark, demonstrating some desired behavior.
-The index takes maximum values for a central hole in the sample distribution.}
+The index takes maximum values for a central hole in the distribution of the projected data.}
 
@@
 
-\textcolor{red}{For simplicity in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
+\textcolor{red}{For simplicity, in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
 }
@@
 Figure \ref{fig:plotV1V2toV3V4} shows the traces representing the index values calculated across a tour between a pair of nuisance projections.
-\textcolor{red}{The tour path is generating by interpolating between the two independent nuisance planes, i.e. from the projection onto $x_1$-$x_2$ to one onto $x_3$-$x_4$.}
+\textcolor{red}{The tour path is generated by interpolating between the two independent nuisance planes, i.e. from the projection onto $x_1$-$x_2$ to one onto $x_3$-$x_4$.}
 The range of each axis is set to be the limits of the index, as might be expected over many different data sets, 0 to 1.
@@
 
-\textcolor{red}{Before applying the new index functions with the guided tour on real examples, we test them on the simulated dataset to understand the performance of the optimization.}
+\textcolor{red}{Before applying the new index functions, with the guided tour on real examples, we test them on the simulated dataset to understand the performance of the optimization.}
 The guided tour combines optimization with interpolation between pairs of planes.
@@
 Such an overview is given in Table \ref{tab:summary}, listing if the criteria is fully met (\checkmark), there are some shortcomings ($\cdot$) or failure ($\times$).
-\textcolor{red}{Notice that the established holes index does not appear in the summary.}
+\textcolor{red}{(The holes index does not appear in the summary because its performance understood, and is not being examined here.)}
 We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fulfilled.
@@
 
-\textcolor{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [@galahr] package, superseding the now archived tourrGui [@tourrGui].
+\textcolor{red}{To apply the existing index functions in practice, we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [@galahr] package.
+This package supersedes the now archived tourrGui [@tourrGui].
 Both packages contain examples to show how the guided tour can be used with different index functions.}

### 2019-08-15 Appendix/appendix.Rmd
Removing color, splitting off appendix.

--- 2019-08-09 scagnostics.Rmd
@@
 ---
-title: Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics
+title: Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics -- Appendix
 
@@
 
-keywords:
-- scagnostics
-- statistical graphics
-- data visualisation
-- exploratory data analysis
-- data science 
-- guided tour
-
 abstract: |
-  Projection pursuit is used to find interesting low-dimensional projections of high-dimensional data by optimizing an index over all possible projections.
-Most indexes have been developed to detect departure from known distributions, such as normality, or to find separations between known groups.
-Here, we are interested in finding projections revealing potentially complex bivariate patterns, using new indexes constructed from scagnostics and a maximum information coefficient, with a purpose to detect unusual relationships between model parameters describing physics phenomena.
-The performance of these indexes is examined with respect to ideal behaviour, using simulated data, and then applied to problems from gravitational wave astronomy.
-The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphic.
-
-bibliography: bibliography.bib
+  Appendix material for the paper "Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics".
+
+bibliography: ../bibliography.bib
 output: rticles::springer_article
@@
 ```
-
-# Introduction
-
-<!-- 
-Main point of paper:
-
-- Some metrics for exploring high-dimensional data by finding interesting pairs of variables, can be used to find interesting projections, but some have problems. 
-- What are the ways to evaluate.
-- What adjustments can be done.
-- How it can be applied
-
-Organisation of this section:
-
-- Motivation first?
-- PP
-- Scagnostics
-- section needs shortening
--->
-
-The term "projection pursuit" (PP) was coined by @FT74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
-The procedure, originally suggested by @kr69, involves defining a criterion function, or index, that measures the "interestingness" of each $d$-dimensional projection of $p$-dimensional data.
-This criterion function is optimized over the space of all $d$-dimensional projections of $p$-space, searching for both global and local maxima.
-It is hoped that the resulting solutions reveal low-dimensional structure in the data not found by methods such as principal component analysis.
-Projection pursuit is primarily used for visualization, with projected data always reported as plots.
-
-A large number of projection pursuit indexes have been developed, primarily based on departure from normality, which includes clusters, outliers and skewness, and also for finding separations between known groups (e.g.
-@f87, @hall89 @CBC92, @naito1997, @lckl2005, @AHC02, @CEM:CEM2568, @JS87, @5508437, @PAN2000153, @FER13, @LOPERFIDO201842).
-Less work has been done on indexes to find nonlinear dependence between variables, when $d>1$.
-
-<!-- check Perisic and Posse (2012) and recent paper by Austrian guys -->
-
-The motivation is based on physics applications, to aid the interpretation of model fits on experimental results.
-A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated.
-(Note here, that while we may have analytic expressions for the predictions, this is not always the case and we often have to rely on numerical computation.) Different sets of model parameters ($n$) found to be  compatible with the experimental results within a selected level of confidence yield the data to be examined using projection pursuit.
-A single prediction can be a complicated function of all of the free parameters, and typically $q \in [100,1000]$ and $p \sim 10$.
-Current practice is to examine pairs of parameters, or combinations produced by intuition or prior knowledge.
-This begs the question, whether important nonlinear associations are missed because tools to search are not available.
-
-\textcolor{red}{PP can also be used to explore data with many more variables, data that is commonly seen today.
-For example, [@Cook:2018mvr] explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit.
-There is not a single solution for working with high-dimensional data.
-PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies but preserved the nonlinear relationships that were interesting to discover.
-Some projection pursuit indexes incorporation penalty terms to naturally deal with noise dimensions.
-Having an efficient optimizer can also important with increasing dimensionality, because there are many more dimensions to search.}
-
-To find appropriate projections pursuit indexes, literature on variable selection was examined.
-With high-dimensional data, even plotting all pairs of variables can lead to too many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs.
-There are eight scagnostics, of which three ("convex", "skinny" and "stringy") are used here.
-The question is whether these can be adapted into projection pursuit indexes, to search for unusual features in two-dimensional projections of high-dimensional data.
-Recent PhD research by @Grimm2016 explored the behavior of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation.
-In addition, two more indexes for measuring dependence have been proposed in the machine learning literature, based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], with computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308].
-These are related to original 1D projection pursuit indexes based on entropy (e.g.
-@huber85, @JS87).
-This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
-
-Projection pursuit can be understood and investigated more substantially when combined with a tour.
-A tour (@As85, @BCAH05) displays a smooth sequence of low dimensional projections from high dimensions to explore multivariate data for structure such as clusters, outliers, and nonlinear relationships.
-@CBCH94 provided an approach combining the tour algorithm with projection pursuit, to interactively both search for interesting projections, and examine the behavior of the indexes.
-The projection pursuit guided tour is available in the R package, tourr [@tourr], and provides optimization routines, and visualization.
-
-\textcolor{red}{This paper is structured as follows.} Section \ref{sec:construct} discusses index construction, and how they can be used in the guided tour.
-Section \ref{sec:investigate} investigates the behavior of the indexes, explored primarily using tour methods.
-The new guided tour with these indexes is applied to two examples from gravitational wave astronomy (Section \ref{sec:phys}).
-\textcolor{red}{The latter two parts are connected in that the application of the new indexes to these problems is the main motivation for the paper, and the simulation study, in the first part, was conducted to better understand the behavior of the indexes in general.
-The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to assess their behavior.
-The paper finishes with a discussion about the limitations of this work, and the potential future directions.}
-
-# Projection pursuit index construction and optimization
-\label{sec:construct}
-
-A projection pursuit index (PPI) is a scalar function $f$ defined on an $d$-dimensional data set, computed by taking a $d$-dimensional projection of an $n\times p$ data matrix.
-Typically the definition is such that larger values of $f$ indicate a more interesting distribution of observations, and therefore maximizing $f$ over all possible $d$ dimensional projections of a data set with $p>d$ variables will find the most interesting projections.
-This section describes the seven indexes that are to be used to explore bivariate association.
-Some data pre-processing, including standardization, is advisable, prior to optimizing the PPI.
-
-## Scaling and standardization
-
-Making a plot always involves some choice of scaling.
-When a scatterplot is made, effectively, albeit under the hood, the data is scaled into a range of $[a, b]$ (often $a=0, b=1$) on both axes to print it on a page or display in a graphics device window.
-The range deliminates page space within which to draw.
-The upshot is that the original data scale is standardized to the range and aspect ratio on the display space.
-It may be that the original range of one variable is $[1, 1000]$ and the other is $[1, 1.6]$ but the display linearly warps this to $[0, 1]$ and $[0, 1]$, say, giving both variables equal visual weight.
-
-With high dimensional data, and particularly projections, it is also necessary to re-scale the original range, and it is important to pay attention to what is conventional, or possible, and the effects.
-The PPIs also may require specific scaling for them to be effectively computed.
-Both of these are addressed here.
-The common pre-processing include:
-
-- Standardizing each variable, to have mean 0 and variance 1, so that individual variable scales do not affect the result.
-Different variable scales are examinable without resorting to projection pursuit, so can be handled prior to searching through high dimensions.
-- Sphering the high-dimensional data is often done to remove linear dependence.
-This is typically done using principal component analysis, and using the principal components as the variables passed to PP.
-If linear dependence is the only structure PP is not needed, and thus this is removed before PP so that the PPIs are not distracted by simple structure.
-- Transform single variables to reduce skewness.
-It is marginal structure, visible in a single variable, which doesn't need a multivariate technique to reveal.
-Skewed distributions will inadvertently affect the PPIs, distracting the search for dependence.
-- Remove outliers, which may be an iterative process, to discover, identify and delete. Extreme values will likely affect PPI performance. Outliers can be examined on a case by case basis later.
-- Possibly remove noise dimensions, which is also likely to be an iterative process.
-Directions where the distribution is purely noise make optimization of a PPI more difficult.
-If a variable is suspected to have little structure and relationship with other variables, conducting PP on the subset of variables without them may improve the efficiency of the search.
-- Centering and scaling of the projected data, can be helpful visually.
-If the data has a small amount of non-normal distribution in some directions, the projected data can appear to wander around the plot window during a tour.
-It doesn't matter what the center of the projected data is, so centering removes a wandering scatterplot.
-Less commonly, it may be useful to scale the projected data to standard values, which would be done to remove any linear dependence remaining in the data.
-
-\begin{table}[htp]
-\caption{Summary of notation}
-\centering
-\begin{tabular}{|l|p{8.2cm}|} \hline
-$p$ & be the number of variables in the data. For physics models the observable space is higher dimension, say $q$, and the data examined is the fitted model space, typically less than 10. \\
-$d (=2) $ & projection dimension. For studying physics models, typically $d=2$ and this is the focus for the index definition. \\
-$n$ & Number of observations, which for the physics models is the number of fitted models being examined and compared. \\
-&\\
-$\bm{X}_{n\times p} = (\bm{X}_1, ..., \bm{X}_p)$ & $n\times p$-dimensional data matrix, where variables $\bm{X}_j$ may be scaled or standardized, and $\bm{X}$ may be sphered\\
-$\bm{Y}_{n\times 2} = (\bm{Y}_1, ..., \bm{Y}_2)$ & projected data matrix. where $Y_j = (\alpha_1 X_1, ..., \alpha_p X_p)$ \\
-$\bm{F}_{p\times 2} = (\bm{\alpha}_1, ..., \bm{\alpha}_2)$ & orthonomal projection matrix\\ 
-&\\
-$H$ & Convex hull \\
-$A$ & Alpha hull \\\hline
-\end{tabular}
-\label{notation}
-\end{table}
-
-## New projection pursuit index functions
-\label{sec:indexDef}
-
-Table \ref{notation} summarizes the notation used for this section.
-Here we give an overview of the functions that are converted into projection pursuit indexes.
-Full details of the functions can be found in the original sources.
-
-* **scagnostics**: The first step to computing the scagnostics is that the bivariate data is binned, and scaled between [0,1] for calculations.
-The convex [@Eddy:1977] and alpha hulls [@1056714], and the minimal spanning tree (MST) [@2033241], are computed.
-    + *convex*: The ratio of the area of alpha to convex hull, $I_{\bm{F}, convex}= \frac{area(A(Y))}{area(H(Y))}$.
-This is the only measure where interesting projections will take low values, with a maximum of 1 if both areas are the same.
-Thus $1-c_{convex}$ is used.
-    + *skinny*: The ratio of the perimeter to the area of the alpha hull, $I_{\bm{F}, skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $I_{\bm{F}, skinny} = 0$ for a full circle.
-Values close to 1 indicate a skinny polygon.
-    + *stringy*: Based on the MST, $I_{\bm{F}, stringy} = \frac{diameter(MST)}{length(MST)}$ where the diameter is the longest connected path, and the length is the total length (sum of all edges).
-If the MST contains no branches $I_{\bm{F}, skinny} = 1$.
-
-* **association**: The index functions are available in @mbgraphic, and are defined to range in [0,1].
-Both functions in the mbgraphic package can bin the data before computing the index, for computational performance.
-    + *dcor2D*: This function is based on distance correlation [@szekely2007], which is designed to find both linear and nonlinear dependencies between variables.
-It involves computing the distances between pairs of observations, conducting an analysis of variance type breakdown of the distances relative to each variable, and the result is then passed to the usual co variance and hence correlation formula.
-The function wdcor, in the R package, extracat [@extracat], computes the statistic, and the mbgraphic package utilises this function.
-    + *splines2D*: Measures nonlinear dependence by fitting a spline model [@wahba1990spline] of $\bm{Y}_2$ on $\bm{Y}_1$ and also $\bm{Y}_1$ on $\bm{Y}_2$, using the \texttt{gam} function in the R package, mgcv [@w16].
-The index compares the variance of the residuals:
-    \begin{equation}
-    I_{\bm{F}, splines2d} = max(1- \frac{Var(res_{\bm{Y}_1\sim \bm{Y}_2})}{Var(\bm{Y}_1)}, 1-\frac{Var(res_{\bm{Y}_2\sim {\bm{Y}_1}})}{Var(\bm{Y}_2)}),
-    \end{equation}
-    which takes large values if functional dependence is strong.
-
-* **information**:  The index functions [@Reshef1518] nonparametricly measure nonlinear association by computing the mutual information, 
-\begin{equation}
-\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p(by_2))),
-\end{equation}
-where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$is the relative bin count in each cell, and $p(by_1), p(by_2)$ are the row and column relative counts, on a range of bin resolutions of the data.
-It is strictly a 2D measure.
-For a fixed binning, e.g.
-$2\times 2$ or $10\times 4$, the optimal binning is found by maximizing $I$.
-The values of I range between $[0,1]$ because they are normalized across bins by dividing by $log(min(\# bins_{y_1}, \# bins_{y_2}))$.
-    - *Maximum Information Coefficient (MIC)*: uses the maximum normalized $I$ across all bin resolutions. 
-    - *Total Information Coefficient (TIC)*: sums the normalized $I$ for all bin resolutions computed.
-This creates a problem of scaling - there is no upper limit, although it is related to number of bins, and number of bin resolutions used.
-In the work below we have made empirical estimates of the maximum and scaled the TIC index using this to get it in the range $[0,1]$.
-This index should be more stable than MIC.
-
-\textcolor{red}{A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016, and is summarized as follows.
-The scagnostics measures are flexible and calculating the full set of measures provides useful guidance in variable selection, however they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning).
-Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation.
-The mutual information based index functions are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present.
-A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.}
-
-\textcolor{red}{In addition to the seven indexes described above, we will also consider the holes index included in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
-This serves as a benchmark, demonstrating some desired behavior.
-The index takes maximum values for a central hole in the sample distribution.}
-
-## Optimization
-Given a PPI we are confronted with the task of finding the maximum over all possible $d$ dimensional projections.
-One challenge is to avoid getting trapped in local maxima that are only a result of sampling fluctuations or a consequence of a noisy index function.
-@posse95b discusses the optimization, in particular that for most index functions and optimizers results are too local, largely dependent on starting point.
-@f87 suggested a two-step procedure: the first step is using a large step size to find the approximate global maximum while stepping over pseudomaxima.
-A second step is then starting from the projection corresponding to the approximate maximum and employing a gradient directed optimization for the identification of the maximum.
-For exploring high-dimensional data, it can be interesting to observe local maxima as well as a global maximum, and thus a hybrid algorithm that still allows lingering but not being trapped by local maxima is ideal.
-In addition, being able to visually monitor the optimization and see the optimal projection in the context of neighboring projections is useful.
-This is provided by combining projection pursuit with the grand tour [@CBCH94].
-The properties of a suitable optimization algorithm include monotonicity of the index value, a variable step-size to avoid overshooting and to increase the chance of reaching the nearest maximum, and a stopping criterion allowing to move out of a local maximum and into a new search region [@tourr].
-A possible algorithm is inspired by simulated annealing and has been described in @lckl2005, this has been implemented in the \texttt{search\_better} and \texttt{search\_better\_random} search functions in the tourr package.
-The tourr package also provides the \texttt{search\_geodesic} function, which first selects a promising direction by comparing index values between a selected number of small random steps, and then optimizes the function over the line along the geodesic in that direction considering projections up to $\pi/4$ away from the current view.
 
@@
 ```
-
-# Investigation of indexes
-\label{sec:investigate}
-
-A useful projection pursuit index needs to have several properties.
-This has been discussed in several seminal papers, e.g.
-@diaconis84, @huber85, @JS87, @posse95b, @hall89.
-The PPI should be  minimized by the normal distribution, because this is not interesting from a data exploration perspective.
-If all projections are normally distributed, good modelling tools already exist.
-A PPI should be approximately affine invariant, regardless how the projection is rotated the index value should be the same, and the scale of each variable shouldn't affect the index value.
-Interestingly the original index proposed by @FT74 was not rotationally invariant.
-A consistent index means that small perturbations to the sample do not dramatically change the index value.
-This is particularly important to making optimization feasible, small angles between projections correspond to small perturbations of the sample, and thus should be small changes to index value.
-@posse95b suggests that indexes should be resistant to features in a tail of the distribution, but this is debatable because one departure from normality that is interesting to detect are anomalous observations.
-Some PPI are designed precisely for these reasons.
-Lastly, because we need to compute the PPI over many projections, it needs to be fast to compute.
-These form the basis of the criteria upon which the scagnostic indexes, and the several alternative indexes are examined, as explained below.
-
-- **smoothness**: This is the consistency property mentioned above.
-The index function values are examined over interpolated tour paths, that is, the value is plotted against time, where time indexes the sequence of projections produced by the tour path.
-The signature of a good PPI is that the plotted function is relatively smooth.
-The interpolation path corresponds to small angle changes between projections, so the value should be very similar.
-
-- **squintability**: @barnett1981interpreting introduced the idea of squint angle to indicate resolution of structure in the data.
-Fine structure like the parallel planes in the infamous RANDU data [@Marsaglia68] has a small squint angle because you have to be very close to the optimal projection plane to be able to see the structure.
-Structures with small squint angle are difficult to find, because the optimization algorithm needs to get very close to begin hill-climbing to the optimum.
-The analyst doesn't have control over the data structure, but does have control over the PPI.
-Squintability is about the shape of the PPI over all projections.
-It should have smooth low values for noise projections and a clearly larger value, ideally with a big squint angle, for structured projections.
-The optimizer should be able to clearly see the optimal projections as different from noise.
-To examine squintability, the PPI values are examined on interpolated tour paths between a noise projection and a distant structured projection.
-
-- **flexibility**: An analyst can have a toolbox of indices that may cover the range of fine and broad structure, which underlies the scagnostics suite.
-Early indexes, based on density estimation could be programmed to detect fine or large structure by varying the binwidth.
-This is examined by using a range of structure in the simulated data examples.
-
-- **rotation invariance**: The orientation of structure within a projection plane should not change the index value.
-This is especially important when using the projection pursuit guided tour, because the tour path is defined between planes, along a geodesic path, not bases within planes.
-If a particular orientation is more optimal, this will get lost as the projection shown pays no attention to orientation.
-@BCAH05 describes alternative interpolation paths based on Givens and Householder rotations which progress from basis to basis.
-It may be possible to ignore rotation invariance with these interpolations but there isn't a current implementation, primarily because the within-plane spin that is generated is distracting from a visualization perspective.
-Rotation invariance is checked for the proposed PPIs by rotating the structured projection, within the plane.
-
-- **speed**: Being fast to compute allows the index to be used in real-time in a guided tour, where the optimization can be watched.
-When the computations are shifted off-line, to watch in replay, computation times matter less.
-This is checked  by comparing times for benchmark scenarios with varying sample size.
-
-## Simulation study setup
-
-### Data construction
-\label{sec:dataOv}
-
-Three families of data simulations are used for examining the behavior of the index functions.
-Each generates  structure in two variables, with the remaining variables containing various types of noise.
-This is a very simple construction, because there is no need for projection pursuit to find the structure, one could simply use the PPIs on pairs of variables.
-However, it serves the purpose to also evaluate the PPIs.
-The three data families are explained below.
-In each set, $n$ is used for the number of points, $p$ is the number of dimensions, and $d=2$ is the projection dimension.
-\textcolor{red}{The three structures were selected to cover both functional and non-functional dependence, different types of nuisance distributions and different structure size and squintability properties.}
-
-- **pipe**: nuisance directions are generated by sampling independently from a uniform distribution between $[-1,1]$, and the circle is generated by sampling randomly on a 2D circle, and adding a small radial noise.
-The circle should be easy to see by some indices because it is large structure, but the nonlinearity creates a complication.
-- **sine**:  nuisance directions are generated by sampling independently from a standard normal distribution, and the sine curve is generated by $x_p = \sin(x_{p-1}) + \mathrm{jittering}$.
-The sine is a medium nonlinear structure, which should be visible to multiple indices.
-- **spiral**:  nuisance directions are generated by sampling independently from a normal distribution, and the structure directions are sampled from an Archimedean spiral, i.e.
-$r = a + b \theta$, with $a=b=0.1$ and we sample angles $\theta$ from a normal distribution with mean 0 and variance $2\pi$, giving a spiral with higher densities at lower radii.
-The absolute value of $\theta$ fixes the direction of the spiral shape.
-This is fine structure which is only visible close to the optimal projection.
 
@@
 ```
-
-```{r data, fig.width=6, fig.height=7, fig.cap="Scatterplots of pairs of variables from samples of each family, showing the nuisance variables and structured variables.", out.width = "0.8\\textwidth"}
-dsn <- spiral1000 %>% select(V1, V2) %>%
-  rename(X1=V1, X2=V2) %>%
-  mutate(family="spiral", type="nuisance")
-dss <- spiral1000 %>% select(V5, V6) %>%
-  rename(X1=V5, X2=V6) %>%
-  mutate(family="spiral", type="structured")
-dpn <- pipe1000 %>% select(V1, V2) %>%
-  rename(X1=V1, X2=V2) %>%
-  mutate(family="pipe", type="nuisance")
-dps <- pipe1000 %>% select(V5, V6) %>%
-  rename(X1=V5, X2=V6) %>%
-  mutate(family="pipe", type="structured")
-din <- sin1000 %>% select(V1, V2) %>%
-  rename(X1=V1, X2=V2) %>%
-  mutate(family="sine", type="nuisance")
-dis <- sin1000 %>% select(V5, V6) %>%
-  rename(X1=V5, X2=V6) %>%
-  mutate(family="sine", type="structured")
-d <- bind_rows(dsn, dss, dpn, dps, din, dis)
-ggplot(d, aes(x=X1, y=X2)) + geom_point(alpha=0.3) + 
-  facet_grid(type~family) + theme(aspect.ratio=1) +
-  xlab("") + ylab("")
-```
-
-```{r indexTable, echo=FALSE}
-tableTemp <- function(){
-  tibble(
-    index=character(),
-    data=character(),
-    size=numeric(),
-    type=character(),
-    value5=numeric(),
-    value95=numeric()
-  )
-}
-
-getSample <- function(dataT, size){
-  if (dataT == "Pipe"){
-    ret <- pipeData(p = size, n = 6) %>% scale() %>% as_tibble()
-  }
-  if (dataT == "Spiral"){
-    ret <- spiralData(p = size, n = 6) %>% scale() %>% as_tibble()
-  }
-  if (dataT == "Sine"){
-    ret <- sinData(p = size, n = 6) %>% scale() %>% as_tibble()
-  }
-  return(ret)
-}
-
-getIdx <- function(dataT, size, type){
-  convex <- numeric(length = 100)
-  skinny <- numeric(length = 100)
-  stringy <- numeric(length = 100)
-  micidx <- numeric(length = 100)
-  ticidx <- numeric(length = 100)
-  dcor2D <- numeric(length = 100)
-  splines2D <- numeric(length = 100)
-  holesI <- numeric(length = 100)
-  for(i in 1:100){
-    cSample <- getSample(dataT, size)
-    if(type == "noise"){
-      x = as.numeric(cSample$V1)
-      y = as.numeric(cSample$V2)
-    } else{
-      x = as.numeric(cSample$V5)
-      y = as.numeric(cSample$V6)
-    }
-    scag <- scagnostics(x=x, y=y)$s
-    midx <- mine(x, y, est = "mic_e")
-    skinny[i] <- scag["Skinny"]
-    stringy[i] <- scag["Stringy"]
-    convex[i] <- scag["Convex"]
-    micidx[i] <- midx$MIC
-    if (size == 100) {ticidx[i] <- midx$TIC / 16}
-    if (size == 1000) {ticidx[i] <- midx$TIC / 148}
-    dcor2D[i] <- dcor2d(x=x, y=y)
-    splines2D[i] <- splines2d(x,y)
-    holesI[i] <- holes(matrix(c(x,y), ncol = 2))
-
-  }
-  #sortedList <- lapply(list(skinny, stringy, convex, micidx, ticidx, dcor2D, splines2D, holesI), sort)
-  skinny <- sort(skinny)
-  stringy <- sort(stringy)
-  convex <- sort(convex)
-  micidx <- sort(micidx)
-  ticidx <- sort(ticidx)
-  dcor2D <- sort(dcor2D)
-  splines2D <- sort(splines2D)
-  holesI <- sort(holesI)
-  res <- tableTemp() %>%
-    add_row(index="skinny", data=dataT, size=size,
-            type=type,
-            value5=skinny[5], value95=skinny[95]) %>%
-    add_row(index="stringy", data=dataT, size=size,
-            type=type,
-            value5=stringy[5], value95=stringy[95]) %>%
-    add_row(index="convex", data=dataT, size=size,
-            type=type,
-            value5=convex[5], value95=convex[95]) %>%
-    add_row(index="MICe", data=dataT, size=size,
-            type=type,
-            value5=micidx[5], value95=micidx[95]) %>%
-    add_row(index="TICe", data=dataT, size=size,
-            type=type,
-            value5=ticidx[5], value95=ticidx[95]) %>%
-    add_row(index="dcor2D", data=dataT, size=size,
-            type=type,
-            value5=dcor2D[5], value95=dcor2D[95]) %>%
-    add_row(index="splines2D", data=dataT, size=size,
-            type=type,
-            value5=splines2D[5], value95=splines2D[95]) %>%
-    add_row(index="holes", data=dataT, size=size,
-            type=type,
-            value5=holesI[5], value95=holesI[95])
-}
-
-if(!file.exists("cache/indexTable.rda")){
-  set.seed(505)
-  res <- tableTemp()
-  for(ds in c("Pipe", "Sine", "Spiral")){
-    for (size in c(100, 1000)){
-      for (type in c("structure", "noise")){
-        res <- bind_rows(res,getIdx(ds, size, type))
-      }
-    }
-  }
-  res <- res %>% arrange(data, size, index, type) %>% select(data, size, index, type, value5, value95)
-  save(res, file = "cache/indexTable.rda")
-} else {
-  load("cache/indexTable.rda")
-}
-#rescale holes index first
-res <- res %>%
-  rowwise() %>%
-  mutate(value5 = if_else(index=="holes", rescaleHoles(value5), value5)) %>%
-  rowwise() %>%
-  mutate(value95 = if_else(index=="holes", rescaleHoles(value95), value95))
-resp <- res %>% filter(size == 1000, data == "Pipe") %>%
-  select(index, type, value5, value95) %>%
-  rename(lower=value5, upper=value95)
-resi <- res %>% filter(size == 1000, data == "Sine") %>%
-  select(index, type, value5, value95) %>%
-  rename(lower=value5, upper=value95)
-ress <- res %>% filter(size == 1000, data == "Spiral") %>%
-  select(index, type, value5, value95) %>%
-  rename(lower=value5, upper=value95)
-res_all <- bind_cols(resp, resi, ress) %>%
-  select(-index1, -type1, -index2, -type2) %>%
-  mutate(indexF = factor(index, c("holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe"))) %>%
-  arrange(indexF) %>%
-  select(-indexF, -index) %>%
-  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", "")) %>%
-  select(index, everything())
-# Good resource for doing tables
-# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
-kable(res_all, format = "latex",  caption = "Comparison of index values between noise projections and structured projections for sample size 1000, using 5th and 95th percentiles from 100 simulated sets.
-Most indices have much larger values for the structured projections, except for convex.",
-      col.names=c("Index","", "lower", "upper", "lower", "upper", "lower", "upper"),
-      digits=2, booktabs = T) %>% 
-      add_header_above(c(" ", " ", "pipe" = 2, "sine" = 2, "spiral"=2)) %>%
-  group_rows("", 1, 2, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 3, 4, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 5, 6, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 7, 8, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 9, 10, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 11, 12, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 13, 14, latex_gap_space = "-0.2cm") %>%
-  group_rows("", 15, 16, latex_gap_space = "-0.2cm") %>% 
-  #row_spec(1:16, font_size = 10) %>%
-  kable_styling(position = "center") #, font_size = 7) 
-```
-
-\textcolor{red}{For simplicity in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
-}
-The structured projection is in variables 5 and 6 ($x_5, x_6$).
-Two samples sizes are used: $n=(100, 1000)$.
-All variables are standardized to have mean 0 and standard deviation 1.
-Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables.
-Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000.
-The lower and upper show the 5th and 95th percentile of values.
-The holes index is sensitive only to the pipe distribution.
-All other indexes, except convex show distinctly higher values for the structured projections.
-The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs.
-The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix).
-Similarly, the TIC index is re-scaled depending on sample size.
-
-### Property assessment {#sec:propAss}
-
-The procedures for assessing the PPI properties of smoothness, squintability, flexibility, rotation invariance, and speed examined for samples from the family of data sets are:
-
-1.
-Compute the PPI values on the tour path along an interpolation between pairs of nuisance variables, $x_1 - x_2$ to $x_3 - x_4$.
-The result is ideally a smooth change in low values.
-This checks the smoothness property.
-2.
-Change to a tour path between a pair of nuisance variables $x_1 - x_2$ and the structured pair of variables $x_5 - x_6$ via the intermediate projection onto $x_1 - x_5$, and compute the PPI along this.
-This examines the squintability, and smoothness.
-If the function is smooth and slowly increases towards the structured projection, then the structure is visible from a distance.
-3.
-Use the guided tour to examine the ease of optimization.
-This depends on having a relatively smooth function, with structure visible from a distance.
-One index is optimized to show how effectively the maximum is attained, and the values for other PPIs is examined along the same path, to examine the similarity between PPIs.
-4. Rotation invariance is checked by computing PPIs on rotations of the structured projection. 
-5. Computational speed for the selected indexes is examined on a range of sample sizes.
-
-```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters $x_1$ and $x_2$, i.e.
-showing a typical uninformed view for each dataset.", eval=FALSE}
-pL <- list()
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
-  pC <- ggplot(ds, aes(V1, V2)) +
-    geom_point() +
-    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
-  pL[[i]] <- pC
-  i <- i+1
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
-             ncol=2)
-```
-
-```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters $x_5$ and $x_6$, i.e.
-showing the special view built into the datasets.", eval=FALSE}
-pL <- list()
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
-  pC <- ggplot(ds, aes(V5, V6)) +
-    geom_point() +
-    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
-  pL[[i]] <- pC
-  i <- i+1
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
-             ncol=2)
-```
-
-```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered.
-Apart from statistical fluctuations we see that for the Pipe datasets the parameters $x_5$ and $x_6$ are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine (and Spiral) dataset the parameter $x_6$ (and $x_5$) follows a very distinct distribution.", eval=FALSE}
-pL <- list()
-i <- 1
-pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
-for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
-  distData <- gather(ds, variable, value)
-  pC <- ggplot(distData, aes(value, color=variable)) +
-    geom_density() +
-    ggtitle(pLabels[i])
-  pL[[i]] <- pC
-  i <- i+1
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
-             ncol=2)
-```
-
-## PPI traces over a tour sequence of interpolated nuisance projections
-\label{sec:smooth}
-
-Figure \ref{fig:plotV1V2toV3V4} shows the traces representing the index values calculated across a tour between a pair of nuisance projections.
-\textcolor{red}{The tour path is generating by interpolating between the two independent nuisance planes, i.e. from the projection onto $x_1$-$x_2$ to one onto $x_3$-$x_4$.}
-The range of each axis is set to be the limits of the index, as might be expected over many different data sets, 0 to 1.
-Each projection in the interpolation will also be noise.
-Two different sample sizes are show, $n=100$ as a dashed line, and $n=1000$ as a solid line.
-The ideal trace is a smooth function, with relatively low values, and no difference between the sample sizes.
-A major feature to notice is that the scagnostics produce noisy functions, which is problematic, because small changes in the projection result in big jumps in the value.
-This will make them difficult to optimize.
-On the other hand holes, dcor2d, splines2d, MIC and TIC are relatively smooth functions.
-
-Several of the indexes are sensitive to sample size also, the same structured projection with differing numbers of points, produces different values.
-
-```{r getProj}
-getProj <- function(df, tPath, nameStr, size){
-  sc <- tibble(
-    holes=numeric(),
-    cmass=numeric(),
-    convex=numeric(),
-    skinny=numeric(),
-    stringy=numeric(),
-    dcor2d=numeric(),
-    splines2d=numeric(),
-    MIC=numeric(),
-    TIC1=numeric(),
-    t=numeric(),
-    name=character(),
-    size=numeric())
-  n <- length(tPath)
-  for (i in 1:n) {
-    dprj <- as.matrix(df) %*% tPath[[i]]
-    scagRes <- scagnostics(dprj)
-    dcorRes <- dcor2d(dprj[,1], dprj[,2])
-    splineRes <- splines2d(dprj[,1], dprj[,2])
-    mineRes <- mine(dprj[,1], dprj[,2])
-    holesRes <- holes(dprj)
-    cmassRes <- cmass(dprj)
-    sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
-                  convex=scagRes[,"Convex"], skinny=scagRes[,"Skinny"],
-                  stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
-                  splines2d=splineRes, MIC=mineRes$MIC,
-                  TIC1=mineRes$TIC, t=i, name=nameStr, size=size)
-  }
-  if (size == 100) {maxTIC <- 16}
-  if (size == 1000) {maxTIC <- 148}
-  sc <- sc %>%
-    mutate(TIC = TIC1/maxTIC) %>%
-    select(-TIC1)
-  return(sc)
-}
-```
-
-```{r V1V2toV3V4}
-if(!file.exists("cache/V1V2toV3V4.rda")){
-  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
-  m2 <- matrix(c(0,0,1,0,0,0,0,0,0,1,0,0),ncol=2)
-  #this is silly but seems that first two entries are being ignored, so need some fake entries
-  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
-  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
-  t1 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
-  t1full <- as.list(interpolate(t1))
-  fullRes <- getProj(sin100, t1full, "sin", 100) %>%
-    rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
-    rbind(getProj(spiral100, t1full, "spiral", 100)) %>%
-    rbind(getProj(spiral1000, t1full, "spiral", 1000)) %>%
-    rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, t1full, "pipe", 1000))
-  save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
-} else {
-  load("cache/V1V2toV3V4.rda")
-}
-```
-
-```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="PPIs for  projections along a geodesic interpolation between two nuisance projections.
-All projections would be nuisance so the PPI are ideally low and smooth, with little difference between sample sizes (solid lines: $n=1000$; dashed: $n=100$).
-The scagnostic PPIs are noisy.
-Some indexes have distinct differences in values between sample sizes.
-(This is not an optimization path, but an interpolation containing 41 projections between two known projections.)"}
-namesFixed <- str_replace_all(fullRes$name, "sin", "sine")
-fullResMelt <- fullRes %>%
-  mutate(name = namesFixed) %>%
-  mutate(convex = 1-convex) %>%
-  rename("1-convex"=convex) %>%
-  rowwise() %>%
-  mutate(holes = rescaleHoles(holes)) %>%
-  gather(PPI, value, -t, -name, -size) %>%
-  filter(PPI != "cmass") %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "cmass", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
-
-fullResMelt %>% 
-  ggplot(aes(x=t, y=value)) + 
-  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
-  ylim(c(0,1)) +
-  facet_grid(PPI~name) + 
-  theme(legend.position="none") + xlab("Sequence of projections (t)") + ylab("PPI value")
-```
-
-## PPI traces over a tour sequence between nuisance and structured projections
-\label{sec:squintability}
-
-Figure \ref{fig:wIntermediate} shows the PPIs for a tour sequence between a nuisance and structured projection. A long sequence is generated where the path 
-\textcolor{red}{interpolates between projections onto}
-$x_1$-$x_2$, $x_1$-$x_5$, $x_5$-$x_6$, in order to see some of the intricacies of holes index.
-Sample size is indicated by line type: dashed being $n=100$ and solid is $n=1000$.
-The beginning of the sequence is the nuisance projection and the end is the structured projection.
-The index values for most PPIs increases substantially nearing the structured projection, indicating that they "see" the structure.
-Some indexes see all three structures: scagnostics, MIC and TIC,  which means that they are flexible indexes capable of detecting a range of structure.
-[@Grimm2016]'s indexes, dcor2d and splines2d, are excellent for detecting the sine, and they can see it from far away, indicated by the long slow increase in index value.
-The holes index easily detects the pipe, and can see it from a distance, but also has local maxima along the tour path.
-The scagnostic index, stringy, can see the structure but is myopic, only when it is very close.
-Interestingly the scagnostic, skinny, sees the spiral from a distance.
-
-```{r wIntermediatePath}
-if(!file.exists("cache/wIntermediate.rda")){
-  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
-  mInt <- matrix(c(1,0,0,0,0,0,0,0,0,0,1,0),ncol=2)
-  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
-  #this is silly but seems that first two entries are being ignored, so need some fake entries
-  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
-  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
-  t3 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,mInt,m2)))
-  t3full <- as.list(interpolate(t3))
-  fullResPlanned <- getProj(sin100, t3full, "sin", 100) %>%
-    rbind(getProj(sin1000, t3full, "sin", 1000)) %>%
-    rbind(getProj(spiral100, t3full, "spiral", 100)) %>%
-    rbind(getProj(spiral1000, t3full, "spiral", 1000)) %>%
-    rbind(getProj(pipe100, t3full, "pipe", 100)) %>%
-    rbind(getProj(pipe1000, t3full, "pipe", 1000))
-  save(t3, t3full, fullResPlanned, file = "cache/wIntermediate.rda")
-} else {
-  load("cache/wIntermediate.rda")
-}
-```
-
-```{r wIntermediate, fig.width=6, fig.height=7, fig.cap="PPIs for projections along an interpolation between nuisance and structured projections, following $x_1$-$x_2$ to $x_1$-$x_5$ to $x_5$-$x_6$ (solid lines: $n=1000$; dashed: $n=100$).
-The vertical blue line indicates the position of the projection onto $x_1$-$x_5$ in the sequence.
-Peaks at the end of the sequence indicate the index sees the structure.
-The scagnostics, MIC and TIC see all three structures, so are more flexible for general pattern detection.
-Holes only responds to the pipe, and is a multimodal function for this data with a local maximum at $x_1$-$x_5$.
-(This is not an optimization path, but an interpolation containing 59 projections between three known projections.) "}
-intermediatePlane <- which(attributes(interpolate(t3))$new_basis)[2]
-namesFixed <- str_replace_all(fullResPlanned$name, "sin", "sine")
-fullResMeltPlanned <- fullResPlanned %>%
-  mutate(name = namesFixed) %>%
-  mutate(convex = 1-convex) %>%
-  rename("1-convex"=convex) %>%
-  rowwise() %>%
-  mutate(holes = rescaleHoles(holes)) %>%
-  gather(PPI, value, -t, -name, -size) %>%
-  filter(PPI != "cmass") %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
-fullResMeltPlanned %>% 
-  ggplot(aes(x=t, y=value)) + 
-  geom_line(aes(group=size, linetype=factor(size, levels=c("1000", "100")))) + 
-  ylim(c(0,1)) +
-  geom_vline(xintercept=intermediatePlane, color="blue", alpha=0.3) +
-  facet_grid(PPI~name) + 
-  theme(legend.position="none") + xlab("Sequence of projections (t)") + ylab("PPI value")
-```
-
-## Optimization check with the guided tour
-\label{sec:guided}
-
-\textcolor{red}{Before applying the new index functions with the guided tour on real examples, we test them on the simulated dataset to understand the performance of the optimization.}
-The guided tour combines optimization with interpolation between pairs of planes.
-Target planes of the path are chosen to maximize the PPI.
-There are three derivative-free optimization methods available in the guided tour: \verb# search_better_random# (1), \verb# search_better# (2), and \verb# search_geodesic# (3).
-Method 1 casts a wide net randomly generating projection planes, computing the PPIs and keeping the best projection, and method 2 conducts a localized maximum search.
-Method 3 is quite different: a local search is conducted to determine a promising direction, and then this direction is followed until the maximum in that direction is found.
-For all methods the optimization is iterative, the best projections form target planes in the tour, the tour path is the interpolation to this target, and then a new search for a better projection is made, followed by the interpolation.
-For each projection during the interpolation steps, the PPI is recorded.
-
-The stopping rule is that no better projections are found after a fixed number of tries, given a fixed tolerance value measuring difference.
-For method 1 and 2 two additional parameters control the optimization: the search window $\alpha$, giving the maximum distance from the current plane in which projections are sampled, and the cooling factor, giving the incremental decrease in search window size.
-Method 3 in principle also has two free parameters, which are however fixed in the current implementation.
-The first is the small step size used when evaluating the most promising direction, it is fixed to 0.01, and the second parameter being the window over which the line search is performed, fixed to $\pm \pi/4$ away from the current plane.
-
-For distributions and indexes with smooth behavior and good squintability, method 3 is the most effective method for optimization.
-If these two criteria are not met the method may still be useful, but only given an informed starting projection.
-In such cases we can follow a method similar to that proposed by Friedman [@f87]: we break the optimization in two distinct steps.
-A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum.
-Note that this likely requires large number of tries, especially as dimension increases, since most randomly picked planes will not be interesting.
-The second step uses method 3 starting from the approximate maximum, which will take small steps to refine the result to be closer to the global maximum.
-
-### Looking down the pipe
-
-Despite the simple structure, the pipe is relatively difficult for the PPIs to find.
-For the TIC index, there is a fairly small squint angle.
-For the holes index, there are several local maxima, that divert the optimizer.
-There is a hint of this from Figure \ref{fig:wIntermediate} because the initial projection (left side of trace) of purely noise variables has a higher index value than the linear combinations of noise and structured variables along the path.
-The uniform distribution was used to generate the noise variables, which has a higher PPI value than a normal distribution, yielding the higher initial value.
-In addition, a local maximum is observed whenever the pair of variables is one structured variable and one noise variable, because there is a lighter density in the center of the projection.
-
-The optimization is done in two stages, a scouting phase using method 2, and a refinement stage using method 3.
-For the scouting we use $\alpha = 0.5$ and stopping condition of maximum 5000 tries, and we optimized the TIC index.
-
-Figure \ref{fig:pipeFirstRun} shows the target projections (points)  selected during the scouting with method 2 on the TIC index.
-The focus is on the target projections rather than the interpolation between them, because the optimization is done off-line, and only the targets are used for the next step.
-The horizontal distance between the points in the plot reflects the relative geodesic distance between the planes in the 6D space.
-All of the other indexes are shown for interest.
-The TIC index value is generally low for this data, although it successfully detects the pipe.
-The holes, convex, skinny, and to some extent MIC, mirror the TIC performance.
-The holes differs in that it has some intermediate high values which are likely the indication of multi-modality of this index on this data.
-
-The final views obtained in each of the two stages are compared in the Appendix, see Fig. \ref{fig:testpipe}.
-
-```{r findpipe, results="hide"}
-if(!file.exists("cache/findpipe.rda")){
-  set.seed(1984)
-  pipeResc <- as.matrix(pipe1000) 
-  pipeTour <- save_history(pipeResc,
-                          guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
-                                     cooling=1, alpha = 0.5, max.tries = 5000))
-  pipeTourFull <- as.list(interpolate(pipeTour))
-  save(pipeResc, pipeTour, pipeTourFull, file = "cache/findpipe.rda")
-} else {
-  load("cache/findpipe.rda")
-}
-```
-
-```{r eval=FALSE}
-# Quick glimpse of optimisations
-myholes <- function(mat) {
-    n <- nrow(mat)
-    d <- ncol(mat)
-    num <- 1 - 1/n * sum(exp(-0.5 * rowSums(mat^2)))
-    den <- 1 - exp(-d/2)
-    (num/den - 0.7) * 1.43
-}
-quartz()
-set.seed(sample(10000:50000, 1))
-animate_xy(pipeResc, guided_tour(scagIndex("Skinny"), search_f = tourr:::search_better, cooling=0.9, alpha = pi, max.tries = 500))
-animate_xy(pipeResc, guided_tour(mineIndex("MIC"), search_f = tourr:::search_better, cooling=1, alpha = pi, max.tries = 500))
-animate_xy(pipeResc, guided_tour(mineIndex("TIC"), search_f = tourr:::search_better, cooling=1, alpha = pi, max.tries = 5000))
-animate_xy(pipeResc, guided_tour(holes, search_f = tourr:::search_geodesic,
-      cooling=0.8, alpha = pi, max.tries = 500))
-```
-
-```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="PPIs for a sequence of projections produced by scouting for the pipe using optimization method 2 on the TIC index.
-Other PPI values are shown for interest.
-Only the values on the target planes are shown.
-Despite the small maximum value of TIC for this data, it identifies the pipe."}
-if(!file.exists("cache/pipeFirstRun.rda")){
-  pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
-  #pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
-  #never used?
-  #save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
-  save(pipeTourRes, file = "cache/pipeFirstRun.rda")
-} else {
-  load("cache/pipeFirstRun.rda")
-}
-newBasisPipeTour <- pipeTourRes %>%
-  mutate(convex = 1-convex) %>%
-  rename("1-convex"=convex) %>%
-  rowwise() %>%
-  mutate(holes = rescaleHoles(holes)) %>%
-  select(-cmass) %>%
-  gather(variable, value, -t, -name, -size) %>%
-  filter(t %in% which(attributes(interpolate(pipeTour))$new_basis)) %>%
-  mutate(variable = fct_relevel(variable, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
-
-ggplot(newBasisPipeTour, aes(x=t, y=value)) +
-  geom_point(mapping=aes(color=(variable=="TIC"))) + 
-  geom_line(mapping=aes(color=(variable=="TIC"))) + 
-  scale_colour_manual(values = c("black", "red")) +
-  facet_wrap(~variable, ncol=1, strip.position = "right") +
-  guides(color=FALSE) +
-  ylim(c(0,1)) +
-  xlab("Number of projections") +
-  ylab("PPI value") 
-```
-
-```{r refinepipe, results="hide"}
-if(!file.exists("cache/refinepipe.rda")){
-  iLast <- length(pipeTourFull)
-  fProj <- pipeTourFull[[iLast]]
-  pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
-  pipeTourFull2 <- as.list(interpolate(pipeTour2))
-  pipeTourRes2 <- getProj(pipe1000, pipeTourFull2, "Pipe", 1000)
-  #pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
-  #save(pipeTour2, pipeTourFull2, pipeTourRes2, pipeTourResMelt2, file = "cache/refinepipe.rda")
-  save(pipeTour2, pipeTourFull2, pipeTourRes2, file = "cache/refinepipe.rda")
-} else {
-  load("cache/refinepipe.rda")
-}
-```
-
-### Finding sine waves
-
-Given the patterns in Figure \ref{fig:wIntermediate} it would be expected that the sine could be found easily, using only optimization method 3 with the splines2d, dcor2d, MIC or TIC indexes.
-This is examined in Figure \ref{fig:findsine}.
-Optimization is conducted using the splines2d index, and the trace of the PPI over the optimization is shown, along with the PPI values for the other indexes over that path.
-The vertical blue lines indicate anchor bases, where the optimizer stops, and does a new search.
-The distance between anchor planes is smaller as the maximum is neared.
-
-The only complications arise from a lack of rotation invariance of the splines2d index.
-It is not easily visible here, but it is possible that the best projection will have a higher PPI.
-The index changes depending on the basis used to define the plane, but the geodesic interpolation conducted by the tour uses any suitable basis to describe the plane, ignoring that which optimizes the PPI.
-This is discussed in section \ref{sec:rot}.
-
-```{r findsine, fig.width=6, fig.height=7, out.width = "0.8\\textwidth", fig.cap="PPIs for sequence of projections produced by a guided tour optimizing  the splines2d index, using optimization method 3, for the sine data, with $n=1000$.
-Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima.
-The sine is found relatively easily, by splines2d, and it is indicated that MIC, TIC, dcor2d and convex would also likely find this structure.", results="hide"}
-set.seed(2018)
-if(!file.exists("cache/findsine.rda")){
-  sineResc <- as.matrix(sin1000)
-  sineTour <- save_history(sineResc,
-                          guided_tour(splineIndex()))
-  sineTourFull <- as.list(interpolate(sineTour))
-  sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
-  #sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
-  #save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
-  save(sineResc, sineTour, sineTourFull, sineTourRes, file = "cache/findsine.rda")
-} else {
-  load("cache/findsine.rda")
-}
-anchorIdx <- as_tibble(which(attributes(interpolate(sineTour))$new_basis))
-sineTourResMelt <-sineTourRes %>%
-  mutate(convex = 1-convex) %>%
-  rename("1-convex"=convex) %>%
-  rowwise() %>%
-  mutate(holes = rescaleHoles(holes)) %>%
-  select(-cmass) %>%
-  gather(variable, value, -t, -name, -size) %>%
-  mutate(variable = fct_relevel(variable, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
-
-ggplot(sineTourResMelt, aes(x=t, y=value)) +
-  geom_line(mapping=aes(color=(variable=="splines2d"))) + 
-  facet_wrap(~variable, ncol=1, scales = "free_y") +
-  scale_colour_manual(values = c("black", "red")) +
-  facet_wrap(~variable, ncol=1, strip.position = "right") +
-  geom_vline(data=anchorIdx, mapping=aes(xintercept=value), color="blue", alpha=0.3) +
-  guides(color=FALSE) +
-  ylim(c(0,1)) + xlab("Number of projections") + ylab("PPI value")
-```
-
-### Spiral detection
-
-The spiral is the most challenging structure to detect because it has a small squint angle [@posse95a], especially as the ratio of noise to structure dimensions increases.
-This is explored using optimization method 2 to scout the space for approximate maxima.
-The skinny scagnostic index is used because it was observed (Figure \ref{fig:wIntermediate}) to be sensitive to this structure, although the noisiness of the index might be problematic.
-The stringy appears to be more sensitive to the spiral, but it has a much smaller squint angle.
-
-The search is conducted for $p=4,5,6$ which would correspond to 2, 3 and 4 noise dimensions respectively.
-In addition we examine the distance between planes, using a Frobenius norm, as defined by Equation 2 of @BCAH05, and available in the ```proj_dist``` function in the tourr package, to compare searches across dimensions.
-The distance between planes is related to squint angle, how far away from the ideal projection can the structure be glimpsed.
-We estimate the squint angle depending on the number of noise dimensions in the appendix.
-In order for the optimizer to find the spiral, the distance between planes would need to be smaller than the squint angle.
-Figure \ref{fig:findspiral} summarizes the results.
-When $p=4$ the scouting method effectively finds the spiral.
-Plot (a) shows the side-by-side boxplots of pairwise distances between planes examined during the optimization, for $p=4,5,6$.
-These are on average smaller for the lower dimension, and gradually increase as dimension increases.
-This is an indication of the extra computation needed to brute force find the spiral as noise dimensions increase.
-Plot (b) shows the distance of the plane in each iteration of the optimization to the ideal plane, where it can be seen that only when $p=4$ does it converge to the ideal.
-Its likely that expanding the search space should result in uncovering the spiral in higher dimensions, which however requires tuning of the stopping conditions and long run times.
-
-```{r findspiral, out.width="0.9\\textwidth", fig.cap="Guided tour optimizing the skinny index for the Sprial dataset with 1000 datapoints, with p = 4, 5, 6.
-The left plot shows the distribution of pairwise distances between planes obtained via the guided tour, the right shows the evolution of distance to the ideal plane as the index is being optimized.", results="hide", fig.width=7, fig.height=2.5}
-
-distanceDist <- function(planes, nn=F){
-  #nn could be used to turn on only distance to nearest neighbour?
-  planes <- as.list(planes)
-  maxI <- length(planes)
-  distL <- numeric(choose(maxI, 2))
-  i <- 1
-  idx <- 1
-  while(i<maxI+1){
-    j <- i+1
-    while(j<maxI+1){
-      cDist <- proj_dist(planes[[i]], planes[[j]])
-      distL[idx] <- cDist
-      j <- j+1
-      idx <- idx+1
-    }
-    i <- i+1
-  }
-  return(distL)
-}
-
-distanceToSp <- function(planes, specialPlane){
-  planes <- as.list(planes)
-  maxI <- length(planes)
-  distL <- numeric(maxI)
-  i <- 1
-  while(i<maxI+1){
-    cDist <- proj_dist(planes[[i]], specialPlane)
-    distL[i] <- cDist
-    i <- i+1
-  }
-  return(distL)
-  }
-
-if(!file.exists("cache/findspiral.rda")){
-  distDf <- tibble(d=numeric(), distance=numeric())
-  distToSp <- tibble(d=numeric(), distance=numeric(), t=numeric())
-
-  maxBases <- 100
-#  maxBases <- 1000
-  fViewL <- list()
-  allPlanes <- list()
-  i <- 1
-  for(d in c(4,5,6)){
-  #d = 6
-    set.seed(58958)
-    specialPlane <- matrix(c(rep(0,d-2),1,0,rep(0,d-1),1),ncol=2)
-    spiralD <- spiral1000 %>%
-      purrr::when(d<6 ~ select(., -V4), ~ .) %>%
-      purrr::when(d<5 ~ select(., -V3), ~ .)
-    spiralTour <- save_history(as.matrix(spiralD),
-                               guided_tour(scagIndex("Skinny"), 
-                                  search_f = tourr:::search_better,
-                                  max.tries = 5000, 
-                                  #max.tries = 50000, 
-                                  cooling = 1, alpha = pi),
-                              max_bases = maxBases)
-    #if(d==4){maxBases <- min(maxBases, length(spiralTour))}
-    nBases <- length(spiralTour)
-    distDf <- rbind(distDf, tibble(d=d, distance=distanceDist(spiralTour)))
-    distToSp <- rbind(distToSp, tibble(d=d, distance=distanceToSp(spiralTour, specialPlane), t=1:nBases))
-    fView <- as_tibble(as.matrix(spiralD) %*% as.list(spiralTour)[[nBases]])
-    fViewL[[i]] <- ggplot(fView, aes(V1, V2)) + geom_point()
-    allPlanes[[i]] <- spiralTour
-    i <- i+1
-  }
-  save(distDf, distToSp, fViewL, allPlanes, file = "cache/findspiral.rda")
-} else {
-  load("cache/findspiral.rda")
-}
-
-distDf <- distDf %>%
-  mutate(p=d) %>%
-  select(-d)
-
-distToSp <- distToSp %>%
-  mutate(p=d) %>%
-  select(-d)
-
-p1 <- ggplot(distDf, aes(group=p, y=distance, x=p)) +
-  geom_boxplot() +
-  scale_x_continuous(breaks = c(4, 5, 6)) + ggtitle("(a)")
-p2 <- ggplot(distToSp, aes(color=as.factor(p),x =t, y=distance)) +
-  geom_point() +
-  labs(y = "Distance to special view", color = "p") + ggtitle("(b)")
-
-fViewL[[1]] <- fViewL[[1]] + xlab("PP1") + ylab("PP2")
-fViewL[[2]] <- fViewL[[2]] + xlab("PP1") + ylab("PP2")
-fViewL[[3]] <- fViewL[[3]] + xlab("PP1") + ylab("PP2")
-
-#grid.arrange(p1, p2, fViewL[[1]], fViewL[[2]], fViewL[[3]], layout_matrix = rbind(c(1,2,2),c(3,4,5)))
-grid.arrange(p1, p2, layout_matrix = rbind(c(1,2,2)))
-```
-
-## Rotational invariance or not
-\label{sec:rot}
-
-Rotational invariance is examined using the sine data ($x_5$-$x_6$), computing PPI for different rotations within the 2D plane, parameterized by angle.
-Results are shown in Figure \ref{fig:rotationDep}.
-Several indexes are invariant, holes, convex and MIC, because their value is constant around rotations.
-The dcor2d, splines2d and TIC index are clearly not rotationally invariant because the value changes depending on the rotation.
-The scagnostics indexes are approximately rotationally invariant, but particularly the skinny index has some random variation depending on rotation.
-
-```{r rotationDep, out.width=".6\\textwidth", fig.cap="PPI for rotations of the sine 1000 data, to examine rotation invariance.
-Most are close to rotation invariant, except for skinny, dcor2d, splines2d and TIC."}
-if(!file.exists("cache/rotationDep.rda")){
-  sineM <- as.matrix(select(sin1000, V5, V6))
-  sc <- tibble(
-      holes=numeric(),
-      cmass=numeric(),
-      convex=numeric(),
-      skinny=numeric(),
-      stringy=numeric(),
-      dcor2d=numeric(),
-      splines2d=numeric(),
-      MIC=numeric(),
-      TIC=numeric(),
-      alpha=numeric()
-  )
-  for (a in seq(0,2*pi, pi/100)){
-    rotM <- matrix(c(cos(a), sin(a), -sin(a), cos(a)), ncol = 2)
-    dprj <- sineM %*% rotM
-    scagRes <- scagnostics(dprj)
-    dcorRes <- dcor2d(dprj[,1], dprj[,2])
-    splineRes <- splines2d(dprj[,1], dprj[,2])
-    mineRes <- mine(dprj[,1], dprj[,2])
-    holesRes <- holes(dprj)
-    cmassRes <- cmass(dprj)
-    sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
-                  convex=scagRes[,"Convex"], skinny=scagRes[,"Skinny"],
-                  stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
-                  splines2d=splineRes, MIC=mineRes$MIC,
-                  TIC=mineRes$TIC/148, alpha=a)
-  }
-  save(sc, file = "cache/rotationDep.rda")
-} else {
-  load("cache/rotationDep.rda")
-}
-
-scMelt <- sc %>% 
-  mutate(convex = 1-convex) %>%
-  rename("1-convex"=convex) %>%
-  rowwise() %>%
-  mutate(holes = rescaleHoles(holes)) %>%
-  #rowwise() %>%
-  #mutate(cmass = rescaleCmass(cmass)) %>%
-  select(-cmass) %>%
-  mutate(angle=alpha*360/(2*pi), TIC = TIC) %>%
-  select(-alpha) %>%
-  gather(PPI, value, -angle) %>%
-  mutate(PPI = fct_relevel(PPI, "holes", "1-convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
-
-ggplot(scMelt, aes(x=angle, y=value)) +
-  geom_line() + 
-  ylim(c(-1, 1.1)) +
-  facet_wrap(~PPI, ncol = 3) +
-  #geom_blank(aes(y = 0)) +
-  #geom_blank(aes(y = 1)) +
-  coord_polar() +
-  xlab("") + ylab("")
-```
-
-## Speed
-Examining the computing time as a function of sample size we find that scagnostics and splines2d are fast even for large samples, while all other index functions slow rapidly with increasing sample size.
-Detailed results are shown in Appendix \ref{app:speed}.
-
-## Parameter choices
-Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure.
-In Appendix \ref{app:param} we examine the dependence of the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
-Sensitivity to the number of bins in the MIC index is also examined, showing that tuning the parameter can improve the final result.
-
-## Index enhancement
-We identified two potential improvements.
-First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in Appendix \ref{app:enh}.
-In addition, rotation dependent indexes may be enhanced by redefining them in a rotation invariant way.
-
-## Summary
-
-Our results can be summarized by evaluating and comparing the advantages and disadvantages of each index function according to the criteria presented above.
-Such an overview is given in Table \ref{tab:summary}, listing if the criteria is fully met (\checkmark), there are some shortcomings ($\cdot$) or failure ($\times$).
-\textcolor{red}{Notice that the established holes index does not appear in the summary.}
-We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fulfilled.
-In addition the limited flexibility of most indexes highlights the importance of index selection in the projection pursuit setup.
-Table \ref{tab:summary} further suggests that there is much room for the improvement of index functions detecting unusual association between model parameters.
-
-\begin{table}
-\begin{center}
-\caption{Summary of findings, showing to what extend the considered new index functions pass the criteria for a good PPI.
-``\checkmark" symbols good behavior, ``$\cdot$" symbols some issues while ``$\times$" symbols failure on the corresponding criteria.
-Each index has particular strengths and drawbacks and selection must be guided by the considered example, see text for details.}
-\label{tab:summary}
-\begin{tabular}{|l|c|c|c|c|c|c|}
-\hline
-Index & smooth & squintability & flexible & rotation invariant & speed \\
-\hline
-convex & $\cdot$ & $\cdot$ & $\cdot$ & \checkmark & \checkmark \\
-skinny & $\cdot$ & $\cdot$ & $\cdot$ & $\times$ & \checkmark \\
-stringy & $\times$ & $\times$ & $\cdot$ & $\cdot$ & \checkmark \\
-\hline
-splines2D & \checkmark & \checkmark & $\cdot$ & $\times$ & \checkmark \\
-dcor2D & \checkmark & \checkmark & $\cdot$ & $\times$ & $\cdot$ \\
-\hline
-MIC & $\cdot$ & \checkmark  & \checkmark & $\cdot$ & $\cdot$ \\
-TIC & $\cdot$ & \checkmark  & \checkmark & $\cdot$ & $\cdot$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-# Application to physics examples
-\label{sec:phys}
-
-This section describes the application of these projection pursuit indices to find two-dimensional structure in two multidimensional gravitational waves problems.
-
-The first example contains 2538 posterior samples obtained by fitting source parameters to the observed gravitational wave signal GW170817 from a neutron star merger [@PhysRevLett119161101].
-Data has been downloaded from @ligoData.
-The fitting procedures are described in detail in @Abbott2018exr.
-We consider six parameters of physical interest (6-D) with some known relationships.
-Projection pursuit is used to find the known relationships.
-
-The second example contains data generated from a simulation study of a binary black hole (BBH) merger event, as described in @Smith:2016qas.
-There are 12 parameters (12-D), with multiple nuisance parameters.
-Projection pursuit uncovers new relationships between parameters.
-
-## Neutron star merger
-
-Figure \ref{fig:neutronStarSPLOM} in the Appendix shows the scatterplot matrix (with transparency) of the six parameters.
-(In astrophysics, scatterplot matrices are often called "corner plots" [@corner].) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables.
-From this it can be seen that m1 and m2 are strongly, and slightly, nonlinearly associated.
-Between the other variables we observe some linear association (R1, R2), some nonlinear association (L1, L2, R1, R2),  heteroskedastic variance in most variables and some bimodality (R1, L1, L2, m1, m2).
-
-The model describes a neutron star merger and contains 6 free parameters, with each neutron star described by its mass $m$ (m1, m2) and radius $R$ (R1, R2), and a so-called tidal deformability parameter $\Lambda$ (L1, L2) which is a function of the mass and radius, approximately proportional to $(m/R)^{-5}$.
-
-### Data pre-pocessing
-
-Because m1 and m2 are very strongly associated, m2 is dropped before doing PP. This relationship is obvious from the scatterplots of pairs of variables and does not need to be re-discovered by PP.
-
-All variables are scaled to range between 0 and 1.
-The purpose is that range differences in individual variables should not affect the detection  of relationships between multiple variables.
-Standardizing the range will still leave differences between the standard deviations of the variables, and for this problem this is preferred.
-Differences in the standard deviations can be important for keeping the non-linear relationships visible to PP.
-
-### Applying PP
-
-With only five parameters, a reasonable start is to examine the 5D space using a grand tour.
-This quickly shows  the strong nonlinear relationships between the parameters.
-PP is then used to extract these relationships.
-The best index for this sort of problem is the splines2d, and it is fast to compute.
-
-Figure \ref{fig:nsePlotOrig} shows the optimal projection found by splines2d, a reconstructed view obtained by manually combining parameters, and a plot of the known relationship between parameters.
-
-```{r neutronStarEq,  results="hide"}
-## Step 1: read data, rescale
-nsD <- read_csv("data/samples.csv") 
-nsM <- nsD %>%
-  select(-m2) %>% # remove m2 to avoid finding well known correlation
-  rescale() # rescaling, this returns a matrix
-## Step 2: select index function (splines2d),
-##         formatting for guided tour is done in the function splineIndex
-## Step 3: recording guided tour via the save_history function
-if(!file.exists("cache/neutronStarEq.rda")){
-  set.seed(2018)
-  nseTour <- save_history(nsM, guided_tour(splineIndex()))
-  nseTourFull <- as.list(interpolate(nseTour))
-  save(nseTour, nseTourFull, file = "cache/neutronStarEq.rda")
-} else {
-  load("cache/neutronStarEq.rda")
-}
-iLast <- length(nseTourFull)
-fProj <- nseTourFull[[iLast]]
-dProj <- as_tibble(nsM %*% fProj)
-colnames(fProj) <- c("PP1", "PP2")
-colnames(dProj) <- c("PP1", "PP2")
-fProj <- as_tibble(fProj) %>%
-  add_column(label=colnames(nsM))
-```
-
-```{r axesDrawing}
-#select cntrx, cntry for placement
-#sclx and scly should be 20% of overall range
-getCircle <- function(cntrx, cntry, sclx, scly){
-  theta <- seq(0, 2 * pi, length = 50)
-  circ <- tibble(x=cos(theta), y=sin(theta))
-  circ <- circ %>% 
-    mutate(x=x*sclx+cntrx, y=y*scly+cntry)
-}
-getAxes <- function(fProj, cntrx, cntry, sclx, scly){
-  x1 <- rep(cntrx, length(fProj$PP1))
-  y1 <- rep(cntry, length(fProj[,1]))
-  x2 <- fProj$PP1*sclx+cntrx
-  y2 <- fProj$PP2*scly+cntry
-  lab <- fProj$label
-  axes <- tibble(x1=x1, y1=y1, x2=x2, y2=y2, label=lab)
-}
-```
-
-```{r nsePlotOrig, fig.cap="Comparison of guided tour final view (left), approximation based on original parameters (middle) and expected relation based on analysis setup (right)."}
-p1 <- ggplot(dProj, aes(PP1, PP2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0,0.6,0.32,0.22), aes(x=x, y=y), color="grey") + 
-  xlim(-1.1, 0.5) + ylim(-0.2, 0.9) +
-  geom_segment(data=getAxes(fProj, 0,0.6,0.32,0.22), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj, 0,0.6,0.32,0.22), aes(x=x2, y=y2, label=label), size=2.5)
-p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
-p3 <- ggplot(nsD, aes(R1/m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
-grid.arrange(p1, p2, p3, ncol=3) 
-```
-
-To further investigate relationships between parameters, $L1$ is removed and PP with the splines2D is applied to the remaining four parameters.
-The dependence of $L2$ on the mass and radius of the lighter neutron star, is revealed (Figure \ref{fig:nseRemL1} left plot).
-A manual reconstruction shows this is a relationship between L2, R1, R2 and m1 (middle plot), but it is effectively the known relationship between L2, R2 and m2 (right plot) -- m2 is latently in the relationship though m1.
-
-```{r nseRemL1, fig.cap="Removing L1 and optimize again of the remaining parameters, where m2 remains removed from the set.
-Because of parameter correlations we can recover clear description of L2 as a function of the other parameters, despite m2 missing."}
-## Step 4: we explore reducing dimensionality, here dropping L1
-## then repeat steps 2+3 as before
-nsM2 <- nsD %>%
-  select(-m2, -L1) %>% # remove m2 to avoid finding well known correlation
-  rescale()
-if(!file.exists("cache/neutronRemL1.rda")){
-  set.seed(2018)
-  nseTour2 <- save_history(nsM2, guided_tour(splineIndex()))
-  save(nseTour2, file = "cache/neutronRemL1.rda")
-} else{
-  load("cache/neutronRemL1.rda")
-}
-tL <- as.list(nseTour2)
-iLast <- length(tL)
-fProj <- tL[[iLast]]
-dProj <- as_tibble(nsM2 %*% fProj)
-colnames(fProj) <- c("PP1", "PP2")
-colnames(dProj) <- c("PP1", "PP2")
-fProj <- as_tibble(fProj) %>%
-  add_column(label=colnames(nsM2))
-p1 <- ggplot(as_tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.1)+ theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.5,0.6,0.3,0.18), aes(x=x, y=y), color="grey") + 
-  xlim(0, 1.5) + ylim(-0.1, 0.8) +
-  geom_segment(data=getAxes(fProj, 0.5,0.6,0.3,0.18), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj, 0.5,0.6,0.3,0.18), aes(x=x2, y=y2, label=label), size=2.5)
-p2 <- ggplot(as_tibble(nsM2), aes(m1+R1+1.5*R2, L2)) + geom_point(alpha=0.1)+ theme(aspect.ratio=1)
-p3 <- ggplot(nsD, aes(R2/m2, L2))+ geom_point(alpha=0.1)+ theme(aspect.ratio=1)
-grid.arrange(p1, p2, p3, ncol=3)
-```
-
-## Black hole simulation
-
-This data contains posterior samples from simulation from a model describing a binary black hole (BBH) merger event. There are twelve model parameters. Flat priors are used for most model parameters.
-
-Figure \ref{fig:bbhSimulation} in the Appendix shows a scatterplot matrix, of nine of the twelve parameters.
-(Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they are uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time.
-The first two describe the position of the event in the sky, and time is the merging time (in GPS units).
-Because of the elliptical relationship between dec and time, the TIC index is used for PP, even though it is slow to compute.
-Between the other parameters, the main structure seen is multimodality and some skewness.
-These patterns are representative of the likelihood function, since most priors are flat, or built to capture growth with volume rather than distance.
-
-### Data pre-processing
-
-The analysis is conducted on 11 of the twelve parameters. One variable is removed, m2, because it is so strongly associated with m1. All parameters are scaled into the range 0 to 1.
-
-### Applying PP
-
-#### Exploring 11D with all PP indexes
-
-All seven PP indexes are applied to the data.
-Figure \ref{fig:bbhGuided} showing the projections that maximize three of the indexes.
-TIC and splines2d indexes identify very similar projections, that are based on the three parameters, dec, time and ra.
-This is to be expected based on the pairwise scatterplots.
-On the other hand, the 1-convex index finds a very different view, but this is because the optimization doesn't adequately reach a maximum for this index.
-
-```{r bbhGuided, results="hide", fig.cap="Projections corresponding to the maxima of three indices: TIC, splines2D and 1-convex.
-Projections a, b found by TIC and splines2d are very similar, and involve the same three parameters, ra, dec and time.
-The 1-convex index finds a very different view.", fig.height = 3.2, dev = "png", dpi=300}
-# Step 1: data cleaning and rescaling
-bbhD <- read_csv("data/posterior_samples.csv")
-bbhM <- rescale(select(bbhD,-m2, -chi_eff))
-# Step 2: select index function (TIC) formatted in mineIndex
-# Step 3: record guided tour with save_history
-if(!file.exists("cache/bbhGuidedTIC.rda")){
-  set.seed(2018)
-  bbhTour1 <- save_history(bbhM, guided_tour(mineIndex("TIC")))
-  bbhTourFull1 <- as.list(interpolate(bbhTour1))
-  save(bbhTour1, bbhTourFull1, file = "cache/bbhGuidedTIC.rda")
-} else {
-  load("cache/bbhGuidedTIC.rda")
-}
-# Step 4: repeating steps 2+3 for different index functions (splines2d, 1-convex)
-if(!file.exists("cache/bbhGuidedSpline.rda")){
-  set.seed(2018)
-  bbhTour2 <- save_history(bbhM, guided_tour(splineIndex()))
-  bbhTourFull2 <- as.list(interpolate(bbhTour2))
-  save(bbhTour2, bbhTourFull2, file = "cache/bbhGuidedSpline.rda")
-} else {
-  load("cache/bbhGuidedSpline.rda")
-}
-if(!file.exists("cache/bbhGuidedConvex.rda")){
-  set.seed(2018)
-  bbhTour3 <- save_history(bbhM, guided_tour(invConvexIndex()))
-  bbhTourFull3 <- as.list(interpolate(bbhTour3))
-  save(bbhTour3, bbhTourFull3, file = "cache/bbhGuidedConvex.rda")
-} else {
-  load("cache/bbhGuidedConvex.rda")
-}
-
-iLast <- length(bbhTourFull1)
-fProj1 <- bbhTourFull1[[iLast]]
-dProj1 <- as_tibble(bbhM %*% fProj1)
-colnames(fProj1) <- c("PP1", "PP2")
-colnames(dProj1) <- c("PP1", "PP2")
-fProj1 <- as_tibble(fProj1) %>%
-  add_column(label=colnames(bbhM))
-p1 <- ggplot(dProj1, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.4,0.5,0.26,0.24), aes(x=x, y=y), color="grey") + 
-  xlim(0.1, 1.4) + ylim(-0.3, 0.9) +
-  geom_segment(data=getAxes(fProj1, 0.4,0.5,0.26,0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj1, 0.4,0.5,0.26,0.24), aes(x=x2, y=y2, label=label), size=2.5) + ggtitle("a. TIC: 779.46")
-iLast <- length(bbhTourFull2)
-fProj2 <- bbhTourFull2[[iLast]]
-dProj2 <- as_tibble(bbhM %*% fProj2)
-colnames(fProj2) <- c("PP1", "PP2")
-colnames(dProj2) <- c("PP1", "PP2")
-fProj2 <- as_tibble(fProj2) %>%
-  add_column(label=colnames(bbhM))
-p2 <- ggplot(dProj2, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.3,-1,0.24,0.22), aes(x=x, y=y), color="grey") + 
-  xlim(-0.1, 1.1) + ylim(-1.3, -0.2) +
-  geom_segment(data=getAxes(fProj2,0.3,-1,0.24,0.22), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj2, 0.3,-1,0.24,0.22), aes(x=x2, y=y2, label=label), size=2.5) + ggtitle("b. splines2D: 0.97")
-iLast <- length(bbhTourFull3)
-fProj3 <- bbhTourFull3[[iLast]]
-dProj3 <- as_tibble(bbhM %*% fProj3)
-colnames(fProj3) <- c("PP1", "PP2")
-colnames(dProj3) <- c("PP1", "PP2")
-fProj3 <- as_tibble(fProj3) %>%
-  add_column(label=colnames(bbhM))
-p3 <- ggplot(dProj3, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(0.75, 1, 0.3, 0.28), aes(x=x, y=y), color="grey") + 
-  xlim(-0.2, 1.3) + ylim(-0.1, 1.3) +
-  geom_segment(data=getAxes(fProj3,0.75, 1, 0.3, 0.28), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj3, 0.75, 1, 0.3, 0.28), aes(x=x2, y=y2, label=label), size=2.5) + ggtitle("c. 1-convex: 0.50")
-grid.arrange(p1, p2, p3, ncol=3) 
-```
-
-```{r indexTableBBH, eval=FALSE}
-getIdxComp <- function(dProj, n){
-  ticI <- mine(dProj$PP1, dProj$PP2)$TIC
-  splineI <- splines2d(dProj$PP1, dProj$PP2)
-  invConvI <- 1-scagnostics(as.matrix(dProj))[,"Convex"]
-  return(tibble(TIC=ticI, splines2D=splineI, invConv=invConvI, nView=n))
-}
-idxT <- tibble(TIC=numeric(),
-               splines2D=numeric(),
-               invConv=numeric(),
-               nView=numeric())
-idxT <- rbind(idxT, getIdxComp(dProj1, 1),
-      getIdxComp(dProj2, 2),
-      getIdxComp(dProj3, 3))
-colnames(idxT) <- c("TIC", "splines2D", "1-convex", "Projection")
-knitr::kable(idxT,  caption = "Matrix of index values, for all three indices and all three final projections. \\label{tab:indexTableBBH}", digits=2) 
-```
-
-#### Exploring reduced space
-
-The variables time, dec and ra are dropped from the data, and PP is applied to the remaining 8D space.
-Figure \ref{fig:bbhGuided2} shows the projections which maximize the TIC, splines2D and 1-convex indices.
-The results provide similar information as already learned from the scatterplot matrix (Figure \ref{fig:bbhSimulation}).
-The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2d index.
-The 1-convex index finds nothing interesting.
-
-```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2d and 1-convex.", fig.height = 3.2, dev = "png", dpi=300}
-## Step 4: exploring reduced parameter space (here simply dropping several variables)
-bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
-if(!file.exists("cache/bbhGuidedTIC2.rda")){
-  set.seed(2018)
-  bbhTour12 <- save_history(bbhM2, guided_tour(mineIndex("TIC")))
-  bbhTourFull12 <- as.list(interpolate(bbhTour12))
-  save(bbhTour12, bbhTourFull12, file = "cache/bbhGuidedTIC2.rda")
-} else {
-  load("cache/bbhGuidedTIC2.rda")
-}
-if(!file.exists("cache/bbhGuidedSpline2.rda")){
-  set.seed(2018)
-  bbhTour22 <- save_history(bbhM2, guided_tour(splineIndex()))
-  bbhTourFull22 <- as.list(interpolate(bbhTour22))
-  save(bbhTour22, bbhTourFull22, file = "cache/bbhGuidedSpline2.rda")
-} else {
-  load("cache/bbhGuidedSpline2.rda")
-}
-if(!file.exists("cache/bbhGuidedConvex2.rda")){
-  set.seed(2018)
-  bbhTour32 <- save_history(bbhM2, guided_tour(invConvexIndex()))
-  bbhTourFull32 <- as.list(interpolate(bbhTour32))
-  save(bbhTour32, bbhTourFull32, file = "cache/bbhGuidedConvex2.rda")
-} else {
-  load("cache/bbhGuidedConvex2.rda")
-}
-
-iLast <- length(bbhTourFull12)
-fProj1 <- bbhTourFull12[[iLast]]
-dProj <- as_tibble(bbhM2 %*% fProj1)
-colnames(fProj1) <- c("PP1", "PP2")
-colnames(dProj) <- c("PP1", "PP2")
-fProj1 <- as_tibble(fProj1) %>%
-  add_column(label=colnames(bbhM2))
-iv <- mine(dProj$PP1,dProj$PP2)["TIC"]
-p1 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1) + 
-  geom_path(data=getCircle(1, 0, 0.34, 0.25), aes(x=x, y=y), color="grey") + 
-  xlim(-0.2, 1.5) + ylim(-0.25, 1) +
-  geom_segment(data=getAxes(fProj1,1, 0, 0.34, 0.25), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj1, 1, 0, 0.34, 0.25), aes(x=x2, y=y2, label=label), size=2.5) +
-  ggtitle(paste0("a, TIC: ", toString(format(iv, digits=2))))
-iLast <- length(bbhTourFull22)
-fProj2 <- bbhTourFull22[[iLast]]
-dProj <- as_tibble(bbhM2 %*% fProj2)
-colnames(fProj2) <- c("PP1", "PP2")
-colnames(dProj) <- c("PP1", "PP2")
-fProj2 <- as_tibble(fProj2) %>%
-  add_column(label=colnames(bbhM2))
-iv <- splines2d(dProj$PP1,dProj$PP2)
-p2 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(0.75, 0, 0.25, 0.24), aes(x=x, y=y), color="grey") + 
-  xlim(-0.25, 1) + ylim(-0.9, 0.3) +
-  geom_segment(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj2, 0.75, 0, 0.25, 0.24), aes(x=x2, y=y2, label=label), size=2.5) +
-  ggtitle(paste0("b, splines2d: ", toString(format(iv, digits=2))))
-iLast <- length(bbhTourFull32)
-fProj3 <- bbhTourFull32[[iLast]]
-dProj <- as_tibble(bbhM2 %*% fProj3)
-colnames(fProj3) <- c("PP1", "PP2")
-colnames(dProj) <- c("PP1", "PP2")
-fProj3 <- as_tibble(fProj3) %>%
-  add_column(label=colnames(bbhM2))
-iv <- 1-scagnostics(as.matrix(dProj))[,"Convex"]
-p3 <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(0, 1.25, 0.32, 0.32), aes(x=x, y=y), color="grey") + 
-  xlim(-0.4, 1.2) + ylim(0, 1.6) +
-  geom_segment(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj3, 0, 1.25, 0.32, 0.32), aes(x=x2, y=y2, label=label), size=2.5) +
-  ggtitle(paste0("c, 1-convex: ", toString(format(iv, digits=2))))
-grid.arrange(p1, p2, p3, ncol=3) 
-```
-
-#### Effect of random starts, and subsets used
-
-The initial conditions for the optimization, and the subset of variables used, can have a large effect on the projections returned.
-We illustrate this using only the splines2d index, and find that there is one more association that can be learned that was masked earlier.
-
-Figure \ref{fig:bbhGuided3} shows six maxima obtained by different starts, for two types of parameters: first, spin related parameters (i.e.
-alpha, theta\_jn, chi\_tot and chi\_p), and second position related parameters (i.e.
-ra, dec and distance).
-Four of the six (a-d) are almost identical, but not interesting projections.
-Projection f has the highest PP index value but it is primarily the view seen in the bivariate plot of dec and ra.
-While none of these projections are particularly interesting on their own, moving between them can be revealing.
-
-Choosing a different subset of variables reveals something new.
-The subspace of m1, ra, chi\_tot, alpha, distance, dec produces a more refined view of Figure \ref{fig:bbhGuided3} projection f.
-When alpha contributes in contrast to dec, the relationship between the points is almost perfectly on a curve.
-This is shown in Figure \ref{fig:constructedExample}.
-Manually reconstructing the optimal projection (left plot) can be done by differencing the two parameters, in their original units.
-This highlights the importance of improved optimization, that would use tiny final steps to polish the view to a finer optimal projection and possibly remove noise induced by small contributions of many variables.
-
-```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane."), dev = "png", dpi=300}
-## Step 4: exploring dependence on randomly selected starting planes
-bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
-getResWithSeed <- function(s){
-  set.seed(s)
-  startM <- orthonormalise(matrix(runif(14),ncol = 2))
-  tourRes <- save_history(bbhM3, guided_tour(splineIndex()))
-}
-
-if(!file.exists("cache/bbhGuidedSpline3.rda")){
-  set.seed(2018)
-  seedVals <- sample(1:10000, 6)
-  bbhRes1 <- getResWithSeed(seedVals[1])
-  bbhRes2 <- getResWithSeed(seedVals[2])
-  bbhRes3 <- getResWithSeed(seedVals[3])
-  bbhRes4 <- getResWithSeed(seedVals[4])
-  bbhRes5 <- getResWithSeed(seedVals[5])
-  bbhRes6 <- getResWithSeed(seedVals[6])
-  save(bbhRes1, bbhRes2, bbhRes3, bbhRes4, bbhRes5, bbhRes6, file = "cache/bbhGuidedSpline3.rda")
-} else {
-  load("cache/bbhGuidedSpline3.rda")
-}
-resList <- list(as.list(bbhRes1), as.list(bbhRes2), as.list(bbhRes3),
-                as.list(bbhRes4), as.list(bbhRes5), as.list(bbhRes6))
-pList <- list()
-idxV <- NULL
-i <- 1
-xmins <- c(-0.5,-0.5,-0.4,-0.9,-1,0.25)
-xmaxs <- c(0.75,0.75,0.8,0.4,0.25,1.5)
-ymins <- c(-0.8,-0.8,-0.7,-0.8,-1.2,-0.6)
-ymaxs <- c(0.7,0.6,0.7,0.7,0.,0.9)
-xscales <- c(0.25,0.25,0.24,0.26,0.25,0.25)
-yscales <- c(0.3,0.28,0.28,0.3,0.24,0.3)
-xcents <- c(0.25,-0.25,-0.1,0.1,-0.75,1.25)
-ycents <- c(0.4,-0.4,-0.4,-0.4,-0.75,0.5)
-titleL <- c("a: ","b: ","c: ","d: ","f: ","e: ")
-for (r in resList){
-  iLast <- length(r)
-  fProj <- r[[iLast]]
-  dProj <- as_tibble(bbhM3 %*% fProj)
-  colnames(fProj) <- c("PP1", "PP2")
-  colnames(dProj) <- c("PP1", "PP2")
-  fProj <- as_tibble(fProj) %>%
-    add_column(label=colnames(bbhM3))
-  idxV <- splines2d(dProj$PP1, dProj$PP2)
-  pList[[i]] <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x, y=y), color="grey") + 
-  xlim(xmins[i], xmaxs[i]) + ylim(ymins[i], ymaxs[i]) +
-  geom_segment(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x2, y=y2, label=label), size=2.5) +
-    ggtitle(paste0(titleL[i], toString(format(idxV, digits=2))))
-  i <- i+1
-}
-grid.arrange(pList[[1]], pList[[2]], pList[[3]],pList[[4]], pList[[6]], pList[[5]], ncol=3)
-```
-
-```{r eval=FALSE}
-m3 <- matrix(c(1,0,0,0,1,0,0,0,1,0,0,0,1,0),ncol=2)
-m4 <- matrix(c(1,1,1,0,0,0,0,0,1,1,0,0,0,0),ncol=2)
-planeL <- list(m3, m4)
-li <- 3
-for(i in c(1, 3, 6, 5)){
-  r <- resList[[i]]
-  iLast <- length(r)
-  fProj <- r[[iLast]]
-  planeL[[li]] <- fProj
-  li <- li+1
-}
-tP <- tourr::planned_tour(planeL[3:6]) # only need dummy planes when using history tour
-dPoints <- attributes(r)$data
-quartz()
-animate_xy(dPoints, tP, pch=".")
-```
-
-```{r constructedExample, results="hide", fig.cap="Manual reconstruction of an optimal projection (left), constructed by differencing alpha from dec in the original units against ra (middle), compared with the two main variables (right).", fig.height = 3.2, dev = "png", dpi=300}
-bbhC <- rescale(select(bbhD, m1, ra, chi_tot, alpha, distance, dec))
-if(!file.exists("cache/constructedExample.rda")){
-  set.seed(1999)
-  tC <- save_history(bbhC, guided_tour(splineIndex()))
-  save(tC, file = "cache/constructedExample.rda")
-} else {
-  load("cache/constructedExample.rda")
-}
-iL <- length(as.list(tC))
-dProj <- bbhC %*% as.list(tC)[[iL]]
-fProj <- as.list(tC)[[iL]]
-colnames(fProj) <- c("PP1", "PP2")
-colnames(dProj) <- c("PP1", "PP2")
-fProj <- as_tibble(fProj) %>%
-  add_column(label=colnames(bbhC))
-p1 <- ggplot(as_tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)+ 
-  geom_path(data=getCircle(-0.7, 0.6, 0.2, 0.24), aes(x=x, y=y), color="grey") + 
-  xlim(-0.9, 0.1) + ylim(-0.1, 1.1) +
-  geom_segment(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
-  geom_text(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
-p2 <- ggplot(as_tibble(bbhD),aes(dec-alpha, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
-p3 <- ggplot(as_tibble(bbhD),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
-grid.arrange(p1, p2, p3, ncol=3)
-```
-
-## Instructions for applying to new data
-
-Applying these procedures to new datasets can be done using the guided tour available in the `tourr` package. The typical steps required are:
-
-1. Scale, standardize or sphere (principal components) the data. 
-2. Index function selection matching the type of structure that is interesting to detect. Any new function can be used as long as it takes a matrix as input and returns a single index value.
-3. Call the guided tour with the data and index function:
-    + for exploration this can be done via the `tourr::animate` function
-    + for recording the results `tourr::save_history` should be used
-4. Explore how the results depend on choices made (index function, starting planes, optimization method, prior dimension reduction).
-
-These are steps followed in the above two applications and are documented in the comments of the source code [@paperSRC].
-For simple usage examples, see documentation of the `tourr::guided_tour` function.
-
-<!--
-
-# Applications to collections of time series
-
-This could be another example.
-
-```{r eval=FALSE}
-music <- read_csv("data/tigs_music.csv")
-music <- apply(music, 2, function(x) (x-mean(x))/sd(x))
-musicTour <- save_history(music,
-                         guided_tour(mineIndex("TIC"), 
-                            search_f = tourr:::search_better,
-                            cooling=1, alpha = 0.5, max.tries = 5000))
-musicTour <- save_history(music,
-                         guided_tour(dcorIndex(), 
-                            search_f = tourr:::search_better,
-                            cooling=1, alpha = 0.5, max.tries = 5000))
-musicTourFull <- as.list(interpolate(musicTour))
-iLast <- length(musicTourFull)
-fProj <- musicTourFull[[iLast]]
-dProj <- as_tibble(as.matrix(music) %*% fProj)
-ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
-fullRes <- getProj(as.matrix(music), musicTourFull, "TIC", 100)
-fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
-ggplot(fullResMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(size))) + 
-  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
-  guides(color=FALSE)
-```
-
--->
-
-# Discussion
-
-<!--
-Reminder what the paper is about
-summarise main findings
-Future plans
--->
-
-The motivation for this work was to discover dependencies between estimated parameters in multiple model fits in physics problems.
-This paper shows how projection pursuit with the new indexes can help address this problem.
-The results are encouraging, showing large potential for discovering unanticipated relations between multiple variables.
-
-All of the indexes fall short against some aspect of the ideal properties of smoothness, squintability, flexibility, rotation invariance and speed.
-The paper describes how these properties can be assessed using tour methodology.
-Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
-\textcolor{red}{We recommend to use the spinebil [@spinebil] package when developing new indexes.
-It includes the functionalities needed to reproduce the assessments presented in this paper.
-While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.}
-
-The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools. Exploring this area would help improve the guided tours.
-\textcolor{red}{As new optimization techniques become available, adapting these to the guided tour would extend the technique to a broader range of problems.
-The current recommended approach is to first reduce the dimensionality, for example by using PCA, taking care to preserve nonlinear structure, prior to applying PP.}
-
-\textcolor{red}{To use the existing index functions on real-world datasets we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [@galahr] package, superseding the now archived tourrGui [@tourrGui].
-Both packages contain examples to show how the guided tour can be used with different index functions.}
-
-# Supplementary material
-
-- This article was created with R Markdown [@rmarkdown], the code for the paper is available at [@paperSRC].
-- Methods for testing new index functions as presented in this work are implemented in the R package spinebil [@spinebil].
-- The R package galahr [@galahr] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
-- Additional explanations are available in the Appendix, covering details of 
-    - how the holes index was rescaled, 
-    - estimations of the squint angle, 
-    - a comparison of the computational performance of the index functions, 
-    - testing the effect of index parameters on the results and 
-    - suggestions how the index functions may be refined.
-
-# Acknowledgments {-}
-The authors gratefully acknowledge the support of the Australian Research Council.
-We thank Rory Smith for help with the gravitational wave examples, and German Valencia for useful comments.
-This article was created with knitr [@knitr], R Markdown [@rmarkdown] and bookdown [@bookdown] with embedded code.
 
@@
 
-if(!file.exists("cache/squintAngle.rda")){
+if(!file.exists("../cache/squintAngle.rda")){
   allAngles <- tibble(d=numeric(), angle=numeric())
@@
   }
-  save(allAngles, file = "cache/squintAngle.rda")
+  save(allAngles, file = "../cache/squintAngle.rda")
 } else {
-  load("cache/squintAngle.rda")
+  load("../cache/squintAngle.rda")
 }
@@
 These, along with dcor2D, are slower with larger sample sizes.", out.width=".6\\textwidth"}
-if(!file.exists("cache/timer.rda")){
+if(!file.exists("../cache/timer.rda")){
   set.seed(2018)
@@
   }
-  save(dfTimer, file = "cache/timer.rda")
+  save(dfTimer, file = "../cache/timer.rda")
 } else {
-  load("cache/timer.rda")
+  load("../cache/timer.rda")
 }
@@
 }
-if(!file.exists("cache/spiralScagBinning.rda")){
+if(!file.exists("../cache/spiralScagBinning.rda")){
   scagBinningDf <- getScagComp(spiral1000, t2full)
@@
     mutate(idx = str_sub(variable, end = -3))
-  save(scagBinningDf, scagMelt, file = "cache/spiralScagBinning.rda")
+  save(scagBinningDf, scagMelt, file = "../cache/spiralScagBinning.rda")
 } else {
-  load("cache/spiralScagBinning.rda")
+  load("../cache/spiralScagBinning.rda")
 }
@@
 The best $\\delta = 0.7$, and the result indicates that this parameter does affect MIC performance."}
-if(!file.exists("cache/randuEx.rda")){
+if(!file.exists("../cache/randuEx.rda")){
   pL <- list()
@@
   }
-  save(pL, file = "cache/randuEx.rda")
+  save(pL, file = "../cache/randuEx.rda")
 } else {
-  load("cache/randuEx.rda")
+  load("../cache/randuEx.rda")
 }
@@
 
-if(!file.exists("cache/spiralScagSmoothing.rda")){
+if(!file.exists("../cache/spiralScagSmoothing.rda")){
   scagSmooth <- getScagSmooth(spiral1000, as.list(t1full))
   scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
-  save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")
+  save(scagSmooth, scagSmoothMelt, file = "../cache/spiralScagSmoothing.rda")
 } else {
-  load("cache/spiralScagSmoothing.rda")
+  load("../cache/spiralScagSmoothing.rda")
 }
@@
 #getting non-smoothed result
-load("cache/V1V2toV3V4.rda")
+load("../cache/V1V2toV3V4.rda")
 noSmoothing <- fullRes %>%
@@
 ```{r testpipe,  fig.height=4, fig.width=8, out.width = "0.8\\textwidth",  fig.cap="Projections returned by TIC optimization: by the scouting phase (left) and refined by  optimization method 3 (right), starting from the scouting phase projection."}
+load("../cache/findpipe.rda")
 iLast <- length(pipeTourFull)
@@
 ```{r neutronStarSPLOM, fig.height=8, fig.width=8, fig.cap="Scatter plot matrix of the neutron star dataset, darker regions represent higher marginalised posterior densities."}
+nsD <- read_csv("../data/samples.csv") 
 ggpairs(nsD, lower=list(continuous = wrap("points", alpha = 0.05)))
@@
 Strong correlation between the parameters time, dec and ra can be observed."}
+bbhD <- read_csv("../data/posterior_samples.csv")
 bbhDsmall <- select(bbhD, -phi_jl, -m2, -psi, -chi_eff)

### 2019-08-15 scagnostics.Rmd
Removing color, splitting off appendix.

@@
 
-\textcolor{red}{PP can be combined with other dimension reduction methods when $p$ is very high.
+PP can be combined with other dimension reduction methods when $p$ is very high.
 For example, it can be beneficial to first do principal component analysis prior to PP, especially to remove linear dependencies before searching for other types of association.
@@
 Some projection pursuit indexes do incorporate penalty terms to automate removing noise dimensions.
-It can also be important to have an efficient PP optimizer, particularly when working with high dimensions, because the search space increases exponentially with dimension.}
+It can also be important to have an efficient PP optimizer, particularly when working with high dimensions, because the search space increases exponentially with dimension.
 
@@
 
-\textcolor{red}{This paper is structured as follows.} Section \ref{sec:construct} discusses index construction, and how they can be used in the guided tour.
+This paper is structured as follows.
+Section \ref{sec:construct} discusses index construction, and how they can be used in the guided tour.
 Section \ref{sec:investigate} investigates the behavior of the indexes, explored primarily using tour methods.
 The new guided tour with these indexes is applied to two examples from gravitational wave astronomy (Section \ref{sec:phys}).
-\textcolor{red}{The latter two parts are connected in that the application of the new indexes to these problems is the main motivation for the paper, and the simulation study, in the first part, was conducted to better understand the behavior of the indexes in general.
+The latter two parts are connected in that the application of the new indexes to these problems is the main motivation for the paper, and the simulation study, in the first part, was conducted to better understand the behavior of the indexes in general.
 The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to visually assess their behavior.
 Visual methods to diagnose the index behavior is important because PP is primarily used for visualistion.
-The paper finishes with a discussion about the limitations of this work, and the potential future directions.}
+The paper finishes with a discussion about the limitations of this work, and the potential future directions.
 
@@
 
-\textcolor{red}{A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016.
+A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016.
 It is only available in German, we we summarise the main findings here.
@@
 The mutual information based index functions (MIC< TIC) are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present.
-A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.}
-
-\textcolor{red}{In addition to the seven indexes described above, we will also include the holes index available in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
+A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.
+
+In addition to the seven indexes described above, we will also include the holes index available in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
 This serves as a benchmark, demonstrating some desired behavior.
-The index takes maximum values for a central hole in the distribution of the projected data.}
+The index takes maximum values for a central hole in the distribution of the projected data.
 
@@
 In each set, $n$ is used for the number of points, $p$ is the number of dimensions, and $d=2$ is the projection dimension.
-\textcolor{red}{The three structures were selected to cover both functional and non-functional dependence, different types of nuisance distributions and different structure size and squintability properties.}
+The three structures were selected to cover both functional and non-functional dependence, different types of nuisance distributions and different structure size and squintability properties.
 
@@
 
-\textcolor{red}{For simplicity, in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
-}
+For simplicity, in the investigations of the index behavior, we fix $p=6$, which corresponds to two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
 The structured projection is in variables 5 and 6 ($x_5, x_6$).
@@
 Figure \ref{fig:plotV1V2toV3V4} shows the traces representing the index values calculated across a tour between a pair of nuisance projections.
-\textcolor{red}{The tour path is generated by interpolating between the two independent nuisance planes, i.e. from the projection onto $x_1$-$x_2$ to one onto $x_3$-$x_4$.}
+The tour path is generated by interpolating between the two independent nuisance planes, i.e. from the projection onto $x_1$-$x_2$ to one onto $x_3$-$x_4$.
 The range of each axis is set to be the limits of the index, as might be expected over many different data sets, 0 to 1.
@@
 Figure \ref{fig:wIntermediate} shows the PPIs for a tour sequence between a nuisance and structured projection. A long sequence is generated where the path 
-\textcolor{red}{interpolates between projections onto}
+interpolates between projections onto
 $x_1$-$x_2$, $x_1$-$x_5$, $x_5$-$x_6$, in order to see some of the intricacies of holes index.
@@
 
-\textcolor{red}{Before applying the new index functions, with the guided tour on real examples, we test them on the simulated dataset to understand the performance of the optimization.}
+Before applying the new index functions, with the guided tour on real examples, we test them on the simulated dataset to understand the performance of the optimization.
 The guided tour combines optimization with interpolation between pairs of planes.
@@
 Such an overview is given in Table \ref{tab:summary}, listing if the criteria is fully met (\checkmark), there are some shortcomings ($\cdot$) or failure ($\times$).
-\textcolor{red}{(The holes index does not appear in the summary because its performance understood, and is not being examined here.)}
+(The holes index does not appear in the summary because its performance understood, and is not being examined here.)
 We find that none of the indexes considered meet all criteria, and in particular rotation invariance is often not fulfilled.
@@
 Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
-\textcolor{red}{We recommend to use the spinebil [@spinebil] package when developing new indexes.
+We recommend to use the spinebil [@spinebil] package when developing new indexes.
 It includes the functionalities needed to reproduce the assessments presented in this paper.
-While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.}
+While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.
 
 The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools. Exploring this area would help improve the guided tours.
-\textcolor{red}{As new optimization techniques become available, adapting these to the guided tour would extend the technique to a broader range of problems.
-The current recommended approach is to first reduce the dimensionality, for example by using PCA, taking care to preserve nonlinear structure, prior to applying PP.}
-
-\textcolor{red}{To apply the existing index functions in practice, we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [@galahr] package.
+As new optimization techniques become available, adapting these to the guided tour would extend the technique to a broader range of problems.
+The current recommended approach is to first reduce the dimensionality, for example by using PCA, taking care to preserve nonlinear structure, prior to applying PP.
+
+To apply the existing index functions in practice, we recommend to either use the tourr package directly, or if interaction is required to call the guided tour via the graphical interface available in the galahr [@galahr] package.
 This package supersedes the now archived tourrGui [@tourrGui].
-Both packages contain examples to show how the guided tour can be used with different index functions.}
+Both packages contain examples to show how the guided tour can be used with different index functions.
 
@@
 
-\appendix
-
-# Appendix {-}
-
-# Re-scaling of holes index
-
-Holes and cmass indexes are derived from $I_0^N$ of [@CBC93].
-As noted in proposition 1 of that paper the index takes local maxima for the minimum and maximum of $a_0$ which are achieved by central hole and central mass distributions respectively.
-@Cook:2007:IDG:1557227 then gives explicit index functions defined for sphered data (zero mean, identity variance-covariance matrix).
-They are defined such that each one is maximized for central holes or central mass type distributions, with maximum=1, and cmass=1-holes.
-It follows that for either index both large and small values signal deviation from the normal distribution, and given a normal distribution we expect to find "average" index values rather than values close to zero.
-
-We can estimate the values found for normal distribution by comparing values of $a_{00}$ of Sec 7.1 in [@CBC93].
-The maximum value is $1/(2\pi)$ found for cmass type distributions, the minimum is $1/(2\pi e)$ found for hole type distributions.
-Evaluating for normal distributions gives $1/(4\pi)$, rescaling such that the index values range from 0 to 1 then puts the value for normal distributions at approximately $0.2$.
-This is consistent with the results we found, i.e.
-for normal distributions the cmass index is about $0.2$, and the holes index (=1-cmass) is about $0.8$.
-
-We therefore rescale as follows: first have cut-off at the respective value for the normal distribution, i.e.
-any value below 0.2 (0.8) is set to zero for cmass (holes) index, and we rescale the remaining range to be between zero and one.
-
-# Estimating the squint angle of the spiral
-
-We estimate the squint angle of the spiral (for p = 4, 5, 6) as follows.
-First pick a random starting plane and generate a tour path from the starting plane to the ideal plane containing the spiral.
-Using skinny as the reference index we fix a lower index value that is attributed to indicate squintability at $0.6$, and we move along the tour path towards the ideal plane until this value is reached.
-The distance between the thus identified plane and the ideal plane is used as an estimate of the squint angle in this direction.
-Since this will strongly depend on the random starting plane, i.e.
-the considered direction, we repeat the estimation 100 times and present the results in the form of a box plot in Figure \ref{fig:squintAngle}.
-The result shows a large drop in squint angle when going from p = 4 to higher dimensions, and generally a large spread of squint angles depending on the direction.
-
-```{r squintAngle, out.width="0.5\\textwidth", fig.cap="Estimated squint angles for the Spiral dataset with 1000 datapoints, with p = 4, 5, 6, containing estimates evaluated for 100 randomly selected directions each.", results="hide"}
-
-squintAngleEstimat <- function(data, indexF, cutoff, structurePlane, n = 100, stepSize = 0.01){
-  # data = numerical input with p > 2 dimensions
-  # indexF = the index function used to estimate the squint angle
-  # cutoff = lower bound on the index value to be considered structured (i.e. all lower index values are outside the squint angle)
-  # structuredPlane = projection matrix onto plane containing the 2-d structure
-  # n = number of random starting directions over which the squint angle estimate is averaged
-  # stepSize = accuracy, step size determines where the index is evaluated to find the first plane above cutoff
-  data <- as.matrix(data) # make sure data is in matrix format
-  angles <- numeric(length = n) # collecting all individual estimates
-  i <- 1
-  p <- ncol(data)
-  while(i <= n){
-    # first generate random direction, make sure it is not too close to structure plane
-    dist <- 0.
-    while(dist < 0.1){
-      rBasis <- tourr::basis_random(p)
-      dist <- tourr::proj_dist(rBasis, structurePlane)
-    }
-    # now interpolate from rBasis to structure plane with selected step size
-    # since planned tour ignores first two entries, generate some random planes to be ignored
-    notUsed1 <- tourr::basis_random(p)
-    notUsed2 <- tourr::basis_random(p)
-    tourHist <- save_history(data,tour_path=planned_tour(list(notUsed1, notUsed2, rBasis, structurePlane)))
-    allBases <- as.list(interpolate(tourHist, angle = stepSize))
-    cIndex <- 0.
-    j <- 1
-    while(cIndex < cutoff){
-      cProj <- data %*% allBases[[j]]
-      cIndex <- indexF(cProj)
-      j <- j+1
-    }
-    cDist <- tourr::proj_dist(allBases[[j]], structurePlane)
-    angles[i] <- cDist
-    i <- i+1
-  }
-  return(angles)
-}
-
-if(!file.exists("cache/squintAngle.rda")){
-  allAngles <- tibble(d=numeric(), angle=numeric())
-  for(d in c(4,5,6)){
-    set.seed(58958)
-    specialPlane <- matrix(c(rep(0,d-2),1,0,rep(0,d-1),1),ncol=2)
-    spiralD <- spiral1000 %>%
-      purrr::when(d<6 ~ select(., -V4), ~ .) %>%
-      purrr::when(d<5 ~ select(., -V3), ~ .)
-    cAngles <- squintAngleEstimat(spiralD, scagIndex("Skinny"), 0.6, specialPlane)
-    allAngles <- bind_rows(allAngles, tibble(d=d, angle=cAngles))
-  }
-  save(allAngles, file = "cache/squintAngle.rda")
-} else {
-  load("cache/squintAngle.rda")
-}
-
-allAngles <- rename(allAngles, p=d)
-
-ggplot(allAngles, aes(p, angle, group=p)) + geom_boxplot()
-```
-
-# Computational performance {#app:speed}
-Computational time is important for using the PPIs with the guided tour, online.
-Figure \ref{fig:timer} summarizes performance for each PPI.
-For simplicity, data with sample sizes ranging from 100-10000 are drawn from a 6-d solid sphere, using the geozoo package [@geozoo].
-The time to compute the PPIs over 100 interpolated grand tour projections is recorded.
-The scagnostics PPI are computed as a bundle, since this is the code base, and that major computational constraint is common to all the scagnostics.
-There are two versions of the MIC and TIC algorithm, labelled MINE and MINE E, the second being a newer algorithm which improves their computational performance.
-
-The results are interesting. The scagnostic indexes and splines2d are very fast regardless of sample size. MIC, TIC (both versions) and dcor2d slow rapidly as sample size increases.
-
-```{r getTimer}
-
-mineE <- function(x,y){
-  return(mine(x=x, y=y, est = "mic_e"))
-}
-
-timeThis <- function(d, t, idx, pmax, n, idxName){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections, n=sample size, idxName=str name of idx funciton
-    i <- 1
-    dfTimer = data.frame( t= numeric(), i=numeric(), n = numeric(), name=character())
-    for(pMatrix in t){
-      if(i>pmax) break
-      tic.clearlog()
-      tic() #start timer
-      dProj <- d %*% pMatrix
-      sgnst <- idx(dProj[,1],dProj[,2])
-      toc(log=TRUE,quiet=TRUE)
-      scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
-      dfTimer <- add_row(dfTimer, t=scTd, i=i, n= n, name=idxName)
-      i = i+1
-    }
-    return(dfTimer)
-}
-
-```
-
-```{r timer, fig.cap="Computational perfoamce for PPIs, using sample sizes 100-10000.
-Colour indicates the PPI.
-Because the scagnostics calculation is bundled together, the values are the same for all these indexes, and they are really fast to compute.
-MINE includes the MIC and TIC indexes, and MINE E are computationally more efficient algorithms for these.
-These, along with dcor2D, are slower with larger sample sizes.", out.width=".6\\textwidth"}
-if(!file.exists("cache/timer.rda")){
-  set.seed(2018)
-  grandTour100 <- save_history(sphereData(6,100), grand_tour(2), max=4) %>%
-    interpolate() %>%
-    as.list()
-
-  sizeL <- c(10, 100, 500, 1000, 5000, 10000)
-  t <- 1
-  dfTimer <- NULL
-  for(sampleSize in sizeL){
-    set.seed(sampleSize + 11) # new seed for each sample size
-    if(sampleSize == 0) sampleSize = 10 # smallest considered sample has 10 points
-    thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
-    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100, sampleSize, "scagnostics")
-    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100, sampleSize, "dcor2D")
-    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100, sampleSize, "splines2D")
-    mineT <- timeThis(thisMatrix, grandTour100, mine, 100, sampleSize, "MINE")
-    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100, sampleSize, "MINE E")
-    dfTimerC <- list(scagT, dcorT, splineT, mineT,mineeT) %>%
-      reduce(rbind)
-    dfTimer <- rbind(dfTimer, dfTimerC)
-    t <-  t+1
-  }
-  save(dfTimer, file = "cache/timer.rda")
-} else {
-  load("cache/timer.rda")
-}
-
-timerMeans <- dfTimer %>%
-  group_by(name, n) %>%
-  summarise(t=mean(t))
-
-ggplot(dfTimer, aes(x = n, y = t)) + 
-  geom_point(alpha=0.1, aes(color=name, shape=name)) +
-  geom_point(data=timerMeans, aes(color=name, shape=name)) +
-  geom_line(data=timerMeans, aes(linetype=name, color=name)) +
-  scale_x_log10(limits=c(90,10100)) +
-#  scale_y_log10(limits=c(0.00001,1)) + 
-  labs(x = "Sample Size", y = "Time [s]", color = "Index family", 
-       linetype = "Index family", shape = "Index family") + 
-  scale_colour_brewer(palette="Dark2")
-```
-
-# Effect of parameter choice in index value {#app:param}
-Some parameters must be provided for some PPIs.
-This can be advantageous, allowing the index to more flexibly work for different types of structure, controlling trade-offs between noise and fine structure detection, and affecting computing time and precision.
-
-- Binning:
-    - Scagnostics: the number of bins can be controlled by the user, note however that internally the implementation will reduce the number of bins if too many non-empty bins are found (more than 250).
-    - MINE: the maximum number of bins considered is fixed by the user as a function of the number of data points.
-The default is chosen as a trade-off between resolution and noise dependence, but it may be tuned based on requirements dictated by specific datasets.
-Apart from sensitivity to noise computing time may also be a consideration here.
-- Spline knots: for the splines2d measure we need to fix the number of knots.
-By default it is fixed to be 10 (or lower if appropriate based on the data values).
-In our examples we find the number to be appropriate to identify functional dependence while rejecting noise, but some distributions may require tuning of this parameter.
-
-<!--### Effect of binning in scagnostics -->
-
-The bins argument for the scagnostics might be reasonably expected to affect the smoothness of the index: a small number of bins should provide a smooth index function, but may affect its ability to detect fine structure.
-Figure \ref{fig:spiralScagBinning} examines this.
-Scagnostics PPIs are computed for the spiral1000 data on a tour path between $x_1$-$x_2$ to $x_5$-$x_6$, for number of bins equal to 10, 20, 50.
-The interesting observation to make is that even with small bin size the functions are all relatively noisy.
-The problem with the small bin size is that the spiral becomes invisible to the PPI.
-
-```{r spiralScagBinning, fig.cap="Comparing the traces of the three scagnostics indices when changing the binning via the bins parameter set to 10, 20 and 50 in this example.", out.width=".7\\textwidth"}
-getScagComp <- function(df, tPath){
-  sc <- tibble(
-    convex10=numeric(),
-    skinny10=numeric(),
-    stringy10=numeric(),
-    convex20=numeric(),
-    skinny20=numeric(),
-    stringy20=numeric(),
-    convex50=numeric(),
-    skinny50=numeric(),
-    stringy50=numeric(),
-    t=numeric())
-  n <- length(tPath)
-  for (i in 1:n) {
-    dprj <- as.matrix(df) %*% tPath[[i]]
-    scagRes10 <- scagnostics.default(dprj[,1], dprj[,2], bins=10)$s
-    scagRes20 <- scagnostics.default(dprj[,1], dprj[,2], bins=20)$s
-    scagRes50 <- scagnostics.default(dprj[,1], dprj[,2], bins=50)$s
-    sc <- add_row(sc,
-                  convex10=scagRes10["Convex"], skinny10=scagRes10["Skinny"],
-                  stringy10=scagRes10["Stringy"],
-                  convex20=scagRes20["Convex"], skinny20=scagRes20["Skinny"],
-                  stringy20=scagRes20["Stringy"],
-                  convex50=scagRes50["Convex"], skinny50=scagRes50["Skinny"],
-                  stringy50=scagRes50["Stringy"],
-                  t=i)
-  }
-
-  return(sc)
-}
-if(!file.exists("cache/spiralScagBinning.rda")){
-  scagBinningDf <- getScagComp(spiral1000, t2full)
-  scagMelt <- gather(scagBinningDf, variable, value, -t) %>%
-    mutate(nbin = as.numeric(str_sub(variable, start = -2))) %>%
-    mutate(idx = str_sub(variable, end = -3))
-  save(scagBinningDf, scagMelt, file = "cache/spiralScagBinning.rda")
-} else {
-  load("cache/spiralScagBinning.rda")
-}
-ggplot(scagMelt, aes(x=t, y=value)) +
-  geom_line(aes(color=factor(nbin))) + 
-  facet_wrap(~idx, ncol=1, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
-  scale_colour_discrete("bins") + xlab("Number of projections") + 
-  ylab("PPI value")
-
-```
-
-## Binning sensitivity of MIC index
-To examine the sensitivity of binning in the MIC PPI, the classic RANDU data [@Marsaglia68], available in R, is used.
-Binning is defined by $\delta$, where $B(n) = n^{\delta}$.
-Figure \ref{fig:randuEx} shows the best projection, index value and computing time obtained when optimizing the MIC index with values $\delta = 0.6, 0.7, 0.8$.
-With small $\delta$, less bins, the structure isn't visible, and with larger $\delta$ the structure is confounded with noise.
-It does appear that this parameter affects the performance of the MIC index.
-
-```{r randuEx, results="hide", fig.cap="Best projection obtained by optimizing the MIC index on the RANDU data, using different number of bins, defined by $\\delta$.
-The smaller the value the fewer bins.
-Above each plot is written the value of $\\delta$, time required to optimize (seconds) and the MIC index value.
-The best $\\delta = 0.7$, and the result indicates that this parameter does affect MIC performance."}
-if(!file.exists("cache/randuEx.rda")){
-  pL <- list()
-  i <- 1
-  for (alpha in c( 0.6, 0.7, 0.8)){
-    set.seed(556677)
-    tic.clearlog()
-    tic() #start timer
-    tP <- save_history(randu, guided_tour(mineIndexAlpha("MIC", alpha)),max_bases = 100000)
-    toc(log=TRUE,quiet=TRUE)
-    optT <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
-    imax <- length(as.list(tP))
-    fV <- as_tibble(as.matrix(randu) %*% as.list(tP)[[imax]])
-    colnames(fV) <- c("PP1", "PP2")
-    idxV <- mine(fV$PP1, fV$PP2, alpha = alpha)[["MIC"]]
-    pL[[i]] <- ggplot(fV, aes(PP1,PP2)) + geom_point() + theme(aspect.ratio=1) + 
-      ggtitle(paste(alpha, toString(format(optT, digits=2)), toString(format(idxV, digits=2)), sep = ", "))
-    i <- i+1
-  }
-  save(pL, file = "cache/randuEx.rda")
-} else {
-  load("cache/randuEx.rda")
-}
-grid.arrange(pL[[1]], pL[[2]], pL[[3]], ncol=3)
-```
-
-# Ways to refine the PPIs {#app:enh}
-
-The biggest issues revealed by the investigation into the new PPIs are a lack of smoothness, particularly for the scagnostics indexes, and the rotation invariance of Grimm's indexes.
-To fix the smoothness of an index function, it is possible to calculate the PPI for a small neighborhood of projections and average the value, or alternatively average the PPI for several jittered projections.
-This is investigated in Fig.
-\ref{fig:spiralScagSmoothing}.
-Rotation invariance is more difficult to fix, but an alternative tour interpolation method could be useful.
-The geodesic interpolation transitions between planes, and it ignores the basis defining the plane, creating a problem with rotation invariant indices.
-Alternative interpolations based on Givens or Householder rotations could be implemented to transition between bases, which should alleviate the need for rotationally invariant indices.
-
-Two different methods are considered for smoothing the index values:
-
-* Jittering points: using the jitter function we move each point by a random amount drawn from a uniform distribution between $\pm \beta$.
-* Jittering angles: using the tourr implementation we can draw a random plane and move some small amount $\epsilon$ in that direction.
-
-The mean value from a sample of projections is recorded as the PPI value. This could be robsitufied by dropping the most extreme values.
-
-This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered.
-Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using the tour path between noise variables of the spiral1000 dataset and different $\epsilon$ and $\beta$ values.
-Both methods appear to be promising in smoothing the function.
-Because the scagnostics are fast to compute, either of these methods is feasible.
-For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.
-
-```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indexes skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line).
-For comparison the red line in the background shows the trace without any smoothing applied."}
-
-jitterScagnostics <-function(proj, d, alpha){
-  newProj <- tourr:::basis_nearby(proj, alpha = alpha, method = "geodesic")
-  newD <- d %*% newProj
-  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
-}
-
-jitterPointsScagnostics <-function(projData, alpha){
-  newD <- jitter(projData, amount=alpha)
-  return(as.vector(scagnostics.default(newD[,1],newD[,2])$s))
-}
-
-getSgnMean <- function(proj, d, alpha, method){
-  dProj <- d %*% proj
-  orig <- as.vector(scagnostics.default(dProj[,1],dProj[,2])$s)
-  if(method == "jitterAngle"){
-    sgnVec <- replicate(10, jitterScagnostics(proj, d, alpha))
-  } else if (method=="jitterPoints"){
-    sgnVec <- replicate(10, jitterPointsScagnostics(dProj, alpha))
-  }
-  return(rowMeans(cbind(orig, sgnVec)))
-}
-
-getScagSmooth <- function(df, tPath){
-  sc <- tibble(
-    convex=numeric(),
-    skinny=numeric(),
-    stringy=numeric(),
-    t=numeric(),
-    method=character(),
-    alpha=numeric())
-  n <- length(tPath)
-  for (method in c("jitterAngle", "jitterPoints")){
-    for (alpha in c(0.01, 0.05, 0.1)){
-      for (i in 1:n) {
-        scagMean <- getSgnMean(tPath[[i]], as.matrix(df), alpha, method)
-        convex <- scagMean[6]
-        skinny <- scagMean[7]
-        stringy <- scagMean[8]
-        sc <- sc %>% add_row(convex=convex, skinny=skinny, stringy=stringy,
-                             t=i, method=method, alpha=alpha)
-      }
-    }
-  }
-
-  return(sc)
-}
-
-if(!file.exists("cache/spiralScagSmoothing.rda")){
-  scagSmooth <- getScagSmooth(spiral1000, as.list(t1full))
-  scagSmoothMelt <- gather(scagSmooth, variable, value, -t, -method, -alpha)
-  save(scagSmooth, scagSmoothMelt, file = "cache/spiralScagSmoothing.rda")
-} else {
-  load("cache/spiralScagSmoothing.rda")
-}
-
-#getting non-smoothed result
-load("cache/V1V2toV3V4.rda")
-noSmoothing <- fullRes %>%
-  filter(size==1000) %>%
-  filter(name=="spiral") %>%
-  select(skinny, stringy, t) %>%
-  gather(variable, value, -t)
-
-scagSmoothMelt <- scagSmoothMelt %>%
-  filter(variable != "convex") %>%
-  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
-ggplot(scagSmoothMelt, aes(x=t, y=value)) +
-  geom_line(data=noSmoothing, color="red", alpha=0.5) +
-  geom_line(aes(linetype=factor(method))) + 
-  theme(legend.position="none") +
-  ylim(0,1) +
-  facet_grid(alpha~variable) +
-  xlab("Sequence of projections (t)") +
-  ylab("PPI value")
-
-```
-
-# Additional Figures
-
-## Final projection of pipe guided tour
-Figure \ref{fig:testpipe} shows the projection returned during the scouting phase (left) and the refinement phase (right).
-It was important to start the method 3 optimizer at the best projection returned during the scouting phase, to smoothly converge more closely to the ideal projection.
-
-```{r testpipe,  fig.height=4, fig.width=8, out.width = "0.8\\textwidth",  fig.cap="Projections returned by TIC optimization: by the scouting phase (left) and refined by  optimization method 3 (right), starting from the scouting phase projection."}
-iLast <- length(pipeTourFull)
-fProj <- pipeTourFull[[iLast]]
-dProj <- as_tibble(pipeResc %*% fProj)
-p1 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
-  xlab("PP1") + ylab("PP2")
-iLast <- length(pipeTourFull2)
-fProj <- pipeTourFull2[[iLast]]
-dProj <- as_tibble(pipeResc %*% fProj)
-p2 <- ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1) +
-  xlab("PP1") + ylab("PP2")
-grid.arrange(p1, p2, ncol=2)
-```
-
-## Dataset overview scatter plot matrices
-
-Figures \ref{fig:neutronStarSPLOM} and \ref{fig:bbhSimulation} show the scatter plot matrices of the gravitational wave datasets considered, see Section \ref{sec:phys} for details.
-
-```{r neutronStarSPLOM, fig.height=8, fig.width=8, fig.cap="Scatter plot matrix of the neutron star dataset, darker regions represent higher marginalised posterior densities."}
-ggpairs(nsD, lower=list(continuous = wrap("points", alpha = 0.05)))
-```
-
-```{r bbhSimulation, fig.height=8, fig.width=8, dev = "png", dpi=300, fig.cap="Scatter plot matrix showing most of the variables included in the BBH dataset.
-Strong correlation between the parameters time, dec and ra can be observed."}
-bbhDsmall <- select(bbhD, -phi_jl, -m2, -psi, -chi_eff)
-ggpairs(bbhDsmall, lower=list(continuous = wrap("points", alpha = 0.02)))
-```
-
 # References {-}

### 2019-08-15 Appendix/build.R

@@ -1,0 +1,3 @@
+setwd("Appendix/")
+bookdown::render_book("appendix.Rmd")
+setwd("..")

### 2019-08-15 review/response.Rmd
Preparing resubmission.

@@
 
-**PP can also be used to explore data with many more variables, data that is commonly seen today.
-For example, [@Cook:2018mvr] explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit.
-There is not a single solution for working with high-dimensional data.
-PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies but preserved the nonlinear relationships that were interesting to discover.
-Some projection pursuit indexes incorporation penalty terms to naturally deal with noise dimensions.
-Having an efficient optimizer can also important with increasing dimensionality, because there are many more dimensions to search.**
+**PP can be combined with other dimension reduction methods when p is very high.
+For example, it can be beneficial to first do principal component analysis prior to PP, especially to remove linear dependencies before searching for other types of association.
+This is the approach used in Cook, Laa, and Valencia (2018), which explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit.
+PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies while preserving the nonlinear relationships that were interesting to discover.
+Some projection pursuit indexes do incorporate penalty terms to automate removing noise dimensions.
+It can also be important to have an efficient PP optimizer, particularly when working with high dimensions, because the search space increases exponentially with dimension.**
 
@@
 The holes index is not included in Table 3 because the focus is on summarising our findings on the new index functions.
-Because the MINE indexes (MIC, TIC) that have very similar behaviour in all assessments,  we combined for this summary.
+Because the MINE indexes (MIC, TIC) have very similar behaviour in all assessments, we had combined them for this summary.
 We have now added another line for TIC, because it doesn't use much extra space, and saves confusion about the differing number of indexes.
@@
 Thirdly, the simulations refer to sequences of projections (the horizontal axis in plots in Figures 2, 3, and 4.
-As I understand these sequences are a consequence of optimising the corresponding projection pursuit index.
+As I understand these sequences are a consequence of optimizing the corresponding projection pursuit index.
 How can you then actually control the two different scenarios referred to in section 3.2 and 3.3?
@@
 
-The optimisation methods are further discussed in section 3.4 and depend on the structure present in the data.
+The optimization methods are further discussed in section 3.4 and depend on the structure present in the data.
 What is the value of these investigations for more realistic settings in which the structure is not known.

### 2019-08-15 Appendix/appendix.Rmd
Preparing resubmission.

@@
 
-```{r timer, fig.cap="Computational perfoamce for PPIs, using sample sizes 100-10000.
+```{r timer, fig.cap="Computational performance for PPIs, using sample sizes 100-10000.
 Colour indicates the PPI.
@@
 
-Figures \ref{fig:neutronStarSPLOM} and \ref{fig:bbhSimulation} show the scatter plot matrices of the gravitational wave datasets considered, see Section \ref{sec:phys} for details.
+Figures \ref{fig:neutronStarSPLOM} and \ref{fig:bbhSimulation} show the scatter plot matrices of the gravitational wave datasets considered.
 

### 2019-08-15 scagnostics.Rmd
Preparing resubmission.

@@
 The performance of these indexes is examined with respect to ideal behaviour, using simulated data, and then applied to problems from gravitational wave astronomy.
-The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphic.
+The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, binostics, minerva and mbgraphic.
 
@@
 
-<!-- 
-Main point of paper:
-
-- Some metrics for exploring high-dimensional data by finding interesting pairs of variables, can be used to find interesting projections, but some have problems. 
-- What are the ways to evaluate.
-- What adjustments can be done.
-- How it can be applied
-
-Organisation of this section:
-
-- Motivation first?
-- PP
-- Scagnostics
-- section needs shortening
--->
-
 The term "projection pursuit" (PP) was coined by @FT74 to describe a procedure for searching high (say $p-$)dimensional data for "interesting" low-dimensional projections ($d=1$ or $2$ usually).
@@
 
-<!-- check Perisic and Posse (2012) and recent paper by Austrian guys -->
-
 The driving application is from physics, to aid the interpretation of model fits on experimental results.
@@
 Current practice is to examine pairs of parameters, or combinations produced by intuition or prior knowledge.
-This begs the question, whether important nonlinear associations are missed because because they are hidden in linear combinations of more than two variables.
+This begs the question, whether important nonlinear associations are missed because they are hidden in linear combinations of more than two variables.
 
@@
 @huber85, @JS87).
-This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
+This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: binostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].
 
@@
 The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to visually assess their behavior.
-Visual methods to diagnose the index behavior is important because PP is primarily used for visualistion.
+Visual methods to diagnose the index behavior is important because PP is primarily used for visualization.
 The paper finishes with a discussion about the limitations of this work, and the potential future directions.
@@
 A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016.
-It is only available in German, we we summarise the main findings here.
+It is only available in German, we we summarize the main findings here.
 The scagnostics measures are flexible, and calculating the full set of measures provides useful guidance in variable selection.
@@
 Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation.
-The mutual information based index functions (MIC< TIC) are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present.
+The mutual information based index functions (MIC, TIC) are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present.
 A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.
 
-In addition to the seven indexes described above, we will also include the holes index available in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227].
+In addition to the seven indexes described above, we will also include the holes index available in the `tourr` package, see (@CBC93, @Cook:2007:IDG:1557227).
 This serves as a benchmark, demonstrating some desired behavior.
@@
   select(-indexF, -index) %>%
-  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", "")) %>%
+  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MIC", "", "TIC", "")) %>%
   select(index, everything())
@@
 kable(res_all, format = "latex",  caption = "Comparison of index values between noise projections and structured projections for sample size 1000, using 5th and 95th percentiles from 100 simulated sets.
-Most indices have much larger values for the structured projections, except for convex.",
+Most indexes have much larger values for the structured projections, except for convex.",
       col.names=c("Index","", "lower", "upper", "lower", "upper", "lower", "upper"),
@@
 The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs.
-The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix).
+The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details are given in the Appendix).
 Similarly, the TIC index is re-scaled depending on sample size.
@@
 If these two criteria are not met the method may still be useful, but only given an informed starting projection.
-In such cases we can follow a method similar to that proposed by Friedman [@f87]: we break the optimization in two distinct steps.
+In such cases we can follow a method similar to that proposed by @f87: we break the optimization in two distinct steps.
 A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum.
@@
 
-The final views obtained in each of the two stages are compared in the Appendix, see Fig. \ref{fig:testpipe}.
+The final views obtained in each of the two stages are compared in the Appendix.
 
@@
 Examining the computing time as a function of sample size we find that scagnostics and splines2d are fast even for large samples, while all other index functions slow rapidly with increasing sample size.
-Detailed results are shown in Appendix \ref{app:speed}.
+Detailed results are shown in the Appendix.
 
@@
 Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure.
-In Appendix \ref{app:param} we examine the dependence of the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
+In the Appendix we examine the dependence of the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure.
 Sensitivity to the number of bins in the MIC index is also examined, showing that tuning the parameter can improve the final result.
@@
 We identified two potential improvements.
-First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in Appendix \ref{app:enh}.
+First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in the Appendix.
 In addition, rotation dependent indexes may be enhanced by redefining them in a rotation invariant way.
@@
 
-Figure \ref{fig:neutronStarSPLOM} in the Appendix shows the scatterplot matrix (with transparency) of the six parameters.
+A scatterplot matrix (with transparency) of the six parameters is shown in the Appendix.
 (In astrophysics, scatterplot matrices are often called "corner plots" [@corner].) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables.
@@
 
-Figure \ref{fig:bbhSimulation} in the Appendix shows a scatterplot matrix, of nine of the twelve parameters.
+A scatterplot matrix, of nine of the twelve parameters, is shown in the Appendix.
 (Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they are uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time.
@@
 All seven PP indexes are applied to the data.
-Figure \ref{fig:bbhGuided} showing the projections that maximize three of the indexes.
+Figure \ref{fig:bbhGuided} shows the projections that maximize three of the indexes.
 TIC and splines2d indexes identify very similar projections, that are based on the three parameters, dec, time and ra.
@@
 Figure \ref{fig:bbhGuided2} shows the projections which maximize the TIC, splines2D and 1-convex indices.
-The results provide similar information as already learned from the scatterplot matrix (Figure \ref{fig:bbhSimulation}).
+The results provide similar information as already learned from the scatterplot matrix.
 The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2d index.

### 2019-12-18 scagnostics.Rmd
Fixed references.

@@
 
-bibliography: bibliography.bib
+bibliography: [bibliography.bib]
 output: rticles::springer_article

### 2019-12-19 scagnostics.Rmd
Simplified notation.

@@
 $\bm{Y}_{n\times 2} = (\bm{Y}_1, ..., \bm{Y}_2)$ & projected data matrix. where $Y_j = (\alpha_1 X_1, ..., \alpha_p X_p)$ \\
-$\bm{F}_{p\times 2} = (\bm{\alpha}_1, ..., \bm{\alpha}_2)$ & orthonomal projection matrix\\ 
 &\\
 $H$ & Convex hull \\
-$A$ & Alpha hull \\\hline
+$A$ & Alpha hull \\
+$MST$ & Minimal spanning tree \\\hline
 \end{tabular}
@@
 The convex [@Eddy:1977] and alpha hulls [@1056714], and the minimal spanning tree (MST) [@2033241], are computed.
-    + *convex*: The ratio of the area of alpha to convex hull, $I_{\bm{F}, convex}= \frac{area(A(Y))}{area(H(Y))}$.
+    + *convex*: The ratio of the area of alpha to convex hull, $I_{convex}= \frac{area(A)}{area(H)}$.
 This is the only measure where interesting projections will take low values, with a maximum of 1 if both areas are the same.
 Thus $1-c_{convex}$ is used.
-    + *skinny*: The ratio of the perimeter to the area of the alpha hull, $I_{\bm{F}, skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $I_{\bm{F}, skinny} = 0$ for a full circle.
+    + *skinny*: The ratio of the perimeter to the area of the alpha hull, $I_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalization is chosen such that $I_{skinny} = 0$ for a full circle.
 Values close to 1 indicate a skinny polygon.
-    + *stringy*: Based on the MST, $I_{\bm{F}, stringy} = \frac{diameter(MST)}{length(MST)}$ where the diameter is the longest connected path, and the length is the total length (sum of all edges).
-If the MST contains no branches $I_{\bm{F}, skinny} = 1$.
+    + *stringy*: Based on the MST, $I_{stringy} = \frac{diameter(MST)}{length(MST)}$ where the diameter is the longest connected path, and the length is the total length (sum of all edges).
+If the MST contains no branches $I_{skinny} = 1$.
 
@@
 The function wdcor, in the R package, extracat [@extracat], computes the statistic, and the mbgraphic package utilises this function.
-    + *splines2D*: Measures nonlinear dependence by fitting a spline model [@wahba1990spline] of $\bm{Y}_2$ on $\bm{Y}_1$ and also $\bm{Y}_1$ on $\bm{Y}_2$, using the \texttt{gam} function in the R package, mgcv [@w16].
+    + *splines2D*: Measures nonlinear dependence by fitting a spline model [@wahba1990spline] of the projected data matrix entries $\bm{Y}_2$ on $\bm{Y}_1$ and also $\bm{Y}_1$ on $\bm{Y}_2$, using the \texttt{gam} function in the R package, mgcv [@w16].
 The index compares the variance of the residuals:
     \begin{equation}
-    I_{\bm{F}, splines2d} = max(1- \frac{Var(res_{\bm{Y}_1\sim \bm{Y}_2})}{Var(\bm{Y}_1)}, 1-\frac{Var(res_{\bm{Y}_2\sim {\bm{Y}_1}})}{Var(\bm{Y}_2)}),
+    I_{splines2d} = max(1- \frac{Var(res_{\bm{Y}_1\sim \bm{Y}_2})}{Var(\bm{Y}_1)}, 1-\frac{Var(res_{\bm{Y}_2\sim {\bm{Y}_1}})}{Var(\bm{Y}_2)}),
     \end{equation}
